### 2022-07-29 additional_files/additional_file_1.tex
Initial commit; first version of additional file 1 with Manhattan plots and a few signal plots

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc} % declare encoding as utf8
\usepackage{graphicx} % to enable \includegraphics
\usepackage{flafter} % to make sure floats appear after their position in text
\usepackage{pdflscape} % allow some pages to be in landscape orientation
\usepackage{layout} % for debugging purposes, looking at page layout
\usepackage[margin=1in]{geometry} % leaves more space for text body
\usepackage[hidelinks]{hyperref} % hyperlinks from list of figures/tables
\usepackage[labelfont=bf]{caption} % format captions and make links go to figure instead of caption
\usepackage{csvsimple} % read and display csv files as tables
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables
\usepackage{textcomp} % makes the \textdegree command available
\usepackage{tocloft} % formatting the lists of tables and figures
\usepackage{xcolor} % enable text coloring
\usepackage{calc} % makes \widthof available to set the width of the last column
\usepackage{array} % using command \arraybackslash to restore the definition of \\ after \centering
\usepackage{longtable} % allow tables to span more than one page
\usepackage{threeparttablex} % threeparttable for longtable
\usepackage{capt-of} % put caption on next page

% Formatting the table of contents/subsection
\renewcommand{\contentsname}{Supplemental methods}

% Formatting the list of tables
\renewcommand{\listtablename}{Supplemental tables} % Write "Supplemental tables" instead of "List of tables"
\renewcommand\cfttabpresnum{Table } % Write "Table" before table number in list of tables
\setlength{\cfttabnumwidth}{6em} % Spacing before captions
\setlength{\cftbeforefigskip}{1ex} % Vertical spacing between entries

% Formatting the list of figures
\renewcommand{\listfigurename}{Supplemental figures} % Write "Supplemental figures" instead of "List of figures"
\renewcommand\cftfigpresnum{Figure } % Write "Figure" before fig number in list of figures
\setlength{\cftfignumwidth}{6em} % Spacing before captions
\setlength{\cftbeforetabskip}{1ex} % Vertical spacing between entries

\graphicspath{ {../figures/} } % tell Latex to look for figures in figures/

% Format figure and table numbers to follow S1, S2, ...
\renewcommand\thefigure{S\arabic{figure}} 
\renewcommand\thetable{S\arabic{table}}

% Defining new environments as I will always want the figures centered
\newenvironment{cfigure}
	{\begin{figure} \centering}
	{\end{figure}}

\newenvironment{lsfigure}
	{\begin{landscape} \begin{figure} \centering}
	{\end{figure} \end{landscape}}

\setcounter{totalnumber}{1} % Allow a maximum of one float (table/figure) per page

\setlength{\parskip}{2ex}

% Defining a command for manhattan plot figures
\newcommand{\manhattanplot}[1]{
\begin{cfigure}
	\caption[Figure caption]{Figure caption}
	\includegraphics[width = 6in]{#1_manhattan}
	\label{#1_manhattan}
\end{cfigure}

\clearpage}

\newcommand{\signalplot}[1]{
\begin{cfigure}
	\caption[Figure caption]{Figure caption}
	\includegraphics[width = 6in]{#1_signal}
	\label{#1_signal}
\end{cfigure}

\clearpage}

% Information for the title page
\title{Supplemental data to ``ARTICLE TITLE''}
\author{Marc-André Lemay \and François Belzile}
\date{}

% END OF PREAMBLE

\begin{document}

\maketitle \thispagestyle{empty}

\tableofcontents

\vspace{4ex}

\listoftables

\thispagestyle{empty}
%\clearpage

\vspace{4ex}

\listoffigures

%\thispagestyle{empty}

\clearpage

% \caption[Number of SVs per Oxford Nanopore sample]{{{{

%\begin{table}
%	\caption[Number of SVs per Oxford Nanopore sample]{}
%	\label{table_s1}
%
%	\begin{threeparttable}\small
%			   \begin{tablenotes}
%			   \end{tablenotes}
%	\end{threeparttable}
%
%\end{table}
%\clearpage%}}}

% Manhattan plots of all traits
\manhattanplot{stem_termination_all}
\manhattanplot{stem_termination_sn}
\manhattanplot{pubescence_color_all}
\manhattanplot{pubescence_color_nogray}
\manhattanplot{protein}
\manhattanplot{pubescence_form_all}
\manhattanplot{pubescence_form_noerect}
\manhattanplot{oil}
\manhattanplot{pubescence_density}
\manhattanplot{seed_coat_luster_all}
\manhattanplot{seed_coat_luster_nointermediate}
\manhattanplot{seed_coat_luster_dullshiny}
\manhattanplot{pod_color_all}
\manhattanplot{pod_color_blbr}
\manhattanplot{flower_color}
\manhattanplot{seed_coat_color_all}
\manhattanplot{seed_coat_color_greenyellow}
\manhattanplot{hilum_color_all}
\manhattanplot{hilum_color_blackbrown}
\manhattanplot{hilum_color_rbr}
\manhattanplot{maturity_group}
\manhattanplot{corrected_dry_weight}

\signalplot{pca_t}
\signalplot{fc_w}
\signalplot{sda_dt1}

\end{document}

### 2022-07-29 figures/manhattan_plot.R

@@ -1,0 +1,94 @@
+# Loading the required packages
+library(grid)
+library(ggplot2)
+library(gwastools)
+library(GenomicRanges)
+
+# The location of the reference genome fasta
+# DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
+refgenome <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
+
+# A vector of Glycine max chromosome names
+chromosomes <- paste0("Gm", ifelse(1:20 < 10, "0", ""), 1:20)
+
+i <- commandArgs(trailingOnly = TRUE)[1]
+
+#-------------------------------------------------- 
+message("Reading in the GWAS results from vg on ", i)
+
+# Loading the results from the CSV file and formatting them as a GRanges object
+vg_results <- format_gapit_gwas(filename = paste0("gwas/vg_gwas/", i, "/GAPIT.MLM.", i, ".GWAS.Results.csv"),
+				ref_fasta = refgenome,
+				chromosomes = chromosomes,
+				pattern = "^Gm[0-9]{2}$")
+
+# Plotting the results using the gwastools::manhattan_plot function
+vg_plot <- manhattan_plot(vg_results,
+			  gwas_type = "gapit",
+			  threshold = -log10(as.numeric(readLines(paste0("gwas/vg_gwas/",
+									 i, "/", i,
+									 "_threshold_5per.txt"))))) +
+	ggplot2::ggtitle(paste0("vg-", i)) +
+	theme(text = element_text(size = 8))
+
+#-------------------------------------------------- 
+message("Reading in the GWAS results from platypus on ", i)
+
+# Loading the results from the CSV file and formatting them as a GRanges object
+platypus_results <- format_gapit_gwas(filename = paste0("gwas/platypus/", i, "/GAPIT.MLM.", i, ".GWAS.Results.csv"),
+				      ref_fasta = refgenome,
+				      chromosomes = chromosomes,
+				      pattern = "^Gm[0-9]{2}$")
+
+# Plotting the results using the gwastools::manhattan_plot function
+platypus_plot <- manhattan_plot(platypus_results,
+				gwas_type = "gapit",
+				threshold = -log10(as.numeric(readLines(paste0("gwas/platypus/",
+									       i, "/", i,
+									       "_threshold_5per.txt"))))) +
+	ggplot2::ggtitle(paste0("platypus-", i)) +
+	theme(text = element_text(size = 8))
+
[…]

### 2022-07-29 figures/signal_plot.R

@@ -1,0 +1,106 @@
+# Loading the required libraries
+library(gwastools)
+library(grid)
+library(GenomicRanges)
+
+### ---------- ANALYSIS PARAMETERS
+# Getting the name of the trait-locus combination from the command line
+id <- commandArgs(trailingOnly = TRUE)[1]
+
+loci <- read.table("utilities/loci.txt", header = TRUE, sep = ",", stringsAsFactors = FALSE)
+gene_v1 <- loci[loci$id == id, "gene_v1"]
+trait <- loci[loci$id == id, "trait"]
+### ----------
+
+# Loading the lookup table from Gmax v1 gene names to Gmax v4 gene names
+# DEPENDENCY: refgenome/lookup_v1_to_v4.rds
+lookup_v1_to_v4 <- readRDS("refgenome/lookup_v1_to_v4.rds")
+
+# Getting the set of genes, transcripts, exons and coding sequences
+# DEPENDENCY: refgenome/*.rds
+genes <- readRDS("refgenome/gmax_v4_genes.rds")
+transcripts <- readRDS("refgenome/gmax_v4_transcripts.rds")
+exons <- readRDS("refgenome/gmax_v4_exons.rds")
+cds <- readRDS("refgenome/gmax_v4_cds.rds")
+
+# Reading the data.frame of significantly associated k-mers and formatting it
+# DEPENDENCY : refgenome/Gmax_508_v4.0_mit_chlp.fasta.fai
+# DEPENDENCY: kmer positions
+kmer_results <- readRDS(paste0("gwas/kmers/", trait, "/katcher_results/", trait, "_kmer_positions.rds"))
+
+# Extracting the signal at the location of the gene
+kmer_signals <- extract_signals(kmer_results,
+				threshold = as.numeric(readLines(paste0("gwas/kmers/", trait, "/kmers/threshold_5per"))),
+				distance = 10^4)
+
+# The next steps depend on whether there is a known gene model associated with the phenotype
+if(!is.na(gene_v1)) {
+
+	# Looking up the name of the gene in version 4
+	gene_v4 <- lookup_v1_to_v4[gene_v1]
+
+	# Getting the one signal associated with this gene
+	signal <- subsetByOverlaps(kmer_signals, genes[gene_v4])
+
+	if(!length(signal) == 1) {
+		stop("There is no signal overlapping the gene ", gene_v4, " for phenotype ", trait)
+	}
+
[…]

### 2022-07-29 figures/signal_plot.R
Add signal plots for several of the signals documented in Bandillo et al. (2017)

@@ -31,13 +31,10 @@ kmer_results <- readRDS(paste0("gwas/kmers/", trait, "/katcher_results/", trait,
 # Extracting the signal at the location of the gene
 kmer_signals <- extract_signals(kmer_results,
 				threshold = as.numeric(readLines(paste0("gwas/kmers/", trait, "/kmers/threshold_5per"))),
-				distance = 10^4)
+				distance = 10^5)
 
 # The next steps depend on whether there is a known gene model associated with the phenotype
-if(!is.na(gene_v1)) {
-
-	# Looking up the name of the gene in version 4
-	gene_v4 <- lookup_v1_to_v4[gene_v1]
+if(!is.na(gene_v1) && !is.na(gene_v4 <- lookup_v1_to_v4[gene_v1])) {
 
 	# Getting the one signal associated with this gene
 	signal <- subsetByOverlaps(kmer_signals, genes[gene_v4])

### 2022-08-01 figures/gene_plot.R
Add plotting of all signal and gene plots with integration into Makefile

@@ -7,15 +7,15 @@ id <- commandArgs(trailingOnly = TRUE)[1]
 programs <- c("platypus", "vg", "paragraph", "kmers")
 
 # Outputting to a PNG file
-png(paste0("figures/", id, "_signal.png"), width = 9, height = 10, units = "in", res = 300)
+png(paste0("figures/", id, "_gene.png"), width = 9, height = 10, units = "in", res = 300)
 
 # Resetting the viewport
 grid.newpage()
 
-# DEPENDENCY: Paragraph signal subplot
-# DEPENDENCY: vg signal subplot
-# DEPENDENCY: Platypus signal subplot
-# DEPENDENCY: kmers signal subplot
+# DEPENDENCY: Paragraph gene subplot
+# DEPENDENCY: vg gene subplot
+# DEPENDENCY: Platypus gene subplot
+# DEPENDENCY: kmers gene subplot
 
 # Setting a viewport with 4 rows, one for each of the programs
 pushViewport(viewport(layout = grid.layout(nrow = 4)))
@@ -23,7 +23,7 @@ pushViewport(viewport(layout = grid.layout(nrow = 4)))
 # Reading and printing each of the subplots in turn in their viewport row
 for(i in 1:4){
 	pushViewport(viewport(layout.pos.row = i))
-	grid.draw(readRDS(paste0("figures/ggplots/", programs[i], "_", id, "_signal.rds")))
+	grid.draw(readRDS(paste0("figures/ggplots/", programs[i], "_", id, "_gene.rds")))
 	grid.text(programs[i], 0.1, 0.9, hjust = 0)
 	upViewport()
 }

### 2022-08-01 figures/gene_subplot.R

@@ -1,0 +1,79 @@
+# Loading the required libraries
+library(gwastools)
+library(grid)
+library(GenomicRanges)
+
+# Getting the name of the trait-locus combination and of the program that called the genotypes from the command line
+id <- commandArgs(trailingOnly = TRUE)[1]
+program <- commandArgs(trailingOnly = TRUE)[2]
+
+# Getting the ID of the gene associated with the trait and the name of the trait from the utilities/loci.txt file
+# DEPENDENCY: utilities/loci.txt
+loci <- read.table("utilities/loci.txt", header = TRUE, sep = ",", stringsAsFactors = FALSE)
+gene_v1 <- loci[loci$id == id, "gene_v1"]
+trait <- loci[loci$id == id, "trait"]
+
+# The reference genome is needed for formatting the GWAS results from programs other than k-mers
+# DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
+refgenome <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
+
+# A vector of Glycine max chromosome names
+chromosomes <- paste0("Gm", ifelse(1:20 < 10, "0", ""), 1:20)
+### ----------
+
+# Loading the lookup table from Gmax v1 gene names to Gmax v4 gene names
+# DEPENDENCY: refgenome/lookup_v1_to_v4.rds
+lookup_v1_to_v4 <- readRDS("refgenome/lookup_v1_to_v4.rds")
+
+# Getting the set of genes, transcripts, exons and coding sequences
+# DEPENDENCY: refgenome/*.rds
+genes <- readRDS("refgenome/gmax_v4_genes.rds")
+transcripts <- readRDS("refgenome/gmax_v4_transcripts.rds")
+exons <- readRDS("refgenome/gmax_v4_exons.rds")
+cds <- readRDS("refgenome/gmax_v4_cds.rds")
+
+# Reading the data.frame of marker associations; treatment is different for k-mers than for other approaches
+# DEPENDENCY: GWAS association results
+# DEPENDENCY: GWAS thresholds
+if(program == "kmers") {
+	gwas_results <- readRDS(paste0("gwas_results/kmers/", trait, "_kmer_positions.rds"))
+	threshold <- as.numeric(readLines(paste0("gwas_results/kmers/", trait, "_threshold_5per")))
+} else if(program %in% c("platypus", "paragraph", "vg")) {
[…]

### 2022-08-01 figures/manhattan_subplot.R

@@ -44,5 +44,5 @@ if(program == "kmers") {
 }
 
 # Outputting to an rds file; actual plotting will be done in a different script
-saveRDS(gwas_plot, file = paste0("figures/ggplots/", program, "_", trait, "_manhattan.rds"))
+saveRDS(gwas_plot, file = paste0("figures/ggplots/", program, "_", trait, "_manhattan.rds"), compress = FALSE)
 

### 2022-08-01 figures/signal_subplot.R

@@ -1,0 +1,125 @@
+# Loading the required libraries
+library(gwastools)
+library(grid)
+library(GenomicRanges)
+
+# Getting the name of the trait-locus combination and of the program that called the genotypes from the command line
+id <- commandArgs(trailingOnly = TRUE)[1]
+program <- commandArgs(trailingOnly = TRUE)[2]
+
+# Getting the ID of the gene associatde with the trait and the name of the trait from the utilities/loci.txt file
+# DEPENDENCY: utilities/loci.txt
+loci <- read.table("utilities/loci.txt", header = TRUE, sep = ",", stringsAsFactors = FALSE)
+gene_v1 <- loci[loci$id == id, "gene_v1"]
+trait <- loci[loci$id == id, "trait"]
+
+# The reference genome is needed for formatting the GWAS results from programs other than k-mers
+# DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
+refgenome <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
+
+# A vector of Glycine max chromosome names
+chromosomes <- paste0("Gm", ifelse(1:20 < 10, "0", ""), 1:20)
+### ----------
+
+# Loading the lookup table from Gmax v1 gene names to Gmax v4 gene names
+# DEPENDENCY: refgenome/lookup_v1_to_v4.rds
+lookup_v1_to_v4 <- readRDS("refgenome/lookup_v1_to_v4.rds")
+
+# Getting the set of genes, transcripts, exons and coding sequences
+# DEPENDENCY: refgenome/*.rds
+genes <- readRDS("refgenome/gmax_v4_genes.rds")
+transcripts <- readRDS("refgenome/gmax_v4_transcripts.rds")
+exons <- readRDS("refgenome/gmax_v4_exons.rds")
+cds <- readRDS("refgenome/gmax_v4_cds.rds")
+
+# Reading the data.frame of marker associations; treatment is different for k-mers than for other approaches
+# DEPENDENCY: GWAS association results
+# DEPENDENCY: GWAS thresholds
+if(program == "kmers") {
+	gwas_results <- readRDS(paste0("gwas_results/kmers/", trait, "_kmer_positions.rds"))
+	threshold <- as.numeric(readLines(paste0("gwas_results/kmers/", trait, "_threshold_5per")))
+} else if(program %in% c("platypus", "paragraph", "vg")) {
[…]

### 2022-08-01 make_symlinks.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+for i in $(tail -n+2 utilities/loci.txt | cut -d "," -f1)
+do
+	trait=$(grep ^$i utilities/loci.txt | cut -d "," -f2)
+	
+	for j in platypus vg paragraph
+	do
+		cd gwas_results/${j}
+		ln -s ${trait}_gwas.csv ${i}_gwas.csv
+		ln -s ${trait}_threshold_5per.txt ${i}_threshold_5per.txt
+		cd ../..
+
+	done
+
+	cd gwas_results/kmers
+	ln -s ${trait}_kmer_positions.rds ${i}_kmer_positions.rds
+	ln -s ${trait}_threshold_5per ${i}_threshold_5per
+	cd ../..
+
+done
+

### 2022-08-03 figures/signal_subplot.R
Plot signals for all loci and all traits based on reference signals

@@ -1,17 +1,26 @@
 # Loading the required libraries
-library(gwastools)
 library(grid)
+library(gwastools)
 library(GenomicRanges)
 
 # Getting the name of the trait-locus combination and of the program that called the genotypes from the command line
 id <- commandArgs(trailingOnly = TRUE)[1]
 program <- commandArgs(trailingOnly = TRUE)[2]
 
-# Getting the ID of the gene associatde with the trait and the name of the trait from the utilities/loci.txt file
-# DEPENDENCY: utilities/loci.txt
-loci <- read.table("utilities/loci.txt", header = TRUE, sep = ",", stringsAsFactors = FALSE)
-gene_v1 <- loci[loci$id == id, "gene_v1"]
-trait <- loci[loci$id == id, "trait"]
+# A parameter that sets the padding of the x-scale when the signal consists of only one marker
+xscale_padding <- 500
+# A parameter that sets the distance between markers for merging signals with extract_signals
+signal_distance <- 250000
+
+# Getting the ID of the gene associated with the trait and the name of the trait from the all_signals.rds file
+# DEPENDENCY: utilities/all_signals.rds
+target_signal <- readRDS("utilities/all_signals.rds")
+target_signal <- target_signal[id]
+
+stopifnot(length(target_signal) == 1)
+
+gene_name <- target_signal$gene_name_v4
+trait <- target_signal$trait
 
 # The reference genome is needed for formatting the GWAS results from programs other than k-mers
 # DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
@@ -21,16 +30,23 @@ refgenome <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
 chromosomes <- paste0("Gm", ifelse(1:20 < 10, "0", ""), 1:20)
 ### ----------
 
-# Loading the lookup table from Gmax v1 gene names to Gmax v4 gene names
-# DEPENDENCY: refgenome/lookup_v1_to_v4.rds
-lookup_v1_to_v4 <- readRDS("refgenome/lookup_v1_to_v4.rds")
-
 # Getting the set of genes, transcripts, exons and coding sequences
 # DEPENDENCY: refgenome/*.rds
 genes <- readRDS("refgenome/gmax_v4_genes.rds")
 transcripts <- readRDS("refgenome/gmax_v4_transcripts.rds")
[…]

### 2022-08-03 make_symlinks.sh

@@ -20,3 +20,25 @@ do
 
 done
 
+for i in $(seq 1 $(wc -l utilities/signal_ids.txt | awk '{print $1}'))
+do
+	line=$(head -n $i utilities/signal_ids.txt | tail -n1)
+	id=$(echo $line | cut -d "," -f1)
+	trait=$(echo $line | cut -d "," -f2)
+	
+	for j in platypus vg paragraph
+	do
+		cd gwas_results/${j}
+		ln -s ${trait}_gwas.csv ${id}_gwas.csv
+		ln -s ${trait}_threshold_5per.txt ${id}_threshold_5per.txt
+		cd ../..
+
+	done
+
+	cd gwas_results/kmers
+	ln -s ${trait}_kmer_positions.rds ${id}_kmer_positions.rds
+	ln -s ${trait}_threshold_5per ${id}_threshold_5per
+	cd ../..
+
+done
+

### 2022-08-03 utilities/make_signals_granges.R

@@ -1,0 +1,32 @@
+# This script combines the reference signals from various sources into a GRanges object
+# that can be used to extract signals found by GWAS analyses and plot them
+
+library(GenomicRanges)
+
+bandillo_signals <- readRDS("utilities/bandillo_signals.rds")
+
+# We need to duplicate every entry with the various coded traits that were used in analyses
+trait_names <- readLines("utilities/trait_names.txt")
+
+all_signals <- lapply(trait_names,
+       FUN = function(x, signals) {
+	       signals <- signals[substring(x, 1, nchar(signals$trait)) == signals$trait]
+	       if(length(signals)) {
+		       signals$original_trait <- signals$trait
+		       signals$trait <- x
+		       signals$signal_id <- paste0(signals$trait, "_", signals$locus)
+		       names(signals) <- signals$signal_id
+	       }
+		       signals
+       },
+       signals = bandillo_signals)
+
+all_signals <- do.call("c", all_signals)
+stopifnot(!anyDuplicated(all_signals$signal_id))
+
+# Saving the list of signal names to file
+writeLines(paste(names(all_signals), all_signals$trait, sep = ","), con = "utilities/signal_ids.txt")
+
+# Saving the GRanges object to an .rds file
+saveRDS(all_signals, file = "utilities/all_signals.rds", compress = FALSE)
+

### 2022-08-03 figures/manhattan_subplot.R
Add vertical lines for reference signals to Manhattan plots

@@ -8,6 +8,11 @@ suppressMessages(library(GenomicRanges))
 trait <- commandArgs(trailingOnly = TRUE)[1]
 program <- commandArgs(trailingOnly = TRUE)[2]
 
+# Reading the reference signals for this trait
+signals <- readRDS("utilities/all_signals.rds")
+signals <- signals[signals$trait == trait]
+if(!length(signals)) signals <- NULL
+
 # The location of the reference genome fasta
 # DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
 refgenome <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
@@ -21,8 +26,9 @@ if(program == "kmers") {
 
 	gwas_plot <- manhattanGrob(gwas_results,
 				   threshold = as.numeric(readLines(paste0("gwas_results/kmers/", trait, "_threshold_5per"))),
+				   signals = signals,
 				   numeric_chrom = TRUE,
-				   margins = c(5.1, 3.1, 0.1, 0.1))
+				   margins = c(5.1, 4.1, 0.1, 0.1))
 
 } else if(program %in% c("platypus", "paragraph", "vg")) {
 	# Loading the results from the CSV file and formatting them as a GRanges object
@@ -34,8 +40,9 @@ if(program == "kmers") {
 	# Plotting the results using the gwastools::manhattan_plot function
 	gwas_plot <- manhattanGrob(gwas_results,
 				   threshold = -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", trait, "_threshold_5per.txt")))),
+				   signals = signals,
 				   numeric_chrom = TRUE,
-				   margins = c(5.1, 3.1, 0.1, 0.1))
+				   margins = c(5.1, 4.1, 0.1, 0.1))
 } else {
 	stop("Unrecognized program option")
 }

### 2022-08-04 figures/gene_plot.R
Generate gwas results separately from plotting scripts; integrate gene plots with the updated signal nomenclature

@@ -1,5 +1,5 @@
 # Loading the required packages
-library(grid)
+suppressMessages(library(grid))
 
 # Reading the ID being analyzed from the command line
 id <- commandArgs(trailingOnly = TRUE)[1]
@@ -7,7 +7,7 @@ id <- commandArgs(trailingOnly = TRUE)[1]
 programs <- c("platypus", "vg", "paragraph", "kmers")
 
 # Outputting to a PNG file
-png(paste0("figures/", id, "_gene.png"), width = 9, height = 10, units = "in", res = 300)
+png(paste0("figures/", id, "_gene.png"), width = 9, height = 10, units = "in", res = 100)
 
 # Resetting the viewport
 grid.newpage()
@@ -23,7 +23,7 @@ pushViewport(viewport(layout = grid.layout(nrow = 4)))
 # Reading and printing each of the subplots in turn in their viewport row
 for(i in 1:4){
 	pushViewport(viewport(layout.pos.row = i))
-	grid.draw(readRDS(paste0("figures/ggplots/", programs[i], "_", id, "_gene.rds")))
+	grid.draw(readRDS(paste0("figures/grobs/", programs[i], "_", id, "_gene.rds")))
 	grid.text(programs[i], 0.1, 0.9, hjust = 0)
 	upViewport()
 }

### 2022-08-04 figures/manhattan_plot.R

@@ -8,7 +8,7 @@ trait <- commandArgs(trailingOnly = TRUE)[1]
 programs <- c("platypus", "vg", "paragraph", "kmers")
 
 # Outputting to a PNG file
-png(paste0("figures/", trait, "_manhattan.png"), width = 9, height = 10, units = "in", res = 300)
+png(paste0("figures/", trait, "_manhattan.png"), width = 9, height = 10, units = "in", res = 100)
 
 # Resetting the viewport
 grid.newpage()
@@ -24,7 +24,7 @@ pushViewport(viewport(layout = grid.layout(nrow = 4)))
 # Reading and printing each of the subplots in turn in their viewport row
 for(i in 1:4){
 	pushViewport(viewport(layout.pos.row = i))
-	grid.draw(readRDS(paste0("figures/ggplots/", programs[i], "_", trait, "_manhattan.rds")))
+	grid.draw(readRDS(paste0("figures/grobs/", programs[i], "_", trait, "_manhattan.rds")))
 	grid.text(programs[i], 0.1, 0.9, hjust = 0)
 	upViewport()
 }

### 2022-08-04 figures/manhattan_subplot.R

@@ -9,44 +9,31 @@ trait <- commandArgs(trailingOnly = TRUE)[1]
 program <- commandArgs(trailingOnly = TRUE)[2]
 
 # Reading the reference signals for this trait
+# DEPENDENCY: utilities/all_signals.rds
 signals <- readRDS("utilities/all_signals.rds")
 signals <- signals[signals$trait == trait]
 if(!length(signals)) signals <- NULL
-
-# The location of the reference genome fasta
-# DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
-refgenome <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
 
 # A vector of Glycine max chromosome names
 chromosomes <- paste0("Gm", ifelse(1:20 < 10, "0", ""), 1:20)
 
 # It is a special case if we are analyzing k-mers
 if(program == "kmers") {
-	gwas_results <- readRDS(paste0("gwas_results/kmers/", trait, "_kmer_positions.rds"))
-
-	gwas_plot <- manhattanGrob(gwas_results,
-				   threshold = as.numeric(readLines(paste0("gwas_results/kmers/", trait, "_threshold_5per"))),
-				   signals = signals,
-				   numeric_chrom = TRUE,
-				   margins = c(5.1, 4.1, 0.1, 0.1))
-
-} else if(program %in% c("platypus", "paragraph", "vg")) {
-	# Loading the results from the CSV file and formatting them as a GRanges object
-	gwas_results <- format_gapit_gwas(filename = paste0("gwas_results/", program, "/", trait, "_gwas.csv"),
-					  ref_fasta = refgenome,
-					  chromosomes = chromosomes,
-					  pattern = "^Gm[0-9]{2}$")
-
-	# Plotting the results using the gwastools::manhattan_plot function
-	gwas_plot <- manhattanGrob(gwas_results,
-				   threshold = -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", trait, "_threshold_5per.txt")))),
-				   signals = signals,
-				   numeric_chrom = TRUE,
-				   margins = c(5.1, 4.1, 0.1, 0.1))
+	threshold <- as.numeric(readLines(paste0("gwas_results/kmers/", trait, "_threshold_5per")))
 } else {
-	stop("Unrecognized program option")
+	threshold <- -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", trait, "_threshold_5per.txt"))))
 }
 
[…]

### 2022-08-04 figures/signal_plot.R

@@ -7,7 +7,7 @@ id <- commandArgs(trailingOnly = TRUE)[1]
 programs <- c("platypus", "vg", "paragraph", "kmers")
 
 # Outputting to a PNG file
-png(paste0("figures/", id, "_signal.png"), width = 9, height = 10, units = "in", res = 300)
+png(paste0("figures/", id, "_signal.png"), width = 9, height = 10, units = "in", res = 100)
 
 # Resetting the viewport
 grid.newpage()
@@ -23,7 +23,7 @@ pushViewport(viewport(layout = grid.layout(nrow = 4)))
 # Reading and printing each of the subplots in turn in their viewport row
 for(i in 1:4){
 	pushViewport(viewport(layout.pos.row = i))
-	grid.draw(readRDS(paste0("figures/ggplots/", programs[i], "_", id, "_signal.rds")))
+	grid.draw(readRDS(paste0("figures/grobs/", programs[i], "_", id, "_signal.rds")))
 	grid.text(programs[i], 0.1, 0.9, hjust = 0)
 	upViewport()
 }

### 2022-08-04 figures/signal_subplot.R

@@ -12,7 +12,7 @@ xscale_padding <- 500
 # A parameter that sets the distance between markers for merging signals with extract_signals
 signal_distance <- 250000
 
-# Getting the ID of the gene associated with the trait and the name of the trait from the all_signals.rds file
+# Getting the ID of the gene associated with the locus
 # DEPENDENCY: utilities/all_signals.rds
 target_signal <- readRDS("utilities/all_signals.rds")
 target_signal <- target_signal[id]
@@ -20,11 +20,6 @@ target_signal <- target_signal[id]
 stopifnot(length(target_signal) == 1)
 
 gene_name <- target_signal$gene_name_v4
-trait <- target_signal$trait
-
-# The reference genome is needed for formatting the GWAS results from programs other than k-mers
-# DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
-refgenome <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
 
 # A vector of Glycine max chromosome names
 chromosomes <- paste0("Gm", ifelse(1:20 < 10, "0", ""), 1:20)
@@ -52,16 +47,13 @@ if(!is.na(gene_name) && grepl(";", gene_name)) {
 # DEPENDENCY: GWAS association results
 # DEPENDENCY: GWAS thresholds
 if(program == "kmers") {
-	gwas_results <- readRDS(paste0("gwas_results/kmers/", trait, "_kmer_positions.rds"))
-	threshold <- as.numeric(readLines(paste0("gwas_results/kmers/", trait, "_threshold_5per")))
+	threshold <- as.numeric(readLines(paste0("gwas_results/kmers/", id, "_locus_threshold_5per")))
 } else if(program %in% c("platypus", "paragraph", "vg")) {
-	# Loading the results from the CSV file and formatting them as a GRanges object
-	gwas_results <- format_gapit_gwas(filename = paste0("gwas_results/", program, "/", trait, "_gwas.csv"),
-					  ref_fasta = refgenome,
-					  chromosomes = chromosomes,
-					  pattern = "^Gm[0-9]{2}$")
-	threshold <- -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", trait, "_threshold_5per.txt"))))
+	threshold <- -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", id, "_locus_threshold_5per.txt"))))
 }
+
+# Reading the GWAS results from the RDS file
[…]

### 2022-08-04 make_symlinks.sh

@@ -1,44 +1,29 @@
 #!/bin/bash
 
-for i in $(tail -n+2 utilities/loci.txt | cut -d "," -f1)
-do
-	trait=$(grep ^$i utilities/loci.txt | cut -d "," -f2)
-	
-	for j in platypus vg paragraph
-	do
-		cd gwas_results/${j}
-		ln -s ${trait}_gwas.csv ${i}_gwas.csv
-		ln -s ${trait}_threshold_5per.txt ${i}_threshold_5per.txt
-		cd ../..
+# Creating the symlinks from the main results in ~/sv_gwas/usda_lines/gwas to the local gwas_results directory
 
+# First for the three genotyping programs
+for i in platypus vg paragraph
+do
+	mkdir -p gwas_results/${i}
+	cd gwas_results/${i}
+
+	for trait in $(cat ../../utilities/trait_names.txt)
+	do
+		ln -s ~/sv_gwas/usda_lines/gwas/${i}/${trait}/GAPIT.MLM.${trait}.GWAS.Results.csv ${trait}_gwas.csv
+		ln -s ~/sv_gwas/usda_lines/gwas/${i}/${trait}/${trait}_threshold_5per.txt
 	done
 
-	cd gwas_results/kmers
-	ln -s ${trait}_kmer_positions.rds ${i}_kmer_positions.rds
-	ln -s ${trait}_threshold_5per ${i}_threshold_5per
 	cd ../..
-
 done
 
-for i in $(seq 1 $(wc -l utilities/signal_ids.txt | awk '{print $1}'))
+# And then for the kmers approach
+mkdir -p gwas_results/kmers
+cd gwas_results/kmers
+
+for trait in $(cat ../../utilities/trait_names.txt)
 do
-	line=$(head -n $i utilities/signal_ids.txt | tail -n1)
-	id=$(echo $line | cut -d "," -f1)
-	trait=$(echo $line | cut -d "," -f2)
-	
-	for j in platypus vg paragraph
-	do
-		cd gwas_results/${j}
-		ln -s ${trait}_gwas.csv ${id}_gwas.csv
-		ln -s ${trait}_threshold_5per.txt ${id}_threshold_5per.txt
-		cd ../..
-
-	done
-
-	cd gwas_results/kmers
-	ln -s ${trait}_kmer_positions.rds ${id}_kmer_positions.rds
-	ln -s ${trait}_threshold_5per ${id}_threshold_5per
[…]

### 2022-08-04 utilities/make_signals_granges.R

@@ -27,6 +27,10 @@ stopifnot(!anyDuplicated(all_signals$signal_id))
 # Saving the list of signal names to file
 writeLines(paste(names(all_signals), all_signals$trait, sep = ","), con = "utilities/signal_ids.txt")
 
+# Also saving a list of signals for which the causal gene is known
+gene_signals <- all_signals[!is.na(all_signals$gene_name_v4)]
+writeLines(names(gene_signals), con = "utilities/gene_signal_ids.txt")
+
 # Saving the GRanges object to an .rds file
 saveRDS(all_signals, file = "utilities/all_signals.rds", compress = FALSE)
 

### 2022-08-10 figures/gene_plot.R
Plot signals on the same x-scale for all programs; only a single transcriptsGrob per plot; allow for several signals to be plotted with different colors for a given reference signal

@@ -1,10 +1,55 @@
 # Loading the required packages
 suppressMessages(library(grid))
+suppressMessages(library(gwastools))
+suppressMessages(library(GenomicRanges))
 
 # Reading the ID being analyzed from the command line
 id <- commandArgs(trailingOnly = TRUE)[1]
 
-programs <- c("platypus", "vg", "paragraph", "kmers")
+# Reading the information on genes and transcripts
+genes <- readRDS("refgenome/gmax_v4_genes.rds")
+transcripts <- readRDS("refgenome/gmax_v4_transcripts.rds")
+cds <- readRDS("refgenome/gmax_v4_cds.rds")
+exons <- readRDS("refgenome/gmax_v4_exons.rds")
+
+# Getting the ID of the gene associated with the trait and the name of the trait from the all_signals.rds file
+# DEPENDENCY: utilities/all_signals.rds
+target_signal <- readRDS("utilities/all_signals.rds")
+target_signal <- target_signal[id]
+
+gene_name <- target_signal$gene_name_v4
+### ----------
+
+# Handling the special case where the gene_name value contains more than one gene or no gene at all
+if(!is.na(gene_name) && grepl(";", gene_name)) {
+	gene_names <- strsplit(gene_name, ";")[[1]]
+	gene <- genes[gene_names]
+	gene <- reduce(gene, min.gapwidth = 10^6, ignore.strand = TRUE)
+} else if (!is.na(gene_name)) {
+	gene <- genes[gene_name]
+} else {
+	gene <- NULL
+}
+
+if(is.null(gene)) stop("No gene found where there should be one")
+
+# DEPENDENCY: Paragraph gene subplot
+# DEPENDENCY: vg gene subplot
+# DEPENDENCY: Platypus gene subplot
+# DEPENDENCY: kmers gene subplot
+pvalue_grobs <- list(readRDS(paste0("figures/grobs/platypus_", id, "_gene.rds")),
+		     readRDS(paste0("figures/grobs/vg_", id, "_gene.rds")),
+		     readRDS(paste0("figures/grobs/paragraph_", id, "_gene.rds")),
+		     readRDS(paste0("figures/grobs/kmers_", id, "_gene.rds")))
+
+output_grob <- pvalue_tx_grob(pvalue_grobs = pvalue_grobs,
+			      xrange = gene,
+			      genes = genes,
+			      transcripts = transcripts,
[…]

### 2022-08-10 figures/gene_subplot.R

@@ -49,7 +49,6 @@ if(!length(signal) == 1) warning("There is no signal overlapping the gene ", gen
 ptx_plot <- pvalueGrob(gwas_results = gwas_results,
 		       interval = gene,
 		       feature = NULL,
-		       pvalue_margins = c(5.1, 3.6, 0.5, 1.1),
 		       yexpand = c(0.1, 0.1))
 
 # Saving the grob to an RDS file for retrieval later on

### 2022-08-10 figures/signal_plot.R

@@ -1,10 +1,53 @@
 # Loading the required packages
 suppressMessages(library(grid))
+suppressMessages(library(gwastools))
+suppressMessages(library(GenomicRanges))
 
 # Reading the ID being analyzed from the command line
 id <- commandArgs(trailingOnly = TRUE)[1]
 
-programs <- c("platypus", "vg", "paragraph", "kmers")
+# Reading the information on genes and transcripts
+genes <- readRDS("refgenome/gmax_v4_genes.rds")
+transcripts <- readRDS("refgenome/gmax_v4_transcripts.rds")
+cds <- readRDS("refgenome/gmax_v4_cds.rds")
+exons <- readRDS("refgenome/gmax_v4_exons.rds")
+
+# Getting the ID of the gene associated with the trait and the name of the trait from the all_signals.rds file
+# DEPENDENCY: utilities/all_signals.rds
+target_signal <- readRDS("utilities/all_signals.rds")
+target_signal <- target_signal[id]
+
+gene_name <- target_signal$gene_name_v4
+
+# Handling the special case where the gene_name value contains more than one gene or no gene at all
+if(!is.na(gene_name) && grepl(";", gene_name)) {
+	gene_names <- strsplit(gene_name, ";")[[1]]
+	gene <- genes[gene_names]
+	gene <- reduce(gene, min.gapwidth = 10^6, ignore.strand = TRUE)
+} else if (!is.na(gene_name)) {
+	gene <- genes[gene_name]
+} else {
+	gene <- NULL
+}
+
+# DEPENDENCY: Paragraph signal subplot
+# DEPENDENCY: vg signal subplot
+# DEPENDENCY: Platypus signal subplot
+# DEPENDENCY: kmers signal subplot
+pvalue_grobs <- list(readRDS(paste0("figures/grobs/platypus_", id, "_signal.rds")),
+		     readRDS(paste0("figures/grobs/vg_", id, "_signal.rds")),
+		     readRDS(paste0("figures/grobs/paragraph_", id, "_signal.rds")),
+		     readRDS(paste0("figures/grobs/kmers_", id, "_signal.rds")))
+
+output_grob <- pvalue_tx_grob(pvalue_grobs = pvalue_grobs,
+			      xrange = NULL,
+			      xchrom = seqnames(target_signal),
+			      genes = genes,
+			      transcripts = transcripts,
+			      exons = exons,
+			      cds = cds,
[…]

### 2022-08-10 figures/signal_subplot.R

@@ -45,14 +45,9 @@ signal <- subsetByOverlaps(gwas_signals, target_signal)
 if(!length(signal)) {
 	warning("No signal found by ", program, " for signal ID ", id)
 	ptx_plot <- grid::gTree(children = gList(textGrob(paste0("No signal found by ", program, " for signal ID ", id))))
-
-} else if (length(signal) > 1) {
-	warning(length(signal), " signals found by ", program, " for signal ID ", id)
-	ptx_plot <- grid::gTree(children = gList(textGrob(paste0(length(signal), " signals found by ", program, " for signal ID ", id))))
-
 } else {
 	# Setting the x-scale while handling special cases where there is no signal or the signal is only 1 bp wide
-	if(width(signal) == 1) {
+	if(length(signal) == 1 && width(signal) == 1) {
 		start(signal) <- start(signal) - xscale_padding
 		end(signal)   <- end(signal) + xscale_padding
 	}
@@ -61,7 +56,6 @@ if(!length(signal)) {
 	ptx_plot <- pvalueGrob(gwas_results = gwas_results,
 			       interval = signal,
 			       feature = gene,
-			       pvalue_margins = c(5.1, 3.6, 0.5, 1.1),
 			       yexpand = c(0.1, 0.1))
 }
 

### 2022-08-15 reference_signals/format_bandillo2015_signals.R
Add reference signals for oil, protein and corrected dry weight. Exclude irrelevant figures from additional file 1 (but still generate all figures in case we eventually do need them)

@@ -1,0 +1,167 @@
+# This script formats the signals found for oil and protein by Bandillo et al. (2015)
+# as GRanges objects with coordinates relative to Williams82 genome assembly version 4
+
+# Loading required libraries
+library(parallel)
+library(GenomicRanges)
+library(Rsamtools)
+library(Biostrings)
+
+# The padding (number of nucleotides to be extracted from either side of the position) to use as a parameter
+padding <- 20
+
+# The path to reference genome version 1
+gmax_v1_path <- "~/refgenome/Gmax_v1/glyma.Wm82.gnm1.FCtY.genome_main.fna"
+gmax_v4_path <- "~/refgenome/Gmax_v4/Gmax_508_v4.0_mit_chlp.fasta"
+
+bandillo2015_snps <- read.table("bandillo2015_table1_curated.csv", sep = ",", header = TRUE,
+			       na.strings = "ns", stringsAsFactors = FALSE)
+
+# Transforming the dataset from a "wide" dataset to a "long" dataset
+oil_snps <- protein_snps <- bandillo2015_snps
+
+oil_snps$trait <- "oil"
+oil_snps$protein_logp <- oil_snps$protein_allelic_effect <- NULL
+oil_snps <- oil_snps[, c("trait", "SNP", "Chromosome", "oil_logp", "oil_allelic_effect")]
+colnames(oil_snps) <- c("trait", "SNP", "Chr", "logp", "allelic_effect")
+
+protein_snps$trait <- "protein"
+protein_snps$oil_logp <- protein_snps$oil_allelic_effect <- NULL
+protein_snps <- protein_snps[, c("trait", "SNP", "Chromosome", "protein_logp", "protein_allelic_effect")]
+colnames(protein_snps) <- c("trait", "SNP", "Chr", "logp", "allelic_effect")
+
+bandillo2015_snps <- rbind(oil_snps, protein_snps)
+
+# Keeping only associations of interest on chromsomes 15 and 20, and removing those with missing data
+bandillo2015_snps <- bandillo2015_snps[complete.cases(bandillo2015_snps) & bandillo2015_snps$Chr %in% c(15, 20),]
+
+# Formatting a few columns
+bandillo2015_snps$Chr <- paste0("glyma.Wm82.gnm1.Gm", bandillo2015_snps$Chr)
+bandillo2015_snps$signal_id <- paste0(bandillo2015_snps$trait, "_", bandillo2015_snps$Chr)
[…]

### 2022-08-15 utilities/make_signals_granges.R

@@ -3,7 +3,22 @@
 
 library(GenomicRanges)
 
-bandillo_signals <- readRDS("utilities/bandillo_signals.rds")
+# DEPENDENCY: reference_signals/bandillo2015_signals.rds
+# DEPENDENCY: reference_signals/bandillo2017_signals.rds
+bandillo2015_signals <- readRDS("reference_signals/bandillo2015_signals.rds")
+bandillo2017_signals <- readRDS("reference_signals/bandillo2017_signals.rds")
+
+# Manually entering data for the signal found by de Ronne et al. (2021) with BLINK on Gm15
+deronne_signal <- GRanges(seqnames = "Gm15",
+			  ranges = IRanges(start = 36764744, end = 36764744))
+deronne_signal$signal_id <- "cdwGm15"
+deronne_signal$trait <- "corrected_dry_weight"
+deronne_signal$locus <- "cdwGm15"
+deronne_signal$n_snps <- 1
+deronne_signal$log_pvalue <- -log10(7.1 * 10^-13)
+deronne_signal$gene_name_v4 <- NA
+deronne_signal$common_name <- NA
+names(deronne_signal) <- deronne_signal$signal_id
 
 # We need to duplicate every entry with the various coded traits that were used in analyses
 trait_names <- readLines("utilities/trait_names.txt")
@@ -19,7 +34,7 @@ all_signals <- lapply(trait_names,
 	       }
 		       signals
        },
-       signals = bandillo_signals)
+       signals = c(deronne_signal, bandillo2015_signals, bandillo2017_signals))
 
 all_signals <- do.call("c", all_signals)
 stopifnot(!anyDuplicated(all_signals$signal_id))

### 2022-09-07 additional_files/additional_file_1.tex
Add signal plots for several of the signals documented in Bandillo et al. (2017)
Add plotting of all signal and gene plots with integration into Makefile
Plot signals for all loci and all traits based on reference signals
Add vertical lines for reference signals to Manhattan plots
Generate gwas results separately from plotting scripts; integrate gene plots with the updated signal nomenclature
Plot signals on the same x-scale for all programs; only a single transcriptsGrob per plot; allow for several signals to be plotted with different colors for a given reference signal
Add reference signals

@@
 \usepackage{capt-of} % put caption on next page
+\usepackage{listings}
 
@@
 
+% Defining a command to generate a table of phenotypic data for a single gwas
+\newcommand{\gwastableone}[1]{
+\begin{table}
+	\caption{Numerical coding of phenotypic data used for the GWAS analysis of #1}
+	\label{#1_gwas_table}
+
+	\centering
+	\begin{threeparttable}\small
+
+		\csvreader[head to column names,
+		/csv/separator=semicolon,
+		tabular = llc,
+		table head = \toprule Phenotype & Frequency & GWAS \#1 \\ \midrule,
+		table foot = \bottomrule]%
+		{../tables/#1_gwas_table.csv}%
+		{}%
+		{\Value & \Frequency & \GWASa}
+
+			   \begin{tablenotes}
+			   \item[a] test
+			   \end{tablenotes}
+	\end{threeparttable}
+
+\end{table}
+\clearpage
+}
+
+% Defining a command to generate a table of phenotypic data for two gwas
+\newcommand{\gwastabletwo}[1]{
+\begin{table}
+	\caption{Numerical coding of phenotypic data used for the GWAS analysis of #1}
+	\label{#1_gwas_table}
+
+	\centering
+	\begin{threeparttable}\small
+
+		\csvreader[head to column names,
+		/csv/separator=semicolon,
+		tabular = llcc,
+		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2\\ \midrule,
+		table foot = \bottomrule]%
+		{../tables/#1_gwas_table.csv}%
+		{}%
+		{\Value & \Frequency & \GWASa & \GWASb}
+
+			   \begin{tablenotes}
+			   \item[a] test
+			   \end{tablenotes}
+	\end{threeparttable}
+
+\end{table}
+\clearpage
+}
+
+% Defining a command to generate a table of phenotypic data for three gwas
+\newcommand{\gwastablethree}[1]{
+\begin{table}
+	\caption{Numerical coding of phenotypic data used for the GWAS analysis of #1}
+	\label{#1_gwas_table}
+
+	\centering
+	\begin{threeparttable}\small
+
+		\csvreader[head to column names,
+		/csv/separator=semicolon,
+		tabular = llccc,
+		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2 & GWAS \# 3\\ \midrule,
+		table foot = \bottomrule]%
+		{../tables/#1_gwas_table.csv}%
+		{}%
+		{\Value & \Frequency & \GWASa & \GWASb & \GWASc}
+
+			   \begin{tablenotes}
+			   \item[a] test
+			   \end{tablenotes}
+	\end{threeparttable}
+
+\end{table}
+\clearpage
+}
+
 % Defining a command for manhattan plot figures
@@
 \begin{cfigure}
-	\caption[Figure caption]{Figure caption}
 	\includegraphics[width = 6in]{#1_manhattan}
+	\caption{#1\_manhattan}
 	\label{#1_manhattan}
@@
 \begin{cfigure}
-	\caption[Figure caption]{Figure caption}
 	\includegraphics[width = 6in]{#1_signal}
+	\caption{#1\_signal}
 	\label{#1_signal}
@@
 \clearpage}
+
+\newcommand{\geneplot}[1]{
+\begin{cfigure}
+	\includegraphics[width = 6in]{#1_gene}
+	\caption{#1\_gene}
+	\label{#1_gene}
+\end{cfigure}
+
+\clearpage}
+
+% Allows underscores in captions but is not a great solution
+\catcode`_=12
 
@@
 
-% \caption[Number of SVs per Oxford Nanopore sample]{{{{
+\gwastableone{FLOWER.COLOR}
+\gwastabletwo{PUBESCENCE.COLOR}
+\gwastabletwo{SEED.COAT.COLOR}
+\gwastabletwo{STEM.TERMINATION.TYPE}
+\gwastablethree{HILUM.COLOR}
+\gwastabletwo{POD.COLOR}
+\gwastabletwo{PUBESCENCE.FORM}
+\gwastableone{PUBESCENCE.DENSITY}
+\gwastablethree{SEED.COAT.LUSTER}
+\gwastableone{MATURITY.GROUP}
 
-%\begin{table}
-%	\caption[Number of SVs per Oxford Nanopore sample]{}
-%	\label{table_s1}
-%
-%	\begin{threeparttable}\small
-%			   \begin{tablenotes}
-%			   \end{tablenotes}
-%	\end{threeparttable}
-%
-%\end{table}
-%\clearpage%}}}
+\manhattanplot{flower_color}
+\signalplot{flower_color_W1}
+\geneplot{flower_color_W1}
+%\signalplot{flower_color_L1}
+%\geneplot{flower_color_L1}
 
-% Manhattan plots of all traits
+\manhattanplot{pubescence_color_all}
+%\signalplot{pubescence_color_all_Td}
+\signalplot{pubescence_color_all_T}
+\geneplot{pubescence_color_all_T}
+
+\manhattanplot{pubescence_color_nogray}
+\signalplot{pubescence_color_nogray_Td}
+%\signalplot{pubescence_color_nogray_T}
+%\geneplot{pubescence_color_nogray_T}
+
+\manhattanplot{seed_coat_color_all}
+\signalplot{seed_coat_color_all_G}
+\signalplot{seed_coat_color_all_T}
+\geneplot{seed_coat_color_all_T}
+\signalplot{seed_coat_color_all_O}
+\signalplot{seed_coat_color_all_I}
+\geneplot{seed_coat_color_all_I}
+%\signalplot{seed_coat_color_all_R}
+%\geneplot{seed_coat_color_all_R}
+
+\manhattanplot{seed_coat_color_greenyellow}
+\signalplot{seed_coat_color_greenyellow_G}
+%\signalplot{seed_coat_color_greenyellow_T}
+%\geneplot{seed_coat_color_greenyellow_T}
+%\signalplot{seed_coat_color_greenyellow_O}
+%\signalplot{seed_coat_color_greenyellow_I}
+%\geneplot{seed_coat_color_greenyellow_I}
+%\signalplot{seed_coat_color_greenyellow_R}
+%\geneplot{seed_coat_color_greenyellow_R}
+
 \manhattanplot{stem_termination_all}
+%\signalplot{stem_termination_all_Dt2}
+%\geneplot{stem_termination_all_Dt2}
+\signalplot{stem_termination_all_Dt1}
+\geneplot{stem_termination_all_Dt1}
+%\signalplot{stem_termination_all_E3}
+%\geneplot{stem_termination_all_E3}
+
 \manhattanplot{stem_termination_sn}
-\manhattanplot{pubescence_color_all}
-\manhattanplot{pubescence_color_nogray}
+%\signalplot{stem_termination_sn_Dt2}
+%\geneplot{stem_termination_sn_Dt2}
+\signalplot{stem_termination_sn_Dt1}
+\geneplot{stem_termination_sn_Dt1}
+%\signalplot{stem_termination_sn_E3}
+%\geneplot{stem_termination_sn_E3}
+
+\manhattanplot{hilum_color_all}
+\signalplot{hilum_color_all_T}
+\geneplot{hilum_color_all_T}
+%\signalplot{hilum_color_all_O}
+\signalplot{hilum_color_all_I}
+\geneplot{hilum_color_all_I}
+\signalplot{hilum_color_all_R}
+\geneplot{hilum_color_all_R}
+%\signalplot{hilum_color_all_W1}
+%\geneplot{hilum_color_all_W1}
+
+\manhattanplot{hilum_color_blackbrown}
+%\signalplot{hilum_color_blackbrown_T}
+%\geneplot{hilum_color_blackbrown_T}
+%\signalplot{hilum_color_blackbrown_O}
+%\signalplot{hilum_color_blackbrown_I}
+%\geneplot{hilum_color_blackbrown_I}
+\signalplot{hilum_color_blackbrown_R}
+\geneplot{hilum_color_blackbrown_R}
+%\signalplot{hilum_color_blackbrown_W1}
+%\geneplot{hilum_color_blackbrown_W1}
+
+\manhattanplot{hilum_color_rbr}
+%\signalplot{hilum_color_rbr_T}
+%\geneplot{hilum_color_rbr_T}
+%\signalplot{hilum_color_rbr_O}
+%\signalplot{hilum_color_rbr_I}
+%\geneplot{hilum_color_rbr_I}
+%\signalplot{hilum_color_rbr_R}
+%\geneplot{hilum_color_rbr_R}
+%\signalplot{hilum_color_rbr_W1}
+%\geneplot{hilum_color_rbr_W1}
+
+\manhattanplot{pod_color_all}
+\signalplot{pod_color_all_L2}
+\signalplot{pod_color_all_L1}
+\geneplot{pod_color_all_L1}
+
+\manhattanplot{pod_color_blbr}
+%\signalplot{pod_color_blbr_L2}
+\signalplot{pod_color_blbr_L1}
+\geneplot{pod_color_blbr_L1}
+
+\manhattanplot{pubescence_form_all}
+\signalplot{pubescence_form_all_Pa1}
+\signalplot{pubescence_form_all_Pa2}
+
+\manhattanplot{pubescence_form_noerect}
+\signalplot{pubescence_form_noerect_Pa1}
+%\signalplot{pubescence_form_noerect_Pa2}
+
+\manhattanplot{pubescence_density}
+%\signalplot{pubescence_density_Pd1}
+\signalplot{pubescence_density_Ps}
+%\signalplot{pubescence_density_P1}
+%\geneplot{pubescence_density_P1}
+
+\manhattanplot{seed_coat_luster_all}
+%\signalplot{seed_coat_luster_all_I}
+%\geneplot{seed_coat_luster_all_I}
+%\signalplot{seed_coat_luster_all_Bq}
+%\signalplot{seed_coat_luster_all_B1}
+\signalplot{seed_coat_luster_all_Hps}
+\geneplot{seed_coat_luster_all_Hps}
+\signalplot{seed_coat_luster_all_B}
+
+\manhattanplot{seed_coat_luster_nointermediate}
+%\signalplot{seed_coat_luster_nointermediate_I}
+%\geneplot{seed_coat_luster_nointermediate_I}
+%\signalplot{seed_coat_luster_nointermediate_Bq}
+%\signalplot{seed_coat_luster_nointermediate_B1}
+\signalplot{seed_coat_luster_nointermediate_Hps}
+\geneplot{seed_coat_luster_nointermediate_Hps}
+\signalplot{seed_coat_luster_nointermediate_B}
+
+\manhattanplot{seed_coat_luster_dullshiny}
+%\signalplot{seed_coat_luster_dullshiny_I}
+%\geneplot{seed_coat_luster_dullshiny_I}
+%\signalplot{seed_coat_luster_dullshiny_Bq}
+%\signalplot{seed_coat_luster_dullshiny_B1}
+\signalplot{seed_coat_luster_dullshiny_Hps}
+\geneplot{seed_coat_luster_dullshiny_Hps}
+\signalplot{seed_coat_luster_dullshiny_B}
+
+\manhattanplot{corrected_dry_weight}
+\signalplot{corrected_dry_weight_cdwGm17}
+
+\manhattanplot{maturity_group}
+%\signalplot{maturity_group_E1}
+%\geneplot{maturity_group_E1}
+%\signalplot{maturity_group_E2}
+%\geneplot{maturity_group_E2}
+%\signalplot{maturity_group_E3}
+%\geneplot{maturity_group_E3}
+%\signalplot{maturity_group_E4}
+%\geneplot{maturity_group_E4}
+
+\manhattanplot{oil}
+\signalplot{oil_oilGm15}
+%\signalplot{oil_oilGm20}
+
 \manhattanplot{protein}
-\manhattanplot{pubescence_form_all}
-\manhattanplot{pubescence_form_noerect}
-\manhattanplot{oil}
-\manhattanplot{pubescence_density}
-\manhattanplot{seed_coat_luster_all}
-\manhattanplot{seed_coat_luster_nointermediate}
-\manhattanplot{seed_coat_luster_dullshiny}
-\manhattanplot{pod_color_all}
-\manhattanplot{pod_color_blbr}
-\manhattanplot{flower_color}
-\manhattanplot{seed_coat_color_all}
-\manhattanplot{seed_coat_color_greenyellow}
-\manhattanplot{hilum_color_all}
-\manhattanplot{hilum_color_blackbrown}
-\manhattanplot{hilum_color_rbr}
-\manhattanplot{maturity_group}
-\manhattanplot{corrected_dry_weight}
-
-\signalplot{pca_t}
-\signalplot{fc_w}
-\signalplot{sda_dt1}
+%\signalplot{protein_proteinGm15}
+%\signalplot{protein_proteinGm20}
 

### 2022-09-07 phenotypic_data/lookup_tables.R

@@ -1,0 +1,102 @@
+# This script prepares the lookup tables used to prepare phenotypic data
+# for GWAS analyses
+stem_termination_all_lookup <- c("D – Determinate (stem abruptly terminating)" = 1,
+				 "S – Semi-determinate (intermediate between determinate and indeterminate)" = 2,
+				 "N – Indeterminate (stem tapering gradually toward tip)" = 3)
+
+stem_termination_sn_lookup <- c("S – Semi-determinate (intermediate between determinate and indeterminate)" = 1,
+				"N – Indeterminate (stem tapering gradually toward tip)" = 2)
+
+pubescence_color_all_lookup <- c("G – Gray" = 1,
+				"T – Tawny" = 2,
+				"Lt – Light tawny" = 3)
+
+pubescence_color_nogray_lookup <- c("T – Tawny" = 1,
+				   "Lt – Light tawny" = 2)
+
+pubescence_form_all_lookup <- c("A – Appressed (most hairs flat on leaf surface)" = 1,
+				"Sa – Semi-appressed (between erect and appressed)" = 2,
+				"E – Erect on leaf surface" = 3)
+
+pubescence_form_noerect_lookup <- c("A – Appressed (most hairs flat on leaf surface)" = 1,
+				    "Sa – Semi-appressed (between erect and appressed)" = 2)
+
+pubescence_density_lookup <- c("N – Normal density" = 1,
+			       "Ssp – Semi-sparse (slightly reduced density, especially on pulvinus)" = 2)
+
+seed_coat_luster_all_lookup <- c("B – Bloom (heavy coating of powdery substance adhering to the seed coat)" = 1,
+				 "D – Dull" = 2,
+				 "I – Intermediate (between dull and shiny)" = 3,
+				 "Lb – Light bloom (slight bloom)" = 4,
+				 "S – Shiny" = 5)
+
+seed_coat_luster_nointermediate_lookup <- c("B – Bloom (heavy coating of powdery substance adhering to the seed coat)" = 1,
+					    "D – Dull" = 2,
+					    "Lb – Light bloom (slight bloom)" = 3,
+					    "S – Shiny" = 4)
+
+seed_coat_luster_dullshiny_lookup <- c("D – Dull" = 1,
+				       "S – Shiny" = 2)
+
+pod_color_all_lookup <- c("Tn – Tan" = 1,
[…]

### 2022-09-07 phenotypic_data/trait_names.R

@@ -1,0 +1,18 @@
+# A script that creates a table with the correspondence between
+# trait names in the USDA GRIN database and the ones I used
+trait_names <- c("STEM.TERMINATION.TYPE" = "stem_termination",
+		 "PUBESCENCE.COLOR" = "pubescence_color",
+		 "PROTEIN" = "protein",
+		 "PUBESCENCE.FORM" = "pubescence_form",
+		 "OIL" = "oil",
+		 "PUBESCENCE.DENSITY" = "pubescence_density",
+		 "SEED.COAT.LUSTER" = "seed_coat_luster",
+		 "POD.COLOR" = "pod_color",
+		 "FLOWER.COLOR" = "flower_color",
+		 "SEED.COAT.COLOR" = "seed_coat_color",
+		 "HILUM.COLOR" = "hilum_color",
+		 "MATURITY.GROUP" = "maturity_group")
+
+# Saving the table as an .rds file for retrieval by R scripts
+saveRDS(trait_names, file = "phenotypic_data/trait_names.rds")
+

### 2022-09-08 reference_signals/format_bandillo2015_signals.R
Integrate the generation of signals with the Makefile; add genes for loci that have been cloned after Bandillo et al. (2017)

@@ -10,12 +10,22 @@ library(Biostrings)
 # The padding (number of nucleotides to be extracted from either side of the position) to use as a parameter
 padding <- 20
 
-# The path to reference genome version 1
-gmax_v1_path <- "~/refgenome/Gmax_v1/glyma.Wm82.gnm1.FCtY.genome_main.fna"
-gmax_v4_path <- "~/refgenome/Gmax_v4/Gmax_508_v4.0_mit_chlp.fasta"
+# --- GROUPING ALL EXTERNAL DEPENDENCIES AT THE TOP OF THE FILE
 
-bandillo2015_snps <- read.table("bandillo2015_table1_curated.csv", sep = ",", header = TRUE,
-			       na.strings = "ns", stringsAsFactors = FALSE)
+# The path to reference genome version 1
+# DEPENDENCY: refgenome/glyma.Wm82.gnm1.FCtY.genome_main.fna
+gmax_v1_path <- "refgenome/glyma.Wm82.gnm1.FCtY.genome_main.fna"
+
+# The path to reference genome version 4
+# DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
+gmax_v4_path <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
+
+# A curated table as obtained from the table in Bandillo et al. (2015) processed with Tabula
+# DOI: 10.3835/plantgenome2015.04.0024
+# DEPENDENCY: reference_signals/bandillo2015_table1_curated.csv
+bandillo2015_snps <- read.table("reference_signals/bandillo2015_table1_curated.csv",
+				sep = ",", header = TRUE,
+				na.strings = "ns", stringsAsFactors = FALSE)
 
 # Transforming the dataset from a "wide" dataset to a "long" dataset
 oil_snps <- protein_snps <- bandillo2015_snps
@@ -55,6 +65,10 @@ bandillo2015_signals$glyma_name <- NA
 bandillo2015_signals$common_name <- NA
 bandillo2015_signals$gene_name_v4 <- NA
 
+# Setting the name of the gene for the QTL on chromosome 15
+# according to Zhang et al. (2020), DOI:10.1371/journal.pgen.1009114
+bandillo2015_signals[bandillo2015_signals$Chr == "glyma.Wm82.gnm1.Gm15", "gene_name_v4"] <- tolower("Glyma.15G049200")
+
 # Reordering the columns
[…]

### 2022-09-08 reference_signals/format_bandillo2017_signals.R

@@ -1,0 +1,306 @@
+# This script translates the positions of the signals found by Bandillo et al. (2017)
+# from reference genome version 1 to reference genome version 4
+#
+# This will imply retrieving the positions of the flanking SNPs and maximal SNPs
+# in their dataset, extracting the flanking sequences in reference genome version 1
+# and look for matches in reference genome version 4
+#
+# As part of this script we also set the names of the genes that are associated
+# with various loci based on literature. The references to the literature can be
+# found in the comments.
+
+# Loading the required libraries
+library(parallel)
+library(Rsamtools)
+library(GenomicRanges)
+library(Biostrings)
+
+# The padding (number of nucleotides to be extracted from either side of the position) to use as a parameter
+padding <- 20
+
+# --- GATHERING ALL EXTERNAL DEPENDENCIES AT THE TOP OF THE FILE
+
+# The path to reference genome version 1
+# DEPENDENCY: refgenome/glyma.Wm82.gnm1.FCtY.genome_main.fna
+gmax_v1_path <- "refgenome/glyma.Wm82.gnm1.FCtY.genome_main.fna"
+
+# The path to reference genome version 4
+# DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta
+gmax_v4_path <- "refgenome/Gmax_508_v4.0_mit_chlp.fasta"
+
+# The signals detected by Bandillo et al. (2017), as obtained from the article PDF with Tabula
+# DOI: 10.3835/plantgenome2016.06.0054
+# DEPENDENCY: reference_signals/bandillo2017_signals_curated.tsv
+bandillo_signals <- read.table("reference_signals/bandillo2017_signals_curated.tsv",
+			       sep = "\t", header = TRUE,
+			       na.strings = "-", stringsAsFactors = FALSE)
+
+# Loading a lookup table of gene names from Williams82 genome assembly version 1 to version 4
+# DEPENDENCY: refgenome/lookup_v1_to_v4.rds
+gene_name_lookup <- readRDS("refgenome/lookup_v1_to_v4.rds")
+
+# Getting the names of the genes that are Chalcone synthase genes on chromosome Gm08 in version 4
+# DEPENDENCY: refgenome/soybase_genome_annotation_v4.0_04-20-2021.txt
[…]

### 2022-09-08 reference_signals/format_deronne2022_signal.R

@@ -1,0 +1,22 @@
+# This script creates the signal found by de Ronne et al. (2022)
+# as a GRanges object for use in downstream analyses
+# DOI:10.1002/tpg2.20184
+
+# Loading the GenomicRanges package
+library(GenomicRanges)
+
+# Manually entering data for the signal found by de Ronne et al. (2021) with BLINK on Gm15
+deronne_signal <- GRanges(seqnames = "Gm15",
+			  ranges = IRanges(start = 36395777, end = 36895690))
+deronne_signal$signal_id <- "cdwGm15"
+deronne_signal$trait <- "corrected_dry_weight"
+deronne_signal$locus <- "cdwGm15"
+deronne_signal$n_snps <- 31
+deronne_signal$log_pvalue <- -log10(7.1 * 10^-13)
+deronne_signal$gene_name_v4 <- tolower("Glyma.15G217100")
+deronne_signal$common_name <- NA
+names(deronne_signal) <- deronne_signal$signal_id
+
+# Saving this signal to file for retrieval in downstream analyses
+saveRDS(deronne_signal, "reference_signals/deronne2022_signal.rds")
+

### 2022-09-08 utilities/make_signals_granges.R

@@ -5,20 +5,10 @@ library(GenomicRanges)
 
 # DEPENDENCY: reference_signals/bandillo2015_signals.rds
 # DEPENDENCY: reference_signals/bandillo2017_signals.rds
+# DEPENDENCY: reference_signals/deronne2022_signal.rds
 bandillo2015_signals <- readRDS("reference_signals/bandillo2015_signals.rds")
 bandillo2017_signals <- readRDS("reference_signals/bandillo2017_signals.rds")
-
-# Manually entering data for the signal found by de Ronne et al. (2021) with BLINK on Gm15
-deronne_signal <- GRanges(seqnames = "Gm15",
-			  ranges = IRanges(start = 36764744, end = 36764744))
-deronne_signal$signal_id <- "cdwGm15"
-deronne_signal$trait <- "corrected_dry_weight"
-deronne_signal$locus <- "cdwGm15"
-deronne_signal$n_snps <- 1
-deronne_signal$log_pvalue <- -log10(7.1 * 10^-13)
-deronne_signal$gene_name_v4 <- NA
-deronne_signal$common_name <- NA
-names(deronne_signal) <- deronne_signal$signal_id
+deronne_signal <- readRDS("reference_signals/deronne2022_signal.rds")
 
 # We need to duplicate every entry with the various coded traits that were used in analyses
 trait_names <- readLines("utilities/trait_names.txt")

### 2022-09-09 figures/kmer_plot.R
Add functionality for plotting the haplotypes and variants deduced from the k-mer analysis; work in progress

@@ -1,0 +1,313 @@
+# Loading the required libraries
+library(grid)
+library(Biostrings)
+library(IRanges)
+library(stringr)
+
+max_kmers <- 10000
+kmer_length <- 31
+
+# A function that maps a set of numeric values onto a color palette
+# Values at the minimum are set to black
+map_color <- function(values, max, min = 0, pal = "YlOrRd", n.colors = 9) {
+	# Breaking the numeric values into a factor
+	index <- cut(values, breaks = seq(min, max, length.out = n.colors))
+	palette <- RColorBrewer::brewer.pal(n = n.colors, name = pal)
+	output <- palette[as.integer(index)]
+	# Setting the values at the minimum to black
+	output[values <= min] <- "black"
+	output
+}
+
+# This script generates a k-mer plot from the consensus sequences of a group of samples
+locus <- commandArgs(trailingOnly = TRUE)[1]
+locus_data <- read.table("utilities/kmer_plot_ranges.txt")
+trait <- locus_data[locus_data[[1]] == locus, 2]
+
+# DEPENDENCY: phenotypic_data/phenotypic_data.csv
+phenotypes <- read.table("phenotypic_data/phenotypic_data.csv",
+			 sep = ";", header = TRUE,)
+
+# DEPENDENCY: phenotypic_data/lookup_tables.rds
+lookup_tables <- readRDS("phenotypic_data/lookup_tables.rds")
+lookup_table  <- lookup_tables[[paste0(trait, "_lookup")]]
+
+# DEPENDENCY: k-mer p-values
+kmer_pvalues <- read.table(paste0("gwas_results/kmer_data/", trait, "/katcher_results/pass_threshold_5per_sorted.txt"),
+			   colClasses = c("NULL", "character", rep("NULL", 6), "numeric"),
+			   nrows = max_kmers)
+names(kmer_pvalues) <- c("kmer_id", "pvalue")
+
+kmer_pvalues$kmer <- sub("[^ATGC]+", "", kmer_pvalues$kmer_id)
+kmer_pvalues$kmer_reverse <- as.character(reverseComplement(DNAStringSet(kmer_pvalues$kmer)))
+
+# DEPENDENCY: consensus sequence for a given locus
+fasta_file <- readLines(paste0("gwas_results/kmer_consensus/", locus, "_sequences.fa"))
+
+# Using conensus sequences from bwa in case the assemblies did not yield great results
+if(length(fasta_file) < 200) {
[…]

### 2022-09-09 make_symlinks.sh

@@ -23,8 +23,20 @@ cd gwas_results/kmers
 
 for trait in $(cat ../../utilities/trait_names.txt)
 do
+	# Creating symlinks to the processed results of the k-mer analysis
 	ln -s ~/sv_gwas/usda_lines/gwas/kmers/${trait}/katcher_results/${trait}_kmer_positions.rds
 	ln -s ~/sv_gwas/usda_lines/gwas/kmers/${trait}/kmers/threshold_5per ${trait}_threshold_5per
+done
+
+cd ../..
+
+# Creating symlinks to the main directories of the k-mer analysis
+mkdir -p gwas_results/kmer_data
+cd gwas_results/kmer_data
+
+for trait in $(cat ../../utilities/trait_names.txt)
+do
+	ln -s ~/sv_gwas/usda_lines/gwas/kmers/${trait}
 done
 
 cd ../..

### 2022-09-16 figures/gene_subplot.R
Add manhattan plots for k-mers located on scaffolds; remove some figures from supplementary material; adjust some parameters in defining signal width and top p-value regions; adjust the signal coordinates for corrected dry weight; make some aesthetic changes to plots

@@ -49,6 +49,8 @@ if(!length(signal) == 1) warning("There is no signal overlapping the gene ", gen
 ptx_plot <- pvalueGrob(gwas_results = gwas_results,
 		       interval = gene,
 		       feature = NULL,
+		       col = "blue",
+		       pruned_col = if(program == "platypus") "firebrick2" else NULL,
 		       yexpand = c(0.1, 0.1))
 
 # Saving the grob to an RDS file for retrieval later on

### 2022-09-16 figures/manhattan_subplot.R

@@ -19,14 +19,16 @@ if(!length(signals)) signals <- NULL
 found_signals <- readRDS(paste0("gwas_results/", program, "/", trait, "_signal.rds"))
 if(!length(signals)) found_signals <- NULL
 
-# A vector of Glycine max chromosome names
-chromosomes <- paste0("Gm", ifelse(1:20 < 10, "0", ""), 1:20)
-
 # It is a special case if we are analyzing k-mers
 threshold <- -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", trait, "_threshold_5per.txt"))))
 
 # Loading the results from the rds file
 gwas_results <- readRDS(paste0("gwas_results/", program, "/", trait, "_gwas.rds"))
+
+# If the program is the k-mers, we filter out variants that are located on unachored scaffolds
+if(program == "kmers") {
+	GenomeInfoDb::seqlevels(gwas_results, pruning.mode = "coarse") <- grep("^Gm[0-9]{2}$", GenomeInfoDb::seqlevels(gwas_results), value = TRUE)
+}
 	
 # Plotting the results using the gwastools::manhattan_plot function
 gwas_plot <- manhattanGrob(gwas_results,

### 2022-09-16 figures/scaffolds_manhattan.R

@@ -1,0 +1,43 @@
+# Loading the required packages
+suppressMessages(library(grid))
+suppressMessages(library(gwastools))
+suppressMessages(library(GenomeInfoDb))
+suppressMessages(library(GenomicRanges))
+
+# Reading in the command line arguments; the first one is the trait, the second is the genotyping program
+trait <- commandArgs(trailingOnly = TRUE)[1]
+program <- commandArgs(trailingOnly = TRUE)[2]
+
+# Reading the threshold from file
+# DEPENDENCY: significance threshold
+threshold <- -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", trait, "_threshold_5per.txt"))))
+
+# Loading the results from the rds file
+# DEPENDENCY: GWAS results
+gwas_results <- readRDS(paste0("gwas_results/", program, "/", trait, "_gwas.rds"))
+
+# We filter out variants other than unachored scaffolds
+GenomeInfoDb::seqlevels(gwas_results, pruning.mode = "coarse") <- grep("^Gm[0-9]{2}$", GenomeInfoDb::seqlevels(gwas_results), value = TRUE, invert = TRUE)
+# And we keep only the reference sequences that are present
+GenomeInfoDb::seqlevels(gwas_results, pruning.mode = "coarse") <- seqlevels(gwas_results)[seqlevels(gwas_results) %in% seqnames(gwas_results)]
+
+# We first check if there are any markers left
+if(!length(gwas_results)) {
+	gwas_plot <- textGrob(paste0("No significant associations found for ", trait, " using ", program))
+} else {
+	# Plotting the results using the gwastools::manhattan_plot function
+	gwas_plot <- manhattanGrob(gwas_results,
+				   threshold = threshold,
+				   ref_signals = NULL,
+				   new_signals = NULL,
+				   numeric_chrom = FALSE,
+				   margins = c(5.1, 4.1, 0.1, 0.1))
+}
+
+# Plotting directly as we are only using a single program (the k-mers)
+# Outputting to a PNG file
+png(paste0("figures/", trait, "_scaffolds_manhattan.png"), width = 9, height = 3, units = "in", res = 100)
+grid.newpage()
+grid.draw(gwas_plot)
+dev.off()
+

### 2022-09-16 figures/signal_subplot.R

@@ -57,6 +57,8 @@ if(!length(signal)) {
 	ptx_plot <- pvalueGrob(gwas_results = gwas_results,
 			       interval = signal,
 			       feature = gene,
+			       col = "blue",
+			       pruned_col = if(program == "platypus") "firebrick2" else NULL,
 			       shading = top_markers,
 			       yexpand = c(0.1, 0.1))
 }

### 2022-09-16 gwas_results/gene_analysis.R

@@ -11,8 +11,13 @@ suppressMessages(library(GenomicRanges))
 id <- commandArgs(trailingOnly = TRUE)[1]
 program <- commandArgs(trailingOnly = TRUE)[2]
 
-# The distance at which signals overlapping the reference signal will be merged
-gapwidth <- 10^6
+#----- ANALYSIS PARAMETERS
+gapwidth <- 10^6 # The distance at which signals overlapping the reference signal will be merged
+fractions <- c("platypus" = 0.01, # the fraction of top k-mers considered to be delimiting the top region
+	       "vg" = 0.05,       # depends on the approach used given the size of the datasets
+	       "paragraph" = 0.05,
+	       "kmers" = 0.01)
+top_fraction <- fractions[program]
 
 # Getting the ID of the gene associated with the locus
 # DEPENDENCY: utilities/all_signals.rds
@@ -172,7 +177,7 @@ subset_annotation <- function(region, markers, genes, annotation, causal_genes, 
 	return(subset_annot)
 }
 
-top_granges <- top_region(gwas_markers, fraction = 0.05)
+top_granges <- top_region(gwas_markers, fraction = top_fraction)
 
 # Getting a data.frame of all the genes in the region
 all_genes_df <- subset_annotation(region = locus_signal, markers = gwas_markers,

### 2022-09-16 reference_signals/format_deronne2022_signal.R

@@ -7,7 +7,7 @@ library(GenomicRanges)
 
 # Manually entering data for the signal found by de Ronne et al. (2021) with BLINK on Gm15
 deronne_signal <- GRanges(seqnames = "Gm15",
-			  ranges = IRanges(start = 36395777, end = 36895690))
+			  ranges = IRanges(start = 36000000, end = 40000000))
 deronne_signal$signal_id <- "cdwGm15"
 deronne_signal$trait <- "corrected_dry_weight"
 deronne_signal$locus <- "cdwGm15"

### 2022-09-16 figures/gene_subplot.R
Add new signals that were discovered from our work; remove some figures from supplementary material; add thresholds to signal and gene plots; use all k-mers instead of top 10,000 in k-mer plots

@@ -40,6 +40,11 @@ if(is.null(gene)) stop("No gene found where there should be one")
 gwas_results <- readRDS(paste0("gwas_results/", program, "/", id, "_gwas_locus.rds"))
 gwas_signals <- readRDS(paste0("gwas_results/", program, "/", id, "_signal_locus.rds"))
 
+# Getting the name of the trait associated with the signal, and from it the significance threshold
+trait <- target_signal$trait
+# DEPENDENCY: significance thresholds (should not need to be included in Makefile, because the dependence is implicit)
+threshold <- -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", trait, "_threshold_5per.txt"))))
+
 # Getting the one signal associated with this gene
 signal <- subsetByOverlaps(gwas_signals, gene)
 
@@ -49,6 +54,7 @@ if(!length(signal) == 1) warning("There is no signal overlapping the gene ", gen
 ptx_plot <- pvalueGrob(gwas_results = gwas_results,
 		       interval = gene,
 		       feature = NULL,
+		       threshold = threshold,
 		       col = "blue",
 		       pruned_col = if(program == "platypus") "firebrick2" else NULL,
 		       yexpand = c(0.1, 0.1))

### 2022-09-16 figures/kmer_plot.R

@@ -5,7 +5,7 @@ library(grid)
 library(gwastools)
 
 # Setting some analysis parameters
-max_kmers <- 10000 # the maximum number of k-mers to map onto the haplotype sequences
+max_kmers <- -1 # the maximum number of k-mers to map onto the haplotype sequences; -1 means all k-mers
 kmer_length <- 31 # the length of the k-mers used in the analysis
 min_sequences <- 80 # min number of samples that must have a consensus sequence from assembly; otherwise obtained from bwa
 min_frequency <- 5 # min number of times that a haplotype must occur to be used for plotting

### 2022-09-16 figures/signal_subplot.R

@@ -40,8 +40,14 @@ gwas_results <- readRDS(paste0("gwas_results/", program, "/", id, "_gwas_locus.r
 gwas_signals <- readRDS(paste0("gwas_results/", program, "/", id, "_signal_locus.rds"))
 top_markers <- readRDS(paste0("gwas_results/", program, "/", id, "_top_markers.rds"))
 
+
 # Extracting the signal for the locus of interest from those just read
 signal <- subsetByOverlaps(gwas_signals, target_signal, ignore.strand = TRUE)
+
+# Getting the name of the trait associated with the signal, and from it the significance threshold
+trait <- target_signal$trait
+# DEPENDENCY: significance thresholds (should not need to be included in Makefile, because the dependencs is implicit)
+threshold <- -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", trait, "_threshold_5per.txt"))))
 
 if(!length(signal)) {
 	warning("No signal found by ", program, " for signal ID ", id)
@@ -57,6 +63,7 @@ if(!length(signal)) {
 	ptx_plot <- pvalueGrob(gwas_results = gwas_results,
 			       interval = signal,
 			       feature = gene,
+			       threshold = threshold,
 			       col = "blue",
 			       pruned_col = if(program == "platypus") "firebrick2" else NULL,
 			       shading = top_markers,

### 2022-09-16 reference_signals/format_bandillo2017_signals.R

@@ -109,7 +109,9 @@ bandillo_signals[bandillo_signals$locus == "I", "gene_name_v4"] <- chalcone_gene
 # The other gene not found in Gmax v1 annotation is glyma19g27460 and its
 # Gmax v4 equivalent is Glyma.19G101700 according to Sedivy et al. (2017)
 # DOI:10.1111/nph.14418
-bandillo_signals[bandillo_signals$locus == "L1", "gene_name_v4"] <- tolower("Glyma.19G101700")
+# bandillo_signals[bandillo_signals$locus == "L1", "gene_name_v4"] <- tolower("Glyma.19G101700")
+# this gene, however, does not appear to be the right one. We suggest Glyma.19G120300 instead
+bandillo_signals[bandillo_signals$locus == "L1", "gene_name_v4"] <- tolower("Glyma.19G120300")
 
 # Let us also add the genes that have been cloned after Bandillo et al. (2017) and
 # therefore not originally listed in their work

### 2022-09-16 reference_signals/format_custom_signals.R

@@ -1,0 +1,103 @@
+# This script creates signals that have not been documented before
+# but were found through our analyses as a GRanges object for use in
+# downstream analyses
+
+# Loading the GenomicRanges package
+library(GenomicRanges)
+
+# Manually entering data for the signal found on chromosome 16 for pubescence color
+# This signal goes from 1429763-1436787 for kmers, 1430025-1434520 for Paragraph, and 1380791-1485581 for Platypus
+pubescence_color_Gm16_signal <- GRanges(seqnames = "Gm16", ranges = IRanges(start = 1380791, end = 1485581))
+pubescence_color_Gm16_signal$signal_id <- "pubescence_color_all_Gm16"
+pubescence_color_Gm16_signal$trait <- "pubescence_color"
+pubescence_color_Gm16_signal$locus <- "pcGm16"
+pubescence_color_Gm16_signal$n_snps <- NA
+pubescence_color_Gm16_signal$log_pvalue <- 10 # Choosing a random p-value just so it shows up on Manhattan plots
+pubescence_color_Gm16_signal$gene_name_v4 <- NA
+pubescence_color_Gm16_signal$common_name <- NA
+names(pubescence_color_Gm16_signal) <- pubescence_color_Gm16_signal$signal_id
+
+# Manually entering data for the signal found on chromosome 11 for stem_termination_sn
+# This signal goes from 15127026-15402077 for kmers and 15159161-15259163 for Platypus
+stem_termination_Gm11_signal <- GRanges(seqnames = "Gm11", ranges = IRanges(start = 15127026, end = 15402077))
+stem_termination_Gm11_signal$signal_id <- "stem_termination_sn_Gm11"
+stem_termination_Gm11_signal$trait <- "stem_termination"
+stem_termination_Gm11_signal$locus <- "stGm11"
+stem_termination_Gm11_signal$n_snps <- NA
+stem_termination_Gm11_signal$log_pvalue <- 10 # Choosing a random p-value just so it shows up on Manhattan plots
+stem_termination_Gm11_signal$gene_name_v4 <- NA
+stem_termination_Gm11_signal$common_name <- NA
+names(stem_termination_Gm11_signal) <- stem_termination_Gm11_signal$signal_id
+
+# Manually entering data for the signal found on chromosome 15 for pod_color_blbr
[…]

### 2022-09-19 additional_files/additional_file_1.tex
Integrate the generation of signals with the Makefile; add genes for loci that have been cloned after Bandillo et al. (2017)
Add functionality for plotting the haplotypes and variants deduced from the k-mer analysis; work in progress
Test for existence of an alignment before trying to gather sequences at a locus; add k-mer plot for pubescence form, locus Pa1
Add manhattan plots for k-mers located on scaffolds; remove some figures from supplementary material; adjust some parameters in defining signal width and top p-value regions; adjust the signal coordinates for corrected dry weight; make som

@@
 
+% Defining a new command for signal plot figures
 \newcommand{\signalplot}[1]{
@@
 
+% Defining a new command for gene plot figures
 \newcommand{\geneplot}[1]{
@@
 	\label{#1_gene}
+\end{cfigure}
+
+\clearpage}
+
+% Defining a new command for k-mer plot figures
+\newcommand{\kmerplot}[1]{
+	\begin{cfigure}
+		\includegraphics[width = 6.5in]{#1_kmers}
+		\caption{#1\_kmers}
+		\label{#1_kmers}
 	\end{cfigure}
@@
 
+\begin{landscape} \footnotesize
+	\begin{ThreePartTable}
+
+		\begin{TableNotes}
+		\item[a] test
+		\end{TableNotes}
+
+		\begin{longtable}{llcclccccl}
+
+			\caption{Signals}\\
+			\label{signals_table}\\
+
+			\toprule
+			Trait & Locus & Chromosome & -log10(p) & Gene & Platypus & vg & Paragraph & k-mers  & Study \\
+			\midrule \endfirsthead
+
+			\caption[]{Signals \emph{(continued)}}\\
+
+			\toprule
+			Trait & Locus & Chromosome & -log10(p) & Gene & Platypus & vg & Paragraph & k-mers  & Study \\
+			\midrule
+			\endhead
+
+			\bottomrule
+			Continued on next page \\
+			\endfoot
+
+			\bottomrule \insertTableNotes
+			\endlastfoot
+			\csvreader[head to column names,
+			late after line=\\,
+			late after last line=,
+			before reading={\catcode`\#=12},
+			after reading={\catcode`\#=6},
+			/csv/separator=semicolon]%
+			{../tables/signals_table.csv}{}{\Trait & \Locus & \Chromosome & \Pvalues & \Gene & \Platypus & \Vg & \Paragraph & \Kmers & \Study}
+
+		\end{longtable}
+
+	\end{ThreePartTable}
+
+\end{landscape}
+\clearpage
+
 \gwastableone{FLOWER.COLOR}
@@
 \manhattanplot{flower_color}
+%\manhattanplot{flower_color_scaffolds}
 \signalplot{flower_color_W1}
 \geneplot{flower_color_W1}
+\kmerplot{flower_color_W1}
 %\signalplot{flower_color_L1}
@@
 \manhattanplot{pubescence_color_all}
+\manhattanplot{pubescence_color_all_scaffolds}
 %\signalplot{pubescence_color_all_Td}
@@
 \geneplot{pubescence_color_all_T}
+\kmerplot{pubescence_color_all_T}
+\signalplot{pubescence_color_all_pcGm16}
 
 \manhattanplot{pubescence_color_nogray}
+%\manhattanplot{pubescence_color_nogray_scaffolds}
 \signalplot{pubescence_color_nogray_Td}
+\geneplot{pubescence_color_nogray_Td}
+\kmerplot{pubescence_color_nogray_Td}
 %\signalplot{pubescence_color_nogray_T}
@@
 \manhattanplot{seed_coat_color_all}
-\signalplot{seed_coat_color_all_G}
-\signalplot{seed_coat_color_all_T}
-\geneplot{seed_coat_color_all_T}
-\signalplot{seed_coat_color_all_O}
+%\manhattanplot{seed_coat_color_all_scaffolds}
+%\signalplot{seed_coat_color_all_G}
+%\geneplot{seed_coat_color_all_G}
+%\signalplot{seed_coat_color_all_T}
+%\geneplot{seed_coat_color_all_T}
+%\signalplot{seed_coat_color_all_O}
 \signalplot{seed_coat_color_all_I}
@@
 \manhattanplot{seed_coat_color_greenyellow}
+%\manhattanplot{seed_coat_color_greenyellow_scaffolds}
 \signalplot{seed_coat_color_greenyellow_G}
+\geneplot{seed_coat_color_greenyellow_G}
+\kmerplot{seed_coat_color_greenyellow_G}
 %\signalplot{seed_coat_color_greenyellow_T}
@@
 \manhattanplot{stem_termination_all}
+%\manhattanplot{stem_termination_all_scaffolds}
 %\signalplot{stem_termination_all_Dt2}
@@
 \geneplot{stem_termination_all_Dt1}
+\kmerplot{stem_termination_all_Dt1}
 %\signalplot{stem_termination_all_E3}
@@
 \manhattanplot{stem_termination_sn}
+\signalplot{stem_termination_sn_stGm11}
+%\manhattanplot{stem_termination_sn_scaffolds}
 %\signalplot{stem_termination_sn_Dt2}
 %\geneplot{stem_termination_sn_Dt2}
-\signalplot{stem_termination_sn_Dt1}
-\geneplot{stem_termination_sn_Dt1}
+%\signalplot{stem_termination_sn_Dt1}
+%\geneplot{stem_termination_sn_Dt1}
 %\signalplot{stem_termination_sn_E3}
@@
 \manhattanplot{hilum_color_all}
+\manhattanplot{hilum_color_all_scaffolds}
 \signalplot{hilum_color_all_T}
 \geneplot{hilum_color_all_T}
+\kmerplot{hilum_color_all_T}
 %\signalplot{hilum_color_all_O}
@@
 \geneplot{hilum_color_all_I}
-\signalplot{hilum_color_all_R}
-\geneplot{hilum_color_all_R}
+\kmerplot{hilum_color_all_I}
+%\signalplot{hilum_color_all_R}
+%\geneplot{hilum_color_all_R}
 %\signalplot{hilum_color_all_W1}
@@
 \manhattanplot{hilum_color_blackbrown}
+%\manhattanplot{hilum_color_blackbrown_scaffolds}
 %\signalplot{hilum_color_blackbrown_T}
@@
 \manhattanplot{hilum_color_rbr}
+%\manhattanplot{hilum_color_rbr_scaffolds}
 %\signalplot{hilum_color_rbr_T}
@@
 \manhattanplot{pod_color_all}
-\signalplot{pod_color_all_L2}
+%\manhattanplot{pod_color_all_scaffolds}
 \signalplot{pod_color_all_L1}
 \geneplot{pod_color_all_L1}
+\signalplot{pod_color_all_L2}
 
 \manhattanplot{pod_color_blbr}
+\signalplot{pod_color_blbr_pdcGm15}
+%\manhattanplot{pod_color_blbr_scaffolds}
 %\signalplot{pod_color_blbr_L2}
-\signalplot{pod_color_blbr_L1}
-\geneplot{pod_color_blbr_L1}
+%\signalplot{pod_color_blbr_L1}
+%\geneplot{pod_color_blbr_L1}
 
 \manhattanplot{pubescence_form_all}
+%\manhattanplot{pubescence_form_all_scaffolds}
 \signalplot{pubescence_form_all_Pa1}
+\geneplot{pubescence_form_all_Pa1}
+\kmerplot{pubescence_form_all_Pa1}
 \signalplot{pubescence_form_all_Pa2}
+\signalplot{pubescence_form_all_pfGm04}
+\signalplot{pubescence_form_all_pfGm15}
 
 \manhattanplot{pubescence_form_noerect}
+%\manhattanplot{pubescence_form_noerect_scaffolds}
 \signalplot{pubescence_form_noerect_Pa1}
+\geneplot{pubescence_form_noerect_Pa1}
 %\signalplot{pubescence_form_noerect_Pa2}
@@
 \manhattanplot{pubescence_density}
+%\manhattanplot{pubescence_density_scaffolds}
 %\signalplot{pubescence_density_Pd1}
 \signalplot{pubescence_density_Ps}
+\geneplot{pubescence_density_Ps}
 %\signalplot{pubescence_density_P1}
@@
 \manhattanplot{seed_coat_luster_all}
+%\manhattanplot{seed_coat_luster_all_scaffolds}
 %\signalplot{seed_coat_luster_all_I}
@@
 %\signalplot{seed_coat_luster_all_B1}
-\signalplot{seed_coat_luster_all_Hps}
-\geneplot{seed_coat_luster_all_Hps}
+%\signalplot{seed_coat_luster_all_Hps}
+%\geneplot{seed_coat_luster_all_Hps}
 \signalplot{seed_coat_luster_all_B}
+\geneplot{seed_coat_luster_all_B}
+\signalplot{seed_coat_luster_all_sclGm09}
 
 \manhattanplot{seed_coat_luster_nointermediate}
+%\manhattanplot{seed_coat_luster_nointermediate_scaffolds}
 %\signalplot{seed_coat_luster_nointermediate_I}
@@
 %\signalplot{seed_coat_luster_nointermediate_B1}
-\signalplot{seed_coat_luster_nointermediate_Hps}
-\geneplot{seed_coat_luster_nointermediate_Hps}
-\signalplot{seed_coat_luster_nointermediate_B}
+%\signalplot{seed_coat_luster_nointermediate_Hps}
+%\geneplot{seed_coat_luster_nointermediate_Hps}
+%\signalplot{seed_coat_luster_nointermediate_B}
+%\geneplot{seed_coat_luster_nointermediate_B}
 
 \manhattanplot{seed_coat_luster_dullshiny}
+%\manhattanplot{seed_coat_luster_dullshiny_scaffolds}
 %\signalplot{seed_coat_luster_dullshiny_I}
@@
 %\signalplot{seed_coat_luster_dullshiny_B1}
-\signalplot{seed_coat_luster_dullshiny_Hps}
-\geneplot{seed_coat_luster_dullshiny_Hps}
-\signalplot{seed_coat_luster_dullshiny_B}
+%\signalplot{seed_coat_luster_dullshiny_Hps}
+%\geneplot{seed_coat_luster_dullshiny_Hps}
+%\signalplot{seed_coat_luster_dullshiny_B}
+\geneplot{seed_coat_luster_dullshiny_B}
+\signalplot{seed_coat_luster_dullshiny_sclGm20}
 
 \manhattanplot{corrected_dry_weight}
-\signalplot{corrected_dry_weight_cdwGm17}
+\manhattanplot{corrected_dry_weight_scaffolds}
+\signalplot{corrected_dry_weight_cdwGm15}
+\geneplot{corrected_dry_weight_cdwGm15}
 
 \manhattanplot{maturity_group}
+%\manhattanplot{maturity_group_scaffolds}
 %\signalplot{maturity_group_E1}
@@
 \manhattanplot{oil}
+%\manhattanplot{oil_scaffolds}
 \signalplot{oil_oilGm15}
+\geneplot{oil_oilGm15}
 %\signalplot{oil_oilGm20}
@@
 \manhattanplot{protein}
+%\manhattanplot{protein_scaffolds}
 %\signalplot{protein_proteinGm15}

### 2022-09-29 figures/kmer_plot.R
Improve the contingency tables for haplotypes and phenotypes in k-mer plots; adjust the the genomic range of some k-mer plots

@@ -1,9 +1,9 @@
 # This script generates a k-mer plot from the consensus sequences of a group of samples at a given locus
 
 # Loading the required libraries
-library(grid)
-library(gwastools)
-library(GenomicRanges)
+suppressMessages(library(grid))
+suppressMessages(library(gwastools))
+suppressMessages(library(GenomicRanges))
 
 # Setting some analysis parameters
 max_kmers <- -1 # the maximum number of k-mers to map onto the haplotype sequences; -1 means all k-mers
@@ -29,6 +29,9 @@ grange <- as.numeric(strsplit(sub(".*:", "", plotting_range), "-")[[1]])
 # DEPENDENCY: phenotypic_data/trait_names.rds
 trait_names <- readRDS("phenotypic_data/trait_names.rds")
 usda_trait <- names(trait_names)[sapply(trait_names, function(x) grepl(paste0("^", x), trait))]
+
+# DEPENDENCY: phenotypic_data/pheno_names_lookup.rds
+pheno_names_lookup <- readRDS("phenotypic_data/pheno_names_lookup.rds")[[usda_trait]]
 
 # Loading the reference gene and transcript data for plotting the transcript
 genes <- readRDS("refgenome/gmax_v4_genes.rds")
@@ -63,6 +66,12 @@ end(causal_gene) <- end(causal_gene) + xextend * window_size
 # Getting the phenotypic data
 # DEPENDENCY: phenotypic_data/phenotypic_data.csv
 phenotypes <- read.table("phenotypic_data/phenotypic_data.csv", sep = ";", header = TRUE,)
+
+# Setting the values of the phenotype to that in pheno_names_lookup
+phenotypes[[usda_trait]] <- pheno_names_lookup[phenotypes[[usda_trait]]]
+
+# Setting to NA the values that were NA in the analysis being considered
+phenotypes[[usda_trait]][is.na(phenotypes[[trait]])] <- NA
 
 # Reading the k-mers associated with that analysis and their p-values
 # DEPENDENCY: k-mer p-values

### 2022-09-29 figures/ld_plot.R
Add functionality for plotting the pairwise LD between k-mers; add such a plot for flower color

@@ -1,0 +1,50 @@
+# This scripts plots the pairwise LD between a set of k-mers based
+# on a pre-computed LD matrix
+
+# Loading the required packages
+suppressMessages(library(gwastools))
+suppressMessages(library(grid))
+suppressMessages(library(GenomicRanges))
+
+# Getting the trait to analyze from the command line
+trait <- commandArgs(trailingOnly = TRUE)[1]
+
+# Reading the LD matrix from file
+# DEPENDENCY: clustered LD matrix for trait considered
+clustered_ld_file <- paste0("gwas_results/kmers/", trait, "_clustered_ld.txt")
+clustered_ld <- as.matrix(read.table(clustered_ld_file, header = FALSE, row.names = 1, sep = "\t"))
+colnames(clustered_ld) <- rownames(clustered_ld)
+
+# Reading the aligned significant k-mers for this trait
+# DEPENDENCY: positions of the significant k-mers for trait
+kmer_positions <- readRDS(paste0("gwas_results/kmers/", trait, "_gwas.rds"))
+names(kmer_positions) <- kmer_positions$kmer_canon
+
+# Sorting the k-mers in the LD matrix based on their genomic position
+gsorted_ld <- ld_sort(clustered_ld, kmer_positions, sort_param = "position")
+
+# Plotting the results using markers sorted by genomic position
+png(paste0("figures/", trait, "_ld.png"), width = 9, height = 10, units = "in", res = 400)
+
+grid.newpage()
+ld_plot(gsorted_ld, kmer_positions)
+
+dev.off()
+
+
+# Getting a GRanges objct describing the whole genome
+# DEPENDENCY: refgenome/Gmax_508_v4.0_mit_chlp.fasta.fai
+#refgenome <- Rsamtools::scanFaIndex("refgenome/Gmax_508_v4.0_mit_chlp.fasta")
+#names(refgenome) <- as.character(seqnames(refgenome))
+
+# Plotting the results using a subset of markers located at certain genomic positions
+#pubescence_form_all_subset <- GRanges(seqnames ="Gm12", ranges = IRanges(start = 38700000, end = 38800000))
+#pubescence_form_all_subset <- c(pubescence_form_all_subset, refgenome[c("Gm04", "Gm13", "Gm15")])
+
+#pubescence_color_all_subset <- GRanges(seqnames = "Gm06", ranges = IRanges(start = 18100000, end = 18800000))
[…]

### 2022-09-29 gwas_results/ld_analysis.R

@@ -1,0 +1,55 @@
+# This script provides functionality for computing and plotting
+# the linkage disequilibrium (LD) for a set of k-mers using
+# their presence/absence pattern.
+
+# Loading the required libraries
+suppressMessages(library(GenomicRanges))
+suppressMessages(library(gwastools))
+
+# Setting the trait to analyze and loading the k-mer results for
+trait <- commandArgs(trailingOnly = TRUE)[1]
+
+# Creating variables for the paths to external programs
+# WARNING: should consider setting these paths in the Makefile
+filter_kmers <- "~/programs/kmers_gwas/bin/filter_kmers "
+kmers_ld <- "~/scripts/kmers_ld/kmers_ld "
+
+# Creating variables for the file paths to improve readability
+# DEPENDENCY: kmers_table/kmers_table.table
+# DEPENDENCY: kmers_table/kmers_table.names
+kmers_table <- "kmers_table/kmers_table"
+
+kmers_file <- paste0("gwas_results/kmers/", trait, "_significant_kmers.txt")
+pav_file <- paste0("gwas_results/kmers/", trait, "_pav_table.txt")
+nodup_pav_file <- paste0("gwas_results/kmers/", trait, "_pav_table_nodup.txt")
+clustered_ld_file <- paste0("gwas_results/kmers/", trait, "_clustered_ld.txt")
+
+# Reading the aligned significant k-mers for this trait
+# DEPENDENCY: positions of the significant k-mers for trait
+kmer_positions <- readRDS(paste0("gwas_results/kmers/", trait, "_gwas.rds"))
+names(kmer_positions) <- kmer_positions$kmer_canon
+
+# Optionally doing some processing to subsample only some k-mers if there are too many
+if(length(kmer_positions) > 2500) {
+	kmer_positions <- subsample_kmers(kmer_positions, nkmers = 2500, npvalue = 1000)
+}
+
+# Writing the significant k-mers to file
+significant_kmers <- kmer_positions$kmer_canon
+writeLines(significant_kmers, con = kmers_file)
+
+# Extracting the presence/absence values for those k-mers using a k-mers GWAS program
+command <- paste0(filter_kmers, " -t ", kmers_table, " -k ", kmers_file, " -o ", pav_file)
+system(command)
+
+# Reading the PAV table from file and removing variants that have exactly the same PAV profile
+pav_table <- read.table(pav_file, header = TRUE)
[…]

### 2022-10-05 cnv_analysis/cdw_analysis.R
Remove some figures from the additional data file; save the correspondence between haplotypes and phenotypes to file in figures/kmer_plot.R; extend the reference signal found by de Ronne et al.

@@ -1,0 +1,118 @@
+# This script aims to delimitate the extent of the duplicated
+# region at the cdwGm15 locus. This will done in the following steps:
+# 1- Identify groups of samples with differing copy numbers
+#    at that locus based on a normalized value of coverage
+#    over the region divided by the average coverage for
+#    that sample.
+# 2- Add together the coverage values in that region (and
+#    extending into non-duplicated regions) for all samples
+#    with higher coverage.
+# 3- Subject the coverage values for duplicated samples to
+#    a segmentation algorithm such that the location of the
+#    copy number variation can be identified as precisely
+#    as possible
+
+# Loading the required libraries
+suppressMessages(library(Rsamtools))
+suppressMessages(library(GenomicRanges))
+suppressMessages(library(parallel))
+suppressMessages(library(robseg))
+suppressMessages(library(GenomicAlignments))
+
+# Sourcing some functions used for this analysis
+# DEPENDENCY: cnv_analysis/cnv_functions.R
+source("cnv_analysis/cnv_functions.R")
+
+# Parameters for the analysis
+gene_name <- "glyma.15g217700"
+
+# Getting the set of samples to analyse
+# DEPENDENCY: utilities/srr_id_correspondence.txt
+samples <- read.table("utilities/srr_id_correspondence.txt")[[1]]
+
+# Getting the names of the bam files to process
+# DEPENDENCY: merged bam files for all samples
+bam_files <- paste0("illumina_data/merged_bams/", samples, "_merged.bam")
+
+# Getting a vector of average coverage per sample (pre-computed)
+# DEPENDENCY: manifest files for all samples
+average_cov <- sapply(samples, function(x) {
+			      read.table(paste0("sv_genotyping/paragraph/manifest_files/", x, "_manifest.txt"),
+					 skip = 1)[[3]]})
+
+# Loading the set of genes annotated in the reference and extracting the coordinates for the gene of interest
+# DEPENDENCY: refgenome/gmax_v4_genes.rds
+genes <- readRDS("refgenome/gmax_v4_genes.rds")
+genes[gene_name]
+
+# A deletion associated with the phenotype was observed at Gm15:37404782-37405538. Let us have a look
+
[…]

### 2022-10-05 gwas_results/gene_analysis.R

@@ -12,7 +12,7 @@ id <- commandArgs(trailingOnly = TRUE)[1]
 program <- commandArgs(trailingOnly = TRUE)[2]
 
 #----- ANALYSIS PARAMETERS
-gapwidth <- 10^6 # The distance at which signals overlapping the reference signal will be merged
+gapwidth <- 5 * 10^6 # The distance at which signals overlapping the reference signal will be merged
 fractions <- c("platypus" = 0.01, # the fraction of top k-mers considered to be delimiting the top region
 	       "vg" = 0.05,       # depends on the approach used given the size of the datasets
 	       "paragraph" = 0.05,

### 2022-10-05 gwas_results/phenodata_analysis.R

@@ -1,0 +1,159 @@
+# This script has a detailed look at the unexpected haplotype/phenotype
+# associations based on k-mers to try and find explanations
+
+# Reading the table of sample names and their SRR accesion numbers
+srr_ids <- read.table("utilities/srr_id_correspondence.txt")
+srr <- srr_ids[[2]]
+names(srr) <- srr_ids[[1]]
+
+# Flower color
+flower_color <- readRDS("gwas_results/kmers/flower_color_W1_phenodata.rds")
+table(flower_color$phenotype, flower_color$haplotype_id)
+flower_color[flower_color$phenotype == "White" & flower_color$haplotype_id == 1, -2]
+#         sample haplotype_id phenotype
+# NA        <NA>           NA      <NA>
+# 83  SRR1533389            1     White
+# 275    USB-480            1     White
+srr["SRR1533389"]
+#   SRR1533389 
+# "SRR1533389" 
+srr["USB-480"]
+#                   USB-480 
+# "SRR12366485;SRR12366488" 
+
+# the fastqc data for SRR1533389 has a weird GC content, which suggests contamination 
+# the two SRA accesions for USB-480 did not suffer from that problem
+
+
+# Pubescence color (T locus)
+pubescence_color_all <- readRDS("gwas_results/kmers/pubescence_color_all_T_phenodata.rds")
+table(pubescence_color_all$phenotype, pubescence_color_all$haplotype_id)
+pubescence_color_all[pubescence_color_all$phenotype == "Tawny" & pubescence_color_all$haplotype_id == 2, -2]
+#    sample haplotype_id phenotype
+# 38  HN057            2     Tawny
+srr["HN057"]
+#                                 HN057 
+# "SRR12190489;SRR12190490;SRR12190491" 
+# There does not seem to be any problem with the SRR IDs associated with this sample
+
+# Pubescence color (Td locus)
+pubescence_color_nogray <- readRDS("gwas_results/kmers/pubescence_color_nogray_Td_phenodata.rds")
+table(pubescence_color_nogray$phenotype, pubescence_color_nogray$haplotype_id)
+pubescence_color_nogray[pubescence_color_nogray$phenotype == "Light tawny" & pubescence_color_nogray$haplotype_id == 1, -2]
+#      sample haplotype_id   phenotype
[…]

### 2022-10-05 reference_signals/format_deronne2022_signal.R

@@ -7,7 +7,7 @@ library(GenomicRanges)
 
 # Manually entering data for the signal found by de Ronne et al. (2021) with BLINK on Gm15
 deronne_signal <- GRanges(seqnames = "Gm15",
-			  ranges = IRanges(start = 36000000, end = 40000000))
+			  ranges = IRanges(start = 36000000, end = 43000000))
 deronne_signal$signal_id <- "cdwGm15"
 deronne_signal$trait <- "corrected_dry_weight"
 deronne_signal$locus <- "cdwGm15"

### 2022-10-12 cnv_analysis/pa1_analysis.R
Refine the results for the final analysis; add a script for the fine analysis of the results that are mentioned in the paper

@@ -1,0 +1,86 @@
+# This script aims to see whether there could be copy number variation
+# at the Pa1 locus
+# 1- Identify groups of samples with differing copy numbers
+#    at that locus based on a normalized value of coverage
+#    over the region divided by the average coverage for
+#    that sample.
+# 2- Add together the coverage values in that region (and
+#    extending into non-duplicated regions) for all samples
+#    with higher coverage.
+# 3- Subject the coverage values for duplicated samples to
+#    a segmentation algorithm such that the location of the
+#    copy number variation can be identified as precisely
+#    as possible
+
+# Loading the required libraries
+suppressMessages(library(Rsamtools))
+suppressMessages(library(GenomicRanges))
+suppressMessages(library(parallel))
+suppressMessages(library(robseg))
+suppressMessages(library(GenomicAlignments))
+
+# Sourcing some functions used for this analysis
+# DEPENDENCY: cnv_analysis/cnv_functions.R
+source("cnv_analysis/cnv_functions.R")
+
+# Parameters for the analysis
+gene_name <- tolower("Glyma.12g213900")
+
+# Getting the set of samples to analyse
+# DEPENDENCY: utilities/srr_id_correspondence.txt
+samples <- read.table("utilities/srr_id_correspondence.txt")[[1]]
+
+# Getting the names of the bam files to process
+# DEPENDENCY: merged bam files for all samples
+bam_files <- paste0("illumina_data/merged_bams/", samples, "_merged.bam")
+
+# Getting a vector of average coverage per sample (pre-computed)
+# DEPENDENCY: manifest files for all samples
+average_cov <- sapply(samples, function(x) {
+			      read.table(paste0("sv_genotyping/paragraph/manifest_files/", x, "_manifest.txt"),
+					 skip = 1)[[3]]})
+
+# Loading the set of genes annotated in the reference and extracting the coordinates for the gene of interest
+# DEPENDENCY: refgenome/gmax_v4_genes.rds
+genes <- readRDS("refgenome/gmax_v4_genes.rds")
+genes[gene_name]
+
+# Plotting the coverage over a window larger than the expected to get an idea as to where the coverage shifts
+if(interactive()) {
+	covplot <- plot_coverage(bam_files, average_coverage = average_cov,
+				 yscale = c(0, 10), cores = 30,
+				 grange = GRanges(seqnames = "Gm12", ranges = IRanges(start = 38675000, end = 38775000)))
+}
+
[…]

### 2022-10-12 cnv_analysis/r_analysis.R

@@ -1,0 +1,107 @@
+# This script aims to find out whether there is copy number variation
+# at the R locus. This will done in the following steps:
+# 1- Identify groups of samples with differing copy numbers
+#    at that locus based on a normalized value of coverage
+#    over the region divided by the average coverage for
+#    that sample.
+# 2- Add together the coverage values in that region (and
+#    extending into non-duplicated regions) for all samples
+#    with higher coverage.
+# 3- Subject the coverage values for duplicated samples to
+#    a segmentation algorithm such that the location of the
+#    copy number variation can be identified as precisely
+#    as possible
+
+# Loading the required libraries
+suppressMessages(library(Rsamtools))
+suppressMessages(library(GenomicRanges))
+suppressMessages(library(parallel))
+suppressMessages(library(robseg))
+suppressMessages(library(GenomicAlignments))
+
+# Sourcing some functions used for this analysis
+# DEPENDENCY: cnv_analysis/cnv_functions.R
+source("cnv_analysis/cnv_functions.R")
+
+# Parameters for the analysis
+gene_name <- tolower("Glyma.09G234900")
+
+# Getting the set of samples to analyse
+# DEPENDENCY: utilities/srr_id_correspondence.txt
+samples <- read.table("utilities/srr_id_correspondence.txt")[[1]]
+
+# Getting the names of the bam files to process
+# DEPENDENCY: merged bam files for all samples
+bam_files <- paste0("illumina_data/merged_bams/", samples, "_merged.bam")
+
+# Getting a vector of average coverage per sample (pre-computed)
+# DEPENDENCY: manifest files for all samples
+average_cov <- sapply(samples, function(x) {
+			      read.table(paste0("sv_genotyping/paragraph/manifest_files/", x, "_manifest.txt"),
+					 skip = 1)[[3]]})
+
+# Loading the set of genes annotated in the reference and extracting the coordinates for the gene of interest
+# DEPENDENCY: refgenome/gmax_v4_genes.rds
+genes <- readRDS("refgenome/gmax_v4_genes.rds")
+genes[gene_name]
+
+# Plotting the coverage over a window larger than the expected to get an idea as to where the coverage shifts
+if(interactive()) {
+	covplot <- plot_coverage(bam_files, average_coverage = average_cov,
[…]

### 2022-10-12 figures/gene_subplot.R

@@ -3,12 +3,14 @@ suppressMessages(library(grid))
 suppressMessages(library(gwastools))
 suppressMessages(library(GenomicRanges))
 
-# Setting some analysis parameters
-xextend <- 0.1 # The expansion factor on either side of the gene for the transcript plot x-scale
-
 # Getting the name of the trait-locus combination and of the program that called the genotypes from the command line
 id <- commandArgs(trailingOnly = TRUE)[1]
 program <- commandArgs(trailingOnly = TRUE)[2]
+
+# Setting some analysis parameters
+xextend <- 0.1 # The expansion factor on either side of the gene for the transcript plot x-scale
+# The R locus is a special case because I want to see its promoter region
+# if(id == "hilum_color_blackbrown_R") xextend <- 2
 
 # Getting the ID of the gene associated with the trait and the name of the trait from the all_signals.rds file
 # DEPENDENCY: utilities/all_signals.rds
@@ -37,7 +39,7 @@ if(!is.na(gene_name) && grepl(";", gene_name)) {
 # DEPENDENCY: utilities/cnv_genes.txt
 cnv_genes <- read.table("utilities/cnv_genes.txt")
 if(id %in% cnv_genes[[1]]) {
-	cnv_range <- readRDS(cnv_genes[cnv_genes[[1]] == id, 2])
+	gene <- cnv_range <- readRDS(cnv_genes[cnv_genes[[1]] == id, 2])
 } else {
 	cnv_range <- NULL
 }

### 2022-10-12 figures/signal_subplot.R

@@ -35,6 +35,16 @@ if(!is.na(gene_name) && grepl(";", gene_name)) {
 	gene <- NULL
 }
 
+# Handling special cases where the causal variant acutally spans several genes
+# In this case we set the "gene" variable to the span of the CNV/variant
+# Reading the CNV range (if applicable)
+# DEPENDENCY: utilities/cnv_genes.txt
+cnv_genes <- read.table("utilities/cnv_genes.txt")
+
+if(id %in% cnv_genes[[1]]) {
+	gene <- readRDS(cnv_genes[cnv_genes[[1]] == id, 2])
+} 
+
 # Reading the GWAS results, signals, and region of top markers from file
 gwas_results <- readRDS(paste0("gwas_results/", program, "/", id, "_gwas_locus.rds"))
 gwas_signals <- readRDS(paste0("gwas_results/", program, "/", id, "_signal_locus.rds"))

### 2022-10-12 gwas_results/fine_analysis.R

@@ -5,6 +5,8 @@
 library(GenomicRanges)
 library(VariantAnnotation)
 library(Biostrings)
+library(GenomicFeatures)
+library(Rsamtools)
 
 # First we will create a few functions that will be used throughout
 
@@ -43,7 +45,7 @@ all_signals <- readRDS("utilities/all_signals.rds")
 genes <- readRDS("refgenome/gmax_v4_genes.rds")
 
 # The paths to the filtered variant files
-platypus_file <- "filtered_variants/platypus/filtered_variants.vcf.gz"
+platypus_file <- "filtered_variants/platypus_full.vcf.gz"
 paragraph_file <- "filtered_variants/paragraph/filtered_variants.vcf.gz"
 vg_file <- "filtered_variants/vg/filtered_variants.vcf.gz"
 
@@ -94,3 +96,154 @@ system("samtools view gwas_results/kmer_data/pubescence_color_all/katcher_result
 system("samtools view gwas_results/kmer_data/pubescence_color_all/katcher_results/USB-512/USB-512_pvalues_sorted.bam | grep AAAAAATAAAAATAAAAAATGAAAACAGAAG | less -S")
 system("samtools view gwas_results/kmer_data/pubescence_color_all/katcher_results/USB-512/USB-512_pvalues_sorted.bam | grep CTTCTGTTTTCATTTTTTATTTTTATTTTTT | less -S")
 # There seems to be an insertion associated with this last k-mer
+
+# I haven't been able to link those k-mers to any variant
+
+# Let us have a look at the two significant markers found using Paragraph
+vcfFixed(paragraph_t <- vcf_extract(paragraph_file, pc_all_top_markers$paragraph[1:2]))
+vcfInfo(paragraph_t)
+width(vcfFixed(paragraph_t)$REF)
+nchar(vcfFixed(paragraph_t)$ALT)
+
+
+
+
+### ---------- PUBESCENCE COLOR NO GRAY
+pubescence_color_nogray <- read_gwas("pubescence_color_nogray")
+pc_nogray_top_markers <- gene_markers(pubescence_color_nogray, all_signals, genes, "pubescence_color_nogray_Td")
+vcfFixed(platypus_td <- vcf_extract(platypus_file, pc_nogray_top_markers$platypus[1]))
+
+
+### ---------- SEED COAT COLOR GREEN YELLOW
+seed_coat_color_greenyellow <- read_gwas("seed_coat_color_greenyellow")
[…]

### 2022-10-12 reference_signals/format_bandillo2017_signals.R

@@ -111,7 +111,9 @@ bandillo_signals[bandillo_signals$locus == "I", "gene_name_v4"] <- chalcone_gene
 # DOI:10.1111/nph.14418
 # bandillo_signals[bandillo_signals$locus == "L1", "gene_name_v4"] <- tolower("Glyma.19G101700")
 # this gene, however, does not appear to be the right one. We suggest Glyma.19G120300 instead
-bandillo_signals[bandillo_signals$locus == "L1", "gene_name_v4"] <- tolower("Glyma.19G120300")
+
+# this gene does not appear sure enough, therefore I will leave it blank
+bandillo_signals[bandillo_signals$locus == "L1", "gene_name_v4"] <- NA
 
 # Let us also add the genes that have been cloned after Bandillo et al. (2017) and
 # therefore not originally listed in their work
@@ -138,6 +140,9 @@ hps_genes <- sort(gmax_v4_genes[hps_genes], ignore.strand = TRUE)
 hps_genes <- subsetByOverlaps(hps_genes,
 			      GRanges(seqnames = "Gm15", ranges = IRanges(start = 9 * 10^6, end = 10 * 10^6)),
 			      ignore.strand = TRUE)
+
+# R locus; my results support another gene as being the causal gene so I'll have a look at it
+# bandillo_signals[bandillo_signals$locus == "R", "gene_name_v4"] <- tolower("glyma.09g234900")
 
 # Let us add the names of the first and last of those genes in the gene_name_v4 column
 hps_gene_ids <- paste0(names(hps_genes)[c(1, length(hps_genes))], collapse = ";")

### 2022-10-12 reference_signals/format_custom_signals.R

@@ -18,8 +18,8 @@ pubescence_color_Gm16_signal$common_name <- NA
 names(pubescence_color_Gm16_signal) <- pubescence_color_Gm16_signal$signal_id
 
 # Manually entering data for the signal found on chromosome 11 for stem_termination_sn
-# This signal goes from 15127026-15402077 for kmers and 15159161-15259163 for Platypus
-stem_termination_Gm11_signal <- GRanges(seqnames = "Gm11", ranges = IRanges(start = 15127026, end = 15402077))
+# This signal goes from 15039148-15413930 for kmers
+stem_termination_Gm11_signal <- GRanges(seqnames = "Gm11", ranges = IRanges(start = 15039148, end = 15413930))
 stem_termination_Gm11_signal$signal_id <- "stem_termination_sn_Gm11"
 stem_termination_Gm11_signal$trait <- "stem_termination"
 stem_termination_Gm11_signal$locus <- "stGm11"
@@ -29,9 +29,33 @@ stem_termination_Gm11_signal$gene_name_v4 <- NA
 stem_termination_Gm11_signal$common_name <- NA
 names(stem_termination_Gm11_signal) <- stem_termination_Gm11_signal$signal_id
 
+# Manually entering data for the signal found on chromosome 18 for stem_termination_sn
+# This signal goes from 50694062-50707043 for kmers
+stem_termination_Gm18_signal <- GRanges(seqnames = "Gm18", ranges = IRanges(start = 50694062, end = 50707043))
+stem_termination_Gm18_signal$signal_id <- "stem_termination_sn_Gm18"
+stem_termination_Gm18_signal$trait <- "stem_termination"
+stem_termination_Gm18_signal$locus <- "stGm18"
+stem_termination_Gm18_signal$n_snps <- NA
+stem_termination_Gm18_signal$log_pvalue <- 10 # Choosing a random p-value just so it shows up on Manhattan plots
+stem_termination_Gm18_signal$gene_name_v4 <- NA
+stem_termination_Gm18_signal$common_name <- NA
+names(stem_termination_Gm18_signal) <- stem_termination_Gm18_signal$signal_id
+
+# Manually entering data for the signal found on chromosome 16 for stem_termination_sn
+# This signal goes from 31899974-31999228 for kmers
[…]

### 2022-10-13 additional_files/additional_file_1.tex
Improve the contingency tables for haplotypes and phenotypes in k-mer plots; adjust the the genomic range of some k-mer plots
Add functionality for plotting the pairwise LD between k-mers; add such a plot for flower color
Specify the figures to be made in the Makefile dynamically from the ones listed in the additional file
Remove some figures from the additional data file; save the correspondence between haplotypes and phenotypes to file in figures/kmer_plot.R; extend the reference signal found by de Ronne et al.
Refine the results for the final analysis; add a script for the fine analysis of

--- 2022-07-29 additional_files/additional_file_1.tex
@@
 
+% Defining a new command for signal plot figures
 \newcommand{\signalplot}[1]{
 \begin{cfigure}
-	\caption[Figure caption]{Figure caption}
 	\includegraphics[width = 6in]{#1_signal}
+	\caption{#1\_signal}
 	\label{#1_signal}
@@
 \clearpage}
+
+% Defining a new command for gene plot figures
+\newcommand{\geneplot}[1]{
+\begin{cfigure}
+	\includegraphics[width = 6in]{#1_gene}
+	\caption{#1\_gene}
+	\label{#1_gene}
+\end{cfigure}
+
+\clearpage}
+
+% Defining a new command for k-mer plot figures
+\newcommand{\kmerplot}[1]{
+	\begin{cfigure}
+		\includegraphics[width = 6.5in]{#1_kmers}
+		\caption{#1\_kmers}
+		\label{#1_kmers}
+	\end{cfigure}
+
+\clearpage}
+
+% Defining a new command for LD plot figures
+\newcommand{\ldplot}[1]{
+	\begin{cfigure}
+		\includegraphics[width = 6.5in]{#1_ld}
+		\caption{#1\_ld}
+		\label{#1_ld}
+	\end{cfigure}
+
+\clearpage}
+
+% Allows underscores in captions but is not a great solution
+\catcode`_=12
 
@@
 
-% \caption[Number of SVs per Oxford Nanopore sample]{{{{
+\begin{landscape} \footnotesize
+	\begin{ThreePartTable}
 
-%\begin{table}
-%	\caption[Number of SVs per Oxford Nanopore sample]{}
-%	\label{table_s1}
-%
-%	\begin{threeparttable}\small
-%			   \begin{tablenotes}
-%			   \end{tablenotes}
-%	\end{threeparttable}
-%
-%\end{table}
-%\clearpage%}}}
+		\begin{TableNotes}
+		\item[a] test
+		\end{TableNotes}
 
-% Manhattan plots of all traits
+		\begin{longtable}{llcclccccl}
+
+			\caption{Signals}\\
+			\label{signals_table}\\
+
+			\toprule
+			Trait & Locus & Chromosome & -log10(p) & Gene & Platypus & vg & Paragraph & k-mers  & Study \\
+			\midrule \endfirsthead
+
+			\caption[]{Signals \emph{(continued)}}\\
+
+			\toprule
+			Trait & Locus & Chromosome & -log10(p) & Gene & Platypus & vg & Paragraph & k-mers  & Study \\
+			\midrule
+			\endhead
+
+			\bottomrule
+			Continued on next page \\
+			\endfoot
+
+			\bottomrule \insertTableNotes
+			\endlastfoot
+			\csvreader[head to column names,
+			late after line=\\,
+			late after last line=,
+			before reading={\catcode`\#=12},
+			after reading={\catcode`\#=6},
+			/csv/separator=semicolon]%
+			{../tables/signals_table.csv}{}{\Trait & \Locus & \Chromosome & \Pvalues & \Gene & \Platypus & \Vg & \Paragraph & \Kmers & \Study}
+
+		\end{longtable}
+
+	\end{ThreePartTable}
+
+\end{landscape}
+\clearpage
+
+\gwastableone{FLOWER.COLOR}
+\gwastabletwo{PUBESCENCE.COLOR}
+\gwastabletwo{SEED.COAT.COLOR}
+\gwastabletwo{STEM.TERMINATION.TYPE}
+\gwastablethree{HILUM.COLOR}
+\gwastabletwo{POD.COLOR}
+\gwastabletwo{PUBESCENCE.FORM}
+\gwastableone{PUBESCENCE.DENSITY}
+\gwastablethree{SEED.COAT.LUSTER}
+\gwastableone{MATURITY.GROUP}
+
+% Figures for flower color
+\manhattanplot{flower_color}
+\ldplot{flower_color}
+\signalplot{flower_color_W1}
+\geneplot{flower_color_W1}
+\kmerplot{flower_color_W1}
+
+% Figures for pubescence color
+\manhattanplot{pubescence_color_all}
+\signalplot{pubescence_color_all_T}
+\geneplot{pubescence_color_all_T}
+\kmerplot{pubescence_color_all_T}
+\ldplot{pubescence_color_all}
+
+\manhattanplot{pubescence_color_nogray}
+\signalplot{pubescence_color_nogray_Td}
+\geneplot{pubescence_color_nogray_Td}
+\kmerplot{pubescence_color_nogray_Td}
+
+% Figures for seed coat color
+\manhattanplot{seed_coat_color_all}
+\signalplot{seed_coat_color_all_I}
+
+\manhattanplot{seed_coat_color_greenyellow}
+\signalplot{seed_coat_color_greenyellow_G}
+\geneplot{seed_coat_color_greenyellow_G}
+\kmerplot{seed_coat_color_greenyellow_G}
+
+% Figures for stem termination
 \manhattanplot{stem_termination_all}
+\signalplot{stem_termination_all_Dt1}
+\geneplot{stem_termination_all_Dt1}
+\kmerplot{stem_termination_all_Dt1}
+
 \manhattanplot{stem_termination_sn}
-\manhattanplot{pubescence_color_all}
-\manhattanplot{pubescence_color_nogray}
+\ldplot{stem_termination_sn}
+\signalplot{stem_termination_sn_stGm11}
+\signalplot{stem_termination_sn_stGm16}
+\signalplot{stem_termination_sn_stGm18}
+
+% Figures for hilum color
+\manhattanplot{hilum_color_all}
+\signalplot{hilum_color_all_T}
+\geneplot{hilum_color_all_T}
+\kmerplot{hilum_color_all_T}
+\signalplot{hilum_color_all_I}
+\geneplot{hilum_color_all_I}
+
+\manhattanplot{hilum_color_blackbrown}
+\signalplot{hilum_color_blackbrown_R}
+\geneplot{hilum_color_blackbrown_R}
+\kmerplot{hilum_color_blackbrown_R}
+
+\manhattanplot{hilum_color_rbr}
+
+% Figures for pod color
+\manhattanplot{pod_color_all}
+\signalplot{pod_color_all_L1}
+\signalplot{pod_color_all_L2}
+
+\manhattanplot{pod_color_blbr}
+\ldplot{pod_color_blbr}
+\signalplot{pod_color_blbr_pdcGm15}
+
+% Figures for pubescence form
+\manhattanplot{pubescence_form_all}
+\signalplot{pubescence_form_all_Pa1}
+\geneplot{pubescence_form_all_Pa1}
+\kmerplot{pubescence_form_all_Pa1}
+\signalplot{pubescence_form_all_Pa2}
+\ldplot{pubescence_form_all}
+
+\manhattanplot{pubescence_form_noerect}
+
+% Figures for pubescence density
+\manhattanplot{pubescence_density}
+\signalplot{pubescence_density_Ps}
+\geneplot{pubescence_density_Ps}
+
+% Figures for seed coat luster
+\manhattanplot{seed_coat_luster_all}
+
+\manhattanplot{seed_coat_luster_nointermediate}
+
+\manhattanplot{seed_coat_luster_dullshiny}
+\ldplot{seed_coat_luster_dullshiny}
+\signalplot{seed_coat_luster_dullshiny_B}
+\geneplot{seed_coat_luster_dullshiny_B}
+
+% Figures for corrected dry weight
+\manhattanplot{corrected_dry_weight}
+\ldplot{corrected_dry_weight}
+\signalplot{corrected_dry_weight_cdwGm15}
+
+% Figures for maturity group
+\manhattanplot{maturity_group}
+
+% Figures for oil
+\manhattanplot{oil}
+
+% Figures for protein
 \manhattanplot{protein}
-\manhattanplot{pubescence_form_all}
-\manhattanplot{pubescence_form_noerect}
-\manhattanplot{oil}
-\manhattanplot{pubescence_density}
-\manhattanplot{seed_coat_luster_all}
-\manhattanplot{seed_coat_luster_nointermediate}
-\manhattanplot{seed_coat_luster_dullshiny}
-\manhattanplot{pod_color_all}
-\manhattanplot{pod_color_blbr}
-\manhattanplot{flower_color}
-\manhattanplot{seed_coat_color_all}
-\manhattanplot{seed_coat_color_greenyellow}
-\manhattanplot{hilum_color_all}
-\manhattanplot{hilum_color_blackbrown}
-\manhattanplot{hilum_color_rbr}
-\manhattanplot{maturity_group}
-\manhattanplot{corrected_dry_weight}
-
-\signalplot{pca_t}
-\signalplot{fc_w}
-\signalplot{sda_dt1}
 

### 2022-10-13 cnv_analysis/cnv_luster_gwas.R

@@ -1,0 +1,29 @@
+# Loading the GAPIT3 package
+library(GAPIT3)
+
+# Loading phenotype data
+phenotypes <- read.table("cnv_analysis/cnv_luster_data.csv",
+			 header = TRUE, sep = ",",
+			 stringsAsFactors = FALSE)[, c("id", "dullshiny")]
+
+colnames(phenotypes) <- c("Taxa", "dullshiny")
+
+# Loading the genotype data
+genotypes <- read.table("~/sv_gwas/usda_lines/gwas/platypus/platypus_formatted.hmp.txt",
+			comment.char = "",
+			header = FALSE)
+
+# Running the GAPIT analysis using 9 PCs and VanRaden kinship
+dir.create("cnv_analysis/cnv_luster_gwas", recursive = TRUE)
+setwd("cnv_analysis/cnv_luster_gwas")
+
+gapit_model <- GAPIT(Y = phenotypes,
+		     G = genotypes,
+		     kinship.algorithm = "VanRaden",
+		     PCA.total = 9,
+		     Model.selection = FALSE, 
+		     model = "MLM",
+		     file.output = TRUE)
+
+save(gapit_model, file = paste0("gapit_MLM_", phenotype, ".RData"))
+

### 2022-10-13 cnv_analysis/hps_analysis.R

@@ -49,7 +49,7 @@ genes[gene_name]
 if(interactive()) {
 	covplot <- plot_coverage(bam_files, average_coverage = average_cov,
 				 yscale = c(0, 15), cores = 30,
-				 grange = GRanges(seqnames = "Gm15", ranges = IRanges(start = 9300000, end = 9450000)))
+				 grange = GRanges(seqnames = "Gm15", ranges = IRanges(start = 9380000, end = 9430000)))
 }
 
 # I will use the window plotted above to calculate the average normalized coverage for all samples
@@ -65,6 +65,33 @@ normalized_cov <- rcoverage / average_cov
 
 # Visualizing the distribution to identify a threshold separating CNV samples from low copy number samples
 if(interactive()) hist(normalized_cov, breaks = 50)
+
+# Having a look at the link between copy number and phenotype
+phenotypic_data <- read.table("phenotypic_data/phenotypic_data.csv", header = TRUE, sep = ";")
+luster <- phenotypic_data[match(names(normalized_cov), phenotypic_data$bayer_id), "seed_coat_luster_all"]
+
+boxplot_data <- data.frame(coverage = normalized_cov,
+			   luster = luster,
+			   stringsAsFactors = FALSE)
+
+boxplot_data$luster <- as.factor(boxplot_data$luster)
+
+boxplot(coverage ~ luster, data = boxplot_data)
+summary(aov(coverage ~ luster, data = boxplot_data))
+TukeyHSD(aov(coverage ~ luster, data = boxplot_data))
+
+# Also plotting the results with only the dull and shiny phenotypes
+boxplot_data$dullshiny <- phenotypic_data[match(names(normalized_cov), phenotypic_data$bayer_id), "seed_coat_luster_dullshiny"]
+boxplot(coverage ~ dullshiny, data = boxplot_data)
+t.test(coverage ~ dullshiny, data = boxplot_data)
+
+# It is weird that some accessions have shiny seed coats, yet high copy number at the B locus
+# Let us make a GWAS contrasting only accessions with high copy number
+cnv_luster_data <- boxplot_data[complete.cases(boxplot_data), ]
+cnv_luster_data <- cnv_luster_data[cnv_luster_data$coverage > 2,]
+cnv_luster_data$id <- rownames(cnv_luster_data)
+write.table(cnv_luster_data, file = "cnv_analysis/cnv_luster_data.csv", sep = ",",
[…]

### 2022-10-13 gwas_results/fine_analysis.R

@@ -247,3 +247,47 @@ pubescence_density$kmers[1]$kmer_canon
 system("samtools view gwas_results/kmer_data/pubescence_density/katcher_results/USB-672/USB-672_pvalues_sorted.bam | grep AAAATTAATGATATTTTTTTGTAAAAATTAT | less -S")
 # This k-mer is clearly associated with the CNV at the Ps locus
 
+
+
+
+# SEED COAT LUSTER (DULL-SHINY)
+seed_coat_luster <- read_gwas("seed_coat_luster_dullshiny")
+seed_coat_luster$kmers[1:10]
+
+# Checking if any of the top k-mers have reads with special patterns
+seed_coat_luster$kmers[1:10]$kmer_canon
+#  [1] "AAAGTTAGATATAGTTTTTTTTTTTTTTTTA" "GCCTAATCTGAATATGGAGTATCAAATAAAA" "CTTTCTTTCTTTTCCTTTCTGGTATGTTTTA" "ATAGTGGTAAAAAAAAAAAAAAAAACTATAT" "GATATAGTTTTTTTTTTTTTTTTTACCACTA"
+#  [6] "ATATATATATATATTATATTTCAAGTTACAT" "CATTTTGAGTTCAGATGGCCTAATCTGAATA" "AAATAAATTTATTATAGCGCGTGTATCTAAA" "TCTTGATTTTGACCGGTCCAGTTGGGTAACA" "GCATTTTTTTTTTGTTCAACTACTAGTTTAA"
+as.character(reverseComplement(DNAStringSet(seed_coat_luster$kmers[1:10]$kmer_canon)))
+#  [1] "TAAAAAAAAAAAAAAAACTATATCTAACTTT" "TTTTATTTGATACTCCATATTCAGATTAGGC" "TAAAACATACCAGAAAGGAAAAGAAAGAAAG" "ATATAGTTTTTTTTTTTTTTTTTACCACTAT" "TAGTGGTAAAAAAAAAAAAAAAAACTATATC"
+#  [6] "ATGTAACTTGAAATATAATATATATATATAT" "TATTCAGATTAGGCCATCTGAACTCAAAATG" "TTTAGATACACGCGCTATAATAAATTTATTT" "TGTTACCCAACTGGACCGGTCAAAATCAAGA" "TTAAACTAGTAGTTGAACAAAAAAAAAATGC"
[…]

### 2022-10-14 figures/kmer_plot.R
Correct an error in determining the extent of signals for Platypus after computing the association of pruned markers; remove a few figures from additional file 1

@@ -53,10 +53,10 @@ if(!is.na(gene_name) && grepl(";", gene_name)) {
 } else if (!is.na(gene_name)) {
 	causal_gene <- genes[gene_name]
 } else {
-	gene <- NULL
+	causal_gene <- NULL
 }
 
-if(is.null(causal_gene)) stop("No gene found where there should be one")
+if(is.null(causal_gene)) warning("No causal gene known for locus ", locus)
 
 # Extending the window for the causal gene
 window_size <- GenomicRanges::width(causal_gene)
@@ -152,17 +152,19 @@ grid::grid.rect()
 grid::pushViewport(grid::viewport(layout = grid::grid.layout(nrow = 5, heights = grid::unit(c(0.1, 0.15, 0.3, 0.15, 0.3), "npc"))))
 
 # Plotting the transcript in the first viewport
-grid::pushViewport(grid::viewport(width = 0.84, layout.pos.row = 1))
-grid.draw(transcriptsGrob(genes = genes,
-			  transcripts = transcripts,
-			  exons = exons,
-			  cds = cds,
-			  xscale = causal_gene,
-			  highlight = GenomicRanges::GRanges(seqnames = chrom,
-							     ranges = IRanges::IRanges(start = grange[1], end = grange[2])),
-			  draw_arrows = TRUE,
-			  first_tx_only = TRUE))
-grid::upViewport()
+if(!is.null(causal_gene)) {
+	grid::pushViewport(grid::viewport(width = 0.84, layout.pos.row = 1))
+	grid.draw(transcriptsGrob(genes = genes,
+				  transcripts = transcripts,
+				  exons = exons,
+				  cds = cds,
+				  xscale = causal_gene,
+				  highlight = GenomicRanges::GRanges(seqnames = chrom,
+								     ranges = IRanges::IRanges(start = grange[1], end = grange[2])),
+				  draw_arrows = TRUE,
+				  first_tx_only = TRUE))
+	grid::upViewport()
+}
 
 # Moving into the viewport associated with the sequences and drawing them
 grid::pushViewport(grid::viewport(layout.pos.row = 3))

### 2022-10-14 gwas_results/find_signals.R

@@ -31,14 +31,15 @@ if(program == "platypus" && length(gwas_signals)) {
 	gwas_subset$pruned <- FALSE
 
 	# Extending each signal to the right and left
-	start(gwas_signals) <- start(gwas_signals) - platypus_extend
-	end(gwas_signals) <- end(gwas_signals) + platypus_extend
+	gwas_focus <- gwas_signals
+	start(gwas_focus) <- start(gwas_focus) - platypus_extend
+	end(gwas_focus) <- end(gwas_focus) + platypus_extend
 
 	# Opening the VCF file that contains the markers before pruning for reading
 	vcf_file <- VariantAnnotation::VcfFile("filtered_variants/platypus_full.vcf.gz")
 
 	# Reading the VCF records found in that region; only keeping those that were not already analyzed
-	subset_vcf <- VariantAnnotation::readVcf(vcf_file, param = VariantAnnotation::ScanVcfParam(fixed = NA, info = NA, geno = "GT", which = gwas_signals))
+	subset_vcf <- VariantAnnotation::readVcf(vcf_file, param = VariantAnnotation::ScanVcfParam(fixed = NA, info = NA, geno = "GT", which = gwas_focus))
 	subset_vcf <- subset_vcf[!names(subset_vcf) %in% names(gwas_subset)]
 
 	# It is only worth computing the GWAS if there are any markers left ; also excluding cases with only one marker as it makes GAPIT fail
@@ -62,6 +63,11 @@ if(program == "platypus" && length(gwas_signals)) {
 		GenomeInfoDb::seqlevels(gapit_results) <- GenomeInfoDb::seqlevels(gwas_subset)
 
 		if(length(gapit_results)) gwas_subset <- c(gwas_subset, gapit_results)
+
+		# We need to re-compute the signals based on this new information
+		gwas_signals <- extract_signals(gwas_subset,
+						threshold = threshold,
+						distance = signal_distance)
 	}
 }
 

### 2022-10-14 gwas_results/fine_analysis.R

@@ -106,7 +106,13 @@ width(vcfFixed(paragraph_t)$REF)
 nchar(vcfFixed(paragraph_t)$ALT)
 
 
+# Let us have a look at the k-mers associated with the signal on chromsomes 16 and 1
+pc_all_signals <- readRDS("gwas_results/kmers/pubescence_color_all_signal.rds")
+pc_all_signals[seqnames(pc_all_signals) %in% c("Gm01", "Gm16") & pc_all_signals$n_markers > 3]
 
+# Let us identify the most significant k-mer at that locus
+subsetByOverlaps(pubescence_color_all$kmers, GRanges(seqnames = "Gm16", ranges = IRanges(start = 1429763, end = 1436822)))
+# One of the k-mers at 1434693-1434723 shows a clear peak that we will use for a k-mer plot
 
 ### ---------- PUBESCENCE COLOR NO GRAY
 pubescence_color_nogray <- read_gwas("pubescence_color_nogray")

### 2022-10-20 additional_files/additional_file_1.tex
Correct an error in determining the extent of signals for Platypus after computing the association of pruned markers; remove a few figures from additional file 1
Start adding detailed captions to figures in Additional file 1; no longer exclude k-mers from LD analysis bbased on identical segregation patterns

@@
 \usepackage{listings}
+\usepackage{datatool} % allows fetching key-value pairs from a text file
+
+% Setting the separator to use with datatool
+\DTLsetseparator{ = }
 
@@
 % Defining a command for manhattan plot figures
-\newcommand{\manhattanplot}[1]{
+\newcommand{\manhattanplot}[3]{
 \begin{cfigure}
 	\includegraphics[width = 6in]{#1_manhattan}
-	\caption{#1\_manhattan}
+	\caption[Manhattan plot of #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
+	(a) Platypus (SNPs and indels), (b) vg (SVs), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
+	The x-axis shows the position along the reference assembly version 4 of Williams82.
+	Each point shows the $-\log_{10} {p-value}$ associated with a particular marker or \emph{k}-mer.
+	Horizontal dotted lines indicate the 5\% family-wise error-rate threshold determined from a randomization approach.
+	Vertical dotted lines indicate the position of signals associated with the trait.
+	Signals documented in previous studies are colored according to whether they were founs (green) or not (red) by a particular approach.
+	The ``Gm'' prefix has been left out of chromosome names for simplicity.
+	}
 	\label{#1_manhattan}
@@
 % Defining a new command for signal plot figures
-\newcommand{\signalplot}[1]{
+\newcommand{\signalplot}[4]{
 \begin{cfigure}
 	\includegraphics[width = 6in]{#1_signal}
-	\caption{#1\_signal}
+	\caption[Zoomed-in Manhattan plot at #3 locus for #2]{Zommed-in Manhattan plots generated from the GWAS analysis of #2 at the #3 locus using four genotype datasets:
+	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
+	Panel (a) shows the location of annotated gene models over the plotting interval.
+	#4
+	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph and vg) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
+	}
 	\label{#1_signal}
@@
 % Defining a new command for gene plot figures
-\newcommand{\geneplot}[1]{
+\newcommand{\geneplot}[4]{
 \begin{cfigure}
 	\includegraphics[width = 6in]{#1_gene}
-	\caption{#1\_gene}
+	\caption[Zommed-in Manhattan plot at gene(s) associated with #3 locus for #2]%
+	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using four genotype datasets:
+	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
+	Panel (a) shows the location of annotated gene models over the plotting interval.
+	Exons are represented by rectangles whereas introns are represented by horizontal lines.
+	Coding sequences are shown in blue and the direction of translation is indicated by arrows.
+	}
 	\label{#1_gene}
@@
 % Defining a new command for LD plot figures
-\newcommand{\ldplot}[1]{
+\newcommand{\ldplot}[3]{
 	\begin{cfigure}
 		\includegraphics[width = 6.5in]{#1_ld}
-		\caption{#1\_ld}
+		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
+		\emph{k}-mers are sorted along the y-axis according to their putative position along the reference assembly version 4 of Williams82,
+		as identified by ``Gm'' chromosome labels.
+		Sequences that lack a ``Gm'' label (top of the y-axis) represent unanchored scaffolds.
+		\emph{k}-mers are represented in the same order along the x- and y-axis.
+		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p-values)$ of each \emph{k}-mer.}
 		\label{#1_ld}
@@
 \begin{document}
+
+% Reading the contents of the variables.txt file first
+\DTLloaddb[noheader, keys={key,value}]{variables}{variables.txt}
+% Defining a new command to fetch the value from the key
+\newcommand{\getvar}[1]{\DTLfetch{variables}{key}{#1}{value}}
 
@@
 % Figures for flower color
-\manhattanplot{flower_color}
-\ldplot{flower_color}
-\signalplot{flower_color_W1}
-\geneplot{flower_color_W1}
+\manhattanplot{flower_color}{flower color}{\getvar{flower_color_nsamples}}
+\ldplot{flower_color}{flower color}{\getvar{flower_color_ldkmers}}
+\signalplot{flower_color_W1}{flower color}{\emph{W1}}%
+{Vertical dotted lines indicate the location of the \getvar{flower_color_W1_gene} gene associated with the locus.}
+
+\geneplot{flower_color_W1}{flower color}{\emph{W1}}{\getvar{flower_color_W1_gene} gene}
 \kmerplot{flower_color_W1}
@@
 % Figures for pubescence color
-\manhattanplot{pubescence_color_all}
-\signalplot{pubescence_color_all_T}
-\geneplot{pubescence_color_all_T}
+\manhattanplot{pubescence_color_all}{pubescence color (first GWAS)}{\getvar{pubescence_color_all_nsamples}}
+\signalplot{pubescence_color_all_T}{pubescence color (first GWAS)}{\emph{T}}%
+{Vertical dotted lines indicate the location of the \getvar{pubescence_color_all_T_gene} gene associated with the locus.}
+\geneplot{pubescence_color_all_T}{pubescence color (first GWAS)}{\emph{T}}{\getvar{pubescence_color_all_T_gene} gene}
 \kmerplot{pubescence_color_all_T}
-\ldplot{pubescence_color_all}
+\ldplot{pubescence_color_all}{pubescence color (first GWAS)}{\getvar{pubescence_color_all_ldkmers}}
 
-\manhattanplot{pubescence_color_nogray}
-\signalplot{pubescence_color_nogray_Td}
-\geneplot{pubescence_color_nogray_Td}
+\manhattanplot{pubescence_color_nogray}{pubescence color (second GWAS)}{\getvar{pubescence_color_nogray_nsamples}}
+\signalplot{pubescence_color_nogray_Td}{pubescence color (second GWAS)}{\emph{Td}}%
+{Vertical dotted lines indicate the location of the \getvar{pubescence_color_nogray_Td_gene} gene associated with the locus.}
+\geneplot{pubescence_color_nogray_Td}{pubescence color (second GWAS)}{\emph{Td}}{\getvar{pubescence_color_nogray_Td_gene} gene}
 \kmerplot{pubescence_color_nogray_Td}
@@
 % Figures for seed coat color
-\manhattanplot{seed_coat_color_all}
-\signalplot{seed_coat_color_all_I}
+\manhattanplot{seed_coat_color_all}{seed coat color (first GWAS)}{\getvar{seed_coat_color_all_nsamples}}
+\signalplot{seed_coat_color_all_I}{seed coat color (first GWAS)}{\emph{I}}%
+{Vertical dotted lines indicate the boundaries of the tandem duplication/inversion identified as the causal variant at this locus.}
 
-\manhattanplot{seed_coat_color_greenyellow}
-\signalplot{seed_coat_color_greenyellow_G}
-\geneplot{seed_coat_color_greenyellow_G}
+\manhattanplot{seed_coat_color_greenyellow}{seed coat color (second GWAS)}{\getvar{seed_coat_color_greenyellow_nsamples}}
+\signalplot{seed_coat_color_greenyellow_G}{seed coat color (second GWAS)}{\emph{G}}%
+{Vertical dotted lines indicate the location of the \getvar{seed_coat_color_greenyellow_G_gene} gene associated with the locus.}
+\geneplot{seed_coat_color_greenyellow_G}{seed coat color (second GWAS)}{\emph{G}}{\getvar{seed_coat_color_greenyellow_G_gene} gene}
 \kmerplot{seed_coat_color_greenyellow_G}
@@
 % Figures for stem termination
-\manhattanplot{stem_termination_all}
-\signalplot{stem_termination_all_Dt1}
-\geneplot{stem_termination_all_Dt1}
+\manhattanplot{stem_termination_all}{stem termination type (first GWAS)}{\getvar{stem_termination_all_nsamples}}
+\signalplot{stem_termination_all_Dt1}{stem termination type (first GWAS)}{\emph{Dt1}}%
+{Vertical dotted lines indicate the location of the \getvar{stem_termination_all_Dt1_gene} gene associated with the locus.}
+\geneplot{stem_termination_all_Dt1}{stem termination type (first GWAS)}{\emph{Dt1}}{\getvar{stem_termination_all_Dt1_gene} gene}
 \kmerplot{stem_termination_all_Dt1}
 
-\manhattanplot{stem_termination_sn}
-\ldplot{stem_termination_sn}
-\signalplot{stem_termination_sn_stGm11}
-\signalplot{stem_termination_sn_stGm16}
-\signalplot{stem_termination_sn_stGm18}
+\manhattanplot{stem_termination_sn}{stem termination type (second GWAS)}{\getvar{stem_termination_sn_nsamples}}
+\ldplot{stem_termination_sn}{stem termination type (second GWAS)}{\getvar{stem_termination_sn_ldkmers}}
 
 % Figures for hilum color
-\manhattanplot{hilum_color_all}
-\signalplot{hilum_color_all_T}
-\geneplot{hilum_color_all_T}
+\manhattanplot{hilum_color_all}{hilum color (first GWAS)}{\getvar{hilum_color_all_nsamples}}
+\signalplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}%
+{Vertical dotted lines indicate the location of the \getvar{hilum_color_all_T_gene} gene associated with the locus.}
+\geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene}
 \kmerplot{hilum_color_all_T}
-\signalplot{hilum_color_all_I}
-\geneplot{hilum_color_all_I}
+\signalplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}%
+{Vertical dotted lines indicate the boundaries of the tandem duplication/inversion identified as the causal variant at this locus.}
+\geneplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}{causal tandem duplication/inversion}
 
-\manhattanplot{hilum_color_blackbrown}
-\signalplot{hilum_color_blackbrown_R}
-\geneplot{hilum_color_blackbrown_R}
+\manhattanplot{hilum_color_blackbrown}{hilum color (second GWAS)}{\getvar{hilum_color_blackbrown_nsamples}}
+\signalplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}%
+{Vertical dotted lines indicate the location of the \getvar{hilum_color_blackbrown_R_gene} gene associated with the locus.}
+\geneplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}{\getvar{hilum_color_blackbrown_R_gene}}
 \kmerplot{hilum_color_blackbrown_R}
 
-\manhattanplot{hilum_color_rbr}
+\manhattanplot{hilum_color_rbr}{hilum color (third GWAS)}{\getvar{hilum_color_rbr_nsamples}}
 
 % Figures for pod color
-\manhattanplot{pod_color_all}
-\signalplot{pod_color_all_L1}
-\signalplot{pod_color_all_L2}
+\manhattanplot{pod_color_all}{pod color (first GWAS)}{\getvar{pod_color_all_nsamples}}
+\signalplot{pod_color_all_L1}{pod color (first GWAS)}{\emph{L1}}{}
+\signalplot{pod_color_all_L2}{pod color (first GWAS)}{\emph{L2}}{}
 
-\manhattanplot{pod_color_blbr}
-\ldplot{pod_color_blbr}
-\signalplot{pod_color_blbr_pdcGm15}
+\manhattanplot{pod_color_blbr}{pod color (second GWAS)}{\getvar{pod_color_blbr_nsamples}}
+\ldplot{pod_color_blbr}{pod color (second GWAS)}{\getvar{pod_color_blbr_ldkmers}}
+\signalplot{pod_color_blbr_pdcGm15}{pod color (second GWAS)}{newly suggested \emph{pdcGm15}}{}
 
 % Figures for pubescence form
-\manhattanplot{pubescence_form_all}
-\signalplot{pubescence_form_all_Pa1}
-\geneplot{pubescence_form_all_Pa1}
+\manhattanplot{pubescence_form_all}{pubescence form (first GWAS)}{\getvar{pubescence_form_all_nsamples}}
+\signalplot{pubescence_form_all_Pa1}{pubescence form (first GWAS)}{\emph{Pa1}}%
+{Vertical dotted lines indicate the location of the \getvar{pubescence_form_all_Pa1_gene} gene that we suggest as a candidate for this locus.}
+\geneplot{pubescence_form_all_Pa1}{pubescence form (first GWAS)}{\emph{Pa1}}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}
 \kmerplot{pubescence_form_all_Pa1}
-\signalplot{pubescence_form_all_Pa2}
-\ldplot{pubescence_form_all}
+\signalplot{pubescence_form_all_Pa2}{pubescence form (first GWAS)}{\emph{Pa2}}{}
+\ldplot{pubescence_form_all}{pubescence form (first GWAS)}{\getvar{pubescence_form_all_ldkmers}}
 
-\manhattanplot{pubescence_form_noerect}
+\manhattanplot{pubescence_form_noerect}{pubescence form (second GWAS)}{\getvar{pubescence_form_noerect_nsamples}}
 
 % Figures for pubescence density
-\manhattanplot{pubescence_density}
-\signalplot{pubescence_density_Ps}
-\geneplot{pubescence_density_Ps}
+\manhattanplot{pubescence_density}{pubescence density}{\getvar{pubescence_density_nsamples}}
+\signalplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}%
+{Vertical dotted lines indicate the boundaries of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene associated with the locus.}
+\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene}
 
 % Figures for seed coat luster
-\manhattanplot{seed_coat_luster_all}
+\manhattanplot{seed_coat_luster_all}{seed coat luster (first GWAS)}{\getvar{seed_coat_luster_all_nsamples}}
 
-\manhattanplot{seed_coat_luster_nointermediate}
+\manhattanplot{seed_coat_luster_nointermediate}{seed coat luster (second GWAS)}{\getvar{seed_coat_luster_nointermediate_nsamples}}
 
-\manhattanplot{seed_coat_luster_dullshiny}
-\ldplot{seed_coat_luster_dullshiny}
-\signalplot{seed_coat_luster_dullshiny_B}
-\geneplot{seed_coat_luster_dullshiny_B}
+\manhattanplot{seed_coat_luster_dullshiny}{seed coat luster (third GWAS)}{\getvar{seed_coat_luster_dullshiny_nsamples}}
+\ldplot{seed_coat_luster_dullshiny}{seed coat luster (third GWAS)}{\getvar{seed_coat_luster_dullshiny_ldkmers}}
+\signalplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}%
+{Vertical dotted lines indicate the boundaries of the causal CNV associated with the locus.}
+\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV}
 
 % Figures for corrected dry weight
-\manhattanplot{corrected_dry_weight}
-\ldplot{corrected_dry_weight}
-\signalplot{corrected_dry_weight_cdwGm15}
+\manhattanplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae}}{\getvar{corrected_dry_weight_nsamples}}
+\ldplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae}}{\getvar{corrected_dry_weight_ldkmers}}
+\signalplot{corrected_dry_weight_cdwGm15}{corrected dry weight (resistance to \emph{P. sojae}}{\emph{cdwGm15}}%
+{Vertical dotted lines indicate the location of the \getvar{corrected_dry_weight_cdwGm15_gene} gene suggested by de Ronne et al. (2022) as associated with the locus.}
 
 % Figures for maturity group
-\manhattanplot{maturity_group}
+\manhattanplot{maturity_group}{maturity group}{\getvar{maturity_group_nsamples}}
 
 % Figures for oil
-\manhattanplot{oil}
+\manhattanplot{oil}{seed oil content}{\getvar{oil_nsamples}}
 
 % Figures for protein
-\manhattanplot{protein}
+\manhattanplot{protein}{seed protein content}{\getvar{protein_nsamples}}
 

### 2022-10-20 additional_files/make_variables.R

@@ -1,0 +1,37 @@
+# Creating a list of key-value pairs that will be made available to
+# additional file 1 so that they can be read in LaTeX and output
+# in the file
+
+suppressMessages(library(GenomicRanges))
+
+# First making a list with the number of k-mers used for LD plots
+# DEPENDENCY: LD analysis
+ld_files <- dir("gwas_results/kmers", pattern = "clustered_ld.txt$", full.names = TRUE)
+
+ld_length <- sapply(ld_files, function(x) length(readLines(x)))
+names(ld_length) <- sub("_clustered_ld.txt", "", basename(names(ld_length)))
+names(ld_length) <- paste0(names(ld_length), "_ldkmers")
+
+# Then making a list of the number of samples used for GWAS analyses
+# DEPENDENCY: phenotypic_data/phenotypic_data.csv
+phenodata <- read.table("phenotypic_data/phenotypic_data.csv", sep = ";", header = TRUE, stringsAsFactors = FALSE)
+# DEPENDENCY: utilities/trait_names.txt
+trait_names <- readLines("utilities/trait_names.txt")
+trait_length <- sapply(trait_names, function(x, dataset) sum(!is.na(dataset[[x]])), dataset = phenodata)
+names(trait_length) <- paste0(names(trait_length), "_nsamples")
+
+# Now making a list of the genes associated with each trait/locus combination
+# DEPENDENCY: utilities/all_signals.rds
+all_signals <- readRDS("utilities/all_signals.rds")
+all_signals <- all_signals[!is.na(all_signals$gene_name_v4)]
+genes <- all_signals$gene_name_v4
+genes <- sub("^g", "G", genes)
+names(genes) <- paste0(names(all_signals), "_gene")
+
+# Grouping all the key-value pairs together
+key_values <- c(ld_length, trait_length, genes)
+
+# Reformatting them using the " = " separator
+key_values <- paste0(names(key_values), " = ", key_values)
+
+writeLines(key_values, con = "additional_files/variables.txt")

### 2022-10-20 gwas_results/ld_analysis.R

@@ -42,9 +42,9 @@ writeLines(significant_kmers, con = kmers_file)
 command <- paste0(filter_kmers, " -t ", kmers_table, " -k ", kmers_file, " -o ", pav_file)
 system(command)
 
-# Reading the PAV table from file and removing variants that have exactly the same PAV profile
+# Reading the PAV table from file and removing variants that have exactly the same PAV profile (removed this: now all variants are considered)
 pav_table <- read.table(pav_file, header = TRUE)
-pav_table <- pav_table[!duplicated(pav_table[, -1]), ]
+#pav_table <- pav_table[!duplicated(pav_table[, -1]), ]
 
 # Writing this de-duplicated PAV table to file
 write.table(pav_table, file = nodup_pav_file, sep = "\t",

### 2022-10-20 additional_files/additional_file_1.tex
Add captions for k-mer plots in additional file 1

@@
 % Defining a new command for k-mer plot figures
-\newcommand{\kmerplot}[1]{
+\newcommand{\kmerplot}[4]{
 	\begin{cfigure}
 		\includegraphics[width = 6.5in]{#1_kmers}
-		\caption{#1\_kmers}
+		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
+		{Identification of variants underlying significant \emph{k}-mers at the #2 associated with the #3 locus for #4.
+		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
+		Coding sequences are shown in blue and the direction of translation is indicated by arrows.
+		The red rectangle identifies the region that is zoomed-in in panel (b).
+		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
+		Individual nucleotides are colored according to the $-\log_{10}(p-value)$ of the most significant \emph{k}-mer overlapping them.
+		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
+		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes corresponds to those shown in panel (b).}
 		\label{#1_kmers}
@@
 \geneplot{flower_color_W1}{flower color}{\emph{W1}}{\getvar{flower_color_W1_gene} gene}
-\kmerplot{flower_color_W1}
+\kmerplot{flower_color_W1}{\getvar{flower_color_W1_gene} gene}{\emph{W1}}{flower color}
 
@@
 \geneplot{pubescence_color_all_T}{pubescence color (first GWAS)}{\emph{T}}{\getvar{pubescence_color_all_T_gene} gene}
-\kmerplot{pubescence_color_all_T}
+\kmerplot{pubescence_color_all_T}{\getvar{pubescence_color_all_T_gene} gene}{\emph{T}}{pubescence color (first GWAS)}
 \ldplot{pubescence_color_all}{pubescence color (first GWAS)}{\getvar{pubescence_color_all_ldkmers}}
@@
 \geneplot{pubescence_color_nogray_Td}{pubescence color (second GWAS)}{\emph{Td}}{\getvar{pubescence_color_nogray_Td_gene} gene}
-\kmerplot{pubescence_color_nogray_Td}
+\kmerplot{pubescence_color_nogray_Td}{\getvar{pubescence_color_nogray_Td_gene} gene}{\emph{Td}}{pubescence color (second GWAS)}
 
@@
 \geneplot{seed_coat_color_greenyellow_G}{seed coat color (second GWAS)}{\emph{G}}{\getvar{seed_coat_color_greenyellow_G_gene} gene}
-\kmerplot{seed_coat_color_greenyellow_G}
+\kmerplot{seed_coat_color_greenyellow_G}{\getvar{seed_coat_color_greenyellow_G_gene} gene}{\emph{G}}{seed coat color (second GWAS)}
 
@@
 \geneplot{stem_termination_all_Dt1}{stem termination type (first GWAS)}{\emph{Dt1}}{\getvar{stem_termination_all_Dt1_gene} gene}
-\kmerplot{stem_termination_all_Dt1}
+\kmerplot{stem_termination_all_Dt1}{\getvar{stem_termination_all_Dt1_gene} gene}{\emph{Dt1}}{stem termination type (first GWAS)}
 
@@
 \geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene}
-\kmerplot{hilum_color_all_T}
+\kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene} gene}{\emph{T}}{hilum color (first GWAS)}
 \signalplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}%
@@
 \geneplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}{\getvar{hilum_color_blackbrown_R_gene}}
-\kmerplot{hilum_color_blackbrown_R}
+\kmerplot{hilum_color_blackbrown_R}{\getvar{hilum_color_blackbrown_R_gene} gene}{\emph{R}}{hilum color (second GWAS)}
 
@@
 \geneplot{pubescence_form_all_Pa1}{pubescence form (first GWAS)}{\emph{Pa1}}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}
-\kmerplot{pubescence_form_all_Pa1}
+\kmerplot{pubescence_form_all_Pa1}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}{\emph{Pa1}}{pubescence form (first GWAS)}
 \signalplot{pubescence_form_all_Pa2}{pubescence form (first GWAS)}{\emph{Pa2}}{}

### 2022-10-20 additional_files/additional_file_1.tex
Complete a first draft of the table and figure captions in Additional file 1

@@
 % Defining a command to generate a table of phenotypic data for a single gwas
-\newcommand{\gwastableone}[1]{
+\newcommand{\gwastableone}[2]{
 \begin{table}
-	\caption{Numerical coding of phenotypic data used for the GWAS analysis of #1}
+	\caption[Phenotypic data used for the GWAS analysis of #2]%
+	{Phenotype frequency and numerical coding of phenotypic data used for the GWAS analysis of #2.
+	``NA'' indicates a missing value in the source data.
+	Dashes indicate that this phenotype was not used in the GWAS analysis.}
 	\label{#1_gwas_table}
@@
 		tabular = llc,
-		table head = \toprule Phenotype & Frequency & GWAS \#1 \\ \midrule,
+		table head = \toprule Phenotype & Frequency & GWAS \\ \midrule,
 		table foot = \bottomrule]%
@@
 
-			   \begin{tablenotes}
-			   \item[a] test
-			   \end{tablenotes}
 	\end{threeparttable}
@@
 % Defining a command to generate a table of phenotypic data for two gwas
-\newcommand{\gwastabletwo}[1]{
+\newcommand{\gwastabletwo}[2]{
 \begin{table}
-	\caption{Numerical coding of phenotypic data used for the GWAS analysis of #1}
+	\caption[Phenotypic data used for the GWAS analyses of #2]%
+	{Phenotype frequency and numerical coding of phenotypic data used in two GWAS analyses of #2.
+	``NA'' indicates a missing value in the source data.
+	Dashes indicate that this phenotype was not used in the GWAS analysis.}
 	\label{#1_gwas_table}
@@
 % Defining a command to generate a table of phenotypic data for three gwas
-\newcommand{\gwastablethree}[1]{
+\newcommand{\gwastablethree}[2]{
 \begin{table}
-	\caption{Numerical coding of phenotypic data used for the GWAS analysis of #1}
+	\caption[Phenotypic data used for the GWAS analyses of #2]%
+	{Phenotype frequency and numerical coding of phenotypic data used in three GWAS analyses of #2.
+	``NA'' indicates a missing value in the source data.
+	Dashes indicate that this phenotype was not used in the GWAS analysis.}
 	\label{#1_gwas_table}
@@
 		\begin{TableNotes}
-		\item[a] test
+		\item[a] Most significant -log\textsubscript{10}(p) reported at this locus by the study in the Study column
+		\item[b] Gene associated with the locus. An question mark following the identifier of the gene indicates that this is simply a candidate
+			which has yet to be confirmed. A dash indicates that no candidate has been suggested yet.
+		\item[c] Most significant -log\textsubscript{10}(p) reported by Platypus at this locus. A dash indicates that no signal
+			was detected by Platypus at this locus. The same applies for other methods (vg, Paragraph, \emph{k}-mers) in
+			their respective columns.
 		\end{TableNotes}
@@
 
-			\caption{Signals}\\
+			\caption[Description of loci associated with the traits studied]%
+			{Description of loci associated with the traits studied and -log\textsubscript{10}(p) of significant
+			signals detected in our study using four different genotypic datasets.}\\
 			\label{signals_table}\\
@@
 			\toprule
-			Trait & Locus & Chromosome & -log10(p) & Gene & Platypus & vg & Paragraph & k-mers  & Study \\
+			Trait & Locus & Chromosome & -log\textsubscript{10}(p)\tnote{a} & Gene\tnote{b} & Platypus\tnote{c} & vg & Paragraph & \emph{k}-mers  & Study \\
 			\midrule \endfirsthead
 
-			\caption[]{Signals \emph{(continued)}}\\
+			\caption[]{Loci associated with the traits studied \emph{(continued)}}\\
 
@@
 
-\gwastableone{FLOWER.COLOR}
-\gwastabletwo{PUBESCENCE.COLOR}
-\gwastabletwo{SEED.COAT.COLOR}
-\gwastabletwo{STEM.TERMINATION.TYPE}
-\gwastablethree{HILUM.COLOR}
-\gwastabletwo{POD.COLOR}
-\gwastabletwo{PUBESCENCE.FORM}
-\gwastableone{PUBESCENCE.DENSITY}
-\gwastablethree{SEED.COAT.LUSTER}
-\gwastableone{MATURITY.GROUP}
+\gwastableone{FLOWER.COLOR}{flower color}
+\gwastabletwo{PUBESCENCE.COLOR}{pubescence color}
+\gwastabletwo{SEED.COAT.COLOR}{seed coat color}
+\gwastabletwo{STEM.TERMINATION.TYPE}{stem termination type}
+\gwastablethree{HILUM.COLOR}{hilum color}
+\gwastabletwo{POD.COLOR}{pod color}
+\gwastabletwo{PUBESCENCE.FORM}{pubescence form}
+\gwastableone{PUBESCENCE.DENSITY}{pubescence density}
+\gwastablethree{SEED.COAT.LUSTER}{seed coat luster}
+\gwastableone{MATURITY.GROUP}{Maturity group}
 

### 2022-10-20 tables/gwas_table.R

@@ -24,6 +24,7 @@ format_table <- function(phenodata, phenotype, lookup_tables) {
 
 	for(i in 1:length(lookup_tables)) {
 		x[[paste0("GWAS", letters[i])]] <- lookup_tables[[i]][x$Value]
+		x[[paste0("GWAS", letters[i])]][is.na(x[[paste0("GWAS", letters[i])]])] <- "-"
 	}
 
 	return(x)

### 2022-10-24 additional_files/additional_file_1.tex
Make various improvements to supplemental tables and figures to make them publication-ready

@@
 	\includegraphics[width = 6in]{#1_manhattan}
-	\caption[Manhattan plot of #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
+	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
 	(a) Platypus (SNPs and indels), (b) vg (SVs), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
@@
 	\includegraphics[width = 6in]{#1_signal}
-	\caption[Zoomed-in Manhattan plot at #3 locus for #2]{Zommed-in Manhattan plots generated from the GWAS analysis of #2 at the #3 locus using four genotype datasets:
+	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zommed-in Manhattan plots generated from the GWAS analysis of #2 at the #3 locus using four genotype datasets:
 	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
@@
 	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph and vg) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
+	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
+	\emph{p}-values were computed after signal discovery.
 	}
@@
 	\includegraphics[width = 6in]{#1_gene}
-	\caption[Zommed-in Manhattan plot at gene(s) associated with #3 locus for #2]%
+	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
 	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using four genotype datasets:
 	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
+	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
+	\emph{p}-values were computed after signal discovery.
 	Panel (a) shows the location of annotated gene models over the plotting interval.
@@
 % Information for the title page
-\title{Supplemental data to ``ARTICLE TITLE''}
-\author{Marc-André Lemay \and François Belzile}
+\title{Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''}
+\author{Marc-André Lemay \and Jonas A. Sibbesen \and Maxime de Ronne \and Richard Bélanger \and François Belzile}
 \date{}
@@
 \maketitle \thispagestyle{empty}
-
-\tableofcontents
-
-\vspace{4ex}
 
@@
 		\begin{TableNotes}
-		\item[a] Most significant -log\textsubscript{10}(p) reported at this locus by the study in the Study column
-		\item[b] Gene associated with the locus. An question mark following the identifier of the gene indicates that this is simply a candidate
-			which has yet to be confirmed. A dash indicates that no candidate has been suggested yet.
-		\item[c] Most significant -log\textsubscript{10}(p) reported by Platypus at this locus. A dash indicates that no signal
+		\item[a] Most significant -log\textsubscript{10}(\emph{p}) reported at this locus by the study in the Study column
+		\item[b] Gene associated with the locus. A question mark following the identifier of the gene indicates that this is simply a candidate
+			that has yet to be confirmed. A dash indicates that no candidate has been suggested yet.
+		\item[c] Most significant -log\textsubscript{10}(\emph{p}) reported by Platypus at this locus. A dash indicates that no signal
 			was detected by Platypus at this locus. The same applies for other methods (vg, Paragraph, \emph{k}-mers) in
 			their respective columns.
+		\item[d] The \emph{p}-value at the W1 locus using the \emph{k}-mers method was so small that it was numerically equal
+			to zero. This value was therefore set to the -log\textsubscript{10} of the smallest value that could be represented
+			in R using our machine.
 		\end{TableNotes}
@@
 			\caption[Description of loci associated with the traits studied]%
-			{Description of loci associated with the traits studied and -log\textsubscript{10}(p) of significant
-			signals detected in our study using four different genotypic datasets.}\\
+			{Description of loci associated with the traits studied and -log\textsubscript{10}(\emph{p}) of significant
+			signals detected in our study using four different genotypic datasets (Platypus, vg, Paragraph, \emph{k}-mers).}\\
 			\label{signals_table}\\
@@
 			\toprule
-			Trait & Locus & Chromosome & -log\textsubscript{10}(p)\tnote{a} & Gene\tnote{b} & Platypus\tnote{c} & vg & Paragraph & \emph{k}-mers  & Study \\
+			Trait & Locus & Chromosome & -log\textsubscript{10}(\emph{p})\tnote{a} & Gene\tnote{b} & Platypus\tnote{c} & vg & Paragraph & \emph{k}-mers  & Study \\
 			\midrule \endfirsthead
@@
 
-\geneplot{flower_color_W1}{flower color}{\emph{W1}}{\getvar{flower_color_W1_gene} gene}
+\geneplot{flower_color_W1}{flower color}{\emph{W1}}{\getvar{flower_color_W1_gene} gene (a)}
 \kmerplot{flower_color_W1}{\getvar{flower_color_W1_gene} gene}{\emph{W1}}{flower color}
@@
 {Vertical dotted lines indicate the location of the \getvar{pubescence_color_all_T_gene} gene associated with the locus.}
-\geneplot{pubescence_color_all_T}{pubescence color (first GWAS)}{\emph{T}}{\getvar{pubescence_color_all_T_gene} gene}
+\geneplot{pubescence_color_all_T}{pubescence color (first GWAS)}{\emph{T}}{\getvar{pubescence_color_all_T_gene} gene (a)}
 \kmerplot{pubescence_color_all_T}{\getvar{pubescence_color_all_T_gene} gene}{\emph{T}}{pubescence color (first GWAS)}
@@
 {Vertical dotted lines indicate the location of the \getvar{pubescence_color_nogray_Td_gene} gene associated with the locus.}
-\geneplot{pubescence_color_nogray_Td}{pubescence color (second GWAS)}{\emph{Td}}{\getvar{pubescence_color_nogray_Td_gene} gene}
+\geneplot{pubescence_color_nogray_Td}{pubescence color (second GWAS)}{\emph{Td}}{\getvar{pubescence_color_nogray_Td_gene} gene (a)}
 \kmerplot{pubescence_color_nogray_Td}{\getvar{pubescence_color_nogray_Td_gene} gene}{\emph{Td}}{pubescence color (second GWAS)}
@@
 {Vertical dotted lines indicate the location of the \getvar{seed_coat_color_greenyellow_G_gene} gene associated with the locus.}
-\geneplot{seed_coat_color_greenyellow_G}{seed coat color (second GWAS)}{\emph{G}}{\getvar{seed_coat_color_greenyellow_G_gene} gene}
+\geneplot{seed_coat_color_greenyellow_G}{seed coat color (second GWAS)}{\emph{G}}{\getvar{seed_coat_color_greenyellow_G_gene} gene (a)}
 \kmerplot{seed_coat_color_greenyellow_G}{\getvar{seed_coat_color_greenyellow_G_gene} gene}{\emph{G}}{seed coat color (second GWAS)}
@@
 {Vertical dotted lines indicate the location of the \getvar{stem_termination_all_Dt1_gene} gene associated with the locus.}
-\geneplot{stem_termination_all_Dt1}{stem termination type (first GWAS)}{\emph{Dt1}}{\getvar{stem_termination_all_Dt1_gene} gene}
+\geneplot{stem_termination_all_Dt1}{stem termination type (first GWAS)}{\emph{Dt1}}{\getvar{stem_termination_all_Dt1_gene} gene (a)}
 \kmerplot{stem_termination_all_Dt1}{\getvar{stem_termination_all_Dt1_gene} gene}{\emph{Dt1}}{stem termination type (first GWAS)}
@@
 {Vertical dotted lines indicate the location of the \getvar{hilum_color_all_T_gene} gene associated with the locus.}
-\geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene}
+\geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene (a)}
 \kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene} gene}{\emph{T}}{hilum color (first GWAS)}
@@
 {Vertical dotted lines indicate the boundaries of the tandem duplication/inversion identified as the causal variant at this locus.}
-\geneplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}{causal tandem duplication/inversion}
+\geneplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}{causal tandem duplication/inversion region (a)}
 
@@
 {Vertical dotted lines indicate the location of the \getvar{hilum_color_blackbrown_R_gene} gene associated with the locus.}
-\geneplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}{\getvar{hilum_color_blackbrown_R_gene}}
+\geneplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}{\getvar{hilum_color_blackbrown_R_gene} gene (a)}
 \kmerplot{hilum_color_blackbrown_R}{\getvar{hilum_color_blackbrown_R_gene} gene}{\emph{R}}{hilum color (second GWAS)}
@@
 {Vertical dotted lines indicate the location of the \getvar{pubescence_form_all_Pa1_gene} gene that we suggest as a candidate for this locus.}
-\geneplot{pubescence_form_all_Pa1}{pubescence form (first GWAS)}{\emph{Pa1}}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}
+\geneplot{pubescence_form_all_Pa1}{pubescence form (first GWAS)}{\emph{Pa1}}{\getvar{pubescence_form_all_Pa1_gene} candidate gene (a) putatively}
 \kmerplot{pubescence_form_all_Pa1}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}{\emph{Pa1}}{pubescence form (first GWAS)}
@@
 {Vertical dotted lines indicate the boundaries of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene associated with the locus.}
-\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene}
+\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene (a)}
 
@@
 {Vertical dotted lines indicate the boundaries of the causal CNV associated with the locus.}
-\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV}
+\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV (a)}
 
 % Figures for corrected dry weight
-\manhattanplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae}}{\getvar{corrected_dry_weight_nsamples}}
-\ldplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae}}{\getvar{corrected_dry_weight_ldkmers}}
-\signalplot{corrected_dry_weight_cdwGm15}{corrected dry weight (resistance to \emph{P. sojae}}{\emph{cdwGm15}}%
+\manhattanplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae})}{\getvar{corrected_dry_weight_nsamples}}
+\ldplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae})}{\getvar{corrected_dry_weight_ldkmers}}
+\signalplot{corrected_dry_weight_cdwGm15}{corrected dry weight (resistance to \emph{P. sojae})}{\emph{cdwGm15}}%
 {Vertical dotted lines indicate the location of the \getvar{corrected_dry_weight_cdwGm15_gene} gene suggested by de Ronne et al. (2022) as associated with the locus.}

### 2022-10-24 cnv_analysis/i_analysis.R

@@ -2,8 +2,8 @@
 # tandem-inverted I locus based on its alignment with bwa
 
 # Loading the required libraries
-library(Rsamtools)
-library(GenomicRanges)
+suppressMessages(library(Rsamtools))
+suppressMessages(library(GenomicRanges))
 
 # Loading the aligned I locus
 # DEPENDENCY: cnv_analysis/i_locus.bam

### 2022-10-24 figures/gene_subplot.R

@@ -72,14 +72,26 @@ signal <- subsetByOverlaps(gwas_signals, gene)
 
 if(!length(signal) == 1) warning("There is no signal overlapping the gene ", gene, " for locus ", id)
 
+# Setting the feature optionally (just for genes for which we want to highlight specific variants)
+# DEPENDENCY: utilities/variant_ranges.txt
+variant_ranges <- read.table("utilities/variant_ranges.txt", head = FALSE, sep = "\t", stringsAsFactors = FALSE)
+key <- paste0(program, "_", id)
+
+if(length(index <- which(key == variant_ranges[[1]]))) {
+	feature <- as(variant_ranges[index, 2], "GRanges")
+} else {
+	feature <- NULL
+}
+
 # Using the right viewport to focus on the p-values over the length of the gene
 ptx_plot <- pvalueGrob(gwas_results = gwas_results,
 		       interval = gene,
 		       shading = cnv_range,
-		       feature = NULL,
+		       feature = feature,
 		       threshold = threshold,
 		       col = "blue",
 		       pruned_col = if(program == "platypus") "firebrick2" else NULL,
+		       cex.points = 0.7,
 		       yexpand = c(0.1, 0.1))
 
 # Saving the grob to an RDS file for retrieval later on

### 2022-10-24 figures/kmer_plot.R

@@ -140,9 +140,6 @@ png(paste0("figures/", locus, "_kmers.png"), width = 8, height = 8, units = "in"
 # Initializing the device
 grid.newpage()
 
-# Drawing a box around the plotting region
-grid::grid.rect()
-
 # Dividing the viewport into five row viewports:
 # - The first viewport is used to plot the transcript for the gene and the highlighted region
 # - The second viewport acts as a buffer between the first and third
@@ -161,6 +158,7 @@ if(!is.null(causal_gene)) {
 				  xscale = causal_gene,
 				  highlight = GenomicRanges::GRanges(seqnames = chrom,
 								     ranges = IRanges::IRanges(start = grange[1], end = grange[2])),
+				  strand_colors = c("dodgerblue", "dodgerblue"),
 				  draw_arrows = TRUE,
 				  first_tx_only = TRUE))
 	grid::upViewport()

### 2022-10-24 figures/ld_plot.R

@@ -27,7 +27,7 @@ gsorted_ld <- ld_sort(clustered_ld, kmer_positions, sort_param = "position")
 png(paste0("figures/", trait, "_ld.png"), width = 9, height = 10, units = "in", res = 400)
 
 grid.newpage()
-ld_plot(gsorted_ld, kmer_positions, top_legend = FALSE, ylabels = TRUE, ylabel_pattern = "^Gm[0-9]{2}$")
+ld_plot(gsorted_ld, kmer_positions, top_legend = FALSE, ylabels = TRUE, ylabel_pattern = "^Gm[0-9]{2}$", fontsize = 10)
 
 dev.off()
 

### 2022-10-24 figures/signal_subplot.R

@@ -61,7 +61,7 @@ threshold <- -log10(as.numeric(readLines(paste0("gwas_results/", program, "/", t
 
 if(!length(signal)) {
 	warning("No signal found by ", program, " for signal ID ", id)
-	ptx_plot <- grid::gTree(children = gList(textGrob(paste0("No signal found by ", program, " for signal ID ", id))))
+	ptx_plot <- grid::gTree(children = gList(textGrob(paste0("No signal found by ", program, " at this locus."))))
 } else {
 	# Setting the x-scale while handling special cases where there is no signal or the signal is only 1 bp wide
 	if(length(signal) == 1 && width(signal) == 1) {
@@ -76,6 +76,7 @@ if(!length(signal)) {
 			       threshold = threshold,
 			       col = "blue",
 			       pruned_col = if(program == "platypus") "firebrick2" else NULL,
+			       cex.points = 0.6,
 			       shading = top_markers,
 			       yexpand = c(0.1, 0.1))
 }

### 2022-10-24 reference_signals/format_bandillo2015_signals.R

@@ -2,10 +2,10 @@
 # as GRanges objects with coordinates relative to Williams82 genome assembly version 4
 
 # Loading required libraries
-library(parallel)
-library(GenomicRanges)
-library(Rsamtools)
-library(Biostrings)
+suppressMessages(library(parallel))
+suppressMessages(library(GenomicRanges))
+suppressMessages(library(Rsamtools))
+suppressMessages(library(Biostrings))
 
 # The padding (number of nucleotides to be extracted from either side of the position) to use as a parameter
 padding <- 20
@@ -68,6 +68,9 @@ bandillo2015_signals$gene_name_v4 <- NA
 # Setting the name of the gene for the QTL on chromosome 15
 # according to Zhang et al. (2020), DOI:10.1371/journal.pgen.1009114
 bandillo2015_signals[bandillo2015_signals$Chr == "glyma.Wm82.gnm1.Gm15", "gene_name_v4"] <- tolower("Glyma.15G049200")
+# Also for the one on chromosome Gm20 according to Fliege et al. (2022) : https://doi.org/10.1111/tpj.15658
+bandillo2015_signals[bandillo2015_signals$Chr == "glyma.Wm82.gnm1.Gm20", "gene_name_v4"] <- tolower("Glyma.20g085100")
+
 
 # Reordering the columns
 bandillo2015_signals <- bandillo2015_signals[, c("trait", "Chr", "n_snps",

### 2022-10-24 reference_signals/format_custom_signals.R

@@ -3,7 +3,7 @@
 # downstream analyses
 
 # Loading the GenomicRanges package
-library(GenomicRanges)
+suppressMessages(library(GenomicRanges))
 
 # Manually entering data for the signal found on chromosome 16 for pubescence color
 # This signal goes from 1429763-1436787 for kmers, 1430025-1434520 for Paragraph, and 1380791-1485581 for Platypus

### 2022-10-24 reference_signals/format_deronne2022_signal.R

@@ -3,7 +3,7 @@
 # DOI:10.1002/tpg2.20184
 
 # Loading the GenomicRanges package
-library(GenomicRanges)
+suppressMessages(library(GenomicRanges))
 
 # Manually entering data for the signal found by de Ronne et al. (2021) with BLINK on Gm15
 deronne_signal <- GRanges(seqnames = "Gm15",

### 2022-10-24 tables/signals_table.R

@@ -1,7 +1,7 @@
 # Creating a table that shows which previously discovered signals were found by our analyses
 
 # Loading required libraries
-library(GenomicRanges)
+suppressMessages(library(GenomicRanges))
 
 # Loading the GRanges object of known signals
 # DEPENDENCY: utilities/all_signals.rds
@@ -18,6 +18,16 @@ signals_table <- unique(as.data.frame(all_signals)[, c("original_trait", "locus"
 						       "start", "end", "log_pvalue", "gene_name_v4")])
 rownames(signals_table) <- NULL
 
+# Ordering the rows of the table in the order in which they are presented in the manuscript
+signals_table$original_trait <- factor(signals_table$original_trait,
+				       levels = c("flower_color", "pubescence_color", "seed_coat_color", "stem_termination",
+						  "hilum_color", "pod_color", "pubescence_form", "pubescence_density",
+						  "seed_coat_luster", "maturity_group", "corrected_dry_weight",
+						  "oil", "protein"))
+
+signals_table <- signals_table[order(signals_table$original_trait, signals_table$log_pvalue, decreasing = c(FALSE, TRUE), method = "radix"), ]
+signals_table$original_trait <- as.character(signals_table$original_trait)
+
 # Creating columns for the log10 p-values observed for each program
 signals_table$platypus <- 0
 signals_table$vg <- 0
@@ -26,7 +36,6 @@ signals_table$kmers <- 0
 
 # For each of the loci, we want to get the most significant p-value detected by each approach
 for(i in 1:nrow(signals_table)) {
-	message("Processing row ", i)
 	i_locus <- signals_table[i, "locus"]
 	locus_signals <- all_signals[all_signals$locus == i_locus & all_signals$original_trait == signals_table[i, "original_trait"]]
 
@@ -35,7 +44,7 @@ for(i in 1:nrow(signals_table)) {
 
 		for(program in programs) {
 			psignals <- readRDS(paste0("gwas_results/", program, "/", j_signal$trait, "_signal.rds"))
-			psignals <- subsetByOverlaps(psignals, j_signal)
+			psignals <- subsetByOverlaps(psignals, j_signal, ignore.strand = TRUE)
 			if(length(psignals)) {
 				signals_table[i, program] <- max(signals_table[i, program], psignals$log10_p)
 			}
@@ -43,7 +52,7 @@ for(i in 1:nrow(signals_table)) {
 	}
 }
 
[…]

### 2022-10-24 utilities/make_signals_granges.R

@@ -1,7 +1,7 @@
 # This script combines the reference signals from various sources into a GRanges object
 # that can be used to extract signals found by GWAS analyses and plot them
 
-library(GenomicRanges)
+suppressMessages(library(GenomicRanges))
 
 # DEPENDENCY: reference_signals/bandillo2015_signals.rds
 # DEPENDENCY: reference_signals/bandillo2017_signals.rds

### 2022-10-24 additional_files/additional_file_1.tex
Improve the captions of gene and k-mer plots; restrict the extent of consensus sequence for k-mer plot of G locus

@@
 	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
-	Panel (a) shows the location of annotated gene models over the plotting interval.
+	Panel (a) shows gene models over the plotting interval.
 	#4
@@
 % Defining a new command for gene plot figures
-\newcommand{\geneplot}[4]{
+\newcommand{\geneplot}[5]{
 \begin{cfigure}
@@
 	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
+	#5
 	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
 	\emph{p}-values were computed after signal discovery.
-	Panel (a) shows the location of annotated gene models over the plotting interval.
+	Panel (a) shows gene models over the plotting interval.
 	Exons are represented by rectangles whereas introns are represented by horizontal lines.
@@
 % Defining a new command for k-mer plot figures
-\newcommand{\kmerplot}[4]{
+\newcommand{\kmerplot}[5]{
 	\begin{cfigure}
@@
 		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
-		{Identification of variants underlying significant \emph{k}-mers at the #2 associated with the #3 locus for #4.
+		{Identification of #5 underlying significant \emph{k}-mers at the #2 associated with the #3 locus for #4.
 		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
@@
 
-\geneplot{flower_color_W1}{flower color}{\emph{W1}}{\getvar{flower_color_W1_gene} gene (a)}
-\kmerplot{flower_color_W1}{\getvar{flower_color_W1_gene} gene}{\emph{W1}}{flower color}
+\geneplot{flower_color_W1}{flower color}{\emph{W1}}{\getvar{flower_color_W1_gene} gene}%
+{Vertical dotted lines in panels (c), (d) and (e) indicate the location of the causal SV at this locus.}
+\kmerplot{flower_color_W1}{\getvar{flower_color_W1_gene}}{\emph{W1}}{flower color}{a causal SV}
 
@@
 {Vertical dotted lines indicate the location of the \getvar{pubescence_color_all_T_gene} gene associated with the locus.}
-\geneplot{pubescence_color_all_T}{pubescence color (first GWAS)}{\emph{T}}{\getvar{pubescence_color_all_T_gene} gene (a)}
-\kmerplot{pubescence_color_all_T}{\getvar{pubescence_color_all_T_gene} gene}{\emph{T}}{pubescence color (first GWAS)}
+\geneplot{pubescence_color_all_T}{pubescence color (first GWAS)}{\emph{T}}{\getvar{pubescence_color_all_T_gene} gene}%
+{Vertical dotted lines in panels (b) and (e) indicate the location of the causal indel at this locus.}
+\kmerplot{pubescence_color_all_T}{\getvar{pubescence_color_all_T_gene} gene}{\emph{T}}{pubescence color (first GWAS)}{a causal indel}
 \ldplot{pubescence_color_all}{pubescence color (first GWAS)}{\getvar{pubescence_color_all_ldkmers}}
@@
 {Vertical dotted lines indicate the location of the \getvar{pubescence_color_nogray_Td_gene} gene associated with the locus.}
-\geneplot{pubescence_color_nogray_Td}{pubescence color (second GWAS)}{\emph{Td}}{\getvar{pubescence_color_nogray_Td_gene} gene (a)}
-\kmerplot{pubescence_color_nogray_Td}{\getvar{pubescence_color_nogray_Td_gene} gene}{\emph{Td}}{pubescence color (second GWAS)}
+\geneplot{pubescence_color_nogray_Td}{pubescence color (second GWAS)}{\emph{Td}}{\getvar{pubescence_color_nogray_Td_gene} gene}%
+{Vertical dotted lines in panels (b) and (e) indicate the location of the causal SNP at this locus.}
+\kmerplot{pubescence_color_nogray_Td}{\getvar{pubescence_color_nogray_Td_gene} gene}{\emph{Td}}{pubescence color (second GWAS)}{a causal SNP}
 
@@
 {Vertical dotted lines indicate the location of the \getvar{seed_coat_color_greenyellow_G_gene} gene associated with the locus.}
-\geneplot{seed_coat_color_greenyellow_G}{seed coat color (second GWAS)}{\emph{G}}{\getvar{seed_coat_color_greenyellow_G_gene} gene (a)}
-\kmerplot{seed_coat_color_greenyellow_G}{\getvar{seed_coat_color_greenyellow_G_gene} gene}{\emph{G}}{seed coat color (second GWAS)}
+\geneplot{seed_coat_color_greenyellow_G}{seed coat color (second GWAS)}{\emph{G}}{\getvar{seed_coat_color_greenyellow_G_gene} gene}%
+{Vertical dotted lines in panels (b) and (e) indicate the location of the causal SNP at this locus.}
+\kmerplot{seed_coat_color_greenyellow_G}{\getvar{seed_coat_color_greenyellow_G_gene} gene}{\emph{G}}{seed coat color (second GWAS)}{a causal SNP}
 
@@
 {Vertical dotted lines indicate the location of the \getvar{stem_termination_all_Dt1_gene} gene associated with the locus.}
-\geneplot{stem_termination_all_Dt1}{stem termination type (first GWAS)}{\emph{Dt1}}{\getvar{stem_termination_all_Dt1_gene} gene (a)}
-\kmerplot{stem_termination_all_Dt1}{\getvar{stem_termination_all_Dt1_gene} gene}{\emph{Dt1}}{stem termination type (first GWAS)}
+\geneplot{stem_termination_all_Dt1}{stem termination type (first GWAS)}{\emph{Dt1}}{\getvar{stem_termination_all_Dt1_gene} gene}%
+{Vertical dotted lines in panels (b) and (e) indicate the location of a causal SNP at this locus.}
+\kmerplot{stem_termination_all_Dt1}{\getvar{stem_termination_all_Dt1_gene} gene}{\emph{Dt1}}{stem termination type (first GWAS)}{a causal SNP}
 
@@
 {Vertical dotted lines indicate the location of the \getvar{hilum_color_all_T_gene} gene associated with the locus.}
-\geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene (a)}
-\kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene} gene}{\emph{T}}{hilum color (first GWAS)}
+\geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene}%
+{Vertical dotted lines in panels (b) and (e) indicate the location of the causal indel at this locus.}
+\kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene} gene}{\emph{T}}{hilum color (first GWAS)}{a causal indel}
 \signalplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}%
 {Vertical dotted lines indicate the boundaries of the tandem duplication/inversion identified as the causal variant at this locus.}
-\geneplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}{causal tandem duplication/inversion region (a)}
+\geneplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}{causal tandem duplication/inversion region}{}
 
@@
 {Vertical dotted lines indicate the location of the \getvar{hilum_color_blackbrown_R_gene} gene associated with the locus.}
-\geneplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}{\getvar{hilum_color_blackbrown_R_gene} gene (a)}
-\kmerplot{hilum_color_blackbrown_R}{\getvar{hilum_color_blackbrown_R_gene} gene}{\emph{R}}{hilum color (second GWAS)}
+\geneplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}{\getvar{hilum_color_blackbrown_R_gene} gene}%
+{Vertical dotted lines indicate the location a SNP (panels b and e) and an indel (panel e) that are putatively causal at this locus.}
+\kmerplot{hilum_color_blackbrown_R}{\getvar{hilum_color_blackbrown_R_gene} gene}{\emph{R}}{hilum color (second GWAS)}{a SNP and an indel that are putatively causal}
 
@@
 {Vertical dotted lines indicate the location of the \getvar{pubescence_form_all_Pa1_gene} gene that we suggest as a candidate for this locus.}
-\geneplot{pubescence_form_all_Pa1}{pubescence form (first GWAS)}{\emph{Pa1}}{\getvar{pubescence_form_all_Pa1_gene} candidate gene (a) putatively}
-\kmerplot{pubescence_form_all_Pa1}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}{\emph{Pa1}}{pubescence form (first GWAS)}
+\geneplot{pubescence_form_all_Pa1}{pubescence form (first GWAS)}{\emph{Pa1}}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}%
+{Vertical dotted lines in panels (b) and (e) indicate the location of two highly significant non-synonymous SNPs at this gene.}
+\kmerplot{pubescence_form_all_Pa1}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}{\emph{Pa1}}{pubescence form (first GWAS)}{a non-synonymous SNP}
 \signalplot{pubescence_form_all_Pa2}{pubescence form (first GWAS)}{\emph{Pa2}}{}
@@
 {Vertical dotted lines indicate the boundaries of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene associated with the locus.}
-\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene (a)}
+\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene}%
+{The vertical dotted line in panel (e) indicates the location of the most significant \emph{k}-mer, which is associated with the causal CNV at this locus.}
 
@@
 {Vertical dotted lines indicate the boundaries of the causal CNV associated with the locus.}
-\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV (a)}
+\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV}{}
 

### 2022-10-26 additional_files/additional_file_1.tex
First manuscript version sent out for review to co-authors

@@
 	The x-axis shows the position along the reference assembly version 4 of Williams82.
-	Each point shows the $-\log_{10} {p-value}$ associated with a particular marker or \emph{k}-mer.
-	Horizontal dotted lines indicate the 5\% family-wise error-rate threshold determined from a randomization approach.
+	Each point shows the $-\log_{10} (p)$ associated with a particular marker or \emph{k}-mer.
+	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
 	Vertical dotted lines indicate the position of signals associated with the trait.
-	Signals documented in previous studies are colored according to whether they were founs (green) or not (red) by a particular approach.
+	Documented loci are colored according to whether they were found (green) or not (red) by a particular approach.
 	The ``Gm'' prefix has been left out of chromosome names for simplicity.
@@
 	\includegraphics[width = 6in]{#1_signal}
-	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zommed-in Manhattan plots generated from the GWAS analysis of #2 at the #3 locus using four genotype datasets:
+	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zommed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
 	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
-	Panel (a) shows gene models over the plotting interval.
+	Panel (a) shows gene models over the genomic interval.
+	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
 	#4
@@
 	#5
+	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
 	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
@@
 	Exons are represented by rectangles whereas introns are represented by horizontal lines.
-	Coding sequences are shown in blue and the direction of translation is indicated by arrows.
+	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
 	}
@@
 		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
-		{Identification of #5 underlying significant \emph{k}-mers at the #2 associated with the #3 locus for #4.
+		{Identification of #5 underlying significant \emph{k}-mers at the #2 gene associated with the #3 locus for #4.
 		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
-		Coding sequences are shown in blue and the direction of translation is indicated by arrows.
+		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
 		The red rectangle identifies the region that is zoomed-in in panel (b).
 		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
-		Individual nucleotides are colored according to the $-\log_{10}(p-value)$ of the most significant \emph{k}-mer overlapping them.
+		Individual nucleotides are colored according to the $-\log_{10}(p)$ of the most significant \emph{k}-mer overlapping them.
 		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
-		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes corresponds to those shown in panel (b).}
+		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes correspond to those shown in panel (b).}
 		\label{#1_kmers}
@@
 		\emph{k}-mers are represented in the same order along the x- and y-axis.
-		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p-values)$ of each \emph{k}-mer.}
+		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p)$ of each \emph{k}-mer.}
 		\label{#1_ld}
@@
 
+\gwastableone{FLOWER.COLOR}{flower color}
+\gwastabletwo{PUBESCENCE.COLOR}{pubescence color}
+\gwastabletwo{SEED.COAT.COLOR}{seed coat color}
+\gwastabletwo{STEM.TERMINATION.TYPE}{stem termination type}
+\gwastablethree{HILUM.COLOR}{hilum color}
+\gwastabletwo{POD.COLOR}{pod color}
+\gwastabletwo{PUBESCENCE.FORM}{pubescence form}
+\gwastableone{PUBESCENCE.DENSITY}{pubescence density}
+\gwastablethree{SEED.COAT.LUSTER}{seed coat luster}
+\gwastableone{MATURITY.GROUP}{maturity group}
+
+% Table with all signals
 \begin{landscape} \footnotesize
@@
 		\begin{TableNotes}
-		\item[a] Most significant -log\textsubscript{10}(\emph{p}) reported at this locus by the study in the Study column
+		\item[a] Most significant -log\textsubscript{10}(\emph{p}) previously reported at this locus by the study in the Study column
 		\item[b] Gene associated with the locus. A question mark following the identifier of the gene indicates that this is simply a candidate
-			that has yet to be confirmed. A dash indicates that no candidate has been suggested yet.
+			that has yet to be confirmed. A dash indicates that no candidate has been suggested yet. We have only included candidate
+			genes that have been suggested in previous studies.
 		\item[c] Most significant -log\textsubscript{10}(\emph{p}) reported by Platypus at this locus. A dash indicates that no signal
@@
 
-\gwastableone{FLOWER.COLOR}{flower color}
-\gwastabletwo{PUBESCENCE.COLOR}{pubescence color}
-\gwastabletwo{SEED.COAT.COLOR}{seed coat color}
-\gwastabletwo{STEM.TERMINATION.TYPE}{stem termination type}
-\gwastablethree{HILUM.COLOR}{hilum color}
-\gwastabletwo{POD.COLOR}{pod color}
-\gwastabletwo{PUBESCENCE.FORM}{pubescence form}
-\gwastableone{PUBESCENCE.DENSITY}{pubescence density}
-\gwastablethree{SEED.COAT.LUSTER}{seed coat luster}
-\gwastableone{MATURITY.GROUP}{Maturity group}
-
 % Figures for flower color
 \manhattanplot{flower_color}{flower color}{\getvar{flower_color_nsamples}}
-\ldplot{flower_color}{flower color}{\getvar{flower_color_ldkmers}}
 \signalplot{flower_color_W1}{flower color}{\emph{W1}}%
@@
 \kmerplot{flower_color_W1}{\getvar{flower_color_W1_gene}}{\emph{W1}}{flower color}{a causal SV}
+\ldplot{flower_color}{flower color}{\getvar{flower_color_ldkmers}}
 
@@
 {Vertical dotted lines in panels (b) and (e) indicate the location of the causal indel at this locus.}
-\kmerplot{pubescence_color_all_T}{\getvar{pubescence_color_all_T_gene} gene}{\emph{T}}{pubescence color (first GWAS)}{a causal indel}
+\kmerplot{pubescence_color_all_T}{\getvar{pubescence_color_all_T_gene}}{\emph{T}}{pubescence color (first GWAS)}{a causal indel}
 \ldplot{pubescence_color_all}{pubescence color (first GWAS)}{\getvar{pubescence_color_all_ldkmers}}
@@
 {Vertical dotted lines in panels (b) and (e) indicate the location of the causal SNP at this locus.}
-\kmerplot{pubescence_color_nogray_Td}{\getvar{pubescence_color_nogray_Td_gene} gene}{\emph{Td}}{pubescence color (second GWAS)}{a causal SNP}
+\kmerplot{pubescence_color_nogray_Td}{\getvar{pubescence_color_nogray_Td_gene}}{\emph{Td}}{pubescence color (second GWAS)}{a causal SNP}
 
@@
 {Vertical dotted lines in panels (b) and (e) indicate the location of the causal SNP at this locus.}
-\kmerplot{seed_coat_color_greenyellow_G}{\getvar{seed_coat_color_greenyellow_G_gene} gene}{\emph{G}}{seed coat color (second GWAS)}{a causal SNP}
+\kmerplot{seed_coat_color_greenyellow_G}{\getvar{seed_coat_color_greenyellow_G_gene}}{\emph{G}}{seed coat color (second GWAS)}{a causal SNP}
 
@@
 {Vertical dotted lines in panels (b) and (e) indicate the location of a causal SNP at this locus.}
-\kmerplot{stem_termination_all_Dt1}{\getvar{stem_termination_all_Dt1_gene} gene}{\emph{Dt1}}{stem termination type (first GWAS)}{a causal SNP}
+\kmerplot{stem_termination_all_Dt1}{\getvar{stem_termination_all_Dt1_gene}}{\emph{Dt1}}{stem termination type (first GWAS)}{a causal SNP}
 
@@
 {Vertical dotted lines in panels (b) and (e) indicate the location of the causal indel at this locus.}
-\kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene} gene}{\emph{T}}{hilum color (first GWAS)}{a causal indel}
+\kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene}}{\emph{T}}{hilum color (first GWAS)}{a causal indel}
 \signalplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}%
 {Vertical dotted lines indicate the boundaries of the tandem duplication/inversion identified as the causal variant at this locus.}
-\geneplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}{causal tandem duplication/inversion region}{}
 
@@
 \geneplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}{\getvar{hilum_color_blackbrown_R_gene} gene}%
-{Vertical dotted lines indicate the location a SNP (panels b and e) and an indel (panel e) that are putatively causal at this locus.}
-\kmerplot{hilum_color_blackbrown_R}{\getvar{hilum_color_blackbrown_R_gene} gene}{\emph{R}}{hilum color (second GWAS)}{a SNP and an indel that are putatively causal}
+{Vertical dotted lines indicate the location an indel (panels b and e) and a SNP (panel e) that are putatively causal at this locus.}
+\kmerplot{hilum_color_blackbrown_R}{\getvar{hilum_color_blackbrown_R_gene}}{\emph{R}}{hilum color (second GWAS)}{a SNP and an indel that are putatively causal}
 
@@
 {Vertical dotted lines in panels (b) and (e) indicate the location of two highly significant non-synonymous SNPs at this gene.}
-\kmerplot{pubescence_form_all_Pa1}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}{\emph{Pa1}}{pubescence form (first GWAS)}{a non-synonymous SNP}
+\kmerplot{pubescence_form_all_Pa1}{\getvar{pubescence_form_all_Pa1_gene} candidate}{\emph{Pa1}}{pubescence form (first GWAS)}{a non-synonymous SNP}
 \signalplot{pubescence_form_all_Pa2}{pubescence form (first GWAS)}{\emph{Pa2}}{}
@@
 {Vertical dotted lines indicate the boundaries of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene associated with the locus.}
-\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene}%
-{The vertical dotted line in panel (e) indicates the location of the most significant \emph{k}-mer, which is associated with the causal CNV at this locus.}
+\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene.}%
+{The vertical dotted line in panel (e) indicates the location of the most significant \emph{k}-mer, which is associated with the causal CNV at this locus.
+The shaded gray rectangles show the boundaries of the causal CNV.}
 
@@
 {Vertical dotted lines indicate the boundaries of the causal CNV associated with the locus.}
-\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV}{}
+\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV}%
+{The shaded gray rectangles show the boundaries of the causal CNV.}
+
+% Figures for maturity group
+\manhattanplot{maturity_group}{maturity group}{\getvar{maturity_group_nsamples}}
 
@@
 {Vertical dotted lines indicate the location of the \getvar{corrected_dry_weight_cdwGm15_gene} gene suggested by de Ronne et al. (2022) as associated with the locus.}
-
-% Figures for maturity group
-\manhattanplot{maturity_group}{maturity group}{\getvar{maturity_group_nsamples}}
 

### 2022-10-26 additional_files/supplemental_file_2.R

@@ -4,7 +4,7 @@ library(parallel)
 # Reading the Bayer IDs and their corresponding SRR IDs
 # DEPENDENCY: utilities/srr_id_correspondence.txt
 bayer_ids <- read.table("utilities/srr_id_correspondence.txt", header = FALSE, stringsAsFactors = FALSE)
-colnames(bayer_ids) <- c("study_ID", "SRA_accessions")
+colnames(bayer_ids) <- c("study_ID", "SRA_runs")
 
 # DEPENDENCY: stats obtained from samtools stats for all samples
 # Now we read the metadata obtained from the samtools stats command to add the number of reads
@@ -28,7 +28,7 @@ average_cov <- sapply(bayer_ids$study_ID, function(x) {
 			      read.table(paste0("sv_genotyping/paragraph/manifest_files/", x, "_manifest.txt"),
 					 skip = 1)[[3]]})
 
-bayer_ids$mapping_depth <- average_cov[bayer_ids$study_ID]
+bayer_ids$average_mapping_depth <- average_cov[bayer_ids$study_ID]
 
 # Adding the results on the concordance rate between the samples and
 # DEPENDENCY: illumina_data/soysnp50k_genotyping/gtcheck_results.tsv

### 2022-10-26 figures/kmer_plot.R

@@ -177,7 +177,7 @@ grid::pushViewport(grid::viewport(layout.pos.row = 3))
 grid.text("(b)", x = -0.03, y = 1.05, just = c(0, 0))
 
 # Plot the haplotype sequences
-grid.haplotypes(hapdata = plotting_data, difflist = difflist, position = plotting_range)
+grid.haplotypes(hapdata = plotting_data, difflist = difflist, position = plotting_range, fontsize = if(trait == "flower_color") 6 else 8)
 grid::upViewport()
 
 

### 2022-10-26 reference_signals/format_bandillo2017_signals.R

@@ -131,7 +131,10 @@ bandillo_signals[bandillo_signals$locus == "Pa1", "gene_name_v4"] <- tolower("gl
 bandillo_signals[bandillo_signals$locus == "Ps", "gene_name_v4"] <- tolower("Glyma.12G187200")
 
 # Pd1 locus: Liu et al. (2020), DOI:10.1016/j.molp.2020.10.004
-bandillo_signals[bandillo_signals$locus == "Ps", "gene_name_v4"] <- tolower("Glyma.01G240100")
+bandillo_signals[bandillo_signals$locus == "Pd1", "gene_name_v4"] <- tolower("Glyma.01G240100")
+
+# P1 locus: Liu et al. (2020), DOI:10.1016/j.molp.2020.10.004
+bandillo_signals[bandillo_signals$locus == "P1", "gene_name_v4"] <- tolower("Glyma.09G278000")
 
 # B locus: Gijzen et al. (2006), DOI:10.1186/1471-2229-6-6
 # This is actually a cluster of HPS genes

### 2023-02-02 additional_files/additional_file_1.tex
Integrate contents from the supplemental data in the thesis into the supplemental file here

@@
 \usepackage{datatool} % allows fetching key-value pairs from a text file
+\usepackage{natbib}
 
@@
 \begin{cfigure}
-	\includegraphics[width = 6in]{#1_manhattan}
+	\includegraphics[width = 5.8in]{#1_manhattan}
 	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
@@
 \begin{cfigure}
-	\includegraphics[width = 6in]{#1_signal}
-	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zommed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
+	\includegraphics[width = 5.8in]{#1_signal}
+	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
 	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
@@
 \begin{cfigure}
-	\includegraphics[width = 6in]{#1_gene}
+	\includegraphics[width = 5.8in]{#1_gene}
 	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
@@
 	\begin{cfigure}
-		\includegraphics[width = 6.5in]{#1_kmers}
+		\includegraphics[width = 5.8in]{#1_kmers}
 		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
@@
 	\begin{cfigure}
-		\includegraphics[width = 6.5in]{#1_ld}
+		\includegraphics[width = 5.8in]{#1_ld}
 		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
@@
 \clearpage}
-
-% Allows underscores in captions but is not a great solution
-\catcode`_=12
 
@@
 
+% Detailed methods for some aspects of the paper
 \maketitle \thispagestyle{empty}
@@
 \listoftables
+\listoffigures
 
 \thispagestyle{empty}
-%\clearpage
 
-\vspace{4ex}
+\section{Detailed methods}
+\label{annexe-sv-gwas-detailed-methods}
 
-\listoffigures
+\subsection{Discovery of structural variation from Illumina data}
+\label{sv-gwas-sv-discovery-illumina}
 
-%\thispagestyle{empty}
+We used the Illumina WGS data of the 389 selected \emph{G. max} accessions as a
+first source of data for SV discovery using methods similar to those used by
+\cite{lemay2022}.  Briefly, we used four different SV discovery programs or
+pipelines:
+
+\begin{enumerate}
+
+\item We used AsmVar \citep[version of 2015-04-16, ][]{liu2015} to call SVs
+	from \emph{de novo} assemblies generated with SOAPdenovo2 v. 2.04
+		\citep{luo2012} and aligned to the reference genome using LAST
+		v. 1047 \citep{kielbasa2011}.
+\item We ran Manta v. 1.6.0 \citep{chen2016} in 78 randomly selected batches of
+	5 samples and 1 batch of 4 samples. We used the candidate SVs
+		identified from each of the batches for downstream analyses.
+\item We used smoove v. 0.2.4 \citep{pedersen2019} with the sequence of
+	commands \texttt{smoove call}, \texttt{smoove merge}, \texttt{smoove
+		genotype} and \texttt{smoove paste} to obtain a dataset of
+		candidate SVs. For three (3) of the 389 samples, we were unable
+		to obtain candidate variants from smoove due to a reproducible
+		cryptic error (segmentation fault) when running \texttt{smoove
+		call}. We nevertheless decided to use smoove given that this
+		represented a very small fraction of the total number of
+		samples and that other programs were able to successfully
+		discover SVs from those samples.
+\item We used SvABA v. 1.1.3 \citep{wala2018} to call SVs from each of the
+	samples separately. We classified the SVs called as breakends by SvABA
+		as either deletions, duplications or inversions, and converted
+		them to explicit sequence using our own R script. Variants
+		called as indels by SvABA were used as is for downstream
+		analyses.
+\end{enumerate}
+
+We filtered the SVs called by all tools to remove those smaller than 50 bp or
+larger than 500 kb in size, those located on unanchored scaffolds or organellar
+genomes, those classified as unresolved breakends (\texttt{SVTYPE=BND}), and
+any variant containing at least one N in either the reference or alternate
+allele sequence.  We converted all variants to a sequence-explicit format and
+normalized them using bcftools \texttt{norm} v. 1.10.2-105
+\citep{li2009-samtools}.
+
+We generated a single VCF file from the variants called using Illumina data by
+running SVmerge \citep{wong2010} in two steps. In the first step, we merged
+variants among the ones called by a given tool by running SVmerge using a
+sliding window of 5 bp to merge together very similar variants called by the
+same tool. In the second step, we used the output of the first merging steps
+for all four tools and merged similar variants across these four datasets using
+a sliding window of 15 bp in SVmerge.  We proceeded this way because using a
+sliding window of 15 bp on the outputs of all tools without pre-merging
+resulted in a computationally intractable number of pairwise comparisons.
+
+\subsection{Discovery of structural variation from Oxford Nanopore sequencing data}
+\label{sv-gwas-sv-discovery-ont}
+
+We used SVs discovered among 17 Canadian soybean cultivars sequenced using
+Oxford Nano\-pore Technologies (ONT) by \cite{lemay2022} as a second source of
+data for SV discovery.  Although that set of cultivars does not overlap the
+ones analyzed in this study, the high quality of the SVs that can be discovered
+from long-read sequencing data motivated the inclusion of this dataset among
+the SV candidates subjected to genotyping.  We merged the SVs detected in the
+long-read (ONT) data with the ones discovered in the short-read (Illumina) data
+using SVmerge and systematically favored SVs discovered by Illumina whenever
+SVs from the two sets were to be merged.  Our rationale for favoring variants
+discovered from the Illumina data was that whenever both technologies discover
+the same variants, Illumina data is likely to be more precise given its higher
+basecalling accuracy.
+
+\subsection{Discovery of structural variation from high-quality assemblies}
+\label{sv-gwas-sv-discovery-assembly}
+
+In addition to using the Illumina data from 389 accessions, we also used
+high-quality genome assemblies of 29 \emph{G. max} or \emph{Glycine soja}
+accessions and called SVs by comparing their sequence to that of Williams82. We
+retrieved 26 high-quality genomes assembled by \cite{liu2020-pangenome} as well
+as that of the cultivar Zhonghuang 13 (ZH13) from the Genome Warehouse
+repository (\url{https://ngdc.cncb.ac.cn/gwh/}).  We also retrieved the genomes
+of \emph{G. soja} accession W05 and \emph{G. max} cultivar Lee from SoyBase
+\citep{grant2010}.
+
+We used methods based on \cite{liu2020-pangenome} to call SVs based on those
+genome assemblies.  First, we used the nucmer program of the MUMmer suite of
+tools v. 4.0.0rc1 \citep{marcais2018} with option \texttt{-c 1000} to align
+each of the assemblies to the Williams82 assembly. Next, we used the
+delta-filter program of the MUMmer suite to keep only one-to-one alignments
+between any genome and the Williams82 reference using the \texttt{-1} option.
+We used these filtered alignments to call SVs using a version of the svmu
+program \citep{chakraborty2019} that we forked from the original repository
+(\url{https://github.com/malemay/svmu}) in order to improve execution time and
+memory usage.
+
+We used a set of tools that we assembled in an R package called svmutools
+\url{https://github.com/malemay/svmutools} to convert the svmu output to VCF
+format. We limited ourselves to processing deletions and insertions because of
+the complexity of resolving other types of variants from whole-genome
+assemblies.  Copy number variants and inversions called by svmu were therefore
+excluded from downstream analyses, although we did take inverted regions into
+account when determining the sequence of insertions in these regions to ensure
+their proper representation in the output VCF file. We also filtered the
+resulting set of variants to remove those smaller than 50 bp or larger than 500
+kb and those with any N in either the reference or alternate allele sequence.
+Finally, we normalized the VCF files using \texttt{bcftools norm} and merged
+them using SVmerge to obtain a single VCF file of variants discovered through
+comparison of high-quality genome assemblies.
+
+\subsection{Genotyping structural variation using Paragraph}
+\label{sv-gwas-sv-genotyping-paragraph}
+
+We used Paragraph v. 2.4a \citep{chen2019} as one of two SV genotyping programs
+to generate a set of SV genotypes for the population of 389 samples for use in
+GWAS analyses.  The set of SVs used as input for Paragraph was the result of
+merging the set of variants discovered from the Illumina and ONT data with
+those discovered from the high-quality assemblies and svmu using SVmerge.
+Variants in this candidate set were prepared for input to Paragraph by removing
+those closer than 150 bp (the maximum read length in our dataset) to any
+chromosome end and padding the variant representation using a script written by
+\cite{hickey2020} for that purpose.
+
+We used the recommendations by the authors of Paragraph for running their
+program at population scale by running Paragraph individually on each sample.
+We computed average sequencing depth for each sample from the output of the
+\texttt{samtools coverage} command and used 20 times that value for the
+\texttt{-M} parameter of Paragraph.  We ran Paragraph using the
+\texttt{multigrmpy} command and merged the resulting VCF files using
+\texttt{bcftools merge} in order to get a  single genotype file for the whole
+population. This VCF was further filtered by setting the genotype calls made
+from less than two reads to missing and by removing variants with minor allele
+frequency < 0.02 or a proportion of missing data > 0.5. The resulting dataset
+of 186,306 SVs was used for GWAS.
+
+\subsection{Genotyping structural variation using the vg toolkit}
+\label{sv-gwas-sv-genotyping-vg}
+
+We also used the vg toolkit v. 1.41.0 \citep{hickey2020} and its Giraffe
+aligner \citep{siren2021} for genotyping our set of candidate SVs on the whole
+population. The reason for also using vg is that its variation graph model allows
+the integration of SNPs and indels and makes for a more robust and integrated
+approach to SV genotyping, even though previous benchmarks comparing those two
+programs in soybean showed a reduced sensitivity of the vg toolkit in
+genotyping insertions \citep{lemay2022}.
+
+We constructed the variation graph using the \texttt{vg autoindex} command with
+the \texttt{-\-workflow giraffe} option. We used the Williams82 reference
+assembly version 4 as a basis for the graph, to which we added the set of
+candidate SVs described above and the set of 7M SNPs and indels called by
+Platypus.  We aligned the trimmed reads of the 741 SRA runs in FASTQ format to
+the variation graph using the \texttt{vg giraffe} command with default
+parameters. Reads that were still paired after adapter- and quality- trimming
+were aligned in paired-end mode, whereas the remaining reads were aligned in
+single-end mode. The resulting files in GAM format were used to compute read
+coverage on the graph using the \texttt{vg pack} command with options
+\texttt{-Q5 -s5}. GAM files belonging to the same accession but originating
+from different SRA runs were concatenated to be used as input for the
+\texttt{vg pack} command such that a single coverage file per accession was
+obtained as output of this step. The command \texttt{vg call} was then used to
+call genotypes based on read coverage using the parameters
+\texttt{-\-genotype-snarls -\-min-length 50} in order to call all variants
+$\geq$ 50 bp supported by the graph.
+
+We merged the genotypes called on all samples using \texttt{bcftools merge} to
+obtain a single VCF file for the population. Multiallelic records were split
+into biallelic records after merging. Genotype calls made with a genotype
+quality (GQ) value < 10 were set to missing, and variants were filtered for a
+fraction of missing genotypes $\leq$ 0.5 and a minor allele frequency $\geq$
+0.02. The resulting VCF file, which was used for GWAS, contained 51,960 SVs.
+
+\subsection{Comparison of WGS and SoySNP50K genotypes}
+\label{annexe-sv-gwas-soysnp50k}
+When we first analyzed the results from the GWAS analysis on all 389 samples,
+we found discrepancies between the haplotypes at some loci and the expected
+phenotype that should correspond to that haplotype for simple traits such as
+flower color and seed coat color. Upon further analysis, we found that a
+mismatch between the sequencing data and the identity of the accession was the
+most probable cause of these discrepancies. In order to filter out all samples
+for which this could be the case, we compared the genotypes of SNPs obtained
+from the SoySNP50K array \cite{song2013} to those obtained from the Illumina WGS
+data of the 385 samples that had a match in the GRIN database. Briefly, we
+retrieved SoySNP50K genotype calls from SoyBase \citep{grant2010} and extracted
+all SNPs with a MAF $\geq$ 0.1 in our population. We translated the positions
+of these SNPs from assembly version 2 to assembly version 4 of Williams by
+finding exact matches of 41-bp sequences surrounding the SNP positions. The
+32,852 SNPs for which unambiguous positions could be found on genome assembly
+version 4 were genotyped from the mapped WGS reads using \texttt{bcftools
+mpileup} and \texttt{bcftools call}. We then used \texttt{bcftools gtcheck} to
+compute the concordance of genotype calls between WGS samples and their
+expected match in the SoySNP50K genotype dataset.  We identified 24 samples
+with < 90\% concordance between the two genotype datasets and removed them from
+the dataset, in addition to another sample for which concordance was over 90\%
+but poor raw sequencing data quality (\%GC content) suggested contamination.
+Following these filtering steps, 363 samples remained for GWAS analyses. The
+concordance between WGS data and SoySNP50K genotypes is included in
+Additional file 2.
+
+% Detailed results that are not meant to fit the main paper
+\section{Detailed results}
+\label{annexe-sv-gwas-detailed-results}
+
+This section provides a more detailed account of some the the results obtained
+by the GWAS analyses on a per-trait basis. Aspects of the results that were
+already discussed in the main text are not repeated here; in such cases,
+readers are simply redirected to specific tables and figures containing
+detailed results from all four GWAS approaches. A summary of the signals found
+at all loci using all four methods is presented in Table \ref{signals_table}.
+Readers are invited to refer to this table as we go over the results obtained
+for each trait below.
+
+\subsection{Flower color}
+\label{sv-gwas-flower-color}
+
+We conducted a single GWAS analysis contrasting purple and white flower colors
+(Table \ref{FLOWER.COLOR_gwas_table}). Of the two classical loci reported by
+\cite{bandillo2017}, \textit{W1} was detected by all four GWAS approaches whereas
+\textit{L1} was detected by none of the approaches (Figure
+\ref{flower_color_manhattan}). These results are unsurprising as W1 is the
+major contributor to flower color in soybean, whereas the \textit{L1} locus
+(typically associated with pod color) was detected with a much less significant
+\textit{p}-value (see Table \ref{signals_table}) and only conjectured as
+associated with flower color by \cite{bandillo2017}. Detailed results for the
+\textit{W1} locus can be viewed in figures \ref{flower-color-main-figure},
+\ref{flower_color_W1_signal} and \ref{flower_color_W1_gene}.
+
+The \textit{k}-mer GWAS analysis also identified presumably spurious significant
+\textit{k}-mers scattered throughout the genome (Figure
+\ref{flower_color_manhattan}). An analysis of the pairwise LD between a subset
+of those \textit{k}-mers showed that all \textit{k}-mers formed a single LD block
+linked to the \textit{W1} locus (Figure \ref{flower_color_ld}), ruling out the
+presence of a second locus controlling that trait in this population.
+
+\subsection{Pubescence color}
+\label{sv-gwas-pubescence-color}
+
+We conducted two GWAS analyses on pubescence color. The first analysis
+contrasted gray, tawny and light tawny pubescence colors (Table
+\ref{PUBESCENCE.COLOR_gwas_table}) and detected a signal overlapping the
+\emph{T} locus using all four approaches (Figure
+\ref{pubescence_color_all_manhattan}). Detailed results at the \textit{T} locus
+van be viewed in figures \ref{pubescence_color_all_T_signal},
+\ref{pubescence_color_all_T_gene} and \ref{pubescence_color_all_T_kmers}. We
+observed that not all accessions with gray pubescence harbored the deleted
+nucleotide, which could indicate either that other causal variants resulting in
+a non-functional F3’H are present in this population, or that the causal
+variant lies elsewhere.
+
+This first GWAS analysis also revealed a second locus on chromosome Gm16
+(Figure \ref{pubescence_color_all_manhattan}). Upon closer analysis, this locus
+appeared to be in strong LD with \emph{k}-mers mapping to chromosome Gm01 and
+most importantly in moderate LD with the \emph{T} locus (Figure
+\ref{pubescence_color_all_ld}), which suggests that this does not represent an
+independent locus controlling pubescence color. This interpretation is also
+supported by the fact that the signal at that locus only spans about 10 kb on
+chromosome Gm16.
+
+In the presence of a functional allele at the \emph{T} locus, the genotype at
+the \emph{Td} locus influences the exact shade of pubescence color.  Therefore,
+we conducted a second GWAS analysis contrasting accessions with tawny and light
+tawny pubescence in order to ignore the epistatic effect of the \emph{T} locus
+and instead focus on the effect of the \emph{Td} locus on pubescence color
+(Table \ref{PUBESCENCE.COLOR_gwas_table}). Detailed results at this locus
+can be viewed in figures \ref{pubescence-color-main-figure}, \ref{pubescence_color_nogray_Td_signal} and
+\ref{pubescence_color_nogray_Td_gene}).
+
+\subsection{Seed coat color}
+\label{sv-gwas-seed-coat-color}
+
+We conducted two different GWAS analyses on seed coat color. The first GWAS
+included five different phenotypes observed in the dataset (Table
+\ref{SEED.COAT.COLOR_gwas_table}) and resulted in the detection of strong
+signals overlapping the \emph{I} locus by all four approaches (Figure
+\ref{seed_coat_color_all_manhattan}). All approaches but vg also detected
+signals at the \emph{G} locus. In addition, the \emph{k}-mer approach also
+detected weak signals overlapping the \emph{T} and \emph{O} loci. However, the
+\emph{T} and \emph{O} loci are likely chance associations given the low \textit{p}-value
+associated with the \emph{T} locus and the proximity of the \emph{O} locus to
+the \emph{I} locus. Detailed results of this first GWAS and the \textit{I}
+locus can be viewed in figures \ref{seed_coat_color_all_manhattan} and
+\ref{seed_coat_color_all_I_signal}.
+
+The second GWAS analysis on seed coat color focused on a subset of accessions
+with yellow or green seed coat only, as this difference is controlled by the
+\emph{G} locus (Table \ref{SEED.COAT.COLOR_gwas_table}). Detailed results of
+this analysis can be viewed in figures \ref{seed-coat-color-main-figure},
+\ref{seed_coat_color_greenyellow_manhattan},
+\ref{seed_coat_color_greenyellow_G_signal} and
+\ref{seed_coat_color_greenyellow_G_gene}.
+
+\subsection{Stem termination type}
+\label{sv-gwas-stem-termination}
+
+We conducted two GWAS analyses on stem termination type. The first GWAS
+included all phenotypes observed in the dataset (determinate, semi-determinate,
+and indeterminate; Table \ref{STEM.TERMINATION.TYPE_gwas_table}). The results
+of this analysis and the \textit{Dt1} locus can be viewed in figures \ref{stem_termination_all_manhattan}
+\ref{stem_termination_all_Dt1_signal}, \ref{stem_termination_all_Dt1_gene} and
+\ref{stem_termination_all_Dt1_kmers}.
+
+The second GWAS analysis that we conducted on this trait included only
+semi-determinate and indeterminate phenotypes in an attempt to detect signals
+at the \emph{Dt2} locus (Table \ref{STEM.TERMINATION.TYPE_gwas_table}).
+However, this was unsuccessful using all four GWAS
+approaches (Figure \ref{stem_termination_sn_manhattan}). Similarly, we failed to
+detect signals at the \textit{E3} locus, which is more commonly associated
+with maturity but was mentioned as associated with stem termination type by
+\cite{bandillo2017}.
+
+\subsection{Hilum color}
+\label{sv-gwas-hilum-color}
+
+We conducted three different GWAS analyses on hilum color. The first GWAS
+included seven colors (Table \ref{HILUM.COLOR_gwas_table}) and detected signals
+at the \emph{I}, \emph{T} and \emph{R} loci using all four approaches except vg
+at the \emph{R} locus (Figure \ref{hilum_color_all_manhattan}) while no signals
+were detected at the \emph{O} and \emph{W1} loci studied by
+\cite{bandillo2017}. The detailed results for the \textit{I} locus can be viewed
+in figure \ref{hilum_color_all_I_signal} whereas those for the \textit{T} locus
+can be viewed in figures \ref{hilum_color_all_T_signal}, \ref{hilum_color_all_T_gene}
+and \ref{hilum_color_all_T_kmers}.
+
+Although \cite{bandillo2017} listed the \textit{W1} locus (typically associated
+with flower color) as putatively associated with hilum color, we found no
+evidence of association in our results.
+
+A second GWAS analysis contrasting accessions with either black or brown hilum
+color was better suited at resolving the signal at the \emph{R} locus (Table
+\ref{HILUM.COLOR_gwas_table}). Accordingly, this GWAS analysis detected strong
+signals overlapping the \emph{R} locus and no other classical loci using all
+four approaches (Figure \ref{hilum_color_blackbrown_manhattan}).
+\cite{gillman2011} identified the R2R3 MYB transcription factor Glyma.09g235100
+as the molecular basis for the \emph{R} locus and documented four different
+loss-of-function alleles linked to brown hilum color. Although our analyses did
+detect strong signals in the vicinity of this gene, the most associated markers
+were not located in the body of the gene (Figure
+\ref{hilum_color_blackbrown_R_signal}). Interestingly, the most significantly
+associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
+another gene (Glyma.09g234900) annotated as a MYB transcription factor and
+putatively involved in the regulation of anthocyanin biosynthesis. This gene
+as well as Glyma.09g235100 and two others are part of a tandem array of four
+R2R3 MYB genes considered candidates by \cite{gillman2011}, and of which
+Glyma.09g235100 was the only one expressed in the seed coat. Given the
+convincing evidence provided by \cite{gillman2011}, it is most likely that
+Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
+our failure to identify the most significantly associated \emph{k}-mers as
+causal variants is due to the coexistence of several variants that can cause
+loss of function at the \emph{R} locus. Indeed, we observed at least two causal
+variants (a 1-bp deletion and a SNP) previously documented by
+\cite{gillman2011} in our dataset (Figure \ref{hilum_color_blackbrown_R_gene},
+\ref{hilum_color_blackbrown_R_kmers}), although they were not among the most
+significantly associated markers or \emph{k}-mers.
+
+We conducted a third GWAS focusing on the brown and reddish brown hilum color
+phenotypes in order to detect a signal at the \emph{O} locus, however this GWAS
+did not detect associations at any of the classical loci (Figure
+\ref{hilum_color_rbr_manhattan}).
+
+\subsection{Pod color}
+\label{sv-gwas-pod-color}
+
+We conducted two GWAS analyses on pod color. The first analysis included four
+different colors (Table \ref{POD.COLOR_gwas_table}) and resulted in the
+detection of signals overlapping both the \emph{L1} and \emph{L2} loci with all
+four methods (Figure \ref{pod_color_all_manhattan}). Detailed results at the
+\textit{L1} locus can be viewed in figure \ref{pod_color_all_L1_signal},
+whereas detailed results at the \textit{L2} locus can be viewed in figure
+\ref{pod_color_all_L2_signal}.
+
+The second GWAS contrasted black and brown pods with the objective of detecting
+signals that provided better resolution at the \emph{L1} locus (Table
+\ref{POD.COLOR_gwas_table}).  Accordingly, all four methods detected a signal
+at the \emph{L1} locus but not at the \emph{L2} locus using this GWAS (Figure
+\ref{pod_color_blbr_manhattan}).  Results at the \emph{L1} locus were similar
+to those observed when using all observed phenotypes, but the \emph{k}-mer
+analysis revealed a previously undocumented signal on chromosome Gm15 which was
+not in LD with any other signal (Figures \ref{pod_color_blbr_manhattan},
+\ref{pod_color_blbr_ld}). The observed signal at this locus overlaps a total of
+27 genes (Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those was
+annotated as having an obvious role in controlling pod color.
+
+\subsection{Pubescence form}
+\label{sv-gwas-pubesence-form}
+
+We conducted two different GWAS on pubescence form. The first GWAS used three
+phenotypes (erect, semi-appressed and appressed; Table
+\ref{PUBESCENCE.FORM_gwas_table}) and detected signals overlapping the
+\emph{Pa1} locus using all four approaches as well as signals overlapping the
+\emph{Pa2} locus using Platypus and the \emph{k}-mer approaches (Figure
+\ref{pubescence_form_all_manhattan}). Detailed results at the \textit{Pa1} locus
+can be viewed in figure \ref{pubescence_form_all_Pa1_signal},
+\ref{pubescence_form_all_Pa1_gene} and \ref{pubescence_form_all_Pa1_kmers},
+whereas detailed results at the \textit{Pa2} locus can be viewed in figure
+\ref{pubescence_form_all_Pa2_signal}.
+
+Signals were also observed on chromosomes Gm04 and Gm15 by the Platypus and
+\emph{k}-mer analyses (Figure \ref{pubescence_form_all_manhattan}), however
+upon further analysis of LD patterns they appeared to be linked to the
+\emph{Pa1} locus (Figure \ref{pubescence_form_all_ld}).
+
+A second GWAS analysis focused only on appressed and semi-appressed phenotypes
+with the objective of enhancing the signal at the \emph{Pa2} locus. However,
+this analysis only detected weak signals at the \emph{Pa1} locus (Figure
+\ref{pubescence_form_noerect_manhattan}) and its results were therefore not
+analyzed further.
+
+\subsection{Pubescence density}
+\label{sv-gwas-pubescence-density}
+
+We conducted a single GWAS analysis on pubescence density (Table
+\ref{PUBESCENCE.DENSITY_gwas_table}), which detected signals at the \emph{Ps}
+locus but not the \emph{Pd1} and \emph{P1} loci using all four approaches
+(Figure \ref{pubescence_density_manhattan}). \cite{liu2020ps} linked the gene
+Glyma.12g187200 to the \emph{Ps} locus and identified copy number variation
+(CNV) of a 25.6-kb sequence including this gene as the causal variant, with
+higher copy number resulting in decreased pubescence. Through a segmentation
+analysis of the sequencing depth in accessions with increased copy number, we
+identified the boundaries of the region varying in copy number as 36,254,989 -
+36,297,368 on chromosome Gm12 of genome assembly version 4 of Williams82, for a
+total of about 42 kb. Detailed results at this locus can be viewed in figures
+\ref{pubescence_density_Ps_signal}, \ref{pubescence_density_Ps_gene}, and
+\ref{pubescence-density-main-figure}.
+
+\subsection{Seed coat luster}
+\label{sv-gwas-seed-coat-luster}
+
+We performed three different GWAS analyses on seed coat luster (Table
+\ref{SEED.COAT.LUSTER_gwas_table}).  However, all three yielded similar results
+(Figure \ref{seed_coat_luster_all_manhattan},
+\ref{seed_coat_luster_nointermediate_manhattan},
+\ref{seed_coat_luster_dullshiny_manhattan}) so we chose to focus on the
+analysis contrasting the dull and shiny phenotypes as it yielded the most
+significant signals and also provided the most consistent results in the
+analysis previously done by \cite{bandillo2017}. This analysis detected
+signals at the \emph{B} locus as the sole classical locus using all four
+methods (Figure \ref{seed_coat_luster_dullshiny_manhattan}). Apparent signals
+on chromosomes Gm20 and Gm09 (which may correspond to the B? locus reported by
+\cite{bandillo2017}) were also detected, but analysis of the LD patterns
+between \emph{k}-mers suggested that these signals were most likely in LD with
+the \emph{B} locus (Figure \ref{seed_coat_luster_dullshiny_ld}). All methods
+also detected signals overlapping the \emph{Hps} locus as defined by
+\cite{bandillo2017}, however results shown in \cite{gijzen2003} and our results
+suggest that \emph{Hps} and \emph{B} are the same locus. Using a segmentation analysis
+similar to the one used for the \emph{Ps} locus, we identified a 31-kb region
+spanning positions 9,386,109-9,417,431 on chromosome Gm15 that exhibited
+variation in copy number across accessions and overlapped four genes annotated
+as containing HPS domains (Figures \ref{seed_coat_luster_dullshiny_B_signal},
+\ref{seed_coat_luster_dullshiny_B_gene}). The top 5\% / 1\% significant
+associations of all approaches other than vg overlapped this CNV region and the
+most significantly associated \emph{k}-mer was located within the CNV region
+(Figures \ref{seed_coat_luster_dullshiny_B_signal},
+\ref{seed_coat_luster_dullshiny_B_gene}), however we were not able to link any
+of the variants found to the presence of copy number variation.
+
+% Seed coat luster - I locus
+
+\cite{bandillo2017} mentioned the \textit{I} locus as possibly associated
+with seed coat luster, however we did not find evidence for such an association
+in our dataset.
+
+\subsection{Maturity group}
+\label{sv-gwas-maturity-group}
+
+A GWAS conducted on maturity group did not yield any significant signals
+overlapping the classical E1-E4 loci using any of the four GWAS approaches,
+although signals at the E1 locus did approach significance in the Platypus
+analysis (Table \ref{MATURITY.GROUP_gwas_table}, Figure \ref{maturity_group_manhattan}).
+
+\subsection{Resistance to \emph{Phytophthora sojae}}
+\label{sv-gwas-resistance}
+
+We conducted a single GWAS analysis on horizontal resistance to
+\textit{Phytophthora sojae} using corrected dry weight (CDW) as a quantitative
+measure of resistance. Detailed results can be viewed in figures
+\ref{corrected_dry_weight_manhattan} and
+\ref{corrected_dry_weight_cdwGm15_signal}.
+
+\subsection{Seed oil content}
+\label{sv-gwas-seed-oil-content}
+
+We conducted a GWAS analysis on the seed oil content data obtained from the
+GRIN database. While several seed oil content quantitative trait loci (QTL)
+have been reported due to high interest in this trait for soybean breeding
+\citep{chaudhary2015}, we focused on two major QTL on chromosomes Gm15 and Gm20
+that have often been reported in \emph{G. max}, and notably observed by
+\cite{bandillo2015} on a collection of USDA lines using SoySNP50K chip
+genotypes and GRIN phenotypic data. We identified a signal located on
+chromosome Gm15 using the \emph{k}-mer approach only (Figure
+\ref{oil_manhattan}). Although this signal did not overlap with the coordinates
+found by \cite{bandillo2017}, it was located only 2 kb away from it and 30 kb
+away from Glyma.15g049200, the sugar transporter gene identified by
+\cite{zhang2020} as associated with that locus. Therefore, the significantly
+associated \emph{k}-mers were not linked to the causal variation at that locus.
+
+\subsection{Seed protein content}
+\label{sv-gwas-seed-protein-content}
+
+Similarly to seed oil content, we conducted a GWAS on seed protein content
+using data obtained from the GRIN database. As is generally the case in
+soybean given the negative correlation between seed oil and protein content,
+the two major QTL considered above and reported by \cite{bandillo2015} for seed
+oil content are also major QTL for seed protein content. We identified a signal
+at the locus located on chromosome Gm20 only, and only using the genotypes
+called by vg (Figure \ref{protein_manhattan}). However, the SV discovered as
+significantly associated was located about 700 kb from the gene cloned by
+\cite{fliege2022} for that locus (Glyma.20g085100) and therefore is unlikely to
+represent the causal variation at that locus.
+
+\clearpage
+
+% Supplemental tables
+\section{Supplemental tables}
+\label{annexe-sv-gwas-tables}
 
@@
 
+% Supplemental figures
+\section{Supplemental figures}
+\label{annexe-sv-gwas-figures}
+
+% Figures mentioned in the main text
 % Figures for flower color
@@
 {Vertical dotted lines indicate the location of the \getvar{flower_color_W1_gene} gene associated with the locus.}
-
 \geneplot{flower_color_W1}{flower color}{\emph{W1}}{\getvar{flower_color_W1_gene} gene}%
 {Vertical dotted lines in panels (c), (d) and (e) indicate the location of the causal SV at this locus.}
-\kmerplot{flower_color_W1}{\getvar{flower_color_W1_gene}}{\emph{W1}}{flower color}{a causal SV}
-\ldplot{flower_color}{flower color}{\getvar{flower_color_ldkmers}}
+%\kmerplot{flower_color_W1}
 
@@
 \kmerplot{pubescence_color_all_T}{\getvar{pubescence_color_all_T_gene}}{\emph{T}}{pubescence color (first GWAS)}{a causal indel}
-\ldplot{pubescence_color_all}{pubescence color (first GWAS)}{\getvar{pubescence_color_all_ldkmers}}
 
@@
 {Vertical dotted lines in panels (b) and (e) indicate the location of the causal SNP at this locus.}
-\kmerplot{pubescence_color_nogray_Td}{\getvar{pubescence_color_nogray_Td_gene}}{\emph{Td}}{pubescence color (second GWAS)}{a causal SNP}
+%\kmerplot{pubescence_color_nogray_Td}
 
@@
 {Vertical dotted lines in panels (b) and (e) indicate the location of the causal SNP at this locus.}
-\kmerplot{seed_coat_color_greenyellow_G}{\getvar{seed_coat_color_greenyellow_G_gene}}{\emph{G}}{seed coat color (second GWAS)}{a causal SNP}
+%\kmerplot{seed_coat_color_greenyellow_G}
 
@@
 \kmerplot{stem_termination_all_Dt1}{\getvar{stem_termination_all_Dt1_gene}}{\emph{Dt1}}{stem termination type (first GWAS)}{a causal SNP}
-
-\manhattanplot{stem_termination_sn}{stem termination type (second GWAS)}{\getvar{stem_termination_sn_nsamples}}
-\ldplot{stem_termination_sn}{stem termination type (second GWAS)}{\getvar{stem_termination_sn_ldkmers}}
 
@@
 
-\manhattanplot{hilum_color_rbr}{hilum color (third GWAS)}{\getvar{hilum_color_rbr_nsamples}}
+% Figures for pubescence density
+\manhattanplot{pubescence_density}{pubescence density}{\getvar{pubescence_density_nsamples}}
+
+% Figures for seed coat luster
+\manhattanplot{seed_coat_luster_dullshiny}{seed coat luster (third GWAS)}{\getvar{seed_coat_luster_dullshiny_nsamples}}
+\signalplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}%
+{Vertical dotted lines indicate the boundaries of the causal CNV associated with the locus.}
+
+% Figures for maturity group
+\manhattanplot{maturity_group}{maturity group}{\getvar{maturity_group_nsamples}}
+
+% Figures for oil
+\manhattanplot{oil}{seed oil content}{\getvar{oil_nsamples}}
+
+% Figures for protein
+\manhattanplot{protein}{seed protein content}{\getvar{protein_nsamples}}
 
@@
 \signalplot{pod_color_all_L2}{pod color (first GWAS)}{\emph{L2}}{}
-
-\manhattanplot{pod_color_blbr}{pod color (second GWAS)}{\getvar{pod_color_blbr_nsamples}}
-\ldplot{pod_color_blbr}{pod color (second GWAS)}{\getvar{pod_color_blbr_ldkmers}}
-\signalplot{pod_color_blbr_pdcGm15}{pod color (second GWAS)}{newly suggested \emph{pdcGm15}}{}
 
@@
 \signalplot{pubescence_form_all_Pa2}{pubescence form (first GWAS)}{\emph{Pa2}}{}
+
+% Figures for corrected dry weight
+\manhattanplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae})}{\getvar{corrected_dry_weight_nsamples}}
+\signalplot{corrected_dry_weight_cdwGm15}{corrected dry weight (resistance to \emph{P. sojae})}{\emph{cdwGm15}}%
+{Vertical dotted lines indicate the location of the \getvar{corrected_dry_weight_cdwGm15_gene} gene suggested by de Ronne et al. (2022) as associated with the locus.}
+%\ldplot{corrected_dry_weight}
+
+% Figures for novel loci
+\manhattanplot{pod_color_blbr}{pod color (second GWAS)}{\getvar{pod_color_blbr_nsamples}}
+\ldplot{pod_color_blbr}{pod color (second GWAS)}{\getvar{pod_color_blbr_ldkmers}}
+
+% Figures mentioned only in supplemental data
+% Flower color
+\ldplot{flower_color}{flower color}{\getvar{flower_color_ldkmers}}
+
+% Pubescence color
+\ldplot{pubescence_color_all}{pubescence color (first GWAS)}{\getvar{pubescence_color_all_ldkmers}}
+
+% Stem termination
+\manhattanplot{stem_termination_sn}{stem termination type (second GWAS)}{\getvar{stem_termination_sn_nsamples}}
+%\ldplot{stem_termination_sn}
+
+% Hilum color
+\manhattanplot{hilum_color_rbr}{hilum color (third GWAS)}{\getvar{hilum_color_rbr_nsamples}}
+
+% Pod color
+\signalplot{pod_color_blbr_pdcGm15}{pod color (second GWAS)}{newly suggested \emph{pdcGm15}}{}
+
+% Pubescence form
 \ldplot{pubescence_form_all}{pubescence form (first GWAS)}{\getvar{pubescence_form_all_ldkmers}}
-
 \manhattanplot{pubescence_form_noerect}{pubescence form (second GWAS)}{\getvar{pubescence_form_noerect_nsamples}}
 
-% Figures for pubescence density
-\manhattanplot{pubescence_density}{pubescence density}{\getvar{pubescence_density_nsamples}}
+% Pubescence density
 \signalplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}%
@@
 
-% Figures for seed coat luster
+% Seed coat luster
 \manhattanplot{seed_coat_luster_all}{seed coat luster (first GWAS)}{\getvar{seed_coat_luster_all_nsamples}}
-
 \manhattanplot{seed_coat_luster_nointermediate}{seed coat luster (second GWAS)}{\getvar{seed_coat_luster_nointermediate_nsamples}}
 
-\manhattanplot{seed_coat_luster_dullshiny}{seed coat luster (third GWAS)}{\getvar{seed_coat_luster_dullshiny_nsamples}}
 \ldplot{seed_coat_luster_dullshiny}{seed coat luster (third GWAS)}{\getvar{seed_coat_luster_dullshiny_ldkmers}}
-\signalplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}%
-{Vertical dotted lines indicate the boundaries of the causal CNV associated with the locus.}
+
 \geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV}%
@@
 
-% Figures for maturity group
-\manhattanplot{maturity_group}{maturity group}{\getvar{maturity_group_nsamples}}
-
-% Figures for corrected dry weight
-\manhattanplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae})}{\getvar{corrected_dry_weight_nsamples}}
-\ldplot{corrected_dry_weight}{corrected dry weight (resistance to \emph{P. sojae})}{\getvar{corrected_dry_weight_ldkmers}}
-\signalplot{corrected_dry_weight_cdwGm15}{corrected dry weight (resistance to \emph{P. sojae})}{\emph{cdwGm15}}%
-{Vertical dotted lines indicate the location of the \getvar{corrected_dry_weight_cdwGm15_gene} gene suggested by de Ronne et al. (2022) as associated with the locus.}
-
-% Figures for oil
-\manhattanplot{oil}{seed oil content}{\getvar{oil_nsamples}}
-
-% Figures for protein
-\manhattanplot{protein}{seed protein content}{\getvar{protein_nsamples}}
+% Bibliography
+\bibliographystyle{genome_research}              % style de la bibliographie
+\bibliography{references}                   % production de la bibliographie
 

### 2023-02-06 additional_files/additional_file_1.tex
Add main text contents to the repository; manuscript.tex file includes both manuscript and supplemental data into one pdf

@@
-\documentclass[12pt]{article}
-
-\usepackage[utf8]{inputenc} % declare encoding as utf8
-\usepackage{graphicx} % to enable \includegraphics
-\usepackage{flafter} % to make sure floats appear after their position in text
-\usepackage{pdflscape} % allow some pages to be in landscape orientation
-\usepackage{layout} % for debugging purposes, looking at page layout
-\usepackage[margin=1in]{geometry} % leaves more space for text body
-\usepackage[hidelinks]{hyperref} % hyperlinks from list of figures/tables
-\usepackage[labelfont=bf]{caption} % format captions and make links go to figure instead of caption
-\usepackage{csvsimple} % read and display csv files as tables
-\usepackage{booktabs} % prettier tables
-\usepackage{threeparttable} % easy addition of footnotes below tables
-\usepackage{textcomp} % makes the \textdegree command available
-\usepackage{tocloft} % formatting the lists of tables and figures
-\usepackage{xcolor} % enable text coloring
-\usepackage{calc} % makes \widthof available to set the width of the last column
-\usepackage{array} % using command \arraybackslash to restore the definition of \\ after \centering
-\usepackage{longtable} % allow tables to span more than one page
-\usepackage{threeparttablex} % threeparttable for longtable
-\usepackage{capt-of} % put caption on next page
-\usepackage{listings}
-\usepackage{datatool} % allows fetching key-value pairs from a text file
-\usepackage{natbib}
-
-% Setting the separator to use with datatool
-\DTLsetseparator{ = }
-
-% Formatting the table of contents/subsection
-\renewcommand{\contentsname}{Supplemental methods}
-
-% Formatting the list of tables
-\renewcommand{\listtablename}{Supplemental tables} % Write "Supplemental tables" instead of "List of tables"
-\renewcommand\cfttabpresnum{Table } % Write "Table" before table number in list of tables
-\setlength{\cfttabnumwidth}{6em} % Spacing before captions
-\setlength{\cftbeforefigskip}{1ex} % Vertical spacing between entries
-
-% Formatting the list of figures
-\renewcommand{\listfigurename}{Supplemental figures} % Write "Supplemental figures" instead of "List of figures"
-\renewcommand\cftfigpresnum{Figure } % Write "Figure" before fig number in list of figures
-\setlength{\cftfignumwidth}{6em} % Spacing before captions
-\setlength{\cftbeforetabskip}{1ex} % Vertical spacing between entries
-
-\graphicspath{ {../figures/} } % tell Latex to look for figures in figures/
-
-% Format figure and table numbers to follow S1, S2, ...
-\renewcommand\thefigure{S\arabic{figure}} 
-\renewcommand\thetable{S\arabic{table}}
-
-% Defining new environments as I will always want the figures centered
-\newenvironment{cfigure}
-	{\begin{figure} \centering}
-	{\end{figure}}
-
-\newenvironment{lsfigure}
-	{\begin{landscape} \begin{figure} \centering}
-	{\end{figure} \end{landscape}}
-
-\setcounter{totalnumber}{1} % Allow a maximum of one float (table/figure) per page
-
-\setlength{\parskip}{2ex}
-
-% Defining a command to generate a table of phenotypic data for a single gwas
-\newcommand{\gwastableone}[2]{
-\begin{table}
-	\caption[Phenotypic data used for the GWAS analysis of #2]%
-	{Phenotype frequency and numerical coding of phenotypic data used for the GWAS analysis of #2.
-	``NA'' indicates a missing value in the source data.
-	Dashes indicate that this phenotype was not used in the GWAS analysis.}
-	\label{#1_gwas_table}
-
-	\centering
-	\begin{threeparttable}\small
-
-		\csvreader[head to column names,
-		/csv/separator=semicolon,
-		tabular = llc,
-		table head = \toprule Phenotype & Frequency & GWAS \\ \midrule,
-		table foot = \bottomrule]%
-		{../tables/#1_gwas_table.csv}%
-		{}%
-		{\Value & \Frequency & \GWASa}
-
-	\end{threeparttable}
-
-\end{table}
-\clearpage
-}
-
-% Defining a command to generate a table of phenotypic data for two gwas
-\newcommand{\gwastabletwo}[2]{
-\begin{table}
-	\caption[Phenotypic data used for the GWAS analyses of #2]%
-	{Phenotype frequency and numerical coding of phenotypic data used in two GWAS analyses of #2.
-	``NA'' indicates a missing value in the source data.
-	Dashes indicate that this phenotype was not used in the GWAS analysis.}
-	\label{#1_gwas_table}
-
-	\centering
-	\begin{threeparttable}\small
-
-		\csvreader[head to column names,
-		/csv/separator=semicolon,
-		tabular = llcc,
-		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2\\ \midrule,
-		table foot = \bottomrule]%
-		{../tables/#1_gwas_table.csv}%
-		{}%
-		{\Value & \Frequency & \GWASa & \GWASb}
-
-	\end{threeparttable}
-
-\end{table}
-\clearpage
-}
-
-% Defining a command to generate a table of phenotypic data for three gwas
-\newcommand{\gwastablethree}[2]{
-\begin{table}
-	\caption[Phenotypic data used for the GWAS analyses of #2]%
-	{Phenotype frequency and numerical coding of phenotypic data used in three GWAS analyses of #2.
-	``NA'' indicates a missing value in the source data.
-	Dashes indicate that this phenotype was not used in the GWAS analysis.}
-	\label{#1_gwas_table}
-
-	\centering
-	\begin{threeparttable}\small
-
-		\csvreader[head to column names,
-		/csv/separator=semicolon,
-		tabular = llccc,
-		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2 & GWAS \# 3\\ \midrule,
-		table foot = \bottomrule]%
-		{../tables/#1_gwas_table.csv}%
-		{}%
-		{\Value & \Frequency & \GWASa & \GWASb & \GWASc}
-
-	\end{threeparttable}
-
-\end{table}
-\clearpage
-}
-
-% Defining a command for manhattan plot figures
-\newcommand{\manhattanplot}[3]{
-\begin{cfigure}
-	\includegraphics[width = 5.8in]{#1_manhattan}
-	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
-	(a) Platypus (SNPs and indels), (b) vg (SVs), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
-	The x-axis shows the position along the reference assembly version 4 of Williams82.
-	Each point shows the $-\log_{10} (p)$ associated with a particular marker or \emph{k}-mer.
-	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
-	Vertical dotted lines indicate the position of signals associated with the trait.
-	Documented loci are colored according to whether they were found (green) or not (red) by a particular approach.
-	The ``Gm'' prefix has been left out of chromosome names for simplicity.
-	}
-	\label{#1_manhattan}
-\end{cfigure}
-
-\clearpage}
-
-% Defining a new command for signal plot figures
-\newcommand{\signalplot}[4]{
-\begin{cfigure}
-	\includegraphics[width = 5.8in]{#1_signal}
-	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
-	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
-	Panel (a) shows gene models over the genomic interval.
-	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
-	#4
-	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph and vg) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
-	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
-	\emph{p}-values were computed after signal discovery.
-	}
-	\label{#1_signal}
-\end{cfigure}
-
-\clearpage}
-
-% Defining a new command for gene plot figures
-\newcommand{\geneplot}[5]{
-\begin{cfigure}
-	\includegraphics[width = 5.8in]{#1_gene}
-	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
-	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using four genotype datasets:
-	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
-	#5
-	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
-	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
-	\emph{p}-values were computed after signal discovery.
-	Panel (a) shows gene models over the plotting interval.
-	Exons are represented by rectangles whereas introns are represented by horizontal lines.
-	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
-	}
-	\label{#1_gene}
-\end{cfigure}
-
-\clearpage}
-
-% Defining a new command for k-mer plot figures
-\newcommand{\kmerplot}[5]{
-	\begin{cfigure}
-		\includegraphics[width = 5.8in]{#1_kmers}
-		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
-		{Identification of #5 underlying significant \emph{k}-mers at the #2 gene associated with the #3 locus for #4.
-		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
-		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
-		The red rectangle identifies the region that is zoomed-in in panel (b).
-		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
-		Individual nucleotides are colored according to the $-\log_{10}(p)$ of the most significant \emph{k}-mer overlapping them.
-		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
-		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes correspond to those shown in panel (b).}
-		\label{#1_kmers}
-	\end{cfigure}
-
-\clearpage}
-
-% Defining a new command for LD plot figures
-\newcommand{\ldplot}[3]{
-	\begin{cfigure}
-		\includegraphics[width = 5.8in]{#1_ld}
-		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
-		\emph{k}-mers are sorted along the y-axis according to their putative position along the reference assembly version 4 of Williams82,
-		as identified by ``Gm'' chromosome labels.
-		Sequences that lack a ``Gm'' label (top of the y-axis) represent unanchored scaffolds.
-		\emph{k}-mers are represented in the same order along the x- and y-axis.
-		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p)$ of each \emph{k}-mer.}
-		\label{#1_ld}
-	\end{cfigure}
-
-\clearpage}
-
-% Information for the title page
-\title{Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''}
-\author{Marc-André Lemay \and Jonas A. Sibbesen \and Maxime de Ronne \and Richard Bélanger \and François Belzile}
-\date{}
-
-% END OF PREAMBLE
-
-\begin{document}
+\chapter{Supplemental data}
+\label{supdata}
 
@@
 % Detailed methods for some aspects of the paper
-\maketitle \thispagestyle{empty}
-
-\listoftables
-\listoffigures
-
-\thispagestyle{empty}
-
 \section{Detailed methods}
@@
 \bibliography{references}                   % production de la bibliographie
-
-\end{document}

### 2023-02-06 additional_files/main_text.tex
Add main text contents to the repository; manuscript.tex file includes both manuscript and supplemental data into one pdf

\chapter{\textit{k}-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean}
\label{kmer-gwas}

Marc-André Lemay\textsuperscript{1,2,3},
Jonas A. Sibbesen\textsuperscript{4},
Maxime de Ronne\textsuperscript{1,2,3},
Richard Bélanger\textsuperscript{1,2,3},
François Belzile\textsuperscript{1,2,3}

\textsuperscript{1} Département de phytologie, Université Laval, Québec, QC, Canada

\textsuperscript{2} Institut de biologie intégrative et des systèmes, Université Laval, Québec, QC, Canada

\textsuperscript{3} Centre de recherche et d’innovation sur les végétaux, Université Laval, Québec, QC, Canada

\textsuperscript{4} Center for Health Data Science, University of Copenhagen, Denmark

\clearpage

\section{Abstract}

Genome-wide association studies (GWAS) are powerful statistical methods
that detect associations between genotype and phenotype at genome scale. Despite
their power, GWAS often fail to pinpoint the causal variant or the gene
controlling a trait at a given locus.  Assessing genetic variants other
than single-nucleotide polymorphisms (SNPs) may alleviate this problem
by including functionally relevant structural variants (SVs) that are
typically omitted from GWAS.  In this study, we assessed the potential
of SV- and \emph{k}-mer-based GWAS by applying these methods to 13
traits on which SNP-based GWAS had been performed previously and for
which the causal genes and variants are known in some cases. We found
that \emph{k}-mer-based GWAS had an impressive ability to pinpoint
causal variants or candidate genes based on the most significantly
associated \emph{k}-mers. The ability of SNP/indel-based and SV-based
GWAS to similarly pinpoint known variants or genes was almost
systematically reduced compared to \emph{k}-mers.  We suggest that
SV-based GWAS may not generally be worth the investment in time and
computational resources required to discover and genotype such variants
given the results obtained with \emph{k}-mer-based GWAS, which requires
comparatively lower resources.  Moreover, analyses based on
\emph{k}-mer GWAS identified promising candidate genes for loci related
to pod color, pubescence form, and resistance to the oomycete
\textit{Phytophthora sojae}.  Despite these encouraging results,
significant challenges remain regarding the downstream analysis of
\emph{k}-mer GWAS, notably in linking significant \emph{k}-mers to
sequence variation and ensuring that all variation types are adequately
detected from \emph{k}-mer occurrence patterns.  Our results suggest
that coupling \emph{k}-mer- and SNP/indel-based GWAS may represent a
powerful approach for discovering candidate genes in crop species.

\section{Introduction}
\label{sv-gwas-introduction}

Genome-wide association studies (GWAS) are analytical approaches that detect
statistical associations between phenotypic observations for a trait of
interest and the genotypes of markers or variants distributed throughout the
genome \citep{tam2019}.  Such studies have been used for over 15 years
\citep{visscher2012} and have enabled significant advances in our understanding
of the genetic architecture of traits involved in human health
\citep[e.g.][]{wellcome2007gwas} and crop \citep{tibbs2021} or animal
\citep{ma2019} production, among others.

Despite their power and convenience for finding significantly associated loci,
GWAS analyses often fail to pinpoint the gene(s) associated with a trait and
the causal variant(s) involved \citep{tam2019}.  One possible reason for this
failure is that most GWAS do not include the causal variant itself among the
genotypes assessed \citep{tibbs2021}. This could be either because the analysis
is restricted to a particular type of variant, such as single-nucleotide
polymorphisms (SNPs), or because the set of markers/variants genotyped in the
association panel is limited by the technology being used
\citep[e.g.][]{bandillo2015, sonah2015}.  Despite these limitations, variants
located near the gene controlling the trait should appear statistically
associated with the phenotype through linkage disequilibrium (LD) with the
causal variant \citep{korte2013}.  A typical workflow then involves the
identification of candidate genes within the haplotype blocks defined by the
statistically associated markers and follow-up functional analyses to confirm
that a given gene controls the trait \citep[e.g.][]{wang2018, liu2020ps}.
Other limitations that may preclude the identification of causal variants from
the output of GWAS include incomplete gene annotation, genetic divergence
between the accessions and the reference genome used, or the existence of
several different causal variants at a locus.

One way of overcoming those limitations would be to broaden the set of genetic
variants used in GWAS by also including structural variants (SVs) such as
deletions, insertions, inversions, or any other type of rearrangement that
exceeds a certain size \cite[typically 50 bp; ][]{ho2020}.  SVs are known to
have large phenotypic impacts by disrupting coding or regulatory sequences
\citep{marroni2014} and have been identified as causal variants in traits such
as resistance to soybean cyst nematode \citep{cook2012}, aluminum tolerance in
wheat \citep{maron2013}, and branching in maize \citep{studer2011}.  SVs have
already been used in GWAS in studies that have shown promising results
\citep[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
their inclusion in GWAS is not systematic.

Despite their high potential, SVs are difficult to discover and genotype with
accuracy. Comprehensive detection of SVs requires a combination of several
methods that often need to be tailored to the study species and data available
\citep{alkan2011, ho2020}. Even then, assessments of the performance of SV
discovery and genotyping generally reveal subpar sensitivity and precision
\citep[e.g.][]{cameron2019, chaisson2019, kosugi2019}. Moreover, despite recent
improvements in computational approaches for genotyping SVs
\citep[e.g.][]{siren2021, ebler2022}, such approaches remain computationally
expensive compared to SNP genotyping workflows. Therefore, discovering and
genotyping SVs at a scale required for GWAS may prove daunting even when
sufficient computational resources and skilled analysts are available.

In this context, alternative approaches that allow to assess a large spectrum
of variants in GWAS without genotyping SVs at population scale could be useful.
\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
represent such an alternative method. This approach relies on the presence or
absence of \emph{k}-mers observed in sequence reads as genotypes for use in
GWAS. Once a presence/absence table of \emph{k}-mers has been generated for the
population under study, this table can be used in association analyses
similarly to SNP or SV genotype calls. One advantage of using \emph{k}-mers is
that they can act as a molecular signature for any type of variant as long as
these variants result in presence/absence of \emph{k}-mers of a given length.
Another advantage is that \emph{k}-mers are not tied to a specific genomic
location and can thus be used to query associations between genotype and
phenotype at genomic locations that are not found in the reference genome.  A
few studies have applied this \emph{k}-mer-based GWAS approach
\citep[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
developed \citep{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
such approaches have not been widely adopted yet.

In this study, we assessed the potential of SV- and \emph{k}-mer-based GWAS in
a major crop, soybean (\emph{Glycine max}).  In order to test these methods, we
analyzed ten qualitative traits \citep{bandillo2017} and two quantitative
traits \citep{bandillo2015} that have been previously studied using a SNP array
developed for soybean \citep{song2013} as well as a quantitative trait
(horizontal resistance to \emph{Phytophthora sojae}) that has been previously
studied using whole-genome sequencing (WGS) data \citep{deronne2022}.  The fact
that the underlying genes and causal variants are already known for several of
these traits provides a good test case for the use of newer GWAS methods, as
the expected result is already known. Our objectives as part of this work were
to:

\begin{enumerate}
	\item Assess the potential of SV- and \emph{k}-mer-based GWAS to
		pinpoint genes and causal variants associated with loci whose
		underlying genes or variants are already known.
	\item Identify potential candidate genes and/or causal variants at loci
		whose underlying genes or candidate variants are not yet known.
	\item Develop a set of computational tools for the downstream analysis
		of significantly associated \emph{k}-mers once these have been
		identified.
\end{enumerate}

\section{Methods}
\label{sv-gwas-methods}

\subsection{Sample selection and processing of sequencing data}
\label{sv-gwas-sequencing-data}

We selected 389 inbred \emph{G. max} accessions based on the availability of
Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
Illumina paired-end sequencing data for those 389 accessions among the ones
analyzed by \cite{bayer2021} and downloaded them from the SRA.  We obtained
reads in FASTQ format using the sratoolkit command fastq-dump v. 2.9.6
(\citeauthor[][\url{https://github.com/ncbi/sra-tools}]{sratoolkit}) with the
option \texttt{-\-split-3} as some of the runs contained reads that were
unpaired. Reads that were not paired in the raw sequencing data were not used
for downstream analyses.  The runs downloaded from the SRA and associated
metadata are listed in Additional file 1.

We filtered the sequencing data for quality and the presence of sequencing
adapters using bbduk from BBtools v. 38.25
(\citeauthor[][\url{https://sourceforge.net/projects/bbmap/}]{bbtools}). We
then aligned reads to the reference assembly version 4 of \emph{G. max}
cultivar Williams82 \citep{valliyodan2019} to which we concatenated chloroplast
and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to align the
reads in paired-end mode. Reads that were left unpaired following adapter- and
quality-trimming were aligned separately in single-end mode using bwa-mem and
merged with the reads aligned in paired-end mode using samtools \texttt{merge}
v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
individual SRA runs to the BAM files and merged the reads belonging to the same
accession using bamaddrg
(\citeauthor[][\url{https://github.com/ekg/bamaddrg}]{bamaddrg}). We used the
resulting BAM files for all downstream analyses requiring aligned reads.
Metadata on mapping depth of aligned data can be found in Additional file 2.

\subsection{Discovery and genotyping of SVs}
\label{sv-gwas-sv-discovery-genotyping}

Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
The \emph{discovery} step consisted in methods leading to a set of candidate SVs,
whereas the \emph{genotyping} step consisted in determining the genotype
of the candidate SVs for all samples in the population. We used three different
sources of data in the SV discovery step to generate a set of candidate variants:

\begin{itemize}
	\item We used Illumina WGS data of the 389 accessions mentioned above
		to call SVs in the population following methods outlined
		in \cite{lemay2022}. In brief, we combined information from
		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
		a set of candidate SVs from Illumina data.
	\item We used SVs discovered by \cite{lemay2022} from Oxford Nanopore
		sequencing data of 17 Canadian soybean cultivars as a second
		source of SVs.
	\item We called SVs from 26 high-quality genome assemblies published by
		\cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
		We used methods similar to those of \cite{liu2020-pangenome} to call
		SVs from the comparison of these assemblies to that of Williams82.
\end{itemize}

We merged the SVs called using each of these approaches using SVmerge
\citep{wong2010} and used the resulting set of candidate SVs for genotyping. We
used two programs with different genotyping models for genotyping this set of
SVs, Paragraph \citep{chen2019} and vg \citep{hickey2020}. In brief, Paragraph
genotypes each SV individually whereas vg builds a variation graph from a set
of candidate variants and then uses this graph to align sequence reads and
genotype variants.  Detailed methods regarding the SV discovery and genotyping
steps can be found in section \ref{annexe-sv-gwas-detailed-methods} of Appendix
\ref{annexe-sv-gwas}.

\subsection{Discovery and genotyping of SNPs and indels}
\label{sv-gwas-snp-discovery}

We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels on the
mapped reads of the 389 cultivars. We initially filtered the 21.1 M SNPs and
indels from the raw output of Platypus to keep only those with:

\begin{enumerate}
	\item \texttt{FILTER} field set to \texttt{PASS}
	\item minor allele frequency (MAF) $\geq$ 0.005
	\item fraction of missing genotypes $\leq$ 0.8
	\item heterozygosity rate $\leq$ 0.2.
\end{enumerate}

The $\sim$7M SNPs and indels kept until this point were included in
the variation graph used for SV genotyping with the vg toolkit (see section
\ref{sv-gwas-sv-genotyping-vg} in Appendix \ref{annexe-sv-gwas}).

The dataset of SNPs and indels used for GWAS was further filtered to achieve a
MAF $\geq$ 0.02 and a fraction of missing genotypes $\leq$ 0.5. Moreover, we
pruned the dataset using the \texttt{-\-indep-pairwise} option of PLINK v.
1.90b5.3 \citep{purcell2007} using a window of 1,000 markers with a step of 100
markers and an $r^2$ LD threshold of 0.9. Markers located on unanchored
scaffolds were also removed. The resulting dataset used for GWAS comprised
924,630 SNPs and indels.

\subsection{Computing the presence/absence table of \emph{k}-mers}
\label{sv-gwas-kmer-pav}

We used the approach outlined by \cite{voichek2020} for \emph{k}-mer-based GWAS
as well as their binary executables (v. 0.2-beta) and recommendations on the
associated GitHub page (\url{https://github.com/voichek/kmersGWAS}) to generate
a table of \emph{k}-mer presence/absence for use in GWAS.  The first step of
the approach required counting the \emph{k}-mers of length 31 (31-mers) present
in the trimmed FASTQ sequencing files of all samples using KMC3 v. 3.2.1
\citep{kokot2017}. We counted \emph{k}-mers twice, the first time using their
canonized form (the first of either the observed \textit{k}-mer or its reverse
complement in lexicographical order) and the second time using the observed
\emph{k}-mers themselves (non-canonized form). After the information from both KMC3 runs was
combined, we filtered the set of \emph{k}-mers to be used for the analysis by
keeping only those seen in at least 5 samples and observed in both canonized
and non-canonized form in at least 20\% of the samples in which they are found.
We then generated a table indicating the presence/absence in all samples of the
\emph{k}-mers that passed all filters. This table was used to compute a kinship
table for use in GWAS and was also used directly as input genotypes for the
GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
computation of kinship and GWAS.

\subsection{Phenotypic data}
\label{sv-gwas-phenotypic-data}

We analyzed ten qualitative traits studied by \cite{bandillo2017}: flower
color, pubescence color, seed coat color, stem termination type, hilum color,
pod color, pubescence form, pubescence density, seed coat luster, and maturity
group.  We obtained phenotypic data for these traits from the GRIN database
(\url{https://npgsweb.ars-grin.gov/gringlobal/search}) by querying the database
using the PI identifiers of the 389 accessions used.  Four out of the 389
accessions did not match the GRIN database and were therefore not used for the
analysis of qualitative traits. We recoded the phenotypes of qualitative traits
to numerical values for GWAS analysis similarly to the numerical coding used by
\cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
using targeted subsets of the observed phenotypes in order to focus on specific
loci. The phenotypes and numerical values used for all analyses as well as the
number of observations in the dataset that we used for GWAS are listed in
Tables \ref{FLOWER.COLOR_gwas_table} to \ref{MATURITY.GROUP_gwas_table}.

We analyzed three quantitative traits in addition to the ten qualitative traits
mentioned above.  We retrieved seed oil and protein content from the GRIN
database. In cases where more than one value was listed for a given accession,
we computed the average of those values and used it in GWAS analyses. We also
used the horizontal resistance to the oomycete \emph{P. sojae} (corrected dry
weight, CDW) assessed in a hydroponic assay \citep{deronne2022} as a third
quantitative trait. Our analysis included 340 accessions whose phenotypic data
was already published by \cite{deronne2022} plus 49 additional accessions for
which sequencing data was not yet available when their study was conducted.
Because of ambiguity regarding the identity of one of the samples (HN019), the
number of samples for which CDW data was available was 388 and this sample
was also dropped from all other GWAS analyses. Similarly to seed oil and
protein content, the CDW values for some samples for which more than one
observation was available were averaged.

Because of clear discrepancies between the observed haplotypes and the
reported phenotypes for some simple traits such as flower color and seed coat
color, we suspected possible errors in the identification of some of the
accessions. We compared the SNP genotypes derived from the SoySNP50K array
and our WGS data (see section \ref{annexe-sv-gwas-soysnp50k} in Appendix
\ref{annexe-sv-gwas} for detailed methods) at over 32,000 SNPs, and
identified 24 samples with < 90\% concordance between the two genotype
datasets. We removed them from the dataset in addition to another
sample that exhibited an atypical GC content suggesting contamination.
Following these filtering steps, 363 samples remained for GWAS analyses. The
concordance between WGS data and SoySNP50K genotypes is included in
Additional file 2 and averaged 98.3\% among retained lines while the
phenotypic data used in this study can be found in Additional file 3.

\subsection{GWAS analyses}
\label{sv-gwas-gwas-methods}

We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
the SNP/indel and SV datasets described above. SVs and indels required some additional
preprocessing steps detailed in \cite{lemay2022gwas} to make variant
representation suitable for downstream analyses. Briefly, we recoded all
variants as SNPs prior to converting the VCF files to Hapmap diploid format
using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
retrieval of the proper metadata for each variant after GWAS.

We ran GAPIT using an MLM model with 9 principal components and the VanRaden
algorithm for computing the kinship matrix. We used an MLM model for
consistency with the statistical model used by the \textit{k}-mer-based
approach.  We used a randomization approach to determine the 5\% family-wise
error-rate threshold for each GWAS analysis as described in \cite{voichek2020}.
To compute this threshold, we permuted the phenotypic observations 100 times
and computed a GWAS on these permuted phenotypes to obtain a distribution of
top \emph{p}-values under the null hypothesis. The fifth most significant of
the 100 top \emph{p}-values was used as the significance threshold for
inference, i.e. markers with \emph{p}-values lower than this threshold were
considered significant.

We conducted GWAS based on \emph{k}-mers similarly to the method outlined by
\cite{voichek2020} and used their binary executables
(\url{https://github.com/voichek/kmersGWAS}) for doing so.  In brief, the
analysis was conducted in two steps. The first step used an approximate model
and identified 1 million possibly significant \emph{k}-mers that were used in
the second step with an exact model implemented in GEMMA \citep{zhou2012} to
compute the \emph{p}-values associated with each of these \emph{k}-mers. We
used the kinship computed from the \emph{k}-mers to correct for relatedness
between accessions. The program implemented by \cite{voichek2020} automatically
returned a list of \emph{k}-mers that passed a 5\% family-wise error-rate
threshold determined by a randomization approach.  These \emph{k}-mers were
considered significant for the purposes of downstream analyses.

\subsection{Analysis of significant \emph{k}-mers}
\label{sv-gwas-kmers-analysis-methods}

\emph{k}-mers are not intrinsically associated with a particular genomic region
and a tailored analysis was therefore required to associate them with genomic
coordinates. To do so, we queried the aligned reads of all samples for matches
to significant \emph{k}-mers for a given trait. Because this is a
computationally demanding operation (there were typically tens of thousands of
significant \emph{k}-mers per trait), we developed a program called katcher
(\url{https://github.com/malemay/katcher}) that is written in the C language
and uses the htslib library \citep{bonfield2021} for efficient reading and
writing of BAM files. katcher and its associated utilities allow the efficient
retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
of interest.

For one trait (flower color), the significance of some \emph{k}-mers was so
high that their \emph{p}-values were numerically equal to 0. In order to
compute $-\log_{10} (p)$ for visualization on Manhattan plots in these cases,
we set the \emph{p}-value to the smallest numerical value that could be
represented using R on our machine (2.225074e-308). In this particular case,
the values represented on the Manhattan plots represent an upper bound on the
\emph{p}-value and not an exact value.

Once we retrieved all the reads containing significant \emph{k}-mers, we still
needed to gather the mapping information contained in those reads for the
purposes of identifying contiguous signals and generating Manhattan plots. To
simplify the amount of information to process in downstream analyses, we
considered only the most significant \emph{k}-mer contained in a given read and
only considered the position where this \emph{k}-mer was most often observed in
the dataset. Moreover, we limited the analysis to reads with a minimum mapping
quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
least one sample to filter out for spurious \emph{k}-mers that may originate
from sequencing errors. All downstream analyses (Manhattan plots,
identification of signals, analysis of genes located near significant
\emph{k}-mers, etc.) used this processed set of significant \emph{k}-mers.

\subsection{Downstream analyses of significant association signals}
\label{sv-gwas-methods-downstream}

We used contiguous regions comprising significant associations, which we called
``signals'', as a basis to compare our results with those obtained by previous
studies and to query the genome for candidate genes.  We defined the boundaries
of those signals by grouping any significant markers or \emph{k}-mers that
occurred within 250 kb of one another.  We did consider the extent of deletions
or \emph{k}-mers in defining the distance between significant associations.

For the GWAS analyses computed from Platypus genotype calls, we retrieved all
markers that had initially been pruned in a region ranging from 50 kb
downstream to 50 kb upstream of significant signals and computed their
\emph{p}-values by running GAPIT using the same parameters as noted above. The
\emph{p}-values thus computed on markers that were initially left out were not
used for initial signal discovery, but provided finer resolution when analyzing
the results of the Platypus GWAS analyses. Indeed, some markers that had not
been included in the original GWAS may be the causal variant or simply yield
more significant \emph{p}-values than those initially used.

To help with the interpretation of results and the identification of candidate
genes, we defined regions of highly significant associations within each signal
by delimiting a range of coordinates containing the top 5\% most significant
associations (Paragraph and vg) or 1\% most significant associations (Platypus
and \emph{k}-mers) found within the signal. In cases where such a fraction
represented fewer than two markers or \emph{k}-mers, the highly significant
region was simply delimited by the two most significant associations.

Because \cite{bandillo2015} and \cite{bandillo2017} used assembly version 1 of
Williams82, we converted the genomic positions of their signals from assembly
version 1 to version 4 by using the same method that was used for converting
SoySNP50K positions (see section \ref{annexe-sv-gwas-soysnp50k}).  The
boundaries of the signals thus translated to genome assembly version 4 were
used for comparison with the signals found in our analyses. For the analysis of
qualitative traits, we only compared our results to signals at named classical
loci and ignored previously reported minor signals with \emph{p}-values that
just barely made it above the significance threshold.

\subsection{Linking significant \emph{k}-mers to variants}
\label{sv-gwas-methods-linking-kmers}

We used the reads retrieved by katcher as a starting point for identifying the
variants associated with a given set of \emph{k}-mers. For any such
\emph{k}-mer set of interest, we used the QNAME field (in the BAM format) of
the reads identified by katcher to extract all reads with a QNAME corresponding
to a read containing significant \emph{k}-mers. We proceeded this way so as to
retrieve the reads that are paired with a significant read even though that
read may not have matched a significant \emph{k}-mer in the first place. Once
all such reads were identified, we converted them to paired-end reads in FASTQ
format using \texttt{samtools fastq} and used the reads for \emph{de novo}
assembly with SPAdes v. 3.15.4 \citep{spades} with the \texttt{-\-careful}
parameter as advised by \cite{voichek2020}. Assemblies were done individually
for each sample and aligned to the Williams82 reference assembly using bwa mem.
Assemblies targeting a region of interest for samples used in a given GWAS were
aligned using multiple sequence alignment with the MAFFT program v. 7.475
\citep{katoh2002} for visualization purposes. We only considered haplotypes
occurring at least five times in the population in such analyses in order to
leave out potential assembly errors. In a single case (\emph{Dt1} locus),
assembly at the locus succeeded for a very low number of samples (26). In this
case, we used consensus sequences from the reads aligned with bwa for comparing
the haplotypes across samples.

\subsection{Computation of LD between \emph{k}-mers}
\label{sv-gwas-ld-kmers}

In some cases, the analysis of \emph{k}-mers revealed signals at previously
undocumented loci or found a very large number of associations throughout the
genome. To shed light on these situations, we computed the pairwise LD between
\emph{k}-mers in order to identify \emph{k}-mers that co-segregated and
likely corresponded to a single locus. Because of the sheer number of pairwise
comparisons that needed to be made for some traits (> 20,000 significant
\emph{k}-mers kept even after the filtering steps mentioned above), we limited
the number of \emph{k}-mers used for LD calculation to 1,500. For traits where
more than 1,500 \emph{k}-mers were found, we computed LD on a subsample of
these by selecting the 500 \emph{k}-mers with the most significant
\emph{p}-values and sampling the remaining 1000 \emph{k}-mers randomly. This
random sampling was done with a probability inversely proportional to the
number of \emph{k}-mers matching a given chromosome or scaffold such that
\emph{k}-mers were sampled from the whole genome and not simply from the most
significant locations.

\subsection{Software used}
\label{sv-gwas-software-used}

Unless otherwise noted, all analyses were conducted using R v. 4.2.0
\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
\citep{obenchain2014}.  We gathered several functions used for the downstream
analysis of GWAS results into a package called gwask which is available on
GitHub (\url{https://github.com/malemay/gwask}).

\section{Results and discussion}
\label{sv-gwas-results-discussion}

We analyzed a total of 13 traits using four different GWAS approaches, one
based on SNP and indel genotypes (Platypus), two based on SV genotypes (Paragraph and
vg), and another based on the presence/absence of \textit{k}-mers.
Because two or three different GWAS analyses were conducted on some traits, a
total of 22 GWAS analyses were performed using each of the four approaches.
Given the large volume of data that this represents, we will focus here on the
most noteworthy results. Readers interested in seeing detailed results on a
per-trait basis are referred to section \ref{annexe-sv-gwas-detailed-results}
in Appendix \ref{annexe-sv-gwas}. In particular, table \ref{signals_table}
provides a summary of the \textit{p}-values obtained by each approach at each
of the loci considered in this study.

\subsection{Loci with known genes or causal mutations}
\label{sv-gwas-results-cloned-genes}

We were first interested in assessing the performance of various GWAS
approaches by analyzing the results at loci for which the genes have
already been cloned. These loci provide interesting test cases because
the expected results (causal genes or variants) are known. Table \ref{loci-table} shows
a summary of the results discussed below.

% Table 1 with results on all loci with cloned genes
\begin{landscape}
	\begin{table}
		\caption[Performance of four GWAS approaches at detecting causal genes/variants]%
		{Performance of four GWAS approaches at detecting causal genes or variants at cloned loci.}
		\label{loci-table}

		\footnotesize
		\centering
		\begin{threeparttable}

			\csvreader[head to column names,
			tabular = llllcccc,
			table head = \toprule Trait & Locus & Gene & Causal variant & Platypus\tnote{a} & vg\tnote{b} & Paragraph\tnote{b} & \textit{k}-mers \\ \midrule,
			table foot = \bottomrule]%
			{../tables/loci_table.csv}%
			{}%
			{\Trait & \Locus & \Gene & \Causal & \platypus & \vg & \paragraph & \kmers}

			\begin{tablenotes}
			\item[a] Platypus genotypes SNPs and indels
			\item[b] vg and Paragraph genotype SVs
			\item[c] Locus not discussed in main text. See section \ref{annexe-sv-gwas-detailed-results} in Appendix \ref{annexe-sv-gwas} for detailed results.
			\item[]- The approach detected no signal overlapping the gene.
			\item[]* The signal found at this locus overlaps the gene.
			\item[]** The region defined by the top 5\% (vg and Paragraph) or 1\% (Platypus and \textit{k}-mers) variants or \textit{k}-mers overlaps the gene.
			\item[]*** The most significantly associated marker or \textit{k}-mer is the causal variant.
			\end{tablenotes}

		\end{threeparttable}
	\end{table}
\end{landscape}

\subsubsection{Flower color -- \textit{W1} locus}
\label{sv-gwas-main-results-flower-color-l1}

\cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
Glyma.13g\-072100 as associated with the \textit{W1} locus for flower color. In
Williams82 and other accessions with white flowers, a 65-bp insertion into the
third exon of the gene results in a premature stop codon that renders the
F3’5’H enzyme non-functional and prevents pigmentation of the flower. All
GWAS approaches detected a signal overlapping the gene at this locus (Figures
\ref{flower_color_manhattan}, \ref{flower_color_W1_signal}) and the most
significant variants/\textit{k}-mers of all approaches except Platypus identified
the known causal variant (Figures \ref{flower-color-main-figure} and
\ref{flower_color_W1_gene}).
As the catalogue of variants analyzed by Platypus contains only SNPs and small
indels, variants located in the gene proved to be among the most closely
associated, but the causal variant could not be captured given the large size
of this variant.

We should note that when we first analyzed the results of the GWAS on flower
color, we observed that two accessions bearing the haplotype typically
associated with purple flowers were noted as having white flowers. Upon further
analysis, we found that the most likely cause of this discrepancy was a
mismatch between the genotype data and the identity of the accession for which
the phenotypic data was obtained. Indeed, when we compared SNP calls made
directly from the WGS data to those made using the SoySNP50K array, we observed
that these two samples had concordance rates of only 62\% and 70\%. To
prevent such mismatches from affecting our GWAS analyses, we computed the
concordance between genotype calls made from WGS data and the SoySNP50K array
for the whole population (see section \ref{sv-gwas-phenotypic-data} in the Methods) and
excluded 24 samples for which the concordance rate was below 90\%.

% Figure 1 with the results on flower color at locus W1
\begin{figure}
	\centering
	\includegraphics[width = 5in]{flower_color_W1_main_figure.png}
	\caption[Results of SV- and \textit{k}-mer-based GWAS at the
	\textit{W1} locus for flower color]{(Caption on the following page.)}
	\label{flower-color-main-figure}
\end{figure}

\begin{figure}
	\contcaption{Results of SV- and \textit{k}-mer-based GWAS at the
	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots of
	the analyses based on Paragraph (SV) genotype calls and
	\textit{k}-mers.  Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{W1} locus. (b) Zoomed-in Manhattan plots of the analyses based
	on Paragraph genotypes and \textit{k}-mers at the \textit{W1} locus.
	Horizontal dotted lines again indicate the significance threshold
	whereas the vertical dotted lines indicate the location of the
	Glyma.13g072100 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 5\% (Paragraph) and
	1\% (\textit{k}-mers) associations.  (c) Gene model of Glyma.13g072100.
	Rectangles represent exons, colored rectangles represent coding
	sequences, and arrows indicate the direction of transcription. The red
	rectangle highlights the region shown in panel (d). (d) Haplotypes and
	significant \textit{k}-mers at the \textit{W1} locus. Nucleotides are
	colored according to the \textit{p}-value of the most significant
	\textit{k}-mer overlapping them. Vertical lines between haplotypes
	indicate sequence differences, whereas dashes indicate gaps in the
	alignment. (e) Contingency table of the phenotypes and haplotypes found
	in the population at the \textit{W1} locus. The haplotypes are the ones
	represented in panel (d).}
\end{figure}

\subsubsection{Pubescence color -- \textit{T} locus}
\label{sv-gwas-main-results-pubescence-color-t}

% Pubescence color - T locus
\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
\textit{T} locus for pubescence color and found a 1-bp deletion as
a putative causal variant at this locus. We found signals overlapping this gene
using all four approaches (Figure \ref{pubescence_color_all_manhattan}). In
particular, the regions defined by the top 1\% most associated variants identified by the
Platypus and \textit{k}-mer approaches included Glyma.06g202300 (Figure
\ref{pubescence_color_all_T_signal}). While the most significant
\textit{k}-mers did occur within the sequence of this gene, they did not
correspond to the causal variant (Figure \ref{pubescence_color_all_T_gene}).
Still, the fourth and fifth most strongly associated \textit{k}-mers
corresponded to the causal variant (Figures \ref{pubescence_color_all_T_gene},
\ref{pubescence_color_all_T_kmers}). As for Platypus, the most significant
variant within the sequence of Glyma.06g202300 was the causal
variant, however it was not the most significantly associated variant overall
(Figures \ref{pubescence_color_all_T_signal},
\ref{pubescence_color_all_T_gene}). Neither vg nor Paragraph detected the causal
variant at this locus given that these approaches did not target small indels.

\subsubsection{Pubescence color -- \textit{Td} locus}
\label{sv-gwas-main-results-pubescence-color-td}

% Pubescence color - Td locus
We conducted a separate GWAS analysis contrasting only accessions with tawny and light
tawny pubescence because the \textit{Td} locus is known to control
these differences in color. Accordingly, we
detected signals in this region using all four approaches, but only Platypus
and the \textit{k}-mers identified signals that overlapped Glyma.03g258700
(Figures \ref{pubescence-color-main-figure}a-b,
\ref{pubescence_color_nogray_manhattan}, \ref{pubescence_color_nogray_Td_signal}).
While the most significant \textit{k}-mer identified the causal variant documented
at this locus \citep{yan2020}, the fourth most significantly associated variant
found by Platypus corresponded to this variant (Figures
\ref{pubescence-color-main-figure}b-d, \ref{pubescence_color_nogray_Td_gene}).
Interestingly, some of the accessions bearing the haplotype associated with tawny pubescence
displayed light tawny pubescence, suggesting that there may be more than one causal
variant at this locus (Figure \ref{pubescence-color-main-figure}e).

% Figure 2 with the results on pubescence color at locus Td
\begin{figure}
	\centering
	\includegraphics[width = 5in]{pubescence_color_nogray_Td_main_figure.png}
	\caption[Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Td} locus for pubescence color]{(Caption on the following page.)}
	\label{pubescence-color-main-figure}
\end{figure}

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots
	of the analyses based on Platypus (SNP/indel) genotype calls and
	\textit{k}-mers.  Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Td} locus. (b) Zoomed-in Manhattan plots of the analyses based
	on Platypus genotypes and \textit{k}-mers at the \textit{Td} locus.
	Horizontal dotted lines again indicate the significance threshold
	whereas the vertical dotted lines indicate the location of the
	Glyma.03g258700 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of Platypus, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrow indicates the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Haplotypes and significant \textit{k}-mers at
	the \textit{Td} locus. Nucleotides are colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{Td} locus. The haplotypes are the ones represented in panel
	(d).}
\end{figure}

\subsubsection{Seed coat color - \textit{I} locus}
\label{sv-gwas-main-results-seed-coat-color-i}

% Seed coat color - I locus
The causal variant at the \textit{I} locus for seed coat color is a complex SV
that takes the form of a 10.91-kb inverted duplication of three chalcone
synthase (CHS) genes \citep[CHS1, CHS3 and CHS4;][]{tuteja2008}. The presence of
this inverted duplication results in the silencing of all CHS genes in the seed
coat specifically and thus in the absence of seed coat pigmentation
\citep{tuteja2009}.  We did identify signals overlapping this causal variant
using all four GWAS approaches (Figures \ref{seed_coat_color_all_manhattan},
\ref{seed_coat_color_all_I_signal}), however none of the approaches identified
the causal variant. This failure to detect the causal variant is likely due
to its complexity, which makes it difficult to detect and genotype. While it
is likely that at least some of the significantly associated \textit{k}-mers
actually are derived from this causal variant, the lack of systematic methods for
linking \textit{k}-mers to sequence variation hindered our ability to do so.

\subsubsection{Seed coat color - \textit{G} locus}
\label{sv-gwas-main-results-seed-coat-color-g}

% Seed coat color - G locus
We conducted a second GWAS analysis contrasting accessions with yellow and
green seed coats in order to target the \textit{G} locus. Accordingly, this
analysis found strong signals overlapping the \emph{G} locus using all four
approaches (Figure \ref{seed_coat_color_greenyellow_manhattan}).
\cite{wang2018} identified Glyma.01g198500 as the gene associated with
this locus and found an A>G SNP affecting transcript splicing as the causal
variant at that locus.  Consistently with their results, we identified the two
most significant \emph{k}-mers and the second most significant marker from the
Platypus analysis as corresponding to that causal variant (Figures
\ref{seed-coat-color-main-figure},
\ref{seed_coat_color_greenyellow_G_signal},
\ref{seed_coat_color_greenyellow_G_gene}). Unsurprisingly, none of the
SV-based GWAS analyses identified the causal variant given its nature.

% Figure 3 with the results on pubescence color at locus G
\begin{figure}
	\centering
	\includegraphics[width = 5in]{seed_coat_color_greenyellow_G_main_figure.png}
	\caption[Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{G} locus for seed coat color]{(Caption on the following page.)}
	\label{seed-coat-color-main-figure}
\end{figure}

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots
	of the analyses based on Platypus (SNP/indel) genotype calls and
	\textit{k}-mers.  Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{G} locus. (b) Zoomed-in Manhattan plots of the analyses based
	on Platypus genotypes and \textit{k}-mers at the \textit{G} locus.
	Horizontal dotted lines again indicate the significance threshold
	whereas the vertical dotted lines indicate the location of the
	Glyma.01g198500 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of Platypus, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Haplotypes and significant \textit{k}-mers at
	the \textit{G} locus. Nucleotides are colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{G} locus. The haplotypes are the ones represented in panel (d).}
\end{figure}

% Stem termination - Dt1 locus
\subsubsection{Stem termination type - \textit{Dt1} locus}
\label{sv-gwas-main-results-stem-termination-dt1}

While all GWAS approaches found signals near the Glyma.19g194300 gene
associated with the \textit{Dt1} locus for stem termination type, the
\textit{k}-mer and Platypus analyses performed best since the region defined by the most highly associated markers
(top 1\%) included the causal gene (Figures
\ref{stem_termination_all_manhattan}, \ref{stem_termination_all_Dt1_signal}).
We did identify one of the SNPs suggested as putative causal variants for this
locus by \cite{tian2010} and \cite{liu2010}, however this SNP was only the
135\textsuperscript{th} most associated marker found by Platypus and the
12\textsuperscript{th} most associated \emph{k}-mer (Figures
\ref{stem_termination_all_Dt1_gene}, \ref{stem_termination_all_Dt1_kmers}).
Our results are consistent with those of \cite{liu2010} and \cite{tian2010} in
suggesting that there is more than one causal variant at that locus. Indeed, a
few cultivars bearing the allele associated with \emph{dt1} showed
indeterminate or semi-determinate phenotypes and some cultivars showed a
determinate phenotype without bearing the allele associated with \emph{dt1}
(Figure \ref{stem_termination_all_Dt1_kmers}). The existence of several causal
variants might explain why our analyses did not identify a causal variant more
clearly.

% Hilum color - T locus
\subsubsection{Hilum color -- \textit{T} locus}
\label{sv-gwas-main-results-hilum-color-t}

The \textit{T} locus mentioned above for pubescence color is also a determining
locus for hilum color, and we detected signals overlapping the corresponding gene
using all approaches but vg (Figures \ref{hilum_color_all_manhattan},
\ref{hilum_color_all_T_signal}). The regions defined by the top 5\% (Paragraph)
and top 1\% (Platypus and \textit{k}-mers) variants/\textit{k}-mers included
the Glyma.06g202300 gene associated with this locus. Similarly to what we
observed for the analysis of the \textit{T} locus based on pubescence color, we
did identify the documented causal variant at this locus using the Platypus and
\textit{k}-mer analyses, although these variants were not the most
significantly associated (Figures \ref{hilum_color_all_T_gene},
\ref{hilum_color_all_T_kmers}).

% Hilum color - I locus
\subsubsection{Hilum color -- \textit{I} locus}
\label{sv-gwas-main-results-hilum-color-i}

Similarly, the \textit{I} locus discussed above for its effect on seed coat
color also controls hilum color and we detected signals overlapping the
associated inverted duplication using all approaches but vg (Figures
\ref{hilum_color_all_manhattan}, \ref{hilum_color_all_I_signal}). However,
as was the case for the effect of \textit{I} on seed coat color, we were
not able to link any of the variants identified to the causal variation given
its highly complex nature.

% Hilum color - R locus
\subsubsection{Hilum color -- \textit{R} locus}
\label{sv-gwas-main-results-hilum-color-r}

We conducted a GWAS analysis contrasting accessions with black and brown hila
in an attempt to detect stronger signals at the \textit{R} locus.  We
accordingly detected signals overlapping this locus using all approaches,
although these signals overlapped the Glyma.09g235100 gene associated with the
\textit{R} locus only for Platypus and \textit{k}-mers (Figures
\ref{hilum_color_blackbrown_manhattan}, \ref{hilum_color_blackbrown_R_signal}).
We observed two causal variants previously documented by \cite{gillman2011},
but these were not among the most significantly associated variants or
\textit{k}-mers (Figures \ref{hilum_color_blackbrown_R_gene},
\ref{hilum_color_blackbrown_R_kmers}).  This failure to identify causal
variants more directly is probably due to the existence of more than one causal
variant at this locus. Interestingly, the most significantly associated
\textit{k}-mers mapped just upstream of a gene (Glyma.09g234900) identified as
a homolog of Glyma.09g235100 \citep{gillman2011}. This result highlights how
difficult it may be to link significant \textit{k}-mers to causal variation in
some cases.

% Pubescence density - Ps locus
\subsubsection{Pubescence density -- Ps locus}
\label{sv-gwas-main-results-pubescence-density-ps}

We conducted a GWAS analysis contrasting accessions with normal and semi-sparse
pubes\-cence in order to detect loci associated with pubescence density. We
detected signals at the \textit{Ps} locus using all four approaches (Figures
\ref{pubescence-density-main-figure}a-b, \ref{pubescence_density_manhattan}). The causal
variant at this recently cloned locus is a copy number variant overlapping the
Glyma.12g187200 gene, with higher copy number resulting in reduced pubescence \citep{liu2020ps}.
We identified a region of $\sim$42 kb that varied in copy number across
the population and was associated with the trait (Figure
\ref{pubescence-density-main-figure}c), and therefore possibly corresponded to the
25.6-kb region identified as the causal CNV by \cite{liu2020ps}. We found that
the most significantly associated \textit{k}-mer was located on the boundary of this CNV
and that reads containing this \textit{k}-mer mapped to either of the two ends
of the CNV (Figure \ref{pubescence-density-main-figure}b,d). This suggests that
the most associated \textit{k}-mer is indeed linked to the CNV
that is causal at this locus, while \textit{k}-mers within the
CNV region may represent differences in sequence between the copies.

% Figure 4 with the results on pubescence density at locus Ps
\begin{figure}
	\centering
	\includegraphics[width = 6in]{pubescence_density_Ps_main_figure.png}
	\caption[Results of \textit{k}-mer-based GWAS at the
	\textit{Ps} locus for pubescence density.]{Results of \textit{k}-mer-based GWAS at the
	\textit{Ps} locus for pubescence density.
	(a) Genome-wide Manhattan plot of significant \textit{k}-mers. The
	horizontal dotted line indicates the significance threshold whereas the
	vertical dotted line indicates the location of the \textit{Ps} locus.
	(b) Zoomed-in Manhattan plot at the \textit{Ps} locus.  The horizontal
	dotted line represents the significance threshold whereas the
	vertical dotted line indicates the location of the most significant
	\textit{k}-mer. The shaded gray area shows the extent of the causal CNV.
	(c) Estimated copy number (sequencing depth at position divided by
	sample average sequencing depth) of accessions with normal and
	semi-sparse pubescence at the \textit{Ps} locus. Solid lines represent
	the average copy number across 73 (semi-sparse pubescence) and 180
	samples (normal pubescence) computed using a 100-bp sliding window.
	Shaded areas outline the first and third quartiles. The start and end
	of the CNV region can be clearly inferred from the divergence between
	the copy number of samples with contrasting pubescence.
	(d) Histogram of the mapping location of 1,169 reads containing the
	most significant \textit{k}-mer in samples with semi-sparse pubescence.
	Reads containing this \textit{k}-mer map to either end of the CNV
	region, suggesting that they are associated with this variant. Panels
	b, c and d share the same x-axis.}
	\label{pubescence-density-main-figure}
\end{figure}

% Seed coat luster - B locus
\subsubsection{Seed coat luster -- B locus}
\label{sv-gwas-main-results-seed-coat-luster-b}

We conducted a GWAS analysis contrasting dull and shiny seed coat phenotypes
for the seed coat luster trait. Seed coat luster was shown by \cite{gijzen1999}
to be largely caused by the deposition of a hydrophobic protein (HPS)
at the seed surface. Copy number variation of this gene at the \textit{B}
locus was later shown to explain variation in luster, although sequence
variation in the HPS sequence may also play a role \citep{gijzen2006}.
We found signals overlapping the causal CNV using all approaches, and the
most significantly associated \textit{k}-mer was notably located
within the CNV region (Figures \ref{seed_coat_luster_dullshiny_manhattan},
\ref{seed_coat_luster_dullshiny_B_signal}). However, we were not able
to directly link any of the variants or \textit{k}-mers identified to the causal
CNV.

% Maturity group loci
\subsubsection{Maturity}
\label{sv-gwas-main-results-maturity}

We did not detect any signals associated with maturity using our dataset
(Figure \ref{maturity_group_manhattan}). The use of maturity group as a proxy
for time to maturity instead of the number of days, combined with the limited
phenotypic variation at this trait in this collection of accessions (Table
\ref{MATURITY.GROUP_gwas_table}), likely explains our failure to detect
signals.

% Seed oil and protein
\subsubsection{Seed oil and protein content}
\label{sv-gwas-main-results-oil-protein}

We expected to detect signals at two loci controlling seed oil and protein
content on chromosomes 15 and 20 \citep{bandillo2015}. However, none of our
analyses detected signals overlapping the genes cloned for these loci (Figures
\ref{oil_manhattan}, \ref{protein_manhattan}). This is probably due to the fact
that the oil and protein content data that we used for GWAS were not obtained
from orthogonal trials, which was a clear limitation to this GWAS given our
relatively small sample size.

\subsection{Analysis of loci with unknown causal genes}
\label{sv-gwas-results}

% Pod color - L1 and L2 loci
\subsubsection{Pod color -- \textit{L1} and \textit{L2} loci}
\label{sv-gwas-main-results-pod-color-l1-l2}

In addition to analyzing loci for which the genes are already known, we were
able to suggest candidate genes at loci for which the underlying genes are not
known yet. For example, both the \textit{L1} and the \textit{L2} loci for pod
color have yet to be cloned. While \cite{he2015} proposed the gene
Glyma.19g101700 as likely corresponding to \textit{L1} based on fine mapping
and expression analysis, our results as well as those reported by
\cite{bandillo2017} do not support this gene as a candidate due to its location
$\sim$3 Mb away from GWAS signals. Our analyses based on
\textit{k}-mers and Platypus found signals overlapping 14 and 6
genes, respectively (Figures \ref{pod_color_all_manhattan},
\ref{pod_color_all_L1_signal}). Of these, a MATE transporter (Glyma.19g120300)
appears as a prime candidate as the most significant \emph{k}-mer was observed
within its sequence and it is highly expressed in pods according to data hosted
on SoyBase \citep{grant2010, severin2010}. Moreover, the role of some MATE
proteins in transporting flavonoids has been demonstrated \citep{chen2015}.

The signals found by Platypus and \emph{k}-mer analyses at the \textit{L2}
locus overlap 24 and 50 genes, respectively (Figures
\ref{pod_color_all_manhattan}, \ref{pod_color_all_L2_signal}). Despite this
number of genes being somewhat high, we noticed that the closest gene to the
most significant \emph{k}-mer (Glyma.03g005800) is also a MATE transporter,
similarly to our top candidate for \emph{L1}. This gene may therefore represent
an interesting candidate for this locus.

% Pubescence form - Pa1 and Pa2 loci
\subsubsection{Pubescence form -- \textit{Pa1} and \textit{Pa2} loci}
\label{sv-gwas-main-results-pubescence-form-pa1-pa2}

The two known loci for pubescence form in soybean, \textit{Pa1} and
\textit{Pa2}, also have not been cloned yet, although \cite{gilbert2017} did
suggest Glyma.12g213900 as a candidate for \textit{Pa1} based on data published
by \cite{bandillo2017}. This gene encodes a Myb transcription factor annotated
as playing a role in trichome branching and is located near the most
significant signals found by the Platypus, Paragraph and \emph{k}-mer GWAS
analyses (Figures \ref{pubescence-form-main-figure}a-b,
\ref{pubescence_form_all_manhattan}, \ref{pubescence_form_all_Pa1_signal}).
GWAS analyses based on Platypus and \emph{k}-mers identified two highly
significant non-synonymous SNPs associated with appressed pubescence in this
gene (Figures \ref{pubescence-form-main-figure}c-d, \ref{pubescence_form_all_Pa1_gene},
\ref{pubescence_form_all_Pa1_kmers}). The alternative alleles for these two
SNPs co-occurred together in our dataset and may represent causal variants at
the \emph{Pa1} locus. The signals observed at the \emph{Pa2} locus were much
weaker and the signal observed from the \emph{k}-mer GWAS overlapped a total of
20 genes (Figures \ref{pubescence_form_all_manhattan},
\ref{pubescence_form_all_Pa2_signal}), none of which appeared obviously linked
to pubescence form.

% Figure 5 with the results on pubescence form at locus Pa1
\begin{figure}
	\centering
	\includegraphics[width = 5in]{pubescence_form_all_Pa1_main_figure.png}
	\caption[Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Pa1} locus for pubescence form]{(Caption on the following page.)}
	\label{pubescence-form-main-figure}
\end{figure}

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots
	of the analyses based on Platypus (SNP/indel) genotype calls and
	\textit{k}-mers. Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots of the analyses based
	on Platypus genotypes and \textit{k}-mers at the \textit{Pa1} locus.
	Horizontal dotted lines again indicate the significance threshold
	whereas the vertical dotted lines indicate the location of the
	Glyma.12g213900 gene suggested as a candidate for this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of Platypus, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.12g213900. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription.
	(d) Zoomed-in Manhattan plots at the location of the Glyma.12g213900 gene.
	Data is represented on the same x-scale in these plots as the gene model
	in panel (c). Vertical lines indicate the location of two non-synonymous
	SNPs associated with appressed pubescence, which may represent causal variants
	at this locus.}
\end{figure}

% Resistance to Phytophthora sojae
\subsubsection{Resistance to \textit{P. sojae}}
\label{sv-gwas-main-results-psojae}

We used corrected dry weight (CDW) as described by \cite{deronne2022} as a
measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
overlapping the genomic region reported by these authors on Gm15 using all four GWAS
approaches (Figure \ref{corrected_dry_weight_manhattan}). \cite{deronne2022}
proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
variation at that locus based on a combination of functional annotation and
contrasting gene expression in resistant and susceptible lines four days after
inoculation. Although the signals found by our analyses did overlap with
Glyma.15g217100, we found that this gene was located at one end of the signal
and was found in a region with markers or \emph{k}-mers showing a weaker
association (Figure \ref{corrected_dry_weight_cdwGm15_signal}).  Although the
large size of the signal means that the gene associated with the phenotype could
be located almost anywhere in this region, we identified a particularly
interesting region defined by the top 1\% associations found from the Platypus
analysis. This region overlaps only three annotated genes and also contains the
most significant \emph{k}-mer (Figure
\ref{corrected_dry_weight_cdwGm15_signal}), which was located 30 kb away from a
gene (Glyma.15g217700) annotated as a calcium ion binding protein with an
EF-hand domain. Given the important role played by calcium signaling in defense
responses \citep{zhang2014}, we suggest Glyma.15g217700 as an interesting
candidate gene to explain resistance to \emph{P. sojae} in this population.

% Stem terminination and pod color - novel loci
\subsubsection{Novel loci for stem termination and pod color}
\label{sv-gwas-main-results-novel-loci}

In a few cases, we were also able to document possibly novel loci from the
results of \textit{k}-mer GWAS. In a GWAS analysis contrasting accessions
with indeterminate and semi-determinate stem termination types, we observed
previously undocumented signals on chromosomes Gm11, Gm16 and Gm18 in
addition to the known \textit{Dt1} locus (Figure \ref{stem-termination-main-figure}a).
Analysis of the pairwise LD between significant \textit{k}-mers found no
LD between these signals, suggesting that they may represent \textit{bona fide}
loci associated with stem termination type
(Figure \ref{stem-termination-main-figure}b). Similarly, a \textit{k}-mer GWAS
analysis contrasting accessions with black and brown pods found a previously
undocumented signal on chromosome Gm15 that showed no LD with \textit{k}-mers
at the \textit{L1} locus and may therefore represent a different locus (Figures
\ref{pod_color_blbr_manhattan}, \ref{pod_color_blbr_ld}).

% Figure 6 with the results novel loci for stem termination type
\begin{figure}
	\centering
	\includegraphics[width = 6in]{stem_termination_sn_main_figure.png}
	\caption[Detection of novel signals for stem termination type
	using a \textit{k}-mer GWAS analysis contrasting accessions with
	indeterminate and semi-determinate stems]{Detection of novel signals for stem termination type
	using a \textit{k}-mer GWAS analysis contrasting accessions with
	indeterminate and semi-determinate stems.
	(a) Genome-wide Manhattan plot of significant \textit{k}-mers.
	The horizontal dotted line indicates the significance threshold
	whereas vertical dotted lines indicate the location of the \textit{Dt1}
	classical locus as well as three newly documented loci (\textit{stGm11},
	\textit{stGm16}, and \textit{stGm18}).
	(b) Pairwise LD among the 162 significant \textit{k}-mers identified.
	\textit{k}-mers are sorted along the y-axis according to their
	putative position along the reference assembly version 4 of Williams82,
	as identified by ``Gm'' chromosome labels. \textit{k}-mers are
	represented in the same order along the x- and y-axis. The colored
	rectangles drawn below the x-axis represent the
	-log\textsubscript{10}(\textit{p}) of each \textit{k}-mer.}
	\label{stem-termination-main-figure}
\end{figure}

\subsection{Potential of \emph{k}-mer-based GWAS}
\label{sv-gwas-potential-kmers}

Arguably, the most striking result from the analyses presented above is the
impressive performance of \emph{k}-mer-based GWAS compared to SNP/indel- and
SV-based GWAS. Indeed, the GWAS analyses using \emph{k}-mers readily identified
the causal variant as the most significantly associated variant at four loci
with various types of causal variants: \emph{W1} for flower color, \emph{Td}
for pubescence color, \emph{G} for seed coat color, and \emph{Ps} for
pubescence density. In comparison, this result was achieved by SV-based GWAS
at the \textit{W1} locus only, and at none of the loci for SNP/indel-based GWAS.
The \textit{k}-mer approach also systematically performed at least as well as
the other three approaches in identifying signals overlapping known variants or genes
(Table \ref{loci-table}). Moreover, whenever \textit{k}-mer-based GWAS did
detect a signal, the region defined by the 1\% most significant \textit{k}-mers
overlapped the known causal variant or gene.

Even in cases where the most significantly associated \emph{k}-mers did not
pinpoint known causal variants, they were often located near
those variants. For example, the most significantly associated \emph{k}-mers
mapping to the \emph{T} locus for pubescence color mapped to intronic sequences
of the gene known to be associated with this locus, even though they did not
correspond to the documented causal variant. Similarly, the most significantly associated
\emph{k}-mers at the \emph{B} locus for seed coat luster mapped within a
genomic interval known to vary in copy number at that locus, and the most
significant \emph{k}-mers at the \emph{R} locus mapped just upstream of a
homolog of the gene associated with the locus. Even though the \emph{k}-mer
analysis did not identify the causal variants directly in those cases, it still
provided a very strong starting point for candidate gene identification.

As would be expected of any GWAS, the success of the \emph{k}-mer approach
depends on the quality of the phenotypic data and the complexity of the genetic
architecture of the trait. It is clear that the loci for which the
\emph{k}-mer approach was most successful (\emph{W1}, \emph{Td}, \emph{G} and
\emph{Ps}) have a simple genetic architecture and phenotypes that can be easily
determined. For other traits, challenges in determining accurate phenotypes may
result in a decreased ability to identify causal variants or genes. For
example, \cite{bandillo2017} noted that observed phenotypes for stem
termination type in the GRIN database may not correspond to the expected
genotypes at classical loci, and that the intermediate phenotype for seed coat
luster is not always consistent across different environments. For qualitative
traits, the numerical scale used to recode phenotypes for use in GWAS is somewhat
arbitrary and may influence the final results. In the case of quantitative
phenotypes, the usual limitations to the precision and stochastic variation of
phenotypic values will necessarily make the interpretation more difficult than
for simple qualitative traits. Notwithstanding those limitations, it remains
clear that \textit{k}-mer GWAS performed better than any of the other
approaches tested here given the same phenotypic data

\subsection{Limitations and challenges of \emph{k}-mer-based GWAS}
\label{limitations-of-kmers-GWAS}

While we have obtained considerable success when using \emph{k}-mers for GWAS, this
approach does present unique challenges and limitations. One obvious limitation
is that using the presence/absence of \emph{k}-mers may not be appropriate in
species where heterozygosity is common (i.e., most species other than inbred
crops) as the \emph{k}-mers associated with the presence of both alleles will
be observed at heterozygous loci. Moreover, although the presence/absence of
\emph{k}-mers can detect variation independently of its type, some types of
variation may be more difficult to detect than others. More precisely,
variation will only be detected if it results in the presence or absence of a
specific \emph{k}-mer or set of \emph{k}-mers of a given length (here, 31 bases) in a
sequencing dataset. For example, copy number variation at a locus may not be
detected from presence/absence alone if all copies are identical and occur in
tandem, such that no \textit{k}-mer can differentiate between two copies and
more. For similar reasons, \emph{k}-mers may also fail to capture variation
occurring in repeated sequences. Alternative models based on \emph{k}-mer
counts instead of presence/absence have been developed \citep{rahman2018,
he2021} which may solve issues like these and broaden the scope of variants
that can be identified from \emph{k}-mers.

The complexity of analyzing the results of \emph{k}-mer-based GWAS is
another limitation of this approach. Although the method developed by
\cite{voichek2020} is in itself relatively simple and computationally
inexpensive, downstream analysis and interpretation of the output is
challenging. Indeed, there are no state-of-the-art methods for identifying the
putative genomic coordinates of a \emph{k}-mer, grouping significant
\emph{k}-mers together into coherent signals, and linking them to biologically
meaningful sequence variation. The fact that this approach is
reference-agnostic suggests that a paradigm shift will be needed to move away
from reference-based analysis to better suited methods. For example, mapping
the \emph{k}-mers or reads to a graph-based genome \citep[using e.g. the vg
toolkit;][]{siren2021} instead of a linear reference may represent an
improvement. Grouping significant \emph{k}-mers by LD instead of by genomic
position, as was done for some of the analyses in this paper, should also prove
useful in determining how many loci control a trait and as a starting point to
identify the variation underlying presence/absence patterns. While we used an
assembly-based approach similar to what others have previously done
\citep[e.g.][]{voichek2020, rahman2018} and have been able to link
\emph{k}-mers to their underlying variation in simple cases involving SNPs or
indels, we have not been able to link significantly associated \emph{k}-mers to
more challenging causal variants such as those explaining the \emph{I} or
\emph{B} loci. In this context, developing approaches to systematically link
significant \emph{k}-mers to the sequence variation that underlies their
presence is much needed. Finding ways to limit the number of spurious
associations found for some traits, as was the case for flower color and
pubescence color in our study, is also needed in order to obtain robust results
from \emph{k}-mer-based GWAS.

In this work, we aimed to provide a set of tools enabling the downstream
analysis of \emph{k}-mer-based GWAS. Most importantly, we developed the
katcher program for retrieving all reads containing
any of a set of significant \emph{k}-mers in a computationally efficient manner. Previously published studies usually
mapped \emph{k}-mers to a reference genome to identify their reference-based
coordinates \citep[e.g.][]{voichek2020, he2021, rahman2018, tripodi2021}.
However, while this approach may work for \emph{k}-mers that show little
divergence from the reference, it might fail to identify the genomic location
of \emph{k}-mers that diverge significantly or are not found at all in the
reference. Moreover, it does not take into account the sequence context where
the \emph{k}-mer was found, which may provide valuable information for mapping
purposes. Our approach avoids these pitfalls by systematically retrieving reads
containing significant \emph{k}-mers, including those that did not map at all
to the reference assembly. While we were able to position the most significant
\emph{k}-mers on the reference sequence in our analyses, it may not be the case
for highly variable species or if the causal variation at a locus results from
a large novel insertion.

\subsection{Use of SV genotypes for GWAS}
\label{use-of-svs-for-gwas}

In addition to \emph{k}-mers, we also assessed the use of SVs in GWAS. Our
rationale was that since SVs have an important functional impact, they may
facilitate the identification of causal variants when these are in fact SVs.
Although we did find the causal variant as the most significantly associated SV
at the \emph{W1} locus, results have otherwise been rather inconclusive.
Indeed, most of the other known or putative causal variants were SNPs or
indels, and therefore were not represented in the SV datasets. GWAS analyses at
other loci where SVs are known to be the underlying causal variant (e.g.
\emph{I}, \emph{B}, \emph{Ps}) did not find the causal variants, most likely
due to the difficulty of calling and genotyping such variants
\citep{kosugi2019, lemay2022}.

Based on these results, we believe that it may not be worth, at least in
soybean, to conduct GWAS based on SVs. Given the current performance of SV
genotyping programs, it is not clear whether using SVs for GWAS provides any
advantage over using \emph{k}-mer- or SNP-based GWAS, or a combination of both.
Indeed, in the single case where the SV-based analyses identified the causal
variant, the \emph{k}-mer-based GWAS identified it as well. Moreover,
SNP/indel-based GWAS provided better guidance in delimiting the signals found
at a given locus given the much higher density of markers in the SNP dataset
than in the SV dataset. The usefulness of SV-based GWAS compared to SNP-based
GWAS will largely depend on whether SVs can represent variation that is not
otherwise in LD with SNPs. In tomato, \cite{dominguez2020} found that most
transposable element insertions associated with phenotypic variation in
agronomic traits were not tagged by SNPs. On the other hand, in humans,
\cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
is therefore likely that the ``tagability'' of SVs using SNPs will depend on
the particular type of variant and on the species being studied. Even though
SVs remain important components of sequence and functional variation, a more
efficient approach than SV-based GWAS may be to conduct \emph{k}-mer-based GWAS
and only then use SV discovery methods to identify SVs in the population and
link them to significant \emph{k}-mers. While previous studies have found
interesting results using SVs for GWAS \citep[e.g.][]{zhang2015, akakpo2020,
dominguez2020}, we would be curious to see whether these results
could be replicated using \emph{k}-mer-based GWAS.

\section{Conclusion}

In conclusion, we used SV-based and \emph{k}-mer-based GWAS to study ten
qualitative traits and three quantitative traits in soybean cultivars and
compared the results to traditional SNP/indel-based GWAS. While
\emph{k}-mer-based GWAS proved to be a powerful approach in pinpointing the
causal variation or genes associated with known loci, it is unclear whether
conducting SV-based GWAS is worth the investment. We also used the
results from the \emph{k}-mer- and SNP-based GWAS to suggest candidate genes
for a few classical loci that have yet to be cloned. Based on our results, we
believe that an optimal workflow may involve conducting SNP and \emph{k}-mer
GWAS in parallel to identify significant signals and candidate genes, and then
possibly use SV datasets to identify interesting variants. Given the large
number of accessions for which WGS data are now available in several species,
applying \emph{k}-mer-based GWAS to leverage already existing sequencing and
phenotypic data appears feasible and promising. As part of our work, we
developed several computational tools that should help other researchers to
analyze the results of a \emph{k}-mer-based GWAS analysis. However, much work
remains to be done to define state-of-the-art methods for the downstream
analysis of significant \emph{k}-mers. In particular, better approaches are
needed for linking \emph{k}-mers to sequence variation and moving from
presence/absence-based methods to count-based methods.

\section*{Additional files}
\label{sv-gwas-additional-files}

All additional files can be accessed on figshare \citep{kmer-gwas-figshare}.

\begin{description}
	\item[Additional file 1.csv] Metadata on the sequencing runs retrieved
		from the Sequence Read Archive (SRA) (CSV 71 KB)
	\item[Additional file 2.csv] Metadata on individual samples following mapping
		and comparison to SoySNP50K data (CSV 36KB)
	\item[Additional file 3.csv] Phenotypic dataset used for GWAS analyses (CSV 135
		KB)
\end{description}

\section*{Availability of data and materials}
\label{sv-gwas-availability}

Some of the datasets generated as part of this work are available on figshare
\citep{kmer-gwas-figshare}.

The Illumina sequencing data used for the analyses are listed in Additional file 1
and available through NCBI BioProjects \citeauthor{PRJNA257011}, \citeauthor{PRJNA289660}
and \citeauthor{PRJNA639876}.

The high-quality assemblies generated by \cite{liu2020-pangenome} are available
on the Genome Warehouse through Accession Number \citeauthor{PRJCA002030}.

The assemblies of ZH13, W05 and Lee are available on SoyBase
(\url{https://soybase.org/GlycineBlastPages/blast_descriptions.php}).

The SVs identified from Oxford Nanopore data by \cite{lemay2022} are available
on figshare \citep{soybean-sv-figshare}.

The following software used in this work is available on GitHub:

\begin{itemize}
	\item Code used for the analyses:
		\url{https://github.com/malemay/soybean_kmer_gwas}
	\item katcher software for retrieving reads containing \textit{k}-mers:
		\url{https://github.com/malemay/katcher}
	\item gwask R package used for processing the output of GAPIT and
		\textit{k}-mer GWAS analyses, and for plotting:
		\url{https://github.com/malemay/gwask}
	\item Forked svmu version that was used for the analyses shown in this
		paper: \url{https://github.com/malemay/svmutools}
	\item svmutools R package for converting the output of svmu to VCF
		format: \url{https://github.com/malemay/svmutools}
\end{itemize}

\section*{Funding}
\label{sv-gwas-funding}

This work was supported by the SoyaGen grant (\url{https://www.soyagen.ca})
awarded to F. Belzile and R. Bélanger, and funded by Génome Québec, Genome
Canada, the government of Canada, the Ministère de l'Économie, Science et
Innovation du Québec, Semences Prograin Inc., Syngenta Canada Inc., Sevita
Genetics, Coop Fédérée, Grain Farmers of Ontario, Saskatchewan Pulse Growers,
Manitoba Pulse \& Soybean Growers, the Canadian Field Crop Research Alliance
and Producteurs de grains du Québec. M-A. Lemay has been supported by a NSERC
Canada Vanier Graduate Scholarship, a FRQNT doctoral B2X scholarship, a NSERC
Michael Smith Foreign Study Supplement, and a scholarship from the
AgroPhytoSciences NSERC CREATE Training Program. None of the funding bodies
were involved in study design, data acquisition, data analysis, interpretation
of the results, or manuscript writing.

\section*{Authors' contributions}
\label{sv-gwas-contributions}

Conception and design of the study: MAL, JAS, MDR, RB, FB.
Phenotypic data acquisition (resistance to \textit{Phytophthora sojae}): MDR, RB.
Data analysis: MAL.
Data interpretation: MAL, JAS, FB.
Software writing: MAL.
Manuscript drafting: MAL, JAS, FB.
All authors have revised the manuscript and approved its submission.

\section*{Acknowledgements}
\label{sv-gwas-acknowledgwments}

We would like to thank Yoav Voichek for help regarding \textit{k}-mer GWAS
analysis.  We also thank Brian Boyle and Martine Jean for valuable comments
regarding the methodology. We are thankful to the Digital Research Alliance of
Canada for using their high-performance computing servers.

\bibliographystyle{genome_research.bst}
\bibliography{references.bib}

### 2023-02-06 additional_files/manuscript.tex
Add main text contents to the repository; manuscript.tex file includes both manuscript and supplemental data into one pdf

\documentclass[12pt]{report}

\usepackage[utf8]{inputenc} % declare encoding as utf8
\usepackage{graphicx} % to enable \includegraphics
\usepackage{flafter} % to make sure floats appear after their position in text
\usepackage{pdflscape} % allow some pages to be in landscape orientation
\usepackage{layout} % for debugging purposes, looking at page layout
\usepackage[margin=1in]{geometry} % leaves more space for text body
\usepackage[hidelinks]{hyperref} % hyperlinks from list of figures/tables
\usepackage[labelfont=bf]{caption} % format captions and make links go to figure instead of caption
\usepackage{csvsimple} % read and display csv files as tables
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables
\usepackage{textcomp} % makes the \textdegree command available
\usepackage{tocloft} % formatting the lists of tables and figures
\usepackage{xcolor} % enable text coloring
\usepackage{calc} % makes \widthof available to set the width of the last column
\usepackage{array} % using command \arraybackslash to restore the definition of \\ after \centering
\usepackage{longtable} % allow tables to span more than one page
\usepackage{threeparttablex} % threeparttable for longtable
\usepackage{capt-of} % put caption on next page
\usepackage{ccaption}
\usepackage{listings}
\usepackage{datatool} % allows fetching key-value pairs from a text file
\usepackage[sectionbib]{natbib}
\usepackage{chapterbib}

% Remove indent for new paragraphs
\setlength{\parindent}{0pt}

% Setting the separator to use with datatool
\DTLsetseparator{ = }

% Formatting the table of contents/subsection
\renewcommand{\contentsname}{Supplemental methods}

% Formatting the list of tables
\renewcommand{\listtablename}{Supplemental tables} % Write "Supplemental tables" instead of "List of tables"
\renewcommand\cfttabpresnum{Table } % Write "Table" before table number in list of tables
\setlength{\cfttabnumwidth}{6em} % Spacing before captions
\setlength{\cftbeforefigskip}{1ex} % Vertical spacing between entries

% Formatting the list of figures
\renewcommand{\listfigurename}{Supplemental figures} % Write "Supplemental figures" instead of "List of figures"
\renewcommand\cftfigpresnum{Figure } % Write "Figure" before fig number in list of figures
\setlength{\cftfignumwidth}{6em} % Spacing before captions
\setlength{\cftbeforetabskip}{1ex} % Vertical spacing between entries

\graphicspath{ {../figures/} } % tell Latex to look for figures in figures/

% Format figure and table numbers to follow S1, S2, ...
\renewcommand\thefigure{S\arabic{figure}} 
\renewcommand\thetable{S\arabic{table}}

% Defining new environments as I will always want the figures centered
\newenvironment{cfigure}
	{\begin{figure} \centering}
	{\end{figure}}

\newenvironment{lsfigure}
	{\begin{landscape} \begin{figure} \centering}
	{\end{figure} \end{landscape}}

\setcounter{totalnumber}{1} % Allow a maximum of one float (table/figure) per page

\setlength{\parskip}{2ex}

% Defining a command to generate a table of phenotypic data for a single gwas
\newcommand{\gwastableone}[2]{
\begin{table}
	\caption[Phenotypic data used for the GWAS analysis of #2]%
	{Phenotype frequency and numerical coding of phenotypic data used for the GWAS analysis of #2.
	``NA'' indicates a missing value in the source data.
	Dashes indicate that this phenotype was not used in the GWAS analysis.}
	\label{#1_gwas_table}

	\centering
	\begin{threeparttable}\small

		\csvreader[head to column names,
		/csv/separator=semicolon,
		tabular = llc,
		table head = \toprule Phenotype & Frequency & GWAS \\ \midrule,
		table foot = \bottomrule]%
		{../tables/#1_gwas_table.csv}%
		{}%
		{\Value & \Frequency & \GWASa}

	\end{threeparttable}

\end{table}
\clearpage
}

% Defining a command to generate a table of phenotypic data for two gwas
\newcommand{\gwastabletwo}[2]{
\begin{table}
	\caption[Phenotypic data used for the GWAS analyses of #2]%
	{Phenotype frequency and numerical coding of phenotypic data used in two GWAS analyses of #2.
	``NA'' indicates a missing value in the source data.
	Dashes indicate that this phenotype was not used in the GWAS analysis.}
	\label{#1_gwas_table}

	\centering
	\begin{threeparttable}\small

		\csvreader[head to column names,
		/csv/separator=semicolon,
		tabular = llcc,
		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2\\ \midrule,
		table foot = \bottomrule]%
		{../tables/#1_gwas_table.csv}%
		{}%
		{\Value & \Frequency & \GWASa & \GWASb}

	\end{threeparttable}

\end{table}
\clearpage
}

% Defining a command to generate a table of phenotypic data for three gwas
\newcommand{\gwastablethree}[2]{
\begin{table}
	\caption[Phenotypic data used for the GWAS analyses of #2]%
	{Phenotype frequency and numerical coding of phenotypic data used in three GWAS analyses of #2.
	``NA'' indicates a missing value in the source data.
	Dashes indicate that this phenotype was not used in the GWAS analysis.}
	\label{#1_gwas_table}

	\centering
	\begin{threeparttable}\small

		\csvreader[head to column names,
		/csv/separator=semicolon,
		tabular = llccc,
		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2 & GWAS \# 3\\ \midrule,
		table foot = \bottomrule]%
		{../tables/#1_gwas_table.csv}%
		{}%
		{\Value & \Frequency & \GWASa & \GWASb & \GWASc}

	\end{threeparttable}

\end{table}
\clearpage
}

% Defining a command for manhattan plot figures
\newcommand{\manhattanplot}[3]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_manhattan}
	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
	(a) Platypus (SNPs and indels), (b) vg (SVs), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
	The x-axis shows the position along the reference assembly version 4 of Williams82.
	Each point shows the $-\log_{10} (p)$ associated with a particular marker or \emph{k}-mer.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	Vertical dotted lines indicate the position of signals associated with the trait.
	Documented loci are colored according to whether they were found (green) or not (red) by a particular approach.
	The ``Gm'' prefix has been left out of chromosome names for simplicity.
	}
	\label{#1_manhattan}
\end{cfigure}

\clearpage}

% Defining a new command for signal plot figures
\newcommand{\signalplot}[4]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_signal}
	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
	Panel (a) shows gene models over the genomic interval.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	#4
	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph and vg) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	}
	\label{#1_signal}
\end{cfigure}

\clearpage}

% Defining a new command for gene plot figures
\newcommand{\geneplot}[5]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_gene}
	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using four genotype datasets:
	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
	#5
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	Panel (a) shows gene models over the plotting interval.
	Exons are represented by rectangles whereas introns are represented by horizontal lines.
	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
	}
	\label{#1_gene}
\end{cfigure}

\clearpage}

% Defining a new command for k-mer plot figures
\newcommand{\kmerplot}[5]{
	\begin{cfigure}
		\includegraphics[width = 5.8in]{#1_kmers}
		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
		{Identification of #5 underlying significant \emph{k}-mers at the #2 gene associated with the #3 locus for #4.
		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
		The red rectangle identifies the region that is zoomed-in in panel (b).
		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
		Individual nucleotides are colored according to the $-\log_{10}(p)$ of the most significant \emph{k}-mer overlapping them.
		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes correspond to those shown in panel (b).}
		\label{#1_kmers}
	\end{cfigure}

\clearpage}

% Defining a new command for LD plot figures
\newcommand{\ldplot}[3]{
	\begin{cfigure}
		\includegraphics[width = 5.8in]{#1_ld}
		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
		\emph{k}-mers are sorted along the y-axis according to their putative position along the reference assembly version 4 of Williams82,
		as identified by ``Gm'' chromosome labels.
		Sequences that lack a ``Gm'' label (top of the y-axis) represent unanchored scaffolds.
		\emph{k}-mers are represented in the same order along the x- and y-axis.
		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p)$ of each \emph{k}-mer.}
		\label{#1_ld}
	\end{cfigure}

\clearpage}

% END OF PREAMBLE

\begin{document}

\include{main_text}
\include{additional_file_1}

\end{document}

### 2023-02-06 additional_files/additional_file_1.tex
Adjust the format of the manuscript + supplemental data to fit an article manuscript format

@@
-\chapter{Supplemental data}
+\chapter*{Supplemental data}
 \label{supdata}
+
+% Format figure and table numbers to follow S1, S2, ...
+\renewcommand\thefigure{S\arabic{figure}} 
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\renewcommand\thetable{S\arabic{table}}
 
@@
 % Detailed methods for some aspects of the paper
-\section{Detailed methods}
+\section*{Detailed methods}
 \label{annexe-sv-gwas-detailed-methods}
 
-\subsection{Discovery of structural variation from Illumina data}
+\subsection*{Discovery of structural variation from Illumina data}
 \label{sv-gwas-sv-discovery-illumina}
@@
 
-\subsection{Discovery of structural variation from Oxford Nanopore sequencing data}
+\subsection*{Discovery of structural variation from Oxford Nanopore sequencing data}
 \label{sv-gwas-sv-discovery-ont}
@@
 
-\subsection{Discovery of structural variation from high-quality assemblies}
+\subsection*{Discovery of structural variation from high-quality assemblies}
 \label{sv-gwas-sv-discovery-assembly}
@@
 
-\subsection{Genotyping structural variation using Paragraph}
+\subsection*{Genotyping structural variation using Paragraph}
 \label{sv-gwas-sv-genotyping-paragraph}
@@
 
-\subsection{Genotyping structural variation using the vg toolkit}
+\subsection*{Genotyping structural variation using the vg toolkit}
 \label{sv-gwas-sv-genotyping-vg}
@@
 
-\subsection{Comparison of WGS and SoySNP50K genotypes}
+\subsection*{Comparison of WGS and SoySNP50K genotypes}
 \label{annexe-sv-gwas-soysnp50k}
@@
 % Detailed results that are not meant to fit the main paper
-\section{Detailed results}
+\section*{Detailed results}
 \label{annexe-sv-gwas-detailed-results}
@@
 
-\subsection{Flower color}
+\subsection*{Flower color}
 \label{sv-gwas-flower-color}
@@
 
-\subsection{Pubescence color}
+\subsection*{Pubescence color}
 \label{sv-gwas-pubescence-color}
@@
 
-\subsection{Seed coat color}
+\subsection*{Seed coat color}
 \label{sv-gwas-seed-coat-color}
@@
 
-\subsection{Stem termination type}
+\subsection*{Stem termination type}
 \label{sv-gwas-stem-termination}
@@
 
-\subsection{Hilum color}
+\subsection*{Hilum color}
 \label{sv-gwas-hilum-color}
@@
 
-\subsection{Pod color}
+\subsection*{Pod color}
 \label{sv-gwas-pod-color}
@@
 
-\subsection{Pubescence form}
+\subsection*{Pubescence form}
 \label{sv-gwas-pubesence-form}
@@
 
-\subsection{Pubescence density}
+\subsection*{Pubescence density}
 \label{sv-gwas-pubescence-density}
@@
 
-\subsection{Seed coat luster}
+\subsection*{Seed coat luster}
 \label{sv-gwas-seed-coat-luster}
@@
 
-\subsection{Maturity group}
+\subsection*{Maturity group}
 \label{sv-gwas-maturity-group}
@@
 
-\subsection{Resistance to \emph{Phytophthora sojae}}
+\subsection*{Resistance to \emph{Phytophthora sojae}}
 \label{sv-gwas-resistance}
@@
 
-\subsection{Seed oil content}
+\subsection*{Seed oil content}
 \label{sv-gwas-seed-oil-content}
@@
 
-\subsection{Seed protein content}
+\subsection*{Seed protein content}
 \label{sv-gwas-seed-protein-content}
@@
 % Supplemental tables
-\section{Supplemental tables}
+\section*{Supplemental tables}
 \label{annexe-sv-gwas-tables}
@@
 % Supplemental figures
-\section{Supplemental figures}
+\section*{Supplemental figures}
 \label{annexe-sv-gwas-figures}

### 2023-02-06 additional_files/main_text.tex
Adjust the format of the manuscript + supplemental data to fit an article manuscript format

@@
-\chapter{\textit{k}-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean}
+\chapter*{\textit{k}-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean}
 \label{kmer-gwas}
@@
 
-\section{Abstract}
+\section*{Abstract}
 
@@
 
-\section{Introduction}
+\section*{Introduction}
 \label{sv-gwas-introduction}
@@
 
-\section{Methods}
+\section*{Methods}
 \label{sv-gwas-methods}
 
-\subsection{Sample selection and processing of sequencing data}
+\subsection*{Sample selection and processing of sequencing data}
 \label{sv-gwas-sequencing-data}
@@
 
-\subsection{Discovery and genotyping of SVs}
+\subsection*{Discovery and genotyping of SVs}
 \label{sv-gwas-sv-discovery-genotyping}
@@
 genotype variants.  Detailed methods regarding the SV discovery and genotyping
-steps can be found in section \ref{annexe-sv-gwas-detailed-methods} of Appendix
-\ref{annexe-sv-gwas}.
-
-\subsection{Discovery and genotyping of SNPs and indels}
+steps can be found in Supplemental data.
+
+\subsection*{Discovery and genotyping of SNPs and indels}
 \label{sv-gwas-snp-discovery}
@@
 The $\sim$7M SNPs and indels kept until this point were included in
-the variation graph used for SV genotyping with the vg toolkit (see section
-\ref{sv-gwas-sv-genotyping-vg} in Appendix \ref{annexe-sv-gwas}).
+the variation graph used for SV genotyping with the vg toolkit (see
+Supplemental data).
 
@@
 
-\subsection{Computing the presence/absence table of \emph{k}-mers}
+\subsection*{Computing the presence/absence table of \emph{k}-mers}
 \label{sv-gwas-kmer-pav}
@@
 
-\subsection{Phenotypic data}
+\subsection*{Phenotypic data}
 \label{sv-gwas-phenotypic-data}
@@
 accessions. We compared the SNP genotypes derived from the SoySNP50K array
-and our WGS data (see section \ref{annexe-sv-gwas-soysnp50k} in Appendix
-\ref{annexe-sv-gwas} for detailed methods) at over 32,000 SNPs, and
+and our WGS data (see Supplemental data
+for detailed methods) at over 32,000 SNPs, and
 identified 24 samples with < 90\% concordance between the two genotype
@@
 
-\subsection{GWAS analyses}
+\subsection*{GWAS analyses}
 \label{sv-gwas-gwas-methods}
@@
 
-\subsection{Analysis of significant \emph{k}-mers}
+\subsection*{Analysis of significant \emph{k}-mers}
 \label{sv-gwas-kmers-analysis-methods}
@@
 
-\subsection{Downstream analyses of significant association signals}
+\subsection*{Downstream analyses of significant association signals}
 \label{sv-gwas-methods-downstream}
@@
 version 1 to version 4 by using the same method that was used for converting
-SoySNP50K positions (see section \ref{annexe-sv-gwas-soysnp50k}).  The
+SoySNP50K positions (see Supplemental data).  The
 boundaries of the signals thus translated to genome assembly version 4 were
@@
 
-\subsection{Linking significant \emph{k}-mers to variants}
+\subsection*{Linking significant \emph{k}-mers to variants}
 \label{sv-gwas-methods-linking-kmers}
@@
 
-\subsection{Computation of LD between \emph{k}-mers}
+\subsection*{Computation of LD between \emph{k}-mers}
 \label{sv-gwas-ld-kmers}
@@
 
-\subsection{Software used}
+\subsection*{Software used}
 \label{sv-gwas-software-used}
@@
 
-\section{Results and discussion}
+\section*{Results and discussion}
 \label{sv-gwas-results-discussion}
@@
 most noteworthy results. Readers interested in seeing detailed results on a
-per-trait basis are referred to section \ref{annexe-sv-gwas-detailed-results}
-in Appendix \ref{annexe-sv-gwas}. In particular, table \ref{signals_table}
+per-trait basis are referred to
+Supplemental data. In particular, table \ref{signals_table}
 provides a summary of the \textit{p}-values obtained by each approach at each
@@
 
-\subsection{Loci with known genes or causal mutations}
+\subsection*{Loci with known genes or causal mutations}
 \label{sv-gwas-results-cloned-genes}
@@
 			\item[b] vg and Paragraph genotype SVs
-			\item[c] Locus not discussed in main text. See section \ref{annexe-sv-gwas-detailed-results} in Appendix \ref{annexe-sv-gwas} for detailed results.
+			\item[c] Locus not discussed in main text. See Supplemental data for detailed results.
 			\item[]- The approach detected no signal overlapping the gene.
@@
 
-\subsubsection{Flower color -- \textit{W1} locus}
+\subsubsection*{Flower color -- \textit{W1} locus}
 \label{sv-gwas-main-results-flower-color-l1}
@@
 concordance between genotype calls made from WGS data and the SoySNP50K array
-for the whole population (see section \ref{sv-gwas-phenotypic-data} in the Methods) and
+for the whole population (see the Methods section) and
 excluded 24 samples for which the concordance rate was below 90\%.
@@
 
-\subsubsection{Pubescence color -- \textit{T} locus}
+\subsubsection*{Pubescence color -- \textit{T} locus}
 \label{sv-gwas-main-results-pubescence-color-t}
@@
 
-\subsubsection{Pubescence color -- \textit{Td} locus}
+\subsubsection*{Pubescence color -- \textit{Td} locus}
 \label{sv-gwas-main-results-pubescence-color-td}
@@
 
-\subsubsection{Seed coat color - \textit{I} locus}
+\subsubsection*{Seed coat color - \textit{I} locus}
 \label{sv-gwas-main-results-seed-coat-color-i}
@@
 
-\subsubsection{Seed coat color - \textit{G} locus}
+\subsubsection*{Seed coat color - \textit{G} locus}
 \label{sv-gwas-main-results-seed-coat-color-g}
@@
 % Stem termination - Dt1 locus
-\subsubsection{Stem termination type - \textit{Dt1} locus}
+\subsubsection*{Stem termination type - \textit{Dt1} locus}
 \label{sv-gwas-main-results-stem-termination-dt1}
@@
 % Hilum color - T locus
-\subsubsection{Hilum color -- \textit{T} locus}
+\subsubsection*{Hilum color -- \textit{T} locus}
 \label{sv-gwas-main-results-hilum-color-t}
@@
 % Hilum color - I locus
-\subsubsection{Hilum color -- \textit{I} locus}
+\subsubsection*{Hilum color -- \textit{I} locus}
 \label{sv-gwas-main-results-hilum-color-i}
@@
 % Hilum color - R locus
-\subsubsection{Hilum color -- \textit{R} locus}
+\subsubsection*{Hilum color -- \textit{R} locus}
 \label{sv-gwas-main-results-hilum-color-r}
@@
 % Pubescence density - Ps locus
-\subsubsection{Pubescence density -- Ps locus}
+\subsubsection*{Pubescence density -- Ps locus}
 \label{sv-gwas-main-results-pubescence-density-ps}
@@
 % Seed coat luster - B locus
-\subsubsection{Seed coat luster -- B locus}
+\subsubsection*{Seed coat luster -- B locus}
 \label{sv-gwas-main-results-seed-coat-luster-b}
@@
 % Maturity group loci
-\subsubsection{Maturity}
+\subsubsection*{Maturity}
 \label{sv-gwas-main-results-maturity}
@@
 % Seed oil and protein
-\subsubsection{Seed oil and protein content}
+\subsubsection*{Seed oil and protein content}
 \label{sv-gwas-main-results-oil-protein}
@@
 
-\subsection{Analysis of loci with unknown causal genes}
+\subsection*{Analysis of loci with unknown causal genes}
 \label{sv-gwas-results}
@@
 % Pod color - L1 and L2 loci
-\subsubsection{Pod color -- \textit{L1} and \textit{L2} loci}
+\subsubsection*{Pod color -- \textit{L1} and \textit{L2} loci}
 \label{sv-gwas-main-results-pod-color-l1-l2}
@@
 % Pubescence form - Pa1 and Pa2 loci
-\subsubsection{Pubescence form -- \textit{Pa1} and \textit{Pa2} loci}
+\subsubsection*{Pubescence form -- \textit{Pa1} and \textit{Pa2} loci}
 \label{sv-gwas-main-results-pubescence-form-pa1-pa2}
@@
 % Resistance to Phytophthora sojae
-\subsubsection{Resistance to \textit{P. sojae}}
+\subsubsection*{Resistance to \textit{P. sojae}}
 \label{sv-gwas-main-results-psojae}
@@
 % Stem terminination and pod color - novel loci
-\subsubsection{Novel loci for stem termination and pod color}
+\subsubsection*{Novel loci for stem termination and pod color}
 \label{sv-gwas-main-results-novel-loci}
@@
 
-\subsection{Potential of \emph{k}-mer-based GWAS}
+\subsection*{Potential of \emph{k}-mer-based GWAS}
 \label{sv-gwas-potential-kmers}
@@
 
-\subsection{Limitations and challenges of \emph{k}-mer-based GWAS}
+\subsection*{Limitations and challenges of \emph{k}-mer-based GWAS}
 \label{limitations-of-kmers-GWAS}
@@
 
-\subsection{Use of SV genotypes for GWAS}
+\subsection*{Use of SV genotypes for GWAS}
 \label{use-of-svs-for-gwas}
@@
 
-\section{Conclusion}
+\section*{Conclusion}
 

### 2023-02-06 additional_files/manuscript.tex
Adjust the format of the manuscript + supplemental data to fit an article manuscript format

@@
 \usepackage[utf8]{inputenc} % declare encoding as utf8
+\usepackage[T1]{fontenc} % declare encoding as utf8
 \usepackage{graphicx} % to enable \includegraphics
@@
 \usepackage{chapterbib}
+\usepackage{setspace}
 
@@
 
+% Adjust line spacing to 1.5
+\onehalfspacing
+
 % Setting the separator to use with datatool
@@
 
-% Formatting the table of contents/subsection
-\renewcommand{\contentsname}{Supplemental methods}
-
-% Formatting the list of tables
-\renewcommand{\listtablename}{Supplemental tables} % Write "Supplemental tables" instead of "List of tables"
-\renewcommand\cfttabpresnum{Table } % Write "Table" before table number in list of tables
-\setlength{\cfttabnumwidth}{6em} % Spacing before captions
-\setlength{\cftbeforefigskip}{1ex} % Vertical spacing between entries
-
-% Formatting the list of figures
-\renewcommand{\listfigurename}{Supplemental figures} % Write "Supplemental figures" instead of "List of figures"
-\renewcommand\cftfigpresnum{Figure } % Write "Figure" before fig number in list of figures
-\setlength{\cftfignumwidth}{6em} % Spacing before captions
-\setlength{\cftbeforetabskip}{1ex} % Vertical spacing between entries
-
 \graphicspath{ {../figures/} } % tell Latex to look for figures in figures/
-
-% Format figure and table numbers to follow S1, S2, ...
-\renewcommand\thefigure{S\arabic{figure}} 
-\renewcommand\thetable{S\arabic{table}}
 

### 2023-02-06 additional_files/main_text.tex
Integrate suggestions by JA Sibbesen into the Abstract and Introduction

@@
 that detect associations between genotype and phenotype at genome scale. Despite
-their power, GWAS often fail to pinpoint the causal variant or the gene
-controlling a trait at a given locus.  Assessing genetic variants other
-than single-nucleotide polymorphisms (SNPs) may alleviate this problem
-by including functionally relevant structural variants (SVs) that are
-typically omitted from GWAS.  In this study, we assessed the potential
-of SV- and \emph{k}-mer-based GWAS by applying these methods to 13
-traits on which SNP-based GWAS had been performed previously and for
-which the causal genes and variants are known in some cases. We found
-that \emph{k}-mer-based GWAS had an impressive ability to pinpoint
+their power, GWAS frequently fail to pinpoint the causal variant or the gene
+controlling a trait at a given locus in crop species. Assessing genetic variants beyond
+single-nucleotide polymorphisms (SNPs) may alleviate this problem,
+for example by including structural variants (SVs) that may have a greater functional impact.
+In this study, we tested the potential
+of SV- and \emph{k}-mer-based GWAS in soybean by applying these methods to 13
+traits on which SNP-based GWAS had been performed previously.
+We assessed the performance of each GWAS approach based on results at loci for which
+the causal genes or variants were known from previous genetic studies. We found
+that \emph{k}-mer-based GWAS was the most versatile approach and the best at pinpointing
 causal variants or candidate genes based on the most significantly
-associated \emph{k}-mers. The ability of SNP/indel-based and SV-based
-GWAS to similarly pinpoint known variants or genes was almost
-systematically reduced compared to \emph{k}-mers.  We suggest that
-SV-based GWAS may not generally be worth the investment in time and
-computational resources required to discover and genotype such variants
-given the results obtained with \emph{k}-mer-based GWAS, which requires
-comparatively lower resources.  Moreover, analyses based on
-\emph{k}-mer GWAS identified promising candidate genes for loci related
+associated \emph{k}-mers. Given the traits that we analyzed, SV-based GWAS did not add any value
+compared to \textit{k}-mer-based GWAS and may not generally be worth the investment in time and
+computational resources required to discover and genotype such variants.
+Moreover, \textit{k}-mer-based analyses
+identified promising candidate genes for loci related
 to pod color, pubescence form, and resistance to the oomycete
-\textit{Phytophthora sojae}.  Despite these encouraging results,
+\textit{Phytophthora sojae}. Despite promising results,
 significant challenges remain regarding the downstream analysis of
-\emph{k}-mer GWAS, notably in linking significant \emph{k}-mers to
+\emph{k}-mer-based GWAS, notably in associating significant \emph{k}-mers with
 sequence variation and ensuring that all variation types are adequately
-detected from \emph{k}-mer occurrence patterns.  Our results suggest
+detected from \emph{k}-mer occurrence patterns. Together, our results suggest
 that coupling \emph{k}-mer- and SNP/indel-based GWAS may represent a
@@
 statistical associations between phenotypic observations for a trait of
-interest and the genotypes of markers or variants distributed throughout the
-genome \citep{tam2019}.  Such studies have been used for over 15 years
-\citep{visscher2012} and have enabled significant advances in our understanding
+interest and the genotypes of variants distributed throughout the
+genome \citep{tam2019}. This approach has been used for over 15 years
+\citep{visscher2012} and has enabled significant advances in our understanding
 of the genetic architecture of traits involved in human health
-\citep[e.g.][]{wellcome2007gwas} and crop \citep{tibbs2021} or animal
+\citep[e.g.][]{wellcome2007gwas} and crop \citep{tibbs2021} and animal
 \citep{ma2019} production, among others.
@@
 Despite their power and convenience for finding significantly associated loci,
-GWAS analyses often fail to pinpoint the gene(s) associated with a trait and
+GWAS analyses frequently fail to pinpoint the gene(s) associated with a trait and
 the causal variant(s) involved \citep{tam2019}. One possible reason for this
-failure is that most GWAS do not include the causal variant itself among the
-genotypes assessed \citep{tibbs2021}. This could be either because the analysis
+is that many GWAS analyses do not include the causal variant itself in their variant dataset
+\citep{tibbs2021}. This could be either because the analysis
 is restricted to a particular type of variant, such as single-nucleotide
@@
 association panel is limited by the technology being used
-\citep[e.g.][]{bandillo2015, sonah2015}.  Despite these limitations, variants
-located near the gene controlling the trait should appear statistically
+\citep[e.g.][]{bandillo2015, sonah2015}. Despite this limitation, variants
+located near the gene controlling the trait often do appear statistically
 associated with the phenotype through linkage disequilibrium (LD) with the
@@
 that a given gene controls the trait \citep[e.g.][]{wang2018, liu2020ps}.
-Other limitations that may preclude the identification of causal variants from
-the output of GWAS include incomplete gene annotation, genetic divergence
-between the accessions and the reference genome used, or the existence of
-several different causal variants at a locus.
-
-One way of overcoming those limitations would be to broaden the set of genetic
-variants used in GWAS by also including structural variants (SVs) such as
-deletions, insertions, inversions, or any other type of rearrangement that
-exceeds a certain size \cite[typically 50 bp; ][]{ho2020}.  SVs are known to
+
+The inclusion of structural variants (SVs) as genotype data in GWAS may partly overcome
+this issue by providing a set of variants with a potentially high functional impact.
+SVs include any variant involving a difference of at least 50 nucleotides, such as
+deletions, insertions, or inversions \citep{ho2020}. Such variants are known to
 have large phenotypic impacts by disrupting coding or regulatory sequences
-\citep{marroni2014} and have been identified as causal variants in traits such
+\citep{marroni2014}, and have been identified as causal variants in traits such
 as resistance to soybean cyst nematode \citep{cook2012}, aluminum tolerance in
 wheat \citep{maron2013}, and branching in maize \citep{studer2011}.  SVs have
-already been used in GWAS in studies that have shown promising results
+already been used in GWAS in studies with promising results
 \citep[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
-their inclusion in GWAS is not systematic.
-
-Despite their high potential, SVs are difficult to discover and genotype with
+their inclusion in GWAS is the exception rather than the norm.
+
+Despite their phenotypic impact, SVs are still largely understudied compared to SNPs
+because of the difficulty of discovering and genotyping them with
 accuracy. Comprehensive detection of SVs requires a combination of several
@@
 improvements in computational approaches for genotyping SVs
-\citep[e.g.][]{siren2021, ebler2022}, such approaches remain computationally
-expensive compared to SNP genotyping workflows. Therefore, discovering and
-genotyping SVs at a scale required for GWAS may prove daunting even when
-sufficient computational resources and skilled analysts are available.
-
-In this context, alternative approaches that allow to assess a large spectrum
+\citep[e.g.][]{siren2021, ebler2022}, such approaches remain relatively inaccurate
+compared to SNP genotyping workflows.
+
+Alternative approaches that allow one to assess a large spectrum
 of variants in GWAS without genotyping SVs at population scale could be useful.
@@
 represent such an alternative method. This approach relies on the presence or
-absence of \emph{k}-mers observed in sequence reads as genotypes for use in
+absence of \emph{k}-mers observed in sequence reads instead of genotypes for use in
 GWAS. Once a presence/absence table of \emph{k}-mers has been generated for the
@@
 traits \citep{bandillo2015} that have been previously studied using a SNP array
-developed for soybean \citep{song2013} as well as a quantitative trait
+developed for soybean \citep{song2013}. In addition, we analyzed a quantitative trait
 (horizontal resistance to \emph{Phytophthora sojae}) that has been previously
-studied using whole-genome sequencing (WGS) data \citep{deronne2022}.  The fact
-that the underlying genes and causal variants are already known for several of
-these traits provides a good test case for the use of newer GWAS methods, as
-the expected result is already known. Our objectives as part of this work were
-to:
+studied using whole-genome sequencing (WGS) data \citep{deronne2022}.
+Since the underlying genes and causal variants are already known for several of
+these traits, they provide a good test case for the performance of GWAS methods.
+The objectives of our study were to:
 
 \begin{enumerate}
-	\item Assess the potential of SV- and \emph{k}-mer-based GWAS to
+	\item Assess the potential of SV- and \emph{k}-mer-based GWAS approaches to
 		pinpoint genes and causal variants associated with loci whose

### 2023-02-13 additional_files/additional_file_1.tex
Integrate suggestions by JA Sibbesen into the Methods

@@
 Additional file 2.
+
+\subsection*{Analysis of extremely low \textit{p}-values}
+\label{annexe-sv-gwas-pvalues}
+
+For one trait (flower color), the significance of some \emph{k}-mers was so
+high that their \emph{p}-values were numerically equal to 0. In order to
+compute $-\log_{10} (p)$ for visualization in Manhattan plots in these cases,
+we set the \emph{p}-value to the smallest numerical value that could be
+represented using R on our machine (2.225074e-308). In this particular case,
+the values represented on the Manhattan plots represent an upper bound on the
+\emph{p}-value and not an exact value.
 

### 2023-02-13 additional_files/main_text.tex
Integrate suggestions by JA Sibbesen into the Methods

@@
 sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
-Illumina paired-end sequencing data for those 389 accessions among the ones
-analyzed by \cite{bayer2021} and downloaded them from the SRA.  We obtained
-reads in FASTQ format using the sratoolkit command fastq-dump v. 2.9.6
+Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
+and extracted paired-end reads using the fastq-dump command v. 2.9.6
 (\citeauthor[][\url{https://github.com/ncbi/sra-tools}]{sratoolkit}) with the
-option \texttt{-\-split-3} as some of the runs contained reads that were
-unpaired. Reads that were not paired in the raw sequencing data were not used
+option \texttt{-\-split-3}.
+Reads that were not paired in the raw sequencing data were not used
 for downstream analyses.  The runs downloaded from the SRA and associated
@@
 
-We filtered the sequencing data for quality and the presence of sequencing
+We first filtered the sequencing data for quality and the presence of sequencing
 adapters using bbduk from BBtools v. 38.25
-(\citeauthor[][\url{https://sourceforge.net/projects/bbmap/}]{bbtools}). We
-then aligned reads to the reference assembly version 4 of \emph{G. max}
-cultivar Williams82 \citep{valliyodan2019} to which we concatenated chloroplast
+(\citeauthor[][\url{https://sourceforge.net/projects/bbmap/}]{bbtools}). Reads
+were then mapped to the reference assembly version 4 of \emph{G. max}
+cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
 and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
-bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to align the
+bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
 reads in paired-end mode. Reads that were left unpaired following adapter- and
-quality-trimming were aligned separately in single-end mode using bwa-mem and
-merged with the reads aligned in paired-end mode using samtools \texttt{merge}
+quality-trimming were aligned separately in single-end mode with bwa-mem and
+merged with mapped paired-end reads using samtools \texttt{merge}
 v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
@@
 Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
-The \emph{discovery} step consisted in methods leading to a set of candidate SVs,
-whereas the \emph{genotyping} step consisted in determining the genotype
-of the candidate SVs for all samples in the population. We used three different
-sources of data in the SV discovery step to generate a set of candidate variants:
+The \emph{discovery} step identified a set of candidate SVs using various methods,
+whereas the \emph{genotyping} step determined the genotype
+of the candidate SVs for all accessions in the population from the mapped Illumina reads. We used three different
+approaches in the SV discovery step to generate a set of candidate variants:
 
 \begin{itemize}
-	\item We used Illumina WGS data of the 389 accessions mentioned above
-		to call SVs in the population following methods outlined
-		in \cite{lemay2022}. In brief, we combined information from
+	\item We used Illumina WGS data from the 389 accessions mentioned above
+		to call SVs in the population following a pipeline outlined
+		by \cite{lemay2022}. In brief, we combined information from
 		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
 		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
-		a set of candidate SVs from Illumina data.
+		a set of candidate SVs from the mapped Illumina data.
 	\item We used SVs discovered by \cite{lemay2022} from Oxford Nanopore
-		sequencing data of 17 Canadian soybean cultivars as a second
-		source of SVs.
+		sequencing data of 17 Canadian soybean cultivars.
 	\item We called SVs from 26 high-quality genome assemblies published by
@@
 
-We merged the SVs called using each of these approaches using SVmerge
-\citep{wong2010} and used the resulting set of candidate SVs for genotyping. We
-used two programs with different genotyping models for genotyping this set of
-SVs, Paragraph \citep{chen2019} and vg \citep{hickey2020}. In brief, Paragraph
-genotypes each SV individually whereas vg builds a variation graph from a set
-of candidate variants and then uses this graph to align sequence reads and
-genotype variants.  Detailed methods regarding the SV discovery and genotyping
-steps can be found in Supplemental data.
+The SVs called using each of these approaches were merged using SVmerge
+\citep{wong2010} and the resulting set of candidate SVs was used for
+genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
+from the mapped Illumina reads for all accessions. Detailed methods regarding
+the SV discovery and genotyping steps can be found in Supplemental data.
 
@@
 
-We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels on the
-mapped reads of the 389 cultivars. We initially filtered the 21.1 M SNPs and
-indels from the raw output of Platypus to keep only those with:
+We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
+mapped reads of the 389 accessions. The 21.1 M SNPs and
+indels from the raw output of Platypus were filtered to keep only those with:
 
@@
 	\item \texttt{FILTER} field set to \texttt{PASS}
-	\item minor allele frequency (MAF) $\geq$ 0.005
-	\item fraction of missing genotypes $\leq$ 0.8
+	\item minor allele frequency (MAF) $\geq$ 0.02
+	\item fraction of missing genotypes $\leq$ 0.5
 	\item heterozygosity rate $\leq$ 0.2.
@@
 
-The $\sim$7M SNPs and indels kept until this point were included in
-the variation graph used for SV genotyping with the vg toolkit (see
-Supplemental data).
-
-The dataset of SNPs and indels used for GWAS was further filtered to achieve a
-MAF $\geq$ 0.02 and a fraction of missing genotypes $\leq$ 0.5. Moreover, we
-pruned the dataset using the \texttt{-\-indep-pairwise} option of PLINK v.
-1.90b5.3 \citep{purcell2007} using a window of 1,000 markers with a step of 100
-markers and an $r^2$ LD threshold of 0.9. Markers located on unanchored
-scaffolds were also removed. The resulting dataset used for GWAS comprised
-924,630 SNPs and indels.
+Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
+PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
+step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
+implemented to reduce the computational requirements of the GWAS analysis by
+removing co-segregating markers. Markers located on unanchored scaffolds were
+also removed. The resulting dataset was used for GWAS and comprised 773,060
+SNPs and 151,570 indels.
 
@@
 We used the approach outlined by \cite{voichek2020} for \emph{k}-mer-based GWAS
-as well as their binary executables (v. 0.2-beta) and recommendations on the
-associated GitHub page (\url{https://github.com/voichek/kmersGWAS}) to generate
-a table of \emph{k}-mer presence/absence for use in GWAS.  The first step of
-the approach required counting the \emph{k}-mers of length 31 (31-mers) present
-in the trimmed FASTQ sequencing files of all samples using KMC3 v. 3.2.1
-\citep{kokot2017}. We counted \emph{k}-mers twice, the first time using their
+and their recommendations on the
+associated GitHub page (v. 0.2-beta, \url{https://github.com/voichek/kmersGWAS}) to generate
+a presence/absence \emph{k}-mer table for use in GWAS.  The first step of
+the approach involved counting all \emph{k}-mers of length 31 (31-mers) present
+in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
+\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
 canonized form (the first of either the observed \textit{k}-mer or its reverse
 complement in lexicographical order) and the second time using the observed
-\emph{k}-mers themselves (non-canonized form). After the information from both KMC3 runs was
-combined, we filtered the set of \emph{k}-mers to be used for the analysis by
-keeping only those seen in at least 5 samples and observed in both canonized
-and non-canonized form in at least 20\% of the samples in which they are found.
-We then generated a table indicating the presence/absence in all samples of the
-\emph{k}-mers that passed all filters. This table was used to compute a kinship
+\emph{k}-mers themselves (non-canonized form). The two sets of \textit{k}-mers
+were then combined and filtered by
+keeping only those seen in at least 5 accessions, and observed in both canonized
+and non-canonized form in at least 20\% of the accessions in which they are found.
+We then generated a table indicating the presence or absence of each
+\emph{k}-mer in all accessions. This table was used to compute a kinship
 table for use in GWAS and was also used directly as input genotypes for the
@@
 
-We analyzed ten qualitative traits studied by \cite{bandillo2017}: flower
+We analyzed ten qualitative traits previously studied by \cite{bandillo2017}: flower
 color, pubescence color, seed coat color, stem termination type, hilum color,
@@
 accessions did not match the GRIN database and were therefore not used for the
-analysis of qualitative traits. We recoded the phenotypes of qualitative traits
-to numerical values for GWAS analysis similarly to the numerical coding used by
+analysis of qualitative traits. The phenotypes of qualitative traits were recoded
+to numerical values for GWAS analysis following methods similar to those of
 \cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
@@
 We analyzed three quantitative traits in addition to the ten qualitative traits
-mentioned above.  We retrieved seed oil and protein content from the GRIN
+mentioned above. Seed oil and protein content were retrieved from the GRIN
 database. In cases where more than one value was listed for a given accession,
-we computed the average of those values and used it in GWAS analyses. We also
-used the horizontal resistance to the oomycete \emph{P. sojae} (corrected dry
-weight, CDW) assessed in a hydroponic assay \citep{deronne2022} as a third
-quantitative trait. Our analysis included 340 accessions whose phenotypic data
-was already published by \cite{deronne2022} plus 49 additional accessions for
+we computed the average of those values and used it for GWAS analyses. As a third quantitative trait, we
+used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
+weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
+Our analysis included 340 accessions whose phenotypic data
+was already published by \cite{deronne2022} as well as 49 additional accessions for
 which sequencing data was not yet available when their study was conducted.
-Because of ambiguity regarding the identity of one of the samples (HN019), the
-number of samples for which CDW data was available was 388 and this sample
-was also dropped from all other GWAS analyses. Similarly to seed oil and
-protein content, the CDW values for some samples for which more than one
+Because of ambiguity regarding the identity of one of the accessions (HN019), this accession
+was dropped from all GWAS analyses. Similarly to seed oil and
+protein content, the CDW values for some accessions for which more than one
 observation was available were averaged.
 
-Because of clear discrepancies between the observed haplotypes and the
-reported phenotypes for some simple traits such as flower color and seed coat
-color, we suspected possible errors in the identification of some of the
-accessions. We compared the SNP genotypes derived from the SoySNP50K array
-and our WGS data (see Supplemental data
-for detailed methods) at over 32,000 SNPs, and
-identified 24 samples with < 90\% concordance between the two genotype
-datasets. We removed them from the dataset in addition to another
-sample that exhibited an atypical GC content suggesting contamination.
-Following these filtering steps, 363 samples remained for GWAS analyses. The
-concordance between WGS data and SoySNP50K genotypes is included in
-Additional file 2 and averaged 98.3\% among retained lines while the
-phenotypic data used in this study can be found in Additional file 3.
+Because of clear discrepancies between the observed haplotypes and the reported
+phenotypes for some simple traits such as flower color and seed coat color, we
+suspected possible errors in the identity of some of the accessions. To
+investigate this, we compared the SNP genotypes derived from the SoySNP50K
+array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
+90\% concordance between the two genotype datasets (see Supplemental data for
+detailed methods). These samples were removed from the dataset in addition to another
+accession that exhibited an atypical GC content suggesting contamination.
+Following these filtering steps and the removal of HN019, 363 accessions
+remained for GWAS analyses. The concordance between WGS data and SoySNP50K
+genotypes is included in Additional file 2 and averaged 98.3\% among retained
+lines. The phenotypic data used in this study can be found in Additional file 3.
 
@@
 the SNP/indel and SV datasets described above. SVs and indels required some additional
-preprocessing steps detailed in \cite{lemay2022gwas} to make variant
-representation suitable for downstream analyses. Briefly, we recoded all
+preprocessing steps to make variant
+representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
 variants as SNPs prior to converting the VCF files to Hapmap diploid format
@@
 
-We ran GAPIT using an MLM model with 9 principal components and the VanRaden
+We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
 algorithm for computing the kinship matrix. We used an MLM model for
@@
 top \emph{p}-values under the null hypothesis. The fifth most significant of
-the 100 top \emph{p}-values was used as the significance threshold for
+the top 100 \emph{p}-values was used as the significance threshold for
 inference, i.e. markers with \emph{p}-values lower than this threshold were
@@
 
-We conducted GWAS based on \emph{k}-mers similarly to the method outlined by
-\cite{voichek2020} and used their binary executables
-(\url{https://github.com/voichek/kmersGWAS}) for doing so.  In brief, the
+We conducted GWAS based on \emph{k}-mers using the method developed by
+\cite{voichek2020}. In brief, the
 analysis was conducted in two steps. The first step used an approximate model
-and identified 1 million possibly significant \emph{k}-mers that were used in
-the second step with an exact model implemented in GEMMA \citep{zhou2012} to
-compute the \emph{p}-values associated with each of these \emph{k}-mers. We
+to identify 1 million potentially significant \emph{k}-mers. These were then used in
+the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
+in a \emph{p}-value for each \emph{k}-mer. We
 used the kinship computed from the \emph{k}-mers to correct for relatedness
-between accessions. The program implemented by \cite{voichek2020} automatically
-returned a list of \emph{k}-mers that passed a 5\% family-wise error-rate
-threshold determined by a randomization approach.  These \emph{k}-mers were
-considered significant for the purposes of downstream analyses.
+between accessions. The program automatically outputs a list of \emph{k}-mers
+that passed a 5\% family-wise error-rate threshold determined by a
+randomization approach, which we considered significant for the purposes of
+downstream analyses.
 
@@
 and a tailored analysis was therefore required to associate them with genomic
-coordinates. To do so, we queried the aligned reads of all samples for matches
+coordinates. To do so, we queried the reads (in BAM format) of all accessions for matches
 to significant \emph{k}-mers for a given trait. Because this is a
 computationally demanding operation (there were typically tens of thousands of
-significant \emph{k}-mers per trait), we developed a program called katcher
-(\url{https://github.com/malemay/katcher}) that is written in the C language
-and uses the htslib library \citep{bonfield2021} for efficient reading and
-writing of BAM files. katcher and its associated utilities allow the efficient
+significant \emph{k}-mers per trait), we developed a C program called katcher
+(\url{https://github.com/malemay/katcher}) that
+uses the htslib library \citep{bonfield2021} for efficient reading and
+writing of BAM files. katcher and its associated utilities allow for efficient
 retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
@@
 
-For one trait (flower color), the significance of some \emph{k}-mers was so
-high that their \emph{p}-values were numerically equal to 0. In order to
-compute $-\log_{10} (p)$ for visualization on Manhattan plots in these cases,
-we set the \emph{p}-value to the smallest numerical value that could be
-represented using R on our machine (2.225074e-308). In this particular case,
-the values represented on the Manhattan plots represent an upper bound on the
-\emph{p}-value and not an exact value.
-
-Once we retrieved all the reads containing significant \emph{k}-mers, we still
+Once we retrieved all the reads containing significant \emph{k}-mers, we next
 needed to gather the mapping information contained in those reads for the
-purposes of identifying contiguous signals and generating Manhattan plots. To
+purposes of identifying contiguous signals and comparing them to annotations. To
 simplify the amount of information to process in downstream analyses, we
 considered only the most significant \emph{k}-mer contained in a given read and
-only considered the position where this \emph{k}-mer was most often observed in
+the position where this \emph{k}-mer was most often observed in
 the dataset. Moreover, we limited the analysis to reads with a minimum mapping
 quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
-least one sample to filter out for spurious \emph{k}-mers that may originate
+least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
 from sequencing errors. All downstream analyses (Manhattan plots,
@@
 
-We used contiguous regions comprising significant associations, which we called
+We used contiguous regions of significant associations, which we called
 ``signals'', as a basis to compare our results with those obtained by previous
@@
 of those signals by grouping any significant markers or \emph{k}-mers that
-occurred within 250 kb of one another.  We did consider the extent of deletions
+occurred within 250 kb of one another.  We did consider the length of deletions
 or \emph{k}-mers in defining the distance between significant associations.
@@
 
-Because \cite{bandillo2015} and \cite{bandillo2017} used assembly version 1 of
-Williams82, we converted the genomic positions of their signals from assembly
-version 1 to version 4 by using the same method that was used for converting
-SoySNP50K positions (see Supplemental data).  The
-boundaries of the signals thus translated to genome assembly version 4 were
-used for comparison with the signals found in our analyses. For the analysis of
+Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
+Williams82, we converted the genomic coordinates of their signals to assembly version 4
+using the same method that we used for converting
+SoySNP50K positions (see Supplemental data). For the analysis of
 qualitative traits, we only compared our results to signals at named classical
@@
 
-We used the reads retrieved by katcher as a starting point for identifying the
-variants associated with a given set of \emph{k}-mers. For any such
-\emph{k}-mer set of interest, we used the QNAME field (in the BAM format) of
-the reads identified by katcher to extract all reads with a QNAME corresponding
-to a read containing significant \emph{k}-mers. We proceeded this way so as to
-retrieve the reads that are paired with a significant read even though that
-read may not have matched a significant \emph{k}-mer in the first place. Once
-all such reads were identified, we converted them to paired-end reads in FASTQ
-format using \texttt{samtools fastq} and used the reads for \emph{de novo}
-assembly with SPAdes v. 3.15.4 \citep{spades} with the \texttt{-\-careful}
-parameter as advised by \cite{voichek2020}. Assemblies were done individually
-for each sample and aligned to the Williams82 reference assembly using bwa mem.
-Assemblies targeting a region of interest for samples used in a given GWAS were
-aligned using multiple sequence alignment with the MAFFT program v. 7.475
-\citep{katoh2002} for visualization purposes. We only considered haplotypes
-occurring at least five times in the population in such analyses in order to
-leave out potential assembly errors. In a single case (\emph{Dt1} locus),
-assembly at the locus succeeded for a very low number of samples (26). In this
-case, we used consensus sequences from the reads aligned with bwa for comparing
-the haplotypes across samples.
+We used the reads retrieved by katcher and an assembly-based approach for
+identifying the variants associated with a given set of \emph{k}-mers. In
+addition to the reads themselves, we also retrieved their paired reads even
+though they may not have matched a significant \emph{k}-mer in the first place.
+These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
+\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
+individually for each accession and then aligned to the Williams82 reference
+assembly using bwa mem with default parameters.  Aligned assembled sequences
+that overlapped loci of interest were realigned using multiple sequence
+alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
+visualizing the haplotypes at these loci. The identification of variants
+supported by significant \textit{k}-mers was based entirely on visual analysis
+of haplotype alignments. We only considered haplotypes occurring at least five
+times in the population in such analyses in order to leave out potential
+assembly errors. In a single case (\emph{Dt1} locus), assembly at the locus
+succeeded for a very low number of accessions (26). In this case, we instead used
+consensus sequences from the reads aligned with bwa for comparing the
+haplotypes across accessions.
 
@@
 directly from the WGS data to those made using the SoySNP50K array, we observed
-that these two samples had concordance rates of only 62\% and 70\%. To
+that these two accessions had concordance rates of only 62\% and 70\%. To
 prevent such mismatches from affecting our GWAS analyses, we computed the
@@
 for the whole population (see the Methods section) and
-excluded 24 samples for which the concordance rate was below 90\%.
+excluded 24 accessions for which the concordance rate was below 90\%.
 
@@
 	(c) Estimated copy number (sequencing depth at position divided by
-	sample average sequencing depth) of accessions with normal and
+	accession average sequencing depth) of accessions with normal and
 	semi-sparse pubescence at the \textit{Ps} locus. Solid lines represent
 	the average copy number across 73 (semi-sparse pubescence) and 180
-	samples (normal pubescence) computed using a 100-bp sliding window.
+	accessions (normal pubescence) computed using a 100-bp sliding window.
 	Shaded areas outline the first and third quartiles. The start and end
 	of the CNV region can be clearly inferred from the divergence between
-	the copy number of samples with contrasting pubescence.
+	the copy number of accessions with contrasting pubescence.
 	(d) Histogram of the mapping location of 1,169 reads containing the
-	most significant \textit{k}-mer in samples with semi-sparse pubescence.
+	most significant \textit{k}-mer in accessions with semi-sparse pubescence.
 	Reads containing this \textit{k}-mer map to either end of the CNV
@@
 		from the Sequence Read Archive (SRA) (CSV 71 KB)
-	\item[Additional file 2.csv] Metadata on individual samples following mapping
+	\item[Additional file 2.csv] Metadata on individual accessions following mapping
 		and comparison to SoySNP50K data (CSV 36KB)

### 2023-02-20 additional_files/main_text.tex
Integrate suggestions by JA Sibbesen into the main Results section

@@
 
-We analyzed a total of 13 traits using four different GWAS approaches, one
-based on SNP and indel genotypes (Platypus), two based on SV genotypes (Paragraph and
-vg), and another based on the presence/absence of \textit{k}-mers.
-Because two or three different GWAS analyses were conducted on some traits, a
+We analyzed a total of 13 traits using three different GWAS approaches, one
+based on SNP and indel genotypes (Platypus), one based on SV genotypes (Paragraph),
+and one based on the presence/absence of \textit{k}-mers.
+Because two or three different GWAS analyses were conducted for some of the traits, a
 total of 22 GWAS analyses were performed using each of the four approaches.
 Given the large volume of data that this represents, we will focus here on the
-most noteworthy results. Readers interested in seeing detailed results on a
-per-trait basis are referred to
-Supplemental data. In particular, table \ref{signals_table}
+most noteworthy results. Readers interested in detailed results for loci not
+discussed here are referred to Supplemental data. In particular, table \ref{signals_table}
 provides a summary of the \textit{p}-values obtained by each approach at each
@@
 
-We were first interested in assessing the performance of various GWAS
+We first wanted to assess the performance of the various GWAS
 approaches by analyzing the results at loci for which the genes have
@@
 GWAS approaches detected a signal overlapping the gene at this locus (Figures
-\ref{flower_color_manhattan}, \ref{flower_color_W1_signal}) and the most
-significant variants/\textit{k}-mers of all approaches except Platypus identified
-the known causal variant (Figures \ref{flower-color-main-figure} and
-\ref{flower_color_W1_gene}).
+\ref{flower_color_manhattan}, \ref{flower_color_W1_signal}). The most
+significant SV and \textit{k}-mers corresponded to the known causal variant
+(Figures \ref{flower-color-main-figure}a-b and \ref{flower_color_W1_gene}).
+In particular, the assembly of the reads containing the most significant \textit{k}-mers
+and the alignment of the resulting haplotypes showed that these \textit{k}-mers
+directly tagged the causal variant (Figure \ref{flower-color-main-figure}c-e).
 As the catalogue of variants analyzed by Platypus contains only SNPs and small
@@
 associated with purple flowers were noted as having white flowers. Upon further
-analysis, we found that the most likely cause of this discrepancy was a
+inspection, we found that the most likely cause of this discrepancy was a
 mismatch between the genotype data and the identity of the accession for which
@@
 directly from the WGS data to those made using the SoySNP50K array, we observed
-that these two accessions had concordance rates of only 62\% and 70\%. To
-prevent such mismatches from affecting our GWAS analyses, we computed the
+that these two accessions had low concordance rates of only 62\% and 70\%. To
+ensure that such mismatches would not affect our GWAS analyses, we computed the
 concordance between genotype calls made from WGS data and the SoySNP50K array
-for the whole population (see the Methods section) and
+for all samples (see the Methods section) and
 excluded 24 accessions for which the concordance rate was below 90\%.
@@
 \textit{T} locus for pubescence color and found a 1-bp deletion as
-a putative causal variant at this locus. We found signals overlapping this gene
-using all four approaches (Figure \ref{pubescence_color_all_manhattan}). In
+a putative causal variant at this locus. For all three GWAS approaches, we
+found a signal overlapping this gene (Figure
+\ref{pubescence_color_all_manhattan}). In
 particular, the regions defined by the top 1\% most associated variants identified by the
@@
 (Figures \ref{pubescence_color_all_T_signal},
-\ref{pubescence_color_all_T_gene}). Neither vg nor Paragraph detected the causal
-variant at this locus given that these approaches did not target small indels.
+\ref{pubescence_color_all_T_gene}). Paragraph did not detect the causal
+variant at this locus given that this approach does not target small indels.
 
@@
 detected signals in this region using all four approaches, but only Platypus
-and the \textit{k}-mers identified signals that overlapped Glyma.03g258700
+and the \textit{k}-mers identified a signal that overlapped Glyma.03g258700
 (Figures \ref{pubescence-color-main-figure}a-b,
@@
 While the most significant \textit{k}-mer identified the causal variant documented
-at this locus \citep{yan2020}, the fourth most significantly associated variant
-found by Platypus corresponded to this variant (Figures
+at this locus \citep{yan2020}, it was the fourth most significantly associated variant
+for the Platypus genotype calls (Figures
 \ref{pubescence-color-main-figure}b-d, \ref{pubescence_color_nogray_Td_gene}).
@@
 coat specifically and thus in the absence of seed coat pigmentation
-\citep{tuteja2009}.  We did identify signals overlapping this causal variant
-using all four GWAS approaches (Figures \ref{seed_coat_color_all_manhattan},
-\ref{seed_coat_color_all_I_signal}), however none of the approaches identified
+\citep{tuteja2009}. While we did identify signals overlapping this causal variant
+using all three GWAS approaches (Figures \ref{seed_coat_color_all_manhattan},
+\ref{seed_coat_color_all_I_signal}), none of the approaches identified
 the causal variant. This failure to detect the causal variant is likely due
-to its complexity, which makes it difficult to detect and genotype. While it
+to its complexity, which makes it difficult to detect and genotype it. While it
 is likely that at least some of the significantly associated \textit{k}-mers
@@
 green seed coats in order to target the \textit{G} locus. Accordingly, this
-analysis found strong signals overlapping the \emph{G} locus using all four
+analysis found strong signals overlapping the \emph{G} locus using all three
 approaches (Figure \ref{seed_coat_color_greenyellow_manhattan}).
@@
 this locus and found an A>G SNP affecting transcript splicing as the causal
-variant at that locus.  Consistently with their results, we identified the two
+variant. Consistently with their results, we identified the two
 most significant \emph{k}-mers and the second most significant marker from the
@@
 \ref{seed_coat_color_greenyellow_G_signal},
-\ref{seed_coat_color_greenyellow_G_gene}). Unsurprisingly, none of the
-SV-based GWAS analyses identified the causal variant given its nature.
+\ref{seed_coat_color_greenyellow_G_gene}). Unsurprisingly, Paragraph
+did not identify the causal variant given its type.
 
@@
 \ref{stem_termination_all_Dt1_gene}, \ref{stem_termination_all_Dt1_kmers}).
-Our results are consistent with those of \cite{liu2010} and \cite{tian2010} in
-suggesting that there is more than one causal variant at that locus. Indeed, a
+Our results are consistent with previous findings in
+suggesting that there is more than one causal variant at the \textit{Dt1} locus \citep{liu2010, tian2010}. Indeed, a
 few cultivars bearing the allele associated with \emph{dt1} showed
@@
 \textit{k}-mer analyses, although these variants were not the most
-significantly associated (Figures \ref{hilum_color_all_T_gene},
+significantly associated overall (Figures \ref{hilum_color_all_T_gene},
 \ref{hilum_color_all_T_kmers}).
@@
 color also controls hilum color and we detected signals overlapping the
-associated inverted duplication using all approaches but vg (Figures
+associated inverted duplication using all approaches (Figures
 \ref{hilum_color_all_manhattan}, \ref{hilum_color_all_I_signal}). However,
@@
 
-We conducted a GWAS analysis contrasting accessions with black and brown hila
+We conducted a GWAS analysis that only compared accessions with black and brown hila
 in an attempt to detect stronger signals at the \textit{R} locus.  We
-accordingly detected signals overlapping this locus using all approaches,
-although these signals overlapped the Glyma.09g235100 gene associated with the
+accordingly detected a signal overlapping this locus using all approaches,
+but signals overlapped the Glyma.09g235100 gene associated with the
 \textit{R} locus only for Platypus and \textit{k}-mers (Figures
@@
 \ref{hilum_color_blackbrown_R_kmers}).  This failure to identify causal
-variants more directly is probably due to the existence of more than one causal
+variants more directly could be due to the existence of more than one causal
 variant at this locus. Interestingly, the most significantly associated
 \textit{k}-mers mapped just upstream of a gene (Glyma.09g234900) identified as
-a homolog of Glyma.09g235100 \citep{gillman2011}. This result highlights how
-difficult it may be to link significant \textit{k}-mers to causal variation in
-some cases.
+a homolog of Glyma.09g235100 \citep{gillman2011}.
 
@@
 
-We conducted a GWAS analysis contrasting accessions with normal and semi-sparse
-pubes\-cence in order to detect loci associated with pubescence density. We
-detected signals at the \textit{Ps} locus using all four approaches (Figures
+We conducted a GWAS analysis comparing accessions with normal and semi-sparse
+pubes\-cence in order to detect loci associated with pubescence density. This analysis
+detected signals at the \textit{Ps} locus using all three approaches (Figures
 \ref{pubescence-density-main-figure}a-b, \ref{pubescence_density_manhattan}). The causal
@@
 25.6-kb region identified as the causal CNV by \cite{liu2020ps}. We found that
-the most significantly associated \textit{k}-mer was located on the boundary of this CNV
-and that reads containing this \textit{k}-mer mapped to either of the two ends
+reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
 of the CNV (Figure \ref{pubescence-density-main-figure}b,d). This suggests that
@@
 CNV region may represent differences in sequence between the copies.
+
+The difference in size between the CNV reported by \cite{liu2020ps} and our
+results may be due to the presence of some unresolved sequence (a stretch of
+4,635 'N' nucleotides) over that interval in the reference sequence that we used.
+Because our SV filtering process removed any SV overlapping 'N' nucleotides in
+the reference, we reasoned that the CNV may have been found by our pipeline but
+filtered out later on. Indeed, found that a 42.8-kb duplication corresponding
+to the one described above among the genotype calls made by smoove. An association
+analysis of that single variant using GAPIT found a \textit{p}-value of 1.9e-23,
+which would have made it the most significant SV. In this particular case, the
+failure of the SV GWAS analysis to find identify the causal variant was due
+to our filtering parameters and not to limitations of the SV genotype calls per se.
 
@@
 
-We conducted a GWAS analysis contrasting dull and shiny seed coat phenotypes
-for the seed coat luster trait. Seed coat luster was shown by \cite{gijzen1999}
+Next, we conducted a GWAS analysis comparing dull and shiny seed coat phenotypes
+for the seed coat luster trait. Seed coat luster has been shown by \cite{gijzen1999}
 to be largely caused by the deposition of a hydrophobic protein (HPS)
@@
 for time to maturity instead of the number of days, combined with the limited
-phenotypic variation at this trait in this collection of accessions (Table
+phenotypic variation of this trait in this collection of accessions (Table
 \ref{MATURITY.GROUP_gwas_table}), likely explains our failure to detect
-signals.
+any signals.
 
@@
 \label{sv-gwas-results}
+In addition to analyzing loci for which the genes are already known, we were
+able to suggest candidate genes at loci for which the underlying genes are not
+known yet.
 
@@
 
-In addition to analyzing loci for which the genes are already known, we were
-able to suggest candidate genes at loci for which the underlying genes are not
-known yet. For example, both the \textit{L1} and the \textit{L2} loci for pod
+Both the \textit{L1} and the \textit{L2} loci for pod
 color have yet to be cloned. While \cite{he2015} proposed the gene
@@
 \ref{pod_color_all_manhattan}, \ref{pod_color_all_L2_signal}). Despite this
-number of genes being somewhat high, we noticed that the closest gene to the
+somewhat high number, we noticed that the closest gene to the
 most significant \emph{k}-mer (Glyma.03g005800) is also a MATE transporter,
 similarly to our top candidate for \emph{L1}. This gene may therefore represent
-an interesting candidate for this locus.
+an interesting candidate for this locus for future studies.
 
@@
 The two known loci for pubescence form in soybean, \textit{Pa1} and
-\textit{Pa2}, also have not been cloned yet, although \cite{gilbert2017} did
-suggest Glyma.12g213900 as a candidate for \textit{Pa1} based on data published
+\textit{Pa2}, have also not been cloned yet, although \cite{gilbert2017} has suggested
+Glyma.12g213900 as a candidate for \textit{Pa1} based on data published
 by \cite{bandillo2017}. This gene encodes a Myb transcription factor annotated
-as playing a role in trichome branching and is located near the most
-significant signals found by the Platypus, Paragraph and \emph{k}-mer GWAS
+as playing a role in trichome branching. Indeed, the most
+significant signals found by all three approaches were located near this gene.
 analyses (Figures \ref{pubescence-form-main-figure}a-b,
 \ref{pubescence_form_all_manhattan}, \ref{pubescence_form_all_Pa1_signal}).
-GWAS analyses based on Platypus and \emph{k}-mers identified two highly
-significant non-synonymous SNPs associated with appressed pubescence in this
-gene (Figures \ref{pubescence-form-main-figure}c-d, \ref{pubescence_form_all_Pa1_gene},
+The analyses based on Platypus and \emph{k}-mers identified two highly
+significant non-synonymous SNPs in Glyma.12g213900 that were associated with appressed pubescence
+(Figures \ref{pubescence-form-main-figure}c-d, \ref{pubescence_form_all_Pa1_gene},
 \ref{pubescence_form_all_Pa1_kmers}). The alternative alleles for these two
@@
 contrasting gene expression in resistant and susceptible lines four days after
-inoculation. Although the signals found by our analyses did overlap with
-Glyma.15g217100, we found that this gene was located at one end of the signal
-and was found in a region with markers or \emph{k}-mers showing a weaker
-association (Figure \ref{corrected_dry_weight_cdwGm15_signal}).  Although the
-large size of the signal means that the gene associated with the phenotype could
-be located almost anywhere in this region, we identified a particularly
-interesting region defined by the top 1\% associations found from the Platypus
+inoculation. Although the signals found by our analyses did overlap
+Glyma.15g217100, we found that this gene was located at the distal end of the signal
+in a region with markers or \emph{k}-mers showing a weaker
+association (Figure \ref{corrected_dry_weight_cdwGm15_signal}). The
+large size of the observed signal means that the gene associated with the phenotype could
+be located almost anywhere in this region, however, we identified a particularly
+interesting region defined by the top 1\% associations in the Platypus
 analysis. This region overlaps only three annotated genes and also contains the
 most significant \emph{k}-mer (Figure
-\ref{corrected_dry_weight_cdwGm15_signal}), which was located 30 kb away from a
-gene (Glyma.15g217700) annotated as a calcium ion binding protein with an
-EF-hand domain. Given the important role played by calcium signaling in defense
+\ref{corrected_dry_weight_cdwGm15_signal}). One of the genes in this region is
+a calcium ion binding protein with an EF-hand domain (Glyma.15g217700).
+Given the important role played by calcium signaling in defense
 responses \citep{zhang2014}, we suggest Glyma.15g217700 as an interesting
-candidate gene to explain resistance to \emph{P. sojae} in this population.
+candidate gene for explaining resistance to \emph{P. sojae} in this population.
 
@@
 
-In a few cases, we were also able to document possibly novel loci from the
-results of \textit{k}-mer GWAS. In a GWAS analysis contrasting accessions
+In addition to verifying known loci, we also observed potentially novel loci in
+the \textit{k}-mer GWAS analysis of stem termination and pod color.
+When comparing accessions
 with indeterminate and semi-determinate stem termination types, we observed
@@
 (Figure \ref{stem-termination-main-figure}b). Similarly, a \textit{k}-mer GWAS
-analysis contrasting accessions with black and brown pods found a previously
-undocumented signal on chromosome Gm15 that showed no LD with \textit{k}-mers
-at the \textit{L1} locus and may therefore represent a different locus (Figures
+analysis comparing accessions with black and brown pods found a previously
+undocumented signal on chromosome Gm15. \textit{k}-mers at this new locus showed no LD with \textit{k}-mers
+at the known \textit{L1} locus and may therefore represent a distinct locus (Figures
 \ref{pod_color_blbr_manhattan}, \ref{pod_color_blbr_ld}).

### 2023-02-20 cnv_analysis/smoove_cnv_gwas.R

@@ -1,0 +1,33 @@
+# This script carries out an association analysis for a single
+# variant using GAPIT: a CNV at the Ps locus genotyped by smoove
+
+# Loading the required packages
+library(VariantAnnotation)
+library(GAPIT3)
+library(gwask)
+
+# Loading the variant from the VCF file
+smoove_vcf <- VcfFile("~/sv_gwas/usda_lines/sv_calling/smoove/all_samples.smoove.square.vcf.gz")
+ps_cnv <- readVcf(smoove_vcf, param = GRanges(seqnames = "Gm12", ranges = IRanges(start = 36.2 * 10^6, end = 36.3 * 10^6)))
+ps_cnv <- ps_cnv[info(ps_cnv)$SVTYPE != "BND"]
+ps_cnv <- ps_cnv[abs(unlist(info(ps_cnv)$SVLEN)) < 100000 & abs(unlist(info(ps_cnv)$SVLEN)) > 10000]
+
+# Making sure that we end up with a single variant
+stopifnot(length(ps_cnv) == 1)
+
+# Artificially doubling because GAPIT bugs otherwise
+ps_cnv <- ps_cnv[c(1, 1)]
+
+# Reading the kinship and pca results
+kinship <- readRDS("filtered_variants/platypus_gapit_kinship.rds")
+pca <- readRDS("filtered_variants/platypus_gapit_pca.rds")
+
+gapit_results <- gapit_vcf(vcf = ps_cnv,
+			   kinship = kinship,
+			   pca = pca,
+			   phenodata = "phenotypic_data/phenotypic_data.csv",
+			   trait = "pubescence_density",
+			   N.sig = min(length(ps_cnv), 20),
+			   id_column = "bayer_id",
+			   tmproot = "tmpdir")
+

### 2023-02-20 tables/loci_table.R

@@ -151,32 +151,32 @@ loci_table$locus <- paste0("\\textit{", loci_table$locus, "}")
 loci_table$gene_name_v4 <- sub("^g", "G", loci_table$gene_name_v4)
 
 # Adding the type of causal variant using a lookup table
-causal_lookup <- c("flower_color_W1" = "Deletion \\citep{zabala2007}",
-		   "pubescence_color_T" = "Indel \\citep{zabala2003}",
+causal_lookup <- c("flower_color_W1" = "65-bp deletion \\citep{zabala2007}",
+		   "pubescence_color_T" = "1-bp deletion \\citep{zabala2003}",
 		   "pubescence_color_Td" = "SNP \\citep{yan2020}",
-		   "seed_coat_color_I" = "Inversion/duplication \\citep{tuteja2008}",
+		   "seed_coat_color_I" = "10.9-kb inversion/duplication \\citep{tuteja2008}",
 		   "seed_coat_color_G" = "SNP \\citep{wang2018}",
-		   "seed_coat_color_T" = "Indel \\citep{zabala2003}",
+		   "seed_coat_color_T" = "1-bp deletion \\citep{zabala2003}",
 		   "seed_coat_color_R" = "SNPs and indels \\citep{gillman2011}",
 		   "stem_termination_Dt1" = "SNPs \\citep{liu2010}",
 		   "stem_termination_Dt2" = "SNPs \\citep{ping2014}",
 		   "stem_termination_E3" = "Various types \\citep{tardivel2019}",
-		   "hilum_color_T" = "Indel \\citep{zabala2003}",
-		   "hilum_color_I" = "Inversion/duplication \\citep{tuteja2008}",
+		   "hilum_color_T" = "1-bp deletion \\citep{zabala2003}",
+		   "hilum_color_I" = "10.9-kb inversion/duplication \\citep{tuteja2008}",
 		   "hilum_color_R" = "SNPs and indels \\citep{gillman2011}",
-		   "hilum_color_W1" = "Deletion \\citep{zabala2007}",
-		   "pubescence_density_Ps" = "CNV \\citep{liu2020ps}",
+		   "hilum_color_W1" = "65-bp deletion \\citep{zabala2007}",
+		   "pubescence_density_Ps" = "25.6-kb CNV \\citep{liu2020ps}",
 		   "pubescence_density_Pd1" = "SNP \\citep{liu2020ps}",
 		   "pubescence_density_P1" = "SNP \\citep{liu2020ps}",
[…]

### 2023-02-20 additional_files/main_text.tex
Integrate suggestions by JA Sibbesen into the main Discussion section

@@
 
-Arguably, the most striking result from the analyses presented above is the
+One of the most striking results from the analyses presented above is the
 impressive performance of \emph{k}-mer-based GWAS compared to SNP/indel- and
-SV-based GWAS. Indeed, the GWAS analyses using \emph{k}-mers readily identified
+SV-based GWAS. The GWAS analyses using \emph{k}-mers identified
 the causal variant as the most significantly associated variant at four loci
@@
 for pubescence color, \emph{G} for seed coat color, and \emph{Ps} for
-pubescence density. In comparison, this result was achieved by SV-based GWAS
+pubescence density (Table \ref{loci-table}). In comparison, this result was achieved by SV-based GWAS
 at the \textit{W1} locus only, and at none of the loci for SNP/indel-based GWAS.
@@
 those variants. For example, the most significantly associated \emph{k}-mers
-mapping to the \emph{T} locus for pubescence color mapped to intronic sequences
-of the gene known to be associated with this locus, even though they did not
-correspond to the documented causal variant. Similarly, the most significantly associated
+mapping to the \emph{T} locus for pubescence color were located in intronic sequences
+of the gene known to be associated with this locus.
+Similarly, the most significantly associated
 \emph{k}-mers at the \emph{B} locus for seed coat luster mapped within a
-genomic interval known to vary in copy number at that locus, and the most
+genomic interval known to vary in copy number at that locus. In addition, the most
 significant \emph{k}-mers at the \emph{R} locus mapped just upstream of a
@@
 
-As would be expected of any GWAS, the success of the \emph{k}-mer approach
+As would be expected of any GWAS analysis, the success of the \emph{k}-mer approach
 depends on the quality of the phenotypic data and the complexity of the genetic
@@
 phenotypic values will necessarily make the interpretation more difficult than
-for simple qualitative traits. Notwithstanding those limitations, it remains
-clear that \textit{k}-mer GWAS performed better than any of the other
-approaches tested here given the same phenotypic data
+for simple qualitative traits. Notwithstanding those limitations,
+\textit{k}-mer-based GWAS clearly performed better than any of the other
+approaches tested here given the same phenotypic data. This is partly due
+to the fact that the \textit{k}-mer approach encompasses all variant types.
+Moreover, it is likely that the simpler workflow that results from using
+\textit{k}-mer presence/absence avoids some of the errors introduced during
+variant calling and genotyping, and therefore yields more accurate results.
 
@@
 
-While we have obtained considerable success when using \emph{k}-mers for GWAS, this
-approach does present unique challenges and limitations. One obvious limitation
+While we have have shown that using \emph{k}-mers for GWAS in soybean performed
+well compared to other GWAS methods, this
+approach does present some unique challenges and limitations. One obvious limitation
 is that using the presence/absence of \emph{k}-mers may not be appropriate in
@@
 \emph{k}-mers can detect variation independently of its type, some types of
-variation may be more difficult to detect than others. More precisely,
-variation will only be detected if it results in the presence or absence of a
-specific \emph{k}-mer or set of \emph{k}-mers of a given length (here, 31 bases) in a
-sequencing dataset. For example, copy number variation at a locus may not be
+variation may be more difficult to detect than others.
+For example, copy number variation at a locus may not be
 detected from presence/absence alone if all copies are identical and occur in
@@
 \emph{k}-mers together into coherent signals, and linking them to biologically
-meaningful sequence variation. The fact that this approach is
-reference-agnostic suggests that a paradigm shift will be needed to move away
-from reference-based analysis to better suited methods. For example, mapping
+meaningful sequence variation. Going forward, mapping
 the \emph{k}-mers or reads to a graph-based genome \citep[using e.g. the vg
-toolkit;][]{siren2021} instead of a linear reference may represent an
-improvement. Grouping significant \emph{k}-mers by LD instead of by genomic
-position, as was done for some of the analyses in this paper, should also prove
-useful in determining how many loci control a trait and as a starting point to
-identify the variation underlying presence/absence patterns. While we used an
+toolkit;][]{siren2021} may represent an
+improvement to using a single linear genome.
+Grouping significant \emph{k}-mers by LD instead of by genomic
+position, as was done for some of the analyses in this paper, could prove
+useful in determining how many loci control a trait. In this study, we used an
 assembly-based approach similar to what others have previously done
-\citep[e.g.][]{voichek2020, rahman2018} and have been able to link
-\emph{k}-mers to their underlying variation in simple cases involving SNPs or
-indels, we have not been able to link significantly associated \emph{k}-mers to
+\citep[e.g.][]{voichek2020, rahman2018}. This approach allowed us to link
+\emph{k}-mers to their underlying variation for simple cases involving SNPs and
+indels, but we were unable to link significantly associated \emph{k}-mers to
 more challenging causal variants such as those explaining the \emph{I} or
@@
 significant \emph{k}-mers to the sequence variation that underlies their
-presence is much needed. Finding ways to limit the number of spurious
+presence is much needed. Furthermore, finding ways to limit the number of spurious
 associations found for some traits, as was the case for flower color and
-pubescence color in our study, is also needed in order to obtain robust results
-from \emph{k}-mer-based GWAS.
-
-In this work, we aimed to provide a set of tools enabling the downstream
+pubescence color in our study, is also needed in order to obtain robust results.
+
+In this work, we provide a set of tools enabling the downstream
 analysis of \emph{k}-mer-based GWAS. Most importantly, we developed the
 katcher program for retrieving all reads containing
-any of a set of significant \emph{k}-mers in a computationally efficient manner. Previously published studies usually
+any of a set of significant \emph{k}-mers in a computationally efficient manner. Previous studies usually
 mapped \emph{k}-mers to a reference genome to identify their reference-based
@@
 of \emph{k}-mers that diverge significantly or are not found at all in the
-reference. Moreover, it does not take into account the sequence context where
+reference. Moreover, it does not take into account the sequence context of the read where
 the \emph{k}-mer was found, which may provide valuable information for mapping
-purposes. Our approach avoids these pitfalls by systematically retrieving reads
+purposes. Our approach avoids these pitfalls by instead systematically retrieving reads
 containing significant \emph{k}-mers, including those that did not map at all
@@
 
-In addition to \emph{k}-mers, we also assessed the use of SVs in GWAS. Our
-rationale was that since SVs have an important functional impact, they may
-facilitate the identification of causal variants when these are in fact SVs.
+In addition to \emph{k}-mers, we also assessed the use of SVs in GWAS analysis of soybean.
 Although we did find the causal variant as the most significantly associated SV
-at the \emph{W1} locus, results have otherwise been rather inconclusive.
+at the \emph{W1} locus, results were otherwise rather inconclusive.
 Indeed, most of the other known or putative causal variants were SNPs or
-indels, and therefore were not represented in the SV datasets. GWAS analyses at
-other loci where SVs are known to be the underlying causal variant (e.g.
-\emph{I}, \emph{B}, \emph{Ps}) did not find the causal variants, most likely
+indels, and therefore were not represented in the SV datasets. At
+most loci where SVs are known to be the underlying causal variant (e.g.
+\emph{I}, \emph{B}, \emph{Ps}), the analyses based on SVs did not find the causal variants, most likely
 due to the difficulty of calling and genotyping such variants
-\citep{kosugi2019, lemay2022}.
-
-Based on these results, we believe that it may not be worth, at least in
-soybean, to conduct GWAS based on SVs. Given the current performance of SV
+\citep{kosugi2019, lemay2022}. In the case of the \textit{Ps} locus, this limitation
+was due to a filtering step in our pipeline rather than to an inherent inability to
+properly call and genotype the causal variant.
+
+Given the current performance of SV
 genotyping programs, it is not clear whether using SVs for GWAS provides any
@@
 SNP/indel-based GWAS provided better guidance in delimiting the signals found
-at a given locus given the much higher density of markers in the SNP dataset
-than in the SV dataset. The usefulness of SV-based GWAS compared to SNP-based
+at a given locus because of the much higher density of markers compared to SVs.
+The usefulness of SV-based GWAS compared to SNP-based
 GWAS will largely depend on whether SVs can represent variation that is not
@@
 \cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
-is therefore likely that the ``tagability'' of SVs using SNPs will depend on
-the particular type of variant and on the species being studied. Even though
-SVs remain important components of sequence and functional variation, a more
+is therefore likely that how well SVs can be tagged by SNPs will depend on
+the particular type of SV and on the species. A more
 efficient approach than SV-based GWAS may be to conduct \emph{k}-mer-based GWAS
-and only then use SV discovery methods to identify SVs in the population and
-link them to significant \emph{k}-mers. While previous studies have found
-interesting results using SVs for GWAS \citep[e.g.][]{zhang2015, akakpo2020,
-dominguez2020}, we would be curious to see whether these results
+and only then link known SVs to significant \emph{k}-mers.
+Alternatively, \textit{k}-mer-based GWAS can be conducted specifically for
+\textit{k}-mers that are associated with SVs \citep{jayakodi2020}.
+Previous studies have
+obtained noteworthy results by using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
+dominguez2020}, however it would be worthwhile to see whether these results
 could be replicated using \emph{k}-mer-based GWAS.
@@
 causal variation or genes associated with known loci, it is unclear whether
-conducting SV-based GWAS is worth the investment. We also used the
+conducting SV-based GWAS is worth the investment. In addition, we used the
 results from the \emph{k}-mer- and SNP-based GWAS to suggest candidate genes
@@
 believe that an optimal workflow may involve conducting SNP and \emph{k}-mer
-GWAS in parallel to identify significant signals and candidate genes, and then
-possibly use SV datasets to identify interesting variants. Given the large
-number of accessions for which WGS data are now available in several species,
+GWAS in parallel to identify significant signals and candidate genes,
+potentially in combination with SV datasets. Given the large
+number of samples for which WGS data are now available in several species,
 applying \emph{k}-mer-based GWAS to leverage already existing sequencing and
 phenotypic data appears feasible and promising. As part of our work, we
-developed several computational tools that should help other researchers to
-analyze the results of a \emph{k}-mer-based GWAS analysis. However, much work
-remains to be done to define state-of-the-art methods for the downstream
+developed several computational tools that should help other researchers with the downstream
+analysis of a \emph{k}-mer-based GWAS results. However, much work
+remains to be done in developing state-of-the-art methods for the downstream
 analysis of significant \emph{k}-mers. In particular, better approaches are

### 2023-02-20 additional_files/main_text.tex
Integrate suggestions by JA Sibbesen into the main figure captions

@@
 	\contcaption{Results of SV- and \textit{k}-mer-based GWAS at the
-	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots of
-	the analyses based on Paragraph (SV) genotype calls and
-	\textit{k}-mers.  Horizontal dotted lines indicate the significance
+	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
 	threshold whereas vertical dotted lines indicate the location of the
-	\textit{W1} locus. (b) Zoomed-in Manhattan plots of the analyses based
-	on Paragraph genotypes and \textit{k}-mers at the \textit{W1} locus.
-	Horizontal dotted lines again indicate the significance threshold
-	whereas the vertical dotted lines indicate the location of the
+	\textit{W1} locus. (b) Zoomed-in Manhattan plots at the \textit{W1} locus.
+	The vertical dotted lines indicate the location of the
 	Glyma.13g072100 gene associated with this locus. The shaded gray
-	rectangles show the regions delimited by the top 5\% (Paragraph) and
+	rectangles show the regions delimited by the top 5\% (SVs) and
 	1\% (\textit{k}-mers) associations.  (c) Gene model of Glyma.13g072100.
@@
 	sequences, and arrows indicate the direction of transcription. The red
-	rectangle highlights the region shown in panel (d). (d) Haplotypes and
-	significant \textit{k}-mers at the \textit{W1} locus. Nucleotides are
+	rectangle highlights the region shown in panel (d). (d) Pairwise
+	alignment of the two haplotype sequences in the population at the
+	\textit{W1} locus. Nucleotides are
 	colored according to the \textit{p}-value of the most significant
-	\textit{k}-mer overlapping them. Vertical lines between haplotypes
+	\textit{k}-mer overlapping them. Vertical dotted lines between haplotypes
 	indicate sequence differences, whereas dashes indicate gaps in the
 	alignment. (e) Contingency table of the phenotypes and haplotypes found
-	in the population at the \textit{W1} locus. The haplotypes are the ones
-	represented in panel (d).}
+	in the population at the \textit{W1} locus.}
 \end{figure}
@@
 	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
-	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots
-	of the analyses based on Platypus (SNP/indel) genotype calls and
-	\textit{k}-mers.  Horizontal dotted lines indicate the significance
+	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
 	threshold whereas vertical dotted lines indicate the location of the
-	\textit{Td} locus. (b) Zoomed-in Manhattan plots of the analyses based
-	on Platypus genotypes and \textit{k}-mers at the \textit{Td} locus.
-	Horizontal dotted lines again indicate the significance threshold
-	whereas the vertical dotted lines indicate the location of the
+	\textit{Td} locus. (b) Zoomed-in Manhattan plots at the \textit{Td} locus.
+	The vertical dotted lines indicate the location of the
 	Glyma.03g258700 gene associated with this locus. The shaded gray
 	rectangles show the regions delimited by the top 1\% associations.
-	In the case of Platypus, blue points denote markers used in the
+	In the case of SNP/indels, blue points denote markers used in the
 	original analysis, whereas red points denote points that had originally
@@
 	direction of transcription. The red rectangle highlights the region
-	shown in panel (d). (d) Haplotypes and significant \textit{k}-mers at
-	the \textit{Td} locus. Nucleotides are colored according to the
+	shown in panel (d). (d) Pairwise alignment of the two haplotype
+	sequences in the population at the \textit{Td} locus. Nucleotides are
+	colored according to the
 	\textit{p}-value of the most significant \textit{k}-mer overlapping
-	them. The vertical line between haplotypes indicates the location of
+	them. The vertical dotted line between haplotypes indicates the location of
 	the causal SNP at this locus. (e) Contingency table of
 	the phenotypes and haplotypes found in the population at the
-	\textit{Td} locus. The haplotypes are the ones represented in panel
-	(d).}
+	\textit{Td} locus.}
 \end{figure}
@@
 	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
-	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots
-	of the analyses based on Platypus (SNP/indel) genotype calls and
-	\textit{k}-mers.  Horizontal dotted lines indicate the significance
+	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
 	threshold whereas vertical dotted lines indicate the location of the
-	\textit{G} locus. (b) Zoomed-in Manhattan plots of the analyses based
-	on Platypus genotypes and \textit{k}-mers at the \textit{G} locus.
-	Horizontal dotted lines again indicate the significance threshold
-	whereas the vertical dotted lines indicate the location of the
+	\textit{G} locus. (b) Zoomed-in Manhattan plots at the \textit{G} locus.
+	The vertical dotted lines indicate the location of the
 	Glyma.01g198500 gene associated with this locus. The shaded gray
 	rectangles show the regions delimited by the top 1\% associations.
-	In the case of Platypus, blue points denote markers used in the
+	In the case of SNP/indels, blue points denote markers used in the
 	original analysis, whereas red points denote points that had originally
@@
 	direction of transcription. The red rectangle highlights the region
-	shown in panel (d). (d) Haplotypes and significant \textit{k}-mers at
-	the \textit{G} locus. Nucleotides are colored according to the
+	shown in panel (d). (d) Pairwise alignment of the two haplotype
+	sequences in the population at the \textit{G} locus. Nucleotides are
+	colored according to the
 	\textit{p}-value of the most significant \textit{k}-mer overlapping
-	them. The vertical line between haplotypes indicates the location of
+	them. The vertical dotted line between haplotypes indicates the location of
 	the causal SNP at this locus. (e) Contingency table of
 	the phenotypes and haplotypes found in the population at the
-	\textit{G} locus. The haplotypes are the ones represented in panel (d).}
+	\textit{G} locus.}
 \end{figure}
@@
 	\textit{Ps} locus for pubescence density.
-	(a) Genome-wide Manhattan plot of significant \textit{k}-mers. The
+	(a) Genome-wide Manhattan plot showing the location of significant \textit{k}-mers. The
 	horizontal dotted line indicates the significance threshold whereas the
 	vertical dotted line indicates the location of the \textit{Ps} locus.
-	(b) Zoomed-in Manhattan plot at the \textit{Ps} locus.  The horizontal
-	dotted line represents the significance threshold whereas the
+	(b) Zoomed-in Manhattan plot at the \textit{Ps} locus. The
 	vertical dotted line indicates the location of the most significant
-	\textit{k}-mer. The shaded gray area shows the extent of the causal CNV.
+	\textit{k}-mer. The shaded gray area shows the location of the causal CNV.
 	(c) Estimated copy number (sequencing depth at position divided by
@@
 	semi-sparse pubescence at the \textit{Ps} locus. Solid lines represent
-	the average copy number across 73 (semi-sparse pubescence) and 180
+	the median copy number across 73 (semi-sparse pubescence) and 180
 	accessions (normal pubescence) computed using a 100-bp sliding window.
@@
 	the copy number of accessions with contrasting pubescence.
-	(d) Histogram of the mapping location of 1,169 reads containing the
+	(d) Histogram showing the location of 1,169 mapped reads containing the
 	most significant \textit{k}-mer in accessions with semi-sparse pubescence.
@@
 	region, suggesting that they are associated with this variant. Panels
-	b, c and d share the same x-axis.}
+	(b), (c) and (d) share the same x-axis.}
 	\label{pubescence-density-main-figure}
@@
 	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
-	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots
-	of the analyses based on Platypus (SNP/indel) genotype calls and
-	\textit{k}-mers. Horizontal dotted lines indicate the significance
+	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
 	threshold whereas vertical dotted lines indicate the location of the
-	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots of the analyses based
-	on Platypus genotypes and \textit{k}-mers at the \textit{Pa1} locus.
-	Horizontal dotted lines again indicate the significance threshold
-	whereas the vertical dotted lines indicate the location of the
+	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots at the \textit{Pa1} locus.
+	The vertical dotted lines indicate the location of the
 	Glyma.12g213900 gene suggested as a candidate for this locus. The shaded gray
 	rectangles show the regions delimited by the top 1\% associations.
-	In the case of Platypus, blue points denote markers used in the
+	In the case of SNP/indels, blue points denote markers used in the
 	original analysis, whereas red points denote points that had originally
@@
 	direction of transcription.
-	(d) Zoomed-in Manhattan plots at the location of the Glyma.12g213900 gene.
-	Data is represented on the same x-scale in these plots as the gene model
-	in panel (c). Vertical lines indicate the location of two non-synonymous
+	(d) Manhattan plots at the location of the Glyma.12g213900 gene.
+	Vertical lines indicate the location of two non-synonymous
 	SNPs associated with appressed pubescence, which may represent causal variants
-	at this locus.}
+	at this locus. Panels (c) and (d) share the same axis.}
 \end{figure}
@@
 	indeterminate and semi-determinate stems]{Detection of novel signals for stem termination type
-	using a \textit{k}-mer GWAS analysis contrasting accessions with
+	using a \textit{k}-mer-based GWAS analysis contrasting accessions with
 	indeterminate and semi-determinate stems.
-	(a) Genome-wide Manhattan plot of significant \textit{k}-mers.
+	(a) Genome-wide Manhattan plot showing the location of significant \textit{k}-mers.
 	The horizontal dotted line indicates the significance threshold
 	whereas vertical dotted lines indicate the location of the \textit{Dt1}
-	classical locus as well as three newly documented loci (\textit{stGm11},
+	classical locus as well as three new loci of interest (\textit{stGm11},
 	\textit{stGm16}, and \textit{stGm18}).
-	(b) Pairwise LD among the 162 significant \textit{k}-mers identified.
-	\textit{k}-mers are sorted along the y-axis according to their
-	putative position along the reference assembly version 4 of Williams82,
-	as identified by ``Gm'' chromosome labels. \textit{k}-mers are
-	represented in the same order along the x- and y-axis. The colored
-	rectangles drawn below the x-axis represent the
+	(b) Pairwise LD between the 162 significant \textit{k}-mers identified.
+	\textit{k}-mers are sorted along the axes according to their
+	putative position on the reference assembly version 4 of Williams82.
+	\textit{k}-mers are
+	represented in the same order along the x- and y-axis. The blue
+	rectangles below the x-axis represent the
 	-log\textsubscript{10}(\textit{p}) of each \textit{k}-mer.}

### 2023-02-20 figures/pubescence_density_main_figure.R

@@ -66,7 +66,7 @@ roll_averaged_cov$pos <- (kmers_zoomed_manhattan$vp$xscale[1] - 100):(kmers_zoom
 
 coverage_data <- list(normal = roll_averaged_cov[, samples[phenotype == 1]], semisparse = roll_averaged_cov[, samples[phenotype == 2]])
 
-average <- lapply(coverage_data, function(x) apply(x, 1, mean))
+average <- lapply(coverage_data, function(x) apply(x, 1, median))
 lower_bound <- lapply(coverage_data, function(x) apply(x, 1, quantile, probs = 0.25))
 upper_bound <- lapply(coverage_data, function(x) apply(x, 1, quantile, probs = 0.75))
 
@@ -112,7 +112,7 @@ grid.text("(a)", x = 0.02, y = 0.95)
 draw_manhattan(list(kmers_gwide_manhattan), 
 	       sigline_regexp = "sigline_3",
 	       siglabel_regexp = "siglabel_3",
-	       labels = "",
+	       labels = "k-mers",
 	       fontsize = 10)
 
 
@@ -120,7 +120,7 @@ draw_manhattan(list(kmers_gwide_manhattan),
 seekViewport("main")
 pushViewport(viewport(layout.pos.row = 3, name = "zoomed_manhattan"))
 grid.text("(b)", x = 0.02, y = 0.95)
-draw_zoomed(list(kmers_zoomed_manhattan), labels = "", fontsize = 10, xaxis = FALSE)
+draw_zoomed(list(kmers_zoomed_manhattan), labels = "k-mers", fontsize = 10, xaxis = FALSE)
 
 # Push a viewport to plot the coverage data
 seekViewport("main")

### 2023-02-25 additional_files/additional_file_1.tex
Integrate suggestions by JA Sibbesen into the supplemental methods

@@
 
-We used the Illumina WGS data of the 389 selected \emph{G. max} accessions as a
-first source of data for SV discovery using methods similar to those used by
+We used the Illumina whole-genome sequencing (WGS) data from the 389 selected \emph{G. max} accessions
+for SV discovery using methods similar to those described by
 \cite{lemay2022}.  Briefly, we used four different SV discovery programs or
@@
 		v. 1047 \citep{kielbasa2011}.
-\item We ran Manta v. 1.6.0 \citep{chen2016} in 78 randomly selected batches of
+\item We ran Manta v. 1.6.0 \citep{chen2016} in randomly selected batches: 77 batches of
 	5 samples and 1 batch of 4 samples. We used the candidate SVs
 		identified from each of the batches for downstream analyses.
-\item We used smoove v. 0.2.4 \citep{pedersen2019} with the sequence of
-	commands \texttt{smoove call}, \texttt{smoove merge}, \texttt{smoove
-		genotype} and \texttt{smoove paste} to obtain a dataset of
-		candidate SVs. For three (3) of the 389 samples, we were unable
-		to obtain candidate variants from smoove due to a reproducible
-		cryptic error (segmentation fault) when running \texttt{smoove
-		call}. We nevertheless decided to use smoove given that this
-		represented a very small fraction of the total number of
-		samples and that other programs were able to successfully
-		discover SVs from those samples.
+\item We used smoove v. 0.2.4 \citep{pedersen2019} to obtain a dataset of candidate SVs
+	using the following sequence of commands:
+	\texttt{smoove call}, \texttt{smoove merge}, \texttt{smoove
+		genotype} and \texttt{smoove paste}.
+		For three (3) of the 389 samples, we were unable
+		to obtain candidate variants due to a reproducible
+		error (segmentation fault) when running \texttt{smoove
+		call}.
 	\item We used SvABA v. 1.1.3 \citep{wala2018} to call SVs from each of the
@@
 		as either deletions, duplications or inversions, and converted
-		them to explicit sequence using our own R script. Variants
+		them to explicit sequence using a custom R script. Variants
 		called as indels by SvABA were used as is for downstream
@@
 
-We filtered the SVs called by all tools to remove those smaller than 50 bp or
-larger than 500 kb in size, those located on unanchored scaffolds or organellar
-genomes, those classified as unresolved breakends (\texttt{SVTYPE=BND}), and
-any variant containing at least one N in either the reference or alternate
-allele sequence.  We converted all variants to a sequence-explicit format and
-normalized them using bcftools \texttt{norm} v. 1.10.2-105
-\citep{li2009-samtools}.
+We filtered the SVs called by each tool to remove those meeting any of the following conditions:
 
-We generated a single VCF file from the variants called using Illumina data by
+\begin{itemize}
+	\item smaller than 50 bp or larger than 500 kb in size
+	\item located on unanchored scaffolds or organellar genomes
+	\item classified as unresolved breakends (\texttt{SVTYPE=BND})
+	\item containing at least one N in either the reference or alternate allele sequence
+\end{itemize}
+
+Next, we converted all filtered variants to a sequence-explicit format using
+\texttt{bayesTyperTools convertAllele} \citep{sibbesen2018} and normalized them
+using bcftools \texttt{norm} v. 1.10.2-105 \citep{li2009-samtools}.
+
+A single VCF file was generated from the variants called using Illumina data by
 running SVmerge \citep{wong2010} in two steps. In the first step, we merged
-variants among the ones called by a given tool by running SVmerge using a
-sliding window of 5 bp to merge together very similar variants called by the
-same tool. In the second step, we used the output of the first merging steps
-for all four tools and merged similar variants across these four datasets using
-a sliding window of 15 bp in SVmerge.  We proceeded this way because using a
+very similar variants among the ones called by the same tool using a
+sliding window of 5 bp. Next, we used the output of the first merging steps
+for all four tools and merged similar variants across the four datasets using
+a sliding window of 15 bp. This approach was used because a
 sliding window of 15 bp on the outputs of all tools without pre-merging
@@
 We used SVs discovered among 17 Canadian soybean cultivars sequenced using
-Oxford Nano\-pore Technologies (ONT) by \cite{lemay2022} as a second source of
-data for SV discovery.  Although that set of cultivars does not overlap the
-ones analyzed in this study, the high quality of the SVs that can be discovered
-from long-read sequencing data motivated the inclusion of this dataset among
-the SV candidates subjected to genotyping.  We merged the SVs detected in the
+Oxford Nano\-pore Technologies (ONT) by \cite{lemay2022} as an additional
+source of candidate SVs. Although this set of cultivars does not overlap the
+ones analyzed in this study, it is likely that many of the SVs occurring
+in Canadian germplasm also occur in the germplasm used for this study given that modern soybean
+cultivars are derived from a rather small set of progenitors. We merged the SVs detected in the
 long-read (ONT) data with the ones discovered in the short-read (Illumina) data
@@
 discovered from the Illumina data was that whenever both technologies discover
-the same variants, Illumina data is likely to be more precise given its higher
-basecalling accuracy.
+the same variants, the breakpoints of SVs discovered from Illumina data are
+likely to be more precise given the higher basecalling accuracy.
 
@@
 
-In addition to using the Illumina data from 389 accessions, we also used
+In addition to Illumina data from 389 accessions, we also used
 high-quality genome assemblies of 29 \emph{G. max} or \emph{Glycine soja}
-accessions and called SVs by comparing their sequence to that of Williams82. We
+accessions to call SVs. We
 retrieved 26 high-quality genomes assembled by \cite{liu2020-pangenome} as well
-as that of the cultivar Zhonghuang 13 (ZH13) from the Genome Warehouse
-repository (\url{https://ngdc.cncb.ac.cn/gwh/}).  We also retrieved the genomes
+as the genome of the cultivar Zhonghuang 13 (ZH13) from the Genome Warehouse
+repository (\url{https://ngdc.cncb.ac.cn/gwh/}). Furthermore, we retrieved the genomes
 of \emph{G. soja} accession W05 and \emph{G. max} cultivar Lee from SoyBase
@@
 
-We used methods based on \cite{liu2020-pangenome} to call SVs based on those
+We used methods based on \cite{liu2020-pangenome} to call SVs from the
 genome assemblies.  First, we used the nucmer program of the MUMmer suite of
@@
 between any genome and the Williams82 reference using the \texttt{-1} option.
-We used these filtered alignments to call SVs using a version of the svmu
+The filtered alignments were used to call SVs with a version of the svmu
 program \citep{chakraborty2019} that we forked from the original repository
@@
 assemblies.  Copy number variants and inversions called by svmu were therefore
-excluded from downstream analyses, although we did take inverted regions into
-account when determining the sequence of insertions in these regions to ensure
-their proper representation in the output VCF file. We also filtered the
-resulting set of variants to remove those smaller than 50 bp or larger than 500
+excluded from downstream analyses. The
+resulting set of variants was filtered to remove those smaller than 50 bp or larger than 500
 kb and those with any N in either the reference or alternate allele sequence.
 Finally, we normalized the VCF files using \texttt{bcftools norm} and merged
-them using SVmerge to obtain a single VCF file of variants discovered through
+them using SVmerge to obtain a single file with variants discovered through
 comparison of high-quality genome assemblies.
@@
 
-We used Paragraph v. 2.4a \citep{chen2019} as one of two SV genotyping programs
-to generate a set of SV genotypes for the population of 389 samples for use in
-GWAS analyses.  The set of SVs used as input for Paragraph was the result of
-merging the set of variants discovered from the Illumina and ONT data with
-those discovered from the high-quality assemblies and svmu using SVmerge.
-Variants in this candidate set were prepared for input to Paragraph by removing
+We used Paragraph v. 2.4a \citep{chen2019}
+to generate a set of SV genotypes from the Illumina data of the 389 samples used in
+GWAS analyses.  The set of SVs used as input for Paragraph was generated using SVmerge by
+merging variants discovered from the Illumina and ONT data with
+those discovered from the high-quality assemblies and svmu.
+Variants in this candidate set were prepared for input to Paragraph by first removing
 those closer than 150 bp (the maximum read length in our dataset) to any
-chromosome end and padding the variant representation using a script written by
-\cite{hickey2020} for that purpose.
+chromosome end and then padding the variant representation using a script written by
+\cite{hickey2020}. The latter step was necessary for the variants to be used
+as input to Paragraph.
 
-We used the recommendations by the authors of Paragraph for running their
-program at population scale by running Paragraph individually on each sample.
-We computed average sequencing depth for each sample from the output of the
+We ran Paragraph individually on each sample as recommended by the authors.
+We computed the average sequencing depth for each sample from the output of the
 \texttt{samtools coverage} command and used 20 times that value for the
@@
 \texttt{bcftools merge} in order to get a  single genotype file for the whole
-population. This VCF was further filtered by setting the genotype calls made
-from less than two reads to missing and by removing variants with minor allele
+population. Finally, we filtered the genotyped SVs by setting the genotype calls made
+from less than two reads to missing and by removing variants with a minor allele
 frequency < 0.02 or a proportion of missing data > 0.5. The resulting dataset
-of 186,306 SVs was used for GWAS.
-
-\subsection*{Genotyping structural variation using the vg toolkit}
-\label{sv-gwas-sv-genotyping-vg}
-
-We also used the vg toolkit v. 1.41.0 \citep{hickey2020} and its Giraffe
-aligner \citep{siren2021} for genotyping our set of candidate SVs on the whole
-population. The reason for also using vg is that its variation graph model allows
-the integration of SNPs and indels and makes for a more robust and integrated
-approach to SV genotyping, even though previous benchmarks comparing those two
-programs in soybean showed a reduced sensitivity of the vg toolkit in
-genotyping insertions \citep{lemay2022}.
-
-We constructed the variation graph using the \texttt{vg autoindex} command with
-the \texttt{-\-workflow giraffe} option. We used the Williams82 reference
-assembly version 4 as a basis for the graph, to which we added the set of
-candidate SVs described above and the set of 7M SNPs and indels called by
-Platypus.  We aligned the trimmed reads of the 741 SRA runs in FASTQ format to
-the variation graph using the \texttt{vg giraffe} command with default
-parameters. Reads that were still paired after adapter- and quality- trimming
-were aligned in paired-end mode, whereas the remaining reads were aligned in
-single-end mode. The resulting files in GAM format were used to compute read
-coverage on the graph using the \texttt{vg pack} command with options
-\texttt{-Q5 -s5}. GAM files belonging to the same accession but originating
-from different SRA runs were concatenated to be used as input for the
-\texttt{vg pack} command such that a single coverage file per accession was
-obtained as output of this step. The command \texttt{vg call} was then used to
-call genotypes based on read coverage using the parameters
-\texttt{-\-genotype-snarls -\-min-length 50} in order to call all variants
-$\geq$ 50 bp supported by the graph.
-
-We merged the genotypes called on all samples using \texttt{bcftools merge} to
-obtain a single VCF file for the population. Multiallelic records were split
-into biallelic records after merging. Genotype calls made with a genotype
-quality (GQ) value < 10 were set to missing, and variants were filtered for a
-fraction of missing genotypes $\leq$ 0.5 and a minor allele frequency $\geq$
-0.02. The resulting VCF file, which was used for GWAS, contained 51,960 SVs.
+of 186,306 called SVs was used for GWAS.
 
@@
 mismatch between the sequencing data and the identity of the accession was the
-most probable cause of these discrepancies. In order to filter out all samples
+most probable cause of these discrepancies. In order to identify the samples
 for which this could be the case, we compared the genotypes of SNPs obtained
-from the SoySNP50K array \cite{song2013} to those obtained from the Illumina WGS
-data of the 385 samples that had a match in the GRIN database. Briefly, we
+from the SoySNP50K array \cite{song2013} to those predicted from the Illumina WGS
+data. We did this for the 385 samples that had a match in the GRIN database. Briefly, we
 retrieved SoySNP50K genotype calls from SoyBase \citep{grant2010} and extracted
-all SNPs with a MAF $\geq$ 0.1 in our population. We translated the positions
+all SNPs with a MAF $\geq$ 0.1 in our population. We then translated the positions
 of these SNPs from assembly version 2 to assembly version 4 of Williams by
@@
 mpileup} and \texttt{bcftools call}. We then used \texttt{bcftools gtcheck} to
-compute the concordance of genotype calls between WGS samples and their
-expected match in the SoySNP50K genotype dataset.  We identified 24 samples
+compute the concordance between the WGS genotype calls and the
+SoySNP50K genotype dataset. We identified 24 samples
 with < 90\% concordance between the two genotype datasets and removed them from
-the dataset, in addition to another sample for which concordance was over 90\%
+the dataset. We also excluded an additional sample for which concordance was over 90\%
 but poor raw sequencing data quality (\%GC content) suggested contamination.
-Following these filtering steps, 363 samples remained for GWAS analyses. The
-concordance between WGS data and SoySNP50K genotypes is included in
-Additional file 2.
+The concordance between WGS data and SoySNP50K genotypes is included in
+Additional file 2 and shown in Figure \ref{concordance-histogram}.
 
@@
 % Figures mentioned only in supplemental data
+
+% Histogram with concordance rates
+\begin{cfigure}
+	\includegraphics[width = 6in]{concordance_histogram}
+	\caption{Histogram of concordance rates between genotypes derived from WGS data and SoySNP50K genotypes for 385 soybean accessions.
+	The vertical line indicates the 0.9 concordance threshold that was used for filtering out mismatching samples.}
+	\label{concordance-histogram}
+
+\end{cfigure}
+
+\clearpage
+
 % Flower color

### 2023-02-25 figures/concordance_histogram.R

@@ -1,0 +1,12 @@
+# A histogram with the concordance values from the WGS/SoySNP50K comparison
+
+# Reading the bcftools gtcheck results
+# DEPENDENCY: illumina_data/soysnp50k_genotyping/gtcheck_results.tsv
+gtcheck <- read.table("illumina_data/soysnp50k_genotyping/gtcheck_results.tsv", sep = "\t", header = TRUE)
+
+png("figures/concordance_histogram.png", width = 6, height = 6, units = "in", res = 200)
+
+hist(gtcheck$conc_rate, breaks = 30, main = NULL, xlab = "Concordance rate", ylab = "Number of accessions")
+abline(v = 0.9, lty = 2)
+
+dev.off()

### 2023-02-28 additional_files/additional_file_1.tex
Integrate suggestions by JA Sibbesen into the supplemental results

@@
 
-This section provides a more detailed account of some the the results obtained
-by the GWAS analyses on a per-trait basis. Aspects of the results that were
+This section provides a more detailed per-trait account of some of the results obtained
+by the GWAS analyses. Aspects of the results that were
 already discussed in the main text are not repeated here; in such cases,
@@
 at all loci using all four methods is presented in Table \ref{signals_table}.
-Readers are invited to refer to this table as we go over the results obtained
-for each trait below.
 
@@
 
-We conducted a single GWAS analysis contrasting purple and white flower colors
+We conducted a GWAS analysis contrasting purple and white flower colors
 (Table \ref{FLOWER.COLOR_gwas_table}). Of the two classical loci reported by
@@
 \ref{flower_color_manhattan}). An analysis of the pairwise LD between a subset
-of those \textit{k}-mers showed that all \textit{k}-mers formed a single LD block
-linked to the \textit{W1} locus (Figure \ref{flower_color_ld}), ruling out the
-presence of a second locus controlling that trait in this population.
+of those \textit{k}-mers showed that most \textit{k}-mers formed a single LD block
+linked to the \textit{W1} locus (Figure \ref{flower_color_ld}). A small LD block
+occurred on chromosome Gm09, however it was in moderate LD with the one on chromosome
+Gm13 and therefore probably does not represent a separate locus.
 
@@
 
-We conducted two GWAS analyses on pubescence color. The first analysis
+We conducted two different GWAS analyses on pubescence color. The first analysis
 contrasted gray, tawny and light tawny pubescence colors (Table
@@
 \ref{pubescence_color_all_manhattan}). Detailed results at the \textit{T} locus
-van be viewed in figures \ref{pubescence_color_all_T_signal},
+can be viewed in figures \ref{pubescence_color_all_T_signal},
 \ref{pubescence_color_all_T_gene} and \ref{pubescence_color_all_T_kmers}. We
 observed that not all accessions with gray pubescence harbored the deleted
-nucleotide, which could indicate either that other causal variants resulting in
-a non-functional F3’H are present in this population, or that the causal
+nucleotide, which could indicate either that other causal variants are resulting in
+a non-functional F3’H in this population, or that the causal
 variant lies elsewhere.
@@
 
-The second GWAS analysis on seed coat color focused on a subset of accessions
-with yellow or green seed coat only, as this difference is controlled by the
-\emph{G} locus (Table \ref{SEED.COAT.COLOR_gwas_table}). Detailed results of
+The second GWAS analysis on seed coat color focused on the subset of accessions
+with yellow or green seed coat only (Table \ref{SEED.COAT.COLOR_gwas_table}),
+as this difference is controlled by the \emph{G} locus. Detailed results of
 this analysis can be viewed in figures \ref{seed-coat-color-main-figure},
@@
 
-We conducted two GWAS analyses on stem termination type. The first GWAS
+We conducted two different GWAS analyses on stem termination type. The first GWAS
 included all phenotypes observed in the dataset (determinate, semi-determinate,
 and indeterminate; Table \ref{STEM.TERMINATION.TYPE_gwas_table}). The results
-of this analysis and the \textit{Dt1} locus can be viewed in figures \ref{stem_termination_all_manhattan}
+of this analysis at the \textit{Dt1} locus can be viewed in figures \ref{stem_termination_all_manhattan},
 \ref{stem_termination_all_Dt1_signal}, \ref{stem_termination_all_Dt1_gene} and
@@
 
-The second GWAS analysis that we conducted on this trait included only
-semi-determinate and indeterminate phenotypes in an attempt to detect signals
+The second GWAS analysis included only
+semi-determinate and indeterminate phenotypes. We performed this analysis in an attempt to also detect signals
 at the \emph{Dt2} locus (Table \ref{STEM.TERMINATION.TYPE_gwas_table}).
@@
 at the \emph{I}, \emph{T} and \emph{R} loci using all four approaches except vg
-at the \emph{R} locus (Figure \ref{hilum_color_all_manhattan}) while no signals
-were detected at the \emph{O} and \emph{W1} loci studied by
+at the \emph{R} locus (Figure \ref{hilum_color_all_manhattan}). No signals
+were detected at the \emph{W1} and \emph{O} loci studied by
 \cite{bandillo2017}. The detailed results for the \textit{I} locus can be viewed
@@
 
-A second GWAS analysis contrasting accessions with either black or brown hilum
+The second GWAS analysis contrasting accessions with either black or brown hilum
 color was better suited at resolving the signal at the \emph{R} locus (Table
@@
 as well as Glyma.09g235100 and two others are part of a tandem array of four
-R2R3 MYB genes considered candidates by \cite{gillman2011}, and of which
+R2R3 MYB genes considered as candidates by \cite{gillman2011}, and of which
 Glyma.09g235100 was the only one expressed in the seed coat. Given the
@@
 \cite{gillman2011} in our dataset (Figure \ref{hilum_color_blackbrown_R_gene},
-\ref{hilum_color_blackbrown_R_kmers}), although they were not among the most
-significantly associated markers or \emph{k}-mers.
+\ref{hilum_color_blackbrown_R_kmers}).
 
@@
 
-We conducted two GWAS analyses on pod color. The first analysis included four
+We conducted two different GWAS analyses on pod color. The first analysis included four
 different colors (Table \ref{POD.COLOR_gwas_table}) and resulted in the
-detection of signals overlapping both the \emph{L1} and \emph{L2} loci with all
+detection of signals overlapping both the \emph{L1} and \emph{L2} loci using all
 four methods (Figure \ref{pod_color_all_manhattan}). Detailed results at the
@@
 
-The second GWAS contrasted black and brown pods with the objective of detecting
-signals that provided better resolution at the \emph{L1} locus (Table
+The second GWAS analysis compared black and brown pods with the objective of detecting stronger
+signals at the \emph{L1} locus (Table
 \ref{POD.COLOR_gwas_table}).  Accordingly, all four methods detected a signal
@@
 \ref{pod_color_blbr_ld}). The observed signal at this locus overlaps a total of
-27 genes (Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those was
+27 genes (Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those are
 annotated as having an obvious role in controlling pod color.
@@
 
-We conducted two different GWAS on pubescence form. The first GWAS used three
+We conducted two different GWAS analyses on pubescence form. The first analysis used three
 phenotypes (erect, semi-appressed and appressed; Table
@@
 \emph{Pa1} locus using all four approaches as well as signals overlapping the
-\emph{Pa2} locus using Platypus and the \emph{k}-mer approaches (Figure
+\emph{Pa2} locus using Platypus and the \emph{k}-mer approach (Figure
 \ref{pubescence_form_all_manhattan}). Detailed results at the \textit{Pa1} locus
@@
 \ref{pubescence_form_all_Pa1_gene} and \ref{pubescence_form_all_Pa1_kmers},
-whereas detailed results at the \textit{Pa2} locus can be viewed in figure
+whereas results at the \textit{Pa2} locus can be viewed in figure
 \ref{pubescence_form_all_Pa2_signal}.
@@
 
-A second GWAS analysis focused only on appressed and semi-appressed phenotypes
+The second GWAS analysis focused only on appressed and semi-appressed phenotypes
 with the objective of enhancing the signal at the \emph{Pa2} locus. However,
@@
 
-We conducted a single GWAS analysis on pubescence density (Table
+We conducted a GWAS analysis on pubescence density (Table
 \ref{PUBESCENCE.DENSITY_gwas_table}), which detected signals at the \emph{Ps}
@@
 (Figure \ref{pubescence_density_manhattan}). \cite{liu2020ps} linked the gene
-Glyma.12g187200 to the \emph{Ps} locus and identified copy number variation
-(CNV) of a 25.6-kb sequence including this gene as the causal variant, with
-higher copy number resulting in decreased pubescence. Through a segmentation
-analysis of the sequencing depth in accessions with increased copy number, we
-identified the boundaries of the region varying in copy number as 36,254,989 -
-36,297,368 on chromosome Gm12 of genome assembly version 4 of Williams82, for a
-total of about 42 kb. Detailed results at this locus can be viewed in figures
-\ref{pubescence_density_Ps_signal}, \ref{pubescence_density_Ps_gene}, and
-\ref{pubescence-density-main-figure}.
+Glyma.12g187200 to the \emph{Ps} locus and identified ia 25.6-kb copy number variation
+containing this gene as the causal variant, with
+higher copy number resulting in decreased pubescence. By
+analyzing the sequencing depth in accessions with increased copy number, we
+identified the boundaries of the CNV region as 36,254,989 -
+36,297,368 (42 kb) on chromosome Gm12.
+Detailed results at this locus can be viewed in figures \ref{pubescence-density-main-figure},
+\ref{pubescence_density_Ps_signal} and \ref{pubescence_density_Ps_gene}.
 
@@
 We performed three different GWAS analyses on seed coat luster (Table
-\ref{SEED.COAT.LUSTER_gwas_table}).  However, all three yielded similar results
+\ref{SEED.COAT.LUSTER_gwas_table}). Since all three analyses yielded similar results
 (Figure \ref{seed_coat_luster_all_manhattan},
 \ref{seed_coat_luster_nointermediate_manhattan},
-\ref{seed_coat_luster_dullshiny_manhattan}) so we chose to focus on the
-analysis contrasting the dull and shiny phenotypes as it yielded the most
-significant signals and also provided the most consistent results in the
-analysis previously done by \cite{bandillo2017}. This analysis detected
+\ref{seed_coat_luster_dullshiny_manhattan}), we chose to focus on the
+GWAS contrasting the dull and shiny phenotypes as it yielded the most
+significant signals. In addition, it also provided results that were most consistent with the
+analysis previously done by \cite{bandillo2017}. The analysis detected
 signals at the \emph{B} locus as the sole classical locus using all four
@@
 also detected signals overlapping the \emph{Hps} locus as defined by
-\cite{bandillo2017}, however results shown in \cite{gijzen2003} and our results
-suggest that \emph{Hps} and \emph{B} are the same locus. Using a segmentation analysis
+\cite{bandillo2017}, however results shown by \cite{gijzen2003} and our results
+suggest that \emph{Hps} and \emph{B} are the same locus. Using a sequencing depth-based segmentation analysis
 similar to the one used for the \emph{Ps} locus, we identified a 31-kb region
@@
 (Figures \ref{seed_coat_luster_dullshiny_B_signal},
-\ref{seed_coat_luster_dullshiny_B_gene}), however we were not able to link any
+\ref{seed_coat_luster_dullshiny_B_gene}). However we were not able to link any
 of the variants found to the presence of copy number variation.
@@
 
-A GWAS conducted on maturity group did not yield any significant signals
-overlapping the classical E1-E4 loci using any of the four GWAS approaches,
-although signals at the E1 locus did approach significance in the Platypus
+We conducted on a GWAS analysis on maturity group which did not yield any significant signals
+overlapping the classical E1-E4 loci using any of the four GWAS approaches.
+However, signals at the E1 locus did approach significance in the Platypus
 analysis (Table \ref{MATURITY.GROUP_gwas_table}, Figure \ref{maturity_group_manhattan}).
@@
 
-We conducted a single GWAS analysis on horizontal resistance to
+We conducted a GWAS analysis on horizontal resistance to
 \textit{Phytophthora sojae} using corrected dry weight (CDW) as a quantitative
@@
 
-We conducted a GWAS analysis on the seed oil content data obtained from the
+We conducted a GWAS analysis on seed oil content using data obtained from the
 GRIN database. While several seed oil content quantitative trait loci (QTL)
 have been reported due to high interest in this trait for soybean breeding
-\citep{chaudhary2015}, we focused on two major QTL on chromosomes Gm15 and Gm20
-that have often been reported in \emph{G. max}, and notably observed by
+\citep{chaudhary2015}, we focused on two major QTL on chromosomes Gm15 and Gm20.
+These loci have often been reported in \emph{G. max} and were notably observed by
 \cite{bandillo2015} on a collection of USDA lines using SoySNP50K chip
@@
 the two major QTL considered above and reported by \cite{bandillo2015} for seed
-oil content are also major QTL for seed protein content. We identified a signal
-at the locus located on chromosome Gm20 only, and only using the genotypes
+oil content are also major QTL for seed protein content. We only identified a signal
+at the locus located on chromosome Gm20, and only using the genotypes
 called by vg (Figure \ref{protein_manhattan}). However, the SV discovered as
 significantly associated was located about 700 kb from the gene cloned by
-\cite{fliege2022} for that locus (Glyma.20g085100) and therefore is unlikely to
-represent the causal variation at that locus.
+\cite{fliege2022} for that locus (Glyma.20g085100). It is therefore unlikely that this SV
+represents the causal variation at that locus.
 

### 2023-02-28 figures/ld_plot.R

@@ -23,11 +23,17 @@ names(kmer_positions) <- kmer_positions$kmer_canon
 # Sorting the k-mers in the LD matrix based on their genomic position
 gsorted_ld <- ld_sort(clustered_ld, kmer_positions, sort_param = "position")
 
+# The ylabel_pattern depends on the locus ; we do not want to clutter the figure
+ypattern <- "^Gm[0-9]{2}$"
+if(trait == "pod_color_blbr") ypattern <- "(Gm13|Gm15|Gm18|Gm19)"
+if(trait == "pubescence_form_all") ypattern <- "(Gm04|Gm12|Gm13|Gm15|Gm20)"
+if(trait == "seed_coat_luster_dullshiny") ypattern <- "(Gm09|Gm15|Gm20)"
+
 # Plotting the results using markers sorted by genomic position
 png(paste0("figures/", trait, "_ld.png"), width = 9, height = 10, units = "in", res = 400)
 
 grid.newpage()
-ld_plot(gsorted_ld, kmer_positions, top_legend = FALSE, ylabels = TRUE, ylabel_pattern = "^Gm[0-9]{2}$", fontsize = 10)
+ld_plot(gsorted_ld, kmer_positions, top_legend = FALSE, ylabels = TRUE, ylabel_pattern = ypattern, fontsize = 10)
 
 dev.off()
 

### 2023-02-28 additional_files/main_text.tex
Remove vg results from Table 1

@@
 			\csvreader[head to column names,
-			tabular = llllcccc,
-			table head = \toprule Trait & Locus & Gene & Causal variant & Platypus\tnote{a} & vg\tnote{b} & Paragraph\tnote{b} & \textit{k}-mers \\ \midrule,
+			tabular = llllccc,
+			table head = \toprule Trait & Locus & Gene & Causal variant & Platypus\tnote{a} & Paragraph\tnote{b} & \textit{k}-mers \\ \midrule,
 			table foot = \bottomrule]%
@@
 			{}%
-			{\Trait & \Locus & \Gene & \Causal & \platypus & \vg & \paragraph & \kmers}
+			{\Trait & \Locus & \Gene & \Causal & \platypus & \paragraph & \kmers}
 
@@
 			\item[a] Platypus genotypes SNPs and indels
-			\item[b] vg and Paragraph genotype SVs
+			\item[b] Paragraph genotypes SVs
 			\item[c] Locus not discussed in main text. See Supplemental data for detailed results.
@@
 			\item[]* The signal found at this locus overlaps the gene.
-			\item[]** The region defined by the top 5\% (vg and Paragraph) or 1\% (Platypus and \textit{k}-mers) variants or \textit{k}-mers overlaps the gene.
+			\item[]** The region defined by the top 5\% (Paragraph) or 1\% (Platypus and \textit{k}-mers) variants or \textit{k}-mers overlaps the gene.
 			\item[]*** The most significantly associated marker or \textit{k}-mer is the causal variant.

### 2023-02-28 additional_files/additional_file_1.tex
Remove vg results from table S11 (signals table)

@@
 		\item[c] Most significant -log\textsubscript{10}(\emph{p}) reported by Platypus at this locus. A dash indicates that no signal
-			was detected by Platypus at this locus. The same applies for other methods (vg, Paragraph, \emph{k}-mers) in
+			was detected by Platypus at this locus. The same applies for other methods (Paragraph, \emph{k}-mers) in
 			their respective columns.
@@
 
-		\begin{longtable}{llcclccccl}
+		\begin{longtable}{llcclcccl}
 
@@
 			{Description of loci associated with the traits studied and -log\textsubscript{10}(\emph{p}) of significant
-			signals detected in our study using four different genotypic datasets (Platypus, vg, Paragraph, \emph{k}-mers).}\\
+			signals detected in our study using four different genotypic datasets (Platypus, Paragraph, \emph{k}-mers).}\\
 			\label{signals_table}\\
@@
 			\toprule
-			Trait & Locus & Chromosome & -log\textsubscript{10}(\emph{p})\tnote{a} & Gene\tnote{b} & Platypus\tnote{c} & vg & Paragraph & \emph{k}-mers  & Study \\
+			Trait & Locus & Chromosome & -log\textsubscript{10}(\emph{p})\tnote{a} & Gene\tnote{b} & Platypus\tnote{c} & Paragraph & \emph{k}-mers  & Study \\
 			\midrule \endfirsthead
@@
 			\toprule
-			Trait & Locus & Chromosome & -log10(p) & Gene & Platypus & vg & Paragraph & k-mers  & Study \\
+			Trait & Locus & Chromosome & -log10(p) & Gene & Platypus & Paragraph & k-mers  & Study \\
 			\midrule
@@
 			/csv/separator=semicolon]%
-			{../tables/signals_table.csv}{}{\Trait & \Locus & \Chromosome & \Pvalues & \Gene & \Platypus & \Vg & \Paragraph & \Kmers & \Study}
+			{../tables/signals_table.csv}{}{\Trait & \Locus & \Chromosome & \Pvalues & \Gene & \Platypus & \Paragraph & \Kmers & \Study}
 

### 2023-02-28 additional_files/manuscript.tex
Remove vg results from the main Manhattan plots

@@
 	\includegraphics[width = 5.8in]{#1_manhattan}
-	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
-	(a) Platypus (SNPs and indels), (b) vg (SVs), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
+	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using three genotype datasets :
+	(a) Platypus (SNPs and indels), (b) Paragraph (SVs), (c) \emph{k}-mers presence/absence.
 	The x-axis shows the position along the reference assembly version 4 of Williams82.

### 2023-02-28 additional_files/additional_file_1.tex
Remove vg results from gene plots

@@
 \geneplot{flower_color_W1}{flower color}{\emph{W1}}{\getvar{flower_color_W1_gene} gene}%
-{Vertical dotted lines in panels (c), (d) and (e) indicate the location of the causal SV at this locus.}
+{Vertical dotted lines in panels (c) and (d) indicate the location of the causal SV at this locus.}
 %\kmerplot{flower_color_W1}
@@
 \geneplot{pubescence_color_all_T}{pubescence color (first GWAS)}{\emph{T}}{\getvar{pubescence_color_all_T_gene} gene}%
-{Vertical dotted lines in panels (b) and (e) indicate the location of the causal indel at this locus.}
+{Vertical dotted lines in panels (b) and (d) indicate the location of the causal indel at this locus.}
 \kmerplot{pubescence_color_all_T}{\getvar{pubescence_color_all_T_gene}}{\emph{T}}{pubescence color (first GWAS)}{a causal indel}
@@
 \geneplot{pubescence_color_nogray_Td}{pubescence color (second GWAS)}{\emph{Td}}{\getvar{pubescence_color_nogray_Td_gene} gene}%
-{Vertical dotted lines in panels (b) and (e) indicate the location of the causal SNP at this locus.}
+{Vertical dotted lines in panels (b) and (d) indicate the location of the causal SNP at this locus.}
 %\kmerplot{pubescence_color_nogray_Td}
@@
 \geneplot{seed_coat_color_greenyellow_G}{seed coat color (second GWAS)}{\emph{G}}{\getvar{seed_coat_color_greenyellow_G_gene} gene}%
-{Vertical dotted lines in panels (b) and (e) indicate the location of the causal SNP at this locus.}
+{Vertical dotted lines in panels (b) and (d) indicate the location of the causal SNP at this locus.}
 %\kmerplot{seed_coat_color_greenyellow_G}
@@
 \geneplot{stem_termination_all_Dt1}{stem termination type (first GWAS)}{\emph{Dt1}}{\getvar{stem_termination_all_Dt1_gene} gene}%
-{Vertical dotted lines in panels (b) and (e) indicate the location of a causal SNP at this locus.}
+{Vertical dotted lines in panels (b) and (d) indicate the location of a causal SNP at this locus.}
 \kmerplot{stem_termination_all_Dt1}{\getvar{stem_termination_all_Dt1_gene}}{\emph{Dt1}}{stem termination type (first GWAS)}{a causal SNP}
@@
 \geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene}%
-{Vertical dotted lines in panels (b) and (e) indicate the location of the causal indel at this locus.}
+{Vertical dotted lines in panels (b) and (d) indicate the location of the causal indel at this locus.}
 \kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene}}{\emph{T}}{hilum color (first GWAS)}{a causal indel}
@@
 \geneplot{hilum_color_blackbrown_R}{hilum color (second GWAS)}{\emph{R}}{\getvar{hilum_color_blackbrown_R_gene} gene}%
-{Vertical dotted lines indicate the location an indel (panels b and e) and a SNP (panel e) that are putatively causal at this locus.}
+{Vertical dotted lines indicate the location an indel (panels b and d) and a SNP (panel d) that are putatively causal at this locus.}
 \kmerplot{hilum_color_blackbrown_R}{\getvar{hilum_color_blackbrown_R_gene}}{\emph{R}}{hilum color (second GWAS)}{a SNP and an indel that are putatively causal}
@@
 \geneplot{pubescence_form_all_Pa1}{pubescence form (first GWAS)}{\emph{Pa1}}{\getvar{pubescence_form_all_Pa1_gene} candidate gene putatively}%
-{Vertical dotted lines in panels (b) and (e) indicate the location of two highly significant non-synonymous SNPs at this gene.}
+{Vertical dotted lines in panels (b) and (d) indicate the location of two highly significant non-synonymous SNPs at this gene.}
 \kmerplot{pubescence_form_all_Pa1}{\getvar{pubescence_form_all_Pa1_gene} candidate}{\emph{Pa1}}{pubescence form (first GWAS)}{a non-synonymous SNP}
@@
 \geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene.}%
-{The vertical dotted line in panel (e) indicates the location of the most significant \emph{k}-mer, which is associated with the causal CNV at this locus.
+{The vertical dotted line in panel (d) indicates the location of the most significant \emph{k}-mer, which is associated with the causal CNV at this locus.
 The shaded gray rectangles show the boundaries of the causal CNV.}

### 2023-02-28 additional_files/manuscript.tex
Remove vg results from gene plots

@@
 	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
-	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using four genotype datasets:
-	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
+	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using three genotype datasets:
+	(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
 	#5

### 2023-02-28 figures/gene_plot.R

@@ -19,11 +19,9 @@ target_signal <- readRDS("utilities/all_signals.rds")
 xchrom <- seqnames(target_signal[id])
 
 # DEPENDENCY: Paragraph gene subplot
-# DEPENDENCY: vg gene subplot
 # DEPENDENCY: Platypus gene subplot
 # DEPENDENCY: kmers gene subplot
 pvalue_grobs <- list(readRDS(paste0("figures/grobs/platypus_", id, "_gene.rds")),
-		     readRDS(paste0("figures/grobs/vg_", id, "_gene.rds")),
 		     readRDS(paste0("figures/grobs/paragraph_", id, "_gene.rds")),
 		     readRDS(paste0("figures/grobs/kmers_", id, "_gene.rds")))
 
@@ -47,14 +45,14 @@ grid.newpage()
 grid.draw(output_grob)
 
 # Add labels for the panels
-grid.text(c("(a)", "(b) Platypus", "(c) vg", "(d) Paragraph"),
+grid.text(c("(a)", "(b) Platypus", "(c) Paragraph"),
 	  x = unit(0.09, "npc"),
-	  y = unit(c(0.993, 0.8725, 0.645, 0.4175), "npc"),
+	  y = unit(c(0.993, 0.8725, 0.57), "npc"),
 	  hjust = 0,
 	  gp = gpar(fontsize = 14))
-grid.text(expression(paste("(e) ", italic(k)-mers)),
+grid.text(expression(paste("(d) ", italic(k)-mers)),
 	  x = unit(0.09, "npc"),
-	  y = unit(0.19, "npc"),
+	  y = unit(0.2675, "npc"),
 	  hjust = 0,
 	  gp = gpar(fontsize = 14))
 

### 2023-02-28 additional_files/manuscript.tex
Remove vg results from signal plots

@@
 	\includegraphics[width = 5.8in]{#1_signal}
-	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
-	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
+	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using three genotype datasets:
+	(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
 	Panel (a) shows gene models over the genomic interval.
@@
 	#4
-	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph and vg) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
+	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
 	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose

### 2023-02-28 figures/signal_plot.R

@@ -31,11 +31,9 @@ if(!is.na(gene_name) && grepl(";", gene_name)) {
 }
 
 # DEPENDENCY: Paragraph signal subplot
-# DEPENDENCY: vg signal subplot
 # DEPENDENCY: Platypus signal subplot
 # DEPENDENCY: kmers signal subplot
 pvalue_grobs <- list(readRDS(paste0("figures/grobs/platypus_", id, "_signal.rds")),
-		     readRDS(paste0("figures/grobs/vg_", id, "_signal.rds")),
 		     readRDS(paste0("figures/grobs/paragraph_", id, "_signal.rds")),
 		     readRDS(paste0("figures/grobs/kmers_", id, "_signal.rds")))
 
@@ -59,14 +57,14 @@ grid.newpage()
 grid.draw(output_grob)
 
 # Add labels for the panels
-grid.text(c("(a)", "(b) Platypus", "(c) vg", "(d) Paragraph"),
+grid.text(c("(a)", "(b) Platypus", "(c) Paragraph"),
 	  x = unit(0.09, "npc"),
-	  y = unit(c(0.993, 0.8725, 0.645, 0.4175), "npc"),
+	  y = unit(c(0.993, 0.8725, 0.57), "npc"),
 	  hjust = 0,
 	  gp = gpar(fontsize = 14))
-grid.text(expression(paste("(e) ", italic(k)-mers)),
+grid.text(expression(paste("(d) ", italic(k)-mers)),
 	  x = unit(0.09, "npc"),
-	  y = unit(0.19, "npc"),
+	  y = unit(0.2675, "npc"),
 	  hjust = 0,
 	  gp = gpar(fontsize = 14))
 

### 2023-02-28 additional_files/additional_file_1.tex
Adjust the text to remove any mention of vg

@@
 signals overlapping the \emph{I} locus by all four approaches (Figure
-\ref{seed_coat_color_all_manhattan}). All approaches but vg also detected
+\ref{seed_coat_color_all_manhattan}). All approaches also detected
 signals at the \emph{G} locus. In addition, the \emph{k}-mer approach also
@@
 included seven colors (Table \ref{HILUM.COLOR_gwas_table}) and detected signals
-at the \emph{I}, \emph{T} and \emph{R} loci using all four approaches except vg
-at the \emph{R} locus (Figure \ref{hilum_color_all_manhattan}). No signals
+at the \emph{I}, \emph{T} and \emph{R} loci using all four approaches
+(Figure \ref{hilum_color_all_manhattan}). No signals
 were detected at the \emph{W1} and \emph{O} loci studied by
@@
 \ref{seed_coat_luster_dullshiny_B_gene}). The top 5\% / 1\% significant
-associations of all approaches other than vg overlapped this CNV region and the
+associations of all approaches overlapped this CNV region and the
 most significantly associated \emph{k}-mer was located within the CNV region
@@
 the two major QTL considered above and reported by \cite{bandillo2015} for seed
-oil content are also major QTL for seed protein content. We only identified a signal
-at the locus located on chromosome Gm20, and only using the genotypes
-called by vg (Figure \ref{protein_manhattan}). However, the SV discovered as
-significantly associated was located about 700 kb from the gene cloned by
-\cite{fliege2022} for that locus (Glyma.20g085100). It is therefore unlikely that this SV
-represents the causal variation at that locus.
+oil content are also major QTL for seed protein content. However, none of the
+approaches detected a signal at this locus.
 

### 2023-02-28 additional_files/main_text.tex
Adjust the text to remove any mention of vg

@@
 by delimiting a range of coordinates containing the top 5\% most significant
-associations (Paragraph and vg) or 1\% most significant associations (Platypus
+associations (Paragraph) or 1\% most significant associations (Platypus
 and \emph{k}-mers) found within the signal. In cases where such a fraction
@@
 Because two or three different GWAS analyses were conducted for some of the traits, a
-total of 22 GWAS analyses were performed using each of the four approaches.
+total of 22 GWAS analyses were performed using each of the all approaches.
 Given the large volume of data that this represents, we will focus here on the
@@
 	\begin{table}
-		\caption[Performance of four GWAS approaches at detecting causal genes/variants]%
-		{Performance of four GWAS approaches at detecting causal genes or variants at cloned loci.}
+		\caption[Performance of three GWAS approaches at detecting causal genes/variants]%
+		{Performance of three GWAS approaches at detecting causal genes or variants at cloned loci.}
 		\label{loci-table}
@@
 these differences in color. Accordingly, we
-detected signals in this region using all four approaches, but only Platypus
+detected signals in this region using all three approaches, but only Platypus
 and the \textit{k}-mers identified a signal that overlapped Glyma.03g258700
@@
 The \textit{T} locus mentioned above for pubescence color is also a determining
-locus for hilum color, and we detected signals overlapping the corresponding gene
-using all approaches but vg (Figures \ref{hilum_color_all_manhattan},
+locus for hilum color. We detected signals overlapping the corresponding gene
+using all approaches (Figures \ref{hilum_color_all_manhattan},
 \ref{hilum_color_all_T_signal}). The regions defined by the top 5\% (Paragraph)
@@
 measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
-overlapping the genomic region reported by these authors on Gm15 using all four GWAS
+overlapping the genomic region reported by these authors on Gm15 using all all GWAS
 approaches (Figure \ref{corrected_dry_weight_manhattan}). \cite{deronne2022}

### 2023-02-28 additional_files/additional_file_1.tex
Adjust the contents of the main and supplementary results to reduce redundancy ; figure order has yet to be adjusted accordingly

@@
 % Detailed results that are not meant to fit the main paper
-\section*{Detailed results}
+\section*{Supplemental results}
 \label{annexe-sv-gwas-detailed-results}
 
-This section provides a more detailed per-trait account of some of the results obtained
-by the GWAS analyses. Aspects of the results that were
-already discussed in the main text are not repeated here; in such cases,
-readers are simply redirected to specific tables and figures containing
-detailed results from all four GWAS approaches. A summary of the signals found
-at all loci using all four methods is presented in Table \ref{signals_table}.
+This section describes results at additional loci that were not discussed in
+the main text. A summary of the signals found at all loci using all four
+methods is presented in Table \ref{signals_table}.
 
-\subsection*{Flower color}
-\label{sv-gwas-flower-color}
+\subsection*{Flower color -- \textit{L1} locus}
+\label{annexe-sv-gwas-flower-color-l1}
 
-We conducted a GWAS analysis contrasting purple and white flower colors
-(Table \ref{FLOWER.COLOR_gwas_table}). Of the two classical loci reported by
-\cite{bandillo2017}, \textit{W1} was detected by all four GWAS approaches whereas
-\textit{L1} was detected by none of the approaches (Figure
-\ref{flower_color_manhattan}). These results are unsurprising as W1 is the
-major contributor to flower color in soybean, whereas the \textit{L1} locus
+In addition to the \textit{W1} locus, \cite{bandillo2017} reported a signal for
+flower color near the \textit{L1} locus.  None of the GWAS approaches that we
+tested found a signal for flower color near this locus (Figure
+\ref{flower_color_manhattan}). This result is unsurprising as W1 is the major
+contributor to flower color in soybean, whereas the \textit{L1} locus
 (typically associated with pod color) was detected with a much less significant
 \textit{p}-value (see Table \ref{signals_table}) and only conjectured as
-associated with flower color by \cite{bandillo2017}. Detailed results for the
-\textit{W1} locus can be viewed in figures \ref{flower-color-main-figure},
-\ref{flower_color_W1_signal} and \ref{flower_color_W1_gene}.
+associated with flower color by \cite{bandillo2017}.
 
-The \textit{k}-mer GWAS analysis also identified presumably spurious significant
-\textit{k}-mers scattered throughout the genome (Figure
-\ref{flower_color_manhattan}). An analysis of the pairwise LD between a subset
-of those \textit{k}-mers showed that most \textit{k}-mers formed a single LD block
-linked to the \textit{W1} locus (Figure \ref{flower_color_ld}). A small LD block
-occurred on chromosome Gm09, however it was in moderate LD with the one on chromosome
-Gm13 and therefore probably does not represent a separate locus.
+\subsection*{Pubescence color -- locus on chromosome Gm16}
+\label{annexe-sv-gwas-pubescence-color-gm16}
 
-\subsection*{Pubescence color}
-\label{sv-gwas-pubescence-color}
-
-We conducted two different GWAS analyses on pubescence color. The first analysis
-contrasted gray, tawny and light tawny pubescence colors (Table
-\ref{PUBESCENCE.COLOR_gwas_table}) and detected a signal overlapping the
-\emph{T} locus using all four approaches (Figure
-\ref{pubescence_color_all_manhattan}). Detailed results at the \textit{T} locus
-can be viewed in figures \ref{pubescence_color_all_T_signal},
-\ref{pubescence_color_all_T_gene} and \ref{pubescence_color_all_T_kmers}. We
-observed that not all accessions with gray pubescence harbored the deleted
-nucleotide, which could indicate either that other causal variants are resulting in
-a non-functional F3’H in this population, or that the causal
-variant lies elsewhere.
-
-This first GWAS analysis also revealed a second locus on chromosome Gm16
+In addition to the \textit{T} locus, the first GWAS analysis conducted on pubescence color
+revealed a second locus on chromosome Gm16
 (Figure \ref{pubescence_color_all_manhattan}). Upon closer analysis, this locus
@@
 
-In the presence of a functional allele at the \emph{T} locus, the genotype at
-the \emph{Td} locus influences the exact shade of pubescence color.  Therefore,
-we conducted a second GWAS analysis contrasting accessions with tawny and light
-tawny pubescence in order to ignore the epistatic effect of the \emph{T} locus
-and instead focus on the effect of the \emph{Td} locus on pubescence color
-(Table \ref{PUBESCENCE.COLOR_gwas_table}). Detailed results at this locus
-can be viewed in figures \ref{pubescence-color-main-figure}, \ref{pubescence_color_nogray_Td_signal} and
-\ref{pubescence_color_nogray_Td_gene}).
+\subsection*{Seed coat color -- loci \textit{T}, \textit{O} and \textit{R}}
+\label{annexe-sv-gwas-seed-coat-color-t-o}
 
-\subsection*{Seed coat color}
-\label{sv-gwas-seed-coat-color}
+In addition to the \textit{I} and \textit{G} loci, the \textit{k}-mer-based
+GWAS analysis of five different seed coat color phenotypes (first GWAS in Table
+\ref{SEED.COAT.COLOR_gwas_table}) detected weak signals overlapping the
+\emph{T} and \emph{O} loci (Figure \ref{seed_coat_color_all_manhattan}c).
+However, we interpreted those results as chance associations given the low
+\textit{p}-value associated with the \emph{T} locus and the proximity of the
+\emph{O} locus to the \emph{I} locus. None of the three GWAS approaches detected
+a signal overlapping the \textit{R} locus.
 
-We conducted two different GWAS analyses on seed coat color. The first GWAS
-included five different phenotypes observed in the dataset (Table
-\ref{SEED.COAT.COLOR_gwas_table}) and resulted in the detection of strong
-signals overlapping the \emph{I} locus by all four approaches (Figure
-\ref{seed_coat_color_all_manhattan}). All approaches also detected
-signals at the \emph{G} locus. In addition, the \emph{k}-mer approach also
-detected weak signals overlapping the \emph{T} and \emph{O} loci. However, the
-\emph{T} and \emph{O} loci are likely chance associations given the low \textit{p}-value
-associated with the \emph{T} locus and the proximity of the \emph{O} locus to
-the \emph{I} locus. Detailed results of this first GWAS and the \textit{I}
-locus can be viewed in figures \ref{seed_coat_color_all_manhattan} and
-\ref{seed_coat_color_all_I_signal}.
-
-The second GWAS analysis on seed coat color focused on the subset of accessions
-with yellow or green seed coat only (Table \ref{SEED.COAT.COLOR_gwas_table}),
-as this difference is controlled by the \emph{G} locus. Detailed results of
-this analysis can be viewed in figures \ref{seed-coat-color-main-figure},
-\ref{seed_coat_color_greenyellow_manhattan},
-\ref{seed_coat_color_greenyellow_G_signal} and
-\ref{seed_coat_color_greenyellow_G_gene}.
-
-\subsection*{Stem termination type}
+\subsection*{Stem termination type -- loci \textit{Dt2} and \textit{E3}}
 \label{sv-gwas-stem-termination}
 
-We conducted two different GWAS analyses on stem termination type. The first GWAS
-included all phenotypes observed in the dataset (determinate, semi-determinate,
-and indeterminate; Table \ref{STEM.TERMINATION.TYPE_gwas_table}). The results
-of this analysis at the \textit{Dt1} locus can be viewed in figures \ref{stem_termination_all_manhattan},
-\ref{stem_termination_all_Dt1_signal}, \ref{stem_termination_all_Dt1_gene} and
-\ref{stem_termination_all_Dt1_kmers}.
+We performed a GWAS analysis comparing only semi-determinate and indeterminate
+stem termination phenotypes in an attempt to detect signals at the \emph{Dt2}
+locus (Table \ref{STEM.TERMINATION.TYPE_gwas_table}).  However, this was
+unsuccessful using all three GWAS approaches (Figure
+\ref{stem_termination_sn_manhattan}). Similarly, we failed to detect signals at
+the \textit{E3} locus, which is more commonly associated with maturity but was
+mentioned as associated with stem termination type by \cite{bandillo2017}.
 
-The second GWAS analysis included only
-semi-determinate and indeterminate phenotypes. We performed this analysis in an attempt to also detect signals
-at the \emph{Dt2} locus (Table \ref{STEM.TERMINATION.TYPE_gwas_table}).
-However, this was unsuccessful using all four GWAS
-approaches (Figure \ref{stem_termination_sn_manhattan}). Similarly, we failed to
-detect signals at the \textit{E3} locus, which is more commonly associated
-with maturity but was mentioned as associated with stem termination type by
-\cite{bandillo2017}.
-
-\subsection*{Hilum color}
+\subsection*{Hilum color -- loci \textit{W1} and \textit{O}}
 \label{sv-gwas-hilum-color}
 
-We conducted three different GWAS analyses on hilum color. The first GWAS
-included seven colors (Table \ref{HILUM.COLOR_gwas_table}) and detected signals
-at the \emph{I}, \emph{T} and \emph{R} loci using all four approaches
-(Figure \ref{hilum_color_all_manhattan}). No signals
-were detected at the \emph{W1} and \emph{O} loci studied by
-\cite{bandillo2017}. The detailed results for the \textit{I} locus can be viewed
-in figure \ref{hilum_color_all_I_signal} whereas those for the \textit{T} locus
-can be viewed in figures \ref{hilum_color_all_T_signal}, \ref{hilum_color_all_T_gene}
-and \ref{hilum_color_all_T_kmers}.
-
-Although \cite{bandillo2017} listed the \textit{W1} locus (typically associated
-with flower color) as putatively associated with hilum color, we found no
-evidence of association in our results.
-
-The second GWAS analysis contrasting accessions with either black or brown hilum
-color was better suited at resolving the signal at the \emph{R} locus (Table
-\ref{HILUM.COLOR_gwas_table}). Accordingly, this GWAS analysis detected strong
-signals overlapping the \emph{R} locus and no other classical loci using all
-four approaches (Figure \ref{hilum_color_blackbrown_manhattan}).
-\cite{gillman2011} identified the R2R3 MYB transcription factor Glyma.09g235100
-as the molecular basis for the \emph{R} locus and documented four different
-loss-of-function alleles linked to brown hilum color. Although our analyses did
-detect strong signals in the vicinity of this gene, the most associated markers
-were not located in the body of the gene (Figure
-\ref{hilum_color_blackbrown_R_signal}). Interestingly, the most significantly
-associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
-another gene (Glyma.09g234900) annotated as a MYB transcription factor and
-putatively involved in the regulation of anthocyanin biosynthesis. This gene
-as well as Glyma.09g235100 and two others are part of a tandem array of four
-R2R3 MYB genes considered as candidates by \cite{gillman2011}, and of which
-Glyma.09g235100 was the only one expressed in the seed coat. Given the
-convincing evidence provided by \cite{gillman2011}, it is most likely that
-Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
-our failure to identify the most significantly associated \emph{k}-mers as
-causal variants is due to the coexistence of several variants that can cause
-loss of function at the \emph{R} locus. Indeed, we observed at least two causal
-variants (a 1-bp deletion and a SNP) previously documented by
-\cite{gillman2011} in our dataset (Figure \ref{hilum_color_blackbrown_R_gene},
-\ref{hilum_color_blackbrown_R_kmers}).
-
-We conducted a third GWAS focusing on the brown and reddish brown hilum color
-phenotypes in order to detect a signal at the \emph{O} locus, however this GWAS
-did not detect associations at any of the classical loci (Figure
+None of the three GWAS analyses that we conducted on hilum color found signals
+at the \emph{W1} and \emph{O} loci studied by \cite{bandillo2017} (Figures
+\ref{hilum_color_all_manhattan}, \ref{hilum_color_blackbrown_manhattan},
 \ref{hilum_color_rbr_manhattan}).
 
-\subsection*{Pod color}
-\label{sv-gwas-pod-color}
-
-We conducted two different GWAS analyses on pod color. The first analysis included four
-different colors (Table \ref{POD.COLOR_gwas_table}) and resulted in the
-detection of signals overlapping both the \emph{L1} and \emph{L2} loci using all
-four methods (Figure \ref{pod_color_all_manhattan}). Detailed results at the
-\textit{L1} locus can be viewed in figure \ref{pod_color_all_L1_signal},
-whereas detailed results at the \textit{L2} locus can be viewed in figure
-\ref{pod_color_all_L2_signal}.
-
-The second GWAS analysis compared black and brown pods with the objective of detecting stronger
-signals at the \emph{L1} locus (Table
-\ref{POD.COLOR_gwas_table}).  Accordingly, all four methods detected a signal
-at the \emph{L1} locus but not at the \emph{L2} locus using this GWAS (Figure
-\ref{pod_color_blbr_manhattan}).  Results at the \emph{L1} locus were similar
-to those observed when using all observed phenotypes, but the \emph{k}-mer
-analysis revealed a previously undocumented signal on chromosome Gm15 which was
-not in LD with any other signal (Figures \ref{pod_color_blbr_manhattan},
-\ref{pod_color_blbr_ld}). The observed signal at this locus overlaps a total of
-27 genes (Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those are
-annotated as having an obvious role in controlling pod color.
-
-\subsection*{Pubescence form}
-\label{sv-gwas-pubesence-form}
-
-We conducted two different GWAS analyses on pubescence form. The first analysis used three
-phenotypes (erect, semi-appressed and appressed; Table
-\ref{PUBESCENCE.FORM_gwas_table}) and detected signals overlapping the
-\emph{Pa1} locus using all four approaches as well as signals overlapping the
-\emph{Pa2} locus using Platypus and the \emph{k}-mer approach (Figure
-\ref{pubescence_form_all_manhattan}). Detailed results at the \textit{Pa1} locus
-can be viewed in figure \ref{pubescence_form_all_Pa1_signal},
-\ref{pubescence_form_all_Pa1_gene} and \ref{pubescence_form_all_Pa1_kmers},
-whereas results at the \textit{Pa2} locus can be viewed in figure
-\ref{pubescence_form_all_Pa2_signal}.
-
-Signals were also observed on chromosomes Gm04 and Gm15 by the Platypus and
-\emph{k}-mer analyses (Figure \ref{pubescence_form_all_manhattan}), however
-upon further analysis of LD patterns they appeared to be linked to the
-\emph{Pa1} locus (Figure \ref{pubescence_form_all_ld}).
-
-The second GWAS analysis focused only on appressed and semi-appressed phenotypes
-with the objective of enhancing the signal at the \emph{Pa2} locus. However,
-this analysis only detected weak signals at the \emph{Pa1} locus (Figure
-\ref{pubescence_form_noerect_manhattan}) and its results were therefore not
-analyzed further.
-
-\subsection*{Pubescence density}
+\subsection*{Pubescence density -- loci \textit{Pd1} and \textit{P1}}
 \label{sv-gwas-pubescence-density}
 
-We conducted a GWAS analysis on pubescence density (Table
-\ref{PUBESCENCE.DENSITY_gwas_table}), which detected signals at the \emph{Ps}
-locus but not the \emph{Pd1} and \emph{P1} loci using all four approaches
-(Figure \ref{pubescence_density_manhattan}). \cite{liu2020ps} linked the gene
-Glyma.12g187200 to the \emph{Ps} locus and identified ia 25.6-kb copy number variation
-containing this gene as the causal variant, with
-higher copy number resulting in decreased pubescence. By
-analyzing the sequencing depth in accessions with increased copy number, we
-identified the boundaries of the CNV region as 36,254,989 -
-36,297,368 (42 kb) on chromosome Gm12.
-Detailed results at this locus can be viewed in figures \ref{pubescence-density-main-figure},
-\ref{pubescence_density_Ps_signal} and \ref{pubescence_density_Ps_gene}.
+Our GWAS analysis on pubescence density did not detect any signals at the
+\emph{Pd1} and \emph{P1} (Figure \ref{pubescence_density_manhattan}). Our
+failure to detect these loci is likely due to our small population size
+compared to \cite{bandillo2017} and to the smaller contribution of those loci
+to pubescence density in our population.
 
-\subsection*{Seed coat luster}
+\subsection*{Seed coat luster -- loci \textit{Hps} \textit{B?} and \textit{I}}
 \label{sv-gwas-seed-coat-luster}
@@
 (Figure \ref{seed_coat_luster_all_manhattan},
-\ref{seed_coat_luster_nointermediate_manhattan},
+\ref{seed_coat_luster_nointermediate_manhattan} and
 \ref{seed_coat_luster_dullshiny_manhattan}), we chose to focus on the
@@
 significant signals. In addition, it also provided results that were most consistent with the
-analysis previously done by \cite{bandillo2017}. The analysis detected
-signals at the \emph{B} locus as the sole classical locus using all four
-methods (Figure \ref{seed_coat_luster_dullshiny_manhattan}). Apparent signals
-on chromosomes Gm20 and Gm09 (which may correspond to the B? locus reported by
-\cite{bandillo2017}) were also detected, but analysis of the LD patterns
-between \emph{k}-mers suggested that these signals were most likely in LD with
-the \emph{B} locus (Figure \ref{seed_coat_luster_dullshiny_ld}). All methods
-also detected signals overlapping the \emph{Hps} locus as defined by
-\cite{bandillo2017}, however results shown by \cite{gijzen2003} and our results
-suggest that \emph{Hps} and \emph{B} are the same locus. Using a sequencing depth-based segmentation analysis
-similar to the one used for the \emph{Ps} locus, we identified a 31-kb region
-spanning positions 9,386,109-9,417,431 on chromosome Gm15 that exhibited
-variation in copy number across accessions and overlapped four genes annotated
-as containing HPS domains (Figures \ref{seed_coat_luster_dullshiny_B_signal},
-\ref{seed_coat_luster_dullshiny_B_gene}). The top 5\% / 1\% significant
-associations of all approaches overlapped this CNV region and the
-most significantly associated \emph{k}-mer was located within the CNV region
-(Figures \ref{seed_coat_luster_dullshiny_B_signal},
-\ref{seed_coat_luster_dullshiny_B_gene}). However we were not able to link any
-of the variants found to the presence of copy number variation.
+analysis previously done by \cite{bandillo2017}.
 
-% Seed coat luster - I locus
+The analysis detected signals at the \emph{B} locus as the sole classical locus
+using all four methods (see main text and Figure
+\ref{seed_coat_luster_dullshiny_manhattan}).  All methods also detected signals
+overlapping the \emph{Hps} locus as defined by \cite{bandillo2017}, however
+results shown by \cite{gijzen2003} and our results suggest that \emph{Hps} and
+\emph{B} are the same locus.
+
+Apparent signals on chromosomes Gm20 and Gm09 (which may correspond to the B?
+locus reported by \cite{bandillo2017}) were also detected, but analysis of the
+LD patterns between \emph{k}-mers suggested that these signals are most likely
+in LD with the \emph{B} locus (Figure \ref{seed_coat_luster_dullshiny_ld}).
 
@@
 
-\subsection*{Maturity group}
-\label{sv-gwas-maturity-group}
+\subsubsection*{Maturity group}
+\label{annexe-sv-gwas-maturity}
 
-We conducted on a GWAS analysis on maturity group which did not yield any significant signals
-overlapping the classical E1-E4 loci using any of the four GWAS approaches.
-However, signals at the E1 locus did approach significance in the Platypus
-analysis (Table \ref{MATURITY.GROUP_gwas_table}, Figure \ref{maturity_group_manhattan}).
+We did not detect any signals associated with maturity using our dataset
+(Figure \ref{maturity_group_manhattan}). The use of maturity group as a proxy
+for time to maturity instead of the number of days, combined with the limited
+phenotypic variation of this trait in this collection of accessions (Table
+\ref{MATURITY.GROUP_gwas_table}), likely explains our failure to detect
+any signals.
 
-\subsection*{Resistance to \emph{Phytophthora sojae}}
-\label{sv-gwas-resistance}
-
-We conducted a GWAS analysis on horizontal resistance to
-\textit{Phytophthora sojae} using corrected dry weight (CDW) as a quantitative
-measure of resistance. Detailed results can be viewed in figures
-\ref{corrected_dry_weight_manhattan} and
-\ref{corrected_dry_weight_cdwGm15_signal}.
-
-\subsection*{Seed oil content}
-\label{sv-gwas-seed-oil-content}
+\subsection*{Seed oil and protein content}
+\label{sv-gwas-seed-oil-protein-content}
 
@@
 
-\subsection*{Seed protein content}
-\label{sv-gwas-seed-protein-content}
-
 Similarly to seed oil content, we conducted a GWAS on seed protein content
@@
 approaches detected a signal at this locus.
+
+Our failure to detect signals overlapping the known genes at these two loci for
+oil and protein is probably due to the fact that the oil and protein content
+data that we used for GWAS were not obtained from orthogonal trials, which was
+a clear limitation to this GWAS given our relatively small sample size.
+
+\subsection*{Pubescence form -- additional loci}
+\label{sv-gwas-pubesence-form}
+
+The first GWAS analysis on pubescence form found were signals on chromosomes Gm04 and Gm15 with Platypus
+\emph{k}-merss (Figure \ref{pubescence_form_all_manhattan}). However,
+upon further analysis of LD patterns, they appeared to be linked to the
+\emph{Pa1} locus (Figure \ref{pubescence_form_all_ld}).
 

### 2023-02-28 additional_files/main_text.tex
Adjust the contents of the main and supplementary results to reduce redundancy ; figure order has yet to be adjusted accordingly

@@
 
+The \textit{k}-mer GWAS analysis also identified presumably spurious significant
+\textit{k}-mers scattered throughout the genome (Figure
+\ref{flower_color_manhattan}). An analysis of the pairwise LD between a subset
+of those \textit{k}-mers showed that most \textit{k}-mers formed a single LD block
+linked to the \textit{W1} locus (Figure \ref{flower_color_ld}). A small LD block
+occurred on chromosome Gm09, however it was in moderate LD with the one on chromosome
+Gm13 and therefore probably does not represent a separate locus.
+
 We should note that when we first analyzed the results of the GWAS on flower
@@
 corresponded to the causal variant (Figures \ref{pubescence_color_all_T_gene},
-\ref{pubescence_color_all_T_kmers}). As for Platypus, the most significant
+\ref{pubescence_color_all_T_kmers}). We
+observed that not all accessions with gray pubescence harbored the deleted
+nucleotide (Figure \ref{pubescence_color_all_T_kmers}c), which could indicate either that other causal variants are resulting in
+a non-functional F3’H in this population, or that the causal
+variant lies elsewhere. As for Platypus, the most significant
 variant within the sequence of Glyma.06g202300 was the causal
@@
 
-We conducted a GWAS analysis that only compared accessions with black and brown hila
-in an attempt to detect stronger signals at the \textit{R} locus.  We
-accordingly detected a signal overlapping this locus using all approaches,
-but signals overlapped the Glyma.09g235100 gene associated with the
-\textit{R} locus only for Platypus and \textit{k}-mers (Figures
+We conducted a GWAS analysis that only compared accessions with black and brown
+hila in an attempt to detect stronger signals at the \textit{R} locus.  We
+accordingly detected a signal overlapping this locus using all approaches, but
+only Platypus and \textit{k}-mers detected signals overlapping the
+Glyma.09g235100 gene associated with \textit{R} (Figures
 \ref{hilum_color_blackbrown_manhattan}, \ref{hilum_color_blackbrown_R_signal}).
-We observed two causal variants previously documented by \cite{gillman2011},
-but these were not among the most significantly associated variants or
-\textit{k}-mers (Figures \ref{hilum_color_blackbrown_R_gene},
-\ref{hilum_color_blackbrown_R_kmers}).  This failure to identify causal
-variants more directly could be due to the existence of more than one causal
-variant at this locus. Interestingly, the most significantly associated
-\textit{k}-mers mapped just upstream of a gene (Glyma.09g234900) identified as
-a homolog of Glyma.09g235100 \citep{gillman2011}.
+\cite{gillman2011} identified this R2R3 MYB transcription factor as the
+molecular basis for the \textit{R} locus and documented four different
+loss-of-function alleles linked to brown hilum color. Although our analyses did
+detect strong signals in the vicinity of this gene, the most associated markers
+were not located in the body of the gene (Figure
+\ref{hilum_color_blackbrown_R_signal}).  Interestingly, the most significantly
+associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
+another gene (Glyma.09g234900) annotated as a MYB transcription factor and
+putatively involved in the regulation of anthocyanin biosynthesis. This gene as
+well as Glyma.09g235100 and two others are part of a tandem array of four R2R3
+MYB genes considered as candidates by \cite{gillman2011}, and of which
+Glyma.09g235100 was the only one expressed in the seed coat. Given the
+convincing evidence provided by \cite{gillman2011}, it is most likely that
+Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
+our failure to identify the most significantly associated \emph{k}-mers as
+causal variants is due to the coexistence of several variants that can cause
+loss of function at the \emph{R} locus.  Still, we observed two causal variants
+previously documented by \cite{gillman2011} in our dataset (Figures
+\ref{hilum_color_blackbrown_R_gene}, \ref{hilum_color_blackbrown_R_kmers}).
 
@@
 detected signals at the \textit{Ps} locus using all three approaches (Figures
-\ref{pubescence-density-main-figure}a-b, \ref{pubescence_density_manhattan}). The causal
+\ref{pubescence-density-main-figure}a-b, \ref{pubescence_density_manhattan}, \ref{pubescence_density_Ps_signal}
+and \ref{pubescence_density_Ps_gene}). The causal
 variant at this recently cloned locus is a copy number variant overlapping the
@@
 
-Next, we conducted a GWAS analysis comparing dull and shiny seed coat phenotypes
-for the seed coat luster trait. Seed coat luster has been shown by \cite{gijzen1999}
-to be largely caused by the deposition of a hydrophobic protein (HPS)
-at the seed surface. Copy number variation of this gene at the \textit{B}
-locus was later shown to explain variation in luster, although sequence
-variation in the HPS sequence may also play a role \citep{gijzen2006}.
-We found signals overlapping the causal CNV using all approaches, and the
-most significantly associated \textit{k}-mer was notably located
-within the CNV region (Figures \ref{seed_coat_luster_dullshiny_manhattan},
-\ref{seed_coat_luster_dullshiny_B_signal}). However, we were not able
-to directly link any of the variants or \textit{k}-mers identified to the causal
+Next, we conducted a GWAS analysis comparing dull and shiny seed coat
+phenotypes for the seed coat luster trait. Seed coat luster has been shown by
+\cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
+protein (HPS) at the seed surface. Copy number variation of this gene at the
+\textit{B} locus was later shown to explain variation in luster, although
+sequence variation in the HPS sequence may also play a role \citep{gijzen2006}.
+Using a sequencing depth-based analysis, we identified a 31-kb region spanning
+positions 9,386,109-9,417,431 on chromosome Gm15 that exhibited variation in
+copy number across accessions. This CNV overlapped four genes annotated as
+containing HPS domains and therefore likely represented the causal variant
+(Figures \ref{seed_coat_luster_dullshiny_B_signal},
+\ref{seed_coat_luster_dullshiny_B_gene}).  We found signals overlapping this
+CNV using all approaches, and the most significantly associated \textit{k}-mer
+was notably located within the CNV region (Figures
+\ref{seed_coat_luster_dullshiny_manhattan},
+\ref{seed_coat_luster_dullshiny_B_signal}). However, we were not able to
+directly link any of the variants or \textit{k}-mers identified to the causal
 CNV.
-
-% Maturity group loci
-\subsubsection*{Maturity}
-\label{sv-gwas-main-results-maturity}
-
-We did not detect any signals associated with maturity using our dataset
-(Figure \ref{maturity_group_manhattan}). The use of maturity group as a proxy
-for time to maturity instead of the number of days, combined with the limited
-phenotypic variation of this trait in this collection of accessions (Table
-\ref{MATURITY.GROUP_gwas_table}), likely explains our failure to detect
-any signals.
-
-% Seed oil and protein
-\subsubsection*{Seed oil and protein content}
-\label{sv-gwas-main-results-oil-protein}
-
-We expected to detect signals at two loci controlling seed oil and protein
-content on chromosomes 15 and 20 \citep{bandillo2015}. However, none of our
-analyses detected signals overlapping the genes cloned for these loci (Figures
-\ref{oil_manhattan}, \ref{protein_manhattan}). This is probably due to the fact
-that the oil and protein content data that we used for GWAS were not obtained
-from orthogonal trials, which was a clear limitation to this GWAS given our
-relatively small sample size.
 
@@
 as playing a role in trichome branching. Indeed, the most
-significant signals found by all three approaches were located near this gene.
-analyses (Figures \ref{pubescence-form-main-figure}a-b,
+significant signals found by all three approaches were located near this gene
+(Figures \ref{pubescence-form-main-figure}a-b,
 \ref{pubescence_form_all_manhattan}, \ref{pubescence_form_all_Pa1_signal}).
@@
 to pubescence form.
+
+A second GWAS analysis focused only on appressed and semi-appressed phenotypes
+with the objective of enhancing the signal at the \emph{Pa2} locus. However,
+this analysis only detected weak signals at the \emph{Pa1} locus (Figure
+\ref{pubescence_form_noerect_manhattan}) and its results were therefore not
+analyzed further.
 
@@
 at the known \textit{L1} locus and may therefore represent a distinct locus (Figures
-\ref{pod_color_blbr_manhattan}, \ref{pod_color_blbr_ld}).
+\ref{pod_color_blbr_manhattan}, \ref{pod_color_blbr_ld}). The observed signal overlaps a total of
+27 genes (Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those are
+annotated as having an obvious role in controlling pod color.
 

### 2023-02-28 tables/loci_table.R

@@ -184,7 +184,9 @@ loci_table$causal <- causal_lookup[loci_table$id]
 # Adding a note for loci that are not discussed in the main text
 nomain <- c("seed_coat_color_T", "seed_coat_color_R", "stem_termination_Dt2",
 	    "stem_termination_E3", "hilum_color_W1", "pubescence_density_Pd1",
-	    "pubescence_density_P1", "seed_coat_luster_I")
+	    "pubescence_density_P1", "seed_coat_luster_I", "maturity_group_E1",
+	    "maturity_group_E2", "maturity_group_E3", "maturity_group_E4",
+	    "oil_oilGm20", "oil_oilGm15", "protein_proteinGm20", "protein_proteinGm15")
 
 loci_table$original_trait <- paste0(loci_table$original_trait, ifelse(loci_table$id %in% nomain, "\\tnote{c}", ""))
 

### 2023-03-01 additional_files/main_text.tex
Invert the Methods and Results section; provide more context at the beginning of the results to compensate

@@
 location and can thus be used to query associations between genotype and
-phenotype at genomic locations that are not found in the reference genome.  A
+phenotype at genomic locations that are not found in a reference genome.  A
 few studies have applied this \emph{k}-mer-based GWAS approach
@@
 
-\section*{Methods}
-\label{sv-gwas-methods}
-
-\subsection*{Sample selection and processing of sequencing data}
-\label{sv-gwas-sequencing-data}
-
-We selected 389 inbred \emph{G. max} accessions based on the availability of
-Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
-sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
-Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
-and extracted paired-end reads using the fastq-dump command v. 2.9.6
-(\citeauthor[][\url{https://github.com/ncbi/sra-tools}]{sratoolkit}) with the
-option \texttt{-\-split-3}.
-Reads that were not paired in the raw sequencing data were not used
-for downstream analyses.  The runs downloaded from the SRA and associated
-metadata are listed in Additional file 1.
-
-We first filtered the sequencing data for quality and the presence of sequencing
-adapters using bbduk from BBtools v. 38.25
-(\citeauthor[][\url{https://sourceforge.net/projects/bbmap/}]{bbtools}). Reads
-were then mapped to the reference assembly version 4 of \emph{G. max}
-cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
-and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
-bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
-reads in paired-end mode. Reads that were left unpaired following adapter- and
-quality-trimming were aligned separately in single-end mode with bwa-mem and
-merged with mapped paired-end reads using samtools \texttt{merge}
-v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
-individual SRA runs to the BAM files and merged the reads belonging to the same
-accession using bamaddrg
-(\citeauthor[][\url{https://github.com/ekg/bamaddrg}]{bamaddrg}). We used the
-resulting BAM files for all downstream analyses requiring aligned reads.
-Metadata on mapping depth of aligned data can be found in Additional file 2.
-
-\subsection*{Discovery and genotyping of SVs}
-\label{sv-gwas-sv-discovery-genotyping}
-
-Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
-The \emph{discovery} step identified a set of candidate SVs using various methods,
-whereas the \emph{genotyping} step determined the genotype
-of the candidate SVs for all accessions in the population from the mapped Illumina reads. We used three different
-approaches in the SV discovery step to generate a set of candidate variants:
-
-\begin{itemize}
-	\item We used Illumina WGS data from the 389 accessions mentioned above
-		to call SVs in the population following a pipeline outlined
-		by \cite{lemay2022}. In brief, we combined information from
-		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
-		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
-		a set of candidate SVs from the mapped Illumina data.
-	\item We used SVs discovered by \cite{lemay2022} from Oxford Nanopore
-		sequencing data of 17 Canadian soybean cultivars.
-	\item We called SVs from 26 high-quality genome assemblies published by
-		\cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
-		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
-		We used methods similar to those of \cite{liu2020-pangenome} to call
-		SVs from the comparison of these assemblies to that of Williams82.
-\end{itemize}
-
-The SVs called using each of these approaches were merged using SVmerge
-\citep{wong2010} and the resulting set of candidate SVs was used for
-genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
-from the mapped Illumina reads for all accessions. Detailed methods regarding
-the SV discovery and genotyping steps can be found in Supplemental data.
-
-\subsection*{Discovery and genotyping of SNPs and indels}
-\label{sv-gwas-snp-discovery}
-
-We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
-mapped reads of the 389 accessions. The 21.1 M SNPs and
-indels from the raw output of Platypus were filtered to keep only those with:
-
-\begin{enumerate}
-	\item \texttt{FILTER} field set to \texttt{PASS}
-	\item minor allele frequency (MAF) $\geq$ 0.02
-	\item fraction of missing genotypes $\leq$ 0.5
-	\item heterozygosity rate $\leq$ 0.2.
-\end{enumerate}
-
-Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
-PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
-step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
-implemented to reduce the computational requirements of the GWAS analysis by
-removing co-segregating markers. Markers located on unanchored scaffolds were
-also removed. The resulting dataset was used for GWAS and comprised 773,060
-SNPs and 151,570 indels.
-
-\subsection*{Computing the presence/absence table of \emph{k}-mers}
-\label{sv-gwas-kmer-pav}
-
-We used the approach outlined by \cite{voichek2020} for \emph{k}-mer-based GWAS
-and their recommendations on the
-associated GitHub page (v. 0.2-beta, \url{https://github.com/voichek/kmersGWAS}) to generate
-a presence/absence \emph{k}-mer table for use in GWAS.  The first step of
-the approach involved counting all \emph{k}-mers of length 31 (31-mers) present
-in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
-\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
-canonized form (the first of either the observed \textit{k}-mer or its reverse
-complement in lexicographical order) and the second time using the observed
-\emph{k}-mers themselves (non-canonized form). The two sets of \textit{k}-mers
-were then combined and filtered by
-keeping only those seen in at least 5 accessions, and observed in both canonized
-and non-canonized form in at least 20\% of the accessions in which they are found.
-We then generated a table indicating the presence or absence of each
-\emph{k}-mer in all accessions. This table was used to compute a kinship
-table for use in GWAS and was also used directly as input genotypes for the
-GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
-computation of kinship and GWAS.
-
-\subsection*{Phenotypic data}
-\label{sv-gwas-phenotypic-data}
-
-We analyzed ten qualitative traits previously studied by \cite{bandillo2017}: flower
-color, pubescence color, seed coat color, stem termination type, hilum color,
-pod color, pubescence form, pubescence density, seed coat luster, and maturity
-group.  We obtained phenotypic data for these traits from the GRIN database
-(\url{https://npgsweb.ars-grin.gov/gringlobal/search}) by querying the database
-using the PI identifiers of the 389 accessions used.  Four out of the 389
-accessions did not match the GRIN database and were therefore not used for the
-analysis of qualitative traits. The phenotypes of qualitative traits were recoded
-to numerical values for GWAS analysis following methods similar to those of
-\cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
-using targeted subsets of the observed phenotypes in order to focus on specific
-loci. The phenotypes and numerical values used for all analyses as well as the
-number of observations in the dataset that we used for GWAS are listed in
-Tables \ref{FLOWER.COLOR_gwas_table} to \ref{MATURITY.GROUP_gwas_table}.
-
-We analyzed three quantitative traits in addition to the ten qualitative traits
-mentioned above. Seed oil and protein content were retrieved from the GRIN
-database. In cases where more than one value was listed for a given accession,
-we computed the average of those values and used it for GWAS analyses. As a third quantitative trait, we
-used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
-weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
-Our analysis included 340 accessions whose phenotypic data
-was already published by \cite{deronne2022} as well as 49 additional accessions for
-which sequencing data was not yet available when their study was conducted.
-Because of ambiguity regarding the identity of one of the accessions (HN019), this accession
-was dropped from all GWAS analyses. Similarly to seed oil and
-protein content, the CDW values for some accessions for which more than one
-observation was available were averaged.
-
-Because of clear discrepancies between the observed haplotypes and the reported
-phenotypes for some simple traits such as flower color and seed coat color, we
-suspected possible errors in the identity of some of the accessions. To
-investigate this, we compared the SNP genotypes derived from the SoySNP50K
-array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
-90\% concordance between the two genotype datasets (see Supplemental data for
-detailed methods). These samples were removed from the dataset in addition to another
-accession that exhibited an atypical GC content suggesting contamination.
-Following these filtering steps and the removal of HN019, 363 accessions
-remained for GWAS analyses. The concordance between WGS data and SoySNP50K
-genotypes is included in Additional file 2 and averaged 98.3\% among retained
-lines. The phenotypic data used in this study can be found in Additional file 3.
-
-\subsection*{GWAS analyses}
-\label{sv-gwas-gwas-methods}
-
-We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
-the SNP/indel and SV datasets described above. SVs and indels required some additional
-preprocessing steps to make variant
-representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
-variants as SNPs prior to converting the VCF files to Hapmap diploid format
-using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
-retrieval of the proper metadata for each variant after GWAS.
-
-We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
-algorithm for computing the kinship matrix. We used an MLM model for
-consistency with the statistical model used by the \textit{k}-mer-based
-approach.  We used a randomization approach to determine the 5\% family-wise
-error-rate threshold for each GWAS analysis as described in \cite{voichek2020}.
-To compute this threshold, we permuted the phenotypic observations 100 times
-and computed a GWAS on these permuted phenotypes to obtain a distribution of
-top \emph{p}-values under the null hypothesis. The fifth most significant of
-the top 100 \emph{p}-values was used as the significance threshold for
-inference, i.e. markers with \emph{p}-values lower than this threshold were
-considered significant.
-
-We conducted GWAS based on \emph{k}-mers using the method developed by
-\cite{voichek2020}. In brief, the
-analysis was conducted in two steps. The first step used an approximate model
-to identify 1 million potentially significant \emph{k}-mers. These were then used in
-the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
-in a \emph{p}-value for each \emph{k}-mer. We
-used the kinship computed from the \emph{k}-mers to correct for relatedness
-between accessions. The program automatically outputs a list of \emph{k}-mers
-that passed a 5\% family-wise error-rate threshold determined by a
-randomization approach, which we considered significant for the purposes of
-downstream analyses.
-
-\subsection*{Analysis of significant \emph{k}-mers}
-\label{sv-gwas-kmers-analysis-methods}
-
-\emph{k}-mers are not intrinsically associated with a particular genomic region
-and a tailored analysis was therefore required to associate them with genomic
-coordinates. To do so, we queried the reads (in BAM format) of all accessions for matches
-to significant \emph{k}-mers for a given trait. Because this is a
-computationally demanding operation (there were typically tens of thousands of
-significant \emph{k}-mers per trait), we developed a C program called katcher
-(\url{https://github.com/malemay/katcher}) that
-uses the htslib library \citep{bonfield2021} for efficient reading and
-writing of BAM files. katcher and its associated utilities allow for efficient
-retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
-of interest.
-
-Once we retrieved all the reads containing significant \emph{k}-mers, we next
-needed to gather the mapping information contained in those reads for the
-purposes of identifying contiguous signals and comparing them to annotations. To
-simplify the amount of information to process in downstream analyses, we
-considered only the most significant \emph{k}-mer contained in a given read and
-the position where this \emph{k}-mer was most often observed in
-the dataset. Moreover, we limited the analysis to reads with a minimum mapping
-quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
-least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
-from sequencing errors. All downstream analyses (Manhattan plots,
-identification of signals, analysis of genes located near significant
-\emph{k}-mers, etc.) used this processed set of significant \emph{k}-mers.
-
-\subsection*{Downstream analyses of significant association signals}
-\label{sv-gwas-methods-downstream}
-
-We used contiguous regions of significant associations, which we called
-``signals'', as a basis to compare our results with those obtained by previous
-studies and to query the genome for candidate genes. We defined the boundaries
-of those signals by grouping any significant markers or \emph{k}-mers that
-occurred within 250 kb of one another.  We did consider the length of deletions
-or \emph{k}-mers in defining the distance between significant associations.
-
-For the GWAS analyses computed from Platypus genotype calls, we retrieved all
-markers that had initially been pruned in a region ranging from 50 kb
-downstream to 50 kb upstream of significant signals and computed their
-\emph{p}-values by running GAPIT using the same parameters as noted above. The
-\emph{p}-values thus computed on markers that were initially left out were not
-used for initial signal discovery, but provided finer resolution when analyzing
-the results of the Platypus GWAS analyses. Indeed, some markers that had not
-been included in the original GWAS may be the causal variant or simply yield
-more significant \emph{p}-values than those initially used.
-
-To help with the interpretation of results and the identification of candidate
-genes, we defined regions of highly significant associations within each signal
-by delimiting a range of coordinates containing the top 5\% most significant
-associations (Paragraph) or 1\% most significant associations (Platypus
-and \emph{k}-mers) found within the signal. In cases where such a fraction
-represented fewer than two markers or \emph{k}-mers, the highly significant
-region was simply delimited by the two most significant associations.
-
-Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
-Williams82, we converted the genomic coordinates of their signals to assembly version 4
-using the same method that we used for converting
-SoySNP50K positions (see Supplemental data). For the analysis of
-qualitative traits, we only compared our results to signals at named classical
-loci and ignored previously reported minor signals with \emph{p}-values that
-just barely made it above the significance threshold.
-
-\subsection*{Linking significant \emph{k}-mers to variants}
-\label{sv-gwas-methods-linking-kmers}
-
-We used the reads retrieved by katcher and an assembly-based approach for
-identifying the variants associated with a given set of \emph{k}-mers. In
-addition to the reads themselves, we also retrieved their paired reads even
-though they may not have matched a significant \emph{k}-mer in the first place.
-These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
-\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
-individually for each accession and then aligned to the Williams82 reference
-assembly using bwa mem with default parameters.  Aligned assembled sequences
-that overlapped loci of interest were realigned using multiple sequence
-alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
-visualizing the haplotypes at these loci. The identification of variants
-supported by significant \textit{k}-mers was based entirely on visual analysis
-of haplotype alignments. We only considered haplotypes occurring at least five
-times in the population in such analyses in order to leave out potential
-assembly errors. In a single case (\emph{Dt1} locus), assembly at the locus
-succeeded for a very low number of accessions (26). In this case, we instead used
-consensus sequences from the reads aligned with bwa for comparing the
-haplotypes across accessions.
-
-\subsection*{Computation of LD between \emph{k}-mers}
-\label{sv-gwas-ld-kmers}
-
-In some cases, the analysis of \emph{k}-mers revealed signals at previously
-undocumented loci or found a very large number of associations throughout the
-genome. To shed light on these situations, we computed the pairwise LD between
-\emph{k}-mers in order to identify \emph{k}-mers that co-segregated and
-likely corresponded to a single locus. Because of the sheer number of pairwise
-comparisons that needed to be made for some traits (> 20,000 significant
-\emph{k}-mers kept even after the filtering steps mentioned above), we limited
-the number of \emph{k}-mers used for LD calculation to 1,500. For traits where
-more than 1,500 \emph{k}-mers were found, we computed LD on a subsample of
-these by selecting the 500 \emph{k}-mers with the most significant
-\emph{p}-values and sampling the remaining 1000 \emph{k}-mers randomly. This
-random sampling was done with a probability inversely proportional to the
-number of \emph{k}-mers matching a given chromosome or scaffold such that
-\emph{k}-mers were sampled from the whole genome and not simply from the most
-significant locations.
-
-\subsection*{Software used}
-\label{sv-gwas-software-used}
-
-Unless otherwise noted, all analyses were conducted using R v. 4.2.0
-\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
-packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
-\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
-1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
-\citep{obenchain2014}.  We gathered several functions used for the downstream
-analysis of GWAS results into a package called gwask which is available on
-GitHub (\url{https://github.com/malemay/gwask}).
-
 \section*{Results and discussion}
@@
 
-We analyzed a total of 13 traits using three different GWAS approaches, one
-based on SNP and indel genotypes (Platypus), one based on SV genotypes (Paragraph),
-and one based on the presence/absence of \textit{k}-mers.
-Because two or three different GWAS analyses were conducted for some of the traits, a
-total of 22 GWAS analyses were performed using each of the all approaches.
-Given the large volume of data that this represents, we will focus here on the
-most noteworthy results. Readers interested in detailed results for loci not
+We analyzed a total of 13 traits using three GWAS approaches that differed in
+the genotypes used. The first approach used SNP and indel
+genotypes called by Platypus. Despite both SNPs and indels being included in
+this approach, we will refer to this approach as SNP-based for simplicity. Our
+SV-based workflow used various sources of data for SV discovery and a single
+program (Paragraph) for genotyping SVs at population scale. Finally, the third
+GWAS approach relied on the presence/absence of \textit{k}-mers as genotypes.
+Because of the large number of \textit{k}-mers analyzed, we only report results
+(e.g. in Manhattan plots) for those that were considered significantly
+associated with traits of interest.
+
+Since two or three different GWAS analyses were conducted for some of the traits, a
+total of 22 GWAS analyses were performed using each of the three approaches.
+Given the large volume of data that this represents, we focus here on the
+most noteworthy results. Readers interested in results at loci that are not
 discussed here are referred to Supplemental data. In particular, table \ref{signals_table}
@@
 
+\section*{Methods}
+\label{sv-gwas-methods}
+
+\subsection*{Sample selection and processing of sequencing data}
+\label{sv-gwas-sequencing-data}
+
+We selected 389 inbred \emph{G. max} accessions based on the availability of
+Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
+sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
+Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
+and extracted paired-end reads using the fastq-dump command v. 2.9.6
+(\citeauthor[][\url{https://github.com/ncbi/sra-tools}]{sratoolkit}) with the
+option \texttt{-\-split-3}.
+Reads that were not paired in the raw sequencing data were not used
+for downstream analyses.  The runs downloaded from the SRA and associated
+metadata are listed in Additional file 1.
+
+We first filtered the sequencing data for quality and the presence of sequencing
+adapters using bbduk from BBtools v. 38.25
+(\citeauthor[][\url{https://sourceforge.net/projects/bbmap/}]{bbtools}). Reads
+were then mapped to the reference assembly version 4 of \emph{G. max}
+cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
+and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
+bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
+reads in paired-end mode. Reads that were left unpaired following adapter- and
+quality-trimming were aligned separately in single-end mode with bwa-mem and
+merged with mapped paired-end reads using samtools \texttt{merge}
+v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
+individual SRA runs to the BAM files and merged the reads belonging to the same
+accession using bamaddrg
+(\citeauthor[][\url{https://github.com/ekg/bamaddrg}]{bamaddrg}). We used the
+resulting BAM files for all downstream analyses requiring aligned reads.
+Metadata on mapping depth of aligned data can be found in Additional file 2.
+
+\subsection*{Discovery and genotyping of SVs}
+\label{sv-gwas-sv-discovery-genotyping}
+
+Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
+The \emph{discovery} step identified a set of candidate SVs using various methods,
+whereas the \emph{genotyping} step determined the genotype
+of the candidate SVs for all accessions in the population from the mapped Illumina reads. We used three different
+approaches in the SV discovery step to generate a set of candidate variants:
+
+\begin{itemize}
+	\item We used Illumina WGS data from the 389 accessions mentioned above
+		to call SVs in the population following a pipeline outlined
+		by \cite{lemay2022}. In brief, we combined information from
+		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
+		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
+		a set of candidate SVs from the mapped Illumina data.
+	\item We used SVs discovered by \cite{lemay2022} from Oxford Nanopore
+		sequencing data of 17 Canadian soybean cultivars.
+	\item We called SVs from 26 high-quality genome assemblies published by
+		\cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
+		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
+		We used methods similar to those of \cite{liu2020-pangenome} to call
+		SVs from the comparison of these assemblies to that of Williams82.
+\end{itemize}
+
+The SVs called using each of these approaches were merged using SVmerge
+\citep{wong2010} and the resulting set of candidate SVs was used for
+genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
+from the mapped Illumina reads for all accessions. Detailed methods regarding
+the SV discovery and genotyping steps can be found in Supplemental data.
+
+\subsection*{Discovery and genotyping of SNPs and indels}
+\label{sv-gwas-snp-discovery}
+
+We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
+mapped reads of the 389 accessions. The 21.1 M SNPs and
+indels from the raw output of Platypus were filtered to keep only those with:
+
+\begin{enumerate}
+	\item \texttt{FILTER} field set to \texttt{PASS}
+	\item minor allele frequency (MAF) $\geq$ 0.02
+	\item fraction of missing genotypes $\leq$ 0.5
+	\item heterozygosity rate $\leq$ 0.2.
+\end{enumerate}
+
+Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
+PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
+step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
+implemented to reduce the computational requirements of the GWAS analysis by
+removing co-segregating markers. Markers located on unanchored scaffolds were
+also removed. The resulting dataset was used for GWAS and comprised 773,060
+SNPs and 151,570 indels.
+
+\subsection*{Computing the presence/absence table of \emph{k}-mers}
+\label{sv-gwas-kmer-pav}
+
+We used the approach outlined by \cite{voichek2020} for \emph{k}-mer-based GWAS
+and their recommendations on the
+associated GitHub page (v. 0.2-beta, \url{https://github.com/voichek/kmersGWAS}) to generate
+a presence/absence \emph{k}-mer table for use in GWAS.  The first step of
+the approach involved counting all \emph{k}-mers of length 31 (31-mers) present
+in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
+\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
+canonized form (the first of either the observed \textit{k}-mer or its reverse
+complement in lexicographical order) and the second time using the observed
+\emph{k}-mers themselves (non-canonized form). The two sets of \textit{k}-mers
+were then combined and filtered by
+keeping only those seen in at least 5 accessions, and observed in both canonized
+and non-canonized form in at least 20\% of the accessions in which they are found.
+We then generated a table indicating the presence or absence of each
+\emph{k}-mer in all accessions. This table was used to compute a kinship
+table for use in GWAS and was also used directly as input genotypes for the
+GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
+computation of kinship and GWAS.
+
+\subsection*{Phenotypic data}
+\label{sv-gwas-phenotypic-data}
+
+We analyzed ten qualitative traits previously studied by \cite{bandillo2017}: flower
+color, pubescence color, seed coat color, stem termination type, hilum color,
+pod color, pubescence form, pubescence density, seed coat luster, and maturity
+group.  We obtained phenotypic data for these traits from the GRIN database
+(\url{https://npgsweb.ars-grin.gov/gringlobal/search}) by querying the database
+using the PI identifiers of the 389 accessions used.  Four out of the 389
+accessions did not match the GRIN database and were therefore not used for the
+analysis of qualitative traits. The phenotypes of qualitative traits were recoded
+to numerical values for GWAS analysis following methods similar to those of
+\cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
+using targeted subsets of the observed phenotypes in order to focus on specific
+loci. The phenotypes and numerical values used for all analyses as well as the
+number of observations in the dataset that we used for GWAS are listed in
+Tables \ref{FLOWER.COLOR_gwas_table} to \ref{MATURITY.GROUP_gwas_table}.
+
+We analyzed three quantitative traits in addition to the ten qualitative traits
+mentioned above. Seed oil and protein content were retrieved from the GRIN
+database. In cases where more than one value was listed for a given accession,
+we computed the average of those values and used it for GWAS analyses. As a third quantitative trait, we
+used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
+weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
+Our analysis included 340 accessions whose phenotypic data
+was already published by \cite{deronne2022} as well as 49 additional accessions for
+which sequencing data was not yet available when their study was conducted.
+Because of ambiguity regarding the identity of one of the accessions (HN019), this accession
+was dropped from all GWAS analyses. Similarly to seed oil and
+protein content, the CDW values for some accessions for which more than one
+observation was available were averaged.
+
+Because of clear discrepancies between the observed haplotypes and the reported
+phenotypes for some simple traits such as flower color and seed coat color, we
+suspected possible errors in the identity of some of the accessions. To
+investigate this, we compared the SNP genotypes derived from the SoySNP50K
+array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
+90\% concordance between the two genotype datasets (see Supplemental data for
+detailed methods). These samples were removed from the dataset in addition to another
+accession that exhibited an atypical GC content suggesting contamination.
+Following these filtering steps and the removal of HN019, 363 accessions
+remained for GWAS analyses. The concordance between WGS data and SoySNP50K
+genotypes is included in Additional file 2 and averaged 98.3\% among retained
+lines. The phenotypic data used in this study can be found in Additional file 3.
+
+\subsection*{GWAS analyses}
+\label{sv-gwas-gwas-methods}
+
+We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
+the SNP/indel and SV datasets described above. SVs and indels required some additional
+preprocessing steps to make variant
+representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
+variants as SNPs prior to converting the VCF files to Hapmap diploid format
+using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
+retrieval of the proper metadata for each variant after GWAS.
+
+We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
+algorithm for computing the kinship matrix. We used an MLM model for
+consistency with the statistical model used by the \textit{k}-mer-based
+approach.  We used a randomization approach to determine the 5\% family-wise
+error-rate threshold for each GWAS analysis as described in \cite{voichek2020}.
+To compute this threshold, we permuted the phenotypic observations 100 times
+and computed a GWAS on these permuted phenotypes to obtain a distribution of
+top \emph{p}-values under the null hypothesis. The fifth most significant of
+the top 100 \emph{p}-values was used as the significance threshold for
+inference, i.e. markers with \emph{p}-values lower than this threshold were
+considered significant.
+
+We conducted GWAS based on \emph{k}-mers using the method developed by
+\cite{voichek2020}. In brief, the
+analysis was conducted in two steps. The first step used an approximate model
+to identify 1 million potentially significant \emph{k}-mers. These were then used in
+the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
+in a \emph{p}-value for each \emph{k}-mer. We
+used the kinship computed from the \emph{k}-mers to correct for relatedness
+between accessions. The program automatically outputs a list of \emph{k}-mers
+that passed a 5\% family-wise error-rate threshold determined by a
+randomization approach, which we considered significant for the purposes of
+downstream analyses.
+
+\subsection*{Analysis of significant \emph{k}-mers}
+\label{sv-gwas-kmers-analysis-methods}
+
+\emph{k}-mers are not intrinsically associated with a particular genomic region
+and a tailored analysis was therefore required to associate them with genomic
+coordinates. To do so, we queried the reads (in BAM format) of all accessions for matches
+to significant \emph{k}-mers for a given trait. Because this is a
+computationally demanding operation (there were typically tens of thousands of
+significant \emph{k}-mers per trait), we developed a C program called katcher
+(\url{https://github.com/malemay/katcher}) that
+uses the htslib library \citep{bonfield2021} for efficient reading and
+writing of BAM files. katcher and its associated utilities allow for efficient
+retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
+of interest.
+
+Once we retrieved all the reads containing significant \emph{k}-mers, we next
+needed to gather the mapping information contained in those reads for the
+purposes of identifying contiguous signals and comparing them to annotations. To
+simplify the amount of information to process in downstream analyses, we
+considered only the most significant \emph{k}-mer contained in a given read and
+the position where this \emph{k}-mer was most often observed in
+the dataset. Moreover, we limited the analysis to reads with a minimum mapping
+quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
+least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
+from sequencing errors. All downstream analyses (Manhattan plots,
+identification of signals, analysis of genes located near significant
+\emph{k}-mers, etc.) used this processed set of significant \emph{k}-mers.
+
+\subsection*{Downstream analyses of significant association signals}
+\label{sv-gwas-methods-downstream}
+
+We used contiguous regions of significant associations, which we called
+``signals'', as a basis to compare our results with those obtained by previous
+studies and to query the genome for candidate genes. We defined the boundaries
+of those signals by grouping any significant markers or \emph{k}-mers that
+occurred within 250 kb of one another.  We did consider the length of deletions
+or \emph{k}-mers in defining the distance between significant associations.
+
+For the GWAS analyses computed from Platypus genotype calls, we retrieved all
+markers that had initially been pruned in a region ranging from 50 kb
+downstream to 50 kb upstream of significant signals and computed their
+\emph{p}-values by running GAPIT using the same parameters as noted above. The
+\emph{p}-values thus computed on markers that were initially left out were not
+used for initial signal discovery, but provided finer resolution when analyzing
+the results of the Platypus GWAS analyses. Indeed, some markers that had not
+been included in the original GWAS may be the causal variant or simply yield
+more significant \emph{p}-values than those initially used.
+
+To help with the interpretation of results and the identification of candidate
+genes, we defined regions of highly significant associations within each signal
+by delimiting a range of coordinates containing the top 5\% most significant
+associations (Paragraph) or 1\% most significant associations (Platypus
+and \emph{k}-mers) found within the signal. In cases where such a fraction
+represented fewer than two markers or \emph{k}-mers, the highly significant
+region was simply delimited by the two most significant associations.
+
+Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
+Williams82, we converted the genomic coordinates of their signals to assembly version 4
+using the same method that we used for converting
+SoySNP50K positions (see Supplemental data). For the analysis of
+qualitative traits, we only compared our results to signals at named classical
+loci and ignored previously reported minor signals with \emph{p}-values that
+just barely made it above the significance threshold.
+
+\subsection*{Linking significant \emph{k}-mers to variants}
+\label{sv-gwas-methods-linking-kmers}
+
+We used the reads retrieved by katcher and an assembly-based approach for
+identifying the variants associated with a given set of \emph{k}-mers. In
+addition to the reads themselves, we also retrieved their paired reads even
+though they may not have matched a significant \emph{k}-mer in the first place.
+These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
+\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
+individually for each accession and then aligned to the Williams82 reference
+assembly using bwa mem with default parameters.  Aligned assembled sequences
+that overlapped loci of interest were realigned using multiple sequence
+alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
+visualizing the haplotypes at these loci. The identification of variants
+supported by significant \textit{k}-mers was based entirely on visual analysis
+of haplotype alignments. We only considered haplotypes occurring at least five
+times in the population in such analyses in order to leave out potential
+assembly errors. In a single case (\emph{Dt1} locus), assembly at the locus
+succeeded for a very low number of accessions (26). In this case, we instead used
+consensus sequences from the reads aligned with bwa for comparing the
+haplotypes across accessions.
+
+\subsection*{Computation of LD between \emph{k}-mers}
+\label{sv-gwas-ld-kmers}
+
+In some cases, the analysis of \emph{k}-mers revealed signals at previously
+undocumented loci or found a very large number of associations throughout the
+genome. To shed light on these situations, we computed the pairwise LD between
+\emph{k}-mers in order to identify \emph{k}-mers that co-segregated and
+likely corresponded to a single locus. Because of the sheer number of pairwise
+comparisons that needed to be made for some traits (> 20,000 significant
+\emph{k}-mers kept even after the filtering steps mentioned above), we limited
+the number of \emph{k}-mers used for LD calculation to 1,500. For traits where
+more than 1,500 \emph{k}-mers were found, we computed LD on a subsample of
+these by selecting the 500 \emph{k}-mers with the most significant
+\emph{p}-values and sampling the remaining 1000 \emph{k}-mers randomly. This
+random sampling was done with a probability inversely proportional to the
+number of \emph{k}-mers matching a given chromosome or scaffold such that
+\emph{k}-mers were sampled from the whole genome and not simply from the most
+significant locations.
+
+\subsection*{Software used}
+\label{sv-gwas-software-used}
+
+Unless otherwise noted, all analyses were conducted using R v. 4.2.0
+\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
+packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
+\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
+1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
+\citep{obenchain2014}.  We gathered several functions used for the downstream
+analysis of GWAS results into a package called gwask which is available on
+GitHub (\url{https://github.com/malemay/gwask}).
+
 \section*{Additional files}

### 2023-03-01 additional_files/additional_file_1.tex
Make minor changes to the results (mostly) so that the manuscript reads more easily

@@
 
+% Hilum color - T locus
+\subsection*{Hilum color -- \textit{T} locus}
+\label{annexe-sv-gwas-results-hilum-color-t}
+
+The \textit{T} locus discussed in the main text for pubescence color is also a determining
+locus for hilum color. We detected signals overlapping the corresponding gene
+using all approaches (Figures \ref{hilum_color_all_manhattan},
+\ref{hilum_color_all_T_signal}). The regions defined by the top 5\% (SVs)
+and top 1\% (SNPs and \textit{k}-mers) variants included
+the Glyma.06g202300 gene associated with this locus. Similarly to what we
+observed for the analysis of the \textit{T} locus based on pubescence color, we
+did identify the documented causal variant at this locus using the Platypus and
+\textit{k}-mer analyses, although these variants were not the most
+significantly associated overall (Figures \ref{hilum_color_all_T_gene},
+\ref{hilum_color_all_T_kmers}).
+
+% Hilum color - I locus
+\subsection*{Hilum color -- \textit{I} locus}
+\label{annexe-sv-gwas-results-hilum-color-i}
+
+Similarly, the \textit{I} locus discussed in the main text for its effect on seed coat
+color also controls hilum color and we detected signals overlapping the
+associated inverted duplication using all approaches (Figures
+\ref{hilum_color_all_manhattan}, \ref{hilum_color_all_I_signal}). However,
+as was the case for the effect of \textit{I} on seed coat color, we were
+not able to link any of the variants identified to the causal variation given
+its highly complex nature.
+
 \subsection*{Hilum color -- loci \textit{W1} and \textit{O}}
@@
 
-\subsubsection*{Maturity group}
+\subsection*{Maturity group}
 \label{annexe-sv-gwas-maturity}
@@
 
-The first GWAS analysis on pubescence form found were signals on chromosomes Gm04 and Gm15 with Platypus
-\emph{k}-merss (Figure \ref{pubescence_form_all_manhattan}). However,
-upon further analysis of LD patterns, they appeared to be linked to the
-\emph{Pa1} locus (Figure \ref{pubescence_form_all_ld}).
+The first GWAS analysis on pubescence form found signals on chromosomes Gm04
+and Gm15 with SNPs and \emph{k}-mers (Figure
+\ref{pubescence_form_all_manhattan}). However, upon further analysis of LD
+patterns, they appeared to be linked to the \emph{Pa1} locus (Figure
+\ref{pubescence_form_all_ld}).
 

### 2023-03-01 additional_files/main_text.tex
Make minor changes to the results (mostly) so that the manuscript reads more easily

@@
 
-We analyzed a total of 13 traits using three GWAS approaches that differed in
-the genotypes used. The first approach used SNP and indel
-genotypes called by Platypus. Despite both SNPs and indels being included in
-this approach, we will refer to this approach as SNP-based for simplicity. Our
-SV-based workflow used various sources of data for SV discovery and a single
-program (Paragraph) for genotyping SVs at population scale. Finally, the third
-GWAS approach relied on the presence/absence of \textit{k}-mers as genotypes.
-Because of the large number of \textit{k}-mers analyzed, we only report results
-(e.g. in Manhattan plots) for those that were considered significantly
-associated with traits of interest.
+We analyzed a total of 13 traits in a population of 363 \textit{G. max}
+accessions using three GWAS approaches that differed in the genotypes used. The
+first approach used SNP and indel genotypes called by Platypus. Despite both
+SNPs and indels being included in this approach, we will refer to this approach
+as SNP-based for simplicity. Our SV-based workflow used various sources of data
+for SV discovery and a single program (Paragraph) for genotyping SVs at
+population scale. Finally, the third GWAS approach relied on the
+presence/absence of \textit{k}-mers as genotypes.  Because of the large number
+of \textit{k}-mers analyzed, we only report results (e.g. in Manhattan plots)
+for those that were considered significantly associated with traits of
+interest.
 
@@
 			tabular = llllccc,
-			table head = \toprule Trait & Locus & Gene & Causal variant & Platypus\tnote{a} & Paragraph\tnote{b} & \textit{k}-mers \\ \midrule,
+			table head = \toprule Trait & Locus & Gene & Causal variant & SNPs\tnote{a} & SVs\tnote{b} & \textit{k}-mers \\ \midrule,
 			table foot = \bottomrule]%
@@
 			\begin{tablenotes}
-			\item[a] Platypus genotypes SNPs and indels
-			\item[b] Paragraph genotypes SVs
+			\item[a] SNPs and indels genotyped by Platypus
+			\item[b] SVs genotyped by Paragraph
 			\item[c] Locus not discussed in main text. See Supplemental data for detailed results.
@@
 			\item[]* The signal found at this locus overlaps the gene.
-			\item[]** The region defined by the top 5\% (Paragraph) or 1\% (Platypus and \textit{k}-mers) variants or \textit{k}-mers overlaps the gene.
-			\item[]*** The most significantly associated marker or \textit{k}-mer is the causal variant.
+			\item[]** The region defined by the top 5\% (SVs) or 1\% (SNPs and \textit{k}-mers) variants overlaps the gene.
+			\item[]*** The most significantly associated variant is the causal variant.
 			\end{tablenotes}
@@
 As the catalogue of variants analyzed by Platypus contains only SNPs and small
-indels, variants located in the gene proved to be among the most closely
-associated, but the causal variant could not be captured given the large size
-of this variant.
+indels, the causal variant could not be captured by this approach. Still, SNPs
+and indels located within the gene sequence were among the most significantly
+associated ones.
 
@@
 \cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
-\textit{T} locus for pubescence color and found a 1-bp deletion as
-a putative causal variant at this locus. For all three GWAS approaches, we
-found a signal overlapping this gene (Figure
-\ref{pubescence_color_all_manhattan}). In
-particular, the regions defined by the top 1\% most associated variants identified by the
-Platypus and \textit{k}-mer approaches included Glyma.06g202300 (Figure
-\ref{pubescence_color_all_T_signal}). While the most significant
-\textit{k}-mers did occur within the sequence of this gene, they did not
-correspond to the causal variant (Figure \ref{pubescence_color_all_T_gene}).
-Still, the fourth and fifth most strongly associated \textit{k}-mers
-corresponded to the causal variant (Figures \ref{pubescence_color_all_T_gene},
-\ref{pubescence_color_all_T_kmers}). We
+\textit{T} locus for pubescence color and found a 1-bp deletion as a putative
+causal variant at this locus. For all three GWAS approaches, we found a signal
+overlapping this gene (Figure \ref{pubescence_color_all_manhattan}). In
+particular, the regions defined by the 1\% most associated variants
+identified by the SNP- and \textit{k}-mer-based approaches included
+Glyma.06g202300 (Figure \ref{pubescence_color_all_T_signal}). While the most
+significant \textit{k}-mers did occur within the sequence of this gene, they
+did not correspond to the causal variant (Figure
+\ref{pubescence_color_all_T_gene}).  Still, the fourth and fifth most strongly
+associated \textit{k}-mers corresponded to the causal variant (Figures
+\ref{pubescence_color_all_T_gene}, \ref{pubescence_color_all_T_kmers}). We
 observed that not all accessions with gray pubescence harbored the deleted
-nucleotide (Figure \ref{pubescence_color_all_T_kmers}c), which could indicate either that other causal variants are resulting in
-a non-functional F3’H in this population, or that the causal
-variant lies elsewhere. As for Platypus, the most significant
-variant within the sequence of Glyma.06g202300 was the causal
-variant, however it was not the most significantly associated variant overall
-(Figures \ref{pubescence_color_all_T_signal},
-\ref{pubescence_color_all_T_gene}). Paragraph did not detect the causal
-variant at this locus given that this approach does not target small indels.
+nucleotide (Figure \ref{pubescence_color_all_T_kmers}c), which could indicate
+either that other causal variants are resulting in a non-functional F3’H in
+this population, or that the causal variant lies elsewhere. As for SNP-based GWAS,
+the most significant variant within the sequence of Glyma.06g202300 was the
+causal variant, however it was not the most significantly associated variant
+overall (Figures \ref{pubescence_color_all_T_signal},
+\ref{pubescence_color_all_T_gene}). SV-based GWAS did not detect the causal variant
+at this locus, as expected from its size.
 
@@
 these differences in color. Accordingly, we
-detected signals in this region using all three approaches, but only Platypus
-and the \textit{k}-mers identified a signal that overlapped Glyma.03g258700
+detected signals in this region using all three approaches, but only SNP-
+and \textit{k}-mer-based GWAS identified a signal that overlapped Glyma.03g258700
 (Figures \ref{pubescence-color-main-figure}a-b,
@@
 at this locus \citep{yan2020}, it was the fourth most significantly associated variant
-for the Platypus genotype calls (Figures
+for the SNP-based approach (Figures
 \ref{pubescence-color-main-figure}b-d, \ref{pubescence_color_nogray_Td_gene}).
@@
 \begin{figure}
-	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
+	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
 	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
@@
 	rectangles show the regions delimited by the top 1\% associations.
-	In the case of SNP/indels, blue points denote markers used in the
+	In the case of SNPs, blue points denote markers used in the
 	original analysis, whereas red points denote points that had originally
@@
 coat specifically and thus in the absence of seed coat pigmentation
-\citep{tuteja2009}. While we did identify signals overlapping this causal variant
+\citep{tuteja2009}. While we did find signals overlapping this causal variant
 using all three GWAS approaches (Figures \ref{seed_coat_color_all_manhattan},
 \ref{seed_coat_color_all_I_signal}), none of the approaches identified
-the causal variant. This failure to detect the causal variant is likely due
+it. This failure to detect the causal variant is likely due
 to its complexity, which makes it difficult to detect and genotype it. While it
@@
 most significant \emph{k}-mers and the second most significant marker from the
-Platypus analysis as corresponding to that causal variant (Figures
+SNP-based analysis as corresponding to that causal variant (Figures
 \ref{seed-coat-color-main-figure},
 \ref{seed_coat_color_greenyellow_G_signal},
-\ref{seed_coat_color_greenyellow_G_gene}). Unsurprisingly, Paragraph
+\ref{seed_coat_color_greenyellow_G_gene}). Unsurprisingly, SV-based GWAS
 did not identify the causal variant given its type.
@@
 \begin{figure}
-	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
+	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
 	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
@@
 While all GWAS approaches found signals near the Glyma.19g194300 gene
-associated with the \textit{Dt1} locus for stem termination type, the
-\textit{k}-mer and Platypus analyses performed best since the region defined by the most highly associated markers
-(top 1\%) included the causal gene (Figures
+associated with the \textit{Dt1} locus for stem termination type, the SNP- and
+\textit{k}-mer-based analyses performed best since the region defined by their
+most highly associated markers (top 1\%) included the causal gene (Figures
 \ref{stem_termination_all_manhattan}, \ref{stem_termination_all_Dt1_signal}).
@@
 \ref{stem_termination_all_Dt1_gene}, \ref{stem_termination_all_Dt1_kmers}).
-Our results are consistent with previous findings in
-suggesting that there is more than one causal variant at the \textit{Dt1} locus \citep{liu2010, tian2010}. Indeed, a
-few cultivars bearing the allele associated with \emph{dt1} showed
-indeterminate or semi-determinate phenotypes and some cultivars showed a
-determinate phenotype without bearing the allele associated with \emph{dt1}
-(Figure \ref{stem_termination_all_Dt1_kmers}). The existence of several causal
-variants might explain why our analyses did not identify a causal variant more
-clearly.
-
-% Hilum color - T locus
-\subsubsection*{Hilum color -- \textit{T} locus}
-\label{sv-gwas-main-results-hilum-color-t}
-
-The \textit{T} locus mentioned above for pubescence color is also a determining
-locus for hilum color. We detected signals overlapping the corresponding gene
-using all approaches (Figures \ref{hilum_color_all_manhattan},
-\ref{hilum_color_all_T_signal}). The regions defined by the top 5\% (Paragraph)
-and top 1\% (Platypus and \textit{k}-mers) variants/\textit{k}-mers included
-the Glyma.06g202300 gene associated with this locus. Similarly to what we
-observed for the analysis of the \textit{T} locus based on pubescence color, we
-did identify the documented causal variant at this locus using the Platypus and
-\textit{k}-mer analyses, although these variants were not the most
-significantly associated overall (Figures \ref{hilum_color_all_T_gene},
-\ref{hilum_color_all_T_kmers}).
-
-% Hilum color - I locus
-\subsubsection*{Hilum color -- \textit{I} locus}
-\label{sv-gwas-main-results-hilum-color-i}
-
-Similarly, the \textit{I} locus discussed above for its effect on seed coat
-color also controls hilum color and we detected signals overlapping the
-associated inverted duplication using all approaches (Figures
-\ref{hilum_color_all_manhattan}, \ref{hilum_color_all_I_signal}). However,
-as was the case for the effect of \textit{I} on seed coat color, we were
-not able to link any of the variants identified to the causal variation given
-its highly complex nature.
+Our results are consistent with previous findings in suggesting that there is
+more than one causal variant at the \textit{Dt1} locus \citep{liu2010,
+tian2010}. Indeed, a few cultivars bearing the allele associated with
+\emph{dt1} showed indeterminate or semi-determinate phenotypes and some
+cultivars showed a determinate phenotype without bearing the allele associated
+with \emph{dt1} (Figure \ref{stem_termination_all_Dt1_kmers}). The existence of
+several causal variants might explain why our analyses did not identify a
+causal variant more clearly.
 
@@
 accordingly detected a signal overlapping this locus using all approaches, but
-only Platypus and \textit{k}-mers detected signals overlapping the
+only SNP- and \textit{k}-mer-based GWAS detected signals overlapping the
 Glyma.09g235100 gene associated with \textit{R} (Figures
@@
 and \ref{pubescence_density_Ps_gene}). The causal
-variant at this recently cloned locus is a copy number variant overlapping the
+variant at this recently cloned locus is a copy number variant (CNV) overlapping the
 Glyma.12g187200 gene, with higher copy number resulting in reduced pubescence \citep{liu2020ps}.
@@
 the reference, we reasoned that the CNV may have been found by our pipeline but
-filtered out later on. Indeed, found that a 42.8-kb duplication corresponding
+filtered out later on. Indeed, found a 42.8-kb duplication corresponding
 to the one described above among the genotype calls made by smoove. An association
@@
 which would have made it the most significant SV. In this particular case, the
-failure of the SV GWAS analysis to find identify the causal variant was due
+failure of the SV-based GWAS analysis to find identify the causal variant was due
 to our filtering parameters and not to limitations of the SV genotype calls per se.
@@
 
-Next, we conducted a GWAS analysis comparing dull and shiny seed coat
-phenotypes for the seed coat luster trait. Seed coat luster has been shown by
+We conducted a GWAS analysis comparing dull and shiny seed coat phenotypes for
+the seed coat luster trait. Seed coat luster has been shown by
 \cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
@@
 \label{sv-gwas-results}
+
 In addition to analyzing loci for which the genes are already known, we were
@@
 
-Both the \textit{L1} and the \textit{L2} loci for pod
-color have yet to be cloned. While \cite{he2015} proposed the gene
-Glyma.19g101700 as likely corresponding to \textit{L1} based on fine mapping
-and expression analysis, our results as well as those reported by
-\cite{bandillo2017} do not support this gene as a candidate due to its location
-$\sim$3 Mb away from GWAS signals. Our analyses based on
-\textit{k}-mers and Platypus found signals overlapping 14 and 6
-genes, respectively (Figures \ref{pod_color_all_manhattan},
+Both the \textit{L1} and the \textit{L2} loci for pod color have yet to be
+cloned. While \cite{he2015} proposed the gene Glyma.19g101700 as likely
+corresponding to \textit{L1} based on fine mapping and expression analysis, our
+results as well as those reported by \cite{bandillo2017} do not support this
+gene as a candidate due to its location $\sim$3 Mb away from GWAS signals. Our
+analyses based on \textit{k}-mers and SNPs found signals overlapping 14 and
+6 genes, respectively (Figures \ref{pod_color_all_manhattan},
 \ref{pod_color_all_L1_signal}). Of these, a MATE transporter (Glyma.19g120300)
@@
 
-The signals found by Platypus and \emph{k}-mer analyses at the \textit{L2}
+The signals found by SNP- and \emph{k}-mer-based analyses at the \textit{L2}
 locus overlap 24 and 50 genes, respectively (Figures
 \ref{pod_color_all_manhattan}, \ref{pod_color_all_L2_signal}). Despite this
-somewhat high number, we noticed that the closest gene to the
-most significant \emph{k}-mer (Glyma.03g005800) is also a MATE transporter,
-similarly to our top candidate for \emph{L1}. This gene may therefore represent
-an interesting candidate for this locus for future studies.
+somewhat high number, we noticed that the closest gene to the most significant
+\emph{k}-mer (Glyma.03g005800) is also a MATE transporter, similarly to our top
+candidate for \emph{L1}. This gene may therefore represent an interesting
+candidate for this locus for future studies.
 
@@
 The two known loci for pubescence form in soybean, \textit{Pa1} and
-\textit{Pa2}, have also not been cloned yet, although \cite{gilbert2017} has suggested
-Glyma.12g213900 as a candidate for \textit{Pa1} based on data published
-by \cite{bandillo2017}. This gene encodes a Myb transcription factor annotated
-as playing a role in trichome branching. Indeed, the most
-significant signals found by all three approaches were located near this gene
-(Figures \ref{pubescence-form-main-figure}a-b,
-\ref{pubescence_form_all_manhattan}, \ref{pubescence_form_all_Pa1_signal}).
-The analyses based on Platypus and \emph{k}-mers identified two highly
-significant non-synonymous SNPs in Glyma.12g213900 that were associated with appressed pubescence
-(Figures \ref{pubescence-form-main-figure}c-d, \ref{pubescence_form_all_Pa1_gene},
+\textit{Pa2}, have also not been cloned yet. Nevertheless, \cite{gilbert2017}
+has suggested Glyma.12g213900 as a candidate for \textit{Pa1} based on data
+published by \cite{bandillo2017}. This gene encodes a Myb transcription factor
+annotated as playing a role in trichome branching. Indeed, the most significant
+signals found by all three approaches were located near this gene (Figures
+\ref{pubescence-form-main-figure}a-b, \ref{pubescence_form_all_manhattan},
+\ref{pubescence_form_all_Pa1_signal}).  The analyses based on SNPs and
+\emph{k}-mers identified two highly significant non-synonymous SNPs in
+Glyma.12g213900 that were associated with appressed pubescence (Figures
+\ref{pubescence-form-main-figure}c-d, \ref{pubescence_form_all_Pa1_gene},
 \ref{pubescence_form_all_Pa1_kmers}). The alternative alleles for these two
@@
 \begin{figure}
-	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
+	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
 	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots.
@@
 measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
-overlapping the genomic region reported by these authors on Gm15 using all all GWAS
+overlapping the genomic region reported by these authors on Gm15 using all three GWAS
 approaches (Figure \ref{corrected_dry_weight_manhattan}). \cite{deronne2022}
@@
 large size of the observed signal means that the gene associated with the phenotype could
-be located almost anywhere in this region, however, we identified a particularly
-interesting region defined by the top 1\% associations in the Platypus
+be located almost anywhere in this region. However, we identified a particularly
+interesting region defined by the top 1\% associations in the SNP-based
 analysis. This region overlaps only three annotated genes and also contains the
@@
 In addition to verifying known loci, we also observed potentially novel loci in
-the \textit{k}-mer GWAS analysis of stem termination and pod color.
-When comparing accessions
-with indeterminate and semi-determinate stem termination types, we observed
-previously undocumented signals on chromosomes Gm11, Gm16 and Gm18 in
-addition to the known \textit{Dt1} locus (Figure \ref{stem-termination-main-figure}a).
-Analysis of the pairwise LD between significant \textit{k}-mers found no
-LD between these signals, suggesting that they may represent \textit{bona fide}
-loci associated with stem termination type
-(Figure \ref{stem-termination-main-figure}b). Similarly, a \textit{k}-mer GWAS
-analysis comparing accessions with black and brown pods found a previously
-undocumented signal on chromosome Gm15. \textit{k}-mers at this new locus showed no LD with \textit{k}-mers
-at the known \textit{L1} locus and may therefore represent a distinct locus (Figures
-\ref{pod_color_blbr_manhattan}, \ref{pod_color_blbr_ld}). The observed signal overlaps a total of
-27 genes (Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those are
-annotated as having an obvious role in controlling pod color.
+the \textit{k}-mer GWAS analysis of stem termination and pod color.  When
+comparing accessions with indeterminate and semi-determinate stem termination
+types, we observed previously undocumented signals on chromosomes Gm11, Gm16
+and Gm18 in addition to the known \textit{Dt1} locus (Figure
+\ref{stem-termination-main-figure}a).  Analysis of the pairwise LD between
+significant \textit{k}-mers found no LD between these signals, suggesting that
+they may represent \textit{bona fide} loci associated with stem termination
+type (Figure \ref{stem-termination-main-figure}b). Similarly, a \textit{k}-mer
+GWAS analysis comparing accessions with black and brown pods found a previously
+undocumented signal on chromosome Gm15. \textit{k}-mers at this new locus
+showed no LD with \textit{k}-mers at the known \textit{L1} locus and may
+therefore represent a distinct locus (Figures \ref{pod_color_blbr_manhattan},
+\ref{pod_color_blbr_ld}). The observed signal overlaps a total of 27 genes
+(Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those are annotated
+as having an obvious role in controlling pod color.
 
@@
 One of the most striking results from the analyses presented above is the
-impressive performance of \emph{k}-mer-based GWAS compared to SNP/indel- and
-SV-based GWAS. The GWAS analyses using \emph{k}-mers identified
-the causal variant as the most significantly associated variant at four loci
-with various types of causal variants: \emph{W1} for flower color, \emph{Td}
-for pubescence color, \emph{G} for seed coat color, and \emph{Ps} for
-pubescence density (Table \ref{loci-table}). In comparison, this result was achieved by SV-based GWAS
-at the \textit{W1} locus only, and at none of the loci for SNP/indel-based GWAS.
+impressive performance of \emph{k}-mer-based GWAS compared to SNP- and SV-based
+GWAS. The GWAS analyses using \emph{k}-mers identified the causal variant as
+the most significantly associated variant at four loci with various types of
+causal variants: \emph{W1} for flower color, \emph{Td} for pubescence color,
+\emph{G} for seed coat color, and \emph{Ps} for pubescence density (Table
+\ref{loci-table}). In comparison, this result was achieved by SV-based GWAS at
+the \textit{W1} locus only, and at none of the loci for SNP-based GWAS.
 The \textit{k}-mer approach also systematically performed at least as well as
-the other three approaches in identifying signals overlapping known variants or genes
-(Table \ref{loci-table}). Moreover, whenever \textit{k}-mer-based GWAS did
-detect a signal, the region defined by the 1\% most significant \textit{k}-mers
-overlapped the known causal variant or gene.
+the other two approaches in identifying signals overlapping known variants or
+genes (Table \ref{loci-table}). Moreover, whenever \textit{k}-mer-based GWAS
+did detect a signal, the region defined by the 1\% most significant
+\textit{k}-mers overlapped the known causal variant or gene.
 
 Even in cases where the most significantly associated \emph{k}-mers did not
-pinpoint known causal variants, they were often located near
-those variants. For example, the most significantly associated \emph{k}-mers
-mapping to the \emph{T} locus for pubescence color were located in intronic sequences
-of the gene known to be associated with this locus.
-Similarly, the most significantly associated
-\emph{k}-mers at the \emph{B} locus for seed coat luster mapped within a
-genomic interval known to vary in copy number at that locus. In addition, the most
-significant \emph{k}-mers at the \emph{R} locus mapped just upstream of a
-homolog of the gene associated with the locus. Even though the \emph{k}-mer
-analysis did not identify the causal variants directly in those cases, it still
-provided a very strong starting point for candidate gene identification.
+pinpoint known causal variants, they were often located near those variants.
+For example, the most significantly associated \emph{k}-mers mapping to the
+\emph{T} locus for pubescence color were located in intronic sequences of the
+gene known to be associated with this locus.  Similarly, the most significantly
+associated \emph{k}-mers at the \emph{B} locus for seed coat luster mapped
+within a genomic interval known to vary in copy number at that locus. In
+addition, the most significant \emph{k}-mers at the \emph{R} locus mapped just
+upstream of a homolog of the gene associated with the locus. Even though the
+\emph{k}-mer analysis did not identify the causal variants directly in those
+cases, it still provided a very strong starting point for candidate gene
+identification.
 
@@
 arbitrary and may influence the final results. In the case of quantitative
-phenotypes, the usual limitations to the precision and stochastic variation of
+traits, the usual limitations to the precision and stochastic variation of
 phenotypic values will necessarily make the interpretation more difficult than
@@
 While we have have shown that using \emph{k}-mers for GWAS in soybean performed
-well compared to other GWAS methods, this
-approach does present some unique challenges and limitations. One obvious limitation
-is that using the presence/absence of \emph{k}-mers may not be appropriate in
-species where heterozygosity is common (i.e., most species other than inbred
-crops) as the \emph{k}-mers associated with the presence of both alleles will
-be observed at heterozygous loci. Moreover, although the presence/absence of
-\emph{k}-mers can detect variation independently of its type, some types of
-variation may be more difficult to detect than others.
-For example, copy number variation at a locus may not be
-detected from presence/absence alone if all copies are identical and occur in
-tandem, such that no \textit{k}-mer can differentiate between two copies and
-more. For similar reasons, \emph{k}-mers may also fail to capture variation
-occurring in repeated sequences. Alternative models based on \emph{k}-mer
-counts instead of presence/absence have been developed \citep{rahman2018,
-he2021} which may solve issues like these and broaden the scope of variants
-that can be identified from \emph{k}-mers.
-
-The complexity of analyzing the results of \emph{k}-mer-based GWAS is
-another limitation of this approach. Although the method developed by
+well compared to other GWAS methods, this approach does present some unique
+challenges and limitations. One obvious limitation is that using the
+presence/absence of \emph{k}-mers may not be appropriate in species where
+heterozygosity is common (i.e., most species other than inbred crops) as the
+\emph{k}-mers associated with the presence of both alleles will be observed at
+heterozygous loci. Moreover, although the presence/absence of \emph{k}-mers can
+detect variation independently of its type, some types of variation may be more
+difficult to detect than others.  For example, copy number variation at a locus
+may not be detected from presence/absence alone if all copies are identical and
+occur in tandem, such that no \textit{k}-mer can differentiate between two
+copies and more. For similar reasons, \emph{k}-mers may also fail to capture
+variation occurring in repeated sequences. Alternative models based on
+\emph{k}-mer counts instead of presence/absence have been developed
+\citep{rahman2018, he2021} which may solve issues like these and broaden the
+scope of variants that can be identified from \emph{k}-mers.
+
+The complexity of analyzing the results of \emph{k}-mer-based GWAS is another
+limitation of this approach. Although the method developed by
 \cite{voichek2020} is in itself relatively simple and computationally
@@
 \emph{k}-mers together into coherent signals, and linking them to biologically
-meaningful sequence variation. Going forward, mapping
-the \emph{k}-mers or reads to a graph-based genome \citep[using e.g. the vg
-toolkit;][]{siren2021} may represent an
-improvement to using a single linear genome.
-Grouping significant \emph{k}-mers by LD instead of by genomic
-position, as was done for some of the analyses in this paper, could prove
-useful in determining how many loci control a trait. In this study, we used an
-assembly-based approach similar to what others have previously done
-\citep[e.g.][]{voichek2020, rahman2018}. This approach allowed us to link
-\emph{k}-mers to their underlying variation for simple cases involving SNPs and
-indels, but we were unable to link significantly associated \emph{k}-mers to
-more challenging causal variants such as those explaining the \emph{I} or
-\emph{B} loci. In this context, developing approaches to systematically link
-significant \emph{k}-mers to the sequence variation that underlies their
-presence is much needed. Furthermore, finding ways to limit the number of spurious
-associations found for some traits, as was the case for flower color and
-pubescence color in our study, is also needed in order to obtain robust results.
-
-In this work, we provide a set of tools enabling the downstream
-analysis of \emph{k}-mer-based GWAS. Most importantly, we developed the
-katcher program for retrieving all reads containing
-any of a set of significant \emph{k}-mers in a computationally efficient manner. Previous studies usually
-mapped \emph{k}-mers to a reference genome to identify their reference-based
-coordinates \citep[e.g.][]{voichek2020, he2021, rahman2018, tripodi2021}.
-However, while this approach may work for \emph{k}-mers that show little
-divergence from the reference, it might fail to identify the genomic location
-of \emph{k}-mers that diverge significantly or are not found at all in the
-reference. Moreover, it does not take into account the sequence context of the read where
-the \emph{k}-mer was found, which may provide valuable information for mapping
-purposes. Our approach avoids these pitfalls by instead systematically retrieving reads
-containing significant \emph{k}-mers, including those that did not map at all
-to the reference assembly. While we were able to position the most significant
-\emph{k}-mers on the reference sequence in our analyses, it may not be the case
-for highly variable species or if the causal variation at a locus results from
-a large novel insertion.
+meaningful sequence variation. Going forward, mapping the \emph{k}-mers or
+reads to a graph-based genome \citep[e.g. using the vg toolkit;][]{siren2021}
+may represent an improvement to using a single linear genome.  Grouping
+significant \emph{k}-mers by LD instead of by genomic position, as was done for
+some of the analyses in this paper, could prove useful in determining how many
+loci control a trait. In this study, we used an assembly-based approach similar
+to what others have previously done \citep[e.g.][]{voichek2020, rahman2018}.
+This approach allowed us to link \emph{k}-mers to their underlying variation
+for simple cases involving SNPs and indels, but we were unable to link
+significantly associated \emph{k}-mers to more challenging causal variants such
+as those explaining the \emph{I} or \emph{B} loci. In this context, developing
+approaches to systematically link significant \emph{k}-mers to the sequence
+variation that underlies their presence is much needed. Furthermore, finding
+ways to limit the number of spurious associations found for some traits, as was
+the case for flower color and pubescence color in our study, is also needed in
+order to obtain robust results.
+
+In this work, we provide a set of tools enabling the downstream analysis of
+\emph{k}-mer-based GWAS. Most importantly, we developed the katcher program for
+retrieving all reads containing any of a set of significant \emph{k}-mers in a
+computationally efficient manner. Previous studies usually mapped \emph{k}-mers
+to a reference genome to identify their reference-based coordinates
+\citep[e.g.][]{voichek2020, he2021, rahman2018, tripodi2021}.  However, while
+this approach may work for \emph{k}-mers that show little divergence from the
+reference, it might fail to identify the genomic location of \emph{k}-mers that
+diverge significantly or are not found at all in the reference. Moreover, it
+does not take into account the sequence context of the read where the
+\emph{k}-mer was found, which may provide valuable information for mapping
+purposes. Our approach avoids these pitfalls by instead systematically
+retrieving reads containing significant \emph{k}-mers, including those that did
+not map at all to the reference assembly. While we were able to position the
+most significant \emph{k}-mers on the reference sequence in our analyses, it
+may not be the case for highly variable species or if the causal variation at a
+locus results from a large novel insertion.
 
@@
 
-In addition to \emph{k}-mers, we also assessed the use of SVs in GWAS analysis of soybean.
-Although we did find the causal variant as the most significantly associated SV
-at the \emph{W1} locus, results were otherwise rather inconclusive.
-Indeed, most of the other known or putative causal variants were SNPs or
-indels, and therefore were not represented in the SV datasets. At
-most loci where SVs are known to be the underlying causal variant (e.g.
-\emph{I}, \emph{B}, \emph{Ps}), the analyses based on SVs did not find the causal variants, most likely
-due to the difficulty of calling and genotyping such variants
-\citep{kosugi2019, lemay2022}. In the case of the \textit{Ps} locus, this limitation
-was due to a filtering step in our pipeline rather than to an inherent inability to
-properly call and genotype the causal variant.
-
-Given the current performance of SV
-genotyping programs, it is not clear whether using SVs for GWAS provides any
-advantage over using \emph{k}-mer- or SNP-based GWAS, or a combination of both.
-Indeed, in the single case where the SV-based analyses identified the causal
-variant, the \emph{k}-mer-based GWAS identified it as well. Moreover,
-SNP/indel-based GWAS provided better guidance in delimiting the signals found
-at a given locus because of the much higher density of markers compared to SVs.
-The usefulness of SV-based GWAS compared to SNP-based
-GWAS will largely depend on whether SVs can represent variation that is not
-otherwise in LD with SNPs. In tomato, \cite{dominguez2020} found that most
-transposable element insertions associated with phenotypic variation in
-agronomic traits were not tagged by SNPs. On the other hand, in humans,
+In addition to \emph{k}-mers, we also assessed the use of SVs in GWAS analysis
+of soybean.  Although we did find the causal variant as the most significantly
+associated SV at the \emph{W1} locus, results were otherwise rather
+inconclusive.  Indeed, most of the other known or putative causal variants were
+SNPs or indels, and therefore were not represented in the SV datasets. At most
+loci where SVs are known to be the underlying causal variant (e.g.  \emph{I},
+\emph{B}, \emph{Ps}), the analyses based on SVs did not find the causal
+variants, most likely due to the difficulty of calling and genotyping such
+variants \citep{kosugi2019, lemay2022}. In the case of the \textit{Ps} locus,
+this limitation was due to a filtering step in our pipeline rather than to an
+inherent inability to properly call and genotype the causal variant.
+
+Given the current performance of SV genotyping programs, it is not clear
+whether using SVs for GWAS provides any advantage over using \emph{k}-mer- or
+SNP-based GWAS, or a combination of both.  Indeed, in the single case where the
+SV-based analyses identified the causal variant, the \emph{k}-mer-based GWAS
+identified it as well. Moreover, SNP-based GWAS provided better guidance
+in delimiting the signals found at a given locus because of the much higher
+density of markers compared to SVs.  The usefulness of SV-based GWAS compared
+to SNP-based GWAS will largely depend on whether SVs can represent variation
+that is not otherwise in LD with SNPs. In tomato, \cite{dominguez2020} found
+that most transposable element insertions associated with phenotypic variation
+in agronomic traits were not tagged by SNPs. On the other hand, in humans,
 \cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
-is therefore likely that how well SVs can be tagged by SNPs will depend on
-the particular type of SV and on the species. A more
-efficient approach than SV-based GWAS may be to conduct \emph{k}-mer-based GWAS
-and only then link known SVs to significant \emph{k}-mers.
-Alternatively, \textit{k}-mer-based GWAS can be conducted specifically for
-\textit{k}-mers that are associated with SVs \citep{jayakodi2020}.
-Previous studies have
-obtained noteworthy results by using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
+is therefore likely that how well SVs can be tagged by SNPs will depend on the
+particular type of SV and on the species. A more efficient approach than
+SV-based GWAS may be to conduct \emph{k}-mer-based GWAS and only then link
+known SVs to significant \emph{k}-mers.  Alternatively, \textit{k}-mer-based
+GWAS can be conducted specifically for \textit{k}-mers that are associated with
+SVs \citep{jayakodi2020}.  Previous studies have obtained noteworthy results by
+using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
 dominguez2020}, however it would be worthwhile to see whether these results
@@
 qualitative traits and three quantitative traits in soybean cultivars and
-compared the results to traditional SNP/indel-based GWAS. While
-\emph{k}-mer-based GWAS proved to be a powerful approach in pinpointing the
-causal variation or genes associated with known loci, it is unclear whether
-conducting SV-based GWAS is worth the investment. In addition, we used the
-results from the \emph{k}-mer- and SNP-based GWAS to suggest candidate genes
-for a few classical loci that have yet to be cloned. Based on our results, we
-believe that an optimal workflow may involve conducting SNP and \emph{k}-mer
-GWAS in parallel to identify significant signals and candidate genes,
-potentially in combination with SV datasets. Given the large
-number of samples for which WGS data are now available in several species,
-applying \emph{k}-mer-based GWAS to leverage already existing sequencing and
-phenotypic data appears feasible and promising. As part of our work, we
-developed several computational tools that should help other researchers with the downstream
-analysis of a \emph{k}-mer-based GWAS results. However, much work
-remains to be done in developing state-of-the-art methods for the downstream
-analysis of significant \emph{k}-mers. In particular, better approaches are
-needed for linking \emph{k}-mers to sequence variation and moving from
-presence/absence-based methods to count-based methods.
+compared the results to traditional SNP-based GWAS. While \emph{k}-mer-based
+GWAS proved to be a powerful approach in pinpointing the causal variation or
+genes associated with known loci, it is unclear whether conducting SV-based
+GWAS is worth the investment. In addition, we used the results from the
+\emph{k}-mer- and SNP-based GWAS to suggest candidate genes for a few classical
+loci that have yet to be cloned. Based on our results, we believe that an
+optimal workflow may involve conducting SNP-based and \emph{k}-mer-based GWAS
+in parallel to identify significant signals and candidate genes, potentially in
+combination with SV datasets. Given the large number of samples for which WGS
+data are now available in several species, applying \emph{k}-mer-based GWAS to
+leverage already existing sequencing and phenotypic data appears feasible and
+promising.  As part of our work, we developed several computational tools that
+should help other researchers with the downstream analysis of
+\emph{k}-mer-based GWAS. However, much work remains to be done in
+developing state-of-the-art methods for the downstream analysis of significant
+\emph{k}-mers. In particular, better approaches are needed for linking
+\emph{k}-mers to sequence variation and moving from presence/absence-based
+methods to count-based methods.
 
@@
 used for initial signal discovery, but provided finer resolution when analyzing
-the results of the Platypus GWAS analyses. Indeed, some markers that had not
+the results of the SNP-based GWAS analyses. Indeed, some markers that had not
 been included in the original GWAS may be the causal variant or simply yield
@@
 by delimiting a range of coordinates containing the top 5\% most significant
-associations (Paragraph) or 1\% most significant associations (Platypus
+associations (Paragraph) or 1\% most significant associations (SNPs/indels
 and \emph{k}-mers) found within the signal. In cases where such a fraction

### 2023-03-01 figures/pubescence_color_nogray_main_figure.R

@@ -68,7 +68,7 @@ draw_manhattan(list(platypus_gwide_manhattan, kmers_gwide_manhattan),
 	       sigline_regexp = "sigline_1",
 	       siglabel_regexp = "siglabel_1",
 	       label_pos = 0.01,
-	       labels = c("SNP/indels", "k-mers"),
+	       labels = c("SNPs/indels", "k-mers"),
 	       fontsize = 10)
 
 
@@ -77,7 +77,7 @@ seekViewport("main")
 pushViewport(viewport(layout.pos.row = 3, name = "zoomed_manhattan"))
 grid.text("(b)", x = 0.02, y = 0.95)
 draw_zoomed(list(platypus_zoomed_manhattan, kmers_zoomed_manhattan),
-	    labels = c("SNP/indels", "k-mers"),
+	    labels = c("SNPs/indels", "k-mers"),
 	    label_pos = 0.01,
 	    fontsize = 10)
 

### 2023-03-01 tables/loci_table.R

@@ -186,7 +186,8 @@ nomain <- c("seed_coat_color_T", "seed_coat_color_R", "stem_termination_Dt2",
 	    "stem_termination_E3", "hilum_color_W1", "pubescence_density_Pd1",
 	    "pubescence_density_P1", "seed_coat_luster_I", "maturity_group_E1",
 	    "maturity_group_E2", "maturity_group_E3", "maturity_group_E4",
-	    "oil_oilGm20", "oil_oilGm15", "protein_proteinGm20", "protein_proteinGm15")
+	    "oil_oilGm20", "oil_oilGm15", "protein_proteinGm20", "protein_proteinGm15",
+	    "hilum_color_T", "hilum_color_I")
 
 loci_table$original_trait <- paste0(loci_table$original_trait, ifelse(loci_table$id %in% nomain, "\\tnote{c}", ""))
 

### 2023-03-01 additional_files/main_text.tex
Adjust figure numbering and order

@@
 most noteworthy results. Readers interested in results at loci that are not
-discussed here are referred to Supplemental data. In particular, table \ref{signals_table}
+discussed here are referred to Supplemental data. In particular, Table \ref{signals_table}
 provides a summary of the \textit{p}-values obtained by each approach at each
@@
 significant SV and \textit{k}-mers corresponded to the known causal variant
-(Figures \ref{flower-color-main-figure}a-b and \ref{flower_color_W1_gene}).
+(Figures \ref{flower-color-main-figure}a-b and \ref{flower_color_W1_gene}c-d).
 In particular, the assembly of the reads containing the most significant \textit{k}-mers
@@
 and indels located within the gene sequence were among the most significantly
-associated ones.
+associated ones (Figure \ref{flower_color_W1_gene}b).
 
@@
 \textit{k}-mers scattered throughout the genome (Figure
-\ref{flower_color_manhattan}). An analysis of the pairwise LD between a subset
+\ref{flower_color_manhattan}c). An analysis of the pairwise LD between a subset
 of those \textit{k}-mers showed that most \textit{k}-mers formed a single LD block
@@
 concordance between genotype calls made from WGS data and the SoySNP50K array
-for all samples (see the Methods section) and
-excluded 24 accessions for which the concordance rate was below 90\%.
+for all samples (see the Methods) and excluded 24 accessions for which the
+concordance rate was below 90\%.
 
@@
 did not correspond to the causal variant (Figure
-\ref{pubescence_color_all_T_gene}).  Still, the fourth and fifth most strongly
+\ref{pubescence_color_all_T_gene}d).  Still, the fourth and fifth most strongly
 associated \textit{k}-mers corresponded to the causal variant (Figures
-\ref{pubescence_color_all_T_gene}, \ref{pubescence_color_all_T_kmers}). We
+\ref{pubescence_color_all_T_gene}d, \ref{pubescence_color_all_T_kmers}). We
 observed that not all accessions with gray pubescence harbored the deleted
@@
 causal variant, however it was not the most significantly associated variant
-overall (Figures \ref{pubescence_color_all_T_signal},
-\ref{pubescence_color_all_T_gene}). SV-based GWAS did not detect the causal variant
+overall (Figures \ref{pubescence_color_all_T_signal}b,
+\ref{pubescence_color_all_T_gene}b). SV-based GWAS did not detect the causal variant
 at this locus, as expected from its size.
@@
 % Pubescence color - Td locus
-We conducted a separate GWAS analysis contrasting only accessions with tawny and light
-tawny pubescence because the \textit{Td} locus is known to control
-these differences in color. Accordingly, we
-detected signals in this region using all three approaches, but only SNP-
-and \textit{k}-mer-based GWAS identified a signal that overlapped Glyma.03g258700
-(Figures \ref{pubescence-color-main-figure}a-b,
-\ref{pubescence_color_nogray_manhattan}, \ref{pubescence_color_nogray_Td_signal}).
-While the most significant \textit{k}-mer identified the causal variant documented
-at this locus \citep{yan2020}, it was the fourth most significantly associated variant
-for the SNP-based approach (Figures
-\ref{pubescence-color-main-figure}b-d, \ref{pubescence_color_nogray_Td_gene}).
-Interestingly, some of the accessions bearing the haplotype associated with tawny pubescence
-displayed light tawny pubescence, suggesting that there may be more than one causal
-variant at this locus (Figure \ref{pubescence-color-main-figure}e).
+We conducted a separate GWAS analysis contrasting only accessions with tawny
+and light tawny pubescence because the \textit{Td} locus is known to control
+these differences in color. Accordingly, we detected signals in this region
+using all three approaches, but only SNP- and \textit{k}-mer-based GWAS
+identified a signal that overlapped Glyma.03g258700 (Figures
+\ref{pubescence-color-main-figure}a-b, \ref{pubescence_color_nogray_manhattan},
+\ref{pubescence_color_nogray_Td_signal}).  While the most significant
+\textit{k}-mer identified the causal variant documented at this locus
+\citep{yan2020}, it was the fourth most significantly associated variant for
+the SNP-based approach (Figures \ref{pubescence-color-main-figure}b-d,
+\ref{pubescence_color_nogray_Td_gene}).  Interestingly, some of the accessions
+bearing the haplotype associated with tawny pubescence displayed light tawny
+pubescence, suggesting that there may be more than one causal variant at this
+locus (Figure \ref{pubescence-color-main-figure}e).
 
@@
 cultivars showed a determinate phenotype without bearing the allele associated
-with \emph{dt1} (Figure \ref{stem_termination_all_Dt1_kmers}). The existence of
+with \emph{dt1} (Figure \ref{stem_termination_all_Dt1_kmers}c). The existence of
 several causal variants might explain why our analyses did not identify a
@@
 reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
-of the CNV (Figure \ref{pubescence-density-main-figure}b,d). This suggests that
+of the CNV (Figure \ref{pubescence-density-main-figure}d). This suggests that
 the most associated \textit{k}-mer is indeed linked to the CNV

### 2023-03-01 additional_files/main_text.tex
Change the numbering of additional files and add references to additional file 1 when mentioning supplemental figures or tables

@@
 most noteworthy results. Readers interested in results at loci that are not
-discussed here are referred to Supplemental data. In particular, Table \ref{signals_table}
+discussed here are referred to Additional file 1. In particular, Table \ref{signals_table} (Additional file 1)
 provides a summary of the \textit{p}-values obtained by each approach at each
@@
 			\item[b] SVs genotyped by Paragraph
-			\item[c] Locus not discussed in main text. See Supplemental data for detailed results.
+			\item[c] Locus not discussed in main text. See Additional file 1 for detailed results.
 			\item[]- The approach detected no signal overlapping the gene.
@@
 F3’5’H enzyme non-functional and prevents pigmentation of the flower. All
-GWAS approaches detected a signal overlapping the gene at this locus (Figures
-\ref{flower_color_manhattan}, \ref{flower_color_W1_signal}). The most
+GWAS approaches detected a signal overlapping the gene at this locus (Additional file 1: Figures
+\ref{flower_color_manhattan} and \ref{flower_color_W1_signal}). The most
 significant SV and \textit{k}-mers corresponded to the known causal variant
-(Figures \ref{flower-color-main-figure}a-b and \ref{flower_color_W1_gene}c-d).
+(Figure \ref{flower-color-main-figure}a-b and Additional file 1: \ref{flower_color_W1_gene}c-d).
 In particular, the assembly of the reads containing the most significant \textit{k}-mers
@@
 and indels located within the gene sequence were among the most significantly
-associated ones (Figure \ref{flower_color_W1_gene}b).
+associated ones (Additional file 1: Figure \ref{flower_color_W1_gene}b).
 
 The \textit{k}-mer GWAS analysis also identified presumably spurious significant
-\textit{k}-mers scattered throughout the genome (Figure
+\textit{k}-mers scattered throughout the genome (Additional file 1: Figure
 \ref{flower_color_manhattan}c). An analysis of the pairwise LD between a subset
 of those \textit{k}-mers showed that most \textit{k}-mers formed a single LD block
-linked to the \textit{W1} locus (Figure \ref{flower_color_ld}). A small LD block
+linked to the \textit{W1} locus (Additional file 1: Figure \ref{flower_color_ld}). A small LD block
 occurred on chromosome Gm09, however it was in moderate LD with the one on chromosome
@@
 causal variant at this locus. For all three GWAS approaches, we found a signal
-overlapping this gene (Figure \ref{pubescence_color_all_manhattan}). In
+overlapping this gene (Additional file 1: Figure \ref{pubescence_color_all_manhattan}). In
 particular, the regions defined by the 1\% most associated variants
 identified by the SNP- and \textit{k}-mer-based approaches included
-Glyma.06g202300 (Figure \ref{pubescence_color_all_T_signal}). While the most
+Glyma.06g202300 (Additional file 1: Figure \ref{pubescence_color_all_T_signal}). While the most
 significant \textit{k}-mers did occur within the sequence of this gene, they
-did not correspond to the causal variant (Figure
+did not correspond to the causal variant (Additional file 1: Figure
 \ref{pubescence_color_all_T_gene}d).  Still, the fourth and fifth most strongly
-associated \textit{k}-mers corresponded to the causal variant (Figures
-\ref{pubescence_color_all_T_gene}d, \ref{pubescence_color_all_T_kmers}). We
+associated \textit{k}-mers corresponded to the causal variant (Additional file 1: Figures
+\ref{pubescence_color_all_T_gene}d and \ref{pubescence_color_all_T_kmers}). We
 observed that not all accessions with gray pubescence harbored the deleted
-nucleotide (Figure \ref{pubescence_color_all_T_kmers}c), which could indicate
+nucleotide (Additional file 1: Figure \ref{pubescence_color_all_T_kmers}c), which could indicate
 either that other causal variants are resulting in a non-functional F3’H in
@@
 causal variant, however it was not the most significantly associated variant
-overall (Figures \ref{pubescence_color_all_T_signal}b,
+overall (Additional file 1: Figures \ref{pubescence_color_all_T_signal}b and
 \ref{pubescence_color_all_T_gene}b). SV-based GWAS did not detect the causal variant
@@
 using all three approaches, but only SNP- and \textit{k}-mer-based GWAS
-identified a signal that overlapped Glyma.03g258700 (Figures
-\ref{pubescence-color-main-figure}a-b, \ref{pubescence_color_nogray_manhattan},
+identified a signal that overlapped Glyma.03g258700 (Figure
+\ref{pubescence-color-main-figure}a-b, and Additional file 1: Figures \ref{pubescence_color_nogray_manhattan} and
 \ref{pubescence_color_nogray_Td_signal}).  While the most significant
@@
 \citep{yan2020}, it was the fourth most significantly associated variant for
-the SNP-based approach (Figures \ref{pubescence-color-main-figure}b-d,
+the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d,Additional file 1: Figure 
 \ref{pubescence_color_nogray_Td_gene}).  Interestingly, some of the accessions
@@
 \citep{tuteja2009}. While we did find signals overlapping this causal variant
-using all three GWAS approaches (Figures \ref{seed_coat_color_all_manhattan},
+using all three GWAS approaches (Additional file 1: Figures \ref{seed_coat_color_all_manhattan} and
 \ref{seed_coat_color_all_I_signal}), none of the approaches identified
@@
 analysis found strong signals overlapping the \emph{G} locus using all three
-approaches (Figure \ref{seed_coat_color_greenyellow_manhattan}).
+approaches (Additional file 1: Figure \ref{seed_coat_color_greenyellow_manhattan}).
 \cite{wang2018} identified Glyma.01g198500 as the gene associated with
@@
 most significant \emph{k}-mers and the second most significant marker from the
-SNP-based analysis as corresponding to that causal variant (Figures
-\ref{seed-coat-color-main-figure},
-\ref{seed_coat_color_greenyellow_G_signal},
+SNP-based analysis as corresponding to that causal variant (Figure
+\ref{seed-coat-color-main-figure} and Additional file 1: Figures
+\ref{seed_coat_color_greenyellow_G_signal} and
 \ref{seed_coat_color_greenyellow_G_gene}). Unsurprisingly, SV-based GWAS
@@
 \textit{k}-mer-based analyses performed best since the region defined by their
-most highly associated markers (top 1\%) included the causal gene (Figures
-\ref{stem_termination_all_manhattan}, \ref{stem_termination_all_Dt1_signal}).
+most highly associated markers (top 1\%) included the causal gene (Additional file 1: Figures
+\ref{stem_termination_all_manhattan} and \ref{stem_termination_all_Dt1_signal}).
 We did identify one of the SNPs suggested as putative causal variants for this
@@
 135\textsuperscript{th} most associated marker found by Platypus and the
-12\textsuperscript{th} most associated \emph{k}-mer (Figures
-\ref{stem_termination_all_Dt1_gene}, \ref{stem_termination_all_Dt1_kmers}).
+12\textsuperscript{th} most associated \emph{k}-mer (Additional file 1: Figures
+\ref{stem_termination_all_Dt1_gene} and \ref{stem_termination_all_Dt1_kmers}).
 Our results are consistent with previous findings in suggesting that there is
@@
 cultivars showed a determinate phenotype without bearing the allele associated
-with \emph{dt1} (Figure \ref{stem_termination_all_Dt1_kmers}c). The existence of
+with \emph{dt1} (Additional file 1: Figure \ref{stem_termination_all_Dt1_kmers}c). The existence of
 several causal variants might explain why our analyses did not identify a
@@
 only SNP- and \textit{k}-mer-based GWAS detected signals overlapping the
-Glyma.09g235100 gene associated with \textit{R} (Figures
-\ref{hilum_color_blackbrown_manhattan}, \ref{hilum_color_blackbrown_R_signal}).
+Glyma.09g235100 gene associated with \textit{R} (Additional file 1: Figures
+\ref{hilum_color_blackbrown_manhattan} and \ref{hilum_color_blackbrown_R_signal}).
 \cite{gillman2011} identified this R2R3 MYB transcription factor as the
@@
 detect strong signals in the vicinity of this gene, the most associated markers
-were not located in the body of the gene (Figure
+were not located in the body of the gene (Additional file 1: Figure
 \ref{hilum_color_blackbrown_R_signal}).  Interestingly, the most significantly
@@
 loss of function at the \emph{R} locus.  Still, we observed two causal variants
-previously documented by \cite{gillman2011} in our dataset (Figures
-\ref{hilum_color_blackbrown_R_gene}, \ref{hilum_color_blackbrown_R_kmers}).
+previously documented by \cite{gillman2011} in our dataset (Additional file 1: Figures
+\ref{hilum_color_blackbrown_R_gene} and \ref{hilum_color_blackbrown_R_kmers}).
 
@@
 pubes\-cence in order to detect loci associated with pubescence density. This analysis
-detected signals at the \textit{Ps} locus using all three approaches (Figures
-\ref{pubescence-density-main-figure}a-b, \ref{pubescence_density_manhattan}, \ref{pubescence_density_Ps_signal}
+detected signals at the \textit{Ps} locus using all three approaches (Figure
+\ref{pubescence-density-main-figure}a-b and Additional file 1: Figures \ref{pubescence_density_manhattan}, \ref{pubescence_density_Ps_signal}
 and \ref{pubescence_density_Ps_gene}). The causal
@@
 containing HPS domains and therefore likely represented the causal variant
-(Figures \ref{seed_coat_luster_dullshiny_B_signal},
+(Additional file 1: Figures \ref{seed_coat_luster_dullshiny_B_signal} and
 \ref{seed_coat_luster_dullshiny_B_gene}).  We found signals overlapping this
 CNV using all approaches, and the most significantly associated \textit{k}-mer
-was notably located within the CNV region (Figures
-\ref{seed_coat_luster_dullshiny_manhattan},
+was notably located within the CNV region (Additional file 1: Figures
+\ref{seed_coat_luster_dullshiny_manhattan} and
 \ref{seed_coat_luster_dullshiny_B_signal}). However, we were not able to
@@
 analyses based on \textit{k}-mers and SNPs found signals overlapping 14 and
-6 genes, respectively (Figures \ref{pod_color_all_manhattan},
+6 genes, respectively (Additional file 1: Figures \ref{pod_color_all_manhattan} and
 \ref{pod_color_all_L1_signal}). Of these, a MATE transporter (Glyma.19g120300)
@@
 The signals found by SNP- and \emph{k}-mer-based analyses at the \textit{L2}
-locus overlap 24 and 50 genes, respectively (Figures
-\ref{pod_color_all_manhattan}, \ref{pod_color_all_L2_signal}). Despite this
+locus overlap 24 and 50 genes, respectively (Additional file 1: Figures
+\ref{pod_color_all_manhattan} and \ref{pod_color_all_L2_signal}). Despite this
 somewhat high number, we noticed that the closest gene to the most significant
@@
 annotated as playing a role in trichome branching. Indeed, the most significant
-signals found by all three approaches were located near this gene (Figures
-\ref{pubescence-form-main-figure}a-b, \ref{pubescence_form_all_manhattan},
+signals found by all three approaches were located near this gene (Figure
+\ref{pubescence-form-main-figure}a-b, and Additional file 1: Figures \ref{pubescence_form_all_manhattan} and
 \ref{pubescence_form_all_Pa1_signal}).  The analyses based on SNPs and
 \emph{k}-mers identified two highly significant non-synonymous SNPs in
-Glyma.12g213900 that were associated with appressed pubescence (Figures
-\ref{pubescence-form-main-figure}c-d, \ref{pubescence_form_all_Pa1_gene},
+Glyma.12g213900 that were associated with appressed pubescence (Figure
+\ref{pubescence-form-main-figure}c-d and Additional file 1: Figures \ref{pubescence_form_all_Pa1_gene} and
 \ref{pubescence_form_all_Pa1_kmers}). The alternative alleles for these two
@@
 weaker and the signal observed from the \emph{k}-mer GWAS overlapped a total of
-20 genes (Figures \ref{pubescence_form_all_manhattan},
+20 genes (Additional file 1: Figures \ref{pubescence_form_all_manhattan} and
 \ref{pubescence_form_all_Pa2_signal}), none of which appeared obviously linked
@@
 with the objective of enhancing the signal at the \emph{Pa2} locus. However,
-this analysis only detected weak signals at the \emph{Pa1} locus (Figure
+this analysis only detected weak signals at the \emph{Pa1} locus (Additional file 1: Figure
 \ref{pubescence_form_noerect_manhattan}) and its results were therefore not
@@
 overlapping the genomic region reported by these authors on Gm15 using all three GWAS
-approaches (Figure \ref{corrected_dry_weight_manhattan}). \cite{deronne2022}
+approaches (Additional file 1: Figure \ref{corrected_dry_weight_manhattan}). \cite{deronne2022}
 proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
@@
 in a region with markers or \emph{k}-mers showing a weaker
-association (Figure \ref{corrected_dry_weight_cdwGm15_signal}). The
+association (Additional file 1: Figure \ref{corrected_dry_weight_cdwGm15_signal}). The
 large size of the observed signal means that the gene associated with the phenotype could
@@
 analysis. This region overlaps only three annotated genes and also contains the
-most significant \emph{k}-mer (Figure
+most significant \emph{k}-mer (Additional file 1: Figure
 \ref{corrected_dry_weight_cdwGm15_signal}). One of the genes in this region is
@@
 showed no LD with \textit{k}-mers at the known \textit{L1} locus and may
-therefore represent a distinct locus (Figures \ref{pod_color_blbr_manhattan},
+therefore represent a distinct locus (Additional file 1: Figures \ref{pod_color_blbr_manhattan} and
 \ref{pod_color_blbr_ld}). The observed signal overlaps a total of 27 genes
-(Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those are annotated
+(Additional file 1: Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those are annotated
 as having an obvious role in controlling pod color.
@@
 for downstream analyses.  The runs downloaded from the SRA and associated
-metadata are listed in Additional file 1.
+metadata are listed in Additional file 2.
 
@@
 resulting BAM files for all downstream analyses requiring aligned reads.
-Metadata on mapping depth of aligned data can be found in Additional file 2.
+Metadata on mapping depth of aligned data can be found in Additional file 3.
 
@@
 from the mapped Illumina reads for all accessions. Detailed methods regarding
-the SV discovery and genotyping steps can be found in Supplemental data.
+the SV discovery and genotyping steps can be found in Additional file 1.
 
@@
 number of observations in the dataset that we used for GWAS are listed in
-Tables \ref{FLOWER.COLOR_gwas_table} to \ref{MATURITY.GROUP_gwas_table}.
+Tables \ref{FLOWER.COLOR_gwas_table} to \ref{MATURITY.GROUP_gwas_table} (Additional file 1).
 
@@
 array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
-90\% concordance between the two genotype datasets (see Supplemental data for
+90\% concordance between the two genotype datasets (see Additional file 1 for
 detailed methods). These samples were removed from the dataset in addition to another
@@
 remained for GWAS analyses. The concordance between WGS data and SoySNP50K
-genotypes is included in Additional file 2 and averaged 98.3\% among retained
-lines. The phenotypic data used in this study can be found in Additional file 3.
+genotypes is included in Additional file 3 and averaged 98.3\% among retained
+lines. The phenotypic data used in this study can be found in Additional file 4.
 
@@
 using the same method that we used for converting
-SoySNP50K positions (see Supplemental data). For the analysis of
+SoySNP50K positions (see Additional file 1). For the analysis of
 qualitative traits, we only compared our results to signals at named classical
@@
 \begin{description}
-	\item[Additional file 1.csv] Metadata on the sequencing runs retrieved
+	\item[Additional file 1.pdf] Supplemental methods and results,
+		including Tables \ref{signals_table} to \ref{MATURITY.GROUP_gwas_table}
+		and Figures \ref{flower_color_manhattan} to \ref{protein_manhattan}
+		(PDF 12.2 MB)
+	\item[Additional file 2.csv] Metadata on the sequencing runs retrieved
 		from the Sequence Read Archive (SRA) (CSV 71 KB)
-	\item[Additional file 2.csv] Metadata on individual accessions following mapping
+	\item[Additional file 3.csv] Metadata on individual accessions following mapping
 		and comparison to SoySNP50K data (CSV 36KB)
-	\item[Additional file 3.csv] Phenotypic dataset used for GWAS analyses (CSV 135
+	\item[Additional file 4.csv] Phenotypic dataset used for GWAS analyses (CSV 135
 		KB)
@@
 
-The Illumina sequencing data used for the analyses are listed in Additional file 1
+The Illumina sequencing data used for the analyses are listed in Additional file 2
 and available through NCBI BioProjects \citeauthor{PRJNA257011}, \citeauthor{PRJNA289660}

### 2023-03-01 additional_files/supplemental_file_1.R

@@ -1,31 +1,0 @@
-# Generating the supplemental data on the SRA accessions metadata
-
-# Reading the metadata from file
-# DEPENDENCY: utilities/correct_sra_metadata.csv
-sra <- read.table("utilities/correct_sra_metadata.csv", sep = ",", header = TRUE, stringsAsFactors = FALSE)
-
-# Keeping only the necessary columns for the output file
-sra <- sra[, c("Run", "spots", "bases", "avgLength", "BioProject", "Model", "BioSample", "SampleName")]
-
-# Adding a column with the ID used as part of this work
-sra$study_ID <- NA
-
-# Loading the correspondence between SRA IDs and sample IDs
-# DEPENDENCY: utilities/srr_id_correspondence.txt
-bayer_ids <- read.table("utilities/srr_id_correspondence.txt", header = FALSE, stringsAsFactors = FALSE)
-
-# Filling the study_ID column
-for(i in 1:nrow(sra)) {
-	matching_row <- which(grepl(sra[i, "Run"], bayer_ids[[2]]))
-	stopifnot(length(matching_row) == 1)
-	sra[i, "study_ID"] <- bayer_ids[matching_row, 1]
-}
-
-# Checking that all SRA IDs matched a sample and that all 389 samples are represented
-stopifnot(all(!is.na(sra$study_ID)))
-stopifnot(length(unique(sra$study_ID)) == 389)
-
-# Writing the resulting table to file in CSV format
-write.table(sra, file = "additional_files/supplemental_file_1.csv", sep = ",",
-	    col.names = TRUE, row.names = FALSE, quote = FALSE)
-

### 2023-03-01 additional_files/supplemental_file_2.R

@@ -1,44 +1,31 @@
-# Supplemental file with metadata on the mapped reads for each sample
-library(parallel)
+# Generating the supplemental data on the SRA accessions metadata
 
-# Reading the Bayer IDs and their corresponding SRR IDs
+# Reading the metadata from file
+# DEPENDENCY: utilities/correct_sra_metadata.csv
+sra <- read.table("utilities/correct_sra_metadata.csv", sep = ",", header = TRUE, stringsAsFactors = FALSE)
+
+# Keeping only the necessary columns for the output file
+sra <- sra[, c("Run", "spots", "bases", "avgLength", "BioProject", "Model", "BioSample", "SampleName")]
+
+# Adding a column with the ID used as part of this work
+sra$study_ID <- NA
+
+# Loading the correspondence between SRA IDs and sample IDs
 # DEPENDENCY: utilities/srr_id_correspondence.txt
 bayer_ids <- read.table("utilities/srr_id_correspondence.txt", header = FALSE, stringsAsFactors = FALSE)
-colnames(bayer_ids) <- c("study_ID", "SRA_runs")
 
-# DEPENDENCY: stats obtained from samtools stats for all samples
-# Now we read the metadata obtained from the samtools stats command to add the number of reads
-# and number of aligned reads
-samtools_stats <- mclapply(bayer_ids$study_ID, function(x) {stats <- readLines(paste0("illumina_data/merged_bams/", x, "_stats.txt"))
-			   stats <- grep("^SN", stats, value = TRUE)
-			   nreads <- as.numeric(strsplit(grep("raw total sequences:", stats, value = TRUE), "\t")[[1]][3])
-			   nmapped <- as.numeric(strsplit(grep("reads mapped:", stats, value = TRUE), "\t")[[1]][3])
-			   list(nreads = nreads, nmapped = nmapped)},
-			   mc.cores = 10)
+# Filling the study_ID column
+for(i in 1:nrow(sra)) {
+	matching_row <- which(grepl(sra[i, "Run"], bayer_ids[[2]]))
+	stopifnot(length(matching_row) == 1)
+	sra[i, "study_ID"] <- bayer_ids[matching_row, 1]
+}
 
-names(samtools_stats) <- bayer_ids$study_ID
+# Checking that all SRA IDs matched a sample and that all 389 samples are represented
+stopifnot(all(!is.na(sra$study_ID)))
[…]

### 2023-03-01 additional_files/supplemental_file_3.R

@@ -1,15 +1,44 @@
-# Preparing a supplemental table with the phenotypic data used bu our GWAS analyses
+# Supplemental file with metadata on the mapped reads for each sample
+library(parallel)
 
-# Reading the main phenotypic data table
-phenodata <- read.table("phenotypic_data/phenotypic_data.csv", sep = ";",
-			header = TRUE, stringsAsFactors = FALSE)
+# Reading the Bayer IDs and their corresponding SRR IDs
+# DEPENDENCY: utilities/srr_id_correspondence.txt
+bayer_ids <- read.table("utilities/srr_id_correspondence.txt", header = FALSE, stringsAsFactors = FALSE)
+colnames(bayer_ids) <- c("study_ID", "SRA_runs")
 
-# Changing the names of some columns
-colnames(phenodata)[colnames(phenodata) == "bayer_id"] <- "study_ID"
+# DEPENDENCY: stats obtained from samtools stats for all samples
+# Now we read the metadata obtained from the samtools stats command to add the number of reads
+# and number of aligned reads
+samtools_stats <- mclapply(bayer_ids$study_ID, function(x) {stats <- readLines(paste0("illumina_data/merged_bams/", x, "_stats.txt"))
+			   stats <- grep("^SN", stats, value = TRUE)
+			   nreads <- as.numeric(strsplit(grep("raw total sequences:", stats, value = TRUE), "\t")[[1]][3])
+			   nmapped <- as.numeric(strsplit(grep("reads mapped:", stats, value = TRUE), "\t")[[1]][3])
+			   list(nreads = nreads, nmapped = nmapped)},
+			   mc.cores = 10)
 
-# Removing some other columns
-phenodata <- phenodata[, !colnames(phenodata) %in% c("TAXONOMY", "N_CDW")]
+names(samtools_stats) <- bayer_ids$study_ID
 
-# Writing the resulting table to file
-write.table(phenodata, file = "additional_files/supplemental_file_3.csv", sep = ";",
+# Adding these data to the data.frame
+bayer_ids$nreads <- sapply(samtools_stats, function(x) x$nreads)[bayer_ids$study_ID]
+bayer_ids$nmapped <- sapply(samtools_stats, function(x) x$nmapped)[bayer_ids$study_ID]
+
+# Reading the mapping coverage from the manifest files
+# DEPENDENCY: manifest files for all samples
[…]

### 2023-03-01 additional_files/supplemental_file_4.R

@@ -1,0 +1,15 @@
+# Preparing a supplemental table with the phenotypic data used bu our GWAS analyses
+
+# Reading the main phenotypic data table
+phenodata <- read.table("phenotypic_data/phenotypic_data.csv", sep = ";",
+			header = TRUE, stringsAsFactors = FALSE)
+
+# Changing the names of some columns
+colnames(phenodata)[colnames(phenodata) == "bayer_id"] <- "study_ID"
+
+# Removing some other columns
+phenodata <- phenodata[, !colnames(phenodata) %in% c("TAXONOMY", "N_CDW")]
+
+# Writing the resulting table to file
+write.table(phenodata, file = "additional_files/supplemental_file_4.csv", sep = ";",
+	    col.names = TRUE, row.names = FALSE, quote = FALSE)

### 2023-03-03 additional_files/additional_file_1.tex
Adjust figure numbering and order
Make a few changes in the text and figures upon proofreading the manuscript

@@
 We used Paragraph v. 2.4a \citep{chen2019}
-to generate a set of SV genotypes from the Illumina data of the 389 samples used in
-GWAS analyses.  The set of SVs used as input for Paragraph was generated using SVmerge by
+to generate a set of SV genotypes from the Illumina data of the 389 accessions.
+The set of SVs used as input for Paragraph was generated using SVmerge by
 merging variants discovered from the Illumina and ONT data with
@@
 chromosome end and then padding the variant representation using a script written by
-\cite{hickey2020}. The latter step was necessary for the variants to be used
-as input to Paragraph.
+\cite{hickey2020}.
 
-We ran Paragraph individually on each sample as recommended by the authors.
-We computed the average sequencing depth for each sample from the output of the
+We ran Paragraph individually on each sample as recommended by the authors.  We
+computed the average sequencing depth for each sample from the output of the
 \texttt{samtools coverage} command and used 20 times that value for the
@@
 \texttt{bcftools merge} in order to get a  single genotype file for the whole
-population. Finally, we filtered the genotyped SVs by setting the genotype calls made
-from less than two reads to missing and by removing variants with a minor allele
-frequency < 0.02 or a proportion of missing data > 0.5. The resulting dataset
-of 186,306 called SVs was used for GWAS.
+population. The resulting genotype calls were filtered as described in the main
+text.
 
@@
 The concordance between WGS data and SoySNP50K genotypes is included in
-Additional file 2 and shown in Figure \ref{concordance-histogram}.
+Additional file 3 and shown in Figure \ref{concordance-histogram}.
 
@@
 This section describes results at additional loci that were not discussed in
-the main text. A summary of the signals found at all loci using all four
+the main text. A summary of the signals found at all loci using all three GWAS
 methods is presented in Table \ref{signals_table}.
@@
 The analysis detected signals at the \emph{B} locus as the sole classical locus
-using all four methods (see main text and Figure
+using all three methods (see main text and Figure
 \ref{seed_coat_luster_dullshiny_manhattan}).  All methods also detected signals
@@
 in our dataset.
+
+\subsection*{Pubescence form -- additional loci}
+\label{sv-gwas-pubesence-form}
+
+The first GWAS analysis on pubescence form found signals on chromosomes Gm04
+and Gm15 with SNPs and \emph{k}-mers (Figure
+\ref{pubescence_form_all_manhattan}). However, upon further analysis of LD
+patterns, they appeared to be linked to the \emph{Pa1} locus (Figure
+\ref{pubescence_form_all_ld}).
 
@@
 oil content are also major QTL for seed protein content. However, none of the
-approaches detected a signal at this locus.
+approaches detected a signal at this locus (Figure \ref{protein_manhattan}).
 
@@
 a clear limitation to this GWAS given our relatively small sample size.
-
-\subsection*{Pubescence form -- additional loci}
-\label{sv-gwas-pubesence-form}
-
-The first GWAS analysis on pubescence form found signals on chromosomes Gm04
-and Gm15 with SNPs and \emph{k}-mers (Figure
-\ref{pubescence_form_all_manhattan}). However, upon further analysis of LD
-patterns, they appeared to be linked to the \emph{Pa1} locus (Figure
-\ref{pubescence_form_all_ld}).
 
@@
 \clearpage
-
-\gwastableone{FLOWER.COLOR}{flower color}
-\gwastabletwo{PUBESCENCE.COLOR}{pubescence color}
-\gwastabletwo{SEED.COAT.COLOR}{seed coat color}
-\gwastabletwo{STEM.TERMINATION.TYPE}{stem termination type}
-\gwastablethree{HILUM.COLOR}{hilum color}
-\gwastabletwo{POD.COLOR}{pod color}
-\gwastabletwo{PUBESCENCE.FORM}{pubescence form}
-\gwastableone{PUBESCENCE.DENSITY}{pubescence density}
-\gwastablethree{SEED.COAT.LUSTER}{seed coat luster}
-\gwastableone{MATURITY.GROUP}{maturity group}
 
@@
 			genes that have been suggested in previous studies.
-		\item[c] Most significant -log\textsubscript{10}(\emph{p}) reported by Platypus at this locus. A dash indicates that no signal
-			was detected by Platypus at this locus. The same applies for other methods (Paragraph, \emph{k}-mers) in
+		\item[c] Most significant -log\textsubscript{10}(\emph{p}) reported from SNPs and indels at this locus. A dash indicates that no signal
+			was detected from SNPs and indels at this locus. The same applies for other methods (SVs, \emph{k}-mers) in
 			their respective columns.
@@
 			{Description of loci associated with the traits studied and -log\textsubscript{10}(\emph{p}) of significant
-			signals detected in our study using four different genotypic datasets (Platypus, Paragraph, \emph{k}-mers).}\\
+			signals detected in our study using three different genotypic datasets (SNPs/indels, SVs, \emph{k}-mers).}\\
 			\label{signals_table}\\
@@
 			\toprule
-			Trait & Locus & Chromosome & -log\textsubscript{10}(\emph{p})\tnote{a} & Gene\tnote{b} & Platypus\tnote{c} & Paragraph & \emph{k}-mers  & Study \\
+			Trait & Locus & Chromosome & -log\textsubscript{10}(\emph{p})\tnote{a} & Gene\tnote{b} & SNPs/indels\tnote{c} & SVs & \emph{k}-mers  & Study \\
 			\midrule \endfirsthead
@@
 			\toprule
-			Trait & Locus & Chromosome & -log10(p) & Gene & Platypus & Paragraph & k-mers  & Study \\
+			Trait & Locus & Chromosome & -log10(p) & Gene & SNPs/indels & SVs & k-mers  & Study \\
 			\midrule
@@
 
+\gwastableone{FLOWER.COLOR}{flower color}
+\gwastabletwo{PUBESCENCE.COLOR}{pubescence color}
+\gwastabletwo{SEED.COAT.COLOR}{seed coat color}
+\gwastabletwo{STEM.TERMINATION.TYPE}{stem termination type}
+\gwastablethree{HILUM.COLOR}{hilum color}
+\gwastabletwo{POD.COLOR}{pod color}
+\gwastabletwo{PUBESCENCE.FORM}{pubescence form}
+\gwastableone{PUBESCENCE.DENSITY}{pubescence density}
+\gwastablethree{SEED.COAT.LUSTER}{seed coat luster}
+\gwastableone{MATURITY.GROUP}{maturity group}
+
 % Supplemental figures
@@
 {Vertical dotted lines in panels (c) and (d) indicate the location of the causal SV at this locus.}
+\ldplot{flower_color}{flower color}{\getvar{flower_color_ldkmers}}
 %\kmerplot{flower_color_W1}
@@
 % Figures for hilum color
-\manhattanplot{hilum_color_all}{hilum color (first GWAS)}{\getvar{hilum_color_all_nsamples}}
-\signalplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}%
-{Vertical dotted lines indicate the location of the \getvar{hilum_color_all_T_gene} gene associated with the locus.}
-\geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene}%
-{Vertical dotted lines in panels (b) and (d) indicate the location of the causal indel at this locus.}
-\kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene}}{\emph{T}}{hilum color (first GWAS)}{a causal indel}
-\signalplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}%
-{Vertical dotted lines indicate the boundaries of the tandem duplication/inversion identified as the causal variant at this locus.}
-
 \manhattanplot{hilum_color_blackbrown}{hilum color (second GWAS)}{\getvar{hilum_color_blackbrown_nsamples}}
@@
 \manhattanplot{pubescence_density}{pubescence density}{\getvar{pubescence_density_nsamples}}
+\signalplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}%
+{Vertical dotted lines indicate the boundaries of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene associated with the locus.}
+\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene.}%
+{The vertical dotted line in panel (d) indicates the location of the most significant \emph{k}-mer, which is associated with the causal CNV at this locus.
+The shaded gray rectangles show the boundaries of the causal CNV.}
 
 % Figures for seed coat luster
-\manhattanplot{seed_coat_luster_dullshiny}{seed coat luster (third GWAS)}{\getvar{seed_coat_luster_dullshiny_nsamples}}
 \signalplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}%
 {Vertical dotted lines indicate the boundaries of the causal CNV associated with the locus.}
-
-% Figures for maturity group
-\manhattanplot{maturity_group}{maturity group}{\getvar{maturity_group_nsamples}}
-
-% Figures for oil
-\manhattanplot{oil}{seed oil content}{\getvar{oil_nsamples}}
-
-% Figures for protein
-\manhattanplot{protein}{seed protein content}{\getvar{protein_nsamples}}
+\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV}%
+{The shaded gray rectangles show the boundaries of the causal CNV.}
+\manhattanplot{seed_coat_luster_dullshiny}{seed coat luster (third GWAS)}{\getvar{seed_coat_luster_dullshiny_nsamples}}
 
@@
 \signalplot{pubescence_form_all_Pa2}{pubescence form (first GWAS)}{\emph{Pa2}}{}
+\manhattanplot{pubescence_form_noerect}{pubescence form (second GWAS)}{\getvar{pubescence_form_noerect_nsamples}}
 
@@
 \ldplot{pod_color_blbr}{pod color (second GWAS)}{\getvar{pod_color_blbr_ldkmers}}
+\signalplot{pod_color_blbr_pdcGm15}{pod color (second GWAS)}{newly suggested \emph{pdcGm15}}{}
 
@@
 % Flower color
-\ldplot{flower_color}{flower color}{\getvar{flower_color_ldkmers}}
 
@@
 % Hilum color
+\manhattanplot{hilum_color_all}{hilum color (first GWAS)}{\getvar{hilum_color_all_nsamples}}
+\signalplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}%
+{Vertical dotted lines indicate the location of the \getvar{hilum_color_all_T_gene} gene associated with the locus.}
+\geneplot{hilum_color_all_T}{hilum color (first GWAS)}{\emph{T}}{\getvar{hilum_color_all_T_gene} gene}%
+{Vertical dotted lines in panels (b) and (d) indicate the location of the causal indel at this locus.}
+\kmerplot{hilum_color_all_T}{\getvar{hilum_color_all_T_gene}}{\emph{T}}{hilum color (first GWAS)}{a causal indel}
+\signalplot{hilum_color_all_I}{hilum color (first GWAS)}{\emph{I}}%
+{Vertical dotted lines indicate the boundaries of the tandem duplication/inversion identified as the causal variant at this locus.}
+
 \manhattanplot{hilum_color_rbr}{hilum color (third GWAS)}{\getvar{hilum_color_rbr_nsamples}}
@@
 % Pod color
-\signalplot{pod_color_blbr_pdcGm15}{pod color (second GWAS)}{newly suggested \emph{pdcGm15}}{}
-
-% Pubescence form
-\ldplot{pubescence_form_all}{pubescence form (first GWAS)}{\getvar{pubescence_form_all_ldkmers}}
-\manhattanplot{pubescence_form_noerect}{pubescence form (second GWAS)}{\getvar{pubescence_form_noerect_nsamples}}
 
 % Pubescence density
-\signalplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}%
-{Vertical dotted lines indicate the boundaries of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene associated with the locus.}
-\geneplot{pubescence_density_Ps}{pubescence density}{\emph{Ps}}{location of the causal CNV overlapping the \getvar{pubescence_density_Ps_gene} gene.}%
-{The vertical dotted line in panel (d) indicates the location of the most significant \emph{k}-mer, which is associated with the causal CNV at this locus.
-The shaded gray rectangles show the boundaries of the causal CNV.}
 
@@
 
-\geneplot{seed_coat_luster_dullshiny_B}{seed coat luster (third GWAS)}{\emph{B}}{location of the causal CNV}%
-{The shaded gray rectangles show the boundaries of the causal CNV.}
+% Pubescence form
+\ldplot{pubescence_form_all}{pubescence form (first GWAS)}{\getvar{pubescence_form_all_ldkmers}}
+
+% Figures for maturity group
+\manhattanplot{maturity_group}{maturity group}{\getvar{maturity_group_nsamples}}
+
+% Figures for oil
+\manhattanplot{oil}{seed oil content}{\getvar{oil_nsamples}}
+
+% Figures for protein
+\manhattanplot{protein}{seed protein content}{\getvar{protein_nsamples}}
 

### 2023-03-03 additional_files/main_text.tex
Make a few changes in the text and figures upon proofreading the manuscript

@@
 
-Genome-wide association studies (GWAS) are powerful statistical methods
-that detect associations between genotype and phenotype at genome scale. Despite
+Genome-wide association studies (GWAS) are powerful statistical methods that
+detect associations between genotype and phenotype at genome scale. Despite
 their power, GWAS frequently fail to pinpoint the causal variant or the gene
-controlling a trait at a given locus in crop species. Assessing genetic variants beyond
-single-nucleotide polymorphisms (SNPs) may alleviate this problem,
-for example by including structural variants (SVs) that may have a greater functional impact.
-In this study, we tested the potential
-of SV- and \emph{k}-mer-based GWAS in soybean by applying these methods to 13
-traits on which SNP-based GWAS had been performed previously.
-We assessed the performance of each GWAS approach based on results at loci for which
-the causal genes or variants were known from previous genetic studies. We found
-that \emph{k}-mer-based GWAS was the most versatile approach and the best at pinpointing
-causal variants or candidate genes based on the most significantly
-associated \emph{k}-mers. Given the traits that we analyzed, SV-based GWAS did not add any value
-compared to \textit{k}-mer-based GWAS and may not generally be worth the investment in time and
-computational resources required to discover and genotype such variants.
-Moreover, \textit{k}-mer-based analyses
-identified promising candidate genes for loci related
-to pod color, pubescence form, and resistance to the oomycete
-\textit{Phytophthora sojae}. Despite promising results,
-significant challenges remain regarding the downstream analysis of
-\emph{k}-mer-based GWAS, notably in associating significant \emph{k}-mers with
-sequence variation and ensuring that all variation types are adequately
-detected from \emph{k}-mer occurrence patterns. Together, our results suggest
-that coupling \emph{k}-mer- and SNP/indel-based GWAS may represent a
+controlling a trait at a given locus in crop species. Assessing genetic
+variants beyond single-nucleotide polymorphisms (SNPs) could alleviate this
+problem, for example by including structural variants (SVs).  In this study, we
+tested the potential of SV- and \emph{k}-mer-based GWAS in soybean by applying
+these methods to 13 traits.  We also performed conventional GWAS analysis based
+on SNPs and small indels for comparison.  We assessed the performance of each
+GWAS approach based on results at loci for which the causal genes or variants
+were known from previous genetic studies. We found that \emph{k}-mer-based GWAS
+was the most versatile approach and the best at pinpointing causal variants or
+candidate genes based on the most significantly associated \emph{k}-mers.
+Moreover, \textit{k}-mer-based analyses identified promising candidate genes
+for loci related to pod color, pubescence form, and resistance to the oomycete
+\textit{Phytophthora sojae}. In our dataset, SV-based GWAS did not add value
+compared to \textit{k}-mer-based GWAS and may not be worth the time and
+computational resources required to genotype SVs at population scale.  Despite
+promising results, significant challenges remain regarding the downstream
+analysis of \emph{k}-mer-based GWAS.  Notably, better methods are needed to
+associate significant \emph{k}-mers with sequence variation. Together, our
+results suggest that coupling \emph{k}-mer- and SNP/indel-based GWAS is a
 powerful approach for discovering candidate genes in crop species.
@@
 wheat \citep{maron2013}, and branching in maize \citep{studer2011}.  SVs have
-already been used in GWAS in studies with promising results
+already been used in GWAS with promising results
 \citep[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
-their inclusion in GWAS is the exception rather than the norm.
-
-Despite their phenotypic impact, SVs are still largely understudied compared to SNPs
-because of the difficulty of discovering and genotyping them with
-accuracy. Comprehensive detection of SVs requires a combination of several
-methods that often need to be tailored to the study species and data available
+their use in GWAS is the exception rather than the norm.
+
+Despite their phenotypic impact, SVs are still largely understudied compared to
+SNPs because discovering and genotyping them with accuracy is difficult.
+Comprehensive detection of SVs requires a combination of several methods that
+often need to be tailored to the study species and data available
 \citep{alkan2011, ho2020}. Even then, assessments of the performance of SV
@@
 improvements in computational approaches for genotyping SVs
-\citep[e.g.][]{siren2021, ebler2022}, such approaches remain relatively inaccurate
-compared to SNP genotyping workflows.
+\citep[e.g.][]{siren2021, ebler2022}, such approaches remain relatively
+inaccurate compared to SNP genotyping workflows.
 
@@
 represent such an alternative method. This approach relies on the presence or
-absence of \emph{k}-mers observed in sequence reads instead of genotypes for use in
+absence of \emph{k}-mers observed in sequence reads in place of variant genotypes for use in
 GWAS. Once a presence/absence table of \emph{k}-mers has been generated for the
@@
 location and can thus be used to query associations between genotype and
-phenotype at genomic locations that are not found in a reference genome.  A
+phenotype at genomic locations that are not included in a reference genome.  A
 few studies have applied this \emph{k}-mer-based GWAS approach
@@
 In this study, we assessed the potential of SV- and \emph{k}-mer-based GWAS in
-a major crop, soybean (\emph{Glycine max}).  In order to test these methods, we
-analyzed ten qualitative traits \citep{bandillo2017} and two quantitative
-traits \citep{bandillo2015} that have been previously studied using a SNP array
-developed for soybean \citep{song2013}. In addition, we analyzed a quantitative trait
-(horizontal resistance to \emph{Phytophthora sojae}) that has been previously
-studied using whole-genome sequencing (WGS) data \citep{deronne2022}.
-Since the underlying genes and causal variants are already known for several of
-these traits, they provide a good test case for the performance of GWAS methods.
-The objectives of our study were to:
+a major crop, soybean (\emph{Glycine max}). We also conducted conventional GWAS
+based on SNPs and indels. In order to test these methods, we analyzed ten
+qualitative traits \citep{bandillo2017} and two quantitative traits
+\citep{bandillo2015} that have been previously studied using a SNP array
+developed for soybean \citep{song2013}. In addition, we analyzed a quantitative
+trait (horizontal resistance to \emph{Phytophthora sojae}) that has been
+previously studied using whole-genome sequencing (WGS) data
+\citep{deronne2022}.  Since the underlying genes and causal variants are
+already known for several of these traits, they provide a good test case for
+the performance of GWAS methods.  The objectives of our study were to:
 
@@
 
-We analyzed a total of 13 traits in a population of 363 \textit{G. max}
-accessions using three GWAS approaches that differed in the genotypes used. The
-first approach used SNP and indel genotypes called by Platypus. Despite both
-SNPs and indels being included in this approach, we will refer to this approach
-as SNP-based for simplicity. Our SV-based workflow used various sources of data
+We analyzed a total of 13 traits in a population of 363 resequenced \textit{G.
+max} accessions using three GWAS approaches that differed in the genotypes
+used. The first approach used SNP and indel genotypes called by Platypus.
+Despite both SNPs and indels being used, we will refer to this approach as
+SNP-based for simplicity. Our SV-based workflow used various sources of data
 for SV discovery and a single program (Paragraph) for genotyping SVs at
-population scale. Finally, the third GWAS approach relied on the
-presence/absence of \textit{k}-mers as genotypes.  Because of the large number
-of \textit{k}-mers analyzed, we only report results (e.g. in Manhattan plots)
+population scale from Illumina WGS data. Finally, the third GWAS approach
+relied on the presence/absence of \textit{k}-mers in the WGS data as genotypes.
+Because of the large number of \textit{k}-mers analyzed, we only report results
 for those that were considered significantly associated with traits of
@@
 
-Since two or three different GWAS analyses were conducted for some of the traits, a
-total of 22 GWAS analyses were performed using each of the three approaches.
-Given the large volume of data that this represents, we focus here on the
-most noteworthy results. Readers interested in results at loci that are not
-discussed here are referred to Additional file 1. In particular, Table \ref{signals_table} (Additional file 1)
-provides a summary of the \textit{p}-values obtained by each approach at each
-of the loci considered in this study.
+Since two or three different GWAS analyses were conducted for some of the
+traits, a total of 22 GWAS analyses were performed using each of the three
+approaches.  Given the large volume of data that this represents, we focus here
+on the most noteworthy results. Readers interested in results at loci that are
+not discussed here are referred to Additional file 1. In particular, Table
+\ref{signals_table} (Additional file 1) provides a summary of the
+\textit{p}-values obtained by each approach at each of the loci considered in
+this study.
 
@@
 significant \textit{k}-mers did occur within the sequence of this gene, they
-did not correspond to the causal variant (Additional file 1: Figure
+did not correspond to the documented causal variant (Additional file 1: Figure
 \ref{pubescence_color_all_T_gene}d).  Still, the fourth and fifth most strongly
-associated \textit{k}-mers corresponded to the causal variant (Additional file 1: Figures
+associated \textit{k}-mers corresponded to this causal variant (Additional file 1: Figures
 \ref{pubescence_color_all_T_gene}d and \ref{pubescence_color_all_T_kmers}). We
@@
 \citep{yan2020}, it was the fourth most significantly associated variant for
-the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d,Additional file 1: Figure 
+the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d, Additional file 1: Figure 
 \ref{pubescence_color_nogray_Td_gene}).  Interestingly, some of the accessions
@@
 the reference, we reasoned that the CNV may have been found by our pipeline but
-filtered out later on. Indeed, found a 42.8-kb duplication corresponding
+filtered out later on. Indeed, we found a 42.8-kb duplication corresponding
 to the one described above among the genotype calls made by smoove. An association
@@
 	SNPs associated with appressed pubescence, which may represent causal variants
-	at this locus. Panels (c) and (d) share the same axis.}
+	at this locus. Panels (c) and (d) share the same x-axis.}
 \end{figure}
@@
 In conclusion, we used SV-based and \emph{k}-mer-based GWAS to study ten
-qualitative traits and three quantitative traits in soybean cultivars and
-compared the results to traditional SNP-based GWAS. While \emph{k}-mer-based
+qualitative traits and three quantitative traits in soybean and
+compared the results to conventional SNP-based GWAS. While \emph{k}-mer-based
 GWAS proved to be a powerful approach in pinpointing the causal variation or
@@
 combination with SV datasets. Given the large number of samples for which WGS
-data are now available in several species, applying \emph{k}-mer-based GWAS to
+data is now available in several species, applying \emph{k}-mer-based GWAS to
 leverage already existing sequencing and phenotypic data appears feasible and
@@
 genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
-from the mapped Illumina reads for all accessions. Detailed methods regarding
-the SV discovery and genotyping steps can be found in Additional file 1.
+from the mapped Illumina reads for all accessions. We filtered the genotyped
+SVs by setting the genotype calls made from less than two reads to missing and
+by removing variants with a minor allele frequency < 0.02 or a proportion of
+missing data > 0.5. The resulting dataset of 186,306 genotyped SVs was used for
+GWAS. Detailed methods regarding the SV discovery and genotyping steps can be
+found in Additional file 1.
 
@@
 	\item fraction of missing genotypes $\leq$ 0.5
-	\item heterozygosity rate $\leq$ 0.2.
+	\item heterozygosity rate $\leq$ 0.2
 \end{enumerate}
@@
 genotypes is included in Additional file 3 and averaged 98.3\% among retained
-lines. The phenotypic data used in this study can be found in Additional file 4.
+lines (Additional file 1: Figure \ref{concordance-histogram}). The phenotypic data used in this study can be found in Additional file 4.
 
@@
 and a tailored analysis was therefore required to associate them with genomic
-coordinates. To do so, we queried the reads (in BAM format) of all accessions for matches
+coordinates. To do so, we queried the mapped reads (in BAM format) of all accessions for matches
 to significant \emph{k}-mers for a given trait. Because this is a
@@
 
-We used contiguous regions of significant associations, which we called
+We used contiguous regions of significant variants, which we called
 ``signals'', as a basis to compare our results with those obtained by previous
@@
 occurred within 250 kb of one another.  We did consider the length of deletions
-or \emph{k}-mers in defining the distance between significant associations.
-
-For the GWAS analyses computed from Platypus genotype calls, we retrieved all
+or \emph{k}-mers in defining the distance between significant variants.
+
+For the GWAS analyses computed from SNP/indel genotype calls, we retrieved all
 markers that had initially been pruned in a region ranging from 50 kb
@@
 
-All additional files can be accessed on figshare \citep{kmer-gwas-figshare}.
-
 \begin{description}

### 2023-03-03 additional_files/manuscript.tex
Make a few changes in the text and figures upon proofreading the manuscript

@@
 	#4
-	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
-	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
+	Gray shaded rectangles indicate the region delimited by the top 5\% (SVs) or top 1\% (SNPs/indels and \emph{k}-mers) associations in the signal region.
+	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
 	\emph{p}-values were computed after signal discovery.
@@
 	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
-	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
+	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
 	\emph{p}-values were computed after signal discovery.

### 2023-03-03 figures/gene_plot.R

@@ -45,7 +45,7 @@ grid.newpage()
 grid.draw(output_grob)
 
 # Add labels for the panels
-grid.text(c("(a)", "(b) Platypus", "(c) Paragraph"),
+grid.text(c("(a)", "(b) SNPs/indels", "(c) SVs"),
 	  x = unit(0.09, "npc"),
 	  y = unit(c(0.993, 0.8725, 0.57), "npc"),
 	  hjust = 0,

### 2023-03-03 figures/manhattan_plot.R

@@ -6,7 +6,7 @@ suppressMessages(library(gwask))
 trait <- commandArgs(trailingOnly = TRUE)[1]
 
 programs <- c("platypus", "paragraph", "kmers")
-program_labels <- c("Platypus", "Paragraph", "kmers")
+program_labels <- c("SNPs/indels", "SVs", "kmers")
 
 # Outputting to a PNG file
 png(paste0("figures/", trait, "_manhattan.png"), width = 9, height = 10, units = "in", res = 100)

### 2023-03-03 figures/signal_plot.R

@@ -57,7 +57,7 @@ grid.newpage()
 grid.draw(output_grob)
 
 # Add labels for the panels
-grid.text(c("(a)", "(b) Platypus", "(c) Paragraph"),
+grid.text(c("(a)", "(b) SNPs/indels", "(c) SVs"),
 	  x = unit(0.09, "npc"),
 	  y = unit(c(0.993, 0.8725, 0.57), "npc"),
 	  hjust = 0,

### 2023-03-03 additional_files/additional_file_1.tex
Pass text through aspell and correct a few other mistakes; manuscript considered ready, version sent to co-authors for approval

@@
 
-\subsection*{Seed coat luster -- loci \textit{Hps} \textit{B?} and \textit{I}}
+\subsection*{Seed coat luster -- loci \textit{Hps}, \textit{B?} and \textit{I}}
 \label{sv-gwas-seed-coat-luster}
@@
 			{Description of loci associated with the traits studied and -log\textsubscript{10}(\emph{p}) of significant
-			signals detected in our study using three different genotypic datasets (SNPs/indels, SVs, \emph{k}-mers).}\\
+			signals detected in our study using three different genotype datasets (SNPs/indels, SVs, \emph{k}-mers).}\\
 			\label{signals_table}\\

### 2023-03-27 additional_files/main_text.tex
Pass text through aspell and correct a few other mistakes; manuscript considered ready, version sent to co-authors for approval
Add README and LICENSE; add reference to Tuteja and Vodkin (2008)'s contig and SoySNP50K data in manuscript

@@
 
-\subsection*{Loci with known genes or causal mutations}
+\subsection*{Loci with known genes or causal variants}
 \label{sv-gwas-results-cloned-genes}
@@
 has suggested Glyma.12g213900 as a candidate for \textit{Pa1} based on data
-published by \cite{bandillo2017}. This gene encodes a Myb transcription factor
+published by \cite{bandillo2017}. This gene encodes a MYB transcription factor
 annotated as playing a role in trichome branching. Indeed, the most significant
@@
 
+The \textit{I} locus contig assembled by \cite{tuteja2008} is available from
+NCBI (\url{https://www.ncbi.nlm.nih.gov/nuccore/EF623854}).
+
+SoySNP50K genotype calls are available from SoyBase (\url{https://soybase.org/snps/}).
+
 The SVs identified from Oxford Nanopore data by \cite{lemay2022} are available

### 2023-03-28 additional_files/main_text.tex
Last changes to manuscript before submitting code to figshare

@@
 Marc-André Lemay\textsuperscript{1,2,3},
-Jonas A. Sibbesen\textsuperscript{4},
 Maxime de Ronne\textsuperscript{1,2,3},
@@
 \textsuperscript{3} Centre de recherche et d’innovation sur les végétaux, Université Laval, Québec, QC, Canada
-
-\textsuperscript{4} Center for Health Data Science, University of Copenhagen, Denmark
 
@@
 		\url{https://github.com/malemay/gwask}
+	\item Software for computing the LD between \textit{k}-mers
+		based on presence/absence: \url{https://github.com/malemay/kmers_ld}
 	\item Forked svmu version that was used for the analyses shown in this
-		paper: \url{https://github.com/malemay/svmutools}
+		paper: \url{https://github.com/malemay/svmu}
 	\item svmutools R package for converting the output of svmu to VCF
@@
 \end{itemize}
+
+The code associated with the results presented in this manuscript is archived
+on figshare \citep{kmer-gwas-code}.
 
@@
 
-Conception and design of the study: MAL, JAS, MDR, RB, FB.
+Conception and design of the study: MAL, MDR, RB, FB.
 Phenotypic data acquisition (resistance to \textit{Phytophthora sojae}): MDR, RB.
 Data analysis: MAL.
-Data interpretation: MAL, JAS, FB.
+Data interpretation: MAL, FB.
 Software writing: MAL.
-Manuscript drafting: MAL, JAS, FB.
+Manuscript drafting: MAL, FB.
 All authors have revised the manuscript and approved its submission.
@@
 
-We would like to thank Yoav Voichek for help regarding \textit{k}-mer GWAS
-analysis.  We also thank Brian Boyle and Martine Jean for valuable comments
+We would like to thank Jonas A. Sibbesen for providing very helpful comments
+on the methodology used and on a previous version of this manuscript.
+We also thank Yoav Voichek for help regarding \textit{k}-mer GWAS
+analysis as well as Brian Boyle and Martine Jean for valuable comments
 regarding the methodology. We are thankful to the Digital Research Alliance of

### 2023-03-28 additional_files/additional_file_1.tex
Fix some typos upon proofreading the manuscript

@@
 	\item classified as unresolved breakends (\texttt{SVTYPE=BND})
-	\item containing at least one N in either the reference or alternate allele sequence
+	\item containing at least one N in the alternate allele sequence
 \end{itemize}
@@
 We used a set of tools that we assembled in an R package called svmutools
-\url{https://github.com/malemay/svmutools} to convert the svmu output to VCF
+(\url{https://github.com/malemay/svmutools}) to convert the svmu output to VCF
 format. We limited ourselves to processing deletions and insertions because of
@@
 for which this could be the case, we compared the genotypes of SNPs obtained
-from the SoySNP50K array \cite{song2013} to those predicted from the Illumina WGS
+from the SoySNP50K array \citep{song2013} to those predicted from the Illumina WGS
 data. We did this for the 385 samples that had a match in the GRIN database. Briefly, we
@@
 at the \emph{W1} and \emph{O} loci studied by \cite{bandillo2017} (Figures
-\ref{hilum_color_all_manhattan}, \ref{hilum_color_blackbrown_manhattan},
+\ref{hilum_color_blackbrown_manhattan} \ref{hilum_color_all_manhattan},
 \ref{hilum_color_rbr_manhattan}).

### 2023-03-28 additional_files/main_text.tex
Fix some typos upon proofreading the manuscript

@@
 	In the case of SNPs, blue points denote markers used in the
-	original analysis, whereas red points denote points that had originally
+	original analysis, whereas red points denote markers that had originally
 	been pruned but whose \emph{p}-values were computed after signal
@@
 % Pubescence density - Ps locus
-\subsubsection*{Pubescence density -- Ps locus}
+\subsubsection*{Pubescence density -- \textit{Ps} locus}
 \label{sv-gwas-main-results-pubescence-density-ps}
@@
 results may be due to the presence of some unresolved sequence (a stretch of
-4,635 'N' nucleotides) over that interval in the reference sequence that we used.
-Because our SV filtering process removed any SV overlapping 'N' nucleotides in
-the reference, we reasoned that the CNV may have been found by our pipeline but
+4,635 N nucleotides) over that interval in the reference sequence that we used.
+Because our SV filtering process removed any alternate allele containing N nucleotides,
+we reasoned that the CNV may have been found by our pipeline but
 filtered out later on. Indeed, we found a 42.8-kb duplication corresponding
@@
 % Seed coat luster - B locus
-\subsubsection*{Seed coat luster -- B locus}
+\subsubsection*{Seed coat luster -- \textit{B} locus}
 \label{sv-gwas-main-results-seed-coat-luster-b}
@@
 candidate for \emph{L1}. This gene may therefore represent an interesting
-candidate for this locus for future studies.
+candidate at this locus for future studies.
 
@@
 GWAS analysis comparing accessions with black and brown pods found a previously
-undocumented signal on chromosome Gm15. \textit{k}-mers at this new locus
+undocumented signal on chromosome Gm15 (Additional file 1: Figure \ref{pod_color_blbr_manhattan}).
+\textit{k}-mers at this new locus
 showed no LD with \textit{k}-mers at the known \textit{L1} locus and may
-therefore represent a distinct locus (Additional file 1: Figures \ref{pod_color_blbr_manhattan} and
+therefore represent a distinct locus (Additional file 1: Figure 
 \ref{pod_color_blbr_ld}). The observed signal overlaps a total of 27 genes
@@
 individually for each accession and then aligned to the Williams82 reference
-assembly using bwa mem with default parameters.  Aligned assembled sequences
+assembly using bwa-mem with default parameters.  Aligned assembled sequences
 that overlapped loci of interest were realigned using multiple sequence

### 2023-03-28 additional_files/manuscript.tex
Fix some typos upon proofreading the manuscript

@@
 	Gray shaded rectangles indicate the region delimited by the top 5\% (SVs) or top 1\% (SNPs/indels and \emph{k}-mers) associations in the signal region.
-	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
+	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote markers that had originally been pruned but whose
 	\emph{p}-values were computed after signal discovery.
@@
 	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
-	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
+	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote markers that had originally been pruned but whose
 	\emph{p}-values were computed after signal discovery.

### 2023-03-28 additional_files/main_text.tex
Make a few changes; version submitted to biorxiv

@@
 \textsuperscript{3} Centre de recherche et d’innovation sur les végétaux, Université Laval, Québec, QC, Canada
+
+\textbf{Contact}: marc-andre.lemay.2@ulaval.ca, francois.belzile@fsaa.ulaval.ca
 
@@
 		and Figures \ref{flower_color_manhattan} to \ref{protein_manhattan}
-		(PDF 12.2 MB)
+		(PDF 12.9 MB)
 	\item[Additional file 2.csv] Metadata on the sequencing runs retrieved

### 2023-03-28 additional_files/manuscript.tex
First version with BMC template; abstract and background sections included

--- 2023-03-03 additional_files/manuscript.tex
@@
-\documentclass[12pt]{report}
+%% BioMed_Central_Tex_Template_v1.06
+%%                                      %
+%  bmc_article.tex            ver: 1.06 %
+%                                       %
 
-\usepackage[utf8]{inputenc} % declare encoding as utf8
-\usepackage[T1]{fontenc} % declare encoding as utf8
-\usepackage{graphicx} % to enable \includegraphics
-\usepackage{flafter} % to make sure floats appear after their position in text
-\usepackage{pdflscape} % allow some pages to be in landscape orientation
-\usepackage{layout} % for debugging purposes, looking at page layout
-\usepackage[margin=1in]{geometry} % leaves more space for text body
-\usepackage[hidelinks]{hyperref} % hyperlinks from list of figures/tables
-\usepackage[labelfont=bf]{caption} % format captions and make links go to figure instead of caption
-\usepackage{csvsimple} % read and display csv files as tables
-\usepackage{booktabs} % prettier tables
-\usepackage{threeparttable} % easy addition of footnotes below tables
-\usepackage{textcomp} % makes the \textdegree command available
-\usepackage{tocloft} % formatting the lists of tables and figures
-\usepackage{xcolor} % enable text coloring
-\usepackage{calc} % makes \widthof available to set the width of the last column
-\usepackage{array} % using command \arraybackslash to restore the definition of \\ after \centering
-\usepackage{longtable} % allow tables to span more than one page
-\usepackage{threeparttablex} % threeparttable for longtable
-\usepackage{capt-of} % put caption on next page
-\usepackage{ccaption}
-\usepackage{listings}
-\usepackage{datatool} % allows fetching key-value pairs from a text file
-\usepackage[sectionbib]{natbib}
-\usepackage{chapterbib}
-\usepackage{setspace}
+%%% additional documentclass options:
+%  [doublespacing]
+%  [linenumbers]   - put the line numbers on margins
 
-% Remove indent for new paragraphs
-\setlength{\parindent}{0pt}
+%%% loading packages, author definitions
 
-% Adjust line spacing to 1.5
-\onehalfspacing
+%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
+\documentclass{bmcart}[doublespacing, linenumbers]
 
-% Setting the separator to use with datatool
-\DTLsetseparator{ = }
+%%% Load packages
+\usepackage{amsthm,amsmath}
+\RequirePackage[numbers]{natbib}
+%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
+\RequirePackage{hyperref}
+\usepackage[utf8]{inputenc} %unicode support
+%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
+%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
 
-\graphicspath{ {../figures/} } % tell Latex to look for figures in figures/
+\def\includegraphic{}
+\def\includegraphics{}
 
-% Defining new environments as I will always want the figures centered
-\newenvironment{cfigure}
-	{\begin{figure} \centering}
-	{\end{figure}}
+%%% Put your definitions there:
+\startlocaldefs
+\endlocaldefs
 
-\newenvironment{lsfigure}
-	{\begin{landscape} \begin{figure} \centering}
-	{\end{figure} \end{landscape}}
+%%% Begin ...
+\begin{document}
 
-\setcounter{totalnumber}{1} % Allow a maximum of one float (table/figure) per page
+%%% Start of article front matter
+\begin{frontmatter}
 
-\setlength{\parskip}{2ex}
+\begin{fmbox}
+\dochead{Research}
 
-% Defining a command to generate a table of phenotypic data for a single gwas
-\newcommand{\gwastableone}[2]{
-\begin{table}
-	\caption[Phenotypic data used for the GWAS analysis of #2]%
-	{Phenotype frequency and numerical coding of phenotypic data used for the GWAS analysis of #2.
-	``NA'' indicates a missing value in the source data.
-	Dashes indicate that this phenotype was not used in the GWAS analysis.}
-	\label{#1_gwas_table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter the title of your article here     %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-	\centering
-	\begin{threeparttable}\small
+\title{k-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean}
 
-		\csvreader[head to column names,
-		/csv/separator=semicolon,
-		tabular = llc,
-		table head = \toprule Phenotype & Frequency & GWAS \\ \midrule,
-		table foot = \bottomrule]%
-		{../tables/#1_gwas_table.csv}%
-		{}%
-		{\Value & \Frequency & \GWASa}
+\author[
+  addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
+  email={marc-andre.lemay.2@ulaval.ca}   % email address
+]{\inits{M.A.L.}\fnm{Marc-André} \snm{Lemay}}
 
-	\end{threeparttable}
+\author[
+  addressref={aff1},
+  email={richard.belanger@fsaa.ulaval.ca}
+]{\inits{R.B.}\fnm{Richard} \snm{Bélanger}}
 
-\end{table}
-\clearpage
+\address[id=aff1]{%                           % unique id
+  \orgdiv{Département de phytologie},             % department, if any
+  \orgname{Université Laval},          % university, etc
+  \city{Québec},                              % city
+  \cny{QC, Canada}                                    % country
 }
 
-% Defining a command to generate a table of phenotypic data for two gwas
-\newcommand{\gwastabletwo}[2]{
-\begin{table}
-	\caption[Phenotypic data used for the GWAS analyses of #2]%
-	{Phenotype frequency and numerical coding of phenotypic data used in two GWAS analyses of #2.
-	``NA'' indicates a missing value in the source data.
-	Dashes indicate that this phenotype was not used in the GWAS analysis.}
-	\label{#1_gwas_table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Enter short notes here                   %%
+%%                                          %%
+%% Short notes will be after addresses      %%
+%% on first page.                           %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-	\centering
-	\begin{threeparttable}\small
+%\begin{artnotes}
+%%\note{Sample of title note}     % note to the article
+%\note[id=n1]{Equal contributor} % note, connected to author
+%\end{artnotes}
 
-		\csvreader[head to column names,
-		/csv/separator=semicolon,
-		tabular = llcc,
-		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2\\ \midrule,
-		table foot = \bottomrule]%
-		{../tables/#1_gwas_table.csv}%
-		{}%
-		{\Value & \Frequency & \GWASa & \GWASb}
+\end{fmbox}% comment this for two column layout
 
-	\end{threeparttable}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                           %%
+%% The Abstract begins here                  %%
+%%                                           %%
+%% Please refer to the Instructions for      %%
+%% authors on https://www.biomedcentral.com/ %%
+%% and include the section headings          %%
+%% accordingly for your article type.        %%
+%%                                           %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\begin{abstractbox}
+
+\begin{abstract} % abstract
+\parttitle{Background}
+Genome-wide association studies (GWAS) are powerful statistical methods that
+detect associations between genotype and phenotype at genome scale. Despite
+their power, GWAS frequently fail to pinpoint the causal variant or the gene
+controlling a trait at a given locus in crop species. Assessing genetic
+variants beyond single-nucleotide polymorphisms (SNPs) could alleviate this
+problem, for example by including structural variants (SVs).
+
+\parttitle{Results}
+In this study, we tested the potential of SV- and \emph{k}-mer-based GWAS in
+soybean by applying these methods to 13 traits.  We also performed
+conventional GWAS analysis based on SNPs and small indels for
+comparison.  We assessed the performance of each GWAS approach based on
+results at loci for which the causal genes or variants were known from
+previous genetic studies. We found that \emph{k}-mer-based GWAS was the
+most versatile approach and the best at pinpointing causal variants or
+candidate genes based on the most significantly associated
+\emph{k}-mers.  Moreover, \textit{k}-mer-based analyses identified
+promising candidate genes for loci related to pod color, pubescence
+form, and resistance to the oomycete \textit{Phytophthora sojae}. In
+our dataset, SV-based GWAS did not add value compared to
+\textit{k}-mer-based GWAS and may not be worth the time and
+computational resources required to genotype SVs at population scale.
+
+\parttitle{Conclusions}
+Despite promising results, significant challenges remain regarding the
+downstream analysis of \emph{k}-mer-based GWAS.  Notably, better
+methods are needed to associate significant \emph{k}-mers with sequence
+variation. Together, our results suggest that coupling \emph{k}-mer-
+and SNP/indel-based GWAS is a powerful approach for discovering
+candidate genes in crop species.
+
+\end{abstract}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% The keywords begin here                  %%
+%%                                          %%
+%% Put each keyword in separate \kwd{}.     %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{keyword}
+\kwd{Genome-wide association studies}
+\kwd{\textit{k}-mer GWAS}
+\kwd{Soybean genomics}
+\kwd{Crop genomics}
+\kwd{\textit{Phytophthora sojae}}
+\kwd{Structural variation}
+\end{keyword}
+
+% MSC classifications codes, if any
+%\begin{keyword}[class=AMS]
+%\kwd[Primary ]{}
+%\kwd{}
+%\kwd[; secondary ]{}
+%\end{keyword}
+
+\end{abstractbox}
+%
+%\end{fmbox}% uncomment this for two column layout
+
+\end{frontmatter}
+
+%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
+% <put your article body there>
+
+%%%%%%%%%%%%%%%%
+%% Background %%
+%%
+\section*{Background}
+Genome-wide association studies (GWAS) are analytical approaches that detect
+statistical associations between phenotypic observations for a trait of
+interest and the genotypes of variants distributed throughout the
+genome \cite{tam2019}. This approach has been used for over 15 years
+\cite{visscher2012} and has enabled significant advances in our understanding
+of the genetic architecture of traits involved in human health
+\cite[e.g.][]{wellcome2007gwas} and crop \cite{tibbs2021} and animal
+\cite{ma2019} production, among others.
+
+Despite their power and convenience for finding significantly associated loci,
+GWAS analyses frequently fail to pinpoint the gene(s) associated with a trait and
+the causal variant(s) involved \cite{tam2019}. One possible reason for this
+is that many GWAS analyses do not include the causal variant itself in their variant dataset
+\cite{tibbs2021}. This could be either because the analysis
+is restricted to a particular type of variant, such as single-nucleotide
+polymorphisms (SNPs), or because the set of markers/variants genotyped in the
+association panel is limited by the technology being used
+\cite[e.g.][]{bandillo2015, sonah2015}. Despite this limitation, variants
+located near the gene controlling the trait often do appear statistically
+associated with the phenotype through linkage disequilibrium (LD) with the
+causal variant \cite{korte2013}.  A typical workflow then involves the
+identification of candidate genes within the haplotype blocks defined by the
+statistically associated markers and follow-up functional analyses to confirm
+that a given gene controls the trait \cite[e.g.][]{wang2018, liu2020ps}.
+
+The inclusion of structural variants (SVs) as genotype data in GWAS may partly overcome
+this issue by providing a set of variants with a potentially high functional impact.
+SVs include any variant involving a difference of at least 50 nucleotides, such as
+deletions, insertions, or inversions \cite{ho2020}. Such variants are known to
+have large phenotypic impacts by disrupting coding or regulatory sequences
+\cite{marroni2014}, and have been identified as causal variants in traits such
+as resistance to soybean cyst nematode \cite{cook2012}, aluminum tolerance in
+wheat \cite{maron2013}, and branching in maize \cite{studer2011}.  SVs have
+already been used in GWAS with promising results
+\cite[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
+their use in GWAS is the exception rather than the norm.
+
+Despite their phenotypic impact, SVs are still largely understudied compared to
+SNPs because discovering and genotyping them with accuracy is difficult.
+Comprehensive detection of SVs requires a combination of several methods that
+often need to be tailored to the study species and data available
+\cite{alkan2011, ho2020}. Even then, assessments of the performance of SV
+discovery and genotyping generally reveal subpar sensitivity and precision
+\cite[e.g.][]{cameron2019, chaisson2019, kosugi2019}. Moreover, despite recent
+improvements in computational approaches for genotyping SVs
+\cite[e.g.][]{siren2021, ebler2022}, such approaches remain relatively
+inaccurate compared to SNP genotyping workflows.
+
+Alternative approaches that allow one to assess a large spectrum
+of variants in GWAS without genotyping SVs at population scale could be useful.
+\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
+represent such an alternative method. This approach relies on the presence or
+absence of \emph{k}-mers observed in sequence reads in place of variant genotypes for use in
+GWAS. Once a presence/absence table of \emph{k}-mers has been generated for the
+population under study, this table can be used in association analyses
+similarly to SNP or SV genotype calls. One advantage of using \emph{k}-mers is
+that they can act as a molecular signature for any type of variant as long as
+these variants result in presence/absence of \emph{k}-mers of a given length.
+Another advantage is that \emph{k}-mers are not tied to a specific genomic
+location and can thus be used to query associations between genotype and
+phenotype at genomic locations that are not included in a reference genome.  A
+few studies have applied this \emph{k}-mer-based GWAS approach
+\cite[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
+developed \cite{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
+such approaches have not been widely adopted yet.
+
+In this study, we assessed the potential of SV- and \emph{k}-mer-based GWAS in
+a major crop, soybean (\emph{Glycine max}). We also conducted conventional GWAS
+based on SNPs and indels. In order to test these methods, we analyzed ten
+qualitative traits \cite{bandillo2017} and two quantitative traits
+\cite{bandillo2015} that have been previously studied using a SNP array
+developed for soybean \cite{song2013}. In addition, we analyzed a quantitative
+trait (horizontal resistance to \emph{Phytophthora sojae}) that has been
+previously studied using whole-genome sequencing (WGS) data
+\cite{deronne2022}.  Since the underlying genes and causal variants are
+already known for several of these traits, they provide a good test case for
+the performance of GWAS methods.  The objectives of our study were to:
+
+\begin{enumerate}
+	\item Assess the potential of SV- and \emph{k}-mer-based GWAS approaches to
+		pinpoint genes and causal variants associated with loci whose
+		underlying genes or variants are already known.
+	\item Identify potential candidate genes and/or causal variants at loci
+		whose underlying genes or candidate variants are not yet known.
+	\item Develop a set of computational tools for the downstream analysis
+		of significantly associated \emph{k}-mers once these have been
+		identified.
+\end{enumerate}
+
+\section*{Results and discussion}
+Text for this section\ldots
+\subsection*{Sub-heading for section}
+Text for this sub-heading\ldots
+\subsubsection*{Sub-sub heading for section}
+Text for this sub-sub-heading\ldots
+\paragraph*{Sub-sub-sub heading for section}
+Text for this sub-sub-sub-heading\ldots
+
+\section*{Conclusions}
+
+\section*{Methods}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Backmatter begins here                   %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{backmatter}
+
+\section*{Acknowledgements}%% if any
+Text for this section\ldots
+
+\section*{Funding}%% if any
+Text for this section\ldots
+
+\section*{Abbreviations}%% if any
+Text for this section\ldots
+
+\section*{Availability of data and materials}%% if any
+Text for this section\ldots
+
+\section*{Ethics approval and consent to participate}%% if any
+Text for this section\ldots
+
+\section*{Competing interests}
+The authors declare that they have no competing interests.
+
+\section*{Consent for publication}%% if any
+Text for this section\ldots
+
+\section*{Authors' contributions}
+Text for this section \ldots
+
+\section*{Authors' information}%% if any
+Text for this section\ldots
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                  The Bibliography                       %%
+%%                                                         %%
+%%  Bmc_mathpys.bst  will be used to                       %%
+%%  create a .BBL file for submission.                     %%
+%%  After submission of the .TEX file,                     %%
+%%  you will be prompted to submit your .BBL file.         %%
+%%                                                         %%
+%%                                                         %%
+%%  Note that the displayed Bibliography will not          %%
+%%  necessarily be rendered by Latex exactly as specified  %%
+%%  in the online Instructions for Authors.                %%
+%%                                                         %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% if your bibliography is in bibtex format, use those commands:
+\bibliographystyle{vancouver} % Style BST file (bmc-mathphys, vancouver, spbasic).
+\bibliography{references}      % Bibliography file (usually '*.bib' )
+% for author-year bibliography (bmc-mathphys or spbasic)
+% a) write to bib file (bmc-mathphys only)
+% @settings{label, options="nameyear"}
+% b) uncomment next line
+%\nocite{label}
+
+% or include bibliography directly:
+% \begin{thebibliography}
+% \bibitem{b1}
+% \end{thebibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Figures                       %%
+%%                               %%
+%% NB: this is for captions and  %%
+%% Titles. All graphics must be  %%
+%% submitted separately and NOT  %%
+%% included in the Tex document  %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% Do not use \listoffigures as most will included as separate files
+
+\section*{Figures}
+  \begin{figure}[h!]
+  \caption{Sample figure title}
+\end{figure}
+
+\begin{figure}[h!]
+  \caption{Sample figure title}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Tables                        %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Use of \listoftables is discouraged.
+%%
+\section*{Tables}
+\begin{table}[h!]
+\caption{Sample table title. This is where the description of the table should go}
+  \begin{tabular}{cccc}
+    \hline
+    & B1  &B2   & B3\\ \hline
+    A1 & 0.1 & 0.2 & 0.3\\
+    A2 & ... & ..  & .\\
+    A3 & ..  & .   & .\\ \hline
+  \end{tabular}
 \end{table}
-\clearpage
-}
 
-% Defining a command to generate a table of phenotypic data for three gwas
-\newcommand{\gwastablethree}[2]{
-\begin{table}
-	\caption[Phenotypic data used for the GWAS analyses of #2]%
-	{Phenotype frequency and numerical coding of phenotypic data used in three GWAS analyses of #2.
-	``NA'' indicates a missing value in the source data.
-	Dashes indicate that this phenotype was not used in the GWAS analysis.}
-	\label{#1_gwas_table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                               %%
+%% Additional Files              %%
+%%                               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-	\centering
-	\begin{threeparttable}\small
+\section*{Additional Files}
+  \subsection*{Additional file 1 --- Sample additional file title}
+    Additional file descriptions text (including details of how to
+    view the file, if it is in a non-standard format or the file extension).  This might
+    refer to a multi-page table or a figure.
 
-		\csvreader[head to column names,
-		/csv/separator=semicolon,
-		tabular = llccc,
-		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2 & GWAS \# 3\\ \midrule,
-		table foot = \bottomrule]%
-		{../tables/#1_gwas_table.csv}%
-		{}%
-		{\Value & \Frequency & \GWASa & \GWASb & \GWASc}
+  \subsection*{Additional file 2 --- Sample additional file title}
+    Additional file descriptions text.
 
-	\end{threeparttable}
-
-\end{table}
-\clearpage
-}
-
-% Defining a command for manhattan plot figures
-\newcommand{\manhattanplot}[3]{
-\begin{cfigure}
-	\includegraphics[width = 5.8in]{#1_manhattan}
-	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using three genotype datasets :
-	(a) Platypus (SNPs and indels), (b) Paragraph (SVs), (c) \emph{k}-mers presence/absence.
-	The x-axis shows the position along the reference assembly version 4 of Williams82.
-	Each point shows the $-\log_{10} (p)$ associated with a particular marker or \emph{k}-mer.
-	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
-	Vertical dotted lines indicate the position of signals associated with the trait.
-	Documented loci are colored according to whether they were found (green) or not (red) by a particular approach.
-	The ``Gm'' prefix has been left out of chromosome names for simplicity.
-	}
-	\label{#1_manhattan}
-\end{cfigure}
-
-\clearpage}
-
-% Defining a new command for signal plot figures
-\newcommand{\signalplot}[4]{
-\begin{cfigure}
-	\includegraphics[width = 5.8in]{#1_signal}
-	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using three genotype datasets:
-	(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
-	Panel (a) shows gene models over the genomic interval.
-	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
-	#4
-	Gray shaded rectangles indicate the region delimited by the top 5\% (SVs) or top 1\% (SNPs/indels and \emph{k}-mers) associations in the signal region.
-	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
-	\emph{p}-values were computed after signal discovery.
-	}
-	\label{#1_signal}
-\end{cfigure}
-
-\clearpage}
-
-% Defining a new command for gene plot figures
-\newcommand{\geneplot}[5]{
-\begin{cfigure}
-	\includegraphics[width = 5.8in]{#1_gene}
-	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
-	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using three genotype datasets:
-	(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
-	#5
-	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
-	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
-	\emph{p}-values were computed after signal discovery.
-	Panel (a) shows gene models over the plotting interval.
-	Exons are represented by rectangles whereas introns are represented by horizontal lines.
-	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
-	}
-	\label{#1_gene}
-\end{cfigure}
-
-\clearpage}
-
-% Defining a new command for k-mer plot figures
-\newcommand{\kmerplot}[5]{
-	\begin{cfigure}
-		\includegraphics[width = 5.8in]{#1_kmers}
-		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
-		{Identification of #5 underlying significant \emph{k}-mers at the #2 gene associated with the #3 locus for #4.
-		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
-		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
-		The red rectangle identifies the region that is zoomed-in in panel (b).
-		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
-		Individual nucleotides are colored according to the $-\log_{10}(p)$ of the most significant \emph{k}-mer overlapping them.
-		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
-		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes correspond to those shown in panel (b).}
-		\label{#1_kmers}
-	\end{cfigure}
-
-\clearpage}
-
-% Defining a new command for LD plot figures
-\newcommand{\ldplot}[3]{
-	\begin{cfigure}
-		\includegraphics[width = 5.8in]{#1_ld}
-		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
-		\emph{k}-mers are sorted along the y-axis according to their putative position along the reference assembly version 4 of Williams82,
-		as identified by ``Gm'' chromosome labels.
-		Sequences that lack a ``Gm'' label (top of the y-axis) represent unanchored scaffolds.
-		\emph{k}-mers are represented in the same order along the x- and y-axis.
-		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p)$ of each \emph{k}-mer.}
-		\label{#1_ld}
-	\end{cfigure}
-
-\clearpage}
-
-% END OF PREAMBLE
-
-\begin{document}
-
-\include{main_text}
-\include{additional_file_1}
-
+\end{backmatter}
 \end{document}

### 2023-03-28 additional_files/manuscript.tex
Add text for Genome Biology manuscript up to and including the Conclusions

@@
 %\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
-\documentclass{bmcart}[doublespacing, linenumbers]
+\documentclass[doublespacing]{bmcart}
 
@@
 %\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
+
+% Packages added by M-A Lemay
+\usepackage{csvsimple} % read and display csv files as tables
+\usepackage{booktabs} % prettier tables
+\usepackage{threeparttable} % easy addition of footnotes below tables
 
@@
 \author[
-  addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
+  addressref={aff1,aff2,aff3},                   % id's of addresses, e.g. {aff1,aff2}
   email={marc-andre.lemay.2@ulaval.ca}   % email address
@@
 \author[
-  addressref={aff1},
+  addressref={aff1,aff2,aff3},
+  email={maxime.de-ronne.1@ulaval.ca}
+]{\inits{M.D.R.}\fnm{Maxime} \snm{de Ronne}}
+
+\author[
+  addressref={aff1,aff2,aff3},
   email={richard.belanger@fsaa.ulaval.ca}
@@
 
-\address[id=aff1]{%                           % unique id
-  \orgdiv{Département de phytologie},             % department, if any
-  \orgname{Université Laval},          % university, etc
-  \city{Québec},                              % city
-  \cny{QC, Canada}                                    % country
+\author[
+  addressref={aff1,aff2,aff3},
+  email={francois.belzile@fsaa.ulaval.ca}
+]{\inits{F.B.}\fnm{François} \snm{Belzile}}
+
+\address[id=aff1]{%
+  \orgdiv{Département de phytologie},
+  \orgname{Université Laval},
+  \city{Québec},
+  \cny{QC, Canada}
+}
+
+\address[id=aff2]{%
+  \orgdiv{Institut de biologie intégrative et des systèmes},
+  \orgname{Université Laval},
+  \city{Québec},
+  \cny{QC, Canada}
+}
+
+\address[id=aff3]{%
+  \orgdiv{Centre de recherche et d’innovation sur les végétaux},
+  \orgname{Université Laval},
+  \city{Québec},
+  \cny{QC, Canada}
 }
@@
 \section*{Results and discussion}
-Text for this section\ldots
-\subsection*{Sub-heading for section}
-Text for this sub-heading\ldots
-\subsubsection*{Sub-sub heading for section}
-Text for this sub-sub-heading\ldots
-\paragraph*{Sub-sub-sub heading for section}
-Text for this sub-sub-sub-heading\ldots
+
+We analyzed a total of 13 traits in a population of 363 resequenced \textit{G.
+max} accessions using three GWAS approaches that differed in the genotypes
+used. The first approach used SNP and indel genotypes called by Platypus.
+Despite both SNPs and indels being used, we will refer to this approach as
+SNP-based for simplicity. Our SV-based workflow used various sources of data
+for SV discovery and a single program (Paragraph) for genotyping SVs at
+population scale from Illumina WGS data. Finally, the third GWAS approach
+relied on the presence/absence of \textit{k}-mers in the WGS data as genotypes.
+Because of the large number of \textit{k}-mers analyzed, we only report results
+for those that were considered significantly associated with traits of
+interest.
+
+Since two or three different GWAS analyses were conducted for some of the
+traits, a total of 22 GWAS analyses were performed using each of the three
+approaches.  Given the large volume of data that this represents, we focus here
+on the most noteworthy results. Readers interested in results at loci that are
+not discussed here are referred to Additional file 1. In particular, Table
+S1 (Additional file 1) provides a summary of the
+\textit{p}-values obtained by each approach at each of the loci considered in
+this study.
+
+\subsection*{Loci with known genes or causal variants}
+
+We first wanted to assess the performance of the various GWAS
+approaches by analyzing the results at loci for which the genes have
+already been cloned. These loci provide interesting test cases because
+the expected results (causal genes or variants) are known. Table \ref{loci-table} shows
+a summary of the results discussed below.
+
+\subsubsection*{Flower color -- \textit{W1} locus}
+
+\cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
+Glyma.13g\-072100 as associated with the \textit{W1} locus for flower color. In
+Williams82 and other accessions with white flowers, a 65-bp insertion into the
+third exon of the gene results in a premature stop codon that renders the
+F3’5’H enzyme non-functional and prevents pigmentation of the flower. All
+GWAS approaches detected a signal overlapping the gene at this locus (Additional file 1: Figures
+S1 and S2). The most
+significant SV and \textit{k}-mers corresponded to the known causal variant
+(Figure \ref{flower-color-main-figure}a-b and Additional file 1: S3c-d).
+In particular, the assembly of the reads containing the most significant \textit{k}-mers
+and the alignment of the resulting haplotypes showed that these \textit{k}-mers
+directly tagged the causal variant (Figure \ref{flower-color-main-figure}c-e).
+As the catalogue of variants analyzed by Platypus contains only SNPs and small
+indels, the causal variant could not be captured by this approach. Still, SNPs
+and indels located within the gene sequence were among the most significantly
+associated ones (Additional file 1: Figure S3b).
+
+The \textit{k}-mer GWAS analysis also identified presumably spurious significant
+\textit{k}-mers scattered throughout the genome (Additional file 1: Figure
+S1c). An analysis of the pairwise LD between a subset
+of those \textit{k}-mers showed that most \textit{k}-mers formed a single LD block
+linked to the \textit{W1} locus (Additional file 1: Figure S4). A small LD block
+occurred on chromosome Gm09, however it was in moderate LD with the one on chromosome
+Gm13 and therefore probably does not represent a separate locus.
+
+We should note that when we first analyzed the results of the GWAS on flower
+color, we observed that two accessions bearing the haplotype typically
+associated with purple flowers were noted as having white flowers. Upon further
+inspection, we found that the most likely cause of this discrepancy was a
+mismatch between the genotype data and the identity of the accession for which
+the phenotypic data was obtained. Indeed, when we compared SNP calls made
+directly from the WGS data to those made using the SoySNP50K array, we observed
+that these two accessions had low concordance rates of only 62\% and 70\%. To
+ensure that such mismatches would not affect our GWAS analyses, we computed the
+concordance between genotype calls made from WGS data and the SoySNP50K array
+for all samples (see the Methods) and excluded 24 accessions for which the
+concordance rate was below 90\%.
+
+\subsubsection*{Pubescence color -- \textit{T} locus}
+
+\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
+\textit{T} locus for pubescence color and found a 1-bp deletion as a putative
+causal variant at this locus. For all three GWAS approaches, we found a signal
+overlapping this gene (Additional file 1: Figure S5). In
+particular, the regions defined by the 1\% most associated variants
+identified by the SNP- and \textit{k}-mer-based approaches included
+Glyma.06g202300 (Additional file 1: Figure S6). While the most
+significant \textit{k}-mers did occur within the sequence of this gene, they
+did not correspond to the documented causal variant (Additional file 1: Figure
+S7d).  Still, the fourth and fifth most strongly
+associated \textit{k}-mers corresponded to this causal variant (Additional file 1: Figures
+S7d and S8). We
+observed that not all accessions with gray pubescence harbored the deleted
+nucleotide (Additional file 1: Figure S8c), which could indicate
+either that other causal variants are resulting in a non-functional F3’H in
+this population, or that the causal variant lies elsewhere. As for SNP-based GWAS,
+the most significant variant within the sequence of Glyma.06g202300 was the
+causal variant, however it was not the most significantly associated variant
+overall (Additional file 1: Figures S6b and
+S7b). SV-based GWAS did not detect the causal variant
+at this locus, as expected from its size.
+
+\subsubsection*{Pubescence color -- \textit{Td} locus}
+
+% Pubescence color - Td locus
+We conducted a separate GWAS analysis contrasting only accessions with tawny
+and light tawny pubescence because the \textit{Td} locus is known to control
+these differences in color. Accordingly, we detected signals in this region
+using all three approaches, but only SNP- and \textit{k}-mer-based GWAS
+identified a signal that overlapped Glyma.03g258700 (Figure
+\ref{pubescence-color-main-figure}a-b, and Additional file 1: Figures S9 and
+S10).  While the most significant
+\textit{k}-mer identified the causal variant documented at this locus
+\citep{yan2020}, it was the fourth most significantly associated variant for
+the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d, Additional file 1: Figure 
+S11).  Interestingly, some of the accessions
+bearing the haplotype associated with tawny pubescence displayed light tawny
+pubescence, suggesting that there may be more than one causal variant at this
+locus (Figure \ref{pubescence-color-main-figure}e).
+
+\subsubsection*{Seed coat color - \textit{I} locus}
+
+% Seed coat color - I locus
+The causal variant at the \textit{I} locus for seed coat color is a complex SV
+that takes the form of a 10.91-kb inverted duplication of three chalcone
+synthase (CHS) genes \citep[CHS1, CHS3 and CHS4;][]{tuteja2008}. The presence of
+this inverted duplication results in the silencing of all CHS genes in the seed
+coat specifically and thus in the absence of seed coat pigmentation
+\citep{tuteja2009}. While we did find signals overlapping this causal variant
+using all three GWAS approaches (Additional file 1: Figures S12 and
+S13), none of the approaches identified
+it. This failure to detect the causal variant is likely due
+to its complexity, which makes it difficult to detect and genotype it. While it
+is likely that at least some of the significantly associated \textit{k}-mers
+actually are derived from this causal variant, the lack of systematic methods for
+linking \textit{k}-mers to sequence variation hindered our ability to do so.
+
+\subsubsection*{Seed coat color - \textit{G} locus}
+
+We conducted a second GWAS analysis contrasting accessions with yellow and
+green seed coats in order to target the \textit{G} locus. Accordingly, this
+analysis found strong signals overlapping the \emph{G} locus using all three
+approaches (Additional file 1: Figure S14).
+\cite{wang2018} identified Glyma.01g198500 as the gene associated with
+this locus and found an A>G SNP affecting transcript splicing as the causal
+variant. Consistently with their results, we identified the two
+most significant \emph{k}-mers and the second most significant marker from the
+SNP-based analysis as corresponding to that causal variant (Figure
+\ref{seed-coat-color-main-figure} and Additional file 1: Figures
+S15 and
+S16). Unsurprisingly, SV-based GWAS
+did not identify the causal variant given its type.
+
+\subsubsection*{Stem termination type - \textit{Dt1} locus}
+
+While all GWAS approaches found signals near the Glyma.19g194300 gene
+associated with the \textit{Dt1} locus for stem termination type, the SNP- and
+\textit{k}-mer-based analyses performed best since the region defined by their
+most highly associated markers (top 1\%) included the causal gene (Additional file 1: Figures
+S17 and S18).
+We did identify one of the SNPs suggested as putative causal variants for this
+locus by \cite{tian2010} and \cite{liu2010}, however this SNP was only the
+135\textsuperscript{th} most associated marker found by Platypus and the
+12\textsuperscript{th} most associated \emph{k}-mer (Additional file 1: Figures
+S19 and S20).
+Our results are consistent with previous findings in suggesting that there is
+more than one causal variant at the \textit{Dt1} locus \citep{liu2010,
+tian2010}. Indeed, a few cultivars bearing the allele associated with
+\emph{dt1} showed indeterminate or semi-determinate phenotypes and some
+cultivars showed a determinate phenotype without bearing the allele associated
+with \emph{dt1} (Additional file 1: Figure S20c). The existence of
+several causal variants might explain why our analyses did not identify a
+causal variant more clearly.
+
+\subsubsection*{Hilum color -- \textit{R} locus}
+
+We conducted a GWAS analysis that only compared accessions with black and brown
+hila in an attempt to detect stronger signals at the \textit{R} locus.  We
+accordingly detected a signal overlapping this locus using all approaches, but
+only SNP- and \textit{k}-mer-based GWAS detected signals overlapping the
+Glyma.09g235100 gene associated with \textit{R} (Additional file 1: Figures
+S21 and S22).
+\cite{gillman2011} identified this R2R3 MYB transcription factor as the
+molecular basis for the \textit{R} locus and documented four different
+loss-of-function alleles linked to brown hilum color. Although our analyses did
+detect strong signals in the vicinity of this gene, the most associated markers
+were not located in the body of the gene (Additional file 1: Figure
+S22).  Interestingly, the most significantly
+associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
+another gene (Glyma.09g234900) annotated as a MYB transcription factor and
+putatively involved in the regulation of anthocyanin biosynthesis. This gene as
+well as Glyma.09g235100 and two others are part of a tandem array of four R2R3
+MYB genes considered as candidates by \cite{gillman2011}, and of which
+Glyma.09g235100 was the only one expressed in the seed coat. Given the
+convincing evidence provided by \cite{gillman2011}, it is most likely that
+Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
+our failure to identify the most significantly associated \emph{k}-mers as
+causal variants is due to the coexistence of several variants that can cause
+loss of function at the \emph{R} locus.  Still, we observed two causal variants
+previously documented by \cite{gillman2011} in our dataset (Additional file 1: Figures
+S23 and S24).
+
+\subsubsection*{Pubescence density -- \textit{Ps} locus}
+
+We conducted a GWAS analysis comparing accessions with normal and semi-sparse
+pubes\-cence in order to detect loci associated with pubescence density. This analysis
+detected signals at the \textit{Ps} locus using all three approaches (Figure
+\ref{pubescence-density-main-figure}a-b and Additional file 1: Figures S25, S26
+and S27). The causal
+variant at this recently cloned locus is a copy number variant (CNV) overlapping the
+Glyma.12g187200 gene, with higher copy number resulting in reduced pubescence \citep{liu2020ps}.
+We identified a region of $\sim$42 kb that varied in copy number across
+the population and was associated with the trait (Figure
+\ref{pubescence-density-main-figure}c), and therefore possibly corresponded to the
+25.6-kb region identified as the causal CNV by \cite{liu2020ps}. We found that
+reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
+of the CNV (Figure \ref{pubescence-density-main-figure}d). This suggests that
+the most associated \textit{k}-mer is indeed linked to the CNV
+that is causal at this locus, while \textit{k}-mers within the
+CNV region may represent differences in sequence between the copies.
+
+The difference in size between the CNV reported by \cite{liu2020ps} and our
+results may be due to the presence of some unresolved sequence (a stretch of
+4,635 N nucleotides) over that interval in the reference sequence that we used.
+Because our SV filtering process removed any alternate allele containing N nucleotides,
+we reasoned that the CNV may have been found by our pipeline but
+filtered out later on. Indeed, we found a 42.8-kb duplication corresponding
+to the one described above among the genotype calls made by smoove. An association
+analysis of that single variant using GAPIT found a \textit{p}-value of 1.9e-23,
+which would have made it the most significant SV. In this particular case, the
+failure of the SV-based GWAS analysis to find identify the causal variant was due
+to our filtering parameters and not to limitations of the SV genotype calls per se.
+
+\subsubsection*{Seed coat luster -- \textit{B} locus}
+
+We conducted a GWAS analysis comparing dull and shiny seed coat phenotypes for
+the seed coat luster trait. Seed coat luster has been shown by
+\cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
+protein (HPS) at the seed surface. Copy number variation of this gene at the
+\textit{B} locus was later shown to explain variation in luster, although
+sequence variation in the HPS sequence may also play a role \citep{gijzen2006}.
+Using a sequencing depth-based analysis, we identified a 31-kb region spanning
+positions 9,386,109-9,417,431 on chromosome Gm15 that exhibited variation in
+copy number across accessions. This CNV overlapped four genes annotated as
+containing HPS domains and therefore likely represented the causal variant
+(Additional file 1: Figures S28 and
+S29).  We found signals overlapping this
+CNV using all approaches, and the most significantly associated \textit{k}-mer
+was notably located within the CNV region (Additional file 1: Figures
+S28 and
+S30). However, we were not able to
+directly link any of the variants or \textit{k}-mers identified to the causal
+CNV.
+
+\subsection*{Analysis of loci with unknown causal genes}
+
+In addition to analyzing loci for which the genes are already known, we were
+able to suggest candidate genes at loci for which the underlying genes are not
+known yet.
+
+\subsubsection*{Pod color -- \textit{L1} and \textit{L2} loci}
+
+Both the \textit{L1} and the \textit{L2} loci for pod color have yet to be
+cloned. While \cite{he2015} proposed the gene Glyma.19g101700 as likely
+corresponding to \textit{L1} based on fine mapping and expression analysis, our
+results as well as those reported by \cite{bandillo2017} do not support this
+gene as a candidate due to its location $\sim$3 Mb away from GWAS signals. Our
+analyses based on \textit{k}-mers and SNPs found signals overlapping 14 and
+6 genes, respectively (Additional file 1: S31 and
+S32). Of these, a MATE transporter (Glyma.19g120300)
+appears as a prime candidate as the most significant \emph{k}-mer was observed
+within its sequence and it is highly expressed in pods according to data hosted
+on SoyBase \citep{grant2010, severin2010}. Moreover, the role of some MATE
+proteins in transporting flavonoids has been demonstrated \citep{chen2015}.
+
+The signals found by SNP- and \emph{k}-mer-based analyses at the \textit{L2}
+locus overlap 24 and 50 genes, respectively (Additional file 1: Figures
+S31 and S33). Despite this
+somewhat high number, we noticed that the closest gene to the most significant
+\emph{k}-mer (Glyma.03g005800) is also a MATE transporter, similarly to our top
+candidate for \emph{L1}. This gene may therefore represent an interesting
+candidate at this locus for future studies.
+
+\subsubsection*{Pubescence form -- \textit{Pa1} and \textit{Pa2} loci}
+\label{sv-gwas-main-results-pubescence-form-pa1-pa2}
+
+The two known loci for pubescence form in soybean, \textit{Pa1} and
+\textit{Pa2}, have also not been cloned yet. Nevertheless, \cite{gilbert2017}
+has suggested Glyma.12g213900 as a candidate for \textit{Pa1} based on data
+published by \cite{bandillo2017}. This gene encodes a MYB transcription factor
+annotated as playing a role in trichome branching. Indeed, the most significant
+signals found by all three approaches were located near this gene (Figure
+\ref{pubescence-form-main-figure}a-b, and Additional file 1: Figures S34 and
+S35).  The analyses based on SNPs and
+\emph{k}-mers identified two highly significant non-synonymous SNPs in
+Glyma.12g213900 that were associated with appressed pubescence (Figure
+\ref{pubescence-form-main-figure}c-d and Additional file 1: Figures S36 and
+S37). The alternative alleles for these two
+SNPs co-occurred together in our dataset and may represent causal variants at
+the \emph{Pa1} locus. The signals observed at the \emph{Pa2} locus were much
+weaker and the signal observed from the \emph{k}-mer GWAS overlapped a total of
+20 genes (Additional file 1: Figures S34 and
+S38), none of which appeared obviously linked
+to pubescence form.
+
+A second GWAS analysis focused only on appressed and semi-appressed phenotypes
+with the objective of enhancing the signal at the \emph{Pa2} locus. However,
+this analysis only detected weak signals at the \emph{Pa1} locus (Additional file 1: Figure
+S39) and its results were therefore not
+analyzed further.
+
+\subsubsection*{Resistance to \textit{P. sojae}}
+
+We used corrected dry weight (CDW) as described by \cite{deronne2022} as a
+measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
+overlapping the genomic region reported by these authors on Gm15 using all three GWAS
+approaches (Additional file 1: Figure S40). \cite{deronne2022}
+proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
+variation at that locus based on a combination of functional annotation and
+contrasting gene expression in resistant and susceptible lines four days after
+inoculation. Although the signals found by our analyses did overlap
+Glyma.15g217100, we found that this gene was located at the distal end of the signal
+in a region with markers or \emph{k}-mers showing a weaker
+association (Additional file 1: Figure S41). The
+large size of the observed signal means that the gene associated with the phenotype could
+be located almost anywhere in this region. However, we identified a particularly
+interesting region defined by the top 1\% associations in the SNP-based
+analysis. This region overlaps only three annotated genes and also contains the
+most significant \emph{k}-mer (Additional file 1: Figure
+S41). One of the genes in this region is
+a calcium ion binding protein with an EF-hand domain (Glyma.15g217700).
+Given the important role played by calcium signaling in defense
+responses \citep{zhang2014}, we suggest Glyma.15g217700 as an interesting
+candidate gene for explaining resistance to \emph{P. sojae} in this population.
+
+\subsubsection*{Novel loci for stem termination and pod color}
+
+In addition to verifying known loci, we also observed potentially novel loci in
+the \textit{k}-mer GWAS analysis of stem termination and pod color.  When
+comparing accessions with indeterminate and semi-determinate stem termination
+types, we observed previously undocumented signals on chromosomes Gm11, Gm16
+and Gm18 in addition to the known \textit{Dt1} locus (Figure
+\ref{stem-termination-main-figure}a).  Analysis of the pairwise LD between
+significant \textit{k}-mers found no LD between these signals, suggesting that
+they may represent \textit{bona fide} loci associated with stem termination
+type (Figure \ref{stem-termination-main-figure}b). Similarly, a \textit{k}-mer
+GWAS analysis comparing accessions with black and brown pods found a previously
+undocumented signal on chromosome Gm15 (Additional file 1: Figure S42).
+\textit{k}-mers at this new locus
+showed no LD with \textit{k}-mers at the known \textit{L1} locus and may
+therefore represent a distinct locus (Additional file 1: Figure 
+S43). The observed signal overlaps a total of 27 genes
+(Additional file 1: Figure S44), but none of those are annotated
+as having an obvious role in controlling pod color.
+
+\subsection*{Potential of \emph{k}-mer-based GWAS}
+
+One of the most striking results from the analyses presented above is the
+impressive performance of \emph{k}-mer-based GWAS compared to SNP- and SV-based
+GWAS. The GWAS analyses using \emph{k}-mers identified the causal variant as
+the most significantly associated variant at four loci with various types of
+causal variants: \emph{W1} for flower color, \emph{Td} for pubescence color,
+\emph{G} for seed coat color, and \emph{Ps} for pubescence density (Table
+\ref{loci-table}). In comparison, this result was achieved by SV-based GWAS at
+the \textit{W1} locus only, and at none of the loci for SNP-based GWAS.
+The \textit{k}-mer approach also systematically performed at least as well as
+the other two approaches in identifying signals overlapping known variants or
+genes (Table \ref{loci-table}). Moreover, whenever \textit{k}-mer-based GWAS
+did detect a signal, the region defined by the 1\% most significant
+\textit{k}-mers overlapped the known causal variant or gene.
+
+Even in cases where the most significantly associated \emph{k}-mers did not
+pinpoint known causal variants, they were often located near those variants.
+For example, the most significantly associated \emph{k}-mers mapping to the
+\emph{T} locus for pubescence color were located in intronic sequences of the
+gene known to be associated with this locus.  Similarly, the most significantly
+associated \emph{k}-mers at the \emph{B} locus for seed coat luster mapped
+within a genomic interval known to vary in copy number at that locus. In
+addition, the most significant \emph{k}-mers at the \emph{R} locus mapped just
+upstream of a homolog of the gene associated with the locus. Even though the
+\emph{k}-mer analysis did not identify the causal variants directly in those
+cases, it still provided a very strong starting point for candidate gene
+identification.
+
+As would be expected of any GWAS analysis, the success of the \emph{k}-mer approach
+depends on the quality of the phenotypic data and the complexity of the genetic
+architecture of the trait. It is clear that the loci for which the
+\emph{k}-mer approach was most successful (\emph{W1}, \emph{Td}, \emph{G} and
+\emph{Ps}) have a simple genetic architecture and phenotypes that can be easily
+determined. For other traits, challenges in determining accurate phenotypes may
+result in a decreased ability to identify causal variants or genes. For
+example, \cite{bandillo2017} noted that observed phenotypes for stem
+termination type in the GRIN database may not correspond to the expected
+genotypes at classical loci, and that the intermediate phenotype for seed coat
+luster is not always consistent across different environments. For qualitative
+traits, the numerical scale used to recode phenotypes for use in GWAS is somewhat
+arbitrary and may influence the final results. In the case of quantitative
+traits, the usual limitations to the precision and stochastic variation of
+phenotypic values will necessarily make the interpretation more difficult than
+for simple qualitative traits. Notwithstanding those limitations,
+\textit{k}-mer-based GWAS clearly performed better than any of the other
+approaches tested here given the same phenotypic data. This is partly due
+to the fact that the \textit{k}-mer approach encompasses all variant types.
+Moreover, it is likely that the simpler workflow that results from using
+\textit{k}-mer presence/absence avoids some of the errors introduced during
+variant calling and genotyping, and therefore yields more accurate results.
+
+\subsection*{Limitations and challenges of \emph{k}-mer-based GWAS}
+
+While we have have shown that using \emph{k}-mers for GWAS in soybean performed
+well compared to other GWAS methods, this approach does present some unique
+challenges and limitations. One obvious limitation is that using the
+presence/absence of \emph{k}-mers may not be appropriate in species where
+heterozygosity is common (i.e., most species other than inbred crops) as the
+\emph{k}-mers associated with the presence of both alleles will be observed at
+heterozygous loci. Moreover, although the presence/absence of \emph{k}-mers can
+detect variation independently of its type, some types of variation may be more
+difficult to detect than others.  For example, copy number variation at a locus
+may not be detected from presence/absence alone if all copies are identical and
+occur in tandem, such that no \textit{k}-mer can differentiate between two
+copies and more. For similar reasons, \emph{k}-mers may also fail to capture
+variation occurring in repeated sequences. Alternative models based on
+\emph{k}-mer counts instead of presence/absence have been developed
+\citep{rahman2018, he2021} which may solve issues like these and broaden the
+scope of variants that can be identified from \emph{k}-mers.
+
+The complexity of analyzing the results of \emph{k}-mer-based GWAS is another
+limitation of this approach. Although the method developed by
+\cite{voichek2020} is in itself relatively simple and computationally
+inexpensive, downstream analysis and interpretation of the output is
+challenging. Indeed, there are no state-of-the-art methods for identifying the
+putative genomic coordinates of a \emph{k}-mer, grouping significant
+\emph{k}-mers together into coherent signals, and linking them to biologically
+meaningful sequence variation. Going forward, mapping the \emph{k}-mers or
+reads to a graph-based genome \citep[e.g. using the vg toolkit;][]{siren2021}
+may represent an improvement to using a single linear genome.  Grouping
+significant \emph{k}-mers by LD instead of by genomic position, as was done for
+some of the analyses in this paper, could prove useful in determining how many
+loci control a trait. In this study, we used an assembly-based approach similar
+to what others have previously done \citep[e.g.][]{voichek2020, rahman2018}.
+This approach allowed us to link \emph{k}-mers to their underlying variation
+for simple cases involving SNPs and indels, but we were unable to link
+significantly associated \emph{k}-mers to more challenging causal variants such
+as those explaining the \emph{I} or \emph{B} loci. In this context, developing
+approaches to systematically link significant \emph{k}-mers to the sequence
+variation that underlies their presence is much needed. Furthermore, finding
+ways to limit the number of spurious associations found for some traits, as was
+the case for flower color and pubescence color in our study, is also needed in
+order to obtain robust results.
+
+In this work, we provide a set of tools enabling the downstream analysis of
+\emph{k}-mer-based GWAS. Most importantly, we developed the katcher program for
+retrieving all reads containing any of a set of significant \emph{k}-mers in a
+computationally efficient manner. Previous studies usually mapped \emph{k}-mers
+to a reference genome to identify their reference-based coordinates
+\citep[e.g.][]{voichek2020, he2021, rahman2018, tripodi2021}.  However, while
+this approach may work for \emph{k}-mers that show little divergence from the
+reference, it might fail to identify the genomic location of \emph{k}-mers that
+diverge significantly or are not found at all in the reference. Moreover, it
+does not take into account the sequence context of the read where the
+\emph{k}-mer was found, which may provide valuable information for mapping
+purposes. Our approach avoids these pitfalls by instead systematically
+retrieving reads containing significant \emph{k}-mers, including those that did
+not map at all to the reference assembly. While we were able to position the
+most significant \emph{k}-mers on the reference sequence in our analyses, it
+may not be the case for highly variable species or if the causal variation at a
+locus results from a large novel insertion.
+
+\subsection*{Use of SV genotypes for GWAS}
+
+In addition to \emph{k}-mers, we also assessed the use of SVs in GWAS analysis
+of soybean.  Although we did find the causal variant as the most significantly
+associated SV at the \emph{W1} locus, results were otherwise rather
+inconclusive.  Indeed, most of the other known or putative causal variants were
+SNPs or indels, and therefore were not represented in the SV datasets. At most
+loci where SVs are known to be the underlying causal variant (e.g.  \emph{I},
+\emph{B}, \emph{Ps}), the analyses based on SVs did not find the causal
+variants, most likely due to the difficulty of calling and genotyping such
+variants \citep{kosugi2019, lemay2022}. In the case of the \textit{Ps} locus,
+this limitation was due to a filtering step in our pipeline rather than to an
+inherent inability to properly call and genotype the causal variant.
+
+Given the current performance of SV genotyping programs, it is not clear
+whether using SVs for GWAS provides any advantage over using \emph{k}-mer- or
+SNP-based GWAS, or a combination of both.  Indeed, in the single case where the
+SV-based analyses identified the causal variant, the \emph{k}-mer-based GWAS
+identified it as well. Moreover, SNP-based GWAS provided better guidance
+in delimiting the signals found at a given locus because of the much higher
+density of markers compared to SVs.  The usefulness of SV-based GWAS compared
+to SNP-based GWAS will largely depend on whether SVs can represent variation
+that is not otherwise in LD with SNPs. In tomato, \cite{dominguez2020} found
+that most transposable element insertions associated with phenotypic variation
+in agronomic traits were not tagged by SNPs. On the other hand, in humans,
+\cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
+is therefore likely that how well SVs can be tagged by SNPs will depend on the
+particular type of SV and on the species. A more efficient approach than
+SV-based GWAS may be to conduct \emph{k}-mer-based GWAS and only then link
+known SVs to significant \emph{k}-mers.  Alternatively, \textit{k}-mer-based
+GWAS can be conducted specifically for \textit{k}-mers that are associated with
+SVs \citep{jayakodi2020}.  Previous studies have obtained noteworthy results by
+using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
+dominguez2020}, however it would be worthwhile to see whether these results
+could be replicated using \emph{k}-mer-based GWAS.
 
 \section*{Conclusions}
+
+In conclusion, we used SV-based and \emph{k}-mer-based GWAS to study ten
+qualitative traits and three quantitative traits in soybean and
+compared the results to conventional SNP-based GWAS. While \emph{k}-mer-based
+GWAS proved to be a powerful approach in pinpointing the causal variation or
+genes associated with known loci, it is unclear whether conducting SV-based
+GWAS is worth the investment. In addition, we used the results from the
+\emph{k}-mer- and SNP-based GWAS to suggest candidate genes for a few classical
+loci that have yet to be cloned. Based on our results, we believe that an
+optimal workflow may involve conducting SNP-based and \emph{k}-mer-based GWAS
+in parallel to identify significant signals and candidate genes, potentially in
+combination with SV datasets. Given the large number of samples for which WGS
+data is now available in several species, applying \emph{k}-mer-based GWAS to
+leverage already existing sequencing and phenotypic data appears feasible and
+promising.  As part of our work, we developed several computational tools that
+should help other researchers with the downstream analysis of
+\emph{k}-mer-based GWAS. However, much work remains to be done in
+developing state-of-the-art methods for the downstream analysis of significant
+\emph{k}-mers. In particular, better approaches are needed for linking
+\emph{k}-mers to sequence variation and moving from presence/absence-based
+methods to count-based methods.
 
@@
 \section*{Figures}
-  \begin{figure}[h!]
-  \caption{Sample figure title}
+
+% Figure 1
+\begin{figure}
+	\caption{Results of SV- and \textit{k}-mer-based GWAS at the
+	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
+	threshold whereas vertical dotted lines indicate the location of the
+	\textit{W1} locus. (b) Zoomed-in Manhattan plots at the \textit{W1} locus.
+	The vertical dotted lines indicate the location of the
+	Glyma.13g072100 gene associated with this locus. The shaded gray
+	rectangles show the regions delimited by the top 5\% (SVs) and
+	1\% (\textit{k}-mers) associations.  (c) Gene model of Glyma.13g072100.
+	Rectangles represent exons, colored rectangles represent coding
+	sequences, and arrows indicate the direction of transcription. The red
+	rectangle highlights the region shown in panel (d). (d) Pairwise
+	alignment of the two haplotype sequences in the population at the
+	\textit{W1} locus. Nucleotides are
+	colored according to the \textit{p}-value of the most significant
+	\textit{k}-mer overlapping them. Vertical dotted lines between haplotypes
+	indicate sequence differences, whereas dashes indicate gaps in the
+	alignment. (e) Contingency table of the phenotypes and haplotypes found
+	in the population at the \textit{W1} locus.}
+	\label{flower-color-main-figure}
 \end{figure}
 
-\begin{figure}[h!]
-  \caption{Sample figure title}
+% Figure 2
+\begin{figure}
+	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
+	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
+	threshold whereas vertical dotted lines indicate the location of the
+	\textit{Td} locus. (b) Zoomed-in Manhattan plots at the \textit{Td} locus.
+	The vertical dotted lines indicate the location of the
+	Glyma.03g258700 gene associated with this locus. The shaded gray
+	rectangles show the regions delimited by the top 1\% associations.
+	In the case of SNPs, blue points denote markers used in the
+	original analysis, whereas red points denote markers that had originally
+	been pruned but whose \emph{p}-values were computed after signal
+	discovery.
+	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
+	rectangles represent coding sequences, and the arrow indicates the
+	direction of transcription. The red rectangle highlights the region
+	shown in panel (d). (d) Pairwise alignment of the two haplotype
+	sequences in the population at the \textit{Td} locus. Nucleotides are
+	colored according to the
+	\textit{p}-value of the most significant \textit{k}-mer overlapping
+	them. The vertical dotted line between haplotypes indicates the location of
+	the causal SNP at this locus. (e) Contingency table of
+	the phenotypes and haplotypes found in the population at the
+	\textit{Td} locus.}
+	\label{pubescence-color-main-figure}
+\end{figure}
+
+% Figure 3 with the results on pubescence color at locus G
+\begin{figure}
+	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
+	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
+	threshold whereas vertical dotted lines indicate the location of the
+	\textit{G} locus. (b) Zoomed-in Manhattan plots at the \textit{G} locus.
+	The vertical dotted lines indicate the location of the
+	Glyma.01g198500 gene associated with this locus. The shaded gray
+	rectangles show the regions delimited by the top 1\% associations.
+	In the case of SNPs, blue points denote markers used in the
+	original analysis, whereas red points denote markers that had originally
+	been pruned but whose \emph{p}-values were computed after signal
+	discovery.
+	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
+	rectangles represent coding sequences, and the arrows indicate the
+	direction of transcription. The red rectangle highlights the region
+	shown in panel (d). (d) Pairwise alignment of the two haplotype
+	sequences in the population at the \textit{G} locus. Nucleotides are
+	colored according to the
+	\textit{p}-value of the most significant \textit{k}-mer overlapping
+	them. The vertical dotted line between haplotypes indicates the location of
+	the causal SNP at this locus. (e) Contingency table of
+	the phenotypes and haplotypes found in the population at the
+	\textit{G} locus.}
+	\label{seed-coat-color-main-figure}
+\end{figure}
+
+% Figure 4 with the results on pubescence density at locus Ps
+\begin{figure}
+	\caption[Results of \textit{k}-mer-based GWAS at the
+	\textit{Ps} locus for pubescence density.]{Results of \textit{k}-mer-based GWAS at the
+	\textit{Ps} locus for pubescence density.
+	(a) Genome-wide Manhattan plot showing the location of significant \textit{k}-mers. The
+	horizontal dotted line indicates the significance threshold whereas the
+	vertical dotted line indicates the location of the \textit{Ps} locus.
+	(b) Zoomed-in Manhattan plot at the \textit{Ps} locus. The
+	vertical dotted line indicates the location of the most significant
+	\textit{k}-mer. The shaded gray area shows the location of the causal CNV.
+	(c) Estimated copy number (sequencing depth at position divided by
+	accession average sequencing depth) of accessions with normal and
+	semi-sparse pubescence at the \textit{Ps} locus. Solid lines represent
+	the median copy number across 73 (semi-sparse pubescence) and 180
+	accessions (normal pubescence) computed using a 100-bp sliding window.
+	Shaded areas outline the first and third quartiles. The start and end
+	of the CNV region can be clearly inferred from the divergence between
+	the copy number of accessions with contrasting pubescence.
+	(d) Histogram showing the location of 1,169 mapped reads containing the
+	most significant \textit{k}-mer in accessions with semi-sparse pubescence.
+	Reads containing this \textit{k}-mer map to either end of the CNV
+	region, suggesting that they are associated with this variant. Panels
+	(b), (c) and (d) share the same x-axis.}
+	\label{pubescence-density-main-figure}
+\end{figure}
+
+% Figure 5 with the results on pubescence form at locus Pa1
+\begin{figure}
+	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
+	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
+	threshold whereas vertical dotted lines indicate the location of the
+	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots at the \textit{Pa1} locus.
+	The vertical dotted lines indicate the location of the
+	Glyma.12g213900 gene suggested as a candidate for this locus. The shaded gray
+	rectangles show the regions delimited by the top 1\% associations.
+	In the case of SNPs, blue points denote markers used in the
+	original analysis, whereas red points denote markers that had originally
+	been pruned but whose \emph{p}-values were computed after signal
+	discovery.
+	(c) Gene model of Glyma.12g213900. Rectangles represent exons, colored
+	rectangles represent coding sequences, and the arrows indicate the
+	direction of transcription.
+	(d) Manhattan plots at the location of the Glyma.12g213900 gene.
+	Vertical lines indicate the location of two non-synonymous
+	SNPs associated with appressed pubescence, which may represent causal variants
+	at this locus. Panels (c) and (d) share the same x-axis.}
+	\label{pubescence-form-main-figure}
+\end{figure}
+
+% Figure 6 with the results novel loci for stem termination type
+\begin{figure}
+	\caption[Detection of novel signals for stem termination type
+	using a \textit{k}-mer GWAS analysis contrasting accessions with
+	indeterminate and semi-determinate stems]{Detection of novel signals for stem termination type
+	using a \textit{k}-mer-based GWAS analysis contrasting accessions with
+	indeterminate and semi-determinate stems.
+	(a) Genome-wide Manhattan plot showing the location of significant \textit{k}-mers.
+	The horizontal dotted line indicates the significance threshold
+	whereas vertical dotted lines indicate the location of the \textit{Dt1}
+	classical locus as well as three new loci of interest (\textit{stGm11},
+	\textit{stGm16}, and \textit{stGm18}).
+	(b) Pairwise LD between the 162 significant \textit{k}-mers identified.
+	\textit{k}-mers are sorted along the axes according to their
+	putative position on the reference assembly version 4 of Williams82.
+	\textit{k}-mers are
+	represented in the same order along the x- and y-axis. The blue
+	rectangles below the x-axis represent the
+	-log\textsubscript{10}(\textit{p}) of each \textit{k}-mer.}
+	\label{stem-termination-main-figure}
 \end{figure}
@@
 \section*{Tables}
+
+% Table 1 with results on all loci with cloned genes
+%\begin{landscape}
+	\begin{table}
+		\caption[Performance of three GWAS approaches at detecting causal genes/variants]%
+		{Performance of three GWAS approaches at detecting causal genes or variants at cloned loci.}
+		\label{loci-table}
+
+		\footnotesize
+		\centering
+		\begin{threeparttable}
+
+			\csvreader[head to column names,
+			tabular = llllccc,
+			table head = \toprule Trait & Locus & Gene & Causal variant & SNPs\tnote{a} & SVs\tnote{b} & \textit{k}-mers \\ \midrule,
+			table foot = \bottomrule]%
+			{../tables/loci_table.csv}%
+			{}%
+			{\Trait & \Locus & \Gene & \Causal & \platypus & \paragraph & \kmers}
+
+			\begin{tablenotes}
+			\item[a] SNPs and indels genotyped by Platypus
+			\item[b] SVs genotyped by Paragraph
+			\item[c] Locus not discussed in main text. See Additional file 1 for detailed results.
+			\item[]- The approach detected no signal overlapping the gene.
+			\item[]* The signal found at this locus overlaps the gene.
+			\item[]** The region defined by the top 5\% (SVs) or 1\% (SNPs and \textit{k}-mers) variants overlaps the gene.
+			\item[]*** The most significantly associated variant is the causal variant.
+			\end{tablenotes}
+
+		\end{threeparttable}
+	\end{table}
+%\end{landscape}
+
 \begin{table}[h!]

### 2023-03-31 additional_files/manuscript.tex
Add methods and other miscellaneous information to the manuscript

@@
 
+\subsection*{Sample selection and processing of sequencing data}
+
+We selected 389 inbred \emph{G. max} accessions based on the availability of
+Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
+sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
+Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
+and extracted paired-end reads using the fastq-dump command v. 2.9.6
+(\citeauthor[][\url{https://github.com/ncbi/sra-tools}]{sratoolkit}) with the
+option \texttt{-\-split-3}.
+Reads that were not paired in the raw sequencing data were not used
+for downstream analyses.  The runs downloaded from the SRA and associated
+metadata are listed in Additional file 2.
+
+We first filtered the sequencing data for quality and the presence of sequencing
+adapters using bbduk from BBtools v. 38.25
+(\citeauthor[][\url{https://sourceforge.net/projects/bbmap/}]{bbtools}). Reads
+were then mapped to the reference assembly version 4 of \emph{G. max}
+cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
+and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
+bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
+reads in paired-end mode. Reads that were left unpaired following adapter- and
+quality-trimming were aligned separately in single-end mode with bwa-mem and
+merged with mapped paired-end reads using samtools \texttt{merge}
+v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
+individual SRA runs to the BAM files and merged the reads belonging to the same
+accession using bamaddrg
+(\citeauthor[][\url{https://github.com/ekg/bamaddrg}]{bamaddrg}). We used the
+resulting BAM files for all downstream analyses requiring aligned reads.
+Metadata on mapping depth of aligned data can be found in Additional file 3.
+
+\subsection*{Discovery and genotyping of SVs}
+
+Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
+The \emph{discovery} step identified a set of candidate SVs using various methods,
+whereas the \emph{genotyping} step determined the genotype
+of the candidate SVs for all accessions in the population from the mapped Illumina reads. We used three different
+approaches in the SV discovery step to generate a set of candidate variants:
+
+\begin{itemize}
+	\item We used Illumina WGS data from the 389 accessions mentioned above
+		to call SVs in the population following a pipeline outlined
+		by \cite{lemay2022}. In brief, we combined information from
+		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
+		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
+		a set of candidate SVs from the mapped Illumina data.
+	\item We used SVs discovered by \cite{lemay2022} from Oxford Nanopore
+		sequencing data of 17 Canadian soybean cultivars.
+	\item We called SVs from 26 high-quality genome assemblies published by
+		\cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
+		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
+		We used methods similar to those of \cite{liu2020-pangenome} to call
+		SVs from the comparison of these assemblies to that of Williams82.
+\end{itemize}
+
+The SVs called using each of these approaches were merged using SVmerge
+\citep{wong2010} and the resulting set of candidate SVs was used for
+genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
+from the mapped Illumina reads for all accessions. We filtered the genotyped
+SVs by setting the genotype calls made from less than two reads to missing and
+by removing variants with a minor allele frequency < 0.02 or a proportion of
+missing data > 0.5. The resulting dataset of 186,306 genotyped SVs was used for
+GWAS. Detailed methods regarding the SV discovery and genotyping steps can be
+found in Additional file 1.
+
+\subsection*{Discovery and genotyping of SNPs and indels}
+
+We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
+mapped reads of the 389 accessions. The 21.1 M SNPs and
+indels from the raw output of Platypus were filtered to keep only those with:
+
+\begin{enumerate}
+	\item \texttt{FILTER} field set to \texttt{PASS}
+	\item minor allele frequency (MAF) $\geq$ 0.02
+	\item fraction of missing genotypes $\leq$ 0.5
+	\item heterozygosity rate $\leq$ 0.2
+\end{enumerate}
+
+Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
+PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
+step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
+implemented to reduce the computational requirements of the GWAS analysis by
+removing co-segregating markers. Markers located on unanchored scaffolds were
+also removed. The resulting dataset was used for GWAS and comprised 773,060
+SNPs and 151,570 indels.
+
+\subsection*{Computing the presence/absence table of \emph{k}-mers}
+
+We used the approach outlined by \cite{voichek2020} for \emph{k}-mer-based GWAS
+and their recommendations on the
+associated GitHub page (v. 0.2-beta, \url{https://github.com/voichek/kmersGWAS}) to generate
+a presence/absence \emph{k}-mer table for use in GWAS.  The first step of
+the approach involved counting all \emph{k}-mers of length 31 (31-mers) present
+in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
+\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
+canonized form (the first of either the observed \textit{k}-mer or its reverse
+complement in lexicographical order) and the second time using the observed
+\emph{k}-mers themselves (non-canonized form). The two sets of \textit{k}-mers
+were then combined and filtered by
+keeping only those seen in at least 5 accessions, and observed in both canonized
+and non-canonized form in at least 20\% of the accessions in which they are found.
+We then generated a table indicating the presence or absence of each
+\emph{k}-mer in all accessions. This table was used to compute a kinship
+table for use in GWAS and was also used directly as input genotypes for the
+GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
+computation of kinship and GWAS.
+
+\subsection*{Phenotypic data}
+
+We analyzed ten qualitative traits previously studied by \cite{bandillo2017}: flower
+color, pubescence color, seed coat color, stem termination type, hilum color,
+pod color, pubescence form, pubescence density, seed coat luster, and maturity
+group.  We obtained phenotypic data for these traits from the GRIN database
+(\url{https://npgsweb.ars-grin.gov/gringlobal/search}) by querying the database
+using the PI identifiers of the 389 accessions used.  Four out of the 389
+accessions did not match the GRIN database and were therefore not used for the
+analysis of qualitative traits. The phenotypes of qualitative traits were recoded
+to numerical values for GWAS analysis following methods similar to those of
+\cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
+using targeted subsets of the observed phenotypes in order to focus on specific
+loci. The phenotypes and numerical values used for all analyses as well as the
+number of observations in the dataset that we used for GWAS are listed in
+Tables S2 to S11 (Additional file 1).
+
+We analyzed three quantitative traits in addition to the ten qualitative traits
+mentioned above. Seed oil and protein content were retrieved from the GRIN
+database. In cases where more than one value was listed for a given accession,
+we computed the average of those values and used it for GWAS analyses. As a third quantitative trait, we
+used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
+weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
+Our analysis included 340 accessions whose phenotypic data
+was already published by \cite{deronne2022} as well as 49 additional accessions for
+which sequencing data was not yet available when their study was conducted.
+Because of ambiguity regarding the identity of one of the accessions (HN019), this accession
+was dropped from all GWAS analyses. Similarly to seed oil and
+protein content, the CDW values for some accessions for which more than one
+observation was available were averaged.
+
+Because of clear discrepancies between the observed haplotypes and the reported
+phenotypes for some simple traits such as flower color and seed coat color, we
+suspected possible errors in the identity of some of the accessions. To
+investigate this, we compared the SNP genotypes derived from the SoySNP50K
+array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
+90\% concordance between the two genotype datasets (see Additional file 1 for
+detailed methods). These samples were removed from the dataset in addition to another
+accession that exhibited an atypical GC content suggesting contamination.
+Following these filtering steps and the removal of HN019, 363 accessions
+remained for GWAS analyses. The concordance between WGS data and SoySNP50K
+genotypes is included in Additional file 3 and averaged 98.3\% among retained
+lines (Additional file 1: Figure S45). The phenotypic data used in this study can be found in Additional file 4.
+
+\subsection*{GWAS analyses}
+
+We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
+the SNP/indel and SV datasets described above. SVs and indels required some additional
+preprocessing steps to make variant
+representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
+variants as SNPs prior to converting the VCF files to Hapmap diploid format
+using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
+retrieval of the proper metadata for each variant after GWAS.
+
+We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
+algorithm for computing the kinship matrix. We used an MLM model for
+consistency with the statistical model used by the \textit{k}-mer-based
+approach.  We used a randomization approach to determine the 5\% family-wise
+error-rate threshold for each GWAS analysis as described in \cite{voichek2020}.
+To compute this threshold, we permuted the phenotypic observations 100 times
+and computed a GWAS on these permuted phenotypes to obtain a distribution of
+top \emph{p}-values under the null hypothesis. The fifth most significant of
+the top 100 \emph{p}-values was used as the significance threshold for
+inference, i.e. markers with \emph{p}-values lower than this threshold were
+considered significant.
+
+We conducted GWAS based on \emph{k}-mers using the method developed by
+\cite{voichek2020}. In brief, the
+analysis was conducted in two steps. The first step used an approximate model
+to identify 1 million potentially significant \emph{k}-mers. These were then used in
+the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
+in a \emph{p}-value for each \emph{k}-mer. We
+used the kinship computed from the \emph{k}-mers to correct for relatedness
+between accessions. The program automatically outputs a list of \emph{k}-mers
+that passed a 5\% family-wise error-rate threshold determined by a
+randomization approach, which we considered significant for the purposes of
+downstream analyses.
+
+\subsection*{Analysis of significant \emph{k}-mers}
+
+\emph{k}-mers are not intrinsically associated with a particular genomic region
+and a tailored analysis was therefore required to associate them with genomic
+coordinates. To do so, we queried the mapped reads (in BAM format) of all accessions for matches
+to significant \emph{k}-mers for a given trait. Because this is a
+computationally demanding operation (there were typically tens of thousands of
+significant \emph{k}-mers per trait), we developed a C program called katcher
+(\url{https://github.com/malemay/katcher}) that
+uses the htslib library \citep{bonfield2021} for efficient reading and
+writing of BAM files. katcher and its associated utilities allow for efficient
+retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
+of interest.
+
+Once we retrieved all the reads containing significant \emph{k}-mers, we next
+needed to gather the mapping information contained in those reads for the
+purposes of identifying contiguous signals and comparing them to annotations. To
+simplify the amount of information to process in downstream analyses, we
+considered only the most significant \emph{k}-mer contained in a given read and
+the position where this \emph{k}-mer was most often observed in
+the dataset. Moreover, we limited the analysis to reads with a minimum mapping
+quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
+least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
+from sequencing errors. All downstream analyses (Manhattan plots,
+identification of signals, analysis of genes located near significant
+\emph{k}-mers, etc.) used this processed set of significant \emph{k}-mers.
+
+\subsection*{Downstream analyses of significant association signals}
+
+We used contiguous regions of significant variants, which we called
+``signals'', as a basis to compare our results with those obtained by previous
+studies and to query the genome for candidate genes. We defined the boundaries
+of those signals by grouping any significant markers or \emph{k}-mers that
+occurred within 250 kb of one another.  We did consider the length of deletions
+or \emph{k}-mers in defining the distance between significant variants.
+
+For the GWAS analyses computed from SNP/indel genotype calls, we retrieved all
+markers that had initially been pruned in a region ranging from 50 kb
+downstream to 50 kb upstream of significant signals and computed their
+\emph{p}-values by running GAPIT using the same parameters as noted above. The
+\emph{p}-values thus computed on markers that were initially left out were not
+used for initial signal discovery, but provided finer resolution when analyzing
+the results of the SNP-based GWAS analyses. Indeed, some markers that had not
+been included in the original GWAS may be the causal variant or simply yield
+more significant \emph{p}-values than those initially used.
+
+To help with the interpretation of results and the identification of candidate
+genes, we defined regions of highly significant associations within each signal
+by delimiting a range of coordinates containing the top 5\% most significant
+associations (Paragraph) or 1\% most significant associations (SNPs/indels
+and \emph{k}-mers) found within the signal. In cases where such a fraction
+represented fewer than two markers or \emph{k}-mers, the highly significant
+region was simply delimited by the two most significant associations.
+
+Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
+Williams82, we converted the genomic coordinates of their signals to assembly version 4
+using the same method that we used for converting
+SoySNP50K positions (see Additional file 1). For the analysis of
+qualitative traits, we only compared our results to signals at named classical
+loci and ignored previously reported minor signals with \emph{p}-values that
+just barely made it above the significance threshold.
+
+\subsection*{Linking significant \emph{k}-mers to variants}
+
+We used the reads retrieved by katcher and an assembly-based approach for
+identifying the variants associated with a given set of \emph{k}-mers. In
+addition to the reads themselves, we also retrieved their paired reads even
+though they may not have matched a significant \emph{k}-mer in the first place.
+These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
+\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
+individually for each accession and then aligned to the Williams82 reference
+assembly using bwa-mem with default parameters.  Aligned assembled sequences
+that overlapped loci of interest were realigned using multiple sequence
+alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
+visualizing the haplotypes at these loci. The identification of variants
+supported by significant \textit{k}-mers was based entirely on visual analysis
+of haplotype alignments. We only considered haplotypes occurring at least five
+times in the population in such analyses in order to leave out potential
+assembly errors. In a single case (\emph{Dt1} locus), assembly at the locus
+succeeded for a very low number of accessions (26). In this case, we instead used
+consensus sequences from the reads aligned with bwa for comparing the
+haplotypes across accessions.
+
+\subsection*{Computation of LD between \emph{k}-mers}
+
+In some cases, the analysis of \emph{k}-mers revealed signals at previously
+undocumented loci or found a very large number of associations throughout the
+genome. To shed light on these situations, we computed the pairwise LD between
+\emph{k}-mers in order to identify \emph{k}-mers that co-segregated and
+likely corresponded to a single locus. Because of the sheer number of pairwise
+comparisons that needed to be made for some traits (> 20,000 significant
+\emph{k}-mers kept even after the filtering steps mentioned above), we limited
+the number of \emph{k}-mers used for LD calculation to 1,500. For traits where
+more than 1,500 \emph{k}-mers were found, we computed LD on a subsample of
+these by selecting the 500 \emph{k}-mers with the most significant
+\emph{p}-values and sampling the remaining 1000 \emph{k}-mers randomly. This
+random sampling was done with a probability inversely proportional to the
+number of \emph{k}-mers matching a given chromosome or scaffold such that
+\emph{k}-mers were sampled from the whole genome and not simply from the most
+significant locations.
+
+\subsection*{Software used}
+\label{sv-gwas-software-used}
+
+Unless otherwise noted, all analyses were conducted using R v. 4.2.0
+\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
+packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
+\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
+1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
+\citep{obenchain2014}.  We gathered several functions used for the downstream
+analysis of GWAS results into a package called gwask which is available on
+GitHub (\url{https://github.com/malemay/gwask}).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section*{Acknowledgements}%% if any
-Text for this section\ldots
+
+We would like to thank Jonas A. Sibbesen for providing very helpful comments
+on the methodology used and on a previous version of this manuscript.
+We also thank Yoav Voichek for help regarding \textit{k}-mer GWAS
+analysis as well as Brian Boyle and Martine Jean for valuable comments
+regarding the methodology. We are thankful to the Digital Research Alliance of
+Canada for using their high-performance computing servers.
 
 \section*{Funding}%% if any
-Text for this section\ldots
+
+This work was supported by the SoyaGen grant (\url{https://www.soyagen.ca})
+awarded to F. Belzile and R. Bélanger, and funded by Génome Québec, Genome
+Canada, the government of Canada, the Ministère de l'Économie, Science et
+Innovation du Québec, Semences Prograin Inc., Syngenta Canada Inc., Sevita
+Genetics, Coop Fédérée, Grain Farmers of Ontario, Saskatchewan Pulse Growers,
+Manitoba Pulse \& Soybean Growers, the Canadian Field Crop Research Alliance
+and Producteurs de grains du Québec. M-A. Lemay has been supported by a NSERC
+Canada Vanier Graduate Scholarship, a FRQNT doctoral B2X scholarship, a NSERC
+Michael Smith Foreign Study Supplement, and a scholarship from the
+AgroPhytoSciences NSERC CREATE Training Program. None of the funding bodies
+were involved in study design, data acquisition, data analysis, interpretation
+of the results, or manuscript writing.
 
 \section*{Abbreviations}%% if any
-Text for this section\ldots
-
-\section*{Availability of data and materials}%% if any
-Text for this section\ldots
+BAM: Binary alignment map format;
+CDW: corrected dry weight;
+CHS: chalcone synthas;
+CNV: copy number variant, copy number variation;
+GRIN: Germplasm Resources Information Network;
+GWAS: Genome-wide association study;
+HPS: Hydrophobic protein from soybean;
+LD: Linkage disequilibrium;
+MAF: Minor allele frequency;
+MATE: ultidrug and toxic compound extrusion;
+MLM: Mixed linear model;
+NCBI: National Center for Biotechnology Information;
+SNP: Single-nucleotide polymorphism;
+SRA: Sequence Read Archive;
+SV: Structural variant, structural variation;
+VCF: Variant call format;
+WGS: Whole-genome (re)sequencing
+
+\section*{Availability of data and materials}
+
+Some of the datasets generated as part of this work are available on figshare
+\citep{kmer-gwas-figshare}.
+
+The Illumina sequencing data used for the analyses are listed in Additional file 2
+and available through NCBI BioProjects \citeauthor{PRJNA257011}, \citeauthor{PRJNA289660}
+and \citeauthor{PRJNA639876}.
+
+The high-quality assemblies generated by \cite{liu2020-pangenome} are available
+on the Genome Warehouse through Accession Number \citeauthor{PRJCA002030}.
+
+The assemblies of ZH13, W05 and Lee are available on SoyBase
+(\url{https://soybase.org/GlycineBlastPages/blast_descriptions.php}).
+
+The \textit{I} locus contig assembled by \cite{tuteja2008} is available from
+NCBI (\url{https://www.ncbi.nlm.nih.gov/nuccore/EF623854}).
+
+SoySNP50K genotype calls are available from SoyBase (\url{https://soybase.org/snps/}).
+
+The SVs identified from Oxford Nanopore data by \cite{lemay2022} are available
+on figshare \citep{soybean-sv-figshare}.
+
+The following software used in this work is available on GitHub:
+
+\begin{itemize}
+	\item Code used for the analyses:
+		\url{https://github.com/malemay/soybean_kmer_gwas}
+	\item katcher software for retrieving reads containing \textit{k}-mers:
+		\url{https://github.com/malemay/katcher}
+	\item gwask R package used for processing the output of GAPIT and
+		\textit{k}-mer GWAS analyses, and for plotting:
+		\url{https://github.com/malemay/gwask}
+	\item Software for computing the LD between \textit{k}-mers
+		based on presence/absence: \url{https://github.com/malemay/kmers_ld}
+	\item Forked svmu version that was used for the analyses shown in this
+		paper: \url{https://github.com/malemay/svmu}
+	\item svmutools R package for converting the output of svmu to VCF
+		format: \url{https://github.com/malemay/svmutools}
+\end{itemize}
+
+The code associated with the results presented in this manuscript is archived
+on figshare \citep{kmer-gwas-code}.
 
 \section*{Ethics approval and consent to participate}%% if any
-Text for this section\ldots
+Not applicable.
 
@@
 \section*{Consent for publication}%% if any
-Text for this section\ldots
+Not applicable.
 
 \section*{Authors' contributions}
-Text for this section \ldots
-
-\section*{Authors' information}%% if any
-Text for this section\ldots
+Conception and design of the study: MAL, MDR, RB, FB.
+Phenotypic data acquisition (resistance to \textit{Phytophthora sojae}): MDR, RB.
+Data analysis: MAL.
+Data interpretation: MAL, FB.
+Software writing: MAL.
+Manuscript drafting: MAL, FB.
+All authors have revised the manuscript and approved its submission.
 
@@
 % Table 1 with results on all loci with cloned genes
-%\begin{landscape}
 \begin{table}
@@
 \end{table}
-%\end{landscape}
-
-\begin{table}[h!]
-\caption{Sample table title. This is where the description of the table should go}
-  \begin{tabular}{cccc}
-    \hline
-    & B1  &B2   & B3\\ \hline
-    A1 & 0.1 & 0.2 & 0.3\\
-    A2 & ... & ..  & .\\
-    A3 & ..  & .   & .\\ \hline
-  \end{tabular}
-\end{table}
 
@@
 \section*{Additional Files}
-  \subsection*{Additional file 1 --- Sample additional file title}
-    Additional file descriptions text (including details of how to
-    view the file, if it is in a non-standard format or the file extension).  This might
-    refer to a multi-page table or a figure.
-
-  \subsection*{Additional file 2 --- Sample additional file title}
-    Additional file descriptions text.
+
+\subsection*{Additional file 1 --- Additional file 1.pdf}
+Supplemental methods and results,
+including Tables S1 to S11
+and Figures S1 to S60 (PDF 12.9 MB)
+
+\subsection*{Additional file 2 --- Additional file 2.csv}
+Metadata on the sequencing runs retrieved
+from the Sequence Read Archive (SRA) (CSV 71 KB)
+
+\subsection*{Additional file 3 --- Additional file 3.csv}
+Metadata on individual accessions following mapping
+and comparison to SoySNP50K data (CSV 36KB)
+
+\subsection*{Additional file 4 --- Additional file 4.csv}
+Phenotypic dataset used for GWAS analyses (CSV 135 KB)
 

### 2023-03-31 additional_files/manuscript.tex
Fix citations by adding author names where required

@@
 of variants in GWAS without genotyping SVs at population scale could be useful.
-\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
+Voichek and Weigel \cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
 represent such an alternative method. This approach relies on the presence or
@@
 
-\cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
+Zabala and Vodkin \cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
 Glyma.13g\-072100 as associated with the \textit{W1} locus for flower color. In
@@
 
-\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
+Zabala and Vodkin \cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
 \textit{T} locus for pubescence color and found a 1-bp deletion as a putative
@@
 approaches (Additional file 1: Figure S14).
-\cite{wang2018} identified Glyma.01g198500 as the gene associated with
+Wang et al. \cite{wang2018} identified Glyma.01g198500 as the gene associated with
 this locus and found an A>G SNP affecting transcript splicing as the causal
@@
 We did identify one of the SNPs suggested as putative causal variants for this
-locus by \cite{tian2010} and \cite{liu2010}, however this SNP was only the
+locus by Tian et al. \cite{tian2010} and Liu et al. \cite{liu2010}, however this SNP was only the
 135\textsuperscript{th} most associated marker found by Platypus and the
@@
 S21 and S22).
-\cite{gillman2011} identified this R2R3 MYB transcription factor as the
+Gillman et al. \cite{gillman2011} identified this R2R3 MYB transcription factor as the
 molecular basis for the \textit{R} locus and documented four different
@@
 well as Glyma.09g235100 and two others are part of a tandem array of four R2R3
-MYB genes considered as candidates by \cite{gillman2011}, and of which
+MYB genes considered as candidates by Gillman et al. \cite{gillman2011}, and of which
 Glyma.09g235100 was the only one expressed in the seed coat. Given the
-convincing evidence provided by \cite{gillman2011}, it is most likely that
+convincing evidence provided by Gillman et al. \cite{gillman2011}, it is most likely that
 Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
@@
 loss of function at the \emph{R} locus.  Still, we observed two causal variants
-previously documented by \cite{gillman2011} in our dataset (Additional file 1: Figures
+previously documented by Gillman et al. \cite{gillman2011} in our dataset (Additional file 1: Figures
 S23 and S24).
@@
 \ref{pubescence-density-main-figure}c), and therefore possibly corresponded to the
-25.6-kb region identified as the causal CNV by \cite{liu2020ps}. We found that
+25.6-kb region identified as the causal CNV by Liu et al. \cite{liu2020ps}. We found that
 reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
@@
 
-The difference in size between the CNV reported by \cite{liu2020ps} and our
+The difference in size between the CNV reported by Liu et al. \cite{liu2020ps} and our
 results may be due to the presence of some unresolved sequence (a stretch of
@@
 the seed coat luster trait. Seed coat luster has been shown by
-\cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
+Gijzen et al. \cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
 protein (HPS) at the seed surface. Copy number variation of this gene at the
@@
 Both the \textit{L1} and the \textit{L2} loci for pod color have yet to be
-cloned. While \cite{he2015} proposed the gene Glyma.19g101700 as likely
+cloned. While He et al. \cite{he2015} proposed the gene Glyma.19g101700 as likely
 corresponding to \textit{L1} based on fine mapping and expression analysis, our
-results as well as those reported by \cite{bandillo2017} do not support this
+results as well as those reported by Bandillo et al. \cite{bandillo2017} do not support this
 gene as a candidate due to its location $\sim$3 Mb away from GWAS signals. Our
@@
 The two known loci for pubescence form in soybean, \textit{Pa1} and
-\textit{Pa2}, have also not been cloned yet. Nevertheless, \cite{gilbert2017}
+\textit{Pa2}, have also not been cloned yet. Nevertheless, Gilbert \cite{gilbert2017}
 has suggested Glyma.12g213900 as a candidate for \textit{Pa1} based on data
-published by \cite{bandillo2017}. This gene encodes a MYB transcription factor
+published by Bandillo et al. \cite{bandillo2017}. This gene encodes a MYB transcription factor
 annotated as playing a role in trichome branching. Indeed, the most significant
@@
 
-We used corrected dry weight (CDW) as described by \cite{deronne2022} as a
+We used corrected dry weight (CDW) as described by de Ronne et al. \cite{deronne2022} as a
 measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
 overlapping the genomic region reported by these authors on Gm15 using all three GWAS
-approaches (Additional file 1: Figure S40). \cite{deronne2022}
+approaches (Additional file 1: Figure S40). de Ronne et al. \cite{deronne2022}
 proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
@@
 result in a decreased ability to identify causal variants or genes. For
-example, \cite{bandillo2017} noted that observed phenotypes for stem
+example, Bandillo et al. \cite{bandillo2017} noted that observed phenotypes for stem
 termination type in the GRIN database may not correspond to the expected
@@
 limitation of this approach. Although the method developed by
-\cite{voichek2020} is in itself relatively simple and computationally
+Voichek and Weigel \cite{voichek2020} is in itself relatively simple and computationally
 inexpensive, downstream analysis and interpretation of the output is
@@
 to SNP-based GWAS will largely depend on whether SVs can represent variation
-that is not otherwise in LD with SNPs. In tomato, \cite{dominguez2020} found
+that is not otherwise in LD with SNPs. In tomato, Domínguez et al. cite{dominguez2020} found
 that most transposable element insertions associated with phenotypic variation
 in agronomic traits were not tagged by SNPs. On the other hand, in humans,
-\cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
+Maretty et al. \cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
 is therefore likely that how well SVs can be tagged by SNPs will depend on the
@@
 and extracted paired-end reads using the fastq-dump command v. 2.9.6
-(\citeauthor[][\url{https://github.com/ncbi/sra-tools}]{sratoolkit}) with the
+\cite{sratoolkit} with the
 option \texttt{-\-split-3}.
@@
 adapters using bbduk from BBtools v. 38.25
-(\citeauthor[][\url{https://sourceforge.net/projects/bbmap/}]{bbtools}). Reads
+\cite{bbtools}. Reads
 were then mapped to the reference assembly version 4 of \emph{G. max}
@@
 accession using bamaddrg
-(\citeauthor[][\url{https://github.com/ekg/bamaddrg}]{bamaddrg}). We used the
+\cite{bamaddrg}. We used the
 resulting BAM files for all downstream analyses requiring aligned reads.
@@
 		to call SVs in the population following a pipeline outlined
-		by \cite{lemay2022}. In brief, we combined information from
+		by Lemay et al. \cite{lemay2022}. In brief, we combined information from
 		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
@@
 		a set of candidate SVs from the mapped Illumina data.
-	\item We used SVs discovered by \cite{lemay2022} from Oxford Nanopore
+	\item We used SVs discovered by Lemay et al. \cite{lemay2022} from Oxford Nanopore
 		sequencing data of 17 Canadian soybean cultivars.
 	\item We called SVs from 26 high-quality genome assemblies published by
-		\cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
+		Liu et al. \cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
 		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
-		We used methods similar to those of \cite{liu2020-pangenome} to call
+		We used methods similar to those of Liu et al. \cite{liu2020-pangenome} to call
 		SVs from the comparison of these assemblies to that of Williams82.
@@
 
-We used the approach outlined by \cite{voichek2020} for \emph{k}-mer-based GWAS
+We used the approach outlined by Voichek and Weigel \cite{voichek2020} for \emph{k}-mer-based GWAS
 and their recommendations on the
@@
 
-We analyzed ten qualitative traits previously studied by \cite{bandillo2017}: flower
+We analyzed ten qualitative traits previously studied by Bandillo et al. \cite{bandillo2017}: flower
 color, pubescence color, seed coat color, stem termination type, hilum color,
@@
 to numerical values for GWAS analysis following methods similar to those of
-\cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
+Bandillo et al. \cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
 using targeted subsets of the observed phenotypes in order to focus on specific
@@
 Our analysis included 340 accessions whose phenotypic data
-was already published by \cite{deronne2022} as well as 49 additional accessions for
+was already published by de Ronne et al. \cite{deronne2022} as well as 49 additional accessions for
 which sequencing data was not yet available when their study was conducted.
@@
 approach.  We used a randomization approach to determine the 5\% family-wise
-error-rate threshold for each GWAS analysis as described in \cite{voichek2020}.
+error-rate threshold for each GWAS analysis as described by Voichek and Weigel \cite{voichek2020}.
 To compute this threshold, we permuted the phenotypic observations 100 times
@@
 We conducted GWAS based on \emph{k}-mers using the method developed by
-\cite{voichek2020}. In brief, the
+Voichek and Weigel \cite{voichek2020}. In brief, the
 analysis was conducted in two steps. The first step used an approximate model
@@
 The Illumina sequencing data used for the analyses are listed in Additional file 2
-and available through NCBI BioProjects \citeauthor{PRJNA257011}, \citeauthor{PRJNA289660}
-and \citeauthor{PRJNA639876}.
-
-The high-quality assemblies generated by \cite{liu2020-pangenome} are available
-on the Genome Warehouse through Accession Number \citeauthor{PRJCA002030}.
+and available through NCBI BioProjects PRJNA257011 \cite{PRJNA257011}, PRJNA289660 \cite{PRJNA289660}
+and PRJNA639876 \cite{PRJNA639876}.
+
+The high-quality assemblies generated by Liu et al. \cite{liu2020-pangenome} are available
+on the Genome Warehouse through Accession Number PRJCA002030 \cite{PRJCA002030}.
 
@@
 
-The \textit{I} locus contig assembled by \cite{tuteja2008} is available from
+The \textit{I} locus contig assembled by Tuteja and Vodkin \cite{tuteja2008} is available from
 NCBI (\url{https://www.ncbi.nlm.nih.gov/nuccore/EF623854}).
@@
 
-The SVs identified from Oxford Nanopore data by \cite{lemay2022} are available
+The SVs identified from Oxford Nanopore data by Lemay et al. \cite{lemay2022} are available
 on figshare \citep{soybean-sv-figshare}.

### 2023-03-31 additional_files/manuscript.tex
Fix table and figure caption placement
Fix corresponding author and author list

@@
   addressref={aff1,aff2,aff3},                   % id's of addresses, e.g. {aff1,aff2}
-  email={marc-andre.lemay.2@ulaval.ca}   % email address
 ]{\inits{M.A.L.}\fnm{Marc-André} \snm{Lemay}}
-
 \author[
   addressref={aff1,aff2,aff3},
-  email={maxime.de-ronne.1@ulaval.ca}
 ]{\inits{M.D.R.}\fnm{Maxime} \snm{de Ronne}}
-
 \author[
   addressref={aff1,aff2,aff3},
-  email={richard.belanger@fsaa.ulaval.ca}
 ]{\inits{R.B.}\fnm{Richard} \snm{Bélanger}}
-
 \author[
@@
 
+\section*{Additional Files}
+
+\subsection*{Additional file 1 --- Additional file 1.pdf}
+Supplemental methods and results,
+including Tables S1 to S11
+and Figures S1 to S60 (PDF 12.9 MB)
+
+\subsection*{Additional file 2 --- Additional file 2.csv}
+Metadata on the sequencing runs retrieved
+from the Sequence Read Archive (SRA) (CSV 71 KB)
+
+\subsection*{Additional file 3 --- Additional file 3.csv}
+Metadata on individual accessions following mapping
+and comparison to SoySNP50K data (CSV 36KB)
+
+\subsection*{Additional file 4 --- Additional file 4.csv}
+Phenotypic dataset used for GWAS analyses (CSV 135 KB)
+
 \section*{Acknowledgements}%% if any
@@
 % Figure 1
-\begin{figure}
+\begin{figure}[p]
 	\caption{Results of SV- and \textit{k}-mer-based GWAS at the
@@
 % Figure 2
-\begin{figure}
+\begin{figure}[p]
 	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
@@
 % Figure 3 with the results on pubescence color at locus G
-\begin{figure}
+\begin{figure}[p]
 	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
@@
 % Figure 4 with the results on pubescence density at locus Ps
-\begin{figure}
+\begin{figure}[p]
 	\caption[Results of \textit{k}-mer-based GWAS at the
@@
 % Figure 5 with the results on pubescence form at locus Pa1
-\begin{figure}
+\begin{figure}[p]
 	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
@@
 % Figure 6 with the results novel loci for stem termination type
-\begin{figure}
+\begin{figure}[p]
 	\caption[Detection of novel signals for stem termination type
@@
 % Table 1 with results on all loci with cloned genes
-\begin{table}
+\begin{table}[p]
 	\caption[Performance of three GWAS approaches at detecting causal genes/variants]%
@@
 
-\section*{Additional Files}
-
-\subsection*{Additional file 1 --- Additional file 1.pdf}
-Supplemental methods and results,
-including Tables S1 to S11
-and Figures S1 to S60 (PDF 12.9 MB)
-
-\subsection*{Additional file 2 --- Additional file 2.csv}
-Metadata on the sequencing runs retrieved
-from the Sequence Read Archive (SRA) (CSV 71 KB)
-
-\subsection*{Additional file 3 --- Additional file 3.csv}
-Metadata on individual accessions following mapping
-and comparison to SoySNP50K data (CSV 36KB)
-
-\subsection*{Additional file 4 --- Additional file 4.csv}
-Phenotypic dataset used for GWAS analyses (CSV 135 KB)
-
 \end{backmatter}

### 2023-03-31 additional_files/manuscript.tex
Shorten abstract to meet GB requirements and remove dependency on csvsimple for Table 1 by embedding the table directly in the TeX file

@@
 % Packages added by M-A Lemay
-\usepackage{csvsimple} % read and display csv files as tables
 \usepackage{booktabs} % prettier tables
@@
 their power, GWAS frequently fail to pinpoint the causal variant or the gene
-controlling a trait at a given locus in crop species. Assessing genetic
+controlling a given trait in crop species. Assessing genetic
 variants beyond single-nucleotide polymorphisms (SNPs) could alleviate this
@@
 soybean by applying these methods to 13 traits.  We also performed
-conventional GWAS analysis based on SNPs and small indels for
-comparison.  We assessed the performance of each GWAS approach based on
+conventional GWAS analysis based on SNPs and small indels.  We assessed the performance of each GWAS approach based on
 results at loci for which the causal genes or variants were known from
@@
 promising candidate genes for loci related to pod color, pubescence
-form, and resistance to the oomycete \textit{Phytophthora sojae}. In
+form, and resistance to \textit{Phytophthora sojae}. In
 our dataset, SV-based GWAS did not add value compared to
@@
 	\begin{threeparttable}
-
-		\csvreader[head to column names,
-		tabular = llllccc,
-		table head = \toprule Trait & Locus & Gene & Causal variant & SNPs\tnote{a} & SVs\tnote{b} & \textit{k}-mers \\ \midrule,
-		table foot = \bottomrule]%
-		{../tables/loci_table.csv}%
-		{}%
-		{\Trait & \Locus & \Gene & \Causal & \platypus & \paragraph & \kmers}
+		\begin{tabular}{llllccc}
+		\toprule Trait & Locus & Gene & Causal variant & SNPs\tnote{a} & SVs\tnote{b} & \textit{k}-mers \\ \midrule
+		Flower color & \textit{W1} & Glyma.13g072100 & 65-bp deletion \citep{zabala2007} & ** & *** & *** \\
+		Pubescence color & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & ** & * & ** \\
+		Pubescence color & \textit{Td} & Glyma.03g258700 & SNP \citep{yan2020} & * & - & *** \\
+		Seed coat color & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & ** & ** & ** \\
+		Seed coat color & \textit{G} & Glyma.01g198500 & SNP \citep{wang2018} & ** & ** & *** \\
+		Seed coat color\tnote{c} & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & - & - & - \\
+		Seed coat color\tnote{c} & \textit{R} & Glyma.09g235100 & SNPs and indels \citep{gillman2011} & - & - & - \\
+		Stem termination type & \textit{Dt1} & Glyma.19g194300 & SNPs \citep{liu2010} & ** & * & ** \\
+		Stem termination type\tnote{c} & \textit{Dt2} & Glyma.18g273600 & SNPs \citep{ping2014} & - & - & - \\
+		Stem termination type\tnote{c} & \textit{E3} & Glyma.19g224200 & Various types \citep{tardivel2019} & - & - & - \\
+		Hilum color\tnote{c} & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & ** & ** & ** \\
+		Hilum color\tnote{c} & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & * & * & ** \\
+		Hilum color & \textit{R} & Glyma.09g235100 & SNPs and indels \citep{gillman2011} & * & - & ** \\
+		Hilum color\tnote{c} & \textit{W1} & Glyma.13g072100 & 65-bp deletion \citep{zabala2007} & - & - & - \\
+		Pubescence density & \textit{Ps} & Glyma.12g187200 & 25.6-kb CNV \citep{liu2020ps} & ** & ** & *** \\
+		Pubescence density\tnote{c} & \textit{Pd1} & Glyma.01g240100 & SNP \citep{liu2020ps} & - & - & - \\
+		Pubescence density\tnote{c} & \textit{P1} & Glyma.09g278000 & SNP \citep{liu2020ps} & - & - & - \\
+		Seed coat luster & \textit{B} & Duplicated HPS genes & CNV (several kb) \citep{gijzen2006} & ** & ** & ** \\
+		Seed coat luster\tnote{c} & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & - & - & - \\
+		Maturity group\tnote{c} & \textit{E1} & Glyma.06g207800 & Various types \citep{tardivel2019} & - & - & - \\
+		Maturity group\tnote{c} & \textit{E2} & Glyma.10g221500 & Various types \citep{tardivel2019} & - & - & - \\
+		Maturity group\tnote{c} & \textit{E3} & Glyma.19g224200 & Various types \citep{tardivel2019} & - & - & - \\
+		Maturity group\tnote{c} & \textit{E4} & Glyma.20g090000 & Various types \citep{tardivel2019} & - & - & - \\
+		Oil\tnote{c} & \textit{oilGm20} & Glyma.20g085100 & 321-bp deletion \citep{fliege2022} & - & - & - \\
+		Oil\tnote{c} & \textit{oilGm15} & Glyma.15g049200 & SNPs and indels \citep{zhang2020} & - & - & - \\
+		Protein\tnote{c} & \textit{proGm20} & Glyma.20g085100 & 321-bp deletion \citep{fliege2022} & - & - & - \\
+		Protein\tnote{c} & \textit{proGm15} & Glyma.15g049200 & SNPs and indels \citep{zhang2020} & - & - & - \\
+		\bottomrule
+		\end{tabular}
 

### 2023-03-31 additional_files/manuscript.tex
Correct some typos upon re-reading the manuscript

@@
 \usepackage{threeparttable} % easy addition of footnotes below tables
+\usepackage[T1]{fontenc}
 
@@
 to SNP-based GWAS will largely depend on whether SVs can represent variation
-that is not otherwise in LD with SNPs. In tomato, Domínguez et al. cite{dominguez2020} found
+that is not otherwise in LD with SNPs. In tomato, Domínguez et al. \cite{dominguez2020} found
 that most transposable element insertions associated with phenotypic variation
@@
 by delimiting a range of coordinates containing the top 5\% most significant
-associations (Paragraph) or 1\% most significant associations (SNPs/indels
+associations (SVs) or 1\% most significant associations (SNPs/indels
 and \emph{k}-mers) found within the signal. In cases where such a fraction
@@
 CDW: corrected dry weight;
-CHS: chalcone synthas;
+CHS: chalcone synthase;
 CNV: copy number variant, copy number variation;
@@
 MAF: Minor allele frequency;
-MATE: ultidrug and toxic compound extrusion;
+MATE: Multidrug and toxic compound extrusion;
 MLM: Mixed linear model;

### 2023-04-04 additional_files/manuscript.tex
Add line numbers for review
Begin preparing submission for The Plant Genome by using the article class and removing BMC-specific commands

@@
-%% BioMed_Central_Tex_Template_v1.06
-%%                                      %
-%  bmc_article.tex            ver: 1.06 %
-%                                       %
-
-%%% additional documentclass options:
-%  [doublespacing]
-%  [linenumbers]   - put the line numbers on margins
-
-%%% loading packages, author definitions
-
-%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
-\documentclass[doublespacing]{bmcart}
+\documentclass{article}
 
@@
 \usepackage{amsthm,amsmath}
-\RequirePackage[numbers]{natbib}
-%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
+\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
 \RequirePackage{hyperref}
 \usepackage[utf8]{inputenc} %unicode support
-%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
-%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
 
@@
 \usepackage[T1]{fontenc}
-
-\def\includegraphic{}
-\def\includegraphics{}
-
-%%% Put your definitions there:
-\startlocaldefs
-\endlocaldefs
+\usepackage{pdflscape}
+
+\usepackage{lineno}
+\linenumbers
+
+\usepackage{setspace}
+\doublespacing
+
+\title{k-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean}
+\date{}
+
+\author{Marc-André Lemay\textsuperscript{1,2,3} \and
+Maxime de Ronne\textsuperscript{1,2,3} \and
+Richard Bélanger\textsuperscript{1,2,3} \and
+François Belzile\textsuperscript{1,2,3}}
 
@@
 
-%%% Start of article front matter
-\begin{frontmatter}
-
-\begin{fmbox}
-\dochead{Research}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the title of your article here     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{k-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean}
-
-\author[
-  addressref={aff1,aff2,aff3},                   % id's of addresses, e.g. {aff1,aff2}
-]{\inits{M.A.L.}\fnm{Marc-André} \snm{Lemay}}
-\author[
-  addressref={aff1,aff2,aff3},
-]{\inits{M.D.R.}\fnm{Maxime} \snm{de Ronne}}
-\author[
-  addressref={aff1,aff2,aff3},
-]{\inits{R.B.}\fnm{Richard} \snm{Bélanger}}
-\author[
-  addressref={aff1,aff2,aff3},
-  email={francois.belzile@fsaa.ulaval.ca}
-]{\inits{F.B.}\fnm{François} \snm{Belzile}}
-
-\address[id=aff1]{%
-  \orgdiv{Département de phytologie},
-  \orgname{Université Laval},
-  \city{Québec},
-  \cny{QC, Canada}
-}
-
-\address[id=aff2]{%
-  \orgdiv{Institut de biologie intégrative et des systèmes},
-  \orgname{Université Laval},
-  \city{Québec},
-  \cny{QC, Canada}
-}
-
-\address[id=aff3]{%
-  \orgdiv{Centre de recherche et d’innovation sur les végétaux},
-  \orgname{Université Laval},
-  \city{Québec},
-  \cny{QC, Canada}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter short notes here                   %%
-%%                                          %%
-%% Short notes will be after addresses      %%
-%% on first page.                           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{artnotes}
-%%\note{Sample of title note}     % note to the article
-%\note[id=n1]{Equal contributor} % note, connected to author
-%\end{artnotes}
-
-\end{fmbox}% comment this for two column layout
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                           %%
-%% The Abstract begins here                  %%
-%%                                           %%
-%% Please refer to the Instructions for      %%
-%% authors on https://www.biomedcentral.com/ %%
-%% and include the section headings          %%
-%% accordingly for your article type.        %%
-%%                                           %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstractbox}
-
-\begin{abstract} % abstract
-\parttitle{Background}
+\maketitle
+
+\textsuperscript{1} Département de phytologie, Université Laval, Québec, QC, Canada
+
+\textsuperscript{2} Institut de biologie intégrative et des systèmes, Université Laval, Québec, QC, Canada
+
+\textsuperscript{3} Centre de recherche et d’innovation sur les végétaux, Université Laval, Québec, QC, Canada
+
+\section*{Abstract}
+
 Genome-wide association studies (GWAS) are powerful statistical methods that
@@
 
-\parttitle{Results}
 In this study, we tested the potential of SV- and \emph{k}-mer-based GWAS in
@@
 
-\parttitle{Conclusions}
 Despite promising results, significant challenges remain regarding the
@@
 
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{keyword}
-\kwd{Genome-wide association studies}
-\kwd{\textit{k}-mer GWAS}
-\kwd{Soybean genomics}
-\kwd{Crop genomics}
-\kwd{\textit{Phytophthora sojae}}
-\kwd{Structural variation}
-\end{keyword}
-
-% MSC classifications codes, if any
-%\begin{keyword}[class=AMS]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
-
-\end{abstractbox}
-%
-%\end{fmbox}% uncomment this for two column layout
-
-\end{frontmatter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
-% <put your article body there>
-
-%%%%%%%%%%%%%%%%
-%% Background %%
-%%
+\section*{Keywords}
+Genome-wide association studies
+\textit{k}-mer GWAS
+Soybean genomics
+Crop genomics
+\textit{Phytophthora sojae}
+Structural variation
+
 \section*{Background}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Backmatter begins here                   %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{backmatter}
-
 \section*{Additional Files}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  Bmc_mathpys.bst  will be used to                       %%
-%%  create a .BBL file for submission.                     %%
-%%  After submission of the .TEX file,                     %%
-%%  you will be prompted to submit your .BBL file.         %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% if your bibliography is in bibtex format, use those commands:
-\bibliographystyle{vancouver} % Style BST file (bmc-mathphys, vancouver, spbasic).
-\bibliography{references}      % Bibliography file (usually '*.bib' )
-% for author-year bibliography (bmc-mathphys or spbasic)
-% a) write to bib file (bmc-mathphys only)
-% @settings{label, options="nameyear"}
-% b) uncomment next line
-%\nocite{label}
-
-% or include bibliography directly:
-% \begin{thebibliography}
-% \bibitem{b1}
-% \end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Figures                       %%
-%%                               %%
-%% NB: this is for captions and  %%
-%% Titles. All graphics must be  %%
-%% submitted separately and NOT  %%
-%% included in the Tex document  %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Do not use \listoffigures as most will included as separate files
+\bibliographystyle{genome_research}
+\bibliography{references}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Use of \listoftables is discouraged.
-%%
 \section*{Tables}
@@
 % Table 1 with results on all loci with cloned genes
+\begin{landscape}
 \begin{table}[p]
@@
 \end{table}
+\end{landscape}
 
@@
 
-\end{backmatter}
 \end{document}

### 2023-04-04 additional_files/additional_file_1.tex
Make a few changes; version submitted to biorxiv
Set Additional file 1 such that it compiles independently of manuscript.pdf

@@
-\chapter*{Supplemental data}
+\documentclass[12pt]{report}
+
+\usepackage[utf8]{inputenc} % declare encoding as utf8
+\usepackage[T1]{fontenc} % declare encoding as utf8
+\usepackage{graphicx} % to enable \includegraphics
+\usepackage{flafter} % to make sure floats appear after their position in text
+\usepackage{pdflscape} % allow some pages to be in landscape orientation
+\usepackage{layout} % for debugging purposes, looking at page layout
+\usepackage[margin=1in]{geometry} % leaves more space for text body
+\usepackage[hidelinks]{hyperref} % hyperlinks from list of figures/tables
+\usepackage[labelfont=bf]{caption} % format captions and make links go to figure instead of caption
+\usepackage{csvsimple} % read and display csv files as tables
+\usepackage{booktabs} % prettier tables
+\usepackage{threeparttable} % easy addition of footnotes below tables
+\usepackage{textcomp} % makes the \textdegree command available
+\usepackage{tocloft} % formatting the lists of tables and figures
+\usepackage{xcolor} % enable text coloring
+\usepackage{calc} % makes \widthof available to set the width of the last column
+\usepackage{array} % using command \arraybackslash to restore the definition of \\ after \centering
+\usepackage{longtable} % allow tables to span more than one page
+\usepackage{threeparttablex} % threeparttable for longtable
+\usepackage{capt-of} % put caption on next page
+\usepackage{ccaption}
+\usepackage{listings}
+\usepackage{datatool} % allows fetching key-value pairs from a text file
+\usepackage{natbib}
+\usepackage{setspace}
+
+% Remove indent for new paragraphs
+\setlength{\parindent}{0pt}
+
+% Adjust line spacing to 1.5
+\onehalfspacing
+
+% Setting the separator to use with datatool
+\DTLsetseparator{ = }
+\graphicspath{ {../figures/} } % tell Latex to look for figures in figures/
+
+% Defining new environments as I will always want the figures centered
+\newenvironment{cfigure}
+{\begin{figure} \centering}
+{\end{figure}}
+
+\newenvironment{lsfigure}
+{\begin{landscape} \begin{figure} \centering}
+{\end{figure} \end{landscape}}
+
+\setcounter{totalnumber}{1} % Allow a maximum of one float (table/figure) per page
+
+\setlength{\parskip}{2ex}
+
+% Defining a command to generate a table of phenotypic data for a single gwas
+\newcommand{\gwastableone}[2]{
+	\begin{table}
+		\caption[Phenotypic data used for the GWAS analysis of #2]%
+		{Phenotype frequency and numerical coding of phenotypic data used for the GWAS analysis of #2.
+		``NA'' indicates a missing value in the source data.
+		Dashes indicate that this phenotype was not used in the GWAS analysis.}
+		\label{#1_gwas_table}
+
+		\centering
+		\begin{threeparttable}\small
+
+			\csvreader[head to column names,
+			/csv/separator=semicolon,
+			tabular = llc,
+			table head = \toprule Phenotype & Frequency & GWAS \\ \midrule,
+			table foot = \bottomrule]%
+			{../tables/#1_gwas_table.csv}%
+			{}%
+			{\Value & \Frequency & \GWASa}
+
+		\end{threeparttable}
+
+	\end{table}
+	\clearpage
+}
+
+% Defining a command to generate a table of phenotypic data for two gwas
+\newcommand{\gwastabletwo}[2]{
+	\begin{table}
+		\caption[Phenotypic data used for the GWAS analyses of #2]%
+		{Phenotype frequency and numerical coding of phenotypic data used in two GWAS analyses of #2.
+		``NA'' indicates a missing value in the source data.
+		Dashes indicate that this phenotype was not used in the GWAS analysis.}
+		\label{#1_gwas_table}
+
+		\centering
+		\begin{threeparttable}\small
+
+			\csvreader[head to column names,
+			/csv/separator=semicolon,
+			tabular = llcc,
+			table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2\\ \midrule,
+			table foot = \bottomrule]%
+			{../tables/#1_gwas_table.csv}%
+			{}%
+			{\Value & \Frequency & \GWASa & \GWASb}
+
+		\end{threeparttable}
+
+	\end{table}
+	\clearpage
+}
+
+% Defining a command to generate a table of phenotypic data for three gwas
+\newcommand{\gwastablethree}[2]{
+	\begin{table}
+		\caption[Phenotypic data used for the GWAS analyses of #2]%
+		{Phenotype frequency and numerical coding of phenotypic data used in three GWAS analyses of #2.
+		``NA'' indicates a missing value in the source data.
+		Dashes indicate that this phenotype was not used in the GWAS analysis.}
+		\label{#1_gwas_table}
+
+		\centering
+		\begin{threeparttable}\small
+
+			\csvreader[head to column names,
+			/csv/separator=semicolon,
+			tabular = llccc,
+			table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2 & GWAS \# 3\\ \midrule,
+			table foot = \bottomrule]%
+			{../tables/#1_gwas_table.csv}%
+			{}%
+			{\Value & \Frequency & \GWASa & \GWASb & \GWASc}
+
+		\end{threeparttable}
+
+	\end{table}
+	\clearpage
+}
+
+% Defining a command for manhattan plot figures
+\newcommand{\manhattanplot}[3]{
+	\begin{cfigure}
+		\includegraphics[width = 5.8in]{#1_manhattan}
+		\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using three genotype datasets :
+		(a) Platypus (SNPs and indels), (b) Paragraph (SVs), (c) \emph{k}-mers presence/absence.
+		The x-axis shows the position along the reference assembly version 4 of Williams82.
+		Each point shows the $-\log_{10} (p)$ associated with a particular marker or \emph{k}-mer.
+		Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
+		Vertical dotted lines indicate the position of signals associated with the trait.
+		Documented loci are colored according to whether they were found (green) or not (red) by a particular approach.
+		The ``Gm'' prefix has been left out of chromosome names for simplicity.
+		}
+		\label{#1_manhattan}
+	\end{cfigure}
+
+	\clearpage}
+
+% Defining a new command for signal plot figures
+\newcommand{\signalplot}[4]{
+	\begin{cfigure}
+		\includegraphics[width = 5.8in]{#1_signal}
+		\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using three genotype datasets:
+		(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
+		Panel (a) shows gene models over the genomic interval.
+		Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
+		#4
+		Gray shaded rectangles indicate the region delimited by the top 5\% (SVs) or top 1\% (SNPs/indels and \emph{k}-mers) associations in the signal region.
+		In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote markers that had originally been pruned but whose
+		\emph{p}-values were computed after signal discovery.
+		}
+		\label{#1_signal}
+	\end{cfigure}
+
+	\clearpage}
+
+% Defining a new command for gene plot figures
+\newcommand{\geneplot}[5]{
+	\begin{cfigure}
+		\includegraphics[width = 5.8in]{#1_gene}
+		\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
+		{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using three genotype datasets:
+		(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
+		#5
+		Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
+		In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote markers that had originally been pruned but whose
+		\emph{p}-values were computed after signal discovery.
+		Panel (a) shows gene models over the plotting interval.
+		Exons are represented by rectangles whereas introns are represented by horizontal lines.
+		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
+		}
+		\label{#1_gene}
+	\end{cfigure}
+
+	\clearpage}
+
+% Defining a new command for k-mer plot figures
+\newcommand{\kmerplot}[5]{
+	\begin{cfigure}
+		\includegraphics[width = 5.8in]{#1_kmers}
+		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
+		{Identification of #5 underlying significant \emph{k}-mers at the #2 gene associated with the #3 locus for #4.
+		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
+		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
+		The red rectangle identifies the region that is zoomed-in in panel (b).
+		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
+		Individual nucleotides are colored according to the $-\log_{10}(p)$ of the most significant \emph{k}-mer overlapping them.
+		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
+		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes correspond to those shown in panel (b).}
+		\label{#1_kmers}
+	\end{cfigure}
+
+	\clearpage}
+
+% Defining a new command for LD plot figures
+\newcommand{\ldplot}[3]{
+	\begin{cfigure}
+		\includegraphics[width = 5.8in]{#1_ld}
+		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
+		\emph{k}-mers are sorted along the y-axis according to their putative position along the reference assembly version 4 of Williams82,
+		as identified by ``Gm'' chromosome labels.
+		Sequences that lack a ``Gm'' label (top of the y-axis) represent unanchored scaffolds.
+		\emph{k}-mers are represented in the same order along the x- and y-axis.
+		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p)$ of each \emph{k}-mer.}
+		\label{#1_ld}
+	\end{cfigure}
+
+	\clearpage}
+
+% END OF PREAMBLE
+
+\begin{document}
+
+\chapter*{Supplemental data to ``\textit{k}-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean''}
 \label{supdata}
@@
 \newcommand{\getvar}[1]{\DTLfetch{variables}{key}{#1}{value}}
+
+% Turning page numbering off for the supplements
+\setcounter{page}{1}
+
+% List of authors
+\textbf{Authors}: Marc-André Lemay, Maxime de Ronne, Richard Bélanger, François Belzile
 
@@
 \bibliography{references}                   % production de la bibliographie
+
+\end{document}

### 2023-04-04 additional_files/manuscript.tex
Invert Methods and Results and change figure numbers accordingly

@@
 
+\section*{Methods}
+
+\subsection*{Sample selection and processing of sequencing data}
+
+We selected 389 inbred \emph{G. max} accessions based on the availability of
+Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
+sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
+Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
+and extracted paired-end reads using the fastq-dump command v. 2.9.6
+\cite{sratoolkit} with the
+option \texttt{-\-split-3}.
+Reads that were not paired in the raw sequencing data were not used
+for downstream analyses.  The runs downloaded from the SRA and associated
+metadata are listed in Additional file 2.
+
+We first filtered the sequencing data for quality and the presence of sequencing
+adapters using bbduk from BBtools v. 38.25
+\cite{bbtools}. Reads
+were then mapped to the reference assembly version 4 of \emph{G. max}
+cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
+and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
+bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
+reads in paired-end mode. Reads that were left unpaired following adapter- and
+quality-trimming were aligned separately in single-end mode with bwa-mem and
+merged with mapped paired-end reads using samtools \texttt{merge}
+v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
+individual SRA runs to the BAM files and merged the reads belonging to the same
+accession using bamaddrg
+\cite{bamaddrg}. We used the
+resulting BAM files for all downstream analyses requiring aligned reads.
+Metadata on mapping depth of aligned data can be found in Additional file 3.
+
+\subsection*{Discovery and genotyping of SVs}
+
+Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
+The \emph{discovery} step identified a set of candidate SVs using various methods,
+whereas the \emph{genotyping} step determined the genotype
+of the candidate SVs for all accessions in the population from the mapped Illumina reads. We used three different
+approaches in the SV discovery step to generate a set of candidate variants:
+
+\begin{itemize}
+	\item We used Illumina WGS data from the 389 accessions mentioned above
+		to call SVs in the population following a pipeline outlined
+		by Lemay et al. \cite{lemay2022}. In brief, we combined information from
+		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
+		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
+		a set of candidate SVs from the mapped Illumina data.
+	\item We used SVs discovered by Lemay et al. \cite{lemay2022} from Oxford Nanopore
+		sequencing data of 17 Canadian soybean cultivars.
+	\item We called SVs from 26 high-quality genome assemblies published by
+		Liu et al. \cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
+		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
+		We used methods similar to those of Liu et al. \cite{liu2020-pangenome} to call
+		SVs from the comparison of these assemblies to that of Williams82.
+\end{itemize}
+
+The SVs called using each of these approaches were merged using SVmerge
+\citep{wong2010} and the resulting set of candidate SVs was used for
+genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
+from the mapped Illumina reads for all accessions. We filtered the genotyped
+SVs by setting the genotype calls made from less than two reads to missing and
+by removing variants with a minor allele frequency < 0.02 or a proportion of
+missing data > 0.5. The resulting dataset of 186,306 genotyped SVs was used for
+GWAS. Detailed methods regarding the SV discovery and genotyping steps can be
+found in Additional file 1.
+
+\subsection*{Discovery and genotyping of SNPs and indels}
+
+We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
+mapped reads of the 389 accessions. The 21.1 M SNPs and
+indels from the raw output of Platypus were filtered to keep only those with:
+
+\begin{enumerate}
+	\item \texttt{FILTER} field set to \texttt{PASS}
+	\item minor allele frequency (MAF) $\geq$ 0.02
+	\item fraction of missing genotypes $\leq$ 0.5
+	\item heterozygosity rate $\leq$ 0.2
+\end{enumerate}
+
+Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
+PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
+step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
+implemented to reduce the computational requirements of the GWAS analysis by
+removing co-segregating markers. Markers located on unanchored scaffolds were
+also removed. The resulting dataset was used for GWAS and comprised 773,060
+SNPs and 151,570 indels.
+
+\subsection*{Computing the presence/absence table of \emph{k}-mers}
+
+We used the approach outlined by Voichek and Weigel \cite{voichek2020} for \emph{k}-mer-based GWAS
+and their recommendations on the
+associated GitHub page (v. 0.2-beta, \url{https://github.com/voichek/kmersGWAS}) to generate
+a presence/absence \emph{k}-mer table for use in GWAS.  The first step of
+the approach involved counting all \emph{k}-mers of length 31 (31-mers) present
+in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
+\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
+canonized form (the first of either the observed \textit{k}-mer or its reverse
+complement in lexicographical order) and the second time using the observed
+\emph{k}-mers themselves (non-canonized form). The two sets of \textit{k}-mers
+were then combined and filtered by
+keeping only those seen in at least 5 accessions, and observed in both canonized
+and non-canonized form in at least 20\% of the accessions in which they are found.
+We then generated a table indicating the presence or absence of each
+\emph{k}-mer in all accessions. This table was used to compute a kinship
+table for use in GWAS and was also used directly as input genotypes for the
+GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
+computation of kinship and GWAS.
+
+\subsection*{Phenotypic data}
+
+We analyzed ten qualitative traits previously studied by Bandillo et al. \cite{bandillo2017}: flower
+color, pubescence color, seed coat color, stem termination type, hilum color,
+pod color, pubescence form, pubescence density, seed coat luster, and maturity
+group.  We obtained phenotypic data for these traits from the GRIN database
+(\url{https://npgsweb.ars-grin.gov/gringlobal/search}) by querying the database
+using the PI identifiers of the 389 accessions used.  Four out of the 389
+accessions did not match the GRIN database and were therefore not used for the
+analysis of qualitative traits. The phenotypes of qualitative traits were recoded
+to numerical values for GWAS analysis following methods similar to those of
+Bandillo et al. \cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
+using targeted subsets of the observed phenotypes in order to focus on specific
+loci. The phenotypes and numerical values used for all analyses as well as the
+number of observations in the dataset that we used for GWAS are listed in
+Tables S1 to S10 (Additional file 1).
+
+We analyzed three quantitative traits in addition to the ten qualitative traits
+mentioned above. Seed oil and protein content were retrieved from the GRIN
+database. In cases where more than one value was listed for a given accession,
+we computed the average of those values and used it for GWAS analyses. As a third quantitative trait, we
+used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
+weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
+Our analysis included 340 accessions whose phenotypic data
+was already published by de Ronne et al. \cite{deronne2022} as well as 49 additional accessions for
+which sequencing data was not yet available when their study was conducted.
+Because of ambiguity regarding the identity of one of the accessions (HN019), this accession
+was dropped from all GWAS analyses. Similarly to seed oil and
+protein content, the CDW values for some accessions for which more than one
+observation was available were averaged.
+
+Because of clear discrepancies between the observed haplotypes and the reported
+phenotypes for some simple traits such as flower color and seed coat color, we
+suspected possible errors in the identity of some of the accessions. To
+investigate this, we compared the SNP genotypes derived from the SoySNP50K
+array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
+90\% concordance between the two genotype datasets (see Additional file 1 for
+detailed methods). These samples were removed from the dataset in addition to another
+accession that exhibited an atypical GC content suggesting contamination.
+Following these filtering steps and the removal of HN019, 363 accessions
+remained for GWAS analyses. The concordance between WGS data and SoySNP50K
+genotypes is included in Additional file 3 and averaged 98.3\% among retained
+lines (Additional file 1: Figure S1). The phenotypic data used in this study can be found in Additional file 4.
+
+\subsection*{GWAS analyses}
+
+We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
+the SNP/indel and SV datasets described above. SVs and indels required some additional
+preprocessing steps to make variant
+representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
+variants as SNPs prior to converting the VCF files to Hapmap diploid format
+using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
+retrieval of the proper metadata for each variant after GWAS.
+
+We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
+algorithm for computing the kinship matrix. We used an MLM model for
+consistency with the statistical model used by the \textit{k}-mer-based
+approach.  We used a randomization approach to determine the 5\% family-wise
+error-rate threshold for each GWAS analysis as described by Voichek and Weigel \cite{voichek2020}.
+To compute this threshold, we permuted the phenotypic observations 100 times
+and computed a GWAS on these permuted phenotypes to obtain a distribution of
+top \emph{p}-values under the null hypothesis. The fifth most significant of
+the top 100 \emph{p}-values was used as the significance threshold for
+inference, i.e. markers with \emph{p}-values lower than this threshold were
+considered significant.
+
+We conducted GWAS based on \emph{k}-mers using the method developed by
+Voichek and Weigel \cite{voichek2020}. In brief, the
+analysis was conducted in two steps. The first step used an approximate model
+to identify 1 million potentially significant \emph{k}-mers. These were then used in
+the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
+in a \emph{p}-value for each \emph{k}-mer. We
+used the kinship computed from the \emph{k}-mers to correct for relatedness
+between accessions. The program automatically outputs a list of \emph{k}-mers
+that passed a 5\% family-wise error-rate threshold determined by a
+randomization approach, which we considered significant for the purposes of
+downstream analyses.
+
+\subsection*{Analysis of significant \emph{k}-mers}
+
+\emph{k}-mers are not intrinsically associated with a particular genomic region
+and a tailored analysis was therefore required to associate them with genomic
+coordinates. To do so, we queried the mapped reads (in BAM format) of all accessions for matches
+to significant \emph{k}-mers for a given trait. Because this is a
+computationally demanding operation (there were typically tens of thousands of
+significant \emph{k}-mers per trait), we developed a C program called katcher
+(\url{https://github.com/malemay/katcher}) that
+uses the htslib library \citep{bonfield2021} for efficient reading and
+writing of BAM files. katcher and its associated utilities allow for efficient
+retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
+of interest.
+
+Once we retrieved all the reads containing significant \emph{k}-mers, we next
+needed to gather the mapping information contained in those reads for the
+purposes of identifying contiguous signals and comparing them to annotations. To
+simplify the amount of information to process in downstream analyses, we
+considered only the most significant \emph{k}-mer contained in a given read and
+the position where this \emph{k}-mer was most often observed in
+the dataset. Moreover, we limited the analysis to reads with a minimum mapping
+quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
+least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
+from sequencing errors. All downstream analyses (Manhattan plots,
+identification of signals, analysis of genes located near significant
+\emph{k}-mers, etc.) used this processed set of significant \emph{k}-mers.
+
+\subsection*{Downstream analyses of significant association signals}
+
+We used contiguous regions of significant variants, which we called
+``signals'', as a basis to compare our results with those obtained by previous
+studies and to query the genome for candidate genes. We defined the boundaries
+of those signals by grouping any significant markers or \emph{k}-mers that
+occurred within 250 kb of one another.  We did consider the length of deletions
+or \emph{k}-mers in defining the distance between significant variants.
+
+For the GWAS analyses computed from SNP/indel genotype calls, we retrieved all
+markers that had initially been pruned in a region ranging from 50 kb
+downstream to 50 kb upstream of significant signals and computed their
+\emph{p}-values by running GAPIT using the same parameters as noted above. The
+\emph{p}-values thus computed on markers that were initially left out were not
+used for initial signal discovery, but provided finer resolution when analyzing
+the results of the SNP-based GWAS analyses. Indeed, some markers that had not
+been included in the original GWAS may be the causal variant or simply yield
+more significant \emph{p}-values than those initially used.
+
+To help with the interpretation of results and the identification of candidate
+genes, we defined regions of highly significant associations within each signal
+by delimiting a range of coordinates containing the top 5\% most significant
+associations (SVs) or 1\% most significant associations (SNPs/indels
+and \emph{k}-mers) found within the signal. In cases where such a fraction
+represented fewer than two markers or \emph{k}-mers, the highly significant
+region was simply delimited by the two most significant associations.
+
+Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
+Williams82, we converted the genomic coordinates of their signals to assembly version 4
+using the same method that we used for converting
+SoySNP50K positions (see Additional file 1). For the analysis of
+qualitative traits, we only compared our results to signals at named classical
+loci and ignored previously reported minor signals with \emph{p}-values that
+just barely made it above the significance threshold.
+
+\subsection*{Linking significant \emph{k}-mers to variants}
+
+We used the reads retrieved by katcher and an assembly-based approach for
+identifying the variants associated with a given set of \emph{k}-mers. In
+addition to the reads themselves, we also retrieved their paired reads even
+though they may not have matched a significant \emph{k}-mer in the first place.
+These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
+\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
+individually for each accession and then aligned to the Williams82 reference
+assembly using bwa-mem with default parameters.  Aligned assembled sequences
+that overlapped loci of interest were realigned using multiple sequence
+alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
+visualizing the haplotypes at these loci. The identification of variants
+supported by significant \textit{k}-mers was based entirely on visual analysis
+of haplotype alignments. We only considered haplotypes occurring at least five
+times in the population in such analyses in order to leave out potential
+assembly errors. In a single case (\emph{Dt1} locus), assembly at the locus
+succeeded for a very low number of accessions (26). In this case, we instead used
+consensus sequences from the reads aligned with bwa for comparing the
+haplotypes across accessions.
+
+\subsection*{Computation of LD between \emph{k}-mers}
+
+In some cases, the analysis of \emph{k}-mers revealed signals at previously
+undocumented loci or found a very large number of associations throughout the
+genome. To shed light on these situations, we computed the pairwise LD between
+\emph{k}-mers in order to identify \emph{k}-mers that co-segregated and
+likely corresponded to a single locus. Because of the sheer number of pairwise
+comparisons that needed to be made for some traits (> 20,000 significant
+\emph{k}-mers kept even after the filtering steps mentioned above), we limited
+the number of \emph{k}-mers used for LD calculation to 1,500. For traits where
+more than 1,500 \emph{k}-mers were found, we computed LD on a subsample of
+these by selecting the 500 \emph{k}-mers with the most significant
+\emph{p}-values and sampling the remaining 1000 \emph{k}-mers randomly. This
+random sampling was done with a probability inversely proportional to the
+number of \emph{k}-mers matching a given chromosome or scaffold such that
+\emph{k}-mers were sampled from the whole genome and not simply from the most
+significant locations.
+
+\subsection*{Software used}
+\label{sv-gwas-software-used}
+
+Unless otherwise noted, all analyses were conducted using R v. 4.2.0
+\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
+packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
+\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
+1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
+\citep{obenchain2014}.  We gathered several functions used for the downstream
+analysis of GWAS results into a package called gwask which is available on
+GitHub (\url{https://github.com/malemay/gwask}).
+
 \section*{Results and discussion}
@@
 not discussed here are referred to Additional file 1. In particular, Table
-S1 (Additional file 1) provides a summary of the
+S11 (Additional file 1) provides a summary of the
 \textit{p}-values obtained by each approach at each of the loci considered in
@@
 GWAS approaches detected a signal overlapping the gene at this locus (Additional file 1: Figures
-S1 and S2). The most
+S2 and S3). The most
 significant SV and \textit{k}-mers corresponded to the known causal variant
-(Figure \ref{flower-color-main-figure}a-b and Additional file 1: S3c-d).
+(Figure \ref{flower-color-main-figure}a-b and Additional file 1: S4c-d).
 In particular, the assembly of the reads containing the most significant \textit{k}-mers
@@
 and indels located within the gene sequence were among the most significantly
-associated ones (Additional file 1: Figure S3b).
+associated ones (Additional file 1: Figure S4b).
 
@@
 \textit{k}-mers scattered throughout the genome (Additional file 1: Figure
-S1c). An analysis of the pairwise LD between a subset
+S2c). An analysis of the pairwise LD between a subset
 of those \textit{k}-mers showed that most \textit{k}-mers formed a single LD block
-linked to the \textit{W1} locus (Additional file 1: Figure S4). A small LD block
+linked to the \textit{W1} locus (Additional file 1: Figure S5). A small LD block
 occurred on chromosome Gm09, however it was in moderate LD with the one on chromosome
@@
 causal variant at this locus. For all three GWAS approaches, we found a signal
-overlapping this gene (Additional file 1: Figure S5). In
+overlapping this gene (Additional file 1: Figure S6). In
 particular, the regions defined by the 1\% most associated variants
 identified by the SNP- and \textit{k}-mer-based approaches included
-Glyma.06g202300 (Additional file 1: Figure S6). While the most
+Glyma.06g202300 (Additional file 1: Figure S7). While the most
 significant \textit{k}-mers did occur within the sequence of this gene, they
 did not correspond to the documented causal variant (Additional file 1: Figure
-S7d).  Still, the fourth and fifth most strongly
+S8d).  Still, the fourth and fifth most strongly
 associated \textit{k}-mers corresponded to this causal variant (Additional file 1: Figures
-S7d and S8). We
+S8d and S9). We
 observed that not all accessions with gray pubescence harbored the deleted
-nucleotide (Additional file 1: Figure S8c), which could indicate
+nucleotide (Additional file 1: Figure S9c), which could indicate
 either that other causal variants are resulting in a non-functional F3’H in
@@
 causal variant, however it was not the most significantly associated variant
-overall (Additional file 1: Figures S6b and
-S7b). SV-based GWAS did not detect the causal variant
+overall (Additional file 1: Figures S7b and
+S8b). SV-based GWAS did not detect the causal variant
 at this locus, as expected from its size.
@@
 identified a signal that overlapped Glyma.03g258700 (Figure
-\ref{pubescence-color-main-figure}a-b, and Additional file 1: Figures S9 and
-S10).  While the most significant
+\ref{pubescence-color-main-figure}a-b, and Additional file 1: Figures S10 and
+S11).  While the most significant
 \textit{k}-mer identified the causal variant documented at this locus
@@
 the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d, Additional file 1: Figure 
-S11).  Interestingly, some of the accessions
+S12).  Interestingly, some of the accessions
 bearing the haplotype associated with tawny pubescence displayed light tawny
@@
 \citep{tuteja2009}. While we did find signals overlapping this causal variant
-using all three GWAS approaches (Additional file 1: Figures S12 and
-S13), none of the approaches identified
+using all three GWAS approaches (Additional file 1: Figures S13 and
+S14), none of the approaches identified
 it. This failure to detect the causal variant is likely due
@@
 analysis found strong signals overlapping the \emph{G} locus using all three
-approaches (Additional file 1: Figure S14).
+approaches (Additional file 1: Figure S15).
 Wang et al. \cite{wang2018} identified Glyma.01g198500 as the gene associated with
@@
 \ref{seed-coat-color-main-figure} and Additional file 1: Figures
-S15 and
-S16). Unsurprisingly, SV-based GWAS
+S16 and
+S17). Unsurprisingly, SV-based GWAS
 did not identify the causal variant given its type.
@@
 most highly associated markers (top 1\%) included the causal gene (Additional file 1: Figures
-S17 and S18).
+S18 and S19).
 We did identify one of the SNPs suggested as putative causal variants for this
@@
 12\textsuperscript{th} most associated \emph{k}-mer (Additional file 1: Figures
-S19 and S20).
+S20 and S21).
 Our results are consistent with previous findings in suggesting that there is
@@
 cultivars showed a determinate phenotype without bearing the allele associated
-with \emph{dt1} (Additional file 1: Figure S20c). The existence of
+with \emph{dt1} (Additional file 1: Figure S21c). The existence of
 several causal variants might explain why our analyses did not identify a
@@
 Glyma.09g235100 gene associated with \textit{R} (Additional file 1: Figures
-S21 and S22).
+S22 and S23).
 Gillman et al. \cite{gillman2011} identified this R2R3 MYB transcription factor as the
@@
 were not located in the body of the gene (Additional file 1: Figure
-S22).  Interestingly, the most significantly
+S23).  Interestingly, the most significantly
 associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
@@
 previously documented by Gillman et al. \cite{gillman2011} in our dataset (Additional file 1: Figures
-S23 and S24).
+S24 and S25).
 
@@
 detected signals at the \textit{Ps} locus using all three approaches (Figure
-\ref{pubescence-density-main-figure}a-b and Additional file 1: Figures S25, S26
-and S27). The causal
+\ref{pubescence-density-main-figure}a-b and Additional file 1: Figures S26, S27
+and S28). The causal
 variant at this recently cloned locus is a copy number variant (CNV) overlapping the
@@
 containing HPS domains and therefore likely represented the causal variant
-(Additional file 1: Figures S28 and
-S29).  We found signals overlapping this
+(Additional file 1: Figures S29 and
+S30).  We found signals overlapping this
 CNV using all approaches, and the most significantly associated \textit{k}-mer
 was notably located within the CNV region (Additional file 1: Figures
-S28 and
-S30). However, we were not able to
+S29 and
+S31). However, we were not able to
 directly link any of the variants or \textit{k}-mers identified to the causal
@@
 analyses based on \textit{k}-mers and SNPs found signals overlapping 14 and
-6 genes, respectively (Additional file 1: S31 and
-S32). Of these, a MATE transporter (Glyma.19g120300)
+6 genes, respectively (Additional file 1: S32 and
+S33). Of these, a MATE transporter (Glyma.19g120300)
 appears as a prime candidate as the most significant \emph{k}-mer was observed
@@
 locus overlap 24 and 50 genes, respectively (Additional file 1: Figures
-S31 and S33). Despite this
+S32 and S34). Despite this
 somewhat high number, we noticed that the closest gene to the most significant
@@
 signals found by all three approaches were located near this gene (Figure
-\ref{pubescence-form-main-figure}a-b, and Additional file 1: Figures S34 and
-S35).  The analyses based on SNPs and
+\ref{pubescence-form-main-figure}a-b, and Additional file 1: Figures S35 and
+S36).  The analyses based on SNPs and
 \emph{k}-mers identified two highly significant non-synonymous SNPs in
 Glyma.12g213900 that were associated with appressed pubescence (Figure
-\ref{pubescence-form-main-figure}c-d and Additional file 1: Figures S36 and
-S37). The alternative alleles for these two
+\ref{pubescence-form-main-figure}c-d and Additional file 1: Figures S37 and
+S38). The alternative alleles for these two
 SNPs co-occurred together in our dataset and may represent causal variants at
@@
 weaker and the signal observed from the \emph{k}-mer GWAS overlapped a total of
-20 genes (Additional file 1: Figures S34 and
-S38), none of which appeared obviously linked
+20 genes (Additional file 1: Figures S35 and
+S39), none of which appeared obviously linked
 to pubescence form.
@@
 this analysis only detected weak signals at the \emph{Pa1} locus (Additional file 1: Figure
-S39) and its results were therefore not
+S40) and its results were therefore not
 analyzed further.
@@
 overlapping the genomic region reported by these authors on Gm15 using all three GWAS
-approaches (Additional file 1: Figure S40). de Ronne et al. \cite{deronne2022}
+approaches (Additional file 1: Figure S41). de Ronne et al. \cite{deronne2022}
 proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
@@
 in a region with markers or \emph{k}-mers showing a weaker
-association (Additional file 1: Figure S41). The
+association (Additional file 1: Figure S42). The
 large size of the observed signal means that the gene associated with the phenotype could
@@
 most significant \emph{k}-mer (Additional file 1: Figure
-S41). One of the genes in this region is
+S42). One of the genes in this region is
 a calcium ion binding protein with an EF-hand domain (Glyma.15g217700).
@@
 GWAS analysis comparing accessions with black and brown pods found a previously
-undocumented signal on chromosome Gm15 (Additional file 1: Figure S42).
+undocumented signal on chromosome Gm15 (Additional file 1: Figure S43).
 \textit{k}-mers at this new locus
@@
 therefore represent a distinct locus (Additional file 1: Figure 
-S43). The observed signal overlaps a total of 27 genes
-(Additional file 1: Figure S44), but none of those are annotated
+S44). The observed signal overlaps a total of 27 genes
+(Additional file 1: Figure S45), but none of those are annotated
 as having an obvious role in controlling pod color.
@@
 methods to count-based methods.
-
-\section*{Methods}
-
-\subsection*{Sample selection and processing of sequencing data}
-
-We selected 389 inbred \emph{G. max} accessions based on the availability of
-Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
-sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
-Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
-and extracted paired-end reads using the fastq-dump command v. 2.9.6
-\cite{sratoolkit} with the
-option \texttt{-\-split-3}.
-Reads that were not paired in the raw sequencing data were not used
-for downstream analyses.  The runs downloaded from the SRA and associated
-metadata are listed in Additional file 2.
-
-We first filtered the sequencing data for quality and the presence of sequencing
-adapters using bbduk from BBtools v. 38.25
-\cite{bbtools}. Reads
-were then mapped to the reference assembly version 4 of \emph{G. max}
-cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
-and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
-bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
-reads in paired-end mode. Reads that were left unpaired following adapter- and
-quality-trimming were aligned separately in single-end mode with bwa-mem and
-merged with mapped paired-end reads using samtools \texttt{merge}
-v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
-individual SRA runs to the BAM files and merged the reads belonging to the same
-accession using bamaddrg
-\cite{bamaddrg}. We used the
-resulting BAM files for all downstream analyses requiring aligned reads.
-Metadata on mapping depth of aligned data can be found in Additional file 3.
-
-\subsection*{Discovery and genotyping of SVs}
-
-Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
-The \emph{discovery} step identified a set of candidate SVs using various methods,
-whereas the \emph{genotyping} step determined the genotype
-of the candidate SVs for all accessions in the population from the mapped Illumina reads. We used three different
-approaches in the SV discovery step to generate a set of candidate variants:
-
-\begin{itemize}
-	\item We used Illumina WGS data from the 389 accessions mentioned above
-		to call SVs in the population following a pipeline outlined
-		by Lemay et al. \cite{lemay2022}. In brief, we combined information from
-		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
-		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
-		a set of candidate SVs from the mapped Illumina data.
-	\item We used SVs discovered by Lemay et al. \cite{lemay2022} from Oxford Nanopore
-		sequencing data of 17 Canadian soybean cultivars.
-	\item We called SVs from 26 high-quality genome assemblies published by
-		Liu et al. \cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
-		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
-		We used methods similar to those of Liu et al. \cite{liu2020-pangenome} to call
-		SVs from the comparison of these assemblies to that of Williams82.
-\end{itemize}
-
-The SVs called using each of these approaches were merged using SVmerge
-\citep{wong2010} and the resulting set of candidate SVs was used for
-genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
-from the mapped Illumina reads for all accessions. We filtered the genotyped
-SVs by setting the genotype calls made from less than two reads to missing and
-by removing variants with a minor allele frequency < 0.02 or a proportion of
-missing data > 0.5. The resulting dataset of 186,306 genotyped SVs was used for
-GWAS. Detailed methods regarding the SV discovery and genotyping steps can be
-found in Additional file 1.
-
-\subsection*{Discovery and genotyping of SNPs and indels}
-
-We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
-mapped reads of the 389 accessions. The 21.1 M SNPs and
-indels from the raw output of Platypus were filtered to keep only those with:
-
-\begin{enumerate}
-	\item \texttt{FILTER} field set to \texttt{PASS}
-	\item minor allele frequency (MAF) $\geq$ 0.02
-	\item fraction of missing genotypes $\leq$ 0.5
-	\item heterozygosity rate $\leq$ 0.2
-\end{enumerate}
-
-Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
-PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
-step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
-implemented to reduce the computational requirements of the GWAS analysis by
-removing co-segregating markers. Markers located on unanchored scaffolds were
-also removed. The resulting dataset was used for GWAS and comprised 773,060
-SNPs and 151,570 indels.
-
-\subsection*{Computing the presence/absence table of \emph{k}-mers}
-
-We used the approach outlined by Voichek and Weigel \cite{voichek2020} for \emph{k}-mer-based GWAS
-and their recommendations on the
-associated GitHub page (v. 0.2-beta, \url{https://github.com/voichek/kmersGWAS}) to generate
-a presence/absence \emph{k}-mer table for use in GWAS.  The first step of
-the approach involved counting all \emph{k}-mers of length 31 (31-mers) present
-in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
-\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
-canonized form (the first of either the observed \textit{k}-mer or its reverse
-complement in lexicographical order) and the second time using the observed
-\emph{k}-mers themselves (non-canonized form). The two sets of \textit{k}-mers
-were then combined and filtered by
-keeping only those seen in at least 5 accessions, and observed in both canonized
-and non-canonized form in at least 20\% of the accessions in which they are found.
-We then generated a table indicating the presence or absence of each
-\emph{k}-mer in all accessions. This table was used to compute a kinship
-table for use in GWAS and was also used directly as input genotypes for the
-GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
-computation of kinship and GWAS.
-
-\subsection*{Phenotypic data}
-
-We analyzed ten qualitative traits previously studied by Bandillo et al. \cite{bandillo2017}: flower
-color, pubescence color, seed coat color, stem termination type, hilum color,
-pod color, pubescence form, pubescence density, seed coat luster, and maturity
-group.  We obtained phenotypic data for these traits from the GRIN database
-(\url{https://npgsweb.ars-grin.gov/gringlobal/search}) by querying the database
-using the PI identifiers of the 389 accessions used.  Four out of the 389
-accessions did not match the GRIN database and were therefore not used for the
-analysis of qualitative traits. The phenotypes of qualitative traits were recoded
-to numerical values for GWAS analysis following methods similar to those of
-Bandillo et al. \cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
-using targeted subsets of the observed phenotypes in order to focus on specific
-loci. The phenotypes and numerical values used for all analyses as well as the
-number of observations in the dataset that we used for GWAS are listed in
-Tables S2 to S11 (Additional file 1).
-
-We analyzed three quantitative traits in addition to the ten qualitative traits
-mentioned above. Seed oil and protein content were retrieved from the GRIN
-database. In cases where more than one value was listed for a given accession,
-we computed the average of those values and used it for GWAS analyses. As a third quantitative trait, we
-used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
-weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
-Our analysis included 340 accessions whose phenotypic data
-was already published by de Ronne et al. \cite{deronne2022} as well as 49 additional accessions for
-which sequencing data was not yet available when their study was conducted.
-Because of ambiguity regarding the identity of one of the accessions (HN019), this accession
-was dropped from all GWAS analyses. Similarly to seed oil and
-protein content, the CDW values for some accessions for which more than one
-observation was available were averaged.
-
-Because of clear discrepancies between the observed haplotypes and the reported
-phenotypes for some simple traits such as flower color and seed coat color, we
-suspected possible errors in the identity of some of the accessions. To
-investigate this, we compared the SNP genotypes derived from the SoySNP50K
-array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
-90\% concordance between the two genotype datasets (see Additional file 1 for
-detailed methods). These samples were removed from the dataset in addition to another
-accession that exhibited an atypical GC content suggesting contamination.
-Following these filtering steps and the removal of HN019, 363 accessions
-remained for GWAS analyses. The concordance between WGS data and SoySNP50K
-genotypes is included in Additional file 3 and averaged 98.3\% among retained
-lines (Additional file 1: Figure S45). The phenotypic data used in this study can be found in Additional file 4.
-
-\subsection*{GWAS analyses}
-
-We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
-the SNP/indel and SV datasets described above. SVs and indels required some additional
-preprocessing steps to make variant
-representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
-variants as SNPs prior to converting the VCF files to Hapmap diploid format
-using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
-retrieval of the proper metadata for each variant after GWAS.
-
-We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
-algorithm for computing the kinship matrix. We used an MLM model for
-consistency with the statistical model used by the \textit{k}-mer-based
-approach.  We used a randomization approach to determine the 5\% family-wise
-error-rate threshold for each GWAS analysis as described by Voichek and Weigel \cite{voichek2020}.
-To compute this threshold, we permuted the phenotypic observations 100 times
-and computed a GWAS on these permuted phenotypes to obtain a distribution of
-top \emph{p}-values under the null hypothesis. The fifth most significant of
-the top 100 \emph{p}-values was used as the significance threshold for
-inference, i.e. markers with \emph{p}-values lower than this threshold were
-considered significant.
-
-We conducted GWAS based on \emph{k}-mers using the method developed by
-Voichek and Weigel \cite{voichek2020}. In brief, the
-analysis was conducted in two steps. The first step used an approximate model
-to identify 1 million potentially significant \emph{k}-mers. These were then used in
-the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
-in a \emph{p}-value for each \emph{k}-mer. We
-used the kinship computed from the \emph{k}-mers to correct for relatedness
-between accessions. The program automatically outputs a list of \emph{k}-mers
-that passed a 5\% family-wise error-rate threshold determined by a
-randomization approach, which we considered significant for the purposes of
-downstream analyses.
-
-\subsection*{Analysis of significant \emph{k}-mers}
-
-\emph{k}-mers are not intrinsically associated with a particular genomic region
-and a tailored analysis was therefore required to associate them with genomic
-coordinates. To do so, we queried the mapped reads (in BAM format) of all accessions for matches
-to significant \emph{k}-mers for a given trait. Because this is a
-computationally demanding operation (there were typically tens of thousands of
-significant \emph{k}-mers per trait), we developed a C program called katcher
-(\url{https://github.com/malemay/katcher}) that
-uses the htslib library \citep{bonfield2021} for efficient reading and
-writing of BAM files. katcher and its associated utilities allow for efficient
-retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
-of interest.
-
-Once we retrieved all the reads containing significant \emph{k}-mers, we next
-needed to gather the mapping information contained in those reads for the
-purposes of identifying contiguous signals and comparing them to annotations. To
-simplify the amount of information to process in downstream analyses, we
-considered only the most significant \emph{k}-mer contained in a given read and
-the position where this \emph{k}-mer was most often observed in
-the dataset. Moreover, we limited the analysis to reads with a minimum mapping
-quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
-least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
-from sequencing errors. All downstream analyses (Manhattan plots,
-identification of signals, analysis of genes located near significant
-\emph{k}-mers, etc.) used this processed set of significant \emph{k}-mers.
-
-\subsection*{Downstream analyses of significant association signals}
-
-We used contiguous regions of significant variants, which we called
-``signals'', as a basis to compare our results with those obtained by previous
-studies and to query the genome for candidate genes. We defined the boundaries
-of those signals by grouping any significant markers or \emph{k}-mers that
-occurred within 250 kb of one another.  We did consider the length of deletions
-or \emph{k}-mers in defining the distance between significant variants.
-
-For the GWAS analyses computed from SNP/indel genotype calls, we retrieved all
-markers that had initially been pruned in a region ranging from 50 kb
-downstream to 50 kb upstream of significant signals and computed their
-\emph{p}-values by running GAPIT using the same parameters as noted above. The
-\emph{p}-values thus computed on markers that were initially left out were not
-used for initial signal discovery, but provided finer resolution when analyzing
-the results of the SNP-based GWAS analyses. Indeed, some markers that had not
-been included in the original GWAS may be the causal variant or simply yield
-more significant \emph{p}-values than those initially used.
-
-To help with the interpretation of results and the identification of candidate
-genes, we defined regions of highly significant associations within each signal
-by delimiting a range of coordinates containing the top 5\% most significant
-associations (SVs) or 1\% most significant associations (SNPs/indels
-and \emph{k}-mers) found within the signal. In cases where such a fraction
-represented fewer than two markers or \emph{k}-mers, the highly significant
-region was simply delimited by the two most significant associations.
-
-Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
-Williams82, we converted the genomic coordinates of their signals to assembly version 4
-using the same method that we used for converting
-SoySNP50K positions (see Additional file 1). For the analysis of
-qualitative traits, we only compared our results to signals at named classical
-loci and ignored previously reported minor signals with \emph{p}-values that
-just barely made it above the significance threshold.
-
-\subsection*{Linking significant \emph{k}-mers to variants}
-
-We used the reads retrieved by katcher and an assembly-based approach for
-identifying the variants associated with a given set of \emph{k}-mers. In
-addition to the reads themselves, we also retrieved their paired reads even
-though they may not have matched a significant \emph{k}-mer in the first place.
-These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
-\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
-individually for each accession and then aligned to the Williams82 reference
-assembly using bwa-mem with default parameters.  Aligned assembled sequences
-that overlapped loci of interest were realigned using multiple sequence
-alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
-visualizing the haplotypes at these loci. The identification of variants
-supported by significant \textit{k}-mers was based entirely on visual analysis
-of haplotype alignments. We only considered haplotypes occurring at least five
-times in the population in such analyses in order to leave out potential
-assembly errors. In a single case (\emph{Dt1} locus), assembly at the locus
-succeeded for a very low number of accessions (26). In this case, we instead used
-consensus sequences from the reads aligned with bwa for comparing the
-haplotypes across accessions.
-
-\subsection*{Computation of LD between \emph{k}-mers}
-
-In some cases, the analysis of \emph{k}-mers revealed signals at previously
-undocumented loci or found a very large number of associations throughout the
-genome. To shed light on these situations, we computed the pairwise LD between
-\emph{k}-mers in order to identify \emph{k}-mers that co-segregated and
-likely corresponded to a single locus. Because of the sheer number of pairwise
-comparisons that needed to be made for some traits (> 20,000 significant
-\emph{k}-mers kept even after the filtering steps mentioned above), we limited
-the number of \emph{k}-mers used for LD calculation to 1,500. For traits where
-more than 1,500 \emph{k}-mers were found, we computed LD on a subsample of
-these by selecting the 500 \emph{k}-mers with the most significant
-\emph{p}-values and sampling the remaining 1000 \emph{k}-mers randomly. This
-random sampling was done with a probability inversely proportional to the
-number of \emph{k}-mers matching a given chromosome or scaffold such that
-\emph{k}-mers were sampled from the whole genome and not simply from the most
-significant locations.
-
-\subsection*{Software used}
-\label{sv-gwas-software-used}
-
-Unless otherwise noted, all analyses were conducted using R v. 4.2.0
-\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
-packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
-\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
-1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
-\citep{obenchain2014}.  We gathered several functions used for the downstream
-analysis of GWAS results into a package called gwask which is available on
-GitHub (\url{https://github.com/malemay/gwask}).
 

### 2023-04-04 additional_files/manuscript.tex
Change the word Additional to Supplemental to be concordant with The Plant Genome's standard

--- 2023-04-04 additional_files/manuscript.tex
@@
 
+\section*{Methods}
+
+\subsection*{Sample selection and processing of sequencing data}
+
+We selected 389 inbred \emph{G. max} accessions based on the availability of
+Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
+sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
+Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
+and extracted paired-end reads using the fastq-dump command v. 2.9.6
+\cite{sratoolkit} with the
+option \texttt{-\-split-3}.
+Reads that were not paired in the raw sequencing data were not used
+for downstream analyses.  The runs downloaded from the SRA and associated
+metadata are listed in Supplemental file 2.
+
+We first filtered the sequencing data for quality and the presence of sequencing
+adapters using bbduk from BBtools v. 38.25
+\cite{bbtools}. Reads
+were then mapped to the reference assembly version 4 of \emph{G. max}
+cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
+and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
+bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
+reads in paired-end mode. Reads that were left unpaired following adapter- and
+quality-trimming were aligned separately in single-end mode with bwa-mem and
+merged with mapped paired-end reads using samtools \texttt{merge}
+v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
+individual SRA runs to the BAM files and merged the reads belonging to the same
+accession using bamaddrg
+\cite{bamaddrg}. We used the
+resulting BAM files for all downstream analyses requiring aligned reads.
+Metadata on mapping depth of aligned data can be found in Supplemental file 3.
+
+\subsection*{Discovery and genotyping of SVs}
+
+Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
+The \emph{discovery} step identified a set of candidate SVs using various methods,
+whereas the \emph{genotyping} step determined the genotype
+of the candidate SVs for all accessions in the population from the mapped Illumina reads. We used three different
+approaches in the SV discovery step to generate a set of candidate variants:
+
+\begin{itemize}
+	\item We used Illumina WGS data from the 389 accessions mentioned above
+		to call SVs in the population following a pipeline outlined
+		by Lemay et al. \cite{lemay2022}. In brief, we combined information from
+		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
+		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
+		a set of candidate SVs from the mapped Illumina data.
+	\item We used SVs discovered by Lemay et al. \cite{lemay2022} from Oxford Nanopore
+		sequencing data of 17 Canadian soybean cultivars.
+	\item We called SVs from 26 high-quality genome assemblies published by
+		Liu et al. \cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
+		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
+		We used methods similar to those of Liu et al. \cite{liu2020-pangenome} to call
+		SVs from the comparison of these assemblies to that of Williams82.
+\end{itemize}
+
+The SVs called using each of these approaches were merged using SVmerge
+\citep{wong2010} and the resulting set of candidate SVs was used for
+genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
+from the mapped Illumina reads for all accessions. We filtered the genotyped
+SVs by setting the genotype calls made from less than two reads to missing and
+by removing variants with a minor allele frequency < 0.02 or a proportion of
+missing data > 0.5. The resulting dataset of 186,306 genotyped SVs was used for
+GWAS. Detailed methods regarding the SV discovery and genotyping steps can be
+found in Supplemental file 1.
+
+\subsection*{Discovery and genotyping of SNPs and indels}
+
+We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
+mapped reads of the 389 accessions. The 21.1 M SNPs and
+indels from the raw output of Platypus were filtered to keep only those with:
+
+\begin{enumerate}
+	\item \texttt{FILTER} field set to \texttt{PASS}
+	\item minor allele frequency (MAF) $\geq$ 0.02
+	\item fraction of missing genotypes $\leq$ 0.5
+	\item heterozygosity rate $\leq$ 0.2
+\end{enumerate}
+
+Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
+PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
+step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
+implemented to reduce the computational requirements of the GWAS analysis by
+removing co-segregating markers. Markers located on unanchored scaffolds were
+also removed. The resulting dataset was used for GWAS and comprised 773,060
+SNPs and 151,570 indels.
+
+\subsection*{Computing the presence/absence table of \emph{k}-mers}
+
+We used the approach outlined by Voichek and Weigel \cite{voichek2020} for \emph{k}-mer-based GWAS
+and their recommendations on the
+associated GitHub page (v. 0.2-beta, \url{https://github.com/voichek/kmersGWAS}) to generate
+a presence/absence \emph{k}-mer table for use in GWAS.  The first step of
+the approach involved counting all \emph{k}-mers of length 31 (31-mers) present
+in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
+\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
+canonized form (the first of either the observed \textit{k}-mer or its reverse
+complement in lexicographical order) and the second time using the observed
+\emph{k}-mers themselves (non-canonized form). The two sets of \textit{k}-mers
+were then combined and filtered by
+keeping only those seen in at least 5 accessions, and observed in both canonized
+and non-canonized form in at least 20\% of the accessions in which they are found.
+We then generated a table indicating the presence or absence of each
+\emph{k}-mer in all accessions. This table was used to compute a kinship
+table for use in GWAS and was also used directly as input genotypes for the
+GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
+computation of kinship and GWAS.
+
+\subsection*{Phenotypic data}
+
+We analyzed ten qualitative traits previously studied by Bandillo et al. \cite{bandillo2017}: flower
+color, pubescence color, seed coat color, stem termination type, hilum color,
+pod color, pubescence form, pubescence density, seed coat luster, and maturity
+group.  We obtained phenotypic data for these traits from the GRIN database
+(\url{https://npgsweb.ars-grin.gov/gringlobal/search}) by querying the database
+using the PI identifiers of the 389 accessions used.  Four out of the 389
+accessions did not match the GRIN database and were therefore not used for the
+analysis of qualitative traits. The phenotypes of qualitative traits were recoded
+to numerical values for GWAS analysis following methods similar to those of
+Bandillo et al. \cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
+using targeted subsets of the observed phenotypes in order to focus on specific
+loci. The phenotypes and numerical values used for all analyses as well as the
+number of observations in the dataset that we used for GWAS are listed in
+Supplemental Tables S1 to S10.
+
+We analyzed three quantitative traits in addition to the ten qualitative traits
+mentioned above. Seed oil and protein content were retrieved from the GRIN
+database. In cases where more than one value was listed for a given accession,
+we computed the average of those values and used it for GWAS analyses. As a third quantitative trait, we
+used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
+weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
+Our analysis included 340 accessions whose phenotypic data
+was already published by de Ronne et al. \cite{deronne2022} as well as 49 additional accessions for
+which sequencing data was not yet available when their study was conducted.
+Because of ambiguity regarding the identity of one of the accessions (HN019), this accession
+was dropped from all GWAS analyses. Similarly to seed oil and
+protein content, the CDW values for some accessions for which more than one
+observation was available were averaged.
+
+Because of clear discrepancies between the observed haplotypes and the reported
+phenotypes for some simple traits such as flower color and seed coat color, we
+suspected possible errors in the identity of some of the accessions. To
+investigate this, we compared the SNP genotypes derived from the SoySNP50K
+array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
+90\% concordance between the two genotype datasets (see Supplemental file 1 for
+detailed methods). These samples were removed from the dataset in addition to another
+accession that exhibited an atypical GC content suggesting contamination.
+Following these filtering steps and the removal of HN019, 363 accessions
+remained for GWAS analyses. The concordance between WGS data and SoySNP50K
+genotypes is included in Supplemental file 3 and averaged 98.3\% among retained
+lines (Supplemental Figure S1). The phenotypic data used in this study can be found in Supplemental file 4.
+
+\subsection*{GWAS analyses}
+
+We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
+the SNP/indel and SV datasets described above. SVs and indels required some additional
+preprocessing steps to make variant
+representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
+variants as SNPs prior to converting the VCF files to Hapmap diploid format
+using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
+retrieval of the proper metadata for each variant after GWAS.
+
+We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
+algorithm for computing the kinship matrix. We used an MLM model for
+consistency with the statistical model used by the \textit{k}-mer-based
+approach.  We used a randomization approach to determine the 5\% family-wise
+error-rate threshold for each GWAS analysis as described by Voichek and Weigel \cite{voichek2020}.
+To compute this threshold, we permuted the phenotypic observations 100 times
+and computed a GWAS on these permuted phenotypes to obtain a distribution of
+top \emph{p}-values under the null hypothesis. The fifth most significant of
+the top 100 \emph{p}-values was used as the significance threshold for
+inference, i.e. markers with \emph{p}-values lower than this threshold were
+considered significant.
+
+We conducted GWAS based on \emph{k}-mers using the method developed by
+Voichek and Weigel \cite{voichek2020}. In brief, the
+analysis was conducted in two steps. The first step used an approximate model
+to identify 1 million potentially significant \emph{k}-mers. These were then used in
+the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
+in a \emph{p}-value for each \emph{k}-mer. We
+used the kinship computed from the \emph{k}-mers to correct for relatedness
+between accessions. The program automatically outputs a list of \emph{k}-mers
+that passed a 5\% family-wise error-rate threshold determined by a
+randomization approach, which we considered significant for the purposes of
+downstream analyses.
+
+\subsection*{Analysis of significant \emph{k}-mers}
+
+\emph{k}-mers are not intrinsically associated with a particular genomic region
+and a tailored analysis was therefore required to associate them with genomic
+coordinates. To do so, we queried the mapped reads (in BAM format) of all accessions for matches
+to significant \emph{k}-mers for a given trait. Because this is a
+computationally demanding operation (there were typically tens of thousands of
+significant \emph{k}-mers per trait), we developed a C program called katcher
+(\url{https://github.com/malemay/katcher}) that
+uses the htslib library \citep{bonfield2021} for efficient reading and
+writing of BAM files. katcher and its associated utilities allow for efficient
+retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
+of interest.
+
+Once we retrieved all the reads containing significant \emph{k}-mers, we next
+needed to gather the mapping information contained in those reads for the
+purposes of identifying contiguous signals and comparing them to annotations. To
+simplify the amount of information to process in downstream analyses, we
+considered only the most significant \emph{k}-mer contained in a given read and
+the position where this \emph{k}-mer was most often observed in
+the dataset. Moreover, we limited the analysis to reads with a minimum mapping
+quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
+least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
+from sequencing errors. All downstream analyses (Manhattan plots,
+identification of signals, analysis of genes located near significant
+\emph{k}-mers, etc.) used this processed set of significant \emph{k}-mers.
+
+\subsection*{Downstream analyses of significant association signals}
+
+We used contiguous regions of significant variants, which we called
+``signals'', as a basis to compare our results with those obtained by previous
+studies and to query the genome for candidate genes. We defined the boundaries
+of those signals by grouping any significant markers or \emph{k}-mers that
+occurred within 250 kb of one another.  We did consider the length of deletions
+or \emph{k}-mers in defining the distance between significant variants.
+
+For the GWAS analyses computed from SNP/indel genotype calls, we retrieved all
+markers that had initially been pruned in a region ranging from 50 kb
+downstream to 50 kb upstream of significant signals and computed their
+\emph{p}-values by running GAPIT using the same parameters as noted above. The
+\emph{p}-values thus computed on markers that were initially left out were not
+used for initial signal discovery, but provided finer resolution when analyzing
+the results of the SNP-based GWAS analyses. Indeed, some markers that had not
+been included in the original GWAS may be the causal variant or simply yield
+more significant \emph{p}-values than those initially used.
+
+To help with the interpretation of results and the identification of candidate
+genes, we defined regions of highly significant associations within each signal
+by delimiting a range of coordinates containing the top 5\% most significant
+associations (SVs) or 1\% most significant associations (SNPs/indels
+and \emph{k}-mers) found within the signal. In cases where such a fraction
+represented fewer than two markers or \emph{k}-mers, the highly significant
+region was simply delimited by the two most significant associations.
+
+Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
+Williams82, we converted the genomic coordinates of their signals to assembly version 4
+using the same method that we used for converting
+SoySNP50K positions (see Supplemental file 1). For the analysis of
+qualitative traits, we only compared our results to signals at named classical
+loci and ignored previously reported minor signals with \emph{p}-values that
+just barely made it above the significance threshold.
+
+\subsection*{Linking significant \emph{k}-mers to variants}
+
+We used the reads retrieved by katcher and an assembly-based approach for
+identifying the variants associated with a given set of \emph{k}-mers. In
+addition to the reads themselves, we also retrieved their paired reads even
+though they may not have matched a significant \emph{k}-mer in the first place.
+These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
+\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
+individually for each accession and then aligned to the Williams82 reference
+assembly using bwa-mem with default parameters.  Aligned assembled sequences
+that overlapped loci of interest were realigned using multiple sequence
+alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
+visualizing the haplotypes at these loci. The identification of variants
+supported by significant \textit{k}-mers was based entirely on visual analysis
+of haplotype alignments. We only considered haplotypes occurring at least five
+times in the population in such analyses in order to leave out potential
+assembly errors. In a single case (\emph{Dt1} locus), assembly at the locus
+succeeded for a very low number of accessions (26). In this case, we instead used
+consensus sequences from the reads aligned with bwa for comparing the
+haplotypes across accessions.
+
+\subsection*{Computation of LD between \emph{k}-mers}
+
+In some cases, the analysis of \emph{k}-mers revealed signals at previously
+undocumented loci or found a very large number of associations throughout the
+genome. To shed light on these situations, we computed the pairwise LD between
+\emph{k}-mers in order to identify \emph{k}-mers that co-segregated and
+likely corresponded to a single locus. Because of the sheer number of pairwise
+comparisons that needed to be made for some traits (> 20,000 significant
+\emph{k}-mers kept even after the filtering steps mentioned above), we limited
+the number of \emph{k}-mers used for LD calculation to 1,500. For traits where
+more than 1,500 \emph{k}-mers were found, we computed LD on a subsample of
+these by selecting the 500 \emph{k}-mers with the most significant
+\emph{p}-values and sampling the remaining 1000 \emph{k}-mers randomly. This
+random sampling was done with a probability inversely proportional to the
+number of \emph{k}-mers matching a given chromosome or scaffold such that
+\emph{k}-mers were sampled from the whole genome and not simply from the most
+significant locations.
+
+\subsection*{Software used}
+\label{sv-gwas-software-used}
+
+Unless otherwise noted, all analyses were conducted using R v. 4.2.0
+\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
+packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
+\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
+1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
+\citep{obenchain2014}.  We gathered several functions used for the downstream
+analysis of GWAS results into a package called gwask which is available on
+GitHub (\url{https://github.com/malemay/gwask}).
+
 \section*{Results and discussion}
@@
 on the most noteworthy results. Readers interested in results at loci that are
-not discussed here are referred to Additional file 1. In particular, Table
-S1 (Additional file 1) provides a summary of the
+not discussed here are referred to Supplemental file 1. In particular, Supplemental Table
+S11 provides a summary of the
 \textit{p}-values obtained by each approach at each of the loci considered in
@@
 F3’5’H enzyme non-functional and prevents pigmentation of the flower. All
-GWAS approaches detected a signal overlapping the gene at this locus (Additional file 1: Figures
-S1 and S2). The most
+GWAS approaches detected a signal overlapping the gene at this locus (Supplemental Figures
+S2 and S3). The most
 significant SV and \textit{k}-mers corresponded to the known causal variant
-(Figure \ref{flower-color-main-figure}a-b and Additional file 1: S3c-d).
+(Figure \ref{flower-color-main-figure}a-b and Supplemental Figure S4c-d).
 In particular, the assembly of the reads containing the most significant \textit{k}-mers
@@
 and indels located within the gene sequence were among the most significantly
-associated ones (Additional file 1: Figure S3b).
+associated ones (Supplemental Figure S4b).
 
 The \textit{k}-mer GWAS analysis also identified presumably spurious significant
-\textit{k}-mers scattered throughout the genome (Additional file 1: Figure
-S1c). An analysis of the pairwise LD between a subset
+\textit{k}-mers scattered throughout the genome (Supplemental Figure
+S2c). An analysis of the pairwise LD between a subset
 of those \textit{k}-mers showed that most \textit{k}-mers formed a single LD block
-linked to the \textit{W1} locus (Additional file 1: Figure S4). A small LD block
+linked to the \textit{W1} locus (Supplemental Figure S5). A small LD block
 occurred on chromosome Gm09, however it was in moderate LD with the one on chromosome
@@
 causal variant at this locus. For all three GWAS approaches, we found a signal
-overlapping this gene (Additional file 1: Figure S5). In
+overlapping this gene (Supplemental Figure S6). In
 particular, the regions defined by the 1\% most associated variants
 identified by the SNP- and \textit{k}-mer-based approaches included
-Glyma.06g202300 (Additional file 1: Figure S6). While the most
+Glyma.06g202300 (Supplemental Figure S7). While the most
 significant \textit{k}-mers did occur within the sequence of this gene, they
-did not correspond to the documented causal variant (Additional file 1: Figure
-S7d).  Still, the fourth and fifth most strongly
-associated \textit{k}-mers corresponded to this causal variant (Additional file 1: Figures
-S7d and S8). We
+did not correspond to the documented causal variant (Supplemental Figure
+S8d).  Still, the fourth and fifth most strongly
+associated \textit{k}-mers corresponded to this causal variant (Supplemental Figures
+S8d and S9). We
 observed that not all accessions with gray pubescence harbored the deleted
-nucleotide (Additional file 1: Figure S8c), which could indicate
+nucleotide (Supplemental Figure S9c), which could indicate
 either that other causal variants are resulting in a non-functional F3’H in
@@
 causal variant, however it was not the most significantly associated variant
-overall (Additional file 1: Figures S6b and
-S7b). SV-based GWAS did not detect the causal variant
+overall (Supplemental Figures S7b and
+S8b). SV-based GWAS did not detect the causal variant
 at this locus, as expected from its size.
@@
 identified a signal that overlapped Glyma.03g258700 (Figure
-\ref{pubescence-color-main-figure}a-b, and Additional file 1: Figures S9 and
-S10).  While the most significant
+\ref{pubescence-color-main-figure}a-b, and Supplemental Figures S10 and
+S11).  While the most significant
 \textit{k}-mer identified the causal variant documented at this locus
 \citep{yan2020}, it was the fourth most significantly associated variant for
-the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d, Additional file 1: Figure 
-S11).  Interestingly, some of the accessions
+the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d, Supplemental Figure 
+S12).  Interestingly, some of the accessions
 bearing the haplotype associated with tawny pubescence displayed light tawny
@@
 \citep{tuteja2009}. While we did find signals overlapping this causal variant
-using all three GWAS approaches (Additional file 1: Figures S12 and
-S13), none of the approaches identified
+using all three GWAS approaches (Supplemental Figures S13 and
+S14), none of the approaches identified
 it. This failure to detect the causal variant is likely due
@@
 analysis found strong signals overlapping the \emph{G} locus using all three
-approaches (Additional file 1: Figure S14).
+approaches (Supplemental Figure S15).
 Wang et al. \cite{wang2018} identified Glyma.01g198500 as the gene associated with
@@
 SNP-based analysis as corresponding to that causal variant (Figure
-\ref{seed-coat-color-main-figure} and Additional file 1: Figures
-S15 and
-S16). Unsurprisingly, SV-based GWAS
+\ref{seed-coat-color-main-figure} and Supplemental Figures
+S16 and
+S17). Unsurprisingly, SV-based GWAS
 did not identify the causal variant given its type.
@@
 \textit{k}-mer-based analyses performed best since the region defined by their
-most highly associated markers (top 1\%) included the causal gene (Additional file 1: Figures
-S17 and S18).
+most highly associated markers (top 1\%) included the causal gene (Supplemental Figures
+S18 and S19).
 We did identify one of the SNPs suggested as putative causal variants for this
@@
 135\textsuperscript{th} most associated marker found by Platypus and the
-12\textsuperscript{th} most associated \emph{k}-mer (Additional file 1: Figures
-S19 and S20).
+12\textsuperscript{th} most associated \emph{k}-mer (Supplemental Figures
+S20 and S21).
 Our results are consistent with previous findings in suggesting that there is
@@
 cultivars showed a determinate phenotype without bearing the allele associated
-with \emph{dt1} (Additional file 1: Figure S20c). The existence of
+with \emph{dt1} (Supplemental Figure S21c). The existence of
 several causal variants might explain why our analyses did not identify a
@@
 only SNP- and \textit{k}-mer-based GWAS detected signals overlapping the
-Glyma.09g235100 gene associated with \textit{R} (Additional file 1: Figures
-S21 and S22).
+Glyma.09g235100 gene associated with \textit{R} (Supplemental Figures
+S22 and S23).
 Gillman et al. \cite{gillman2011} identified this R2R3 MYB transcription factor as the
@@
 detect strong signals in the vicinity of this gene, the most associated markers
-were not located in the body of the gene (Additional file 1: Figure
-S22).  Interestingly, the most significantly
+were not located in the body of the gene (Supplemental Figure
+S23).  Interestingly, the most significantly
 associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
@@
 loss of function at the \emph{R} locus.  Still, we observed two causal variants
-previously documented by Gillman et al. \cite{gillman2011} in our dataset (Additional file 1: Figures
-S23 and S24).
+previously documented by Gillman et al. \cite{gillman2011} in our dataset (Supplemental Figures
+S24 and S25).
 
@@
 detected signals at the \textit{Ps} locus using all three approaches (Figure
-\ref{pubescence-density-main-figure}a-b and Additional file 1: Figures S25, S26
-and S27). The causal
+\ref{pubescence-density-main-figure}a-b and Supplemental Figures S26, S27
+and S28). The causal
 variant at this recently cloned locus is a copy number variant (CNV) overlapping the
@@
 containing HPS domains and therefore likely represented the causal variant
-(Additional file 1: Figures S28 and
-S29).  We found signals overlapping this
+(Supplemental Figures S29 and
+S30).  We found signals overlapping this
 CNV using all approaches, and the most significantly associated \textit{k}-mer
-was notably located within the CNV region (Additional file 1: Figures
-S28 and
-S30). However, we were not able to
+was notably located within the CNV region (Supplemental Figures
+S29 and
+S31). However, we were not able to
 directly link any of the variants or \textit{k}-mers identified to the causal
@@
 analyses based on \textit{k}-mers and SNPs found signals overlapping 14 and
-6 genes, respectively (Additional file 1: S31 and
-S32). Of these, a MATE transporter (Glyma.19g120300)
+6 genes, respectively (Supplemental figures S32 and
+S33). Of these, a MATE transporter (Glyma.19g120300)
 appears as a prime candidate as the most significant \emph{k}-mer was observed
@@
 The signals found by SNP- and \emph{k}-mer-based analyses at the \textit{L2}
-locus overlap 24 and 50 genes, respectively (Additional file 1: Figures
-S31 and S33). Despite this
+locus overlap 24 and 50 genes, respectively (Supplemental Figures
+S32 and S34). Despite this
 somewhat high number, we noticed that the closest gene to the most significant
@@
 signals found by all three approaches were located near this gene (Figure
-\ref{pubescence-form-main-figure}a-b, and Additional file 1: Figures S34 and
-S35).  The analyses based on SNPs and
+\ref{pubescence-form-main-figure}a-b, and Supplemental Figures S35 and
+S36).  The analyses based on SNPs and
 \emph{k}-mers identified two highly significant non-synonymous SNPs in
 Glyma.12g213900 that were associated with appressed pubescence (Figure
-\ref{pubescence-form-main-figure}c-d and Additional file 1: Figures S36 and
-S37). The alternative alleles for these two
+\ref{pubescence-form-main-figure}c-d and Supplemental Figures S37 and
+S38). The alternative alleles for these two
 SNPs co-occurred together in our dataset and may represent causal variants at
@@
 weaker and the signal observed from the \emph{k}-mer GWAS overlapped a total of
-20 genes (Additional file 1: Figures S34 and
-S38), none of which appeared obviously linked
+20 genes (Supplemental Figures S35 and
+S39), none of which appeared obviously linked
 to pubescence form.
@@
 with the objective of enhancing the signal at the \emph{Pa2} locus. However,
-this analysis only detected weak signals at the \emph{Pa1} locus (Additional file 1: Figure
-S39) and its results were therefore not
+this analysis only detected weak signals at the \emph{Pa1} locus (Supplemental Figure
+S40) and its results were therefore not
 analyzed further.
@@
 overlapping the genomic region reported by these authors on Gm15 using all three GWAS
-approaches (Additional file 1: Figure S40). de Ronne et al. \cite{deronne2022}
+approaches (Supplemental Figure S41). de Ronne et al. \cite{deronne2022}
 proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
@@
 in a region with markers or \emph{k}-mers showing a weaker
-association (Additional file 1: Figure S41). The
+association (Supplemental Figure S42). The
 large size of the observed signal means that the gene associated with the phenotype could
@@
 analysis. This region overlaps only three annotated genes and also contains the
-most significant \emph{k}-mer (Additional file 1: Figure
-S41). One of the genes in this region is
+most significant \emph{k}-mer (Supplemental Figure
+S42). One of the genes in this region is
 a calcium ion binding protein with an EF-hand domain (Glyma.15g217700).
@@
 GWAS analysis comparing accessions with black and brown pods found a previously
-undocumented signal on chromosome Gm15 (Additional file 1: Figure S42).
+undocumented signal on chromosome Gm15 (Supplemental Figure S43).
 \textit{k}-mers at this new locus
 showed no LD with \textit{k}-mers at the known \textit{L1} locus and may
-therefore represent a distinct locus (Additional file 1: Figure 
-S43). The observed signal overlaps a total of 27 genes
-(Additional file 1: Figure S44), but none of those are annotated
+therefore represent a distinct locus (Supplemental Figure 
+S44). The observed signal overlaps a total of 27 genes
+(Supplemental Figure S45), but none of those are annotated
 as having an obvious role in controlling pod color.
@@
 
-\section*{Methods}
-
-\subsection*{Sample selection and processing of sequencing data}
-
-We selected 389 inbred \emph{G. max} accessions based on the availability of
-Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
-sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
-Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
-and extracted paired-end reads using the fastq-dump command v. 2.9.6
-\cite{sratoolkit} with the
-option \texttt{-\-split-3}.
-Reads that were not paired in the raw sequencing data were not used
-for downstream analyses.  The runs downloaded from the SRA and associated
-metadata are listed in Additional file 2.
-
-We first filtered the sequencing data for quality and the presence of sequencing
-adapters using bbduk from BBtools v. 38.25
-\cite{bbtools}. Reads
-were then mapped to the reference assembly version 4 of \emph{G. max}
-cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
-and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
-bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
-reads in paired-end mode. Reads that were left unpaired following adapter- and
-quality-trimming were aligned separately in single-end mode with bwa-mem and
-merged with mapped paired-end reads using samtools \texttt{merge}
-v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
-individual SRA runs to the BAM files and merged the reads belonging to the same
-accession using bamaddrg
-\cite{bamaddrg}. We used the
-resulting BAM files for all downstream analyses requiring aligned reads.
-Metadata on mapping depth of aligned data can be found in Additional file 3.
-
-\subsection*{Discovery and genotyping of SVs}
-
-Our SV analysis pipeline consisted of separate SV discovery and genotyping steps.
-The \emph{discovery} step identified a set of candidate SVs using various methods,
-whereas the \emph{genotyping} step determined the genotype
-of the candidate SVs for all accessions in the population from the mapped Illumina reads. We used three different
-approaches in the SV discovery step to generate a set of candidate variants:
-
-\begin{itemize}
-	\item We used Illumina WGS data from the 389 accessions mentioned above
-		to call SVs in the population following a pipeline outlined
-		by Lemay et al. \cite{lemay2022}. In brief, we combined information from
-		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
-		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
-		a set of candidate SVs from the mapped Illumina data.
-	\item We used SVs discovered by Lemay et al. \cite{lemay2022} from Oxford Nanopore
-		sequencing data of 17 Canadian soybean cultivars.
-	\item We called SVs from 26 high-quality genome assemblies published by
-		Liu et al. \cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
-		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
-		We used methods similar to those of Liu et al. \cite{liu2020-pangenome} to call
-		SVs from the comparison of these assemblies to that of Williams82.
-\end{itemize}
-
-The SVs called using each of these approaches were merged using SVmerge
-\citep{wong2010} and the resulting set of candidate SVs was used for
-genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
-from the mapped Illumina reads for all accessions. We filtered the genotyped
-SVs by setting the genotype calls made from less than two reads to missing and
-by removing variants with a minor allele frequency < 0.02 or a proportion of
-missing data > 0.5. The resulting dataset of 186,306 genotyped SVs was used for
-GWAS. Detailed methods regarding the SV discovery and genotyping steps can be
-found in Additional file 1.
-
-\subsection*{Discovery and genotyping of SNPs and indels}
-
-We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
-mapped reads of the 389 accessions. The 21.1 M SNPs and
-indels from the raw output of Platypus were filtered to keep only those with:
-
-\begin{enumerate}
-	\item \texttt{FILTER} field set to \texttt{PASS}
-	\item minor allele frequency (MAF) $\geq$ 0.02
-	\item fraction of missing genotypes $\leq$ 0.5
-	\item heterozygosity rate $\leq$ 0.2
-\end{enumerate}
-
-Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
-PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
-step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
-implemented to reduce the computational requirements of the GWAS analysis by
-removing co-segregating markers. Markers located on unanchored scaffolds were
-also removed. The resulting dataset was used for GWAS and comprised 773,060
-SNPs and 151,570 indels.
-
-\subsection*{Computing the presence/absence table of \emph{k}-mers}
-
-We used the approach outlined by Voichek and Weigel \cite{voichek2020} for \emph{k}-mer-based GWAS
-and their recommendations on the
-associated GitHub page (v. 0.2-beta, \url{https://github.com/voichek/kmersGWAS}) to generate
-a presence/absence \emph{k}-mer table for use in GWAS.  The first step of
-the approach involved counting all \emph{k}-mers of length 31 (31-mers) present
-in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
-\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
-canonized form (the first of either the observed \textit{k}-mer or its reverse
-complement in lexicographical order) and the second time using the observed
-\emph{k}-mers themselves (non-canonized form). The two sets of \textit{k}-mers
-were then combined and filtered by
-keeping only those seen in at least 5 accessions, and observed in both canonized
-and non-canonized form in at least 20\% of the accessions in which they are found.
-We then generated a table indicating the presence or absence of each
-\emph{k}-mer in all accessions. This table was used to compute a kinship
-table for use in GWAS and was also used directly as input genotypes for the
-GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
-computation of kinship and GWAS.
-
-\subsection*{Phenotypic data}
-
-We analyzed ten qualitative traits previously studied by Bandillo et al. \cite{bandillo2017}: flower
-color, pubescence color, seed coat color, stem termination type, hilum color,
-pod color, pubescence form, pubescence density, seed coat luster, and maturity
-group.  We obtained phenotypic data for these traits from the GRIN database
-(\url{https://npgsweb.ars-grin.gov/gringlobal/search}) by querying the database
-using the PI identifiers of the 389 accessions used.  Four out of the 389
-accessions did not match the GRIN database and were therefore not used for the
-analysis of qualitative traits. The phenotypes of qualitative traits were recoded
-to numerical values for GWAS analysis following methods similar to those of
-Bandillo et al. \cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
-using targeted subsets of the observed phenotypes in order to focus on specific
-loci. The phenotypes and numerical values used for all analyses as well as the
-number of observations in the dataset that we used for GWAS are listed in
-Tables S2 to S11 (Additional file 1).
-
-We analyzed three quantitative traits in addition to the ten qualitative traits
-mentioned above. Seed oil and protein content were retrieved from the GRIN
-database. In cases where more than one value was listed for a given accession,
-we computed the average of those values and used it for GWAS analyses. As a third quantitative trait, we
-used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
-weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
-Our analysis included 340 accessions whose phenotypic data
-was already published by de Ronne et al. \cite{deronne2022} as well as 49 additional accessions for
-which sequencing data was not yet available when their study was conducted.
-Because of ambiguity regarding the identity of one of the accessions (HN019), this accession
-was dropped from all GWAS analyses. Similarly to seed oil and
-protein content, the CDW values for some accessions for which more than one
-observation was available were averaged.
-
-Because of clear discrepancies between the observed haplotypes and the reported
-phenotypes for some simple traits such as flower color and seed coat color, we
-suspected possible errors in the identity of some of the accessions. To
-investigate this, we compared the SNP genotypes derived from the SoySNP50K
-array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
-90\% concordance between the two genotype datasets (see Additional file 1 for
-detailed methods). These samples were removed from the dataset in addition to another
-accession that exhibited an atypical GC content suggesting contamination.
-Following these filtering steps and the removal of HN019, 363 accessions
-remained for GWAS analyses. The concordance between WGS data and SoySNP50K
-genotypes is included in Additional file 3 and averaged 98.3\% among retained
-lines (Additional file 1: Figure S45). The phenotypic data used in this study can be found in Additional file 4.
-
-\subsection*{GWAS analyses}
-
-We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
-the SNP/indel and SV datasets described above. SVs and indels required some additional
-preprocessing steps to make variant
-representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
-variants as SNPs prior to converting the VCF files to Hapmap diploid format
-using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
-retrieval of the proper metadata for each variant after GWAS.
-
-We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
-algorithm for computing the kinship matrix. We used an MLM model for
-consistency with the statistical model used by the \textit{k}-mer-based
-approach.  We used a randomization approach to determine the 5\% family-wise
-error-rate threshold for each GWAS analysis as described by Voichek and Weigel \cite{voichek2020}.
-To compute this threshold, we permuted the phenotypic observations 100 times
-and computed a GWAS on these permuted phenotypes to obtain a distribution of
-top \emph{p}-values under the null hypothesis. The fifth most significant of
-the top 100 \emph{p}-values was used as the significance threshold for
-inference, i.e. markers with \emph{p}-values lower than this threshold were
-considered significant.
-
-We conducted GWAS based on \emph{k}-mers using the method developed by
-Voichek and Weigel \cite{voichek2020}. In brief, the
-analysis was conducted in two steps. The first step used an approximate model
-to identify 1 million potentially significant \emph{k}-mers. These were then used in
-the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
-in a \emph{p}-value for each \emph{k}-mer. We
-used the kinship computed from the \emph{k}-mers to correct for relatedness
-between accessions. The program automatically outputs a list of \emph{k}-mers
-that passed a 5\% family-wise error-rate threshold determined by a
-randomization approach, which we considered significant for the purposes of
-downstream analyses.
-
-\subsection*{Analysis of significant \emph{k}-mers}
-
-\emph{k}-mers are not intrinsically associated with a particular genomic region
-and a tailored analysis was therefore required to associate them with genomic
-coordinates. To do so, we queried the mapped reads (in BAM format) of all accessions for matches
-to significant \emph{k}-mers for a given trait. Because this is a
-computationally demanding operation (there were typically tens of thousands of
-significant \emph{k}-mers per trait), we developed a C program called katcher
-(\url{https://github.com/malemay/katcher}) that
-uses the htslib library \citep{bonfield2021} for efficient reading and
-writing of BAM files. katcher and its associated utilities allow for efficient
-retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
-of interest.
-
-Once we retrieved all the reads containing significant \emph{k}-mers, we next
-needed to gather the mapping information contained in those reads for the
-purposes of identifying contiguous signals and comparing them to annotations. To
-simplify the amount of information to process in downstream analyses, we
-considered only the most significant \emph{k}-mer contained in a given read and
-the position where this \emph{k}-mer was most often observed in
-the dataset. Moreover, we limited the analysis to reads with a minimum mapping
-quality of 20 and to \emph{k}-mers that were observed at least 10 times in at
-least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
-from sequencing errors. All downstream analyses (Manhattan plots,
-identification of signals, analysis of genes located near significant
-\emph{k}-mers, etc.) used this processed set of significant \emph{k}-mers.
-
-\subsection*{Downstream analyses of significant association signals}
-
-We used contiguous regions of significant variants, which we called
-``signals'', as a basis to compare our results with those obtained by previous
-studies and to query the genome for candidate genes. We defined the boundaries
-of those signals by grouping any significant markers or \emph{k}-mers that
-occurred within 250 kb of one another.  We did consider the length of deletions
-or \emph{k}-mers in defining the distance between significant variants.
-
-For the GWAS analyses computed from SNP/indel genotype calls, we retrieved all
-markers that had initially been pruned in a region ranging from 50 kb
-downstream to 50 kb upstream of significant signals and computed their
-\emph{p}-values by running GAPIT using the same parameters as noted above. The
-\emph{p}-values thus computed on markers that were initially left out were not
-used for initial signal discovery, but provided finer resolution when analyzing
-the results of the SNP-based GWAS analyses. Indeed, some markers that had not
-been included in the original GWAS may be the causal variant or simply yield
-more significant \emph{p}-values than those initially used.
-
-To help with the interpretation of results and the identification of candidate
-genes, we defined regions of highly significant associations within each signal
-by delimiting a range of coordinates containing the top 5\% most significant
-associations (SVs) or 1\% most significant associations (SNPs/indels
-and \emph{k}-mers) found within the signal. In cases where such a fraction
-represented fewer than two markers or \emph{k}-mers, the highly significant
-region was simply delimited by the two most significant associations.
-
-Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
-Williams82, we converted the genomic coordinates of their signals to assembly version 4
-using the same method that we used for converting
-SoySNP50K positions (see Additional file 1). For the analysis of
-qualitative traits, we only compared our results to signals at named classical
-loci and ignored previously reported minor signals with \emph{p}-values that
-just barely made it above the significance threshold.
-
-\subsection*{Linking significant \emph{k}-mers to variants}
-
-We used the reads retrieved by katcher and an assembly-based approach for
-identifying the variants associated with a given set of \emph{k}-mers. In
-addition to the reads themselves, we also retrieved their paired reads even
-though they may not have matched a significant \emph{k}-mer in the first place.
-These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
-\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
-individually for each accession and then aligned to the Williams82 reference
-assembly using bwa-mem with default parameters.  Aligned assembled sequences
-that overlapped loci of interest were realigned using multiple sequence
-alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
-visualizing the haplotypes at these loci. The identification of variants
-supported by significant \textit{k}-mers was based entirely on visual analysis
-of haplotype alignments. We only considered haplotypes occurring at least five
-times in the population in such analyses in order to leave out potential
-assembly errors. In a single case (\emph{Dt1} locus), assembly at the locus
-succeeded for a very low number of accessions (26). In this case, we instead used
-consensus sequences from the reads aligned with bwa for comparing the
-haplotypes across accessions.
-
-\subsection*{Computation of LD between \emph{k}-mers}
-
-In some cases, the analysis of \emph{k}-mers revealed signals at previously
-undocumented loci or found a very large number of associations throughout the
-genome. To shed light on these situations, we computed the pairwise LD between
-\emph{k}-mers in order to identify \emph{k}-mers that co-segregated and
-likely corresponded to a single locus. Because of the sheer number of pairwise
-comparisons that needed to be made for some traits (> 20,000 significant
-\emph{k}-mers kept even after the filtering steps mentioned above), we limited
-the number of \emph{k}-mers used for LD calculation to 1,500. For traits where
-more than 1,500 \emph{k}-mers were found, we computed LD on a subsample of
-these by selecting the 500 \emph{k}-mers with the most significant
-\emph{p}-values and sampling the remaining 1000 \emph{k}-mers randomly. This
-random sampling was done with a probability inversely proportional to the
-number of \emph{k}-mers matching a given chromosome or scaffold such that
-\emph{k}-mers were sampled from the whole genome and not simply from the most
-significant locations.
-
-\subsection*{Software used}
-\label{sv-gwas-software-used}
-
-Unless otherwise noted, all analyses were conducted using R v. 4.2.0
-\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
-packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
-\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
-1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
-\citep{obenchain2014}.  We gathered several functions used for the downstream
-analysis of GWAS results into a package called gwask which is available on
-GitHub (\url{https://github.com/malemay/gwask}).
-
-\section*{Additional Files}
-
-\subsection*{Additional file 1 --- Additional file 1.pdf}
+\section*{Supplemental Files}
+
+\subsection*{Supplemental file 1 --- Supplemental file 1.pdf}
 Supplemental methods and results,
-including Tables S1 to S11
-and Figures S1 to S60 (PDF 12.9 MB)
-
-\subsection*{Additional file 2 --- Additional file 2.csv}
+including Supplemental Tables S1 to S11
+and Supplemental Figures S1 to S60 (PDF 12.9 MB)
+
+\subsection*{Supplemental file 2 --- Supplemental file 2.csv}
 Metadata on the sequencing runs retrieved
@@
 
-\subsection*{Additional file 3 --- Additional file 3.csv}
+\subsection*{Supplemental file 3 --- Supplemental file 3.csv}
 Metadata on individual accessions following mapping
@@
 
-\subsection*{Additional file 4 --- Additional file 4.csv}
+\subsection*{Supplemental file 4 --- Supplemental file 4.csv}
 Phenotypic dataset used for GWAS analyses (CSV 135 KB)
@@
 
-The Illumina sequencing data used for the analyses are listed in Additional file 2
+The Illumina sequencing data used for the analyses are listed in Supplemental file 2
 and available through NCBI BioProjects PRJNA257011 \cite{PRJNA257011}, PRJNA289660 \cite{PRJNA289660}
@@
 		\item[b] SVs genotyped by Paragraph
-		\item[c] Locus not discussed in main text. See Additional file 1 for detailed results.
+		\item[c] Locus not discussed in main text. See Supplemental file 1 for detailed results.
 		\item[]- The approach detected no signal overlapping the gene.

### 2023-04-04 additional_files/manuscript.tex
Add figures in the manuscript

@@
 \usepackage{pdflscape}
+\usepackage{graphicx}
+\usepackage{ccaption}
+\usepackage[margin=1in]{geometry} % leaves more space for text body
 
@@
 associated ones (Supplemental Figure S4b).
+
+% Figure 1
+\begin{figure}
+	\centering
+	\includegraphics[width = 5in]{flower_color_W1_main_figure.png}
+	\caption{(Caption on the following page.)}
+	\label{flower-color-main-figure}
+\end{figure}
+
+\begin{figure}
+	\contcaption{Results of SV- and \textit{k}-mer-based GWAS at the
+	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
+	threshold whereas vertical dotted lines indicate the location of the
+	\textit{W1} locus. (b) Zoomed-in Manhattan plots at the \textit{W1} locus.
+	The vertical dotted lines indicate the location of the
+	Glyma.13g072100 gene associated with this locus. The shaded gray
+	rectangles show the regions delimited by the top 5\% (SVs) and
+	1\% (\textit{k}-mers) associations.  (c) Gene model of Glyma.13g072100.
+	Rectangles represent exons, colored rectangles represent coding
+	sequences, and arrows indicate the direction of transcription. The red
+	rectangle highlights the region shown in panel (d). (d) Pairwise
+	alignment of the two haplotype sequences in the population at the
+	\textit{W1} locus. Nucleotides are
+	colored according to the \textit{p}-value of the most significant
+	\textit{k}-mer overlapping them. Vertical dotted lines between haplotypes
+	indicate sequence differences, whereas dashes indicate gaps in the
+	alignment. (e) Contingency table of the phenotypes and haplotypes found
+	in the population at the \textit{W1} locus.}
+\end{figure}
 
@@
 
+% Figure 2
+\begin{figure}
+	\centering
+	\includegraphics[width = 5in]{pubescence_color_nogray_Td_main_figure.png}
+	\caption{(Caption on the following page.)}
+	\label{pubescence-color-main-figure}
+\end{figure}
+
+\begin{figure}
+	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
+	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
+	threshold whereas vertical dotted lines indicate the location of the
+	\textit{Td} locus. (b) Zoomed-in Manhattan plots at the \textit{Td} locus.
+	The vertical dotted lines indicate the location of the
+	Glyma.03g258700 gene associated with this locus. The shaded gray
+	rectangles show the regions delimited by the top 1\% associations.
+	In the case of SNPs, blue points denote markers used in the
+	original analysis, whereas red points denote markers that had originally
+	been pruned but whose \emph{p}-values were computed after signal
+	discovery.
+	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
+	rectangles represent coding sequences, and the arrow indicates the
+	direction of transcription. The red rectangle highlights the region
+	shown in panel (d). (d) Pairwise alignment of the two haplotype
+	sequences in the population at the \textit{Td} locus. Nucleotides are
+	colored according to the
+	\textit{p}-value of the most significant \textit{k}-mer overlapping
+	them. The vertical dotted line between haplotypes indicates the location of
+	the causal SNP at this locus. (e) Contingency table of
+	the phenotypes and haplotypes found in the population at the
+	\textit{Td} locus.}
+\end{figure}
+
 \subsubsection*{Seed coat color - \textit{I} locus}
@@
 did not identify the causal variant given its type.
+
+% Figure 3 with the results on pubescence color at locus G
+\begin{figure}
+	\centering
+	\includegraphics[width = 5in]{seed_coat_color_greenyellow_G_main_figure.png}
+	\caption{(Caption on the following page.)}
+	\label{seed-coat-color-main-figure}
+\end{figure}
+
+\begin{figure}
+	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
+	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
+	Horizontal dotted lines indicate the significance
+	threshold whereas vertical dotted lines indicate the location of the
+	\textit{G} locus. (b) Zoomed-in Manhattan plots at the \textit{G} locus.
+	The vertical dotted lines indicate the location of the
+	Glyma.01g198500 gene associated with this locus. The shaded gray
+	rectangles show the regions delimited by the top 1\% associations.
+	In the case of SNPs, blue points denote markers used in the
+	original analysis, whereas red points denote markers that had originally
+	been pruned but whose \emph{p}-values were computed after signal
+	discovery.
+	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
+	rectangles represent coding sequences, and the arrows indicate the
+	direction of transcription. The red rectangle highlights the region
+	shown in panel (d). (d) Pairwise alignment of the two haplotype
+	sequences in the population at the \textit{G} locus. Nucleotides are
+	colored according to the
+	\textit{p}-value of the most significant \textit{k}-mer overlapping
+	them. The vertical dotted line between haplotypes indicates the location of
+	the causal SNP at this locus. (e) Contingency table of
+	the phenotypes and haplotypes found in the population at the
+	\textit{G} locus.}
+\end{figure}
 
@@
 
-The difference in size between the CNV reported by Liu et al. \cite{liu2020ps} and our
-results may be due to the presence of some unresolved sequence (a stretch of
-4,635 N nucleotides) over that interval in the reference sequence that we used.
-Because our SV filtering process removed any alternate allele containing N nucleotides,
-we reasoned that the CNV may have been found by our pipeline but
-filtered out later on. Indeed, we found a 42.8-kb duplication corresponding
-to the one described above among the genotype calls made by smoove. An association
-analysis of that single variant using GAPIT found a \textit{p}-value of 1.9e-23,
-which would have made it the most significant SV. In this particular case, the
-failure of the SV-based GWAS analysis to find identify the causal variant was due
-to our filtering parameters and not to limitations of the SV genotype calls per se.
-
-\subsubsection*{Seed coat luster -- \textit{B} locus}
-
-We conducted a GWAS analysis comparing dull and shiny seed coat phenotypes for
-the seed coat luster trait. Seed coat luster has been shown by
-Gijzen et al. \cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
-protein (HPS) at the seed surface. Copy number variation of this gene at the
-\textit{B} locus was later shown to explain variation in luster, although
-sequence variation in the HPS sequence may also play a role \citep{gijzen2006}.
-Using a sequencing depth-based analysis, we identified a 31-kb region spanning
-positions 9,386,109-9,417,431 on chromosome Gm15 that exhibited variation in
-copy number across accessions. This CNV overlapped four genes annotated as
-containing HPS domains and therefore likely represented the causal variant
-(Supplemental Figures S29 and
-S30).  We found signals overlapping this
-CNV using all approaches, and the most significantly associated \textit{k}-mer
-was notably located within the CNV region (Supplemental Figures
-S29 and
-S31). However, we were not able to
-directly link any of the variants or \textit{k}-mers identified to the causal
-CNV.
-
-\subsection*{Analysis of loci with unknown causal genes}
-
-In addition to analyzing loci for which the genes are already known, we were
-able to suggest candidate genes at loci for which the underlying genes are not
-known yet.
-
-\subsubsection*{Pod color -- \textit{L1} and \textit{L2} loci}
-
-Both the \textit{L1} and the \textit{L2} loci for pod color have yet to be
-cloned. While He et al. \cite{he2015} proposed the gene Glyma.19g101700 as likely
-corresponding to \textit{L1} based on fine mapping and expression analysis, our
-results as well as those reported by Bandillo et al. \cite{bandillo2017} do not support this
-gene as a candidate due to its location $\sim$3 Mb away from GWAS signals. Our
-analyses based on \textit{k}-mers and SNPs found signals overlapping 14 and
-6 genes, respectively (Supplemental figures S32 and
-S33). Of these, a MATE transporter (Glyma.19g120300)
-appears as a prime candidate as the most significant \emph{k}-mer was observed
-within its sequence and it is highly expressed in pods according to data hosted
-on SoyBase \citep{grant2010, severin2010}. Moreover, the role of some MATE
-proteins in transporting flavonoids has been demonstrated \citep{chen2015}.
-
-The signals found by SNP- and \emph{k}-mer-based analyses at the \textit{L2}
-locus overlap 24 and 50 genes, respectively (Supplemental Figures
-S32 and S34). Despite this
-somewhat high number, we noticed that the closest gene to the most significant
-\emph{k}-mer (Glyma.03g005800) is also a MATE transporter, similarly to our top
-candidate for \emph{L1}. This gene may therefore represent an interesting
-candidate at this locus for future studies.
-
-\subsubsection*{Pubescence form -- \textit{Pa1} and \textit{Pa2} loci}
-\label{sv-gwas-main-results-pubescence-form-pa1-pa2}
-
-The two known loci for pubescence form in soybean, \textit{Pa1} and
-\textit{Pa2}, have also not been cloned yet. Nevertheless, Gilbert \cite{gilbert2017}
-has suggested Glyma.12g213900 as a candidate for \textit{Pa1} based on data
-published by Bandillo et al. \cite{bandillo2017}. This gene encodes a MYB transcription factor
-annotated as playing a role in trichome branching. Indeed, the most significant
-signals found by all three approaches were located near this gene (Figure
-\ref{pubescence-form-main-figure}a-b, and Supplemental Figures S35 and
-S36).  The analyses based on SNPs and
-\emph{k}-mers identified two highly significant non-synonymous SNPs in
-Glyma.12g213900 that were associated with appressed pubescence (Figure
-\ref{pubescence-form-main-figure}c-d and Supplemental Figures S37 and
-S38). The alternative alleles for these two
-SNPs co-occurred together in our dataset and may represent causal variants at
-the \emph{Pa1} locus. The signals observed at the \emph{Pa2} locus were much
-weaker and the signal observed from the \emph{k}-mer GWAS overlapped a total of
-20 genes (Supplemental Figures S35 and
-S39), none of which appeared obviously linked
-to pubescence form.
-
-A second GWAS analysis focused only on appressed and semi-appressed phenotypes
-with the objective of enhancing the signal at the \emph{Pa2} locus. However,
-this analysis only detected weak signals at the \emph{Pa1} locus (Supplemental Figure
-S40) and its results were therefore not
-analyzed further.
-
-\subsubsection*{Resistance to \textit{P. sojae}}
-
-We used corrected dry weight (CDW) as described by de Ronne et al. \cite{deronne2022} as a
-measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
-overlapping the genomic region reported by these authors on Gm15 using all three GWAS
-approaches (Supplemental Figure S41). de Ronne et al. \cite{deronne2022}
-proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
-variation at that locus based on a combination of functional annotation and
-contrasting gene expression in resistant and susceptible lines four days after
-inoculation. Although the signals found by our analyses did overlap
-Glyma.15g217100, we found that this gene was located at the distal end of the signal
-in a region with markers or \emph{k}-mers showing a weaker
-association (Supplemental Figure S42). The
-large size of the observed signal means that the gene associated with the phenotype could
-be located almost anywhere in this region. However, we identified a particularly
-interesting region defined by the top 1\% associations in the SNP-based
-analysis. This region overlaps only three annotated genes and also contains the
-most significant \emph{k}-mer (Supplemental Figure
-S42). One of the genes in this region is
-a calcium ion binding protein with an EF-hand domain (Glyma.15g217700).
-Given the important role played by calcium signaling in defense
-responses \citep{zhang2014}, we suggest Glyma.15g217700 as an interesting
-candidate gene for explaining resistance to \emph{P. sojae} in this population.
-
-\subsubsection*{Novel loci for stem termination and pod color}
-
-In addition to verifying known loci, we also observed potentially novel loci in
-the \textit{k}-mer GWAS analysis of stem termination and pod color.  When
-comparing accessions with indeterminate and semi-determinate stem termination
-types, we observed previously undocumented signals on chromosomes Gm11, Gm16
-and Gm18 in addition to the known \textit{Dt1} locus (Figure
-\ref{stem-termination-main-figure}a).  Analysis of the pairwise LD between
-significant \textit{k}-mers found no LD between these signals, suggesting that
-they may represent \textit{bona fide} loci associated with stem termination
-type (Figure \ref{stem-termination-main-figure}b). Similarly, a \textit{k}-mer
-GWAS analysis comparing accessions with black and brown pods found a previously
-undocumented signal on chromosome Gm15 (Supplemental Figure S43).
-\textit{k}-mers at this new locus
-showed no LD with \textit{k}-mers at the known \textit{L1} locus and may
-therefore represent a distinct locus (Supplemental Figure 
-S44). The observed signal overlaps a total of 27 genes
-(Supplemental Figure S45), but none of those are annotated
-as having an obvious role in controlling pod color.
-
-\subsection*{Potential of \emph{k}-mer-based GWAS}
-
-One of the most striking results from the analyses presented above is the
-impressive performance of \emph{k}-mer-based GWAS compared to SNP- and SV-based
-GWAS. The GWAS analyses using \emph{k}-mers identified the causal variant as
-the most significantly associated variant at four loci with various types of
-causal variants: \emph{W1} for flower color, \emph{Td} for pubescence color,
-\emph{G} for seed coat color, and \emph{Ps} for pubescence density (Table
-\ref{loci-table}). In comparison, this result was achieved by SV-based GWAS at
-the \textit{W1} locus only, and at none of the loci for SNP-based GWAS.
-The \textit{k}-mer approach also systematically performed at least as well as
-the other two approaches in identifying signals overlapping known variants or
-genes (Table \ref{loci-table}). Moreover, whenever \textit{k}-mer-based GWAS
-did detect a signal, the region defined by the 1\% most significant
-\textit{k}-mers overlapped the known causal variant or gene.
-
-Even in cases where the most significantly associated \emph{k}-mers did not
-pinpoint known causal variants, they were often located near those variants.
-For example, the most significantly associated \emph{k}-mers mapping to the
-\emph{T} locus for pubescence color were located in intronic sequences of the
-gene known to be associated with this locus.  Similarly, the most significantly
-associated \emph{k}-mers at the \emph{B} locus for seed coat luster mapped
-within a genomic interval known to vary in copy number at that locus. In
-addition, the most significant \emph{k}-mers at the \emph{R} locus mapped just
-upstream of a homolog of the gene associated with the locus. Even though the
-\emph{k}-mer analysis did not identify the causal variants directly in those
-cases, it still provided a very strong starting point for candidate gene
-identification.
-
-As would be expected of any GWAS analysis, the success of the \emph{k}-mer approach
-depends on the quality of the phenotypic data and the complexity of the genetic
-architecture of the trait. It is clear that the loci for which the
-\emph{k}-mer approach was most successful (\emph{W1}, \emph{Td}, \emph{G} and
-\emph{Ps}) have a simple genetic architecture and phenotypes that can be easily
-determined. For other traits, challenges in determining accurate phenotypes may
-result in a decreased ability to identify causal variants or genes. For
-example, Bandillo et al. \cite{bandillo2017} noted that observed phenotypes for stem
-termination type in the GRIN database may not correspond to the expected
-genotypes at classical loci, and that the intermediate phenotype for seed coat
-luster is not always consistent across different environments. For qualitative
-traits, the numerical scale used to recode phenotypes for use in GWAS is somewhat
-arbitrary and may influence the final results. In the case of quantitative
-traits, the usual limitations to the precision and stochastic variation of
-phenotypic values will necessarily make the interpretation more difficult than
-for simple qualitative traits. Notwithstanding those limitations,
-\textit{k}-mer-based GWAS clearly performed better than any of the other
-approaches tested here given the same phenotypic data. This is partly due
-to the fact that the \textit{k}-mer approach encompasses all variant types.
-Moreover, it is likely that the simpler workflow that results from using
-\textit{k}-mer presence/absence avoids some of the errors introduced during
-variant calling and genotyping, and therefore yields more accurate results.
-
-\subsection*{Limitations and challenges of \emph{k}-mer-based GWAS}
-
-While we have have shown that using \emph{k}-mers for GWAS in soybean performed
-well compared to other GWAS methods, this approach does present some unique
-challenges and limitations. One obvious limitation is that using the
-presence/absence of \emph{k}-mers may not be appropriate in species where
-heterozygosity is common (i.e., most species other than inbred crops) as the
-\emph{k}-mers associated with the presence of both alleles will be observed at
-heterozygous loci. Moreover, although the presence/absence of \emph{k}-mers can
-detect variation independently of its type, some types of variation may be more
-difficult to detect than others.  For example, copy number variation at a locus
-may not be detected from presence/absence alone if all copies are identical and
-occur in tandem, such that no \textit{k}-mer can differentiate between two
-copies and more. For similar reasons, \emph{k}-mers may also fail to capture
-variation occurring in repeated sequences. Alternative models based on
-\emph{k}-mer counts instead of presence/absence have been developed
-\citep{rahman2018, he2021} which may solve issues like these and broaden the
-scope of variants that can be identified from \emph{k}-mers.
-
-The complexity of analyzing the results of \emph{k}-mer-based GWAS is another
-limitation of this approach. Although the method developed by
-Voichek and Weigel \cite{voichek2020} is in itself relatively simple and computationally
-inexpensive, downstream analysis and interpretation of the output is
-challenging. Indeed, there are no state-of-the-art methods for identifying the
-putative genomic coordinates of a \emph{k}-mer, grouping significant
-\emph{k}-mers together into coherent signals, and linking them to biologically
-meaningful sequence variation. Going forward, mapping the \emph{k}-mers or
-reads to a graph-based genome \citep[e.g. using the vg toolkit;][]{siren2021}
-may represent an improvement to using a single linear genome.  Grouping
-significant \emph{k}-mers by LD instead of by genomic position, as was done for
-some of the analyses in this paper, could prove useful in determining how many
-loci control a trait. In this study, we used an assembly-based approach similar
-to what others have previously done \citep[e.g.][]{voichek2020, rahman2018}.
-This approach allowed us to link \emph{k}-mers to their underlying variation
-for simple cases involving SNPs and indels, but we were unable to link
-significantly associated \emph{k}-mers to more challenging causal variants such
-as those explaining the \emph{I} or \emph{B} loci. In this context, developing
-approaches to systematically link significant \emph{k}-mers to the sequence
-variation that underlies their presence is much needed. Furthermore, finding
-ways to limit the number of spurious associations found for some traits, as was
-the case for flower color and pubescence color in our study, is also needed in
-order to obtain robust results.
-
-In this work, we provide a set of tools enabling the downstream analysis of
-\emph{k}-mer-based GWAS. Most importantly, we developed the katcher program for
-retrieving all reads containing any of a set of significant \emph{k}-mers in a
-computationally efficient manner. Previous studies usually mapped \emph{k}-mers
-to a reference genome to identify their reference-based coordinates
-\citep[e.g.][]{voichek2020, he2021, rahman2018, tripodi2021}.  However, while
-this approach may work for \emph{k}-mers that show little divergence from the
-reference, it might fail to identify the genomic location of \emph{k}-mers that
-diverge significantly or are not found at all in the reference. Moreover, it
-does not take into account the sequence context of the read where the
-\emph{k}-mer was found, which may provide valuable information for mapping
-purposes. Our approach avoids these pitfalls by instead systematically
-retrieving reads containing significant \emph{k}-mers, including those that did
-not map at all to the reference assembly. While we were able to position the
-most significant \emph{k}-mers on the reference sequence in our analyses, it
-may not be the case for highly variable species or if the causal variation at a
-locus results from a large novel insertion.
-
-\subsection*{Use of SV genotypes for GWAS}
-
-In addition to \emph{k}-mers, we also assessed the use of SVs in GWAS analysis
-of soybean.  Although we did find the causal variant as the most significantly
-associated SV at the \emph{W1} locus, results were otherwise rather
-inconclusive.  Indeed, most of the other known or putative causal variants were
-SNPs or indels, and therefore were not represented in the SV datasets. At most
-loci where SVs are known to be the underlying causal variant (e.g.  \emph{I},
-\emph{B}, \emph{Ps}), the analyses based on SVs did not find the causal
-variants, most likely due to the difficulty of calling and genotyping such
-variants \citep{kosugi2019, lemay2022}. In the case of the \textit{Ps} locus,
-this limitation was due to a filtering step in our pipeline rather than to an
-inherent inability to properly call and genotype the causal variant.
-
-Given the current performance of SV genotyping programs, it is not clear
-whether using SVs for GWAS provides any advantage over using \emph{k}-mer- or
-SNP-based GWAS, or a combination of both.  Indeed, in the single case where the
-SV-based analyses identified the causal variant, the \emph{k}-mer-based GWAS
-identified it as well. Moreover, SNP-based GWAS provided better guidance
-in delimiting the signals found at a given locus because of the much higher
-density of markers compared to SVs.  The usefulness of SV-based GWAS compared
-to SNP-based GWAS will largely depend on whether SVs can represent variation
-that is not otherwise in LD with SNPs. In tomato, Domínguez et al. \cite{dominguez2020} found
-that most transposable element insertions associated with phenotypic variation
-in agronomic traits were not tagged by SNPs. On the other hand, in humans,
-Maretty et al. \cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
-is therefore likely that how well SVs can be tagged by SNPs will depend on the
-particular type of SV and on the species. A more efficient approach than
-SV-based GWAS may be to conduct \emph{k}-mer-based GWAS and only then link
-known SVs to significant \emph{k}-mers.  Alternatively, \textit{k}-mer-based
-GWAS can be conducted specifically for \textit{k}-mers that are associated with
-SVs \citep{jayakodi2020}.  Previous studies have obtained noteworthy results by
-using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
-dominguez2020}, however it would be worthwhile to see whether these results
-could be replicated using \emph{k}-mer-based GWAS.
-
-\section*{Conclusions}
-
-In conclusion, we used SV-based and \emph{k}-mer-based GWAS to study ten
-qualitative traits and three quantitative traits in soybean and
-compared the results to conventional SNP-based GWAS. While \emph{k}-mer-based
-GWAS proved to be a powerful approach in pinpointing the causal variation or
-genes associated with known loci, it is unclear whether conducting SV-based
-GWAS is worth the investment. In addition, we used the results from the
-\emph{k}-mer- and SNP-based GWAS to suggest candidate genes for a few classical
-loci that have yet to be cloned. Based on our results, we believe that an
-optimal workflow may involve conducting SNP-based and \emph{k}-mer-based GWAS
-in parallel to identify significant signals and candidate genes, potentially in
-combination with SV datasets. Given the large number of samples for which WGS
-data is now available in several species, applying \emph{k}-mer-based GWAS to
-leverage already existing sequencing and phenotypic data appears feasible and
-promising.  As part of our work, we developed several computational tools that
-should help other researchers with the downstream analysis of
-\emph{k}-mer-based GWAS. However, much work remains to be done in
-developing state-of-the-art methods for the downstream analysis of significant
-\emph{k}-mers. In particular, better approaches are needed for linking
-\emph{k}-mers to sequence variation and moving from presence/absence-based
-methods to count-based methods.
-
-\section*{Supplemental Files}
-
-\subsection*{Supplemental file 1 --- Supplemental file 1.pdf}
-Supplemental methods and results,
-including Supplemental Tables S1 to S11
-and Supplemental Figures S1 to S60 (PDF 12.9 MB)
-
-\subsection*{Supplemental file 2 --- Supplemental file 2.csv}
-Metadata on the sequencing runs retrieved
-from the Sequence Read Archive (SRA) (CSV 71 KB)
-
-\subsection*{Supplemental file 3 --- Supplemental file 3.csv}
-Metadata on individual accessions following mapping
-and comparison to SoySNP50K data (CSV 36KB)
-
-\subsection*{Supplemental file 4 --- Supplemental file 4.csv}
-Phenotypic dataset used for GWAS analyses (CSV 135 KB)
-
-\section*{Acknowledgements}%% if any
-
-We would like to thank Jonas A. Sibbesen for providing very helpful comments
-on the methodology used and on a previous version of this manuscript.
-We also thank Yoav Voichek for help regarding \textit{k}-mer GWAS
-analysis as well as Brian Boyle and Martine Jean for valuable comments
-regarding the methodology. We are thankful to the Digital Research Alliance of
-Canada for using their high-performance computing servers.
-
-\section*{Funding}%% if any
-
-This work was supported by the SoyaGen grant (\url{https://www.soyagen.ca})
-awarded to F. Belzile and R. Bélanger, and funded by Génome Québec, Genome
-Canada, the government of Canada, the Ministère de l'Économie, Science et
-Innovation du Québec, Semences Prograin Inc., Syngenta Canada Inc., Sevita
-Genetics, Coop Fédérée, Grain Farmers of Ontario, Saskatchewan Pulse Growers,
-Manitoba Pulse \& Soybean Growers, the Canadian Field Crop Research Alliance
-and Producteurs de grains du Québec. M-A. Lemay has been supported by a NSERC
-Canada Vanier Graduate Scholarship, a FRQNT doctoral B2X scholarship, a NSERC
-Michael Smith Foreign Study Supplement, and a scholarship from the
-AgroPhytoSciences NSERC CREATE Training Program. None of the funding bodies
-were involved in study design, data acquisition, data analysis, interpretation
-of the results, or manuscript writing.
-
-\section*{Abbreviations}%% if any
-BAM: Binary alignment map format;
-CDW: corrected dry weight;
-CHS: chalcone synthase;
-CNV: copy number variant, copy number variation;
-GRIN: Germplasm Resources Information Network;
-GWAS: Genome-wide association study;
-HPS: Hydrophobic protein from soybean;
-LD: Linkage disequilibrium;
-MAF: Minor allele frequency;
-MATE: Multidrug and toxic compound extrusion;
-MLM: Mixed linear model;
-NCBI: National Center for Biotechnology Information;
-SNP: Single-nucleotide polymorphism;
-SRA: Sequence Read Archive;
-SV: Structural variant, structural variation;
-VCF: Variant call format;
-WGS: Whole-genome (re)sequencing
-
-\section*{Availability of data and materials}
-
-Some of the datasets generated as part of this work are available on figshare
-\citep{kmer-gwas-figshare}.
-
-The Illumina sequencing data used for the analyses are listed in Supplemental file 2
-and available through NCBI BioProjects PRJNA257011 \cite{PRJNA257011}, PRJNA289660 \cite{PRJNA289660}
-and PRJNA639876 \cite{PRJNA639876}.
-
-The high-quality assemblies generated by Liu et al. \cite{liu2020-pangenome} are available
-on the Genome Warehouse through Accession Number PRJCA002030 \cite{PRJCA002030}.
-
-The assemblies of ZH13, W05 and Lee are available on SoyBase
-(\url{https://soybase.org/GlycineBlastPages/blast_descriptions.php}).
-
-The \textit{I} locus contig assembled by Tuteja and Vodkin \cite{tuteja2008} is available from
-NCBI (\url{https://www.ncbi.nlm.nih.gov/nuccore/EF623854}).
-
-SoySNP50K genotype calls are available from SoyBase (\url{https://soybase.org/snps/}).
-
-The SVs identified from Oxford Nanopore data by Lemay et al. \cite{lemay2022} are available
-on figshare \citep{soybean-sv-figshare}.
-
-The following software used in this work is available on GitHub:
-
-\begin{itemize}
-	\item Code used for the analyses:
-		\url{https://github.com/malemay/soybean_kmer_gwas}
-	\item katcher software for retrieving reads containing \textit{k}-mers:
-		\url{https://github.com/malemay/katcher}
-	\item gwask R package used for processing the output of GAPIT and
-		\textit{k}-mer GWAS analyses, and for plotting:
-		\url{https://github.com/malemay/gwask}
-	\item Software for computing the LD between \textit{k}-mers
-		based on presence/absence: \url{https://github.com/malemay/kmers_ld}
-	\item Forked svmu version that was used for the analyses shown in this
-		paper: \url{https://github.com/malemay/svmu}
-	\item svmutools R package for converting the output of svmu to VCF
-		format: \url{https://github.com/malemay/svmutools}
-\end{itemize}
-
-The code associated with the results presented in this manuscript is archived
-on figshare \citep{kmer-gwas-code}.
-
-\section*{Ethics approval and consent to participate}%% if any
-Not applicable.
-
-\section*{Competing interests}
-The authors declare that they have no competing interests.
-
-\section*{Consent for publication}%% if any
-Not applicable.
-
-\section*{Authors' contributions}
-Conception and design of the study: MAL, MDR, RB, FB.
-Phenotypic data acquisition (resistance to \textit{Phytophthora sojae}): MDR, RB.
-Data analysis: MAL.
-Data interpretation: MAL, FB.
-Software writing: MAL.
-Manuscript drafting: MAL, FB.
-All authors have revised the manuscript and approved its submission.
-
-\bibliographystyle{genome_research}
-\bibliography{references}
-
-\section*{Figures}
-
-% Figure 1
-\begin{figure}[p]
-	\caption{Results of SV- and \textit{k}-mer-based GWAS at the
-	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots.
-	Horizontal dotted lines indicate the significance
-	threshold whereas vertical dotted lines indicate the location of the
-	\textit{W1} locus. (b) Zoomed-in Manhattan plots at the \textit{W1} locus.
-	The vertical dotted lines indicate the location of the
-	Glyma.13g072100 gene associated with this locus. The shaded gray
-	rectangles show the regions delimited by the top 5\% (SVs) and
-	1\% (\textit{k}-mers) associations.  (c) Gene model of Glyma.13g072100.
-	Rectangles represent exons, colored rectangles represent coding
-	sequences, and arrows indicate the direction of transcription. The red
-	rectangle highlights the region shown in panel (d). (d) Pairwise
-	alignment of the two haplotype sequences in the population at the
-	\textit{W1} locus. Nucleotides are
-	colored according to the \textit{p}-value of the most significant
-	\textit{k}-mer overlapping them. Vertical dotted lines between haplotypes
-	indicate sequence differences, whereas dashes indicate gaps in the
-	alignment. (e) Contingency table of the phenotypes and haplotypes found
-	in the population at the \textit{W1} locus.}
-	\label{flower-color-main-figure}
-\end{figure}
-
-% Figure 2
-\begin{figure}[p]
-	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
-	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
-	Horizontal dotted lines indicate the significance
-	threshold whereas vertical dotted lines indicate the location of the
-	\textit{Td} locus. (b) Zoomed-in Manhattan plots at the \textit{Td} locus.
-	The vertical dotted lines indicate the location of the
-	Glyma.03g258700 gene associated with this locus. The shaded gray
-	rectangles show the regions delimited by the top 1\% associations.
-	In the case of SNPs, blue points denote markers used in the
-	original analysis, whereas red points denote markers that had originally
-	been pruned but whose \emph{p}-values were computed after signal
-	discovery.
-	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
-	rectangles represent coding sequences, and the arrow indicates the
-	direction of transcription. The red rectangle highlights the region
-	shown in panel (d). (d) Pairwise alignment of the two haplotype
-	sequences in the population at the \textit{Td} locus. Nucleotides are
-	colored according to the
-	\textit{p}-value of the most significant \textit{k}-mer overlapping
-	them. The vertical dotted line between haplotypes indicates the location of
-	the causal SNP at this locus. (e) Contingency table of
-	the phenotypes and haplotypes found in the population at the
-	\textit{Td} locus.}
-	\label{pubescence-color-main-figure}
-\end{figure}
-
-% Figure 3 with the results on pubescence color at locus G
-\begin{figure}[p]
-	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
-	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
-	Horizontal dotted lines indicate the significance
-	threshold whereas vertical dotted lines indicate the location of the
-	\textit{G} locus. (b) Zoomed-in Manhattan plots at the \textit{G} locus.
-	The vertical dotted lines indicate the location of the
-	Glyma.01g198500 gene associated with this locus. The shaded gray
-	rectangles show the regions delimited by the top 1\% associations.
-	In the case of SNPs, blue points denote markers used in the
-	original analysis, whereas red points denote markers that had originally
-	been pruned but whose \emph{p}-values were computed after signal
-	discovery.
-	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
-	rectangles represent coding sequences, and the arrows indicate the
-	direction of transcription. The red rectangle highlights the region
-	shown in panel (d). (d) Pairwise alignment of the two haplotype
-	sequences in the population at the \textit{G} locus. Nucleotides are
-	colored according to the
-	\textit{p}-value of the most significant \textit{k}-mer overlapping
-	them. The vertical dotted line between haplotypes indicates the location of
-	the causal SNP at this locus. (e) Contingency table of
-	the phenotypes and haplotypes found in the population at the
-	\textit{G} locus.}
-	\label{seed-coat-color-main-figure}
-\end{figure}
-
 % Figure 4 with the results on pubescence density at locus Ps
-\begin{figure}[p]
+\begin{figure}
+	\centering
+	\includegraphics[width = 6in]{pubescence_density_Ps_main_figure.png}
 	\caption[Results of \textit{k}-mer-based GWAS at the
@@
 
+The difference in size between the CNV reported by Liu et al. \cite{liu2020ps} and our
+results may be due to the presence of some unresolved sequence (a stretch of
+4,635 N nucleotides) over that interval in the reference sequence that we used.
+Because our SV filtering process removed any alternate allele containing N nucleotides,
+we reasoned that the CNV may have been found by our pipeline but
+filtered out later on. Indeed, we found a 42.8-kb duplication corresponding
+to the one described above among the genotype calls made by smoove. An association
+analysis of that single variant using GAPIT found a \textit{p}-value of 1.9e-23,
+which would have made it the most significant SV. In this particular case, the
+failure of the SV-based GWAS analysis to find identify the causal variant was due
+to our filtering parameters and not to limitations of the SV genotype calls per se.
+
+\subsubsection*{Seed coat luster -- \textit{B} locus}
+
+We conducted a GWAS analysis comparing dull and shiny seed coat phenotypes for
+the seed coat luster trait. Seed coat luster has been shown by
+Gijzen et al. \cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
+protein (HPS) at the seed surface. Copy number variation of this gene at the
+\textit{B} locus was later shown to explain variation in luster, although
+sequence variation in the HPS sequence may also play a role \citep{gijzen2006}.
+Using a sequencing depth-based analysis, we identified a 31-kb region spanning
+positions 9,386,109-9,417,431 on chromosome Gm15 that exhibited variation in
+copy number across accessions. This CNV overlapped four genes annotated as
+containing HPS domains and therefore likely represented the causal variant
+(Supplemental Figures S29 and
+S30).  We found signals overlapping this
+CNV using all approaches, and the most significantly associated \textit{k}-mer
+was notably located within the CNV region (Supplemental Figures
+S29 and
+S31). However, we were not able to
+directly link any of the variants or \textit{k}-mers identified to the causal
+CNV.
+
+\subsection*{Analysis of loci with unknown causal genes}
+
+In addition to analyzing loci for which the genes are already known, we were
+able to suggest candidate genes at loci for which the underlying genes are not
+known yet.
+
+\subsubsection*{Pod color -- \textit{L1} and \textit{L2} loci}
+
+Both the \textit{L1} and the \textit{L2} loci for pod color have yet to be
+cloned. While He et al. \cite{he2015} proposed the gene Glyma.19g101700 as likely
+corresponding to \textit{L1} based on fine mapping and expression analysis, our
+results as well as those reported by Bandillo et al. \cite{bandillo2017} do not support this
+gene as a candidate due to its location $\sim$3 Mb away from GWAS signals. Our
+analyses based on \textit{k}-mers and SNPs found signals overlapping 14 and
+6 genes, respectively (Supplemental figures S32 and
+S33). Of these, a MATE transporter (Glyma.19g120300)
+appears as a prime candidate as the most significant \emph{k}-mer was observed
+within its sequence and it is highly expressed in pods according to data hosted
+on SoyBase \citep{grant2010, severin2010}. Moreover, the role of some MATE
+proteins in transporting flavonoids has been demonstrated \citep{chen2015}.
+
+The signals found by SNP- and \emph{k}-mer-based analyses at the \textit{L2}
+locus overlap 24 and 50 genes, respectively (Supplemental Figures
+S32 and S34). Despite this
+somewhat high number, we noticed that the closest gene to the most significant
+\emph{k}-mer (Glyma.03g005800) is also a MATE transporter, similarly to our top
+candidate for \emph{L1}. This gene may therefore represent an interesting
+candidate at this locus for future studies.
+
+\subsubsection*{Pubescence form -- \textit{Pa1} and \textit{Pa2} loci}
+\label{sv-gwas-main-results-pubescence-form-pa1-pa2}
+
+The two known loci for pubescence form in soybean, \textit{Pa1} and
+\textit{Pa2}, have also not been cloned yet. Nevertheless, Gilbert \cite{gilbert2017}
+has suggested Glyma.12g213900 as a candidate for \textit{Pa1} based on data
+published by Bandillo et al. \cite{bandillo2017}. This gene encodes a MYB transcription factor
+annotated as playing a role in trichome branching. Indeed, the most significant
+signals found by all three approaches were located near this gene (Figure
+\ref{pubescence-form-main-figure}a-b, and Supplemental Figures S35 and
+S36).  The analyses based on SNPs and
+\emph{k}-mers identified two highly significant non-synonymous SNPs in
+Glyma.12g213900 that were associated with appressed pubescence (Figure
+\ref{pubescence-form-main-figure}c-d and Supplemental Figures S37 and
+S38). The alternative alleles for these two
+SNPs co-occurred together in our dataset and may represent causal variants at
+the \emph{Pa1} locus. The signals observed at the \emph{Pa2} locus were much
+weaker and the signal observed from the \emph{k}-mer GWAS overlapped a total of
+20 genes (Supplemental Figures S35 and
+S39), none of which appeared obviously linked
+to pubescence form.
+
 % Figure 5 with the results on pubescence form at locus Pa1
-\begin{figure}[p]
-	\caption{Results of SNP- and \textit{k}-mer-based GWAS at the
+\begin{figure}
+	\centering
+	\includegraphics[width = 5in]{pubescence_form_all_Pa1_main_figure.png}
+	\caption{(Caption on the following page.)}
+	\label{pubescence-form-main-figure}
+\end{figure}
+
+\begin{figure}
+	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
 	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots.
@@
 	at this locus. Panels (c) and (d) share the same x-axis.}
-	\label{pubescence-form-main-figure}
 \end{figure}
 
+A second GWAS analysis focused only on appressed and semi-appressed phenotypes
+with the objective of enhancing the signal at the \emph{Pa2} locus. However,
+this analysis only detected weak signals at the \emph{Pa1} locus (Supplemental Figure
+S40) and its results were therefore not
+analyzed further.
+
+\subsubsection*{Resistance to \textit{P. sojae}}
+
+We used corrected dry weight (CDW) as described by de Ronne et al. \cite{deronne2022} as a
+measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
+overlapping the genomic region reported by these authors on Gm15 using all three GWAS
+approaches (Supplemental Figure S41). de Ronne et al. \cite{deronne2022}
+proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
+variation at that locus based on a combination of functional annotation and
+contrasting gene expression in resistant and susceptible lines four days after
+inoculation. Although the signals found by our analyses did overlap
+Glyma.15g217100, we found that this gene was located at the distal end of the signal
+in a region with markers or \emph{k}-mers showing a weaker
+association (Supplemental Figure S42). The
+large size of the observed signal means that the gene associated with the phenotype could
+be located almost anywhere in this region. However, we identified a particularly
+interesting region defined by the top 1\% associations in the SNP-based
+analysis. This region overlaps only three annotated genes and also contains the
+most significant \emph{k}-mer (Supplemental Figure
+S42). One of the genes in this region is
+a calcium ion binding protein with an EF-hand domain (Glyma.15g217700).
+Given the important role played by calcium signaling in defense
+responses \citep{zhang2014}, we suggest Glyma.15g217700 as an interesting
+candidate gene for explaining resistance to \emph{P. sojae} in this population.
+
+\subsubsection*{Novel loci for stem termination and pod color}
+
+In addition to verifying known loci, we also observed potentially novel loci in
+the \textit{k}-mer GWAS analysis of stem termination and pod color.  When
+comparing accessions with indeterminate and semi-determinate stem termination
+types, we observed previously undocumented signals on chromosomes Gm11, Gm16
+and Gm18 in addition to the known \textit{Dt1} locus (Figure
+\ref{stem-termination-main-figure}a).  Analysis of the pairwise LD between
+significant \textit{k}-mers found no LD between these signals, suggesting that
+they may represent \textit{bona fide} loci associated with stem termination
+type (Figure \ref{stem-termination-main-figure}b). Similarly, a \textit{k}-mer
+GWAS analysis comparing accessions with black and brown pods found a previously
+undocumented signal on chromosome Gm15 (Supplemental Figure S43).
+\textit{k}-mers at this new locus
+showed no LD with \textit{k}-mers at the known \textit{L1} locus and may
+therefore represent a distinct locus (Supplemental Figure 
+S44). The observed signal overlaps a total of 27 genes
+(Supplemental Figure S45), but none of those are annotated
+as having an obvious role in controlling pod color.
+
 % Figure 6 with the results novel loci for stem termination type
-\begin{figure}[p]
+\begin{figure}
+	\centering
+	\includegraphics[width = 6in]{stem_termination_sn_main_figure.png}
 	\caption[Detection of novel signals for stem termination type
@@
 
+\subsection*{Potential of \emph{k}-mer-based GWAS}
+
+One of the most striking results from the analyses presented above is the
+impressive performance of \emph{k}-mer-based GWAS compared to SNP- and SV-based
+GWAS. The GWAS analyses using \emph{k}-mers identified the causal variant as
+the most significantly associated variant at four loci with various types of
+causal variants: \emph{W1} for flower color, \emph{Td} for pubescence color,
+\emph{G} for seed coat color, and \emph{Ps} for pubescence density (Table
+\ref{loci-table}). In comparison, this result was achieved by SV-based GWAS at
+the \textit{W1} locus only, and at none of the loci for SNP-based GWAS.
+The \textit{k}-mer approach also systematically performed at least as well as
+the other two approaches in identifying signals overlapping known variants or
+genes (Table \ref{loci-table}). Moreover, whenever \textit{k}-mer-based GWAS
+did detect a signal, the region defined by the 1\% most significant
+\textit{k}-mers overlapped the known causal variant or gene.
+
+Even in cases where the most significantly associated \emph{k}-mers did not
+pinpoint known causal variants, they were often located near those variants.
+For example, the most significantly associated \emph{k}-mers mapping to the
+\emph{T} locus for pubescence color were located in intronic sequences of the
+gene known to be associated with this locus.  Similarly, the most significantly
+associated \emph{k}-mers at the \emph{B} locus for seed coat luster mapped
+within a genomic interval known to vary in copy number at that locus. In
+addition, the most significant \emph{k}-mers at the \emph{R} locus mapped just
+upstream of a homolog of the gene associated with the locus. Even though the
+\emph{k}-mer analysis did not identify the causal variants directly in those
+cases, it still provided a very strong starting point for candidate gene
+identification.
+
+As would be expected of any GWAS analysis, the success of the \emph{k}-mer approach
+depends on the quality of the phenotypic data and the complexity of the genetic
+architecture of the trait. It is clear that the loci for which the
+\emph{k}-mer approach was most successful (\emph{W1}, \emph{Td}, \emph{G} and
+\emph{Ps}) have a simple genetic architecture and phenotypes that can be easily
+determined. For other traits, challenges in determining accurate phenotypes may
+result in a decreased ability to identify causal variants or genes. For
+example, Bandillo et al. \cite{bandillo2017} noted that observed phenotypes for stem
+termination type in the GRIN database may not correspond to the expected
+genotypes at classical loci, and that the intermediate phenotype for seed coat
+luster is not always consistent across different environments. For qualitative
+traits, the numerical scale used to recode phenotypes for use in GWAS is somewhat
+arbitrary and may influence the final results. In the case of quantitative
+traits, the usual limitations to the precision and stochastic variation of
+phenotypic values will necessarily make the interpretation more difficult than
+for simple qualitative traits. Notwithstanding those limitations,
+\textit{k}-mer-based GWAS clearly performed better than any of the other
+approaches tested here given the same phenotypic data. This is partly due
+to the fact that the \textit{k}-mer approach encompasses all variant types.
+Moreover, it is likely that the simpler workflow that results from using
+\textit{k}-mer presence/absence avoids some of the errors introduced during
+variant calling and genotyping, and therefore yields more accurate results.
+
+\subsection*{Limitations and challenges of \emph{k}-mer-based GWAS}
+
+While we have have shown that using \emph{k}-mers for GWAS in soybean performed
+well compared to other GWAS methods, this approach does present some unique
+challenges and limitations. One obvious limitation is that using the
+presence/absence of \emph{k}-mers may not be appropriate in species where
+heterozygosity is common (i.e., most species other than inbred crops) as the
+\emph{k}-mers associated with the presence of both alleles will be observed at
+heterozygous loci. Moreover, although the presence/absence of \emph{k}-mers can
+detect variation independently of its type, some types of variation may be more
+difficult to detect than others.  For example, copy number variation at a locus
+may not be detected from presence/absence alone if all copies are identical and
+occur in tandem, such that no \textit{k}-mer can differentiate between two
+copies and more. For similar reasons, \emph{k}-mers may also fail to capture
+variation occurring in repeated sequences. Alternative models based on
+\emph{k}-mer counts instead of presence/absence have been developed
+\citep{rahman2018, he2021} which may solve issues like these and broaden the
+scope of variants that can be identified from \emph{k}-mers.
+
+The complexity of analyzing the results of \emph{k}-mer-based GWAS is another
+limitation of this approach. Although the method developed by
+Voichek and Weigel \cite{voichek2020} is in itself relatively simple and computationally
+inexpensive, downstream analysis and interpretation of the output is
+challenging. Indeed, there are no state-of-the-art methods for identifying the
+putative genomic coordinates of a \emph{k}-mer, grouping significant
+\emph{k}-mers together into coherent signals, and linking them to biologically
+meaningful sequence variation. Going forward, mapping the \emph{k}-mers or
+reads to a graph-based genome \citep[e.g. using the vg toolkit;][]{siren2021}
+may represent an improvement to using a single linear genome.  Grouping
+significant \emph{k}-mers by LD instead of by genomic position, as was done for
+some of the analyses in this paper, could prove useful in determining how many
+loci control a trait. In this study, we used an assembly-based approach similar
+to what others have previously done \citep[e.g.][]{voichek2020, rahman2018}.
+This approach allowed us to link \emph{k}-mers to their underlying variation
+for simple cases involving SNPs and indels, but we were unable to link
+significantly associated \emph{k}-mers to more challenging causal variants such
+as those explaining the \emph{I} or \emph{B} loci. In this context, developing
+approaches to systematically link significant \emph{k}-mers to the sequence
+variation that underlies their presence is much needed. Furthermore, finding
+ways to limit the number of spurious associations found for some traits, as was
+the case for flower color and pubescence color in our study, is also needed in
+order to obtain robust results.
+
+In this work, we provide a set of tools enabling the downstream analysis of
+\emph{k}-mer-based GWAS. Most importantly, we developed the katcher program for
+retrieving all reads containing any of a set of significant \emph{k}-mers in a
+computationally efficient manner. Previous studies usually mapped \emph{k}-mers
+to a reference genome to identify their reference-based coordinates
+\citep[e.g.][]{voichek2020, he2021, rahman2018, tripodi2021}.  However, while
+this approach may work for \emph{k}-mers that show little divergence from the
+reference, it might fail to identify the genomic location of \emph{k}-mers that
+diverge significantly or are not found at all in the reference. Moreover, it
+does not take into account the sequence context of the read where the
+\emph{k}-mer was found, which may provide valuable information for mapping
+purposes. Our approach avoids these pitfalls by instead systematically
+retrieving reads containing significant \emph{k}-mers, including those that did
+not map at all to the reference assembly. While we were able to position the
+most significant \emph{k}-mers on the reference sequence in our analyses, it
+may not be the case for highly variable species or if the causal variation at a
+locus results from a large novel insertion.
+
+\subsection*{Use of SV genotypes for GWAS}
+
+In addition to \emph{k}-mers, we also assessed the use of SVs in GWAS analysis
+of soybean.  Although we did find the causal variant as the most significantly
+associated SV at the \emph{W1} locus, results were otherwise rather
+inconclusive.  Indeed, most of the other known or putative causal variants were
+SNPs or indels, and therefore were not represented in the SV datasets. At most
+loci where SVs are known to be the underlying causal variant (e.g.  \emph{I},
+\emph{B}, \emph{Ps}), the analyses based on SVs did not find the causal
+variants, most likely due to the difficulty of calling and genotyping such
+variants \citep{kosugi2019, lemay2022}. In the case of the \textit{Ps} locus,
+this limitation was due to a filtering step in our pipeline rather than to an
+inherent inability to properly call and genotype the causal variant.
+
+Given the current performance of SV genotyping programs, it is not clear
+whether using SVs for GWAS provides any advantage over using \emph{k}-mer- or
+SNP-based GWAS, or a combination of both.  Indeed, in the single case where the
+SV-based analyses identified the causal variant, the \emph{k}-mer-based GWAS
+identified it as well. Moreover, SNP-based GWAS provided better guidance
+in delimiting the signals found at a given locus because of the much higher
+density of markers compared to SVs.  The usefulness of SV-based GWAS compared
+to SNP-based GWAS will largely depend on whether SVs can represent variation
+that is not otherwise in LD with SNPs. In tomato, Domínguez et al. \cite{dominguez2020} found
+that most transposable element insertions associated with phenotypic variation
+in agronomic traits were not tagged by SNPs. On the other hand, in humans,
+Maretty et al. \cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
+is therefore likely that how well SVs can be tagged by SNPs will depend on the
+particular type of SV and on the species. A more efficient approach than
+SV-based GWAS may be to conduct \emph{k}-mer-based GWAS and only then link
+known SVs to significant \emph{k}-mers.  Alternatively, \textit{k}-mer-based
+GWAS can be conducted specifically for \textit{k}-mers that are associated with
+SVs \citep{jayakodi2020}.  Previous studies have obtained noteworthy results by
+using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
+dominguez2020}, however it would be worthwhile to see whether these results
+could be replicated using \emph{k}-mer-based GWAS.
+
+\section*{Conclusions}
+
+In conclusion, we used SV-based and \emph{k}-mer-based GWAS to study ten
+qualitative traits and three quantitative traits in soybean and
+compared the results to conventional SNP-based GWAS. While \emph{k}-mer-based
+GWAS proved to be a powerful approach in pinpointing the causal variation or
+genes associated with known loci, it is unclear whether conducting SV-based
+GWAS is worth the investment. In addition, we used the results from the
+\emph{k}-mer- and SNP-based GWAS to suggest candidate genes for a few classical
+loci that have yet to be cloned. Based on our results, we believe that an
+optimal workflow may involve conducting SNP-based and \emph{k}-mer-based GWAS
+in parallel to identify significant signals and candidate genes, potentially in
+combination with SV datasets. Given the large number of samples for which WGS
+data is now available in several species, applying \emph{k}-mer-based GWAS to
+leverage already existing sequencing and phenotypic data appears feasible and
+promising.  As part of our work, we developed several computational tools that
+should help other researchers with the downstream analysis of
+\emph{k}-mer-based GWAS. However, much work remains to be done in
+developing state-of-the-art methods for the downstream analysis of significant
+\emph{k}-mers. In particular, better approaches are needed for linking
+\emph{k}-mers to sequence variation and moving from presence/absence-based
+methods to count-based methods.
+
+\section*{Supplemental Files}
+
+\subsection*{Supplemental file 1 --- Supplemental file 1.pdf}
+Supplemental methods and results,
+including Supplemental Tables S1 to S11
+and Supplemental Figures S1 to S60 (PDF 12.9 MB)
+
+\subsection*{Supplemental file 2 --- Supplemental file 2.csv}
+Metadata on the sequencing runs retrieved
+from the Sequence Read Archive (SRA) (CSV 71 KB)
+
+\subsection*{Supplemental file 3 --- Supplemental file 3.csv}
+Metadata on individual accessions following mapping
+and comparison to SoySNP50K data (CSV 36KB)
+
+\subsection*{Supplemental file 4 --- Supplemental file 4.csv}
+Phenotypic dataset used for GWAS analyses (CSV 135 KB)
+
+\section*{Acknowledgements}%% if any
+
+We would like to thank Jonas A. Sibbesen for providing very helpful comments
+on the methodology used and on a previous version of this manuscript.
+We also thank Yoav Voichek for help regarding \textit{k}-mer GWAS
+analysis as well as Brian Boyle and Martine Jean for valuable comments
+regarding the methodology. We are thankful to the Digital Research Alliance of
+Canada for using their high-performance computing servers.
+
+\section*{Funding}%% if any
+
+This work was supported by the SoyaGen grant (\url{https://www.soyagen.ca})
+awarded to F. Belzile and R. Bélanger, and funded by Génome Québec, Genome
+Canada, the government of Canada, the Ministère de l'Économie, Science et
+Innovation du Québec, Semences Prograin Inc., Syngenta Canada Inc., Sevita
+Genetics, Coop Fédérée, Grain Farmers of Ontario, Saskatchewan Pulse Growers,
+Manitoba Pulse \& Soybean Growers, the Canadian Field Crop Research Alliance
+and Producteurs de grains du Québec. M-A. Lemay has been supported by a NSERC
+Canada Vanier Graduate Scholarship, a FRQNT doctoral B2X scholarship, a NSERC
+Michael Smith Foreign Study Supplement, and a scholarship from the
+AgroPhytoSciences NSERC CREATE Training Program. None of the funding bodies
+were involved in study design, data acquisition, data analysis, interpretation
+of the results, or manuscript writing.
+
+\section*{Abbreviations}%% if any
+BAM: Binary alignment map format;
+CDW: corrected dry weight;
+CHS: chalcone synthase;
+CNV: copy number variant, copy number variation;
+GRIN: Germplasm Resources Information Network;
+GWAS: Genome-wide association study;
+HPS: Hydrophobic protein from soybean;
+LD: Linkage disequilibrium;
+MAF: Minor allele frequency;
+MATE: Multidrug and toxic compound extrusion;
+MLM: Mixed linear model;
+NCBI: National Center for Biotechnology Information;
+SNP: Single-nucleotide polymorphism;
+SRA: Sequence Read Archive;
+SV: Structural variant, structural variation;
+VCF: Variant call format;
+WGS: Whole-genome (re)sequencing
+
+\section*{Availability of data and materials}
+
+Some of the datasets generated as part of this work are available on figshare
+\citep{kmer-gwas-figshare}.
+
+The Illumina sequencing data used for the analyses are listed in Supplemental file 2
+and available through NCBI BioProjects PRJNA257011 \cite{PRJNA257011}, PRJNA289660 \cite{PRJNA289660}
+and PRJNA639876 \cite{PRJNA639876}.
+
+The high-quality assemblies generated by Liu et al. \cite{liu2020-pangenome} are available
+on the Genome Warehouse through Accession Number PRJCA002030 \cite{PRJCA002030}.
+
+The assemblies of ZH13, W05 and Lee are available on SoyBase
+(\url{https://soybase.org/GlycineBlastPages/blast_descriptions.php}).
+
+The \textit{I} locus contig assembled by Tuteja and Vodkin \cite{tuteja2008} is available from
+NCBI (\url{https://www.ncbi.nlm.nih.gov/nuccore/EF623854}).
+
+SoySNP50K genotype calls are available from SoyBase (\url{https://soybase.org/snps/}).
+
+The SVs identified from Oxford Nanopore data by Lemay et al. \cite{lemay2022} are available
+on figshare \citep{soybean-sv-figshare}.
+
+The following software used in this work is available on GitHub:
+
+\begin{itemize}
+	\item Code used for the analyses:
+		\url{https://github.com/malemay/soybean_kmer_gwas}
+	\item katcher software for retrieving reads containing \textit{k}-mers:
+		\url{https://github.com/malemay/katcher}
+	\item gwask R package used for processing the output of GAPIT and
+		\textit{k}-mer GWAS analyses, and for plotting:
+		\url{https://github.com/malemay/gwask}
+	\item Software for computing the LD between \textit{k}-mers
+		based on presence/absence: \url{https://github.com/malemay/kmers_ld}
+	\item Forked svmu version that was used for the analyses shown in this
+		paper: \url{https://github.com/malemay/svmu}
+	\item svmutools R package for converting the output of svmu to VCF
+		format: \url{https://github.com/malemay/svmutools}
+\end{itemize}
+
+The code associated with the results presented in this manuscript is archived
+on figshare \citep{kmer-gwas-code}.
+
+\section*{Ethics approval and consent to participate}%% if any
+Not applicable.
+
+\section*{Competing interests}
+The authors declare that they have no competing interests.
+
+\section*{Consent for publication}%% if any
+Not applicable.
+
+\section*{Authors' contributions}
+Conception and design of the study: MAL, MDR, RB, FB.
+Phenotypic data acquisition (resistance to \textit{Phytophthora sojae}): MDR, RB.
+Data analysis: MAL.
+Data interpretation: MAL, FB.
+Software writing: MAL.
+Manuscript drafting: MAL, FB.
+All authors have revised the manuscript and approved its submission.
+
+\bibliographystyle{genome_research}
+\bibliography{references}
+
+% Table 1 with results on all loci with cloned genes
 \section*{Tables}
-
-% Table 1 with results on all loci with cloned genes
 \begin{landscape}
-\begin{table}[p]
+\begin{table}
 	\caption[Performance of three GWAS approaches at detecting causal genes/variants]%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Additional Files              %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2023-04-04 additional_files/additional_file_1.tex
Invert Methods and Results and change figure numbers accordingly
Change the word Additional to Supplemental to be concordant with The Plant Genome's standard
Format all parts of the manuscript and supplemental material (except references) as required by The Plant Genome

@@
-\documentclass[12pt]{report}
+\documentclass[12pt]{article}
 
@@
 \usepackage{setspace}
+
+\title{Supplementary material to ``\textit{k}-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean''}
+\date{}
+
+\author{Marc-André Lemay \and
+Maxime de Ronne \and
+Richard Bélanger \and
+François Belzile}
 
@@
 
-\chapter*{Supplemental data to ``\textit{k}-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean''}
-\label{supdata}
+\maketitle
 
@@
 \setcounter{page}{1}
-
-% List of authors
-\textbf{Authors}: Marc-André Lemay, Maxime de Ronne, Richard Bélanger, François Belzile
 
@@
 The concordance between WGS data and SoySNP50K genotypes is included in
-Additional file 3 and shown in Figure \ref{concordance-histogram}.
+Supplemental file 3 and shown in Figure \ref{concordance-histogram}.
 
@@
 
+\gwastableone{FLOWER.COLOR}{flower color}
+\gwastabletwo{PUBESCENCE.COLOR}{pubescence color}
+\gwastabletwo{SEED.COAT.COLOR}{seed coat color}
+\gwastabletwo{STEM.TERMINATION.TYPE}{stem termination type}
+\gwastablethree{HILUM.COLOR}{hilum color}
+\gwastabletwo{POD.COLOR}{pod color}
+\gwastabletwo{PUBESCENCE.FORM}{pubescence form}
+\gwastableone{PUBESCENCE.DENSITY}{pubescence density}
+\gwastablethree{SEED.COAT.LUSTER}{seed coat luster}
+\gwastableone{MATURITY.GROUP}{maturity group}
+
 % Table with all signals
@@
 
-\gwastableone{FLOWER.COLOR}{flower color}
-\gwastabletwo{PUBESCENCE.COLOR}{pubescence color}
-\gwastabletwo{SEED.COAT.COLOR}{seed coat color}
-\gwastabletwo{STEM.TERMINATION.TYPE}{stem termination type}
-\gwastablethree{HILUM.COLOR}{hilum color}
-\gwastabletwo{POD.COLOR}{pod color}
-\gwastabletwo{PUBESCENCE.FORM}{pubescence form}
-\gwastableone{PUBESCENCE.DENSITY}{pubescence density}
-\gwastablethree{SEED.COAT.LUSTER}{seed coat luster}
-\gwastableone{MATURITY.GROUP}{maturity group}
-
 % Supplemental figures
@@
 % Figures mentioned in the main text
+
+% Histogram with concordance rates
+\begin{cfigure}
+	\includegraphics[width = 6in]{concordance_histogram}
+	\caption{Histogram of concordance rates between genotypes derived from WGS data and SoySNP50K genotypes for 385 soybean accessions.
+	The vertical line indicates the 0.9 concordance threshold that was used for filtering out mismatching samples.}
+	\label{concordance-histogram}
+\end{cfigure}
+
+\clearpage
+
 % Figures for flower color
@@
 
-% Histogram with concordance rates
-\begin{cfigure}
-	\includegraphics[width = 6in]{concordance_histogram}
-	\caption{Histogram of concordance rates between genotypes derived from WGS data and SoySNP50K genotypes for 385 soybean accessions.
-	The vertical line indicates the 0.9 concordance threshold that was used for filtering out mismatching samples.}
-	\label{concordance-histogram}
-
-\end{cfigure}
-
-\clearpage
-
 % Flower color

### 2023-04-04 additional_files/manuscript.tex
Format all parts of the manuscript and supplemental material (except references) as required by The Plant Genome

@@
 \RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
-\RequirePackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage[utf8]{inputenc} %unicode support
@@
 
-\title{k-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean}
+\title{\textit{k}-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean}
 \date{}
@@
 Richard Bélanger\textsuperscript{1,2,3} \and
-François Belzile\textsuperscript{1,2,3}}
+François Belzile\textsuperscript{1,2,3,*}}
 
@@
 \textsuperscript{3} Centre de recherche et d’innovation sur les végétaux, Université Laval, Québec, QC, Canada
+
+\textsuperscript{*} Corresponding author: francois.belzile@fsaa.ulaval.ca
+
+\clearpage
+
+\section*{Core ideas}
+
+\begin{itemize}
+	\item We compared SNP-, SV- and \textit{k}-mer-based GWAS approaches
+		on 13 traits using a population of 363 cultivated soybean lines.
+	\item \textit{k}-mer-based GWAS performed better than the other two approaches
+		at pinpointing causal variants and genes.
+	\item We developed computational tools to facilitate the downstream analysis
+		of \textit{k}-mer-based GWAS.
+	\item Performing SV-based GWAS in soybean may not be worth the investment
+		of time and computational resources.
+\end{itemize}
+
+\section*{Abbreviations}
+BAM, Binary alignment map format;
+CDW, corrected dry weight;
+CHS, chalcone synthase;
+CNV, copy number variant, copy number variation;
+GRIN, Germplasm Resources Information Network;
+GWAS, Genome-wide association study;
+HPS, Hydrophobic protein from soybean;
+LD, Linkage disequilibrium;
+MAF, Minor allele frequency;
+MATE, Multidrug and toxic compound extrusion;
+MLM, Mixed linear model;
+NCBI, National Center for Biotechnology Information;
+SNP, Single-nucleotide polymorphism;
+SRA, Sequence Read Archive;
+SV, Structural variant, structural variation;
+VCF, Variant call format;
+WGS, Whole-genome (re)sequencing
 
@@
 problem, for example by including structural variants (SVs).
-
 In this study, we tested the potential of SV- and \emph{k}-mer-based GWAS in
@@
 computational resources required to genotype SVs at population scale.
-
 Despite promising results, significant challenges remain regarding the
@@
 
-\section*{Keywords}
-Genome-wide association studies
-\textit{k}-mer GWAS
-Soybean genomics
-Crop genomics
-\textit{Phytophthora sojae}
-Structural variation
+\clearpage
 
@@
 \end{enumerate}
+
+\clearpage
 
@@
 GitHub (\url{https://github.com/malemay/gwask}).
+
+\clearpage
 
@@
 
+\clearpage
+
 \section*{Conclusions}
@@
 
-\section*{Supplemental Files}
-
-\subsection*{Supplemental file 1 --- Supplemental file 1.pdf}
-Supplemental methods and results,
-including Supplemental Tables S1 to S11
-and Supplemental Figures S1 to S60 (PDF 12.9 MB)
-
-\subsection*{Supplemental file 2 --- Supplemental file 2.csv}
-Metadata on the sequencing runs retrieved
-from the Sequence Read Archive (SRA) (CSV 71 KB)
-
-\subsection*{Supplemental file 3 --- Supplemental file 3.csv}
-Metadata on individual accessions following mapping
-and comparison to SoySNP50K data (CSV 36KB)
-
-\subsection*{Supplemental file 4 --- Supplemental file 4.csv}
-Phenotypic dataset used for GWAS analyses (CSV 135 KB)
-
-\section*{Acknowledgements}%% if any
+\clearpage
+
+\section*{Acknowledgments}
 
@@
 
-\section*{Funding}%% if any
-
-This work was supported by the SoyaGen grant (\url{https://www.soyagen.ca})
-awarded to F. Belzile and R. Bélanger, and funded by Génome Québec, Genome
-Canada, the government of Canada, the Ministère de l'Économie, Science et
-Innovation du Québec, Semences Prograin Inc., Syngenta Canada Inc., Sevita
-Genetics, Coop Fédérée, Grain Farmers of Ontario, Saskatchewan Pulse Growers,
-Manitoba Pulse \& Soybean Growers, the Canadian Field Crop Research Alliance
-and Producteurs de grains du Québec. M-A. Lemay has been supported by a NSERC
-Canada Vanier Graduate Scholarship, a FRQNT doctoral B2X scholarship, a NSERC
-Michael Smith Foreign Study Supplement, and a scholarship from the
-AgroPhytoSciences NSERC CREATE Training Program. None of the funding bodies
-were involved in study design, data acquisition, data analysis, interpretation
-of the results, or manuscript writing.
-
-\section*{Abbreviations}%% if any
-BAM: Binary alignment map format;
-CDW: corrected dry weight;
-CHS: chalcone synthase;
-CNV: copy number variant, copy number variation;
-GRIN: Germplasm Resources Information Network;
-GWAS: Genome-wide association study;
-HPS: Hydrophobic protein from soybean;
-LD: Linkage disequilibrium;
-MAF: Minor allele frequency;
-MATE: Multidrug and toxic compound extrusion;
-MLM: Mixed linear model;
-NCBI: National Center for Biotechnology Information;
-SNP: Single-nucleotide polymorphism;
-SRA: Sequence Read Archive;
-SV: Structural variant, structural variation;
-VCF: Variant call format;
-WGS: Whole-genome (re)sequencing
+\section*{Conflict of interest}
+The authors declare no conflict of interest.
+
+\section*{ORCID}
+
+\begin{description}
+	\item[Marc-André Lemay] \url{https://orcid.org/0000-0003-2484-8447}
+	\item[Maxime de Ronne] \url{https://orcid.org/0000-0003-4062-3512}
+	\item[Richard Bélanger] \url{https://orcid.org/0000-0001-8549-975X}
+	\item[François Belzile] \url{https://orcid.org/0000-0001-6893-4529}
+\end{description}
+
+\section*{Supplemental material}
+
+\begin{description}
+	\item[Supplemental file 1.pdf] Supplemental methods and results, including
+		Supplemental Tables S1 to S11 and Supplemental Figures S1 to
+		S60.
+
+	\item[Supplemental file 2.csv] Metadata on the sequencing runs retrieved
+		from the Sequence Read Archive (SRA).
+
+	\item[Supplemental file 3.csv] Metadata on individual accessions following
+		mapping and comparison to SoySNP50K data.
+
+	\item[Supplemental file 4.csv] Phenotypic dataset used for GWAS analyses.
+\end{description}
 
@@
 
-\section*{Ethics approval and consent to participate}%% if any
-Not applicable.
-
-\section*{Competing interests}
-The authors declare that they have no competing interests.
-
-\section*{Consent for publication}%% if any
-Not applicable.
-
-\section*{Authors' contributions}
-Conception and design of the study: MAL, MDR, RB, FB.
-Phenotypic data acquisition (resistance to \textit{Phytophthora sojae}): MDR, RB.
-Data analysis: MAL.
+\section*{Funding}%% if any
+
+This work was supported by the SoyaGen grant (\url{https://www.soyagen.ca})
+awarded to F. Belzile and R. Bélanger, and funded by Génome Québec, Genome
+Canada, the government of Canada, the Ministère de l'Économie, Science et
+Innovation du Québec, Semences Prograin Inc., Syngenta Canada Inc., Sevita
+Genetics, Coop Fédérée, Grain Farmers of Ontario, Saskatchewan Pulse Growers,
+Manitoba Pulse \& Soybean Growers, the Canadian Field Crop Research Alliance
+and Producteurs de grains du Québec. M-A. Lemay has been supported by a NSERC
+Canada Vanier Graduate Scholarship, a FRQNT doctoral B2X scholarship, a NSERC
+Michael Smith Foreign Study Supplement, and a scholarship from the
+AgroPhytoSciences NSERC CREATE Training Program. None of the funding bodies
+were involved in study design, data acquisition, data analysis, interpretation
+of the results, or manuscript writing.
+
+\section*{Author contributions}
+Conceptualization: MAL, MDR, RB, FB.
+Data curation: MAL, MDR.
+Methodology: MAL, FB.
+Investigation: MAL, MDR.
+Formal analysis: MAL.
+Funding acquisition: RB, FB.
 Data interpretation: MAL, FB.
-Software writing: MAL.
-Manuscript drafting: MAL, FB.
-All authors have revised the manuscript and approved its submission.
+Software: MAL.
+Supervision: RB, FB.
+Visualization: MAL.
+Writing -- original draft: MAL, FB.
+Writing -- review \& editing: MAL, MDR, RB, FB.
+
+\clearpage
 
@@
 
+\clearpage
+
+\section*{Tables}
+
 % Table 1 with results on all loci with cloned genes
-\section*{Tables}
 \begin{landscape}

### 2023-04-04 additional_files/manuscript.tex
Format references for submission to The Plant Genome
Last changes before re-reading manuscript prior to submission; shorten abstract and adjust citations

@@
 \usepackage{amsthm,amsmath}
-\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
 \usepackage[hidelinks]{hyperref}
@@
 \usepackage[margin=1in]{geometry} % leaves more space for text body
+\usepackage[natbibapa]{apacite}
 
@@
 
-%%% Begin ...
 \begin{document}
@@
 \textsuperscript{3} Centre de recherche et d’innovation sur les végétaux, Université Laval, Québec, QC, Canada
+
+Received \_\_\_\_\_\_\_\_\_\_\_
 
@@
 their power, GWAS frequently fail to pinpoint the causal variant or the gene
-controlling a given trait in crop species. Assessing genetic
-variants beyond single-nucleotide polymorphisms (SNPs) could alleviate this
-problem, for example by including structural variants (SVs).
-In this study, we tested the potential of SV- and \emph{k}-mer-based GWAS in
-soybean by applying these methods to 13 traits.  We also performed
-conventional GWAS analysis based on SNPs and small indels.  We assessed the performance of each GWAS approach based on
-results at loci for which the causal genes or variants were known from
-previous genetic studies. We found that \emph{k}-mer-based GWAS was the
-most versatile approach and the best at pinpointing causal variants or
-candidate genes based on the most significantly associated
-\emph{k}-mers.  Moreover, \textit{k}-mer-based analyses identified
-promising candidate genes for loci related to pod color, pubescence
-form, and resistance to \textit{Phytophthora sojae}. In
-our dataset, SV-based GWAS did not add value compared to
-\textit{k}-mer-based GWAS and may not be worth the time and
-computational resources required to genotype SVs at population scale.
-Despite promising results, significant challenges remain regarding the
-downstream analysis of \emph{k}-mer-based GWAS.  Notably, better
+controlling a given trait in crop species. Assessing genetic variants other than
+single-nucleotide polymorphisms (SNPs) could alleviate this problem.  In this
+study, we tested the potential of SV- and \emph{k}-mer-based GWAS in soybean by
+applying these methods as well as conventional SNP/indel-based GWAS to 13
+traits. We assessed the performance of each GWAS approach based on loci for
+which the causal genes or variants were known from previous genetic studies. We
+found that \emph{k}-mer-based GWAS was the most versatile approach and the best
+at pinpointing causal variants or candidate genes.  Moreover,
+\textit{k}-mer-based analyses identified promising candidate genes for loci
+related to pod color, pubescence form, and resistance to \textit{Phytophthora
+sojae}. In our dataset, SV-based GWAS did not add value compared to
+\textit{k}-mer-based GWAS and may not be worth the time and computational
+resources invested.  Despite promising results, significant challenges remain
+regarding the downstream analysis of \emph{k}-mer-based GWAS.  Notably, better
 methods are needed to associate significant \emph{k}-mers with sequence
-variation. Together, our results suggest that coupling \emph{k}-mer-
-and SNP/indel-based GWAS is a powerful approach for discovering
-candidate genes in crop species.
+variation. Our results suggest that coupling \emph{k}-mer- and SNP/indel-based
+GWAS is a powerful approach for discovering candidate genes in crop species.
 
@@
 
-\section*{Background}
+\section*{Introduction}
 Genome-wide association studies (GWAS) are analytical approaches that detect
@@
 interest and the genotypes of variants distributed throughout the
-genome \cite{tam2019}. This approach has been used for over 15 years
-\cite{visscher2012} and has enabled significant advances in our understanding
+genome \citep{tam2019}. This approach has been used for over 15 years
+\citep{visscher2012} and has enabled significant advances in our understanding
 of the genetic architecture of traits involved in human health
-\cite[e.g.][]{wellcome2007gwas} and crop \cite{tibbs2021} and animal
-\cite{ma2019} production, among others.
+\citep[e.g.][]{wellcome2007gwas} and crop \citep{tibbs2021} and animal
+\citep{ma2019} production, among others.
 
@@
 GWAS analyses frequently fail to pinpoint the gene(s) associated with a trait and
-the causal variant(s) involved \cite{tam2019}. One possible reason for this
+the causal variant(s) involved \citep{tam2019}. One possible reason for this
 is that many GWAS analyses do not include the causal variant itself in their variant dataset
-\cite{tibbs2021}. This could be either because the analysis
+\citep{tibbs2021}. This could be either because the analysis
 is restricted to a particular type of variant, such as single-nucleotide
@@
 association panel is limited by the technology being used
-\cite[e.g.][]{bandillo2015, sonah2015}. Despite this limitation, variants
+\citep[e.g.][]{bandillo2015, sonah2015}. Despite this limitation, variants
 located near the gene controlling the trait often do appear statistically
 associated with the phenotype through linkage disequilibrium (LD) with the
-causal variant \cite{korte2013}.  A typical workflow then involves the
+causal variant \citep{korte2013}.  A typical workflow then involves the
 identification of candidate genes within the haplotype blocks defined by the
 statistically associated markers and follow-up functional analyses to confirm
-that a given gene controls the trait \cite[e.g.][]{wang2018, liu2020ps}.
+that a given gene controls the trait \citep[e.g.][]{wang2018, liu2020ps}.
 
@@
 SVs include any variant involving a difference of at least 50 nucleotides, such as
-deletions, insertions, or inversions \cite{ho2020}. Such variants are known to
+deletions, insertions, or inversions \citep{ho2020}. Such variants are known to
 have large phenotypic impacts by disrupting coding or regulatory sequences
-\cite{marroni2014}, and have been identified as causal variants in traits such
-as resistance to soybean cyst nematode \cite{cook2012}, aluminum tolerance in
-wheat \cite{maron2013}, and branching in maize \cite{studer2011}.  SVs have
+\citep{marroni2014}, and have been identified as causal variants in traits such
+as resistance to soybean cyst nematode \citep{cook2012}, aluminum tolerance in
+wheat \citep{maron2013}, and branching in maize \citep{studer2011}.  SVs have
 already been used in GWAS with promising results
-\cite[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
+\citep[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
 their use in GWAS is the exception rather than the norm.
@@
 often need to be tailored to the study species and data available
-\cite{alkan2011, ho2020}. Even then, assessments of the performance of SV
+\citep{alkan2011, ho2020}. Even then, assessments of the performance of SV
 discovery and genotyping generally reveal subpar sensitivity and precision
-\cite[e.g.][]{cameron2019, chaisson2019, kosugi2019}. Moreover, despite recent
+\citep[e.g.][]{cameron2019, chaisson2019, kosugi2019}. Moreover, despite recent
 improvements in computational approaches for genotyping SVs
-\cite[e.g.][]{siren2021, ebler2022}, such approaches remain relatively
+\citep[e.g.][]{siren2021, ebler2022}, such approaches remain relatively
 inaccurate compared to SNP genotyping workflows.
@@
 of variants in GWAS without genotyping SVs at population scale could be useful.
-Voichek and Weigel \cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
+\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
 represent such an alternative method. This approach relies on the presence or
@@
 few studies have applied this \emph{k}-mer-based GWAS approach
-\cite[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
-developed \cite{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
+\citep[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
+developed \citep{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
 such approaches have not been widely adopted yet.
@@
 based on SNPs and indels. In order to test these methods, we analyzed ten
-qualitative traits \cite{bandillo2017} and two quantitative traits
-\cite{bandillo2015} that have been previously studied using a SNP array
-developed for soybean \cite{song2013}. In addition, we analyzed a quantitative
+qualitative traits \citep{bandillo2017} and two quantitative traits
+\citep{bandillo2015} that have been previously studied using a SNP array
+developed for soybean \citep{song2013}. In addition, we analyzed a quantitative
 trait (horizontal resistance to \emph{Phytophthora sojae}) that has been
 previously studied using whole-genome sequencing (WGS) data
-\cite{deronne2022}.  Since the underlying genes and causal variants are
+\citep{deronne2022}.  Since the underlying genes and causal variants are
 already known for several of these traits, they provide a good test case for
@@
 
-\section*{Methods}
+\section*{Materials and methods}
 
@@
 and extracted paired-end reads using the fastq-dump command v. 2.9.6
-\cite{sratoolkit} with the
+(\citeauthor{sratoolkit}) with the
 option \texttt{-\-split-3}.
@@
 adapters using bbduk from BBtools v. 38.25
-\cite{bbtools}. Reads
+(\citeauthor{bbtools}). Reads
 were then mapped to the reference assembly version 4 of \emph{G. max}
@@
 accession using bamaddrg
-\cite{bamaddrg}. We used the
+(\citeauthor{bamaddrg}). We used the
 resulting BAM files for all downstream analyses requiring aligned reads.
@@
 		to call SVs in the population following a pipeline outlined
-		by Lemay et al. \cite{lemay2022}. In brief, we combined information from
+		by \cite{lemay2022}. In brief, we combined information from
 		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
@@
 		a set of candidate SVs from the mapped Illumina data.
-	\item We used SVs discovered by Lemay et al. \cite{lemay2022} from Oxford Nanopore
+	\item We used SVs discovered by \cite{lemay2022} from Oxford Nanopore
 		sequencing data of 17 Canadian soybean cultivars.
 	\item We called SVs from 26 high-quality genome assemblies published by
-		Liu et al. \cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
+		\cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
 		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
-		We used methods similar to those of Liu et al. \cite{liu2020-pangenome} to call
+		We used methods similar to those of \cite{liu2020-pangenome} to call
 		SVs from the comparison of these assemblies to that of Williams82.
@@
 
-We used the approach outlined by Voichek and Weigel \cite{voichek2020} for \emph{k}-mer-based GWAS
+We used the approach outlined by \cite{voichek2020} for \emph{k}-mer-based GWAS
 and their recommendations on the
@@
 
-We analyzed ten qualitative traits previously studied by Bandillo et al. \cite{bandillo2017}: flower
+We analyzed ten qualitative traits previously studied by \cite{bandillo2017}: flower
 color, pubescence color, seed coat color, stem termination type, hilum color,
@@
 to numerical values for GWAS analysis following methods similar to those of
-Bandillo et al. \cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
+\cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
 using targeted subsets of the observed phenotypes in order to focus on specific
@@
 Our analysis included 340 accessions whose phenotypic data
-was already published by de Ronne et al. \cite{deronne2022} as well as 49 additional accessions for
+was already published by \cite{deronne2022} as well as 49 additional accessions for
 which sequencing data was not yet available when their study was conducted.
@@
 approach.  We used a randomization approach to determine the 5\% family-wise
-error-rate threshold for each GWAS analysis as described by Voichek and Weigel \cite{voichek2020}.
+error-rate threshold for each GWAS analysis as described by \cite{voichek2020}.
 To compute this threshold, we permuted the phenotypic observations 100 times
@@
 We conducted GWAS based on \emph{k}-mers using the method developed by
-Voichek and Weigel \cite{voichek2020}. In brief, the
+\cite{voichek2020}. In brief, the
 analysis was conducted in two steps. The first step used an approximate model
@@
 
-Zabala and Vodkin \cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
+\cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
 Glyma.13g\-072100 as associated with the \textit{W1} locus for flower color. In
@@
 
-Zabala and Vodkin \cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
+\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
 \textit{T} locus for pubescence color and found a 1-bp deletion as a putative
@@
 approaches (Supplemental Figure S15).
-Wang et al. \cite{wang2018} identified Glyma.01g198500 as the gene associated with
+\cite{wang2018} identified Glyma.01g198500 as the gene associated with
 this locus and found an A>G SNP affecting transcript splicing as the causal
@@
 We did identify one of the SNPs suggested as putative causal variants for this
-locus by Tian et al. \cite{tian2010} and Liu et al. \cite{liu2010}, however this SNP was only the
+locus by \cite{tian2010} and \cite{liu2010}, however this SNP was only the
 135\textsuperscript{th} most associated marker found by Platypus and the
@@
 S22 and S23).
-Gillman et al. \cite{gillman2011} identified this R2R3 MYB transcription factor as the
+\cite{gillman2011} identified this R2R3 MYB transcription factor as the
 molecular basis for the \textit{R} locus and documented four different
@@
 well as Glyma.09g235100 and two others are part of a tandem array of four R2R3
-MYB genes considered as candidates by Gillman et al. \cite{gillman2011}, and of which
+MYB genes considered as candidates by \cite{gillman2011}, and of which
 Glyma.09g235100 was the only one expressed in the seed coat. Given the
-convincing evidence provided by Gillman et al. \cite{gillman2011}, it is most likely that
+convincing evidence provided by \cite{gillman2011}, it is most likely that
 Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
@@
 loss of function at the \emph{R} locus.  Still, we observed two causal variants
-previously documented by Gillman et al. \cite{gillman2011} in our dataset (Supplemental Figures
+previously documented by \cite{gillman2011} in our dataset (Supplemental Figures
 S24 and S25).
@@
 \ref{pubescence-density-main-figure}c), and therefore possibly corresponded to the
-25.6-kb region identified as the causal CNV by Liu et al. \cite{liu2020ps}. We found that
+25.6-kb region identified as the causal CNV by \cite{liu2020ps}. We found that
 reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
@@
 
-The difference in size between the CNV reported by Liu et al. \cite{liu2020ps} and our
+The difference in size between the CNV reported by \cite{liu2020ps} and our
 results may be due to the presence of some unresolved sequence (a stretch of
@@
 the seed coat luster trait. Seed coat luster has been shown by
-Gijzen et al. \cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
+\cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
 protein (HPS) at the seed surface. Copy number variation of this gene at the
@@
 Both the \textit{L1} and the \textit{L2} loci for pod color have yet to be
-cloned. While He et al. \cite{he2015} proposed the gene Glyma.19g101700 as likely
+cloned. While \cite{he2015} proposed the gene Glyma.19g101700 as likely
 corresponding to \textit{L1} based on fine mapping and expression analysis, our
-results as well as those reported by Bandillo et al. \cite{bandillo2017} do not support this
+results as well as those reported by \cite{bandillo2017} do not support this
 gene as a candidate due to its location $\sim$3 Mb away from GWAS signals. Our
@@
 The two known loci for pubescence form in soybean, \textit{Pa1} and
-\textit{Pa2}, have also not been cloned yet. Nevertheless, Gilbert \cite{gilbert2017}
+\textit{Pa2}, have also not been cloned yet. Nevertheless, \cite{gilbert2017}
 has suggested Glyma.12g213900 as a candidate for \textit{Pa1} based on data
-published by Bandillo et al. \cite{bandillo2017}. This gene encodes a MYB transcription factor
+published by \cite{bandillo2017}. This gene encodes a MYB transcription factor
 annotated as playing a role in trichome branching. Indeed, the most significant
@@
 
-We used corrected dry weight (CDW) as described by de Ronne et al. \cite{deronne2022} as a
+We used corrected dry weight (CDW) as described by \cite{deronne2022} as a
 measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
 overlapping the genomic region reported by these authors on Gm15 using all three GWAS
-approaches (Supplemental Figure S41). de Ronne et al. \cite{deronne2022}
+approaches (Supplemental Figure S41). \cite{deronne2022}
 proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
@@
 result in a decreased ability to identify causal variants or genes. For
-example, Bandillo et al. \cite{bandillo2017} noted that observed phenotypes for stem
+example, \cite{bandillo2017} noted that observed phenotypes for stem
 termination type in the GRIN database may not correspond to the expected
@@
 limitation of this approach. Although the method developed by
-Voichek and Weigel \cite{voichek2020} is in itself relatively simple and computationally
+\cite{voichek2020} is in itself relatively simple and computationally
 inexpensive, downstream analysis and interpretation of the output is
@@
 to SNP-based GWAS will largely depend on whether SVs can represent variation
-that is not otherwise in LD with SNPs. In tomato, Domínguez et al. \cite{dominguez2020} found
+that is not otherwise in LD with SNPs. In tomato, \cite{dominguez2020} found
 that most transposable element insertions associated with phenotypic variation
 in agronomic traits were not tagged by SNPs. On the other hand, in humans,
-Maretty et al. \cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
+\cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
 is therefore likely that how well SVs can be tagged by SNPs will depend on the
@@
 
-\section*{Availability of data and materials}
+\section*{Data availability}
 
@@
 The Illumina sequencing data used for the analyses are listed in Supplemental file 2
-and available through NCBI BioProjects PRJNA257011 \cite{PRJNA257011}, PRJNA289660 \cite{PRJNA289660}
-and PRJNA639876 \cite{PRJNA639876}.
-
-The high-quality assemblies generated by Liu et al. \cite{liu2020-pangenome} are available
-on the Genome Warehouse through Accession Number PRJCA002030 \cite{PRJCA002030}.
+and available through NCBI BioProjects \citeauthor{PRJNA257011}, \citeauthor{PRJNA289660}
+and \citeauthor{PRJNA639876}.
+
+The high-quality assemblies generated by \cite{liu2020-pangenome} are available
+on the Genome Warehouse through Accession Number \citeauthor{PRJCA002030}.
 
@@
 
-The \textit{I} locus contig assembled by Tuteja and Vodkin \cite{tuteja2008} is available from
+The \textit{I} locus contig assembled by \cite{tuteja2008} is available from
 NCBI (\url{https://www.ncbi.nlm.nih.gov/nuccore/EF623854}).
@@
 
-The SVs identified from Oxford Nanopore data by Lemay et al. \cite{lemay2022} are available
+The SVs identified from Oxford Nanopore data by \cite{lemay2022} are available
 on figshare \citep{soybean-sv-figshare}.
@@
 
-\section*{Funding}%% if any
+\section*{Funding information}
 
@@
 
-\bibliographystyle{genome_research}
-\bibliography{references}
+\bibliographystyle{apacite}
+\bibliography{references.bib}
 

### 2023-06-13 additional_files/submitted_additional_file_1.tex
Set up repository for automatically generating a track-changes version of the manuscript

--- 2023-04-04 additional_files/additional_file_1.tex
@@
 \usepackage{datatool} % allows fetching key-value pairs from a text file
-\usepackage{natbib}
+\usepackage{apacite}
 \usepackage{setspace}
@@
 for SV discovery using methods similar to those described by
-\cite{lemay2022}.  Briefly, we used four different SV discovery programs or
+\shortciteA{lemay2022}.  Briefly, we used four different SV discovery programs or
 pipelines:
@@
 
-\item We used AsmVar \citep[version of 2015-04-16, ][]{liu2015} to call SVs
+\item We used AsmVar \shortcite<version of 2015-04-16, >[]{liu2015} to call SVs
 	from \emph{de novo} assemblies generated with SOAPdenovo2 v. 2.04
-		\citep{luo2012} and aligned to the reference genome using LAST
-		v. 1047 \citep{kielbasa2011}.
-\item We ran Manta v. 1.6.0 \citep{chen2016} in randomly selected batches: 77 batches of
+		\shortcite{luo2012} and aligned to the reference genome using LAST
+		v. 1047 \shortcite{kielbasa2011}.
+\item We ran Manta v. 1.6.0 \shortcite{chen2016} in randomly selected batches: 77 batches of
 	5 samples and 1 batch of 4 samples. We used the candidate SVs
 		identified from each of the batches for downstream analyses.
-\item We used smoove v. 0.2.4 \citep{pedersen2019} to obtain a dataset of candidate SVs
+\item We used smoove v. 0.2.4 \shortcite{pedersen2019} to obtain a dataset of candidate SVs
 	using the following sequence of commands:
@@
 		call}.
-	\item We used SvABA v. 1.1.3 \citep{wala2018} to call SVs from each of the
+	\item We used SvABA v. 1.1.3 \shortcite{wala2018} to call SVs from each of the
 	samples separately. We classified the SVs called as breakends by SvABA
@@
 Next, we converted all filtered variants to a sequence-explicit format using
-\texttt{bayesTyperTools convertAllele} \citep{sibbesen2018} and normalized them
-using bcftools \texttt{norm} v. 1.10.2-105 \citep{li2009-samtools}.
+\texttt{bayesTyperTools convertAllele} \shortcite{sibbesen2018} and normalized them
+using bcftools \texttt{norm} v. 1.10.2-105 \shortcite{li2009-samtools}.
 
 A single VCF file was generated from the variants called using Illumina data by
-running SVmerge \citep{wong2010} in two steps. In the first step, we merged
+running SVmerge \shortcite{wong2010} in two steps. In the first step, we merged
 very similar variants among the ones called by the same tool using a
@@
 We used SVs discovered among 17 Canadian soybean cultivars sequenced using
-Oxford Nano\-pore Technologies (ONT) by \cite{lemay2022} as an additional
+Oxford Nano\-pore Technologies (ONT) by \shortciteA{lemay2022} as an additional
 source of candidate SVs. Although this set of cultivars does not overlap the
@@
 accessions to call SVs. We
-retrieved 26 high-quality genomes assembled by \cite{liu2020-pangenome} as well
+retrieved 26 high-quality genomes assembled by \shortciteA{liu2020-pangenome} as well
 as the genome of the cultivar Zhonghuang 13 (ZH13) from the Genome Warehouse
@@
 of \emph{G. soja} accession W05 and \emph{G. max} cultivar Lee from SoyBase
-\citep{grant2010}.
+\shortcite{grant2010}.
 
-We used methods based on \cite{liu2020-pangenome} to call SVs from the
+We used methods based on \shortciteA{liu2020-pangenome} to call SVs from the
 genome assemblies.  First, we used the nucmer program of the MUMmer suite of
-tools v. 4.0.0rc1 \citep{marcais2018} with option \texttt{-c 1000} to align
+tools v. 4.0.0rc1 \shortcite{marcais2018} with option \texttt{-c 1000} to align
 each of the assemblies to the Williams82 assembly. Next, we used the
@@
 The filtered alignments were used to call SVs with a version of the svmu
-program \citep{chakraborty2019} that we forked from the original repository
+program \shortcite{chakraborty2019} that we forked from the original repository
 (\url{https://github.com/malemay/svmu}) in order to improve execution time and
@@
 
-We used Paragraph v. 2.4a \citep{chen2019}
+We used Paragraph v. 2.4a \shortcite{chen2019}
 to generate a set of SV genotypes from the Illumina data of the 389 accessions.
@@
 chromosome end and then padding the variant representation using a script written by
-\cite{hickey2020}.
+\shortciteA{hickey2020}.
 
@@
 for which this could be the case, we compared the genotypes of SNPs obtained
-from the SoySNP50K array \citep{song2013} to those predicted from the Illumina WGS
+from the SoySNP50K array \shortcite{song2013} to those predicted from the Illumina WGS
 data. We did this for the 385 samples that had a match in the GRIN database. Briefly, we
-retrieved SoySNP50K genotype calls from SoyBase \citep{grant2010} and extracted
+retrieved SoySNP50K genotype calls from SoyBase \shortcite{grant2010} and extracted
 all SNPs with a MAF $\geq$ 0.1 in our population. We then translated the positions
@@
 
-In addition to the \textit{W1} locus, \cite{bandillo2017} reported a signal for
+In addition to the \textit{W1} locus, \shortciteA{bandillo2017} reported a signal for
 flower color near the \textit{L1} locus.  None of the GWAS approaches that we
@@
 \textit{p}-value (see Table \ref{signals_table}) and only conjectured as
-associated with flower color by \cite{bandillo2017}.
+associated with flower color by \shortciteA{bandillo2017}.
 
@@
 the \textit{E3} locus, which is more commonly associated with maturity but was
-mentioned as associated with stem termination type by \cite{bandillo2017}.
+mentioned as associated with stem termination type by \shortciteA{bandillo2017}.
 
@@
 None of the three GWAS analyses that we conducted on hilum color found signals
-at the \emph{W1} and \emph{O} loci studied by \cite{bandillo2017} (Figures
+at the \emph{W1} and \emph{O} loci studied by \shortciteA{bandillo2017} (Figures
 \ref{hilum_color_blackbrown_manhattan} \ref{hilum_color_all_manhattan},
@@
 failure to detect these loci is likely due to our small population size
-compared to \cite{bandillo2017} and to the smaller contribution of those loci
+compared to \shortciteA{bandillo2017} and to the smaller contribution of those loci
 to pubescence density in our population.
@@
 significant signals. In addition, it also provided results that were most consistent with the
-analysis previously done by \cite{bandillo2017}.
+analysis previously done by \shortciteA{bandillo2017}.
 
@@
 \ref{seed_coat_luster_dullshiny_manhattan}).  All methods also detected signals
-overlapping the \emph{Hps} locus as defined by \cite{bandillo2017}, however
-results shown by \cite{gijzen2003} and our results suggest that \emph{Hps} and
+overlapping the \emph{Hps} locus as defined by \shortciteA{bandillo2017}, however
+results shown by \shortciteA{gijzen2003} and our results suggest that \emph{Hps} and
 \emph{B} are the same locus.
@@
 Apparent signals on chromosomes Gm20 and Gm09 (which may correspond to the B?
-locus reported by \cite{bandillo2017}) were also detected, but analysis of the
+locus reported by \shortciteA{bandillo2017}) were also detected, but analysis of the
 LD patterns between \emph{k}-mers suggested that these signals are most likely
@@
 
-\cite{bandillo2017} mentioned the \textit{I} locus as possibly associated
+\shortciteA{bandillo2017} mentioned the \textit{I} locus as possibly associated
 with seed coat luster, however we did not find evidence for such an association
@@
 have been reported due to high interest in this trait for soybean breeding
-\citep{chaudhary2015}, we focused on two major QTL on chromosomes Gm15 and Gm20.
+\shortcite{chaudhary2015}, we focused on two major QTL on chromosomes Gm15 and Gm20.
 These loci have often been reported in \emph{G. max} and were notably observed by
-\cite{bandillo2015} on a collection of USDA lines using SoySNP50K chip
+\shortciteA{bandillo2015} on a collection of USDA lines using SoySNP50K chip
 genotypes and GRIN phenotypic data. We identified a signal located on
@@
 \ref{oil_manhattan}). Although this signal did not overlap with the coordinates
-found by \cite{bandillo2017}, it was located only 2 kb away from it and 30 kb
+found by \shortciteA{bandillo2017}, it was located only 2 kb away from it and 30 kb
 away from Glyma.15g049200, the sugar transporter gene identified by
-\cite{zhang2020} as associated with that locus. Therefore, the significantly
+\shortciteA{zhang2020} as associated with that locus. Therefore, the significantly
 associated \emph{k}-mers were not linked to the causal variation at that locus.
@@
 soybean given the negative correlation between seed oil and protein content,
-the two major QTL considered above and reported by \cite{bandillo2015} for seed
+the two major QTL considered above and reported by \shortciteA{bandillo2015} for seed
 oil content are also major QTL for seed protein content. However, none of the
@@
 % Bibliography
-\bibliographystyle{genome_research}              % style de la bibliographie
+\bibliographystyle{apacite}              % style de la bibliographie
 \bibliography{references}                   % production de la bibliographie

### 2023-06-13 additional_files/submitted_manuscript.tex
Set up repository for automatically generating a track-changes version of the manuscript

--- 2023-04-04 additional_files/manuscript.tex
@@
 \usepackage[margin=1in]{geometry} % leaves more space for text body
-\usepackage[natbibapa]{apacite}
+\usepackage{apacite}
 
@@
 interest and the genotypes of variants distributed throughout the
-genome \citep{tam2019}. This approach has been used for over 15 years
-\citep{visscher2012} and has enabled significant advances in our understanding
+genome \shortcite{tam2019}. This approach has been used for over 15 years
+\shortcite{visscher2012} and has enabled significant advances in our understanding
 of the genetic architecture of traits involved in human health
-\citep[e.g.][]{wellcome2007gwas} and crop \citep{tibbs2021} and animal
-\citep{ma2019} production, among others.
+\shortcite<e.g.>[]{wellcome2007gwas} and crop \shortcite{tibbs2021} and animal
+\shortcite{ma2019} production, among others.
 
@@
 GWAS analyses frequently fail to pinpoint the gene(s) associated with a trait and
-the causal variant(s) involved \citep{tam2019}. One possible reason for this
+the causal variant(s) involved \shortcite{tam2019}. One possible reason for this
 is that many GWAS analyses do not include the causal variant itself in their variant dataset
-\citep{tibbs2021}. This could be either because the analysis
+\shortcite{tibbs2021}. This could be either because the analysis
 is restricted to a particular type of variant, such as single-nucleotide
@@
 association panel is limited by the technology being used
-\citep[e.g.][]{bandillo2015, sonah2015}. Despite this limitation, variants
+\shortcite<e.g.>[]{bandillo2015, sonah2015}. Despite this limitation, variants
 located near the gene controlling the trait often do appear statistically
 associated with the phenotype through linkage disequilibrium (LD) with the
-causal variant \citep{korte2013}.  A typical workflow then involves the
+causal variant \shortcite{korte2013}.  A typical workflow then involves the
 identification of candidate genes within the haplotype blocks defined by the
 statistically associated markers and follow-up functional analyses to confirm
-that a given gene controls the trait \citep[e.g.][]{wang2018, liu2020ps}.
+that a given gene controls the trait \shortcite<e.g.>[]{wang2018, liu2020ps}.
 
@@
 SVs include any variant involving a difference of at least 50 nucleotides, such as
-deletions, insertions, or inversions \citep{ho2020}. Such variants are known to
+deletions, insertions, or inversions \shortcite{ho2020}. Such variants are known to
 have large phenotypic impacts by disrupting coding or regulatory sequences
-\citep{marroni2014}, and have been identified as causal variants in traits such
-as resistance to soybean cyst nematode \citep{cook2012}, aluminum tolerance in
-wheat \citep{maron2013}, and branching in maize \citep{studer2011}.  SVs have
+\shortcite{marroni2014}, and have been identified as causal variants in traits such
+as resistance to soybean cyst nematode \shortcite{cook2012}, aluminum tolerance in
+wheat \shortcite{maron2013}, and branching in maize \shortcite{studer2011}.  SVs have
 already been used in GWAS with promising results
-\citep[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
+\shortcite<e.g.>[]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
 their use in GWAS is the exception rather than the norm.
@@
 often need to be tailored to the study species and data available
-\citep{alkan2011, ho2020}. Even then, assessments of the performance of SV
+\shortcite{alkan2011, ho2020}. Even then, assessments of the performance of SV
 discovery and genotyping generally reveal subpar sensitivity and precision
-\citep[e.g.][]{cameron2019, chaisson2019, kosugi2019}. Moreover, despite recent
+\shortcite<e.g.>[]{cameron2019, chaisson2019, kosugi2019}. Moreover, despite recent
 improvements in computational approaches for genotyping SVs
-\citep[e.g.][]{siren2021, ebler2022}, such approaches remain relatively
+\shortcite<e.g.>[]{siren2021, ebler2022}, such approaches remain relatively
 inaccurate compared to SNP genotyping workflows.
@@
 of variants in GWAS without genotyping SVs at population scale could be useful.
-\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
+\shortciteA{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
 represent such an alternative method. This approach relies on the presence or
@@
 few studies have applied this \emph{k}-mer-based GWAS approach
-\citep[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
-developed \citep{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
+\shortcite<e.g.>[]{tripodi2021, colque2021} and other similar approaches have been
+developed \shortcite{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
 such approaches have not been widely adopted yet.
@@
 based on SNPs and indels. In order to test these methods, we analyzed ten
-qualitative traits \citep{bandillo2017} and two quantitative traits
-\citep{bandillo2015} that have been previously studied using a SNP array
-developed for soybean \citep{song2013}. In addition, we analyzed a quantitative
+qualitative traits \shortcite{bandillo2017} and two quantitative traits
+\shortcite{bandillo2015} that have been previously studied using a SNP array
+developed for soybean \shortcite{song2013}. In addition, we analyzed a quantitative
 trait (horizontal resistance to \emph{Phytophthora sojae}) that has been
 previously studied using whole-genome sequencing (WGS) data
-\citep{deronne2022}.  Since the underlying genes and causal variants are
+\shortcite{deronne2022}.  Since the underlying genes and causal variants are
 already known for several of these traits, they provide a good test case for
@@
 Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
-sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
-Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
+sojae} \shortcite{deronne2022}.  We identified 741 SRA runs corresponding to
+Illumina paired-end sequencing data for those 389 accessions \shortcite<data from>[]{zhou2015, valliyodan2016, fang2017, bayer2021}
 and extracted paired-end reads using the fastq-dump command v. 2.9.6
@@
 were then mapped to the reference assembly version 4 of \emph{G. max}
-cultivar Williams82 \citep{valliyodan2019} concatenated with chloroplast
-and mitochondrion sequences obtained from SoyBase \citep{grant2010}. We used
-bwa-mem v. 0.7.17-r1188 \citep{li2009-bwa} with default parameters to map the
+cultivar Williams82 \shortcite{valliyodan2019} concatenated with chloroplast
+and mitochondrion sequences obtained from SoyBase \shortcite{grant2010}. We used
+bwa-mem v. 0.7.17-r1188 \shortcite{li2009-bwa} with default parameters to map the
 reads in paired-end mode. Reads that were left unpaired following adapter- and
@@
 merged with mapped paired-end reads using samtools \texttt{merge}
-v. 1.12 \citep{li2009-samtools}.  We then added read groups identifying
+v. 1.12 \shortcite{li2009-samtools}.  We then added read groups identifying
 individual SRA runs to the BAM files and merged the reads belonging to the same
@@
 		to call SVs in the population following a pipeline outlined
-		by \cite{lemay2022}. In brief, we combined information from
-		the SV calling programs AsmVar \citep{liu2015}, Manta \citep{chen2016},
-		smoove \citep{pedersen2019} and SvABA \citep{wala2018} to generate
+		by \shortciteA{lemay2022}. In brief, we combined information from
+		the SV calling programs AsmVar \shortcite{liu2015}, Manta \shortcite{chen2016},
+		smoove \shortcite{pedersen2019} and SvABA \shortcite{wala2018} to generate
 		a set of candidate SVs from the mapped Illumina data.
-	\item We used SVs discovered by \cite{lemay2022} from Oxford Nanopore
+	\item We used SVs discovered by \shortciteA{lemay2022} from Oxford Nanopore
 		sequencing data of 17 Canadian soybean cultivars.
 	\item We called SVs from 26 high-quality genome assemblies published by
-		\cite{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
+		\shortciteA{liu2020-pangenome}, as well as that of the cultivar Zhonghuang 13 (ZH13),
 		\textit{G. soja} accession W05, and \textit{G. max} cultivar Lee.
-		We used methods similar to those of \cite{liu2020-pangenome} to call
+		We used methods similar to those of \shortciteA{liu2020-pangenome} to call
 		SVs from the comparison of these assemblies to that of Williams82.
@@
 The SVs called using each of these approaches were merged using SVmerge
-\citep{wong2010} and the resulting set of candidate SVs was used for
-genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
+\shortcite{wong2010} and the resulting set of candidate SVs was used for
+genotyping. We used Paragraph \shortcite{chen2019} to genotype the candidate SVs
 from the mapped Illumina reads for all accessions. We filtered the genotyped
@@
 
-We used Platypus v. 0.8.1.1 \citep{rimmer2014} to call SNPs and indels from the
+We used Platypus v. 0.8.1.1 \shortcite{rimmer2014} to call SNPs and indels from the
 mapped reads of the 389 accessions. The 21.1 M SNPs and
@@
 Moreover, we pruned the dataset using the \texttt{-\-indep-pairwise} option of
-PLINK v. 1.90b5.3 \citep{purcell2007} with a window of 1,000 markers, a
+PLINK v. 1.90b5.3 \shortcite{purcell2007} with a window of 1,000 markers, a
 step of 100 markers and an $r^2$ LD threshold of 0.9. This pruning step was
@@
 
-We used the approach outlined by \cite{voichek2020} for \emph{k}-mer-based GWAS
+We used the approach outlined by \shortciteA{voichek2020} for \emph{k}-mer-based GWAS
 and their recommendations on the
@@
 in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
-\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
+\shortcite{kokot2017}. \emph{k}-mers were counted twice, the first time using their
 canonized form (the first of either the observed \textit{k}-mer or its reverse
@@
 
-We analyzed ten qualitative traits previously studied by \cite{bandillo2017}: flower
+We analyzed ten qualitative traits previously studied by \shortciteA{bandillo2017}: flower
 color, pubescence color, seed coat color, stem termination type, hilum color,
@@
 to numerical values for GWAS analysis following methods similar to those of
-\cite{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
+\shortciteA{bandillo2017}.  For some traits, we conducted more than one GWAS analysis
 using targeted subsets of the observed phenotypes in order to focus on specific
@@
 used the horizontal resistance to the oomycete \textit{P. sojae} (corrected dry
-weight, CDW) as recently assessed in a hydroponic assay \citep{deronne2022}.
+weight, CDW) as recently assessed in a hydroponic assay \shortcite{deronne2022}.
 Our analysis included 340 accessions whose phenotypic data
-was already published by \cite{deronne2022} as well as 49 additional accessions for
+was already published by \shortciteA{deronne2022} as well as 49 additional accessions for
 which sequencing data was not yet available when their study was conducted.
@@
 
-We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
+We used the GAPIT3 R package v. 3.1.0 \shortcite{wang2021} for conducting GWAS on
 the SNP/indel and SV datasets described above. SVs and indels required some additional
 preprocessing steps to make variant
-representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
+representation suitable for downstream analyses \shortcite{lemay2022gwas}. Briefly, we recoded all
 variants as SNPs prior to converting the VCF files to Hapmap diploid format
-using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
+using the TASSEL command-line tools \shortcite{bradbury2007}; variant IDs enabled
 retrieval of the proper metadata for each variant after GWAS.
@@
 approach.  We used a randomization approach to determine the 5\% family-wise
-error-rate threshold for each GWAS analysis as described by \cite{voichek2020}.
+error-rate threshold for each GWAS analysis as described by \shortciteA{voichek2020}.
 To compute this threshold, we permuted the phenotypic observations 100 times
@@
 We conducted GWAS based on \emph{k}-mers using the method developed by
-\cite{voichek2020}. In brief, the
+\shortciteA{voichek2020}. In brief, the
 analysis was conducted in two steps. The first step used an approximate model
 to identify 1 million potentially significant \emph{k}-mers. These were then used in
-the second step as input to an exact model implemented in GEMMA \citep{zhou2012}, resulting
+the second step as input to an exact model implemented in GEMMA \shortcite{zhou2012}, resulting
 in a \emph{p}-value for each \emph{k}-mer. We
@@
 (\url{https://github.com/malemay/katcher}) that
-uses the htslib library \citep{bonfield2021} for efficient reading and
+uses the htslib library \shortcite{bonfield2021} for efficient reading and
 writing of BAM files. katcher and its associated utilities allow for efficient
@@
 
-Because previous studies \citep{bandillo2015, bandillo2017} used assembly version 1 of
+Because previous studies \shortcite{bandillo2015, bandillo2017} used assembly version 1 of
 Williams82, we converted the genomic coordinates of their signals to assembly version 4
@@
 These read pairs were used for \emph{de novo} assembly with SPAdes v. 3.15.4
-\citep{spades} using the \texttt{--careful} parameter. The assemblies were done
+\shortcite{spades} using the \texttt{--careful} parameter. The assemblies were done
 individually for each accession and then aligned to the Williams82 reference
@@
 that overlapped loci of interest were realigned using multiple sequence
-alignment with the MAFFT program v. 7.475 \citep{katoh2002} for identifying and
+alignment with the MAFFT program v. 7.475 \shortcite{katoh2002} for identifying and
 visualizing the haplotypes at these loci. The identification of variants
@@
 Unless otherwise noted, all analyses were conducted using R v. 4.2.0
-\citep{r2022} and Bioconductor v. 3.15 \citep{huber2015}. We used Bioconductor
-packages Biostrings v. 2.64.0 \citep{pages2022}, GenomicRanges v. 1.48.0
-\citep{lawrence2013}, Rsamtools v. 2.12.0 \citep{rsamtools2022}, rtracklayer v.
-1.56.0 \citep{lawrence2009} and VariantAnnotation v. 1.42.1
-\citep{obenchain2014}.  We gathered several functions used for the downstream
+\shortcite{r2022} and Bioconductor v. 3.15 \shortcite{huber2015}. We used Bioconductor
+packages Biostrings v. 2.64.0 \shortcite{pages2022}, GenomicRanges v. 1.48.0
+\shortcite{lawrence2013}, Rsamtools v. 2.12.0 \shortcite{rsamtools2022}, rtracklayer v.
+1.56.0 \shortcite{lawrence2009} and VariantAnnotation v. 1.42.1
+\shortcite{obenchain2014}.  We gathered several functions used for the downstream
 analysis of GWAS results into a package called gwask which is available on
@@
 
-\cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
+\shortciteA{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
 Glyma.13g\-072100 as associated with the \textit{W1} locus for flower color. In
@@
 
-\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
+\shortciteA{zabala2003} identified Glyma.06g202300 as the gene associated with the
 \textit{T} locus for pubescence color and found a 1-bp deletion as a putative
@@
 \textit{k}-mer identified the causal variant documented at this locus
-\citep{yan2020}, it was the fourth most significantly associated variant for
+\shortcite{yan2020}, it was the fourth most significantly associated variant for
 the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d, Supplemental Figure 
@@
 that takes the form of a 10.91-kb inverted duplication of three chalcone
-synthase (CHS) genes \citep[CHS1, CHS3 and CHS4;][]{tuteja2008}. The presence of
+synthase (CHS) genes \shortcite<CHS1, CHS3 and CHS4;>[]{tuteja2008}. The presence of
 this inverted duplication results in the silencing of all CHS genes in the seed
 coat specifically and thus in the absence of seed coat pigmentation
-\citep{tuteja2009}. While we did find signals overlapping this causal variant
+\shortcite{tuteja2009}. While we did find signals overlapping this causal variant
 using all three GWAS approaches (Supplemental Figures S13 and
@@
 approaches (Supplemental Figure S15).
-\cite{wang2018} identified Glyma.01g198500 as the gene associated with
+\shortciteA{wang2018} identified Glyma.01g198500 as the gene associated with
 this locus and found an A>G SNP affecting transcript splicing as the causal
@@
 We did identify one of the SNPs suggested as putative causal variants for this
-locus by \cite{tian2010} and \cite{liu2010}, however this SNP was only the
+locus by \shortciteA{tian2010} and \shortciteA{liu2010}, however this SNP was only the
 135\textsuperscript{th} most associated marker found by Platypus and the
@@
 Our results are consistent with previous findings in suggesting that there is
-more than one causal variant at the \textit{Dt1} locus \citep{liu2010,
+more than one causal variant at the \textit{Dt1} locus \shortcite{liu2010,
 tian2010}. Indeed, a few cultivars bearing the allele associated with
@@
 S22 and S23).
-\cite{gillman2011} identified this R2R3 MYB transcription factor as the
+\shortciteA{gillman2011} identified this R2R3 MYB transcription factor as the
 molecular basis for the \textit{R} locus and documented four different
@@
 well as Glyma.09g235100 and two others are part of a tandem array of four R2R3
-MYB genes considered as candidates by \cite{gillman2011}, and of which
+MYB genes considered as candidates by \shortciteA{gillman2011}, and of which
 Glyma.09g235100 was the only one expressed in the seed coat. Given the
-convincing evidence provided by \cite{gillman2011}, it is most likely that
+convincing evidence provided by \shortciteA{gillman2011}, it is most likely that
 Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
@@
 loss of function at the \emph{R} locus.  Still, we observed two causal variants
-previously documented by \cite{gillman2011} in our dataset (Supplemental Figures
+previously documented by \shortciteA{gillman2011} in our dataset (Supplemental Figures
 S24 and S25).
@@
 variant at this recently cloned locus is a copy number variant (CNV) overlapping the
-Glyma.12g187200 gene, with higher copy number resulting in reduced pubescence \citep{liu2020ps}.
+Glyma.12g187200 gene, with higher copy number resulting in reduced pubescence \shortcite{liu2020ps}.
 We identified a region of $\sim$42 kb that varied in copy number across
@@
 \ref{pubescence-density-main-figure}c), and therefore possibly corresponded to the
-25.6-kb region identified as the causal CNV by \cite{liu2020ps}. We found that
+25.6-kb region identified as the causal CNV by \shortciteA{liu2020ps}. We found that
 reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
@@
 
-The difference in size between the CNV reported by \cite{liu2020ps} and our
+The difference in size between the CNV reported by \shortciteA{liu2020ps} and our
 results may be due to the presence of some unresolved sequence (a stretch of
@@
 the seed coat luster trait. Seed coat luster has been shown by
-\cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
+\shortciteA{gijzen1999} to be largely caused by the deposition of a hydrophobic
 protein (HPS) at the seed surface. Copy number variation of this gene at the
 \textit{B} locus was later shown to explain variation in luster, although
-sequence variation in the HPS sequence may also play a role \citep{gijzen2006}.
+sequence variation in the HPS sequence may also play a role \shortcite{gijzen2006}.
 Using a sequencing depth-based analysis, we identified a 31-kb region spanning
@@
 Both the \textit{L1} and the \textit{L2} loci for pod color have yet to be
-cloned. While \cite{he2015} proposed the gene Glyma.19g101700 as likely
+cloned. While \shortciteA{he2015} proposed the gene Glyma.19g101700 as likely
 corresponding to \textit{L1} based on fine mapping and expression analysis, our
-results as well as those reported by \cite{bandillo2017} do not support this
+results as well as those reported by \shortciteA{bandillo2017} do not support this
 gene as a candidate due to its location $\sim$3 Mb away from GWAS signals. Our
@@
 within its sequence and it is highly expressed in pods according to data hosted
-on SoyBase \citep{grant2010, severin2010}. Moreover, the role of some MATE
-proteins in transporting flavonoids has been demonstrated \citep{chen2015}.
+on SoyBase \shortcite{grant2010, severin2010}. Moreover, the role of some MATE
+proteins in transporting flavonoids has been demonstrated \shortcite{chen2015}.
 
@@
 The two known loci for pubescence form in soybean, \textit{Pa1} and
-\textit{Pa2}, have also not been cloned yet. Nevertheless, \cite{gilbert2017}
+\textit{Pa2}, have also not been cloned yet. Nevertheless, \shortciteA{gilbert2017}
 has suggested Glyma.12g213900 as a candidate for \textit{Pa1} based on data
-published by \cite{bandillo2017}. This gene encodes a MYB transcription factor
+published by \shortciteA{bandillo2017}. This gene encodes a MYB transcription factor
 annotated as playing a role in trichome branching. Indeed, the most significant
@@
 
-We used corrected dry weight (CDW) as described by \cite{deronne2022} as a
+We used corrected dry weight (CDW) as described by \shortciteA{deronne2022} as a
 measure of horizontal resistance to \emph{P. sojae}. We found a strong signal
 overlapping the genomic region reported by these authors on Gm15 using all three GWAS
-approaches (Supplemental Figure S41). \cite{deronne2022}
+approaches (Supplemental Figure S41). \shortciteA{deronne2022}
 proposed Glyma.15g217100 as a candidate gene for explaining phenotypic
@@
 Given the important role played by calcium signaling in defense
-responses \citep{zhang2014}, we suggest Glyma.15g217700 as an interesting
+responses \shortcite{zhang2014}, we suggest Glyma.15g217700 as an interesting
 candidate gene for explaining resistance to \emph{P. sojae} in this population.
@@
 result in a decreased ability to identify causal variants or genes. For
-example, \cite{bandillo2017} noted that observed phenotypes for stem
+example, \shortciteA{bandillo2017} noted that observed phenotypes for stem
 termination type in the GRIN database may not correspond to the expected
@@
 \emph{k}-mer counts instead of presence/absence have been developed
-\citep{rahman2018, he2021} which may solve issues like these and broaden the
+\shortcite{rahman2018, he2021} which may solve issues like these and broaden the
 scope of variants that can be identified from \emph{k}-mers.
@@
 limitation of this approach. Although the method developed by
-\cite{voichek2020} is in itself relatively simple and computationally
+\shortciteA{voichek2020} is in itself relatively simple and computationally
 inexpensive, downstream analysis and interpretation of the output is
@@
 meaningful sequence variation. Going forward, mapping the \emph{k}-mers or
-reads to a graph-based genome \citep[e.g. using the vg toolkit;][]{siren2021}
+reads to a graph-based genome \shortcite<e.g. using the vg toolkit;>[]{siren2021}
 may represent an improvement to using a single linear genome.  Grouping
@@
 loci control a trait. In this study, we used an assembly-based approach similar
-to what others have previously done \citep[e.g.][]{voichek2020, rahman2018}.
+to what others have previously done \shortcite<e.g.>[]{voichek2020, rahman2018}.
 This approach allowed us to link \emph{k}-mers to their underlying variation
@@
 to a reference genome to identify their reference-based coordinates
-\citep[e.g.][]{voichek2020, he2021, rahman2018, tripodi2021}.  However, while
+\shortcite<e.g.>[]{voichek2020, he2021, rahman2018, tripodi2021}.  However, while
 this approach may work for \emph{k}-mers that show little divergence from the
@@
 variants, most likely due to the difficulty of calling and genotyping such
-variants \citep{kosugi2019, lemay2022}. In the case of the \textit{Ps} locus,
+variants \shortcite{kosugi2019, lemay2022}. In the case of the \textit{Ps} locus,
 this limitation was due to a filtering step in our pipeline rather than to an
@@
 to SNP-based GWAS will largely depend on whether SVs can represent variation
-that is not otherwise in LD with SNPs. In tomato, \cite{dominguez2020} found
+that is not otherwise in LD with SNPs. In tomato, \shortciteA{dominguez2020} found
 that most transposable element insertions associated with phenotypic variation
 in agronomic traits were not tagged by SNPs. On the other hand, in humans,
-\cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
+\shortciteA{maretty2017} found that the majority of SVs were in high LD with SNPs. It
 is therefore likely that how well SVs can be tagged by SNPs will depend on the
@@
 GWAS can be conducted specifically for \textit{k}-mers that are associated with
-SVs \citep{jayakodi2020}.  Previous studies have obtained noteworthy results by
-using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
+SVs \shortcite{jayakodi2020}.  Previous studies have obtained noteworthy results by
+using SVs for GWAS analyses \shortcite<e.g.>[]{zhang2015, akakpo2020,
 dominguez2020}, however it would be worthwhile to see whether these results
@@
 Some of the datasets generated as part of this work are available on figshare
-\citep{kmer-gwas-figshare}.
+\shortcite{kmer-gwas-figshare}.
 
@@
 
-The high-quality assemblies generated by \cite{liu2020-pangenome} are available
+The high-quality assemblies generated by \shortciteA{liu2020-pangenome} are available
 on the Genome Warehouse through Accession Number \citeauthor{PRJCA002030}.
@@
 
-The \textit{I} locus contig assembled by \cite{tuteja2008} is available from
+The \textit{I} locus contig assembled by \shortciteA{tuteja2008} is available from
 NCBI (\url{https://www.ncbi.nlm.nih.gov/nuccore/EF623854}).
@@
 
-The SVs identified from Oxford Nanopore data by \cite{lemay2022} are available
-on figshare \citep{soybean-sv-figshare}.
+The SVs identified from Oxford Nanopore data by \shortciteA{lemay2022} are available
+on figshare \shortcite{soybean-sv-figshare}.
 
@@
 The code associated with the results presented in this manuscript is archived
-on figshare \citep{kmer-gwas-code}.
+on figshare \shortcite{kmer-gwas-code}.
 
@@
 		\toprule Trait & Locus & Gene & Causal variant & SNPs\tnote{a} & SVs\tnote{b} & \textit{k}-mers \\ \midrule
-		Flower color & \textit{W1} & Glyma.13g072100 & 65-bp deletion \citep{zabala2007} & ** & *** & *** \\
-		Pubescence color & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & ** & * & ** \\
-		Pubescence color & \textit{Td} & Glyma.03g258700 & SNP \citep{yan2020} & * & - & *** \\
-		Seed coat color & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & ** & ** & ** \\
-		Seed coat color & \textit{G} & Glyma.01g198500 & SNP \citep{wang2018} & ** & ** & *** \\
-		Seed coat color\tnote{c} & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & - & - & - \\
-		Seed coat color\tnote{c} & \textit{R} & Glyma.09g235100 & SNPs and indels \citep{gillman2011} & - & - & - \\
-		Stem termination type & \textit{Dt1} & Glyma.19g194300 & SNPs \citep{liu2010} & ** & * & ** \\
-		Stem termination type\tnote{c} & \textit{Dt2} & Glyma.18g273600 & SNPs \citep{ping2014} & - & - & - \\
-		Stem termination type\tnote{c} & \textit{E3} & Glyma.19g224200 & Various types \citep{tardivel2019} & - & - & - \\
-		Hilum color\tnote{c} & \textit{T} & Glyma.06g202300 & 1-bp deletion \citep{zabala2003} & ** & ** & ** \\
-		Hilum color\tnote{c} & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & * & * & ** \\
-		Hilum color & \textit{R} & Glyma.09g235100 & SNPs and indels \citep{gillman2011} & * & - & ** \\
-		Hilum color\tnote{c} & \textit{W1} & Glyma.13g072100 & 65-bp deletion \citep{zabala2007} & - & - & - \\
-		Pubescence density & \textit{Ps} & Glyma.12g187200 & 25.6-kb CNV \citep{liu2020ps} & ** & ** & *** \\
-		Pubescence density\tnote{c} & \textit{Pd1} & Glyma.01g240100 & SNP \citep{liu2020ps} & - & - & - \\
-		Pubescence density\tnote{c} & \textit{P1} & Glyma.09g278000 & SNP \citep{liu2020ps} & - & - & - \\
-		Seed coat luster & \textit{B} & Duplicated HPS genes & CNV (several kb) \citep{gijzen2006} & ** & ** & ** \\
-		Seed coat luster\tnote{c} & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \citep{tuteja2008} & - & - & - \\
-		Maturity group\tnote{c} & \textit{E1} & Glyma.06g207800 & Various types \citep{tardivel2019} & - & - & - \\
-		Maturity group\tnote{c} & \textit{E2} & Glyma.10g221500 & Various types \citep{tardivel2019} & - & - & - \\
-		Maturity group\tnote{c} & \textit{E3} & Glyma.19g224200 & Various types \citep{tardivel2019} & - & - & - \\
-		Maturity group\tnote{c} & \textit{E4} & Glyma.20g090000 & Various types \citep{tardivel2019} & - & - & - \\
-		Oil\tnote{c} & \textit{oilGm20} & Glyma.20g085100 & 321-bp deletion \citep{fliege2022} & - & - & - \\
-		Oil\tnote{c} & \textit{oilGm15} & Glyma.15g049200 & SNPs and indels \citep{zhang2020} & - & - & - \\
-		Protein\tnote{c} & \textit{proGm20} & Glyma.20g085100 & 321-bp deletion \citep{fliege2022} & - & - & - \\
-		Protein\tnote{c} & \textit{proGm15} & Glyma.15g049200 & SNPs and indels \citep{zhang2020} & - & - & - \\
+		Flower color & \textit{W1} & Glyma.13g072100 & 65-bp deletion \shortcite{zabala2007} & ** & *** & *** \\
+		Pubescence color & \textit{T} & Glyma.06g202300 & 1-bp deletion \shortcite{zabala2003} & ** & * & ** \\
+		Pubescence color & \textit{Td} & Glyma.03g258700 & SNP \shortcite{yan2020} & * & - & *** \\
+		Seed coat color & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \shortcite{tuteja2008} & ** & ** & ** \\
+		Seed coat color & \textit{G} & Glyma.01g198500 & SNP \shortcite{wang2018} & ** & ** & *** \\
+		Seed coat color\tnote{c} & \textit{T} & Glyma.06g202300 & 1-bp deletion \shortcite{zabala2003} & - & - & - \\
+		Seed coat color\tnote{c} & \textit{R} & Glyma.09g235100 & SNPs and indels \shortcite{gillman2011} & - & - & - \\
+		Stem termination type & \textit{Dt1} & Glyma.19g194300 & SNPs \shortcite{liu2010} & ** & * & ** \\
+		Stem termination type\tnote{c} & \textit{Dt2} & Glyma.18g273600 & SNPs \shortcite{ping2014} & - & - & - \\
+		Stem termination type\tnote{c} & \textit{E3} & Glyma.19g224200 & Various types \shortcite{tardivel2019} & - & - & - \\
+		Hilum color\tnote{c} & \textit{T} & Glyma.06g202300 & 1-bp deletion \shortcite{zabala2003} & ** & ** & ** \\
+		Hilum color\tnote{c} & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \shortcite{tuteja2008} & * & * & ** \\
+		Hilum color & \textit{R} & Glyma.09g235100 & SNPs and indels \shortcite{gillman2011} & * & - & ** \\
+		Hilum color\tnote{c} & \textit{W1} & Glyma.13g072100 & 65-bp deletion \shortcite{zabala2007} & - & - & - \\
+		Pubescence density & \textit{Ps} & Glyma.12g187200 & 25.6-kb CNV \shortcite{liu2020ps} & ** & ** & *** \\
+		Pubescence density\tnote{c} & \textit{Pd1} & Glyma.01g240100 & SNP \shortcite{liu2020ps} & - & - & - \\
+		Pubescence density\tnote{c} & \textit{P1} & Glyma.09g278000 & SNP \shortcite{liu2020ps} & - & - & - \\
+		Seed coat luster & \textit{B} & Duplicated HPS genes & CNV (several kb) \shortcite{gijzen2006} & ** & ** & ** \\
+		Seed coat luster\tnote{c} & \textit{I} & CHS gene cluster & 10.9-kb inversion/duplication \shortcite{tuteja2008} & - & - & - \\
+		Maturity group\tnote{c} & \textit{E1} & Glyma.06g207800 & Various types \shortcite{tardivel2019} & - & - & - \\
+		Maturity group\tnote{c} & \textit{E2} & Glyma.10g221500 & Various types \shortcite{tardivel2019} & - & - & - \\
+		Maturity group\tnote{c} & \textit{E3} & Glyma.19g224200 & Various types \shortcite{tardivel2019} & - & - & - \\
+		Maturity group\tnote{c} & \textit{E4} & Glyma.20g090000 & Various types \shortcite{tardivel2019} & - & - & - \\
+		Oil\tnote{c} & \textit{oilGm20} & Glyma.20g085100 & 321-bp deletion \shortcite{fliege2022} & - & - & - \\
+		Oil\tnote{c} & \textit{oilGm15} & Glyma.15g049200 & SNPs and indels \shortcite{zhang2020} & - & - & - \\
+		Protein\tnote{c} & \textit{proGm20} & Glyma.20g085100 & 321-bp deletion \shortcite{fliege2022} & - & - & - \\
+		Protein\tnote{c} & \textit{proGm15} & Glyma.15g049200 & SNPs and indels \shortcite{zhang2020} & - & - & - \\
 		\bottomrule

### 2023-07-04 additional_files/additional_file_1.tex
Format references for submission to The Plant Genome
Add qq-plots to additional file 1

@@
 	\begin{cfigure}
-		\includegraphics[width = 5.8in]{#1_manhattan}
+		\includegraphics[width = 6.5in]{#1_manhattan}
 		\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using three genotype datasets :

### 2023-07-04 figures/manhattan_plot.R

@@ -1,6 +1,7 @@
 # Loading the required packages
 suppressMessages(library(grid))
 suppressMessages(library(gwask))
+suppressMessages(library(GenomicRanges))
 
 # Reading the trait being analyzed from the command line
 trait <- commandArgs(trailingOnly = TRUE)[1]
@@ -9,7 +10,7 @@ programs <- c("platypus", "paragraph", "kmers")
 program_labels <- c("SNPs/indels", "SVs", "kmers")
 
 # Outputting to a PNG file
-png(paste0("figures/", trait, "_manhattan.png"), width = 9, height = 10, units = "in", res = 100)
+png(paste0("figures/", trait, "_manhattan.png"), width = 10, height = 10, units = "in", res = 100)
 
 # Resetting the viewport
 grid.newpage()
@@ -19,11 +20,12 @@ grid.newpage()
 # DEPENDENCY: kmers manhattan subplot
 
 # Setting a viewport with 3 rows, one for each of the programs
-pushViewport(viewport(height = 0.95, y = 0, just = "bottom", layout = grid.layout(nrow = 3)))
+pushViewport(viewport(height = 0.95, y = 0, just = "bottom",
+		      layout = grid.layout(nrow = 3, ncol = 2, widths = unit(c(72, 28), "null"))))
 
 # Reading and printing each of the subplots in turn in their viewport row
 for(i in 1:3){
-	pushViewport(viewport(layout.pos.row = i))
+	pushViewport(viewport(layout.pos.row = i, layout.pos.col = 1))
 	grid.draw(readRDS(paste0("figures/grobs/", programs[i], "_", trait, "_manhattan.rds")))
 	if(programs[i] == "kmers") {
 		grid.text(expression(paste("(c) ", italic(k)-mers)), 0.1, 1.08, hjust = 0, gp = gpar(fontsize = 14))
@@ -31,6 +33,43 @@ for(i in 1:3){
 		grid.text(paste0("(", letters[i], ") ", program_labels[i]), 0.1, 1.08, hjust = 0, gp = gpar(fontsize = 14))
 	}
 	upViewport()
+}
+
+# Taking care of the qqplots
+for(i in 1:3) {
+	# Generating the data to plot
+	if(programs[i] == "kmers") {
+		pvalues <- scan(paste0("gwas_results/kmer_data/", trait, "/all_pvalues_sorted.txt"),
+				what = numeric(1),
+				sep = "\n")
+	} else {
+		# Formatting the p-values
[…]

### 2023-07-04 additional_files/additional_file_1.tex
Address most of the reviewers' comments

@@
 		\includegraphics[width = 6.5in]{#1_manhattan}
-		\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using three genotype datasets :
+		\caption[Manhattan plots for #2]{Manhattan plots and associated Q-Q plots
+		generated from the GWAS analysis of #2 on #3 samples using three genotype datasets :
 		(a) Platypus (SNPs and indels), (b) Paragraph (SVs), (c) \emph{k}-mers presence/absence.
@@
 		The ``Gm'' prefix has been left out of chromosome names for simplicity.
+		The dashed red line in Q-Q plots represents the identity line ($y = x$)
 		}

### 2023-07-04 additional_files/manuscript.tex
Address most of the reviewers' comments

@@
 single-nucleotide polymorphisms (SNPs) could alleviate this problem.  In this
-study, we tested the potential of SV- and \emph{k}-mer-based GWAS in soybean by
+study, we tested the potential of structural variant (SV)- and \emph{k}-mer-based GWAS in soybean by
 applying these methods as well as conventional SNP/indel-based GWAS to 13
@@
 We selected 389 inbred \emph{G. max} accessions based on the availability of
-Illumina WGS data and phenotypic data for resistance to the oomycete \emph{P.
-sojae} \citep{deronne2022}.  We identified 741 SRA runs corresponding to
-Illumina paired-end sequencing data for those 389 accessions \citep[data from][]{zhou2015, valliyodan2016, fang2017, bayer2021}
+phenotypic data for resistance to the oomycete \emph{P.
+sojae} \shortcite{deronne2022} and matching Illumina WGS data. We identified 741 SRA runs corresponding to
+Illumina paired-end sequencing data for those 389 accessions \shortcite<data from>[]{zhou2015, valliyodan2016, fang2017, bayer2021}
 and extracted paired-end reads using the fastq-dump command v. 2.9.6
@@
 also removed. The resulting dataset was used for GWAS and comprised 773,060
-SNPs and 151,570 indels.
+SNPs and 151,570 indels. Pruned markers were still used for GWAS at targeted
+loci when filtered markers identified significant associations (more details
+in the section describing the analysis of significant signals).
 
@@
 in the trimmed (FASTQ) reads for each accession using KMC3 v. 3.2.1
-\citep{kokot2017}. \emph{k}-mers were counted twice, the first time using their
+\shortcite{kokot2017}. We selected $k = 31$ based on demonstrated performance in previously
+published studies \shortcite{rahman2018, voichek2020, he2021}.
+\emph{k}-mers were counted twice, the first time using their
 canonized form (the first of either the observed \textit{k}-mer or its reverse
@@
 table for use in GWAS and was also used directly as input genotypes for the
-GWAS analyses. We used \emph{k}-mers with a MAF $\geq$ 0.02 for both the
-computation of kinship and GWAS.
+GWAS analyses. We used all \emph{k}-mers with a MAF $\geq$ 0.02 as input genotypes
+for both the computation of kinship and GWAS analysis. The total number of \textit{k}-mers
+in this filtered table was 637,605,529.
 
@@
 
-We used the GAPIT3 R package v. 3.1.0 \citep{wang2021} for conducting GWAS on
+We used the GAPIT3 R package v. 3.1.0 \shortcite{wang2021} for conducting GWAS on
 the SNP/indel and SV datasets described above. SVs and indels required some additional
 preprocessing steps to make variant
-representation suitable for downstream analyses \citep{lemay2022gwas}. Briefly, we recoded all
+representation suitable for downstream analyses \shortcite{lemay2022gwas}. Briefly, we recoded all
 variants as SNPs prior to converting the VCF files to Hapmap diploid format
-using the TASSEL command-line tools \citep{bradbury2007}; variant IDs enabled
+using the TASSEL command-line tools \shortcite{bradbury2007}; variant IDs enabled
 retrieval of the proper metadata for each variant after GWAS.
 
-We ran GAPIT using a mixed linear model (MLM) with 9 principal components and the VanRaden
-algorithm for computing the kinship matrix. We used an MLM model for
+We ran GAPIT using a mixed linear model (MLM) with nine principal components and the VanRaden
+algorithm for computing the kinship matrix. We chose to use nine principal components
+based on previous work by \shortciteA{deronne2022} using the same population.
+Principal components and kinship were computed
+on the fly by the GAPIT function itself from the provided inputs, i.e. based on SNPs/indels
+or on SVs depending on the analysis. We used an MLM model for
 consistency with the statistical model used by the \textit{k}-mer-based
 approach.  We used a randomization approach to determine the 5\% family-wise
-error-rate threshold for each GWAS analysis as described by \cite{voichek2020}.
+error-rate threshold for each GWAS analysis as described by \shortciteA{voichek2020}.
 To compute this threshold, we permuted the phenotypic observations 100 times
@@
 least one accession. The latter was done to remove spurious \emph{k}-mers that may originate
-from sequencing errors. All downstream analyses (Manhattan plots,
+from sequencing errors. Filtering based on mapping quality also removed all reads that
+mapped to multiple locations because bwa assigns a mapping quality of 0 to such reads.
+All downstream analyses (Manhattan plots,
 identification of signals, analysis of genes located near significant
@@
 \emph{k}-mer counts instead of presence/absence have been developed
-\citep{rahman2018, he2021} which may solve issues like these and broaden the
+\shortcite{rahman2018, he2021} which may solve issues like these and broaden the
 scope of variants that can be identified from \emph{k}-mers.
+Nevertheless, \shortciteA{voichek2020} demonstrated that 98.9\% of known translocations and
+inversions as well as 99.6\% of known insertions/deletions in \textit{Arabidopsis thaliana}
+could be tagged by unique 31-mers. It is therefore likely that the vast majority
+of variants can be detected through the presence/absence of \textit{k}-mers.
 
@@
 to SNP-based GWAS will largely depend on whether SVs can represent variation
-that is not otherwise in LD with SNPs. In tomato, \cite{dominguez2020} found
+that is not otherwise in LD with SNPs. In tomato, \shortciteA{dominguez2020} found
 that most transposable element insertions associated with phenotypic variation
 in agronomic traits were not tagged by SNPs. On the other hand, in humans,
-\cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
+\shortciteA{maretty2017} found that the majority of SVs were in high LD with SNPs. It
 is therefore likely that how well SVs can be tagged by SNPs will depend on the
-particular type of SV and on the species. A more efficient approach than
+particular type of SV and on the species. Until more robust SV calling and genotyping
+methods are developed or advances in sequencing technology result in more
+trustworthy SV datasets, a more efficient approach than
 SV-based GWAS may be to conduct \emph{k}-mer-based GWAS and only then link

### 2023-07-05 additional_files/manuscript.tex
Add a figure about sequencing depth and some material in the discussion.

@@
 and non-canonized form in at least 20\% of the accessions in which they are found.
+Only \textit{k}-mers that occurred twice in a given sample were considered to
+reduce the impact of sequencing errors.
 We then generated a table indicating the presence or absence of each
@@
 well compared to other GWAS methods, this approach does present some unique
-challenges and limitations. One obvious limitation is that using the
-presence/absence of \emph{k}-mers may not be appropriate in species where
-heterozygosity is common (i.e., most species other than inbred crops) as the
-\emph{k}-mers associated with the presence of both alleles will be observed at
-heterozygous loci. Moreover, although the presence/absence of \emph{k}-mers can
-detect variation independently of its type, some types of variation may be more
-difficult to detect than others.  For example, copy number variation at a locus
-may not be detected from presence/absence alone if all copies are identical and
-occur in tandem, such that no \textit{k}-mer can differentiate between two
-copies and more. For similar reasons, \emph{k}-mers may also fail to capture
-variation occurring in repeated sequences. Alternative models based on
-\emph{k}-mer counts instead of presence/absence have been developed
-\shortcite{rahman2018, he2021} which may solve issues like these and broaden the
-scope of variants that can be identified from \emph{k}-mers.
-Nevertheless, \shortciteA{voichek2020} demonstrated that 98.9\% of known translocations and
-inversions as well as 99.6\% of known insertions/deletions in \textit{Arabidopsis thaliana}
-could be tagged by unique 31-mers. It is therefore likely that the vast majority
-of variants can be detected through the presence/absence of \textit{k}-mers.
+challenges and limitations. One of these challenges is the
+interpretation of \textit{k}-mer-based GWAS given the greater possibility
+of false positive associations than in conventional GWAS approaches. In this
+study, we noticed apparently spurious associations with \textit{k}-mers
+scattered throughout the genome were common. This was especially the case for
+highly significant loci whose signal spanned several megabases such as \textit{W}
+and \textit{T}. A look at the Q-Q plots derived from \textit{k}-mer-based GWAS
+showed that these usually had the same shape as the Q-Q plots derived from SNP-
+and SV-based GWAS, suggesting that population structure was properly controlled.
+We believe that a more likely culprit for these spurious associations is the
+difficulty in linking \textit{k}-mers to genomic coordinates. At loci that
+span a large interval, it is indeed be highly likely that some of the \textit{k}-mers
+coming from the region map elsewhere in the genome. This interpretation is supported
+by the fact that the LD between presumably spurious \textit{k}-mers and
+\textit{k}-mers from target loci was generally high (e.g. Figures S5 and S47).
+Overall, the risk of false association due to population structure and the identification
+of an appropriate significance threshold seems well controlled, but work remains
+to be done in reducing the occurrence of spurious associations.
+
+There are obvious limitations to the relying on the presence/absence of
+\emph{k}-mers for GWAS. For example, this approach may not be appropriate in
+species where heterozygosity is common (i.e., most species other than inbred
+crops) as the \emph{k}-mers associated with the presence of both alleles will
+be observed at heterozygous loci. Moreover, although the presence/absence of
+\emph{k}-mers can detect variation independently of its type, some types of
+variation may be more difficult to detect than others.  For example, copy
+number variation at a locus may not be detected from presence/absence alone if
+all copies are identical and occur in tandem, such that no \textit{k}-mer can
+differentiate between two copies and more. For similar reasons, \emph{k}-mers
+may also fail to capture variation occurring in repeated sequences. Alternative
+models based on \emph{k}-mer counts instead of presence/absence have been
+developed \shortcite{rahman2018, he2021} which may solve issues like these and
+broaden the scope of variants that can be identified from \emph{k}-mers.
+Nevertheless, \shortciteA{voichek2020} demonstrated that 98.9\% of known
+translocations and inversions as well as 99.6\% of known insertions/deletions
+in \textit{Arabidopsis thaliana} could be tagged by unique 31-mers. It is
+therefore likely that the vast majority of variants can be detected through the
+presence/absence of \textit{k}-mers.
+
+In addition to the link between \textit{k}-mer presence/absence and sequence
+variation, one may wonder to what extent errors may occur in declaring the
+presence or absence of a \textit{k}-mer based on Illumina WGS data. If we
+neglect the possibility of contamination, the most likely error would be to
+declare a \textit{k}-mer as absent simply because sequencing depth was
+insufficient. In this study, we declared \textit{k}-mers as present
+based on a minimum of two occurrences. Therefore, a \textit{k}-mer may be
+wrongly declared as absent if it is observed only once or not at all.  To
+quantify this error, we compared the number of times that a given position was
+observed only once in a sample to the average sequencing depth of that sample.
+We found that most samples with a sequencing depth between 10 and 20X had
+a proportion of only 0.5 to 1.5\% of positions covered by only one read.
+Therefore, the risk of falsely declaring a \textit{k}-mer as absent from
+our dataset was low, and in any case much lower than typical error rates
+involved in calling and genoyping SVs \shortcite<e.g.>[]{kosugi2019, ebler2022}.
+The robustness of \textit{k}-mer presence/absence to sequencing depth is also
+demonstrated by \shortciteA{gaurav2022}.
 

### 2023-07-05 figures/seqdepth_figure.R

@@ -1,0 +1,49 @@
+#!/home/malem420/.local/bin/Rscript
+
+# This figure shows the proportion of the genome covered by
+# only one read as a function of mean coverage for a given sample
+
+# Loading the required libraries
+library(parallel)
+
+# Reading the mean coverage per sample
+# DEPENDENCY: phenotypic_data/phenotypic_data.csv
+samples <- read.table("phenotypic_data/phenotypic_data.csv", sep = ";", header = TRUE)$bayer_id
+
+# DEPENDENCY: sv_genotyping/paragraph/MANIFEST_FILES
+average_cov <- sapply(samples, function(x) {
+			      read.table(paste0("sv_genotyping/paragraph/manifest_files/", x, "_manifest.txt"),
+					 skip = 1)[[3]]})
+
+# Reading the coverage
+# DEPENDENCY: illumina_data/SAMTOOLS_STATS
+stats_files <- paste0("illumina_data/merged_bams/", samples, "_stats.txt")
+stopifnot(all(file.exists(stats_files)))
+
+read_coverage <- function(stats) {
+	output <- readLines(stats)
+	output <- grep("^COV", output, value = TRUE)
+	output <- do.call("rbind", strsplit(output, "\t"))
+	output <- as.data.frame(output)[, -1]
+	names(output) <- c("range", "mean", "count")
+	output$mean <- as.numeric(output$mean)
+	output$count <- as.numeric(output$count)
+	output
+}
+
+one_proportion <- function(stats) {
+	coverage <- read_coverage(stats)
+	stopifnot(sum(coverage$mean == 1) == 1)
+	coverage[coverage$mean == 1, "count"] / sum(coverage$count)
+}
+
+# The following vector contains the proportion of positions in the genome that are covered by only one read
+kmer_one <- unlist(mclapply(stats_files, one_proportion, mc.cores = 10))
+
+# Saving to file
+png("figures/seqdepth.png", width = 6, height = 6, units = "in", res = 300)
+plot(average_cov, kmer_one, main = "",
+     xlab = "Mean sequencing depth of sample",
+     ylab = "Fraction of genome covered by one read")
+dev.off()
+

### 2023-07-20 additional_files/additional_file_1.tex
Add a figure about sequencing depth and some material in the discussion.
Version submitted as revision

@@
 
+% A table with the number of k-mers foundd by each analysis
+\begin{table}
+	\centering
+	\caption[Number of k-mers per GWAS analysis]{Number of significant \textit{k}-mers identified per GWAS analysis}
+	\label{kmer_count_table}
+
+	\begin{threeparttable}\footnotesize
+
+			\csvreader[head to column names,
+			tabular = lrrr@{\hskip 3pt}l,
+			table head = \toprule GWAS analysis\tnote{a} & Raw\tnote{b} & Filtered\tnote{c} & At known loci & (\%)\tnote{d} \\ \midrule,
+			table foot=\bottomrule,
+			before reading={\catcode`\#=12},
+			after reading={\catcode`\#=6},
+			/csv/separator=semicolon]%
+			{../tables/kmer_count_table.csv}{}{\gwas & \raw & \filtered & \nsignals & (\percent)}
+
+				\begin{tablenotes}
+				\item[a] GWAS numbers refer to numbers in tables \ref{FLOWER.COLOR_gwas_table}
+					to \ref{MATURITY.GROUP_gwas_table}.
+				\item[b] Number of significant \textit{k}-mers in raw GWAS output.
+				\item[c] Number of significant \textit{k}-mers after filtering for
+					mapping quality, minimum count and redundancy.
+				\item[d] Number of \textit{k}-mers overlapping known loci and
+					percentage relative to number of filtered \textit{k}-mers.
+				\end{tablenotes}
+
+	\end{threeparttable}
+\end{table}
+
+\clearpage
+
 % Supplemental figures
@@
 
+% Figure for the proportion of genome covered by only one read as a function of sequencing depth
+\begin{cfigure}
+	\includegraphics[width = 6.5in]{seqdepth}
+	\caption[Proportion of the genome covered by one read versus sequencing depth]{%
+		Proportion of the genome covered by only one read as a function
+	of the sequencing depth of 363 Illumina whole-genome sequencing soybean
+	samples used for GWAS in this study.}
+	\label{seqdepth-figure}
+\end{cfigure}
+
+\clearpage
+
 % Figures mentioned only in supplemental data

### 2023-07-20 additional_files/manuscript.tex
Version submitted as revision

@@
 SNPs and 151,570 indels. Pruned markers were still used for GWAS at targeted
-loci when filtered markers identified significant associations (more details
+loci where filtered markers identified significant associations (more details
 in the section describing the analysis of significant signals).
@@
 and non-canonized form in at least 20\% of the accessions in which they are found.
-Only \textit{k}-mers that occurred twice in a given sample were considered to
-reduce the impact of sequencing errors.
+To reduce the potential impact of sequencing errors, only \textit{k}-mers that
+occurred at least twice in a given sample were considered.
 We then generated a table indicating the presence or absence of each
@@
 \textit{p}-values obtained by each approach at each of the loci considered in
-this study.
+this study. Supplemental Table S12 also shows some results specific to \textit{k}-mer-based
+analyses.
 
@@
 they may represent \textit{bona fide} loci associated with stem termination
-type (Figure \ref{stem-termination-main-figure}b). Similarly, a \textit{k}-mer
+type (Figure \ref{stem-termination-main-figure}b). We note that the signal newly
+identified on Gm18 is located 5 Mb away from the already documented \textit{Dt2}
+locus and may therefore correspond to that locus.
+Similarly, a \textit{k}-mer
 GWAS analysis comparing accessions with black and brown pods found a previously
@@
 of false positive associations than in conventional GWAS approaches. In this
-study, we noticed apparently spurious associations with \textit{k}-mers
-scattered throughout the genome were common. This was especially the case for
-highly significant loci whose signal spanned several megabases such as \textit{W}
-and \textit{T}. A look at the Q-Q plots derived from \textit{k}-mer-based GWAS
-showed that these usually had the same shape as the Q-Q plots derived from SNP-
+study, we noticed that spurious associations, appearing as \textit{k}-mers
+scattered throughout the genome, were common. This was especially the case for
+traits with highly significant signals that spanned several megabases, such as flower color
+and pubescence color (Supplemental Table S12). Q-Q plots derived from \textit{k}-mer-based GWAS
+generally displayed the same shape as the Q-Q plots derived from SNP-
 and SV-based GWAS, suggesting that population structure was properly controlled.
+Despite the very large number of tests conducted, the approach developed by
+\shortciteA{voichek2020} for setting significance thresholds and controlling
+for population structure therefore appears appropriate.
 We believe that a more likely culprit for these spurious associations is the
 difficulty in linking \textit{k}-mers to genomic coordinates. At loci that
-span a large interval, it is indeed be highly likely that some of the \textit{k}-mers
-coming from the region map elsewhere in the genome. This interpretation is supported
+span a large interval, it is indeed highly likely that some of the \textit{k}-mers
+coming from the target region will map elsewhere in the genome. This interpretation is supported
 by the fact that the LD between presumably spurious \textit{k}-mers and
-\textit{k}-mers from target loci was generally high (e.g. Figures S5 and S47).
-Overall, the risk of false association due to population structure and the identification
-of an appropriate significance threshold seems well controlled, but work remains
-to be done in reducing the occurrence of spurious associations.
-
-There are obvious limitations to the relying on the presence/absence of
+\textit{k}-mers from target loci was generally high (e.g. Supplemental Figures S5 and S47).
+Until further work resolves these issues, we recommend users of
+\textit{k}-mer-based GWAS to assess their results critically and to be aware
+that spurious associations are likely to occur.
+
+There are obvious limitations to relying on the presence/absence of
 \emph{k}-mers for GWAS. For example, this approach may not be appropriate in
@@
 
-In addition to the link between \textit{k}-mer presence/absence and sequence
-variation, one may wonder to what extent errors may occur in declaring the
+A related problem is the extent to which errors may occur in declaring the
 presence or absence of a \textit{k}-mer based on Illumina WGS data. If we
@@
 wrongly declared as absent if it is observed only once or not at all.  To
-quantify this error, we compared the number of times that a given position was
+quantify this error, we compared the number of positions that were
 observed only once in a sample to the average sequencing depth of that sample.
 We found that most samples with a sequencing depth between 10 and 20X had
-a proportion of only 0.5 to 1.5\% of positions covered by only one read.
+between 0.5 to 1.5\% of positions covered by a single read (Supplemental Figure S46).
 Therefore, the risk of falsely declaring a \textit{k}-mer as absent from
@@
 involved in calling and genoyping SVs \shortcite<e.g.>[]{kosugi2019, ebler2022}.
-The robustness of \textit{k}-mer presence/absence to sequencing depth is also
-demonstrated by \shortciteA{gaurav2022}.
+The robustness to low sequencing depth of GWAS analyses based on \textit{k}-mer
+presence/absence was also demonstrated by \shortciteA{gaurav2022}.
 

### 2023-07-20 figures/seqdepth_figure.R

@@ -44,6 +44,6 @@ kmer_one <- unlist(mclapply(stats_files, one_proportion, mc.cores = 10))
 png("figures/seqdepth.png", width = 6, height = 6, units = "in", res = 300)
 plot(average_cov, kmer_one, main = "",
      xlab = "Mean sequencing depth of sample",
-     ylab = "Fraction of genome covered by one read")
+     ylab = "Fraction of genome covered by a single read")
 dev.off()
 

### 2023-07-20 additional_files/manuscript.tex
Final files sent for publication

@@
 
-\section*{Introduction}
+\section{Introduction}
 Genome-wide association studies (GWAS) are analytical approaches that detect
@@
 
-\section*{Materials and methods}
-
-\subsection*{Sample selection and processing of sequencing data}
+\section{Materials and methods}
+
+\subsection{Sample selection and processing of sequencing data}
 
@@
 
-\subsection*{Discovery and genotyping of SVs}
+\subsection{Discovery and genotyping of SVs}
 
@@
 
-\subsection*{Discovery and genotyping of SNPs and indels}
+\subsection{Discovery and genotyping of SNPs and indels}
 
@@
 
-\subsection*{Computing the presence/absence table of \emph{k}-mers}
+\subsection{Computing the presence/absence table of \emph{k}-mers}
 
@@
 
-\subsection*{Phenotypic data}
+\subsection{Phenotypic data}
 
@@
 
-\subsection*{GWAS analyses}
+\subsection{GWAS analyses}
 
@@
 
-\subsection*{Analysis of significant \emph{k}-mers}
+\subsection{Analysis of significant \emph{k}-mers}
 
@@
 
-\subsection*{Downstream analyses of significant association signals}
+\subsection{Downstream analyses of significant association signals}
 
@@
 
-\subsection*{Linking significant \emph{k}-mers to variants}
+\subsection{Linking significant \emph{k}-mers to variants}
 
@@
 
-\subsection*{Computation of LD between \emph{k}-mers}
+\subsection{Computation of LD between \emph{k}-mers}
 
@@
 
-\subsection*{Software used}
+\subsection{Software used}
 \label{sv-gwas-software-used}
@@
 
-\section*{Results and discussion}
+\section{Results and discussion}
 
@@
 
-\subsection*{Loci with known genes or causal variants}
+\subsection{Loci with known genes or causal variants}
 
@@
 already been cloned. These loci provide interesting test cases because
-the expected results (causal genes or variants) are known. Table \ref{loci-table} shows
+the expected results (causal genes or variants) are known. Table 1 shows
 a summary of the results discussed below.
 
-\subsubsection*{Flower color -- \textit{W1} locus}
+\subsubsection{Flower color -- \textit{W1} locus}
 
@@
 significant SV and \textit{k}-mers corresponded to the known causal variant
-(Figure \ref{flower-color-main-figure}a-b and Supplemental Figure S4c-d).
+(Figure 1a-b and Supplemental Figure S4c-d).
 In particular, the assembly of the reads containing the most significant \textit{k}-mers
 and the alignment of the resulting haplotypes showed that these \textit{k}-mers
-directly tagged the causal variant (Figure \ref{flower-color-main-figure}c-e).
+directly tagged the causal variant (Figure 1c-e).
 As the catalogue of variants analyzed by Platypus contains only SNPs and small
@@
 associated ones (Supplemental Figure S4b).
-
-% Figure 1
-\begin{figure}
-	\centering
-	\includegraphics[width = 5in]{flower_color_W1_main_figure.png}
-	\caption{(Caption on the following page.)}
-	\label{flower-color-main-figure}
-\end{figure}
-
-\begin{figure}
-	\contcaption{Results of SV- and \textit{k}-mer-based GWAS at the
-	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots.
-	Horizontal dotted lines indicate the significance
-	threshold whereas vertical dotted lines indicate the location of the
-	\textit{W1} locus. (b) Zoomed-in Manhattan plots at the \textit{W1} locus.
-	The vertical dotted lines indicate the location of the
-	Glyma.13g072100 gene associated with this locus. The shaded gray
-	rectangles show the regions delimited by the top 5\% (SVs) and
-	1\% (\textit{k}-mers) associations.  (c) Gene model of Glyma.13g072100.
-	Rectangles represent exons, colored rectangles represent coding
-	sequences, and arrows indicate the direction of transcription. The red
-	rectangle highlights the region shown in panel (d). (d) Pairwise
-	alignment of the two haplotype sequences in the population at the
-	\textit{W1} locus. Nucleotides are
-	colored according to the \textit{p}-value of the most significant
-	\textit{k}-mer overlapping them. Vertical dotted lines between haplotypes
-	indicate sequence differences, whereas dashes indicate gaps in the
-	alignment. (e) Contingency table of the phenotypes and haplotypes found
-	in the population at the \textit{W1} locus.}
-\end{figure}
 
@@
 
-\subsubsection*{Pubescence color -- \textit{T} locus}
+\subsubsection{Pubescence color -- \textit{T} locus}
 
@@
 
-\subsubsection*{Pubescence color -- \textit{Td} locus}
+\subsubsection{Pubescence color -- \textit{Td} locus}
 
@@
 identified a signal that overlapped Glyma.03g258700 (Figure
-\ref{pubescence-color-main-figure}a-b, and Supplemental Figures S10 and
+2a-b, and Supplemental Figures S10 and
 S11).  While the most significant
@@
 \shortcite{yan2020}, it was the fourth most significantly associated variant for
-the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d, Supplemental Figure 
+the SNP-based approach (Figure 2b-d, Supplemental Figure 
 S12).  Interestingly, some of the accessions
@@
 pubescence, suggesting that there may be more than one causal variant at this
-locus (Figure \ref{pubescence-color-main-figure}e).
-
-% Figure 2
-\begin{figure}
-	\centering
-	\includegraphics[width = 5in]{pubescence_color_nogray_Td_main_figure.png}
-	\caption{(Caption on the following page.)}
-	\label{pubescence-color-main-figure}
-\end{figure}
-
-\begin{figure}
-	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
-	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
-	Horizontal dotted lines indicate the significance
-	threshold whereas vertical dotted lines indicate the location of the
-	\textit{Td} locus. (b) Zoomed-in Manhattan plots at the \textit{Td} locus.
-	The vertical dotted lines indicate the location of the
-	Glyma.03g258700 gene associated with this locus. The shaded gray
-	rectangles show the regions delimited by the top 1\% associations.
-	In the case of SNPs, blue points denote markers used in the
-	original analysis, whereas red points denote markers that had originally
-	been pruned but whose \emph{p}-values were computed after signal
-	discovery.
-	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
-	rectangles represent coding sequences, and the arrow indicates the
-	direction of transcription. The red rectangle highlights the region
-	shown in panel (d). (d) Pairwise alignment of the two haplotype
-	sequences in the population at the \textit{Td} locus. Nucleotides are
-	colored according to the
-	\textit{p}-value of the most significant \textit{k}-mer overlapping
-	them. The vertical dotted line between haplotypes indicates the location of
-	the causal SNP at this locus. (e) Contingency table of
-	the phenotypes and haplotypes found in the population at the
-	\textit{Td} locus.}
-\end{figure}
-
-\subsubsection*{Seed coat color - \textit{I} locus}
+locus (Figure 2e).
+
+\subsubsection{Seed coat color - \textit{I} locus}
 
@@
 
-\subsubsection*{Seed coat color - \textit{G} locus}
+\subsubsection{Seed coat color - \textit{G} locus}
 
@@
 SNP-based analysis as corresponding to that causal variant (Figure
-\ref{seed-coat-color-main-figure} and Supplemental Figures
+3 and Supplemental Figures
 S16 and
@@
 
-% Figure 3 with the results on pubescence color at locus G
-\begin{figure}
-	\centering
-	\includegraphics[width = 5in]{seed_coat_color_greenyellow_G_main_figure.png}
-	\caption{(Caption on the following page.)}
-	\label{seed-coat-color-main-figure}
-\end{figure}
-
-\begin{figure}
-	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
-	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
-	Horizontal dotted lines indicate the significance
-	threshold whereas vertical dotted lines indicate the location of the
-	\textit{G} locus. (b) Zoomed-in Manhattan plots at the \textit{G} locus.
-	The vertical dotted lines indicate the location of the
-	Glyma.01g198500 gene associated with this locus. The shaded gray
-	rectangles show the regions delimited by the top 1\% associations.
-	In the case of SNPs, blue points denote markers used in the
-	original analysis, whereas red points denote markers that had originally
-	been pruned but whose \emph{p}-values were computed after signal
-	discovery.
-	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
-	rectangles represent coding sequences, and the arrows indicate the
-	direction of transcription. The red rectangle highlights the region
-	shown in panel (d). (d) Pairwise alignment of the two haplotype
-	sequences in the population at the \textit{G} locus. Nucleotides are
-	colored according to the
-	\textit{p}-value of the most significant \textit{k}-mer overlapping
-	them. The vertical dotted line between haplotypes indicates the location of
-	the causal SNP at this locus. (e) Contingency table of
-	the phenotypes and haplotypes found in the population at the
-	\textit{G} locus.}
-\end{figure}
-
-\subsubsection*{Stem termination type - \textit{Dt1} locus}
+\subsubsection{Stem termination type - \textit{Dt1} locus}
 
@@
 
-\subsubsection*{Hilum color -- \textit{R} locus}
+\subsubsection{Hilum color -- \textit{R} locus}
 
@@
 
-\subsubsection*{Pubescence density -- \textit{Ps} locus}
+\subsubsection{Pubescence density -- \textit{Ps} locus}
 
@@
 detected signals at the \textit{Ps} locus using all three approaches (Figure
-\ref{pubescence-density-main-figure}a-b and Supplemental Figures S26, S27
+4a-b and Supplemental Figures S26, S27
 and S28). The causal
@@
 the population and was associated with the trait (Figure
-\ref{pubescence-density-main-figure}c), and therefore possibly corresponded to the
+4c), and therefore possibly corresponded to the
 25.6-kb region identified as the causal CNV by \shortciteA{liu2020ps}. We found that
 reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
-of the CNV (Figure \ref{pubescence-density-main-figure}d). This suggests that
+of the CNV (Figure 4d). This suggests that
 the most associated \textit{k}-mer is indeed linked to the CNV
@@
 CNV region may represent differences in sequence between the copies.
-
-% Figure 4 with the results on pubescence density at locus Ps
-\begin{figure}
-	\centering
-	\includegraphics[width = 6in]{pubescence_density_Ps_main_figure.png}
-	\caption[Results of \textit{k}-mer-based GWAS at the
-	\textit{Ps} locus for pubescence density.]{Results of \textit{k}-mer-based GWAS at the
-	\textit{Ps} locus for pubescence density.
-	(a) Genome-wide Manhattan plot showing the location of significant \textit{k}-mers. The
-	horizontal dotted line indicates the significance threshold whereas the
-	vertical dotted line indicates the location of the \textit{Ps} locus.
-	(b) Zoomed-in Manhattan plot at the \textit{Ps} locus. The
-	vertical dotted line indicates the location of the most significant
-	\textit{k}-mer. The shaded gray area shows the location of the causal CNV.
-	(c) Estimated copy number (sequencing depth at position divided by
-	accession average sequencing depth) of accessions with normal and
-	semi-sparse pubescence at the \textit{Ps} locus. Solid lines represent
-	the median copy number across 73 (semi-sparse pubescence) and 180
-	accessions (normal pubescence) computed using a 100-bp sliding window.
-	Shaded areas outline the first and third quartiles. The start and end
-	of the CNV region can be clearly inferred from the divergence between
-	the copy number of accessions with contrasting pubescence.
-	(d) Histogram showing the location of 1,169 mapped reads containing the
-	most significant \textit{k}-mer in accessions with semi-sparse pubescence.
-	Reads containing this \textit{k}-mer map to either end of the CNV
-	region, suggesting that they are associated with this variant. Panels
-	(b), (c) and (d) share the same x-axis.}
-	\label{pubescence-density-main-figure}
-\end{figure}
 
@@
 
-\subsubsection*{Seed coat luster -- \textit{B} locus}
+\subsubsection{Seed coat luster -- \textit{B} locus}
 
@@
 
-\subsection*{Analysis of loci with unknown causal genes}
+\subsection{Analysis of loci with unknown causal genes}
 
@@
 
-\subsubsection*{Pod color -- \textit{L1} and \textit{L2} loci}
+\subsubsection{Pod color -- \textit{L1} and \textit{L2} loci}
 
@@
 
-\subsubsection*{Pubescence form -- \textit{Pa1} and \textit{Pa2} loci}
+\subsubsection{Pubescence form -- \textit{Pa1} and \textit{Pa2} loci}
 \label{sv-gwas-main-results-pubescence-form-pa1-pa2}
@@
 signals found by all three approaches were located near this gene (Figure
-\ref{pubescence-form-main-figure}a-b, and Supplemental Figures S35 and
+5a-b, and Supplemental Figures S35 and
 S36).  The analyses based on SNPs and
@@
 Glyma.12g213900 that were associated with appressed pubescence (Figure
-\ref{pubescence-form-main-figure}c-d and Supplemental Figures S37 and
+5c-d and Supplemental Figures S37 and
 S38). The alternative alleles for these two
@@
 
-% Figure 5 with the results on pubescence form at locus Pa1
-\begin{figure}
-	\centering
-	\includegraphics[width = 5in]{pubescence_form_all_Pa1_main_figure.png}
-	\caption{(Caption on the following page.)}
-	\label{pubescence-form-main-figure}
-\end{figure}
-
-\begin{figure}
-	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
-	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots.
-	Horizontal dotted lines indicate the significance
-	threshold whereas vertical dotted lines indicate the location of the
-	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots at the \textit{Pa1} locus.
-	The vertical dotted lines indicate the location of the
-	Glyma.12g213900 gene suggested as a candidate for this locus. The shaded gray
-	rectangles show the regions delimited by the top 1\% associations.
-	In the case of SNPs, blue points denote markers used in the
-	original analysis, whereas red points denote markers that had originally
-	been pruned but whose \emph{p}-values were computed after signal
-	discovery.
-	(c) Gene model of Glyma.12g213900. Rectangles represent exons, colored
-	rectangles represent coding sequences, and the arrows indicate the
-	direction of transcription.
-	(d) Manhattan plots at the location of the Glyma.12g213900 gene.
-	Vertical lines indicate the location of two non-synonymous
-	SNPs associated with appressed pubescence, which may represent causal variants
-	at this locus. Panels (c) and (d) share the same x-axis.}
-\end{figure}
-
 A second GWAS analysis focused only on appressed and semi-appressed phenotypes
@@
 
-\subsubsection*{Resistance to \textit{P. sojae}}
+\subsubsection{Resistance to \textit{P. sojae}}
 
@@
 
-\subsubsection*{Novel loci for stem termination and pod color}
+\subsubsection{Novel loci for stem termination and pod color}
 
@@
 and Gm18 in addition to the known \textit{Dt1} locus (Figure
-\ref{stem-termination-main-figure}a).  Analysis of the pairwise LD between
+6a).  Analysis of the pairwise LD between
 significant \textit{k}-mers found no LD between these signals, suggesting that
 they may represent \textit{bona fide} loci associated with stem termination
-type (Figure \ref{stem-termination-main-figure}b). We note that the signal newly
+type (Figure 6b). We note that the signal newly
 identified on Gm18 is located 5 Mb away from the already documented \textit{Dt2}
@@
 
-% Figure 6 with the results novel loci for stem termination type
-\begin{figure}
-	\centering
-	\includegraphics[width = 6in]{stem_termination_sn_main_figure.png}
-	\caption[Detection of novel signals for stem termination type
-	using a \textit{k}-mer GWAS analysis contrasting accessions with
-	indeterminate and semi-determinate stems]{Detection of novel signals for stem termination type
-	using a \textit{k}-mer-based GWAS analysis contrasting accessions with
-	indeterminate and semi-determinate stems.
-	(a) Genome-wide Manhattan plot showing the location of significant \textit{k}-mers.
-	The horizontal dotted line indicates the significance threshold
-	whereas vertical dotted lines indicate the location of the \textit{Dt1}
-	classical locus as well as three new loci of interest (\textit{stGm11},
-	\textit{stGm16}, and \textit{stGm18}).
-	(b) Pairwise LD between the 162 significant \textit{k}-mers identified.
-	\textit{k}-mers are sorted along the axes according to their
-	putative position on the reference assembly version 4 of Williams82.
-	\textit{k}-mers are
-	represented in the same order along the x- and y-axis. The blue
-	rectangles below the x-axis represent the
-	-log\textsubscript{10}(\textit{p}) of each \textit{k}-mer.}
-	\label{stem-termination-main-figure}
-\end{figure}
-
-\subsection*{Potential of \emph{k}-mer-based GWAS}
+\subsection{Potential of \emph{k}-mer-based GWAS}
 
@@
 \emph{G} for seed coat color, and \emph{Ps} for pubescence density (Table
-\ref{loci-table}). In comparison, this result was achieved by SV-based GWAS at
+1). In comparison, this result was achieved by SV-based GWAS at
 the \textit{W1} locus only, and at none of the loci for SNP-based GWAS.
@@
 the other two approaches in identifying signals overlapping known variants or
-genes (Table \ref{loci-table}). Moreover, whenever \textit{k}-mer-based GWAS
+genes (Table 1). Moreover, whenever \textit{k}-mer-based GWAS
 did detect a signal, the region defined by the 1\% most significant
@@
 
-\subsection*{Limitations and challenges of \emph{k}-mer-based GWAS}
+\subsection{Limitations and challenges of \emph{k}-mer-based GWAS}
 
@@
 
-\subsection*{Use of SV genotypes for GWAS}
+\subsection{Use of SV genotypes for GWAS}
 
@@
 
-\section*{Conclusions}
+\section{Conclusions}
 
@@
 	\item[Supplemental file 1.pdf] Supplemental methods and results, including
-		Supplemental Tables S1 to S11 and Supplemental Figures S1 to
-		S60.
+		Supplemental Tables S1 to S12 and Supplemental Figures S1 to
+		S61.
 
@@
 \clearpage
+
+\section*{Figure captions}
+
+% Figure 1
+Figure 1: Results of SV- and \textit{k}-mer-based GWAS at the \textit{W1} locus
+for flower color. (a) Genome-wide Manhattan plots.  Horizontal dotted lines
+indicate the significance threshold whereas vertical dotted lines indicate the
+location of the \textit{W1} locus.  (b) Zoomed-in Manhattan plots at the
+\textit{W1} locus.  The vertical dotted lines indicate the location of the
+Glyma.13g072100 gene associated with this locus. The shaded gray rectangles
+show the regions delimited by the top 5\% (SVs) and 1\% (\textit{k}-mers)
+associations.  (c) Gene model of Glyma.13g072100.  Rectangles represent exons,
+colored rectangles represent coding sequences, and arrows indicate the
+direction of transcription. The red rectangle highlights the region shown in
+panel (d). (d) Pairwise alignment of the two haplotype sequences in the
+population at the \textit{W1} locus. Nucleotides are colored according to the
+\textit{p}-value of the most significant \textit{k}-mer overlapping them.
+Vertical dotted lines between haplotypes indicate sequence differences, whereas
+dashes indicate gaps in the alignment. (e) Contingency table of the phenotypes
+and haplotypes found in the population at the \textit{W1} locus.
+
+% Figure 2
+Figure 2: Results of SNP- and \textit{k}-mer-based GWAS at the \textit{Td}
+locus for pubescence color. (a) Genome-wide Manhattan plots.  Horizontal dotted
+lines indicate the significance threshold whereas vertical dotted lines
+indicate the location of the \textit{Td} locus. (b) Zoomed-in Manhattan plots
+at the \textit{Td} locus.  The vertical dotted lines indicate the location of
+the Glyma.03g258700 gene associated with this locus. The shaded gray rectangles
+show the regions delimited by the top 1\% associations.  In the case of SNPs,
+blue points denote markers used in the original analysis, whereas red points
+denote markers that had originally been pruned but whose \emph{p}-values were
+computed after signal discovery.  (c) Gene model of Glyma.03g258700. Rectangles
+represent exons, colored rectangles represent coding sequences, and the arrow
+indicates the direction of transcription. The red rectangle highlights the
+region shown in panel (d). (d) Pairwise alignment of the two haplotype
+sequences in the population at the \textit{Td} locus. Nucleotides are colored
+according to the \textit{p}-value of the most significant \textit{k}-mer
+overlapping them. The vertical dotted line between haplotypes indicates the
+location of the causal SNP at this locus. (e) Contingency table of the
+phenotypes and haplotypes found in the population at the \textit{Td} locus.
+
+% Figure 3 with the results on pubescence color at locus G
+Figure 3: Results of SNP- and \textit{k}-mer-based GWAS at the \textit{G} locus
+for seed coat color. (a) Genome-wide Manhattan plots.  Horizontal dotted lines
+indicate the significance threshold whereas vertical dotted lines indicate the
+location of the \textit{G} locus. (b) Zoomed-in Manhattan plots at the
+\textit{G} locus.  The vertical dotted lines indicate the location of the
+Glyma.01g198500 gene associated with this locus. The shaded gray rectangles
+show the regions delimited by the top 1\% associations.  In the case of SNPs,
+blue points denote markers used in the original analysis, whereas red points
+denote markers that had originally been pruned but whose \emph{p}-values were
+computed after signal discovery.  (c) Gene model of Glyma.01g198500. Rectangles
+represent exons, colored rectangles represent coding sequences, and the arrows
+indicate the direction of transcription. The red rectangle highlights the
+region shown in panel (d). (d) Pairwise alignment of the two haplotype
+sequences in the population at the \textit{G} locus. Nucleotides are colored
+according to the \textit{p}-value of the most significant \textit{k}-mer
+overlapping them. The vertical dotted line between haplotypes indicates the
+location of the causal SNP at this locus. (e) Contingency table of the
+phenotypes and haplotypes found in the population at the \textit{G} locus.
+
+% Figure 4 with the results on pubescence density at locus Ps
+Figure 4: Results of \textit{k}-mer-based GWAS at the \textit{Ps} locus for
+pubescence density.  (a) Genome-wide Manhattan plot showing the location of
+significant \textit{k}-mers. The horizontal dotted line indicates the
+significance threshold whereas the vertical dotted line indicates the location
+of the \textit{Ps} locus.  (b) Zoomed-in Manhattan plot at the \textit{Ps}
+locus. The vertical dotted line indicates the location of the most significant
+\textit{k}-mer. The shaded gray area shows the location of the causal CNV.  (c)
+Estimated copy number (sequencing depth at position divided by accession
+average sequencing depth) of accessions with normal and semi-sparse pubescence
+at the \textit{Ps} locus. Solid lines represent the median copy number across
+73 (semi-sparse pubescence) and 180 accessions (normal pubescence) computed
+using a 100-bp sliding window.  Shaded areas outline the first and third
+quartiles. The start and end of the CNV region can be clearly inferred from the
+divergence between the copy number of accessions with contrasting pubescence.
+(d) Histogram showing the location of 1,169 mapped reads containing the most
+significant \textit{k}-mer in accessions with semi-sparse pubescence.  Reads
+containing this \textit{k}-mer map to either end of the CNV region, suggesting
+that they are associated with this variant. Panels (b), (c) and (d) share the
+same x-axis.
+
+% Figure 5 with the results on pubescence form at locus Pa1
+Figure 5: Results of SNP- and \textit{k}-mer-based GWAS at the \textit{Pa1}
+locus for pubescence form. (a) Genome-wide Manhattan plots.  Horizontal dotted
+lines indicate the significance threshold whereas vertical dotted lines
+indicate the location of the \textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in
+Manhattan plots at the \textit{Pa1} locus.  The vertical dotted lines indicate
+the location of the Glyma.12g213900 gene suggested as a candidate for this
+locus. The shaded gray rectangles show the regions delimited by the top 1\%
+associations.  In the case of SNPs, blue points denote markers used in the
+original analysis, whereas red points denote markers that had originally been
+pruned but whose \emph{p}-values were computed after signal discovery.  (c)
+Gene model of Glyma.12g213900. Rectangles represent exons, colored rectangles
+represent coding sequences, and the arrows indicate the direction of
+transcription.  (d) Manhattan plots at the location of the Glyma.12g213900
+gene.  Vertical lines indicate the location of two non-synonymous SNPs
+associated with appressed pubescence, which may represent causal variants at
+this locus. Panels (c) and (d) share the same x-axis.
+
+% Figure 6 with the results novel loci for stem termination type
+Figure 6: Detection of novel signals for stem termination type using a
+\textit{k}-mer-based GWAS analysis contrasting accessions with indeterminate
+and semi-determinate stems.  (a) Genome-wide Manhattan plot showing the
+location of significant \textit{k}-mers.  The horizontal dotted line indicates
+the significance threshold whereas vertical dotted lines indicate the location
+of the \textit{Dt1} classical locus as well as three new loci of interest
+(\textit{stGm11}, \textit{stGm16}, and \textit{stGm18}).  (b) Pairwise LD
+between the 162 significant \textit{k}-mers identified.  \textit{k}-mers are
+sorted along the axes according to their putative position on the reference
+assembly version 4 of Williams82.  \textit{k}-mers are represented in the same
+order along the x- and y-axis. The blue rectangles below the x-axis represent
+the -log\textsubscript{10}(\textit{p}) of each \textit{k}-mer.
 
